---
title: "Plot Summary of Gating Results"
output: 
  html_document:
    toc: true
    toc_depth: 4
params:
  data: ""
  data_name: ""
  ind_batch_list: !r list()
  ind_in_batch_gate: 1
  ind_in_batch_uns: 1
  ind_in_batch_lab_vec: ""
  pop_gate: ""
  cut: ""
  high: ""
  gate_combn: !r list()
  pop_man_sub: ""
  pop_man_match_exact: true
  tol: 1
  fdr: 1
  chnl_lab: ""
  pop_sub: "gate"
  noise_sd: NULL
  bias_uns: 1
  bw_min: 1
  cp_min: 1
  boot_sd: NULL
  boot_n: NULL
  min_cell: 1
  max_pos_prob_x: 1
  gate_quant: 1
  tol_ctrl: 1e-3
  tol_gate: 1e-6
  tol_gate_single: 1e-7
  gate_tbl: NULL
---

```{r g-setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
```

```{r g-misc, eval = TRUE}
pop_gate <- params$pop_gate
high <- params$high
pop_sub <- params$pop_sub
```

```{r heading-pop_marker, results = 'asis', eval = TRUE}
pander::pandoc.header(paste0(pop_gate, " - ", params$chnl_lab[params$cut]), 
                      level = 1)
```

## Key for legends

- Gate method
  - Manual: manual gating cutpoint
  - Finak (2014): cytoUtils first-derivative-based tailgate/cytokinegate cutpoint
  - Uns-based (#): Unstim-based method using a false-discovery rate (FDR) of #, where # is a percentage (e.g. 20)
- Gate combination method
  - Ind: gates determined per-stim-sample and not combined to form a single gate for all samples within a batch
  - Batch: gates determined per-batch (i.e. all stim samples concatenated to get the gate)
  - Min: gates determined per-sample and combined afterwards by setting each stim sample to have the lowest gate across all stim samples

For example, `Uns-based (20) - Min` means that the unstim-based method with an FDR of 20 was used to determine the cutpoint for each stimulation sample individually within a batch (the gating method). A common cutpoint for all stimulation samples within that batch was set to be the minimum cutpoint across the stimulation samples. 

```{r g-setup_objects, eval = TRUE}
# base dir
dir_base <- stimgate_dir_base_create(params = params)
# pops for which results are plotted
pop_res <- c("gate", params$pop_sub)
```

```{r g-print_plots_in_dir, eval = TRUE}
.print_plots_in_dir <- function(dir_plot){
    # get plot paths
    plot_path_vec <- list.files(dir_plot, full.names = TRUE)
    plot_path_vec <- plot_path_vec[!stringr::str_detect(plot_path_vec, "count")]
    plot_path_vec <- plot_path_vec[!stringr::str_detect(plot_path_vec, "dodge_gm")]
    #plot_path_vec <- plot_path_vec[!stringr::str_detect(plot_path_vec, "dodge_stim-freq_bs")]
    #plot_path_vec <- plot_path_vec[!stringr::str_detect(plot_path_vec, "dodge_stim-count_bs")]
    # get plot anmes
    plot_name_vec <- list.files(dir_plot, full.names = FALSE) |>
      stringr::str_remove(".png")
    plot_name_vec <- plot_name_vec[!stringr::str_detect(plot_name_vec, "count")]
    plot_name_vec <- plot_name_vec[!stringr::str_detect(plot_name_vec, "dodge_gm")]
    if(length(plot_name_vec) == 0) return(invisible(NULL))
    #plot_name_vec <- plot_name_vec[!stringr::str_detect(plot_name_vec, "dodge_stim-freq_bs")]
    #plot_name_vec <- plot_name_vec[!stringr::str_detect(plot_name_vec, "dodge_stim-count_bs")]
    # replace names with reasonable values
    plot_name_vec <- plot_name_vec |>
      stringr::str_replace("stim_vs_uns_jc", "Proportion of of stim over (stim + unstim) ") |>
      stringr::str_replace("auto_vs_man", "Auto vs manual") |>
      stringr::str_replace("stim_vs_uns_dodge_stim-freq_bs", 
                  "Frequency (background subtracted)") |>
      stringr::str_replace("stim_vs_uns_dodge_stim-freq", 
                  "Stim vs unstim - Frequency") |>
      stringr::str_replace("dodge_gm", " - gate method as dodge") |>
      stringr::str_replace("dodge_stim", " - stim as dodge") |>
      stringr::str_replace("-count_bs", "- count (background subtracted)") |>
      stringr::str_replace("-count", "- count") |>
      stringr::str_replace("-freq_bs", " frequency (background subtracted)") |>
      stringr::str_replace("-freq", "- frequency") |>
      stringr::str_replace("ccc", "Concordance correlation coefficient") |>
      stringr::str_replace("pcc", "Pearson's correlation coefficient") |>
      stringr::str_replace("std_diff_ind-abs", "Absolute standardised difference from manual") |> 
      stringr::str_replace("std_diff_ind", "Standardised difference from manual") |>
      stringr::str_replace("p_s2n_freq_bs", "Signal-to-noise for background-subtracted frequencies") |>
      stringr::str_replace("p_seprop_freq_bs", "Mean signal less one standard error of mean for background-subtracted frequencies") |>
      stringr::str_replace("p_s2n_ecdf", "Survival function for signal-to-noise ratio")
    # print plots for each batch
    for(i in seq_along(plot_name_vec)){
      # print fcs name
      pander::pandoc.header(plot_name_vec[i], level = 5)
      # print plot
      cat(paste0("![](", plot_path_vec[i], ")"), "\n")
    }
    invisible(TRUE)
}
```

## Distribution of thresholds

```{r g-print_plots_in_dir_thresh, ,results = 'asis'}
.print_plots_in_dir(dir_plot =  file.path(dir_base, "gating_plots", pop_res[1],
                                          'plot_hist_gates'))
```

## 2D

```{r g-print_2d, results = 'asis', eval = TRUE}
# loop over pops for which results are required
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 3)
  # loop over high markers
  for(high_marker in names(high)){
    # get common name for high marker
    high_marker_name <- params$chnl_lab[high_marker][[1]]
    # print high marker name
    pander::pandoc.header(high_marker_name, level = 4)
    # get plot directory
    dir_plot <- file.path(dir_base, "gating_plots", pop_res_curr, 
                          paste0("plot2d-", high_marker_name))
    # print plots
    .print_plots_in_dir(dir_plot = dir_plot)
  }
}
```

## Local FDR approach

### Density

```{r g-print_density, results = 'asis'}
# loop over pops for which results are required
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 4)
  # loop over high markers
  for(bias in params$bias_uns){
    # bias level
    # print high marker name
    pander::pandoc.header(paste0("Bias - ", bias), level = 5)
    # get plot directory
    dir_plot <- file.path(dir_base, "gating_plots", pop_res_curr, 
                          paste0("p_locb", bias, "_dens"))
    # print plots
    .print_plots_in_dir(dir_plot = dir_plot)
  }
}
```

### Probability of positivity

```{r g-print_prob, results = 'asis'}
# loop over pops for which results are required
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 4)
  # loop over high markers
  for(bias in params$bias_uns){
    # bias level
    # print high marker name
    pander::pandoc.header(paste0("Bias - ", bias), level = 5)
    # get plot directory
    dir_plot <- file.path(dir_base, "gating_plots", pop_res_curr, 
                          paste0("p_locb", bias, "_prob"))
    # print plots
    .print_plots_in_dir(dir_plot = dir_plot)
  }
}
```

### Contribution to count

```{r g-print_ctb, results = 'asis'}
# loop over pops for which results are required
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 4)
  # loop over high markers
  for(bias in params$bias_uns){
    # bias level
    # print high marker name
    pander::pandoc.header(paste0("Bias - ", bias), level = 5)
    # get plot directory
    dir_plot <- file.path(dir_base, "gating_plots", pop_res_curr, 
                          paste0("p_locb", bias, "_ctb"))
    # print plots
    .print_plots_in_dir(dir_plot = dir_plot)
  }
}
```

## Frequency by cutpoint

```{r g-freq_by_cp, results = 'asis', eval = TRUE}
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 3)
  # get plot directory
  dir_plot <- file.path(dir_base, "gating_plots", pop_res_curr, 
                        "plot_gate_stat-p_freq")
  # print plots
  .print_plots_in_dir(dir_plot = dir_plot)
}
```

## Plots of auto vs manual and stim vs uns counts and frequencies (raw and background-subtracted)

```{r g-auto_vs_man, results = 'asis', eval = TRUE}
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 3)
  # get plot directory
  dir_plot <- file.path(dir_base, "stats", pop_res_curr)
  # print plots
  .print_plots_in_dir(dir_plot = dir_plot)
}
```

Note that the final plot is meant to show the relative size of stim to unstim. A value of 1 means that there were no unstim cells and only stim cells. A value of 0.5 means that there were as many stim cells as unstim cells. A value of 0 means that there were no stim cells and only unstim cells. 

```{r g-perf, results = 'asis', eval = !is.null(params$pop_man_sub)}
pander::pandoc.header("Performance", level = 2)
for(pop_res_curr in pop_res){
  # print each pop
  pander::pandoc.header(ifelse(pop_res_curr == "gate", 
                               "Gating pop", 
                               pop_res_curr), level = 3)
  # get plot directory
  dir_plot <- file.path(dir_base, "perf", pop_res_curr)
  # print plots
  .print_plots_in_dir(dir_plot = dir_plot)
}
```
