---
title: "Setting up coding agents for StimGate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting up coding agents for StimGate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Set up a cloud coding agent with the repository's scripts or workflow, then check that StimGate loads. Agents must read `AGENTS.md` before changing code; it defines coding conventions and required checks.

## Choose the setup

| Agent | Configuration in this repository |
|---|---|
| OpenAI Codex Cloud | `scripts/agents/setup-codex.sh` and `scripts/agents/maintenance-codex.sh` |
| Google Jules | `scripts/agents/setup-jules.sh` |
| GitHub Copilot coding agent | `.github/workflows/copilot-setup-steps.yml` |

Keep setup changes in these files. If a platform requires pasted script text, copy the current repository file rather than maintaining another version.

## Shared environment rules

- **Cloud library:** set persistent `CI=true` so `.Rprofile` skips `renv`. Use the installed development library; missing packages indicate a setup problem. Do not run `renv::restore()` in these cloud environments.
- **Local library:** ordinary local development still uses the project's `renv` profiles when a reproducible library is needed.
- **Dependencies:** Codex/Jules scripts target R 4.6 and Bioconductor 3.23. They install system libraries, Quarto, `DESCRIPTION` development dependencies, and tools including `devtools`, `rcmdcheck`, `roxygen2`, `styler`, `lintr`, `covr` and `pkgdown`.
- **Analysis support:** Codex/Jules also install Python/NumPy, `simcyto`, `future`, `furrr`, `progressr`, `reticulate` and `cytoUtils`. F-beta comparisons use Python through `reticulate`.
- **`simcyto`:** use unpinned `SATVILab/simcyto` for development, including Copilot and analysis CI. Pin a SHA only for a task that requires historical reproduction or a version-specific regression check.
- **Validation:** the scripts check required packages and `devtools::load_all()`. Codex/Jules repair failed installation stages before these final checks; an intermediate failure alone does not establish that setup failed.

Codex/Jules use this Posit CRAN URL for Linux binaries:

```text
https://packagemanager.posit.co/cran/__linux__/noble/latest
```

Keep the `__linux__` form used by the scripts. The `/bin/linux/...` form has been classified as a source repository by pak, causing ordinary CRAN packages to compile from source.

## OpenAI Codex Cloud

Use the universal container image, **Manual** setup and container caching. Configure `CI=true` as a persistent environment variable: setup runs in a separate shell, so an export there does not reach later agent commands. See the [OpenAI environment guide](https://learn.chatgpt.com/docs/environments/cloud-environment) for platform settings.

Setup script:

```bash
bash scripts/agents/setup-codex.sh
```

Maintenance script:

```bash
bash scripts/agents/maintenance-codex.sh
```

Maintenance installs newly declared development dependencies, refreshes `SATVILab/simcyto`, ensures repository tools remain available, and checks that StimGate loads.

### Network access

For task-time documentation lookup or dependency downloads, enable agent internet access with **Common dependencies** and add the domains needed below. Allow additional HTTP methods only when required, for example for Git-over-HTTPS operations. See the [OpenAI internet-access guide](https://learn.chatgpt.com/docs/cloud/internet-access) for allowlist controls.

```text
github.com
api.github.com
codeload.github.com
raw.githubusercontent.com
objects.githubusercontent.com
r-lib.github.io
cloud.r-project.org
cran.r-project.org
r-project.org
packagemanager.posit.co
p3m.dev
bioconductor.org
```

Keep task-time access limited to needed domains. Setup scripts have internet access for installing dependencies.

## Google Jules

Use this setup command:

```bash
bash scripts/agents/setup-jules.sh
```

The script expects the checkout under `/app` and exports `CI=true` during installation. If Jules offers persistent project environment variables, also set `CI=true` there for later R sessions.

Use the same package-source domains listed above if network controls are available. Rebuilding installs unpinned `SATVILab/simcyto`; to refresh a long-lived environment, rerun setup or use the refresh command under Troubleshooting.

## GitHub Copilot coding agent

Configure Copilot through `.github/workflows/copilot-setup-steps.yml`. It installs Linux system libraries, the current R release with Posit Package Manager, package/development dependencies, `devtools`, `rcmdcheck`, `decor`, unpinned `SATVILab/simcyto` and Quarto. Its verification step checks core packages and loads StimGate.

`.Rprofile` recognises Copilot as CI and skips `renv`. Use its installed library directly.

When adding a dependency:

1. Put package dependencies in `DESCRIPTION`.
2. Add agent-only tools to the workflow when Copilot tasks need them.
3. Keep `SATVILab/simcyto` unpinned, as in analysis CI.
4. Extend verification when a missing dependency should fail setup.

Keep configuration in the workflow; add a separate installer only if the workflow cannot express the required setup.

## Check a coding task

Run commands from the repository root. Follow the full checklist in `AGENTS.md`; common package checks are:

```bash
Rscript -e "devtools::load_all(); devtools::test()"
```

When changing `analysis/` or `scripts/r/`, also run:

```bash
Rscript analysis/tests/run_analysis_tests.R
```

The GitHub Actions workflows provide a smaller baseline for agents that only need package tests. Narrow documentation or package tasks may not need Quarto or the comparison stack; the full Codex/Jules setup covers research analyses too.

## Troubleshooting

| Symptom | Check or action |
|---|---|
| R activates `renv` | Check task-time `CI=true`, not just the setup export. For Copilot, check `.Rprofile` still recognises its environment. |
| pak compiles many ordinary CRAN packages | Inspect `pak::repo_status()`. Codex/Jules should use the `__linux__/noble/latest` URL above; check it was not replaced with `/bin/linux/...`. |
| `simcyto` is stale | Run Codex maintenance, or refresh with the R command below. For Copilot, check the workflow is unpinned and provision a new environment. |
| Bioconductor installation fails | If final validation fails after repair, inspect the missing package's system requirements. For Copilot, inspect the system-library and `setup-r-dependencies` steps. |
| StimGate loads but analysis tests fail | Check analysis dependencies, including `simcyto` and comparison tools. Use full Codex/Jules setup, or add the needed dependency to Copilot's workflow. |

Refresh `simcyto` in an existing development library:

```{r refresh-simcyto, eval=FALSE}
pak::pkg_install("SATVILab/simcyto", upgrade = TRUE, ask = FALSE)
```

## Add another agent

1. Add a small script under `scripts/agents/`, or use the platform's repository workflow.
2. Reuse the R/Bioconductor dependency strategy unless the platform has a suitable native equivalent.
3. Keep repository conventions in `AGENTS.md` and add setup instructions to this guide.
4. Avoid duplicating pinned dependency lists across configuration files.
