---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stimgate)
```

# Introduction

The `stimgate` package provides tools to identify cells that have possibly responded to stimulation by comparing unstimulated and stimulated tubes from the same sample.

## Main Functions

The package provides several key functions:

- `stimgate_gate()`: Main function to identify cytokine-positive cells by gating
- `get_stats()`: Generate statistics from gating results
- `stim_gate_plot()`: Visualize identified gates
- `get_gate_tbl()`: Extract gate information
- `stimgate_fcs_write()`: Write FCS files of cytokine-positive cells

## Basic Usage

```{r eval=FALSE}
# Basic gating workflow
result <- stimgate_gate(
  path_project = "/path/to/project",
  .data = gs, # GatingSet object
  batch_list = list(batch1 = 1:10, batch2 = 11:20),
  marker = list(
    list(cut = "IL2", tol = 0.5e-8),
    list(cut = "TNFa", tol = 0.5e-8)
  )
)

# Get statistics
stats <- get_stats("/path/to/project")

# Get gate table
gates <- get_gate_tbl("/path/to/project")

# Plot gates
plots <- stim_gate_plot(
  ind = 1:3,
  .data = gs,
  path_project = "/path/to/project",
  marker = c("IL2", "TNFa")
)
```

For more detailed examples and advanced usage, please refer to the function documentation.

```{r}
sessionInfo()
```