---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stimgate)
```

# Introduction

The `stimgate` package provides tools to identify cells that have possibly responded to stimulation by comparing unstimulated and stimulated tubes from the same sample.

## Main Functions

The package provides several key functions:

- `gateStim()`: Main function to identify cytokine-positive cells by gating
- `getStimStats()`: Get statistics from gating results
- `plotStim()`: Visualise identified gates
- `getStimGates()`: Extract gate information
- `writeStimFCS()`: Write FCS files of cytokine-positive cells

## Specifying batches

`batchList` tells `gateStim()` which samples to compare. Each element is one
batch (typically one donor or sample) and holds the indices of its samples in
the `GatingSet`. **The first index in each element is the unstimulated
sample**; the remaining indices are the stimulated samples from the same batch.
No other argument identifies the unstimulated sample.

```{r eval=FALSE}
# Sample 3 is the unstimulated sample for donor 1 (stimulated: 1, 2),
# and sample 6 for donor 2 (stimulated: 4, 5)
batchList <- list(donor1 = c(3, 1, 2), donor2 = c(6, 4, 5))
```

Every batch needs an unstimulated sample and at least one stimulated sample.
An unstimulated sample may be shared by several batches, provided it is first
in each; a stimulated sample may belong to only one batch. `getBatchList()`
builds a `batchList` in this order from a table of sample metadata.

## Basic Usage

```{r eval=FALSE}
# Basic gating workflow
gateStim(
  pathProject = "/path/to/project",
  .data = gs, # GatingSet object
  batchList = batchList,
  marker = c("IL2", "TNFa")
)

# Get statistics
stats <- getStimStats("/path/to/project")

# Get gate table
gates <- getStimGates("/path/to/project")

# Plot gates for the first batch
plots <- plotStim(
  ind = batchList[[1]],
  .data = gs,
  pathProject = "/path/to/project",
  marker = c("IL2", "TNFa")
)
```

For more detailed examples and advanced usage, please refer to the function documentation.

`gateStim()` takes its data and marker arguments directly (`pathProject`, `.data`, `batchList`, `marker` or `chnl`, `popGate`, `biasUns`, `bw`). All other method, bandwidth and thresholding settings are supplied through its `control` argument as a `stimControl()` object, and per-marker overrides through `markerControl`, keyed by marker label or channel name.

```{r}
sessionInfo()
```
