---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stimgate)
```

# Introduction

The `stimgate` package provides tools to identify cells that have possibly responded to stimulation by comparing unstimulated and stimulated tubes from the same sample.

## Main Functions

The package provides several key functions:

- `stimgate_gate()`: Main function to identify cytokine-positive cells by gating
- `get_stats()`: Generate statistics from gating results
- `stimgate_plot()`: Visualize identified gates
- `stimgate_gate_get()`: Extract gate information
- `stimgate_fcs_write()`: Write FCS files of cytokine-positive cells

## Input formats

`gateStim()`, `plotStim()`, `writeStimFCS()` and `getStimExpr()` accept a
GatingSet, flowSet/cytoset, individual flowFrame/cytoframe, FCS file paths or a
single FCS directory, and a list of numeric matrices or data frames (cells by
channels). Lists use their names as sample names, or `sample1`, `sample2`, etc.
Columns must have matching names; their order is aligned to the first sample.
A long data frame may instead contain a `sample` column: samples follow observed
factor levels or first appearance. Matrix column names work as channels and markers.

FCS directories are searched non-recursively, case-insensitively, with sorted
filenames. Explicit file vectors preserve order; sample names are file basenames.
Only GatingSets provide populations beyond `root`. StimGate adds no arcsinh or
logicle transformation; prepare the desired scale before gating. FCS reading
uses the defaults of `flowWorkspace::load_cytoset_from_fcs()`.

`batchList` accepts indices or sample names, with the unstimulated control first.
Use the same sample order when supplying data to downstream functions.

```{r eval=FALSE}
matrices <- list(control = control_matrix, stimulated = stimulated_matrix)
gateStim(
  pathProject = "matrix-results", .data = matrices,
  batchList = list(donor1 = c("control", "stimulated")),
  chnl = colnames(control_matrix)
)
```

## Basic Usage

```{r eval=FALSE}
# Basic gating workflow
result <- stimgate_gate(
  pathProject = "/path/to/project",
  .data = gs, # GatingSet object
  batchList = list(batch1 = 1:10, batch2 = 11:20),
  marker = list(
    list(cut = "IL2", tol = 0.5e-8),
    list(cut = "TNFa", tol = 0.5e-8)
  )
)

# Get statistics
stats <- get_stats("/path/to/project")

# Get gate table
gates <- get_gate_tbl("/path/to/project")

# Plot gates
plots <- stimgate_plot(
  ind = 1:3,
  .data = gs,
  pathProject = "/path/to/project",
  marker = c("IL2", "TNFa")
)
```

For more detailed examples and advanced usage, please refer to the function documentation.

`gateStim()` takes its data and marker arguments directly (`pathProject`, `.data`, `batchList`, `marker` or `chnl`, `popGate`, `biasUns`, `bw`). All other method, bandwidth and thresholding settings are supplied through its `control` argument as a `stimControl()` object, and per-marker overrides through `markerControl`, keyed by marker label or channel name.

```{r}
sessionInfo()
```
