---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stimgate)
```

# Introduction

The `stimgate` package provides tools to identify cells that have possibly responded to stimulation by comparing unstimulated and stimulated tubes from the same sample.

## Main Functions

The package provides several key functions:

- `gateStim()`: Main function to identify cytokine-positive cells by gating
- `getStimStats()`: Get statistics from gating results
- `plotStim()`: Visualise identified gates
- `getStimGates()`: Extract gate information
- `writeStimFCS()`: Write FCS files of cytokine-positive cells

## Input formats

`gateStim()`, `plotStim()`, `writeStimFCS()` and `getStimExpr()` accept a
GatingSet, flowSet/cytoset, individual flowFrame/cytoframe, FCS file paths or a
single FCS directory, and a list of numeric matrices or data frames (cells by
channels). Lists use their names as sample names, or `sample1`, `sample2`, etc.
Columns must have matching names; their order is aligned to the first sample.
A long data frame may instead contain a `sample` column: samples follow observed
factor levels or first appearance. Matrix column names work as channels and markers.

FCS directories are searched non-recursively, case-insensitively, with sorted
filenames. Explicit file vectors preserve order; sample names are file basenames.
Only GatingSets provide populations beyond `root`. StimGate adds no arcsinh or
logicle transformation; prepare the desired scale before gating. FCS reading
uses the defaults of `flowWorkspace::load_cytoset_from_fcs()`.

Use the same sample order when supplying data to downstream functions.

```{r eval=FALSE}
matrices <- list(control = control_matrix, stimulated = stimulated_matrix)
gateStim(
  pathProject = "matrix-results", .data = matrices,
  batchList = list(donor1 = c("control", "stimulated")),
  chnl = colnames(control_matrix)
)
```

## Specifying batches

`batchList` tells `gateStim()` which samples to compare. Each element is one
batch (typically one donor or sample) and holds its samples, as indices into
`.data` or as sample names. **The first sample in each element is the
unstimulated sample**; the remaining samples are the stimulated samples from
the same batch. No other argument identifies the unstimulated sample.

```{r eval=FALSE}
# Sample 3 is the unstimulated sample for donor 1 (stimulated: 1, 2),
# and sample 6 for donor 2 (stimulated: 4, 5)
batchList <- list(donor1 = c(3, 1, 2), donor2 = c(6, 4, 5))
```

Every batch needs an unstimulated sample and at least one stimulated sample.
An unstimulated sample may be shared by several batches, provided it is first
in each; a stimulated sample may belong to only one batch. `getBatchList()`
builds a `batchList` in this order from a table of sample metadata.

## Basic Usage

```{r eval=FALSE}
# Basic gating workflow
gateStim(
  pathProject = "/path/to/project",
  .data = gs, # GatingSet object
  batchList = batchList,
  marker = c("IL2", "TNFa")
)

# Get statistics
stats <- getStimStats("/path/to/project")

# Get gate table
gates <- getStimGates("/path/to/project")

# Plot gates for the first batch
plots <- plotStim(
  ind = batchList[[1]],
  .data = gs,
  pathProject = "/path/to/project",
  marker = c("IL2", "TNFa")
)
```

For more detailed examples and advanced usage, please refer to the function documentation.

`gateStim()` takes its data and marker arguments directly (`pathProject`, `.data`, `batchList`, `marker` or `chnl`, `popGate`, `biasUns`, `bw`). All other method, bandwidth and thresholding settings are supplied through its `control` argument as a `stimControl()` object, and per-marker overrides through `markerControl`, keyed by marker label or channel name.

```{r}
sessionInfo()
```
