---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stimgate)
```

# Introduction

The `stimgate` package provides tools to identify cells that have possibly responded to stimulation by comparing unstimulated and stimulated tubes from the same sample.

## Main Functions

The package provides several key functions:

- `gateStim()`: Main function to identify cytokine-positive cells by gating, writing the results to a project directory
- `stimControl()`: Collects the tuning settings passed to `gateStim()` through its `control` argument
- `getStimGates()` / `getStimGatesDetailed()`: Read the fitted gates and detailed threshold diagnostics
- `getStimStats()`: Generate statistics from gating results
- `plotStim()`: Visualize identified gates
- `getStimExpr()`: Read the expression data saved by a project, optionally applying the saved gates
- `writeStimFCS()`: Write FCS files of cytokine-positive cells

## Basic Usage

```{r eval=FALSE}
# Basic gating workflow
example_data <- getExampleData()
gs <- flowWorkspace::load_gs(example_data$pathGs)

path_project <- gateStim(
  pathProject = file.path(tempdir(), "stimgate-example"),
  .data = gs, # GatingSet object
  batchList = example_data$batchList,
  marker = example_data$marker
)

# Get gate table and statistics
gates <- getStimGates(path_project)
stats <- getStimStats(path_project)

# Plot gates
plotStim(
  ind = example_data$batchList[[1]],
  .data = gs,
  pathProject = path_project,
  marker = example_data$marker
)
```

All method, bandwidth and thresholding settings are passed as a single `stimControl()` object through `gateStim()`'s `control` argument, and per-marker overrides as a `markerControl` list keyed by marker label or channel name.

```{r eval=FALSE}
gateStim(
  pathProject = path_project,
  .data = gs,
  batchList = example_data$batchList,
  marker = example_data$marker,
  control = stimControl(clusterGates = FALSE, bwMtd = "nrd0"),
  markerControl = list(
    MarkerF1 = list(bw = 0.12, biasUns = 0)
  )
)
```

For more detailed examples and advanced usage, please refer to the function documentation.

```{r}
sessionInfo()
```
