---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`stimgate` finds cells that may have responded to stimulation. For each donor, it compares the stimulated samples with an unstimulated sample from the same donor. This guide gates the example data, reads the results and plots them.

## 1. Load the data

`getExampleData()` loads a small example dataset that comes with the package: two donors, each with an unstimulated and a stimulated sample, and two markers. It returns where the data are saved, the marker names, and `batchList`, which groups each donor's samples.

```{r data, message = FALSE}
library(stimgate)
exampleData <- getExampleData()
gs <- flowWorkspace::load_gs(exampleData$pathGs)
exampleData$batchList
exampleData$marker
```

### Your own data

You can give `gateStim()` any of these:

- a GatingSet, flowSet, cytoset or flowFrame;
- FCS file paths, or a folder of FCS files;
- a list of numeric matrices or data frames, one per sample, with a column per channel;
- one data frame with a `sample` column saying which sample each cell is from.

Each element of `batchList` lists one donor's samples, by number or by name. **The first sample is the unstimulated one**; the rest are that donor's stimulated samples. Several donors can share one unstimulated sample (it must be first in each), but a stimulated sample can belong to only one donor. `getBatchList()` can build `batchList` from a table describing your samples.

```{r own-data, eval = FALSE}
samples <- list(uns = unsMatrix, stim = stimMatrix) # columns named by channel
batchList <- list(donor1 = c("uns", "stim"))
```

StimGate does not transform your data (for example with arcsinh), so give it values on the scale you want to gate. With matrices, the column names are used as both channel and marker names. Only a GatingSet can hold gated populations other than all cells (`"root"`). Use the same data, in the same sample order, when you later call `plotStim()` or `writeStimFCS()`.

## 2. Find the gates

Give `gateStim()` a folder for its results and the markers to gate. It saves everything in that folder and returns the folder's path. By default it uses all cells; set `popGate` to gate within a population already in your GatingSet.

```{r gate, message = FALSE}
pathProject <- gateStim(
  pathProject = tempfile("stimgate_"),
  .data = gs,
  batchList = exampleData$batchList,
  marker = exampleData$marker
)
```

To pick channels by channel name instead, use `chnl` instead of `marker`. The defaults suit most data; see `?stimControl` for the settings you can change, and `?gateStim` for setting them separately for each marker.

## 3. Read the gates and statistics

`getStimGates()` gives the gates for each stimulated sample (`ind`) and marker. `gate` is the gate value. `gateCyt` is the gate used for cells already positive for another cytokine; it can be lower than `gate`.

```{r gates}
gates <- getStimGates(pathProject)
head(gates)
```

`getStimStats()` gives, for each stimulated sample (`ind`), the number and percentage of cells positive for each combination of markers (`cytCombn`), in the stimulated sample and in its unstimulated sample. `freqBs` is the stimulated percentage minus the unstimulated percentage: the response after removing background.

```{r stats}
stats <- getStimStats(pathProject)
head(stats[, c("ind", "cytCombn", "countStim", "freqStim", "freqUns", "freqBs")])
```

## 4. Plot the results

Plot the first donor's samples to compare unstimulated and stimulated expression, with the gate drawn on. One marker gives density curves; two markers also give two-dimensional plots, which need the `hexbin` package.

```{r plots, fig.width = 7, fig.height = 4}
plotStim(
  ind = exampleData$batchList[[1]], .data = gs,
  pathProject = pathProject, marker = exampleData$marker[1]
)
```

The results stay in `pathProject`, so you can read them again in a later R session. To see more detail on how each gate was chosen, see `?getStimGatesDetailed`. To save the positive cells as FCS files, see `?writeStimFCS`.

## Session information

```{r session}
sessionInfo()
```
