---
title: "Getting Started with stimgate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with stimgate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`stimgate` identifies cells that may have responded to stimulation by comparing stimulated tubes with an unstimulated control from the same donor. Run the following chunks in order to save gates, read response statistics and inspect plots.

## 1. Load the data

`getExampleData()` returns a list with a saved GatingSet path, marker and channel labels, and tube indices grouped by donor. Load that GatingSet:

```{r data, eval=FALSE}
library(stimgate)
exampleData <- getExampleData()
gs <- flowWorkspace::load_gs(exampleData$pathGs)
exampleData$batchList
exampleData$marker
```

For your own data, supply a `flowWorkspace` GatingSet with consistent channels across tubes. In each `batchList` element, put the unstimulated tube index first, followed by its matched stimulated tubes.

## 2. Find gates

Choose an output directory and the markers to gate. `gateStim()` saves results there and returns the directory path as a character string. The default `popGate = "root"` uses all cells; choose an existing GatingSet population to restrict gating.

```{r gate, eval=FALSE}
pathProject <- gateStim(
  pathProject = tempfile("stimgate_"),
  .data = gs,
  batchList = exampleData$batchList,
  marker = exampleData$marker
)
```

Use `chnl` instead of `marker` to select channels by name. Start with the defaults; see `?stimControl` for tuning settings and `?gateStim` for per-marker overrides through `markerControl`.

## 3. Read gates and statistics

`getStimGates()` returns a tibble of stimulated-tube thresholds. Columns include `pop`, `marker`, `chnl`, `batch`, `ind` (tube index), `gateName`, `gate` and `gateCyt` (the threshold used for cells positive for another cytokine).

```{r results, eval=FALSE}
gates <- getStimGates(pathProject)
head(gates)
stats <- getStimStats(pathProject)
head(stats)
```

`getStimStats()` reads a data frame with rows for tubes (`ind`), gate methods (`gateName`) and marker combinations (`cytCombn`). `countStim` and `countUns` count cells in each combination; `nCellStim` and `nCellUns` give total cells. `propBs = propStim - propUns` subtracts control from stimulated proportions. `freqStim` and `freqUns` are percentages; `freqBs` is their difference in percentage points.

## 4. Inspect plots

Plot the first donor's tubes to compare control and stimulated expression with the gate lines. One marker gives density plots; two also give bivariate hex plots (requiring `hexbin`). The default output is a ggplot grid; `grid = FALSE` returns a list of ggplot objects.

```{r plots, eval=FALSE}
plotStim(
  ind = exampleData$batchList[[1]], .data = gs,
  pathProject = pathProject, marker = exampleData$marker[1]
)
```

For threshold diagnostics, see `?getStimGatesDetailed`. To export positive cells as FCS files, see `?writeStimFCS`. Keep `pathProject` to read saved gates and statistics in later sessions.
