---
title: Simulated transformation scenarios
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
  simulation_seed: 12345
---

This analysis checks the expression distributions used in the bandwidth simulations. One unstimulated and one stimulated sample are simulated under Gaussian, skewed and gamma transformations, with low and high positive-component means chosen separately for each transformation. The density plots show how the rare responding cells sit relative to the negative population.

## Set-up

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(ggplot2)
  library(tibble)
  library(dplyr)
  library(purrr)
  library(cowplot)
})

if (requireNamespace("projr", quietly = TRUE)) {
  root_dir <- tryCatch(
    projr::projr_path_get("project"),
    error = function(e) getwd()
  )
} else {
  root_dir <- getwd()
}

# Quarto may start in analysis/ when no project execution directory is set.
if (!file.exists(file.path(root_dir, "DESCRIPTION"))) {
  root_dir <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
}
if (!file.exists(file.path(root_dir, "DESCRIPTION"))) {
  stop("Cannot find the stimgate repository root; render from the checkout.")
}

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1)) &&
    requireNamespace("devtools", quietly = TRUE)
) {
  devtools::load_all(root_dir, quiet = TRUE)
} else {
  r_dir <- file.path(root_dir, "R")
  if (dir.exists(r_dir)) {
    fn_vec <- list.files(r_dir, full.names = TRUE, pattern = "[.]R$")
    for (fn in fn_vec) {
      source(fn)
    }
  }
}
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-trans.R for helper utilities.
for (fn in c("sim-misc.R", "sim-trans.R")) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  n_cell = 1e5,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5
)

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl() |>
  dplyr::mutate(
    prob_response = 0.002
  )

# Single definition of the transformation order, used for simulation and plots.
transformation_vec <- c("gaussian", "skew", "gamma")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

main_settings
```

The figures below simulate one unstimulated and one stimulated condition for each selected transformation-specific `mean_pos` setting. The stimulated sample has an excess response probability of `r mean_pos_settings_tbl$prob_response[[1]]`, while the unstimulated sample has background response probability equal to `prob_response * background_relative_to_response`.

## Simulation and collation

Each grid row has its own seed. The cached expression table records the scientific settings so a later plotting render cannot silently use a different configuration.

```{r}
#| label: simulate-mean-pos-grid-univariate
#| include: false

dir_trans <- .analysis_cache_dir(c("sim", "trans"), root_dir, create = run_simulations)
path_uni_tbl <- file.path(dir_trans, "uni_tbl.rds")
cache_settings <- sim_trans_cache_settings(
  main_settings,
  mean_pos_settings_tbl,
  transformation_vec,
  simulation_seed
)

run_trans_row <- function(
    transformation, mean_pos_setting, mean_pos, prob_response, row_index) {
  .analysis_with_seed(as.integer(simulation_seed + row_index - 1L), {
    sim_trans_set_row_seed(simulation_seed, row_index)
    sim_trans_univariate_experiment_one(
      transformation = transformation,
      mean_pos = mean_pos,
      prob_response = prob_response,
      settings = main_settings
    ) |>
      dplyr::mutate(mean_pos_setting = mean_pos_setting) |>
      dplyr::mutate(prob_response = prob_response)
  })
}
uni_grid <- dplyr::mutate(
  mean_pos_settings_tbl,
  row_index = dplyr::row_number()
)

if (isTRUE(run_simulations)) {
  uni_tbl <- purrr::pmap_dfr(
    uni_grid,
    run_trans_row
  ) |>
    dplyr::mutate(
      transformation = factor(
        .data$transformation,
        levels = transformation_vec
      ),
      mean_pos_setting = factor(
        .data$mean_pos_setting,
        levels = c("low", "high")
      ),
      mean_pos_lab = paste0(
        .data$mean_pos_setting,
        " (mean_pos = ",
        .data$mean_pos,
        ")"
      ),
      prob_response = factor(
        .data$prob_response,
        levels = sort(unique(.data$prob_response))
      ),
      condition = factor(
        .data$condition,
        levels = c("stimulated", "unstimulated")
      )
    )

  sim_trans_write_cache(uni_tbl, cache_settings, path_uni_tbl)
} else {
  if (!file.exists(path_uni_tbl)) {
    stop(
      "No transformation cache found at: ", path_uni_tbl,
      ". Run RUN_SIMULATIONS=true RUN_PLOTS=false quarto render ",
      "analysis/1-sim-trans.qmd first."
    )
  }
  uni_tbl <- sim_trans_read_cache(path_uni_tbl, cache_settings)
}

```

```{r}
#| label: rerun-one-simulation
#| eval: false
#| include: false
sim_id_target <- 1L
purrr::pmap_dfr(
  uni_grid |> dplyr::filter(.data$row_index == sim_id_target),
  run_trans_row
)
```

## Results

### Mean-positive transformation grid

Rows show the low and high positive-component means; columns show the transformations. Colour distinguishes stimulated and unstimulated samples, and the square-root density axis makes rare components easier to see. Vertical lines mark realised component means, while the rug marks positive-component cells; compare their separation from the negative density.

```{r}
#| label: fig-mean-pos-grid-univariate
#| echo: false
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Simulated expression distributions at low and high positive-component means."
if (isTRUE(run_plots)) {
  gamma_range_vec <- uni_tbl |>
    dplyr::filter(
      transformation == "gamma"
    ) |>
    dplyr::pull(F1) |>
    range()
  from_gamma <- gamma_range_vec[1] - 0.05 * abs(diff(gamma_range_vec))
  gamma_range <- c(from_gamma, gamma_range_vec[2])
  dens_tbl <- make_density_tbl(uni_tbl, gamma_range)

  mean_line_tbl <- uni_tbl |>
    dplyr::filter(
      (
        .data$condition == "unstimulated" &
          .data$response_class == "gn"
      ) |
        (
          .data$condition == "stimulated" &
            .data$response_class == "gp"
        )
    ) |>
    dplyr::mutate(
      component = dplyr::case_when(
        .data$condition == "unstimulated" &
          .data$response_class == "gn" ~ "negative mean",
        .data$condition == "stimulated" &
          .data$response_class == "gp" ~ "positive mean",
        TRUE ~ NA_character_
      )
    ) |>
    dplyr::group_by(
      .data$mean_pos_setting,
      .data$mean_pos,
      .data$mean_pos_lab,
      .data$transformation,
      .data$component
    ) |>
    dplyr::summarise(
      F1 = mean(.data$F1, na.rm = TRUE),
      .groups = "drop"
    )

  rug_tbl <- uni_tbl |>
    dplyr::filter(.data$response_class == "gp")

  p <- ggplot(
    dens_tbl,
    aes(
      x = .data$F1,
      y = .data$density,
      colour = .data$condition,
      fill = .data$condition
    )
  ) +
    geom_hline(yintercept = 0, linewidth = 0.25, colour = "gray25") +
    geom_area(
      alpha = 0.18,
      position = "identity"
    ) +
    geom_line(linewidth = 0.25, alpha = 0.1) +
    geom_vline(
      data = mean_line_tbl,
      aes(
        xintercept = .data$F1,
        linetype = .data$component,
        group = .data$component
      ),
      inherit.aes = FALSE,
      linewidth = 0.6,
      colour = "gray25"
    ) +
    geom_rug(
      data = rug_tbl,
      aes(
        x = .data$F1,
        colour = .data$condition
      ),
      inherit.aes = FALSE,
      sides = "b",
      alpha = 0.25,
      linewidth = 0.15
    ) +
    facet_grid(
      rows = vars(.data$mean_pos_setting),
      cols = vars(.data$transformation),
      scales = "free_x",
      labeller = labeller(
        transformation = .simMiscGetTransPretty(),
        mean_pos_setting = .simMiscGetMeanPosPretty()
      )
    ) +
    scale_y_sqrt() +
    labs(
      x = "Marker expression",
      y = "Density, square-root scale",
      colour = NULL,
      fill = NULL,
      linetype = NULL
    ) +
    cowplot::theme_cowplot(
      font_size = 12
    ) +
    cowplot::background_grid(major = "x", minor = "none") +
    theme(
      strip.background = element_rect(fill = "white", colour = "gray20"),
    ) +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA),
      strip.text.y = element_text(angle = 0),
      # --- Legend Centering Tweaks ---
      legend.position = "bottom",
      legend.box = "horizontal",
      legend.box.just = "center",
      legend.justification = "center"
    ) +
    coord_cartesian(ylim = c(0, 0.5)) +
    scale_colour_manual(
      values = .simMiscGetStimColVec()
    )

  ext_vec <- c("png", "pdf")
  for (ext in ext_vec) {
    path_p <- file.path(
      dir_trans,
      paste0("mean-pos-grid-univariate.", ext)
    )
    ggsave(
      filename = path_p,
      plot = p,
      width = 30,
      height = 20,
      units = "cm",
      dpi = 300
    )
  }

  print(p)
}
```

### Component counts

Counts and proportions show the realised numbers of negative and positive cells in each condition.

```{r}
#| label: univariate-counts-by-mean-pos
#| echo: false
uni_tbl |>
  dplyr::count(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$transformation,
    .data$condition,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$transformation,
    .data$condition
  ) |>
  dplyr::mutate(
    prop_cell = .data$n_cell / sum(.data$n_cell)
  ) |>
  dplyr::ungroup()
```
