---
title: Simulated transformation scenarios
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(ggplot2)
  library(tibble)
  library(dplyr)
  library(purrr)
  library(cowplot)
})

if (requireNamespace("projr", quietly = TRUE)) {
  root_dir <- tryCatch(
    projr::projr_path_get("project"),
    error = function(e) getwd()
  )
} else {
  root_dir <- getwd()
}

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1)) &&
    requireNamespace("devtools", quietly = TRUE)
) {
  devtools::load_all(root_dir, quiet = TRUE)
} else {
  r_dir <- file.path(root_dir, "R")
  if (dir.exists(r_dir)) {
    fn_vec <- list.files(r_dir, full.names = TRUE, pattern = "[.]R$")
    for (fn in fn_vec) {
      source(fn)
    }
  }
}
```

## Aim

This page checks that the simulated expression distributions look sensible under the three transformation settings used in the bandwidth simulations. The main scenario is:

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  mean_pos = 8,
  n_cell = 1e5,
  prob_response = 0.002,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 2,
  cov_ev_max = 2
)

main_settings
```

The figures below simulate only one stimulated condition at a time. For the univariate plots, the matching unstimulated condition is also shown for comparison.

```{r}
#| label: helpers
#| include: false

transformation_vec <- c("gaussian", "gamma", "skew")

```

## Main univariate transformation check

The first figure shows the univariate scenario for the main settings. The stimulated sample has an excess response probability of `r main_settings$prob_response[[1]]`, while the unstimulated sample has background response probability equal to `prob_response * background_relative_to_response`.
Positive cells are also shown as rug marks along the x-axis.

```{r}
#| label: simulate-main-univariate
#| fig-width: 12
#| fig-height: 4.8
#| fig-cap: "Univariate simulated expression under the three transformations. The y-axis is square-root transformed to make the rare response component easier to see. Dotted vertical lines mark the transformed negative and positive component means."

set.seed(1)

uni_tbl <- purrr::map_dfr(
  transformation_vec,
  sim_trans_univariate_one
)

uni_plot_list <- purrr::map(
  transformation_vec,
  ~ plot_univariate_transformation(uni_tbl, .x)
)

p <- cowplot::plot_grid(
  plotlist = uni_plot_list,
  nrow = 1,
  align = "hv"
)
path_p <- projr::projr_path_get(
  "cache",
  "sim",
  "trans",
  "main-univariate.png"
)
ggsave(
  filename = path_p,
  plot = p,
  width = 30,
  height = 15,
  units = "cm",
  dpi = 300
)
p
```

```{r}
#| label: univariate-counts
uni_tbl |>
  dplyr::count(
    .data$transformation,
    .data$condition,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(.data$transformation, .data$condition) |>
  dplyr::mutate(prop_cell = .data$n_cell / sum(.data$n_cell)) |>
  dplyr::ungroup()
```

## Extra 2-by-3 transformation check

The next figure keeps the univariate comparison in the top row and adds a bivariate stimulated condition in the bottom row. In the bivariate stimulated condition, the excess response probability is split equally across F1-only, F2-only and F1/F2 co-expressing cells. The background response probability is also split equally across those three response classes.

```{r}
#| label: simulate-main-univariate-and-bivariate
#| fig-width: 12
#| fig-height: 9
#| fig-cap: "Top row: univariate stimulated versus unstimulated simulations. Bottom row: bivariate stimulated simulations with equal probability assigned to F1-only, F2-only and F1/F2 co-expressing response classes. For the bivariate row, background cells are down-sampled for display and all response cells are retained."

set.seed(2)

bi_tbl <- purrr::map_dfr(
  transformation_vec,
  sim_trans_bivariate_one
)

top_row <- purrr::map(
  transformation_vec,
  ~ plot_univariate_transformation(uni_tbl, .x) +
    theme(legend.position = "none")
)

bottom_row <- purrr::map(
  transformation_vec,
  ~ plot_bivariate_transformation(bi_tbl, .x) +
    theme(legend.position = "none")
)

p <- cowplot::plot_grid(
  plotlist = c(top_row, bottom_row),
  ncol = 3,
  align = "hv"
)
path_p <- projr::projr_path_get(
  "cache",
  "sim",
  "trans",
  "main-univariate-and-bivariate.png"
)
ggsave(
  filename = path_p,
  plot = p,
  width = 30,
  height = 22.5,
  units = "cm",
  dpi = 300
)
p
```

```{r}
#| label: bivariate-counts

bi_tbl |>
  dplyr::count(
    .data$transformation,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(.data$transformation) |>
  dplyr::mutate(prop_cell = .data$n_cell / sum(.data$n_cell)) |>
  dplyr::ungroup()
```
