---
title: Simulated transformation scenarios
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
  simulation_seed: 12345
---

This page checks that the simulated expression distributions look sensible under the three transformation settings used in the bandwidth simulations. Here, `mean_pos` is chosen separately for each transformation, with low and high settings selected after inspecting the transformed distributions.

## Set-up

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(ggplot2)
  library(tibble)
  library(dplyr)
  library(purrr)
  library(cowplot)
})

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev

run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1)) &&
    requireNamespace("devtools", quietly = TRUE)
) {
  devtools::load_all(root_dir, quiet = TRUE)
} else {
  r_dir <- file.path(root_dir, "R")
  if (dir.exists(r_dir)) {
    fn_vec <- list.files(r_dir, full.names = TRUE, pattern = "[.]R$")
    for (fn in fn_vec) {
      source(fn)
    }
  }
}
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-trans.R for helper utilities.
for (fn in c("sim-misc.R", "sim-trans.R")) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  n_cell = 1e5,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5
)

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl() |>
  dplyr::mutate(
    prob_response = 0.002
  )

# Single definition of the transformation order, used for simulation and plots.
transformation_vec <- c("gaussian", "skew", "gamma")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

main_settings
```

The figures below simulate one unstimulated and one stimulated condition for each selected transformation-specific `mean_pos` setting. The stimulated sample has an excess response probability of `r mean_pos_settings_tbl$prob_response[[1]]`, while the unstimulated sample has background response probability equal to `prob_response * background_relative_to_response`.

## Simulation

Each grid row has its own seed. The cached expression table records the scientific settings so a later plotting render cannot silently use a different configuration.

```{r}
#| label: simulate-mean-pos-grid-univariate
#| include: false

analysis_key <- c("sim", "trans")
analysis_key <- .analysis_mode_key(analysis_key)
dir_trans <- .analysis_cache_dir(analysis_key, root_dir, create = run_simulations)
path_uni_tbl <- file.path(dir_trans, "uni_tbl.rds")

run_trans_row <- function(
    transformation, mean_pos_setting, mean_pos, prob_response, row_index) {
  .analysis_with_seed(as.integer(simulation_seed + row_index - 1L), {
    sim_trans_set_row_seed(simulation_seed, row_index)
    sim_trans_univariate_experiment_one(
      transformation = transformation,
      mean_pos = mean_pos,
      prob_response = prob_response,
      settings = main_settings_run
    ) |>
      dplyr::mutate(mean_pos_setting = mean_pos_setting) |>
      dplyr::mutate(prob_response = prob_response)
  })
}
uni_grid <- dplyr::mutate(
  mean_pos_settings_tbl,
  row_index = dplyr::row_number()
)

uni_grid_full <- uni_grid
main_settings_run <- main_settings
# Quick mode keeps transformation (including gamma), mean position and both conditions/components.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  uni_grid <- uni_grid |>
    dplyr::filter(transformation %in% c("gaussian", "gamma"))
  main_settings_run <- main_settings_run |>
    dplyr::mutate(n_cell = 5e3)
}
cache_settings <- c(
  sim_trans_cache_settings(
    main_settings_run, dplyr::select(uni_grid, -row_index),
    transformation_vec, simulation_seed
  ),
  list(analysis_quick = analysis_quick, analysis_dev = analysis_dev)
)

if (isTRUE(run_simulations)) {
  uni_tbl <- purrr::pmap_dfr(
    uni_grid,
    run_trans_row
  ) |>
    dplyr::mutate(
      transformation = factor(
        .data$transformation,
        levels = transformation_vec
      ),
      mean_pos_setting = factor(
        .data$mean_pos_setting,
        levels = c("low", "high")
      ),
      mean_pos_lab = paste0(
        .data$mean_pos_setting,
        " (mean_pos = ",
        .data$mean_pos,
        ")"
      ),
      prob_response = factor(
        .data$prob_response,
        levels = sort(unique(.data$prob_response))
      ),
      condition = factor(
        .data$condition,
        levels = c("stimulated", "unstimulated")
      )
    )

  sim_trans_write_cache(uni_tbl, cache_settings, path_uni_tbl)
}
```

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L
purrr::pmap_dfr(
  uni_grid_full |> dplyr::filter(.data$row_index == sim_id_target),
  run_trans_row
)
```

## Collation

Read the matching cached expression table when simulations are disabled.

```{r}
#| label: transformation-collate
#| include: false
if (!isTRUE(run_simulations)) {
  uni_tbl <- sim_trans_read_cache(path_uni_tbl, cache_settings)
}

```


## Results

### Mean-positive transformation grid

Each row corresponds to the low or high transformation-specific `mean_pos` setting, and each column corresponds to one transformation. The y-axis is square-root transformed to make rare response components easier to see. Vertical lines (line type distinguishes the component) mark the realised negative component mean in the unstimulated condition and the realised positive component mean in the stimulated condition. The rug along the bottom axis shows the individual positive-component cells.

```{r}
#| label: fig-mean-pos-grid-univariate
#| echo: false
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Simulated expression distributions at low and high positive-component means."
if (isTRUE(run_plots)) {
  gamma_range_vec <- uni_tbl |>
    dplyr::filter(
      transformation == "gamma"
    ) |>
    dplyr::pull(F1) |>
    range()
  from_gamma <- gamma_range_vec[1] - 0.05 * abs(diff(gamma_range_vec))
  gamma_range <- c(from_gamma, gamma_range_vec[2])
  dens_tbl <- make_density_tbl(uni_tbl, gamma_range)

  mean_line_tbl <- uni_tbl |>
    dplyr::filter(
      (
        .data$condition == "unstimulated" &
          .data$response_class == "gn"
      ) |
        (
          .data$condition == "stimulated" &
            .data$response_class == "gp"
        )
    ) |>
    dplyr::mutate(
      component = dplyr::case_when(
        .data$condition == "unstimulated" &
          .data$response_class == "gn" ~ "negative mean",
        .data$condition == "stimulated" &
          .data$response_class == "gp" ~ "positive mean",
        TRUE ~ NA_character_
      )
    ) |>
    dplyr::group_by(
      .data$mean_pos_setting,
      .data$mean_pos,
      .data$mean_pos_lab,
      .data$transformation,
      .data$component
    ) |>
    dplyr::summarise(
      F1 = mean(.data$F1, na.rm = TRUE),
      .groups = "drop"
    )

  rug_tbl <- uni_tbl |>
    dplyr::filter(.data$response_class == "gp")

  p <- ggplot(
    dens_tbl,
    aes(
      x = .data$F1,
      y = .data$density,
      colour = .data$condition,
      fill = .data$condition
    )
  ) +
    geom_hline(yintercept = 0, linewidth = 0.25, colour = "gray25") +
    geom_area(
      alpha = 0.18,
      position = "identity"
    ) +
    geom_line(linewidth = 0.25, alpha = 0.1) +
    geom_vline(
      data = mean_line_tbl,
      aes(
        xintercept = .data$F1,
        linetype = .data$component,
        group = .data$component
      ),
      inherit.aes = FALSE,
      linewidth = 0.6,
      colour = "gray25"
    ) +
    geom_rug(
      data = rug_tbl,
      aes(
        x = .data$F1,
        colour = .data$condition
      ),
      inherit.aes = FALSE,
      sides = "b",
      alpha = 0.25,
      linewidth = 0.15
    ) +
    facet_grid(
      rows = vars(.data$mean_pos_setting),
      cols = vars(.data$transformation),
      scales = "free_x",
      labeller = labeller(
        transformation = .simMiscGetTransPretty(),
        mean_pos_setting = .simMiscGetMeanPosPretty()
      )
    ) +
    scale_y_sqrt() +
    labs(
      x = "Marker expression",
      y = "Density, square-root scale",
      colour = NULL,
      fill = NULL,
      linetype = NULL
    ) +
    cowplot::theme_cowplot(
      font_size = 12
    ) +
    cowplot::background_grid(major = "x", minor = "none") +
    theme(
      strip.background = element_rect(fill = "white", colour = "gray20"),
    ) +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA),
      strip.text.y = element_text(angle = 0),
      # --- Legend Centering Tweaks ---
      legend.position = "bottom",
      legend.box = "horizontal",
      legend.box.just = "center",
      legend.justification = "center"
    ) +
    coord_cartesian(ylim = c(0, 0.5)) +
    scale_colour_manual(
      values = .simMiscGetStimColVec()
    )

  ext_vec <- c("png", "pdf")
  for (ext in ext_vec) {
    path_p <- file.path(
      dir_trans,
      paste0("mean-pos-grid-univariate.", ext)
    )
    ggsave(
      filename = path_p,
      plot = p,
      width = 30,
      height = 20,
      units = "cm",
      dpi = 300
    )
  }

  print(p)
}
```

### Component counts

Counts and proportions show the realised numbers of negative and positive cells in each condition.

```{r}
#| label: univariate-counts-by-mean-pos
#| echo: false
uni_tbl |>
  dplyr::count(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$transformation,
    .data$condition,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$transformation,
    .data$condition
  ) |>
  dplyr::mutate(
    prop_cell = .data$n_cell / sum(.data$n_cell)
  ) |>
  dplyr::ungroup()
```
