---
title: Simulated transformation scenarios
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(ggplot2)
  library(tibble)
  library(dplyr)
  library(purrr)
  library(cowplot)
})

if (requireNamespace("projr", quietly = TRUE)) {
  root_dir <- tryCatch(
    projr::projr_path_get("project"),
    error = function(e) getwd()
  )
} else {
  root_dir <- getwd()
}

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1)) &&
    requireNamespace("devtools", quietly = TRUE)
) {
  devtools::load_all(root_dir, quiet = TRUE)
} else {
  r_dir <- file.path(root_dir, "R")
  if (dir.exists(r_dir)) {
    fn_vec <- list.files(r_dir, full.names = TRUE, pattern = "[.]R$")
    for (fn in fn_vec) {
      source(fn)
    }
  }
}
```

## Aim

This page checks that the simulated expression distributions look sensible under the three transformation settings used in the bandwidth simulations. Here, `mean_pos` is varied separately from 4 to 12 so that we can choose low, medium, and high separation settings for each transformation.

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  n_cell = 1e5,
  prob_response = 0.002,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 2,
  cov_ev_max = 2
)

mean_pos_vec <- 4:12
transformation_vec <- c("gaussian", "gamma", "skew")

main_settings
````

The figures below simulate one unstimulated and one stimulated condition for each combination of `mean_pos` and transformation. The stimulated sample has an excess response probability of `r main_settings$prob_response[[1]]`, while the unstimulated sample has background response probability equal to `prob_response * background_relative_to_response`.

```{r}
#| label: helpers
#| include: false

sim_trans_univariate_one <- function(
  transformation,
  mean_pos,
  settings = main_settings
) {
  n_cell <- settings$n_cell[[1]]
  prob_response <- settings$prob_response[[1]]
  background_relative_to_response <-
    settings$background_relative_to_response[[1]]

  prob_response_uns <- prob_response * background_relative_to_response

  transformation_fun <- .simMiscGetTrans(transformation)
  attr(transformation_fun, "sim_transformation") <- transformation

  out <- simCytExperiment(
    nSample = 1L,
    nMarker = 1L,
    nCondition = 2L,
    nCluster = 2L,
    nCellByCondition = c(n_cell, n_cell),
    transformationFunc = transformation_fun,
    mixtureType = settings$mixture_type[[1]],
    meanExprMat = matrix(
      c(0, mean_pos),
      byrow = TRUE,
      ncol = 1
    ),
    clusterLabelVec = c("gn", "gp"),
    probVecUns = c(
      1 - prob_response_uns,
      prob_response_uns
    ),
    probExact = settings$prob_exact[[1]],
    probResponseVecByStimCondition = list(
      c(-prob_response, prob_response)
    ),
    samplePerturbationSd = 0,
    conditionPerturbationSd = 0,
    clusterPerturbationSd = settings$cluster_perturbation_sd[[1]],
    covEvMin = settings$cov_ev_min[[1]],
    covEvMax = settings$cov_ev_max[[1]]
  )

  flowFrameList <- out$flowFrameList
  labelsList <- out$labelsList

  purrr::map_dfr(seq_along(flowFrameList), function(ind) {
    tibble::tibble(
      transformation = transformation,
      mean_pos = mean_pos,
      condition = c("unstimulated", "stimulated")[[ind]],
      response_class = labelsList[[ind]],
      F1 = as.numeric(flowCore::exprs(flowFrameList[[ind]])[, "F1"])
    )
  })
}

make_density_tbl <- function(.data, n = 2048) {
  .data |>
    dplyr::group_by(
      .data$mean_pos,
      .data$transformation,
      .data$condition
    ) |>
    dplyr::group_modify(function(.x, .y) {
      x <- .x$F1
      x <- x[is.finite(x)]

      if (length(x) < 2L || length(unique(x)) < 2L) {
        return(tibble::tibble(F1 = numeric(), density = numeric()))
      }

      dens <- stats::density(x, n = n)

      tibble::tibble(
        F1 = dens$x,
        density = dens$y
      )
    }) |>
    dplyr::ungroup()
}
```

## Mean-positive transformation grid

Each row corresponds to one latent `mean_pos` value, and each column corresponds to one transformation. The y-axis is square-root transformed to make rare response components easier to see. Dotted vertical lines mark the realised negative component mean in the unstimulated condition and the realised positive component mean in the stimulated condition.

```{r}
#| label: simulate-mean-pos-grid-univariate
#| fig-width: 12
#| fig-height: 18
#| fig-cap: "Univariate simulated expression for mean_pos from 4 to 12 across transformations. The y-axis is square-root transformed to make the rare response component easier to see."

set.seed(1)

uni_grid <- tidyr::expand_grid(
  mean_pos = mean_pos_vec,
  transformation = transformation_vec
)

uni_tbl <- purrr::pmap_dfr(
  uni_grid,
  function(mean_pos, transformation) {
    sim_trans_univariate_one(
      transformation = transformation,
      mean_pos = mean_pos,
      settings = main_settings
    )
  }
) |>
  dplyr::mutate(
    transformation = factor(
      .data$transformation,
      levels = transformation_vec
    ),
    mean_pos_lab = factor(
      paste0("mean_pos = ", .data$mean_pos),
      levels = paste0("mean_pos = ", mean_pos_vec)
    ),
    condition = factor(
      .data$condition,
      levels = c("stimulated", "unstimulated")
    )
  )

dens_tbl <- make_density_tbl(uni_tbl) |>
  dplyr::mutate(
    transformation = factor(
      .data$transformation,
      levels = transformation_vec
    ),
    mean_pos_lab = factor(
      paste0("mean_pos = ", .data$mean_pos),
      levels = paste0("mean_pos = ", mean_pos_vec)
    ),
    condition = factor(
      .data$condition,
      levels = c("stimulated", "unstimulated")
    )
  )

mean_line_tbl <- uni_tbl |>
  dplyr::filter(
    (
      .data$condition == "unstimulated" &
        .data$response_class == "gn"
    ) |
      (
        .data$condition == "stimulated" &
          .data$response_class == "gp"
      )
  ) |>
  dplyr::mutate(
    component = dplyr::case_when(
      .data$condition == "unstimulated" &
        .data$response_class == "gn" ~ "negative mean",
      .data$condition == "stimulated" &
        .data$response_class == "gp" ~ "positive mean",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::group_by(
    .data$mean_pos,
    .data$mean_pos_lab,
    .data$transformation,
    .data$component
  ) |>
  dplyr::summarise(
    F1 = mean(.data$F1, na.rm = TRUE),
    .groups = "drop"
  )

rug_tbl <- uni_tbl |>
  dplyr::filter(.data$response_class == "gp")

p <- ggplot(
  dens_tbl,
  aes(
    x = .data$F1,
    y = .data$density,
    colour = .data$condition,
    fill = .data$condition
  )
) +
  geom_area(
    alpha = 0.18,
    position = "identity"
  ) +
  geom_line(linewidth = 0.25) +
  geom_vline(
    data = mean_line_tbl,
    aes(
      xintercept = .data$F1,
      linetype = .data$component
    ),
    inherit.aes = FALSE,
    linewidth = 0.3
  ) +
  geom_rug(
    data = rug_tbl,
    aes(
      x = .data$F1,
      colour = .data$condition
    ),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.25,
    linewidth = 0.15
  ) +
  facet_grid(
    rows = vars(.data$mean_pos_lab),
    cols = vars(.data$transformation),
    scales = "free_x"
  ) +
  scale_y_sqrt() +
  labs(
    x = "F1 expression",
    y = "Density, square-root scale",
    colour = NULL,
    fill = NULL,
    linetype = NULL
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy", minor = "none") +
  theme(
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    strip.text.y = element_text(angle = 0)
  ) +
  coord_cartesian(ylim = c(0, 1))

path_p <- projr::projr_path_get(
  "output",
  "sim",
  "trans",
  "mean-pos-grid-univariate.png"
)

ggsave(
  filename = path_p,
  plot = p,
  width = 30,
  height = 40,
  units = "cm",
  dpi = 300
)

p
```

```{r}
#| label: univariate-counts-by-mean-pos
uni_tbl |>
  dplyr::count(
    .data$mean_pos,
    .data$transformation,
    .data$condition,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(
    .data$mean_pos,
    .data$transformation,
    .data$condition
  ) |>
  dplyr::mutate(
    prop_cell = .data$n_cell / sum(.data$n_cell)
  ) |>
  dplyr::ungroup()
```


### Test scenarios

```{r}

```

This page checks that the simulated expression distributions look sensible under the three transformation settings used in the bandwidth simulations. The main scenario is:

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  mean_pos = 4,
  n_cell = 1e5,
  prob_response = 0.002,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 2,
  cov_ev_max = 2
)

main_settings
```

The figures below simulate only one stimulated condition at a time. For the univariate plots, the matching unstimulated condition is also shown for comparison.

```{r}
#| label: helpers
#| include: false
transformation_vec <- c("gaussian", "gamma", "skew")
```

## Main univariate transformation check

The first figure shows the univariate scenario for the main settings. The stimulated sample has an excess response probability of `r main_settings$prob_response[[1]]`, while the unstimulated sample has background response probability equal to `prob_response * background_relative_to_response`.
Positive cells are also shown as rug marks along the x-axis.

```{r}
#| label: simulate-main-univariate
#| fig-width: 12
#| fig-height: 4.8
#| fig-cap: "Univariate simulated expression under the three transformations. The y-axis is square-root transformed to make the rare response component easier to see. Dotted vertical lines mark the transformed negative and positive component means."

set.seed(1)

uni_tbl <- purrr::map_dfr(
  transformation_vec,
  sim_trans_univariate_one
)

uni_plot_list <- purrr::map(
  transformation_vec,
  ~ plot_univariate_transformation(uni_tbl, .x)
)

p <- cowplot::plot_grid(
  plotlist = uni_plot_list,
  nrow = 1,
  align = "hv"
)
path_p <- projr::projr_path_get(
  "output",
  "sim",
  "trans",
  "main-univariate.png"
)
ggsave(
  filename = path_p,
  plot = p,
  width = 30,
  height = 15,
  units = "cm",
  dpi = 300
)
p
```

```{r}
#| label: univariate-counts
uni_tbl |>
  dplyr::count(
    .data$transformation,
    .data$condition,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(.data$transformation, .data$condition) |>
  dplyr::mutate(prop_cell = .data$n_cell / sum(.data$n_cell)) |>
  dplyr::ungroup()
```

## Extra 2-by-3 transformation check

The next figure keeps the univariate comparison in the top row and adds a bivariate stimulated condition in the bottom row. In the bivariate stimulated condition, the excess response probability is split equally across F1-only, F2-only and F1/F2 co-expressing cells. The background response probability is also split equally across those three response classes.

```{r}
#| label: simulate-main-univariate-and-bivariate
#| fig-width: 12
#| fig-height: 9
#| fig-cap: "Top row: univariate stimulated versus unstimulated simulations. Bottom row: bivariate stimulated simulations with equal probability assigned to F1-only, F2-only and F1/F2 co-expressing response classes. For the bivariate row, background cells are down-sampled for display and all response cells are retained."

set.seed(2)

bi_tbl <- purrr::map_dfr(
  transformation_vec,
  sim_trans_bivariate_one
)

top_row <- purrr::map(
  transformation_vec,
  ~ plot_univariate_transformation(uni_tbl, .x) +
    theme(legend.position = "none")
)

bottom_row <- purrr::map(
  transformation_vec,
  ~ plot_bivariate_transformation(bi_tbl, .x) +
    theme(legend.position = "none")
)

p <- cowplot::plot_grid(
  plotlist = c(top_row, bottom_row),
  ncol = length(transformation_vec),
  align = "hv"
)
path_p <- projr::projr_path_get(
  "output",
  "sim",
  "trans",
  "main-univariate-and-bivariate.png"
)
ggsave(
  filename = path_p,
  plot = p,
  width = 30,
  height = 22.5,
  units = "cm",
  dpi = 300
)
p
```

```{r}
#| label: bivariate-counts

bi_tbl |>
  dplyr::count(
    .data$transformation,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(.data$transformation) |>
  dplyr::mutate(prop_cell = .data$n_cell / sum(.data$n_cell)) |>
  dplyr::ungroup()
```
