---
title: Simulated transformation scenarios
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(ggplot2)
  library(tibble)
  library(dplyr)
  library(purrr)
  library(cowplot)
})

if (requireNamespace("projr", quietly = TRUE)) {
  root_dir <- tryCatch(
    projr::projr_path_get("project"),
    error = function(e) getwd()
  )
} else {
  root_dir <- getwd()
}

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1)) &&
    requireNamespace("devtools", quietly = TRUE)
) {
  devtools::load_all(root_dir, quiet = TRUE)
} else {
  r_dir <- file.path(root_dir, "R")
  if (dir.exists(r_dir)) {
    fn_vec <- list.files(r_dir, full.names = TRUE, pattern = "[.]R$")
    for (fn in fn_vec) {
      source(fn)
    }
  }
}
scripts_r_dir <- file.path(root_dir, "scripts", "r")
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-trans.R for helper utilities.
for (fn in c("sim-misc.R", "functionsForBenchmarking-Cyt.R", "sim-trans.R")) {
  source(file.path(scripts_r_dir, fn))
}
```

## Aim

This page checks that the simulated expression distributions look sensible under the three transformation settings used in the bandwidth simulations. Here, `mean_pos` is chosen separately for each transformation, with low and high settings selected after inspecting the transformed distributions.

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  n_cell = 1e5,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5
)

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl() |>
  dplyr::mutate(
    prob_response = 0.002
  )

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

main_settings
```

The figures below simulate one unstimulated and one stimulated condition for each selected transformation-specific `mean_pos` setting. The stimulated sample has an excess response probability of `r mean_pos_settings_tbl$prob_response[[1]]`, while the unstimulated sample has background response probability equal to `prob_response * background_relative_to_response`.

```{r}
#| label: helpers
#| include: false
```

## Mean-positive transformation grid

Each row corresponds to the low or high transformation-specific `mean_pos` setting, and each column corresponds to one transformation. The y-axis is square-root transformed to make rare response components easier to see. Dotted vertical lines mark the realised negative component mean in the unstimulated condition and the realised positive component mean in the stimulated condition.

```{r}
#| label: simulate-mean-pos-grid-univariate
#| fig-width: 12
#| fig-height: 18
#| fig-cap: "Univariate simulated expression for transformation-specific low and high mean_pos settings. The y-axis is square-root transformed to make the rare response component easier to see."

set.seed(1)

uni_grid <- mean_pos_settings_tbl

uni_tbl <- purrr::pmap_dfr(
  uni_grid,
  function(transformation, mean_pos_setting, mean_pos, prob_response) {
    sim_trans_univariate_experiment_one(
      transformation = transformation,
      mean_pos = mean_pos,
      prob_response = prob_response,
      settings = main_settings
    ) |>
      dplyr::mutate(mean_pos_setting = mean_pos_setting) |>
      dplyr::mutate(prob_response = prob_response)
  }
) |>
  dplyr::mutate(
    transformation = factor(
      .data$transformation,
      levels = transformation_vec
    ),
    mean_pos_setting = factor(
      .data$mean_pos_setting,
      levels = c("low", "high")
    ),
    mean_pos_lab = paste0(
      .data$mean_pos_setting,
      " (mean_pos = ",
      .data$mean_pos,
      ")"
    ),
    prob_response = factor(
      .data$prob_response,
      levels = sort(unique(.data$prob_response))
    ),
    condition = factor(
      .data$condition,
      levels = c("stimulated", "unstimulated")
    )
  )

gamma_range_vec <- uni_tbl |>
  dplyr::filter(
    transformation == "gamma"
  ) |>
  dplyr::pull(F1) |>
  range()
from_gamma <- gamma_range_vec[1] - 0.05 * abs(diff(gamma_range_vec))
gamma_range <- c(from_gamma, gamma_range_vec[2])
dens_tbl <- make_density_tbl(uni_tbl, gamma_range) |>
  dplyr::mutate(
    transformation = factor(
      .data$transformation,
      levels = c("gaussian", "gamma", "skew")
    ),
    mean_pos_setting = factor(
      .data$mean_pos_setting,
      levels = c("low", "high")
    ),
    mean_pos_lab = paste0(
      .data$mean_pos_setting,
      " (mean_pos = ",
      .data$mean_pos,
      ")"
    ),
    prob_response = factor(
      .data$prob_response,
      levels = sort(unique(.data$prob_response))
    ),
    condition = factor(
      .data$condition,
      levels = c("stimulated", "unstimulated")
    )
  )

mean_line_tbl <- uni_tbl |>
  dplyr::filter(
    (
      .data$condition == "unstimulated" &
        .data$response_class == "gn"
    ) |
      (
        .data$condition == "stimulated" &
          .data$response_class == "gp"
      )
  ) |>
  dplyr::mutate(
    component = dplyr::case_when(
      .data$condition == "unstimulated" &
        .data$response_class == "gn" ~ "negative mean",
      .data$condition == "stimulated" &
        .data$response_class == "gp" ~ "positive mean",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::group_by(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$mean_pos_lab,
    .data$transformation,
    .data$component
  ) |>
  dplyr::summarise(
    F1 = mean(.data$F1, na.rm = TRUE),
    .groups = "drop"
  )

rug_tbl <- uni_tbl |>
  dplyr::filter(.data$response_class == "gp")

p <- ggplot(
  dens_tbl |>
    dplyr::mutate(
      transformation = factor(
        .data$transformation,
        levels = c("gaussian", "skew", "gamma")
      ),
      mean_pos_setting = factor(
        .data$mean_pos_setting,
        levels = c("low", "high")
      )
    ),
  aes(
    x = .data$F1,
    y = .data$density,
    colour = .data$condition,
    fill = .data$condition
  )
) +
  geom_hline(yintercept = 0, linewidth = 0.25, colour = "gray25") +
  geom_area(
    alpha = 0.18,
    position = "identity"
  ) +
  geom_line(linewidth = 0.25, alpha = 0.1) +
  geom_vline(
    data = mean_line_tbl,
    aes(
      xintercept = .data$F1,
      linetype = .data$component
    ),
    inherit.aes = FALSE,
    linewidth = 0.6,
    colour = "gray25"
  ) +
  geom_rug(
    data = rug_tbl,
    aes(
      x = .data$F1,
      colour = .data$condition
    ),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.25,
    linewidth = 0.15
  ) +
  facet_grid(
    rows = vars(.data$mean_pos_setting),
    cols = vars(.data$transformation),
    scales = "free_x",
    labeller = labeller(
      transformation = .simMiscGetTransPretty(),
      mean_pos_setting = .simMiscGetMeanPosPretty()
    )
  ) +
  scale_y_sqrt() +
  labs(
    x = "Marker expression",
    y = "Density, square-root scale",
    colour = NULL,
    fill = NULL,
    linetype = NULL
  ) +
  cowplot::theme_cowplot(
    font_size = 12
  ) +
  cowplot::background_grid(major = "x", minor = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "gray20"),
  ) +
  theme(
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    strip.text.y = element_text(angle = 0),
    # --- Legend Centering Tweaks ---
    legend.position      = "bottom",
    legend.box           = "horizontal", # Places line types and colors next to each other
    legend.box.just      = "center",     # Centers the contents within the box
    legend.justification = "center"
  ) +
  coord_cartesian(ylim = c(0, 0.5)) +
  scale_colour_manual(
    values = .simMiscGetStimColVec()
  )


ext_vec <- c("png", "pdf")
for (ext in ext_vec) {
  path_p <- projr::projr_path_get(
    "cache",
    "sim",
    "trans",
    paste0("mean-pos-grid-univariate.", ext)
  )
  ggsave(
    filename = path_p,
    plot = p,
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300
  )
}
```

```{r}
#| label: univariate-counts-by-mean-pos
uni_tbl |>
  dplyr::count(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$transformation,
    .data$condition,
    .data$response_class,
    name = "n_cell"
  ) |>
  dplyr::group_by(
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$transformation,
    .data$condition
  ) |>
  dplyr::mutate(
    prop_cell = .data$n_cell / sum(.data$n_cell)
  ) |>
  dplyr::ungroup()
```
