---
title: ACS CyTOF validation outputs
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
---

Reproduce the ACS manuscript-style validation plots for StimGate, F-beta and
Tailgate using the comparison table produced by the main ACS analysis.
The figures use saved results from analysis 9 without rerunning the gating methods.

## Set-up

Failed method/marker estimates remain missing and are excluded from agreement
statistics; correlation tables state total, failed and scored counts for each
included stratum. Analysis 9 also saves coverage for all strata and an exclusion
table for unmatched cohort keys.

The saved `acs_cytof/manual-comparison/manual-comparison.rds` table is sufficient; raw FCS files are not needed here. Render analysis 9 with `RUN_SIMULATIONS=true RUN_PLOTS=false` first if that table is missing. `RUN_PLOTS=true` displays and saves the validation figures; this document does not rerun the gating methods.

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
source(file.path(scripts_r_dir, "sim-bandwidth-analysis-plot.R"))
fig_key <- .analysis_mode_key("10-real-compare-acs-cytof-validation", sized = FALSE)
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-manual.R"))
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

path_manual_output <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "manual-comparison",
  format = "absolute"
)
path_comparison <- file.path(
  path_manual_output,
  "manual-comparison.rds"
)
# Must match the StimGate threshold method set in analysis 9.
stimgate_loc_threshold_method <- "cap"
manual_run_manifest <- .analysis_read_rds(
  file.path(path_manual_output, "manifest.rds"), "acs_cytof",
  "analysis/9-real-compare-acs-cytof.qmd"
)
.analysis_check_expected_run(
  manual_run_manifest,
  "acs_cytof", "analysis/9-real-compare-acs-cytof.qmd",
  semantics_version = "acs-cytof-v8"
)
.acsCytofCheckRunManifestMethod(
  manual_run_manifest, stimgate_loc_threshold_method,
  "analysis/9-real-compare-acs-cytof.qmd"
)
manual_comparison_tbl <- .analysis_read_rds(
  path_comparison, "acs_cytof", "analysis/9-real-compare-acs-cytof.qmd"
)
.acsCytofValidateComparisonManifest(
  manual_comparison_tbl,
  locThresholdMethod = stimgate_loc_threshold_method
)
validation_methods <- .acsCytofValidationMethods()
appendix_methods <- .acsCytofValidationAppendixMethods()
.acsCytofValidationValidateComparisonTable(
  manual_comparison_tbl,
  requiredMethods = c(validation_methods, appendix_methods)
)
# The saved table also holds Tailgate and F-beta at their published defaults,
# reported in the appendix; the main results use the tuned comparators.
manual_comparison_all_tbl <- manual_comparison_tbl
manual_comparison_tbl <- .acsCytofMainComparisonTable(manual_comparison_all_tbl)
```

## Scientific inclusion rules

The raw-frequency plots show automated against manual background-subtracted
frequencies for every matched population and cytokine. The correlation outputs
follow the manuscript eligibility rule before calculating Pearson correlation
(PCC) and concordance correlation (CCC): within each population/cytokine/stimulus
stratum, the 75th percentile of the manual stimulated frequency must exceed
three times the larger of 0.01% and the median manual unstimulated frequency,
and the 75th percentile of the manual background-subtracted frequency must
exceed 0.02%. The heatmaps omit the non-secreted-protein (`p4`) stimulation,
matching the manuscript-style display. One set of heatmaps is restricted to
CD4, CD8 and TCRgd T cells and another uses every matched population. This
keeps the original manuscript comparison available while also showing the
additional populations in the current benchmark.

## Frequency-error coverage and donor uncertainty

Manual gating is an imperfect reference, so agreement is not proof of biological
accuracy. Analysis 9 provides absolute frequency-error and relative-error figures
for each stimulus. Absolute error includes zero manual net frequencies; the current
manual and automated net frequencies are truncated at zero after subtraction.
relative error needs a positive manual net frequency, and very small denominators
can dominate it. The table reports matched-row and exclusion denominators.

```{r}
#| label: validation-error-coverage
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .analysis_report_table(.acsCytofManualSummaryTable(manual_comparison_tbl),
    c(fig_key, "validation-error-coverage.csv"),
    "Matched-row, failed, scored and exclusion denominators by stratum.", path_root = root_dir)
}
```

These 95% percentile intervals estimate between-donor uncertainty of mean absolute
frequency and relative errors. All tubes belonging to a donor, including their
shared already-subtracted control, are resampled together using the same donor
draws across stimuli and methods. Finite successful estimates determine the mean;
relative error also requires a positive manual net frequency. Intervals are missing
with fewer than two contributing donors. The bootstrap does not quantify possible
bias or uncertainty in manual gating itself.

```{r}
#| label: validation-donor-uncertainty
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .analysis_report_table(.acsCytofManualUncertainty(manual_comparison_tbl),
    c(fig_key, "validation-donor-uncertainty.csv"),
    "Donor-bootstrap intervals for mean frequency and relative errors.", path_root = root_dir)
}
```

## Results

### Correlation table

```{r}
#| label: correlation-table
#| echo: false
#| results: asis
correlation_tbl <- .acsCytofValidationCorrelationTable(manual_comparison_tbl)
if (isTRUE(run_plots)) {
  .analysis_report_table(correlation_tbl, c(fig_key, "validation-correlations.csv"),
    "Eligible-stratum PCC and CCC with matched-row and exclusion counts.", path_root = root_dir)
}
```

### Unconditional percentiles of signed relative error by stimulus

These figures keep stimuli separate and include every finite signed relative
error, including exact agreement at zero. Relative error is (automated - manual)
/ manual and requires a positive manual net frequency; failed or undefined
estimates remain in the coverage table. Manual gating is an imperfect reference.
Each population panel shows all seven percentiles, with cytokine on the x axis.
The outer pair is 2.5th/97.5th, or 5th/95th throughout a figure if either outer
interval is unavailable at any finite point; the figure labels that fallback.
Quantiles are computed on raw errors before the signed display transformation.

Donor-bootstrap intervals carry all tubes for a donor together, using common
draws across methods, stimuli and strata. They describe between-donor uncertainty,
not uncertainty in manual gating, and are retained in every render.

```{r}
#| label: validation-signed-percentiles-by-stimulus
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  manual_percentiles_by_stimulus <- .acsCytofManualSignedPercentiles(
    manual_comparison_tbl, group_cols = c("method", "pop", "cyt", "stim")
  )
  for (stimulus in unique(as.character(manual_percentiles_by_stimulus$stim))) {
    .analysis_heading(paste("Stimulus:", stimulus), 4)
    for (method in validation_methods) {
      .analysis_heading(paste("Method:", .acsCytofValidationMethodLabels()[[method]]), 5)
      rows <- manual_percentiles_by_stimulus |>
        dplyr::filter(.data$stim == .env$stimulus, .data$method == .env$method)
      p <- .simBandwidthSignedPercentilePlot(
        rows, x = "cyt", x_label = "Cytokine", mcse = TRUE,
        facet = ggplot2::facet_wrap(~pop, scales = "free_y")
      )
      .analysis_print_save_fig(p,
        file.path(.analysis_fig_dir(
          c(fig_key, "signed_percentiles_by_stimulus", stimulus, method), root_dir),
          "manual-comparison-signed-percentiles.png"), height = 22)
      .analysis_report_table(rows |>
        dplyr::select(method, pop, cyt, stim, dplyr::starts_with("n_"),
          dplyr::starts_with("bootstrap_coverage_"), dplyr::starts_with("interval_available_")),
        c(fig_key, "signed_percentiles_by_stimulus", stimulus, paste0(method, ".csv")),
        "Donor-bootstrap validity for unconditional percentiles by stimulus.", path_root = root_dir)
    }
  }
}
```

### Validation figures

```{r}
#| label: save-validation-figures
#| include: false
if (isTRUE(run_plots)) {
  # The figure directory is swapped in whole, so it must not be pre-created.
  path_dir_save <- .analysis_fig_dir(
    c(fig_key, "validation_figures"), root_dir,
    create = FALSE
  )
  .acsCytofValidationSavePlots(
    comparisonTbl = manual_comparison_all_tbl,
    pathDirSave = path_dir_save
  )
}
```

Each figure shows one method. Scatterplots also have a separate figure for each population. The saved figures also cover Tailgate and F-beta at their default settings (see the appendix).

### Raw background-subtracted frequencies

Manual background-subtracted frequency (%) is on the horizontal axis and automated frequency (%) on the vertical axis. Colours and shapes identify stimuli. Each population has its own figure with cytokine panels; both axis ranges differ between panels. Points closer to the identity line indicate closer agreement; horizontal and vertical lines mark zero frequency.

```{r}
#| label: acs-validation-scatter
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (method in validation_methods) {
    .analysis_heading(
      paste0("Method: ", .acsCytofValidationMethodLabels()[[method]]), 4
    )
    method_rows <- manual_comparison_tbl |> dplyr::filter(.data$method == .env$method)
    for (population in unique(as.character(method_rows$pop))) {
      .analysis_heading(paste("Population:", population), 5)
      .analysis_print_fig(
        .acsCytofValidationPlotScatter(
          method_rows |> dplyr::filter(.data$pop == .env$population), method
        )
      )
    }
  }
}
```

### Correlation heatmaps for CD4, CD8 and TCRgd T cells

Cytokines run across columns and the three T-cell populations across rows, with panels for stimuli. Colour and text show Pearson or concordance correlation for each method after applying the eligibility rules. Values closer to one indicate stronger agreement; light-grey cells labelled `excl.` mark strata excluded by the signal eligibility rules, while white cells labelled `NA` mark eligible strata whose correlation is unavailable. Blank positions have no matched stratum.

```{r}
#| label: acs-validation-t-cell-correlations
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (method in validation_methods) {
    .analysis_heading(
      paste0("Method: ", .acsCytofValidationMethodLabels()[[method]]), 4
    )
    for (metric in c("pcc", "ccc")) {
      .analysis_heading(
        if (metric == "pcc") "Pearson correlation" else "Concordance correlation",
        5
      )
      .analysis_print_fig(.acsCytofValidationPlotCorrelation(
        correlationTbl = correlation_tbl,
        method = method,
        metric = metric,
        realPopulationsOnly = TRUE
      ))
    }
  }
}
```

### Correlation heatmaps for all matched populations

These heatmaps use the same cytokine, stimulus and correlation scales for all matched populations, including B and NK cells. Compare methods within the same population and cytokine, keeping the eligibility rules in mind.

```{r}
#| label: acs-validation-all-correlations
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (method in validation_methods) {
    .analysis_heading(
      paste0("Method: ", .acsCytofValidationMethodLabels()[[method]]), 4
    )
    for (metric in c("pcc", "ccc")) {
      .analysis_heading(
        if (metric == "pcc") "Pearson correlation" else "Concordance correlation",
        5
      )
      .analysis_print_fig(.acsCytofValidationPlotCorrelation(
        correlationTbl = correlation_tbl,
        method = method,
        metric = metric,
        realPopulationsOnly = FALSE
      ))
    }
  }
}
```

## Appendix: Tailgate and F-beta at default settings

The main results use Tailgate and F-beta as tuned for the CyTOF data in
Analysis 9: cells with an expression of exactly zero are removed before the
gate is estimated, and Tailgate uses a fixed tolerance and a bias of 0.2. Here
both methods use their published defaults on every cell instead: F-beta with
`beta = 0.8`, `theta = 2`, `width = 10` and pdfs that integrate to one, and
Tailgate with its automatic tolerance and no bias. The eligibility rules and
heatmap layout match the main results.

```{r}
#| label: appendix-correlation-table
#| echo: false
#| results: asis
appendix_correlation_tbl <- .acsCytofValidationCorrelationTable(
  manual_comparison_all_tbl |>
    dplyr::filter(.data$method %in% appendix_methods)
)
if (isTRUE(run_plots)) {
  .analysis_report_table(appendix_correlation_tbl,
    c(fig_key, "validation-correlations-default-settings.csv"),
    "Eligible-stratum PCC and CCC for Tailgate and F-beta at default settings.",
    path_root = root_dir)
}
```

### Correlation heatmaps for CD4, CD8 and TCRgd T cells

```{r}
#| label: appendix-t-cell-correlations
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (method in appendix_methods) {
    .analysis_heading(
      paste0("Method: ", .acsCytofValidationMethodLabels()[[method]]), 4
    )
    for (metric in c("pcc", "ccc")) {
      .analysis_heading(
        if (metric == "pcc") "Pearson correlation" else "Concordance correlation",
        5
      )
      .analysis_print_fig(.acsCytofValidationPlotCorrelation(
        correlationTbl = appendix_correlation_tbl,
        method = method,
        metric = metric,
        realPopulationsOnly = TRUE
      ))
    }
  }
}
```

### Correlation heatmaps for all matched populations

```{r}
#| label: appendix-all-correlations
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (method in appendix_methods) {
    .analysis_heading(
      paste0("Method: ", .acsCytofValidationMethodLabels()[[method]]), 4
    )
    for (metric in c("pcc", "ccc")) {
      .analysis_heading(
        if (metric == "pcc") "Pearson correlation" else "Concordance correlation",
        5
      )
      .analysis_print_fig(.acsCytofValidationPlotCorrelation(
        correlationTbl = appendix_correlation_tbl,
        method = method,
        metric = metric,
        realPopulationsOnly = FALSE
      ))
    }
  }
}
```
