---
title: ACS CyTOF validation outputs
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
---

This analysis compares automated and manual background-subtracted cytokine frequencies in the ACS CyTOF dataset. It uses the matched comparison table saved by analysis 9 to plot frequencies and calculate Pearson and concordance correlations for StimGate, F-beta and Tailgate. The correlation summaries retain the manuscript eligibility rules.

## Set-up

The saved `acs_cytof/manual-comparison/manual-comparison.rds` table is sufficient; raw FCS files are not needed here. Render analysis 9 with `RUN_SIMULATIONS=true RUN_PLOTS=false` first if that table is missing. `RUN_PLOTS=true` displays and saves the validation figures; this document does not rerun the gating methods.

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)
# Quarto may start in analysis/ when no project execution directory is set.
if (!file.exists(file.path(root_dir, "DESCRIPTION"))) {
  root_dir <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
}
if (!file.exists(file.path(root_dir, "DESCRIPTION"))) {
  stop("Cannot find the stimgate repository root; render from the checkout.")
}

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

path_manual_output <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "manual-comparison"
)
path_comparison <- file.path(
  path_manual_output,
  "manual-comparison.rds"
)
if (!file.exists(path_comparison)) {
  stop(
    "ACS manual comparison table not found at: ",
    path_comparison,
    ". Run RUN_SIMULATIONS=true RUN_PLOTS=false quarto render ",
    "analysis/9-real-compare-acs-cytof.qmd first with the ACS data available."
  )
}
manual_comparison_tbl <- readRDS(path_comparison)
validation_methods <- .acsCytofValidationMethods()
.acsCytofValidationValidateComparisonTable(
  manual_comparison_tbl,
  requiredMethods = validation_methods
)
```

## Scientific inclusion rules

The raw-frequency plots show automated against manual background-subtracted
frequencies for every matched population and cytokine. The correlation outputs
follow the manuscript eligibility rule before calculating Pearson correlation
(PCC) and concordance correlation (CCC): within each population/cytokine/stimulus
stratum, the 75th percentile of the manual stimulated frequency must exceed
three times the larger of 0.01% and the median manual unstimulated frequency,
and the 75th percentile of the manual background-subtracted frequency must
exceed 0.02%. The heatmaps omit the non-secreted-protein (`p4`) stimulation,
matching the manuscript-style display. One set of heatmaps is restricted to
CD4, CD8 and TCRgd T cells and another uses every matched population. This
keeps the original manuscript comparison available while also showing the
additional populations in the current benchmark.

## Correlation table

```{r}
#| label: correlation-table
#| echo: false
correlation_tbl <- .acsCytofValidationCorrelationTable(manual_comparison_tbl)
correlation_tbl
```

## Validation figures

```{r}
#| label: save-validation-figures
#| include: false
if (isTRUE(run_plots)) {
  path_dir_save <- file.path(
    path_manual_output, "validation-figures"
  )
  .acsCytofValidationSavePlots(
    comparisonTbl = manual_comparison_tbl,
    pathDirSave = path_dir_save
  )
}
```

### Raw background-subtracted frequencies

Manual background-subtracted frequency is on the horizontal axis and automated frequency on the vertical axis. Colours identify stimuli, and panels separate populations and cytokines within each method. Points closer to the identity line indicate closer agreement; horizontal and vertical lines mark zero frequency.

```{r}
#| label: fig-acs-validation-scatter
#| echo: false
#| fig-width: 12
#| fig-height: 9
#| fig-cap: "Automated versus manual background-subtracted frequencies, one figure per method."
if (isTRUE(run_plots)) {
  for (method in validation_methods) {
    print(.acsCytofValidationPlotScatter(manual_comparison_tbl, method))
  }
}
```

### Correlation heatmaps for CD4, CD8 and TCRgd T cells

Cytokines run across columns and the three T-cell populations across rows, with panels for stimuli. Colour and text show Pearson or concordance correlation for each method after applying the eligibility rules. Values closer to one indicate stronger agreement; grey cells mark unavailable correlations.

```{r}
#| label: fig-acs-validation-t-cell-correlations
#| echo: false
#| fig-width: 12
#| fig-height: 5
#| fig-cap: "Pearson and concordance correlations for CD4, CD8 and TCRgd T cells."
if (isTRUE(run_plots)) {
  for (method in validation_methods) {
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "pcc",
      realPopulationsOnly = TRUE
    ))
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "ccc",
      realPopulationsOnly = TRUE
    ))
  }
}
```

### Correlation heatmaps for all matched populations

These heatmaps use the same cytokine, stimulus and correlation scales for all matched populations, including B and NK cells. Compare methods within the same population and cytokine, keeping the eligibility rules in mind.

```{r}
#| label: fig-acs-validation-all-correlations
#| echo: false
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "Pearson and concordance correlations across all matched populations."
if (isTRUE(run_plots)) {
  for (method in validation_methods) {
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "pcc",
      realPopulationsOnly = FALSE
    ))
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "ccc",
      realPopulationsOnly = FALSE
    ))
  }
}
```
