---
title: Share thresholds within clusters of similar samples
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
---

This analysis illustrates threshold sharing within clusters of similar paired
samples. Section A asks whether large expression location shifts between two
labs lead to separate sample clusters (#558). Section B examines weak-response
sensitivity when thresholds are adjusted using information from other samples
in the same cluster (#559). Both demonstrations assess the observed result;
neither lab membership nor improved sensitivity is assumed in advance.

## Set-up

```{r}
#| label: setup
#| include: false
root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)
scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
run_simulations <- .as_flag(.get_qmd_param_env("run_simulations", "RUN_SIMULATIONS", FALSE))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
analysis_qmd <- "analysis/12-sim-cluster-gates.qmd"
# A fixed demonstration, without draft/final simulation-size settings.
analysis_key <- .analysis_mode_key(c("sim", "cluster-gates"), sized = FALSE)
fig_key <- .analysis_mode_key("12-sim-cluster-gates", sized = FALSE)
devtools::load_all(root_dir)
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "sim-cluster-lab.R"))
source(file.path(scripts_r_dir, "sim-cluster-weak.R"))
if (isTRUE(run_simulations)) .analysis_require_packages("simcyto")
```

## A. Lab location shifts and sample clusters

Each lab contributes six samples, each with its own unstimulated and stimulated
tube. The simulated marker has Gaussian negative and positive populations with
means 10 and 20 and variance 1. The positive fraction is 2% in the unstimulated
tube and 22% in the stimulated tube. Small sample perturbations (SD 0.05) are
allowed, with no additional condition or population perturbations. Lab B receives
an additive shift of four expression units in **both tubes and both populations**.
This is between-lab location variation, with matched controls in each lab.

All samples are gated together. The `batchList` entries identify the individual
sample pairs so each has its own control; lab labels are used only for reporting.
The density bandwidth, clustering bandwidth and unstimulated bias are each 0.3.
Threshold sharing is enabled and cytokine-positive refinement is disabled.
The seed is fixed at 558. Quick and development renders use three samples per
lab and 500 cells per tube; they are smoke checks rather than the full example.

The current method compares paired densities on a common absolute expression
grid over their left-hand region. A gap statistic selects the number of clusters,
then k-means assigns samples. It requires at least one directly generated local-FDR
threshold. Within a cluster, direct thresholds are clipped to its 15th and 85th
percentiles when at least three are available; non-direct thresholds can receive
the 60th percentile of direct thresholds. Clusters without direct thresholds keep
their original thresholds. These threshold-sharing clusters differ from the
clusters used for shared bandwidth estimation.

```{r}
#| label: lab-settings
#| echo: false
lab_settings <- list(
  seed = 558L,
  n_sample_lab = if (analysis_quick || analysis_dev) 3L else 6L,
  n_cell = if (analysis_quick || analysis_dev) 500L else 2000L,
  lab_shift = 4
)
lab_cache_file <- file.path(
  .analysis_cache_dir(c(analysis_key, "lab"), root_dir, create = FALSE),
  "result.rds"
)
```

```{r}
#| label: lab-simulate
#| include: false
if (isTRUE(run_simulations)) {
  lab_new_result <- do.call(.simClusterLabRun, lab_settings)
  .write_rds_atomic(list(
    semantics = "cluster-lab-v1", settings = lab_settings, result = lab_new_result
  ), lab_cache_file)
}
```

```{r}
#| label: lab-read
#| include: false
lab_cache <- .analysis_read_rds(lab_cache_file, c(analysis_key, "lab"), analysis_qmd)
if (!identical(lab_cache$semantics, "cluster-lab-v1") ||
    !identical(lab_cache$settings, lab_settings)) {
  .analysis_cache_error(c(analysis_key, "lab"),
    "Cached lab demonstration has different scientific settings.", analysis_qmd)
}
lab_result <- lab_cache$result
```

### Paired expression distributions

Each line represents one tube; colours identify the lab. The panels use the same
expression scale, so any lab shift should appear in both conditions.

```{r}
#| label: lab-distributions
#| echo: false
if (isTRUE(run_plots)) {
  .analysis_print_save_fig(
    .simClusterLabDistributionPlot(lab_result$expression),
    file.path(.analysis_fig_dir(c(fig_key, "lab", "distributions"), root_dir),
      "paired-expression.png"), height = 14
  )
}
```

### Cluster allocations and thresholds

Cluster numbers are arbitrary labels. Separation is supported if every sample
has an allocation and no assigned cluster contains samples from both labs.
Each lab may occupy more than one cluster. Missing allocations prevent a complete
separation claim and remain visible in the figure and table. The threshold columns
show the original threshold and its value after within-cluster adjustment.

```{r}
#| label: lab-allocations
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .analysis_print_save_fig(
    .simClusterLabAllocationPlot(lab_result$allocations),
    file.path(.analysis_fig_dir(c(fig_key, "lab", "allocations"), root_dir),
      "sample-clusters.png"), height = 14
  )
  .analysis_report_table(lab_result$allocations,
    c(fig_key, "lab", "sample-lab-cluster.csv"),
    "Sample, lab and assigned cluster.", root_dir)
  .analysis_report_table(lab_result$summary,
    c(fig_key, "lab", "cluster-separation-summary.csv"),
    "Cluster separation summary.", root_dir)
  if (isTRUE(lab_result$summary$labs_separated)) {
    cat("\n\nAll samples were allocated, and no cluster contains samples from both labs. ",
      "The observed allocation separates the labs in this seeded example.\n\n", sep = "")
  } else if (is.na(lab_result$summary$labs_separated)) {
    cat("\n\nSome samples have no recorded cluster allocation. ",
      "Complete separation cannot be assessed; inspect the saved gate details.\n\n", sep = "")
  } else {
    cat("\n\nAt least one cluster contains samples from both labs. ",
      "This seeded example does not demonstrate complete lab separation.\n\n", sep = "")
  }
}
```

This single dataset illustrates the current procedure, rather than estimating how
often it separates labs. The common density grid ends at a fraction of a low
quantile of direct thresholds, so a large shift can leave only part of Lab B's
background in the clustering region. Review the saved details and allocations
before interpreting separation as general robustness to lab variation.

## B. Can stronger samples help detect a weak response?

This example asks whether sharing thresholds between comparable samples can
recover true responding cells in samples with a rare response. Strong and weak
samples have the same negative and responding expression distributions. Only
the responding fraction differs. The cells and their truth labels stay fixed
throughout the comparison; only the gates can change.

### Response settings

The strong responding fraction is 1%. The clearly editable weak fraction below
is 0.2%, five times smaller. At 50,000 cells per tube this gives approximately
100 true responders per weak sample, enough to inspect sensitivity without
relying on just a handful of cells. Four samples of each type provide several
potential donors while keeping this a single small demonstration. Responders
have mean expression 4.5, negatives have mean 0, and both have variance 1.5.
This moderate overlap leaves a real contamination tradeoff when lowering a gate.
Unstimulated tubes contain only negatives. There are no location shifts,
response-shape changes or negative-variance changes between samples or tubes.

The fixed bandwidth is 0.25, with an unstimulated bias of 0.25, and
cytokine-positive refinement is disabled for this one-marker example. These
settings isolate threshold sharing from additional refinements. They are not
selected by inspecting the results. Quick and development renders reduce the
cell count to 2,000 but keep both sample groups; they check execution and are
not evidence about sensitivity at a rare response frequency.

```{r}
#| label: weak-settings
weak_settings <- list(
  seed = 559L,
  n_strong = 4L,
  n_weak = 4L,
  strong_prob = 0.01,
  weak_prob = 0.002, # Try 0.001 only as a separately reported scenario.
  n_cell = if (analysis_quick || analysis_dev) 2000L else 50000L,
  mean_pos = 4.5,
  variance = 1.5,
  bw = 0.25,
  bias_uns = 0.25
)
weak_cache_key <- .analysis_mode_key(c("sim", "cluster-gates", "weak"), sized = FALSE)
weak_cache_file <- file.path(
  .analysis_cache_dir(weak_cache_key, root_dir, create = run_simulations), "result.rds"
)
```

### Compare gates on the same cells

**Before** is the pre-cluster sample threshold (`cpOrigQuantMin`) reported by
`getStimGatesDetailed()` during the same run. **After** is its cluster-adjusted
threshold (`cpJoinTgOrig`), checked against the final applied local-FDR
gate with `loc_minClust`. This isolates the adjustment without rerunning simulation or bandwidth
selection. With one stimulated tube per sample and cytokine-positive refinement
disabled, the pre-cluster value is also the applied local-FDR gate with
`loc_min`. Counts for
both versions are checked against the package's own statistics.

Clustering uses paired stimulated and unstimulated densities on a common
absolute expression grid in the background region. Strong samples help only
if they join a weak sample's cluster and supply directly generated thresholds.
With at least three direct donors, their thresholds are clipped to the cluster's
15th and 85th percentiles. A non-direct threshold is replaced by the 60th
percentile of available direct thresholds. A cluster without direct donors
retains its original thresholds. A weak sample with a direct threshold may
keep it or even have it raised; no rule here forces a gate down.

```{r}
#| label: weak-simulate
#| include: false
if (isTRUE(run_simulations)) {
  weak_result <- .simClusterWeakRun(weak_settings)
  .simClusterWeakValidate(weak_result, weak_settings)
  .write_rds_atomic(weak_result, weak_cache_file)
}
```

```{r}
#| label: weak-read
#| include: false
weak_result <- .analysis_read_rds(
  weak_cache_file, weak_cache_key, "analysis/12-sim-cluster-gates.qmd"
)
.simClusterWeakValidate(weak_result, weak_settings)
```

```{r}
#| label: weak-rerun-one-simulation
#| eval: false
# Same scenario and seed as the cached run, without writing the cache.
weak_rerun <- .simClusterWeakRun(weak_settings)
```

### Cluster allocations and threshold changes

The tables record requested response probabilities, sample seeds, cluster
allocations, direct donor counts, quantiles, adjustment actions and threshold
provenance. Negative threshold changes indicate gates moving down. Positive
values for recovered responders are net gains in true positives; negative
values mean that responders were lost. Additional false positives show how
many negative cells changed to positive calls.

```{r}
#| label: weak-tables
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .analysis_report_table(weak_result$settings_table, c(fig_key, "weak", "response-settings.csv"), "Table `response-settings`.", root_dir)
  weak_allocations <- weak_result$allocations |>
    dplyr::select("ind", "grp", "cpOrigQuantMin", "cpJoinTgOrig",
      dplyr::starts_with("locCluster"), "locGenerated", "locGeneratedDirect", "locSource", "locReason") |>
    dplyr::left_join(weak_result$settings_table, by = "ind")
  .analysis_report_table(weak_allocations, c(fig_key, "weak", "cluster-allocations.csv"), "Table `cluster-allocations`.", root_dir)
  .analysis_report_table(weak_result$changes, c(fig_key, "weak", "paired-changes.csv"), "Table `paired-changes`.", root_dir)
  .analysis_report_table(weak_result$scores, c(fig_key, "weak", "truth-and-contamination.csv"), "Table `truth-and-contamination`.", root_dir)
}
```

### Distributions and gates

Each panel shows one sample. Histogram heights are fractions of the whole
tube, so the responder curve retains its small population weight. Dashed lines
show the original gates and solid lines show the adjusted gates; coincident
lines mean no movement. The second figure enlarges the high-expression region
without changing the data or renormalising the responder distribution.

```{r}
#| label: weak-distributions
#| echo: false
#| fig-width: 8
#| fig-height: 10
if (isTRUE(run_plots)) {
  weak_fig_dir <- .analysis_fig_dir(c(fig_key, "weak", "distributions"), root_dir)
  .analysis_print_save_fig(.simClusterWeakDistributions(weak_result),
    file.path(weak_fig_dir, "whole-distribution.png"), width = 22, height = 28, allow_tall = TRUE)
  .analysis_print_save_fig(.simClusterWeakDistributions(weak_result, tail = TRUE),
    file.path(weak_fig_dir, "response-tail.png"), width = 22, height = 28, allow_tall = TRUE)
}
```

### Sensitivity and contamination

Sensitivity is the fraction of true responding cells above the gate. The false
discovery proportion (FDP) is the fraction of positive calls that are actually
negative; the false positive rate is the fraction of all negative cells called
positive. All counts use strict `expression > gate` on the single-precision
GatingSet expression that the package gated. FDP is undefined when there are
no positive calls, and sensitivity is undefined when there are no responders;
these outcomes remain missing rather than being replaced by zero.

```{r}
#| label: weak-outcomes
#| echo: false
if (isTRUE(run_plots)) {
  weak_outcome_dir <- .analysis_fig_dir(c(fig_key, "weak", "outcomes"), root_dir)
  .analysis_print_save_fig(.simClusterWeakOutcomes(weak_result),
    file.path(weak_outcome_dir, "paired-gates-and-purity.png"), height = 16)
}
```

### What the adjustment did

The summary below is computed from the cached result. Strong and weak samples
can share one cluster, because they have the same negative and unstimulated
distributions. Within that cluster, only direct thresholds outside the 15th–85th
percentile range move. Weak samples whose gates already lie inside that range
keep them, so the adjustment can lower only the most extreme weak gates.

```{r}
#| label: weak-interpretation
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  weak_changes <- weak_result$changes
  weak_rows <- weak_changes[weak_changes$response == "Weak", ]
  moved <- weak_rows[weak_rows$threshold_change != 0, ]
  cat("\n\n", nrow(moved), " of ", nrow(weak_rows), " weak samples had their gate changed",
    if (nrow(moved)) paste0(" (", paste0(moved$sample, ": ",
      sprintf("%+.2f", moved$threshold_change), collapse = "; "), ")") else "",
    ". Across weak samples, the adjustment recovered ", sum(weak_rows$responders_recovered),
    " true responders net, added ", sum(weak_rows$additional_false_positives),
    " false positives and changed mean sensitivity from ",
    sprintf("%.3f", mean(weak_rows$sensitivity_before)), " to ",
    sprintf("%.3f", mean(weak_rows$sensitivity_after)), ". ",
    "Strong samples recovered ", sum(weak_changes$responders_recovered[weak_changes$response == "Strong"]),
    " responders net and added ",
    sum(weak_changes$additional_false_positives[weak_changes$response == "Strong"]),
    " false positives.\n\n", sep = "")
}
```

More positive calls alone would not show a benefit, so read any gain beside the
extra false positives and FDP. An undefined FDP means no cells were called
positive; it is not evidence of perfect purity. This is one jointly gated
dataset: it shows what the current method does here, with no estimate of
between-dataset uncertainty or general performance.
