---
title: Where StimGate gates each ACS CyTOF sample (13)
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
  pops: all
  stims: all
  cyts: all
  samples: none
  html_pops: all
  html_stims: all
  html_cyts: all
  html_samples: none
  html_by_group: false
  html_n_largest: 20
  html_n_random: 5
  html_n_closest: 5
  html_seed: 20261008
  n_workers: 2
---

This document shows, for real data, what Analysis 2c shows for simulated
samples: where StimGate's gate falls on each stimulated tube of the ACS CyTOF
dataset, and why. There is one page for every sample (stimulated tube),
cytokine, population and stimulation.

Each population is gated again exactly as in Analysis 9 (same settings,
same data and the same random-number stream), while the local-FDR steps of
every tube are recorded. Each page shows the raw stimulated and unstimulated
densities, the raw and smoothed response probability, its derivative, the
expected responding cells per bin and the taut-string density, with the
filtering boundaries and the initial local-FDR gate. It adds the final gate
StimGate applied after threshold sharing and cytokine-positive refinement
(`loc_minClust`, the gate Analysis 9 scores), Analysis 9's Tailgate and
F-beta gates, and a purple dashed line at the stimulated-tube value where
stimulated minus unstimulated frequency equals the manual frequency. That line
is derived from the manual frequency only; it is not a manual gate. The text
below the plots gives the settings, the bandwidth and `biasUns` actually used,
the gates with their source and reason, and StimGate's and the manual
background-subtracted frequencies.

## Set-up

`RUN_SIMULATIONS=true` re-gates the chosen populations and draws every page
into this analysis's cache; `RUN_PLOTS=true` copies the pages to the output
folder and writes this report. Both read Analysis 9's cached GatingSets, its
StimGate gates, its Tailgate and F-beta results and its manual comparison,
and never change them.

Parameters (or environment variables) choose what is drawn. Lists are comma
separated; `all` selects everything.

- `html_pops`, `html_stims`, `html_cyts` (`ACS_DEBUG_HTML_POPS`,
  `ACS_DEBUG_HTML_STIMS`, `ACS_DEBUG_HTML_CYTS`, default all): the
  combinations considered for this report. Across all of them together
  (`html_by_group = false`, the default; `true` chooses within each
  population, stimulation and cytokine), it shows the `html_n_largest` (20)
  samples where StimGate's Analysis 9 frequency differs most from the manual
  one, the `html_n_closest` (5) that agree best among samples with a manual
  response, and `html_n_random` (5) of the others (seed `html_seed`); without
  Analysis 9's manual comparison, all are chosen at random. `html_samples`
  (`ACS_DEBUG_HTML_SAMPLES`) shows chosen SampleIDs instead. The report
  samples are chosen when re-gating, so change these with
  `RUN_SIMULATIONS=true`.
- `pops` (`ACS_DEBUG_POPS`), `stims` (`ACS_DEBUG_STIMS`), `cyts`
  (`ACS_DEBUG_CYTS`), `samples` (`ACS_DEBUG_SAMPLES`): further pages saved to
  files. `samples` defaults to `none`, so only the report's samples are drawn;
  `manual` adds every donor with manual gating and `all` every donor. Every
  tube of a chosen population is still gated, so the gates match Analysis 9.
- `n_workers` (`ACS_N_WORKERS`): populations gated at once.

Quick and dev renders gate only the Analysis 9 TCRgd tester GatingSet and draw
its IL-2 pages.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
analysis_qmd <- "analysis/13-real-debug-acs-cytof.qmd"
fig_key <- .analysis_mode_key("13-real-debug-acs-cytof", sized = FALSE)
analysis_quick <- .analysis_is_quick() || .analysis_is_dev()
for (fn in c(
  "acs_cytof-helper.R",
  "acs_cytof-gate.R",
  "acs_cytof-methods.R",
  "sim-debug-loc.R",
  "acs_cytof-debug.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations", "RUN_SIMULATIONS", FALSE
))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
param_list <- function(param, env, default = "all") {
  .acsCytofDebugList(.get_qmd_param_env(param, env, default))
}
param_int <- function(param, env, default) {
  int <- as.integer(.get_qmd_param_env(param, env, default))
  if (length(int) != 1L || is.na(int) || int < 0L) {
    stop(param, " must be a non-negative integer.")
  }
  int
}
sel_pops <- param_list("pops", "ACS_DEBUG_POPS")
sel_stims <- param_list("stims", "ACS_DEBUG_STIMS")
sel_cyts <- param_list("cyts", "ACS_DEBUG_CYTS")
sel_samples <- param_list("samples", "ACS_DEBUG_SAMPLES", "none")
html_pops <- param_list("html_pops", "ACS_DEBUG_HTML_POPS")
html_stims <- param_list("html_stims", "ACS_DEBUG_HTML_STIMS")
html_cyts <- param_list("html_cyts", "ACS_DEBUG_HTML_CYTS")
html_samples <- param_list("html_samples", "ACS_DEBUG_HTML_SAMPLES", "none")
if (!length(html_samples)) html_samples <- NULL
html_by_group <- .as_flag(.get_qmd_param_env(
  "html_by_group", "ACS_DEBUG_HTML_BY_GROUP", FALSE
))
html_n_largest <- param_int("html_n_largest", "ACS_DEBUG_HTML_N_LARGEST", 20L)
html_n_random <- param_int("html_n_random", "ACS_DEBUG_HTML_N_RANDOM", 5L)
html_n_closest <- param_int("html_n_closest", "ACS_DEBUG_HTML_N_CLOSEST", 5L)
html_seed <- param_int("html_seed", "ACS_DEBUG_HTML_SEED", 20261008L)
n_workers <- param_int("n_workers", "ACS_N_WORKERS", 2L)
if (n_workers < 1L) stop("n_workers must be at least one.")

.acsCytofEnsureCurrentCheckout(root_dir)
```

## Settings

The populations, StimGate settings and seed are read from Analysis 9's
`scientific-settings` chunk, so the two analyses cannot drift apart.

```{r}
#| label: scientific-settings
settings <- .acsCytofDebugSettings(
  file.path(root_dir, "analysis", "9-real-compare-acs-cytof.qmd")
)
output_group <- NULL
if (analysis_quick) {
  # Analysis 9's tester: its first complete TCRgd batches, gated on their own.
  output_group <- "tester"
  sel_pops <- "tcrgd"
  if (is.null(sel_cyts)) sel_cyts <- "IL2"
  html_pops <- "tcrgd"
  html_cyts <- intersect(html_cyts %||% sel_cyts, sel_cyts)
}
run_pops <- intersect(settings$popVec, sel_pops %||% settings$popVec)
unknown_pops <- setdiff(c(sel_pops, html_pops), settings$popVec)
if (length(unknown_pops)) {
  stop("Unknown populations: ", paste(unknown_pops, collapse = ", "))
}
unknown_cyts <- setdiff(c(sel_cyts, html_cyts), .acsCytofChannelMap())
if (length(unknown_cyts)) {
  stop("Unknown cytokines: ", paste(unknown_cyts, collapse = ", "))
}
str(settings)
```

```{r}
#| label: paths
#| include: false
# Analysis 9's caches are only read: never create or change them here.
path_gs_base <- projr::projr_path_get(
  "cache", "acs_cytof", "gs",
  create = FALSE, format = "absolute"
)
path_scratch_base <- projr::projr_path_get(
  "cache", "acs_cytof", "scratch",
  create = FALSE, format = "absolute"
)
path_comparison <- projr::projr_path_get(
  "cache", "acs_cytof", "manual-comparison", "manual-comparison.rds",
  create = FALSE, format = "absolute"
)
# This analysis's own cache.
path_debug_base <- .analysis_cache_dir(
  c("acs_cytof_debug", fig_key[-1L]), root_dir
)
paths9 <- lapply(stats::setNames(nm = settings$popVec), function(pop) {
  .acsCytofPopulationPaths(
    pop, "unused", path_gs_base, path_scratch_base,
    outputGroup = output_group
  )
})
paths_debug <- lapply(stats::setNames(nm = settings$popVec), function(pop) {
  .acsCytofDebugPaths(pop, path_debug_base, outputGroup = output_group)
})
```

## Re-gating

Every tube of each chosen population is gated, as in Analysis 9, with one
worker per population (`n_workers` at once) and the population's own
random-number stream. Each recorded tube is saved as soon as it is gated;
once the final gates are known, its page is drawn and the record deleted. The
pages are built in a temporary folder that replaces the previous one only
once the population has finished.

```{r}
#| label: run-regating
#| include: false
if (isTRUE(run_simulations)) {
  comparison <- tryCatch(
    .acsCytofDebugReadComparison(path_comparison, settings$locThresholdMethod),
    error = function(e) {
      message("Analysis 9's manual comparison is unusable: ", conditionMessage(e))
      NULL
    }
  )
  if (analysis_quick && !is.null(comparison)) {
    # The tester's gates differ from the full run that Analysis 9 scored.
    comparison[c("freqBsAnalysis9", "diffAnalysis9", "absDiffAnalysis9")] <- NA_real_
  }
  candidates <- dplyr::bind_rows(lapply(intersect(html_pops, run_pops), function(pop) {
    .acsCytofReadPreprocessing(paths9[[pop]]$gs)$sampleMap |>
      dplyr::filter(.data$stim != "uns") |>
      tidyr::expand_grid(cyt = unname(.acsCytofChannelMap())) |>
      dplyr::transmute(pop = .env$pop, .data$stim, .data$cyt, .data$SampleID)
  }))
  if (nrow(candidates)) {
    candidates <- candidates |>
      dplyr::filter(
        is.null(html_stims) | .data$stim %in% html_stims,
        is.null(html_cyts) | .data$cyt %in% html_cyts
      )
  }
  html_keys <- .acsCytofDebugSelectHtml(
    candidates, comparison,
    nLargest = html_n_largest, nRandom = html_n_random,
    nClosest = html_n_closest, seed = html_seed, samples = html_samples,
    byGroup = html_by_group
  )
  sel_samples <- .acsCytofDebugResolveSamples(sel_samples, comparison)
  # Pages for the report's samples are drawn even if not selected here.
  select <- list(stim = sel_stims, cyt = sel_cyts, sample = sel_samples)
  manifest <- list(
    analysis_semantics_version = "acs-cytof-debug-v1",
    run_id = .sanitize_run_id(Sys.getenv("ANALYSIS_RUN_ID", unset = "")),
    settings = settings,
    select = select,
    output_group = output_group
  )
  run_population <- function(pop) {
    if (!pop %in% run_pops) {
      return(list(pop = pop, success = TRUE, error = NULL))
    }
    tryCatch(
      .acsCytofDebugRunPopulation(
        pop = pop,
        paths9 = paths9[[pop]],
        pathsDebug = paths_debug[[pop]],
        settings = settings,
        select = select,
        comparison = comparison,
        htmlKeys = html_keys[html_keys$pop == pop, ],
        manifest = manifest,
        rootDir = root_dir
      ),
      error = function(e) {
        list(pop = pop, success = FALSE, error = conditionMessage(e))
      }
    )
  }
  if (!is.null(output_group)) {
    # Analysis 9 gates its tester in the main session after this seed.
    set.seed(settings$seed)
    tester_run <- run_population("tcrgd")
    if (!isTRUE(tester_run$success)) {
      stop("TCRgd tester re-gating failed: ", tester_run$error)
    }
  } else {
    # Same populations, order and seed as Analysis 9, so each population
    # gets the random-number stream it had there. Unchosen ones return at once.
    .acsCytofMapPopulations(
      settings$popVec,
      run_population,
      nWorkers = n_workers,
      seed = settings$seed,
      label = "ACS per-sample re-gating"
    )
  }
}
```

## Results

```{r}
#| label: read-summaries
#| include: false
summaries <- list()
if (isTRUE(run_plots)) {
  for (pop in run_pops) {
    path <- paths_debug[[pop]]$summary
    if (!file.exists(path)) {
      message("No re-gated pages for ", pop, "; run with RUN_SIMULATIONS=true.")
      next
    }
    obj <- .analysis_read_rds(path, "acs_cytof_debug", analysis_qmd)
    .analysis_check_expected_run(
      obj$manifest, "acs_cytof_debug", analysis_qmd,
      semantics_version = "acs-cytof-debug-v1"
    )
    summaries[[pop]] <- obj$table
  }
}
summary_tbl <- dplyr::bind_rows(summaries)
```

### Gates and pages per population

`same gates as Analysis 9` counts tubes and cytokines whose final gates
(ordinary and cytokine-positive) equal the gates Analysis 9 saved. Any
difference means the re-gating did not reproduce Analysis 9 and the pages
must not be read as Analysis 9's gating. `local FDR ran` counts tubes that
reached local-FDR gating; the others had too few cells.

```{r}
#| label: population-summary
#| echo: false
#| results: asis
if (isTRUE(run_plots) && nrow(summary_tbl)) {
  pop_tbl <- summary_tbl |>
    dplyr::group_by(.data$pop) |>
    dplyr::summarise(
      n = dplyr::n(),
      local_fdr = sum(.data$recorded),
      fallback = sum(!(.data$locGenerated %in% TRUE)),
      same = sum(.data$matchesAnalysis9 %in% TRUE),
      different = sum(.data$matchesAnalysis9 %in% FALSE),
      pages = sum(!is.na(.data$page)),
      failed = sum(!is.na(.data$pageError)),
      .groups = "drop"
    )
  for (i in seq_len(nrow(pop_tbl))) {
    with(pop_tbl[i, ], cat(
      "- **", pop, "**: ", n, " tubes x cytokines; local FDR ran for ",
      local_fdr, "; final gate fallback for ", fallback, "; same gates as ",
      "Analysis 9 for ", same, ", different for ", different, "; ", pages,
      " pages saved (", failed, " failed).\n", sep = ""
    ))
  }
  if (any(summary_tbl$matchesAnalysis9 %in% FALSE)) {
    cat("\n\n**Warning: some re-gated gates differ from Analysis 9's saved gates.**\n\n")
  }
}
```

### Combinations

```{r}
#| label: combination-summary
#| echo: false
#| results: asis
if (isTRUE(run_plots) && nrow(summary_tbl)) {
  summary_tbl |>
    dplyr::group_by(.data$pop, .data$stim, .data$cyt) |>
    dplyr::summarise(
      n = dplyr::n(),
      n_local_fdr = sum(.data$recorded),
      n_fallback = sum(!(.data$locGenerated %in% TRUE)),
      n_same_as_analysis_9 = sum(.data$matchesAnalysis9 %in% TRUE),
      n_manual = sum(is.finite(.data$freqBsManual)),
      median_abs_diff_analysis_9 = stats::median(.data$absDiffAnalysis9, na.rm = TRUE),
      max_abs_diff_analysis_9 = suppressWarnings(max(.data$absDiffAnalysis9, na.rm = TRUE)),
      pdf = paste(unique(stats::na.omit(.data$pdf)), collapse = ";"),
      .groups = "drop"
    ) |>
    dplyr::mutate(max_abs_diff_analysis_9 = dplyr::if_else(
      is.finite(.data$max_abs_diff_analysis_9), .data$max_abs_diff_analysis_9, NA_real_
    )) |>
    .analysis_report_table(
      c(fig_key, "combination-summary.csv"),
      paste(
        "Tubes, local-FDR and fallback counts, agreement with Analysis 9's gates",
        "and absolute StimGate - manual difference (percentage points) per",
        "population, stimulation and cytokine."
      ),
      path_root = root_dir
    )
  summary_tbl |>
    dplyr::arrange(.data$pop, .data$stim, .data$cyt, .data$SampleID) |>
    .analysis_report_table(
      c(fig_key, "per-sample-summary.csv"),
      paste(
        "One row per tube and cytokine: gates, provenance, frequencies and the",
        "page (`pdf`, `page`) showing it."
      ),
      path_root = root_dir
    )
}
```

### Saved pages

Pages are saved under `output/fig/13-real-debug-acs-cytof/per_sample/`, one
PDF per population, stimulation and cytokine
(`<population>/<stimulation>/<cytokine>.pdf`), one page per sample in
SampleID order. `page` and `pdf` in the per-sample table give each sample's
page.

```{r}
#| label: publish-pages
#| include: false
if (isTRUE(run_plots)) {
  path_pages_out <- .analysis_fig_dir(c(fig_key, "per_sample"), root_dir)
  for (pop in names(summaries)) {
    .acsCytofReplaceDir(file.path(path_pages_out, pop), function(path_tmp) {
      from <- list.files(paths_debug[[pop]]$pages, full.names = TRUE)
      if (length(from) && !all(file.copy(from, path_tmp, recursive = TRUE))) {
        stop("Could not copy the pages of ", pop, ".")
      }
    })
  }
}
```

### Chosen samples

Each figure is one saved page. Its heading gives why it was chosen and
StimGate's frequency scored in Analysis 9 (after cytokine-positive
refinement) against the manual frequency, both background subtracted and
clipped at zero, in percent of the population.

```{r}
#| label: html-figures
#| echo: false
#| results: asis
if (isTRUE(run_plots) && nrow(summary_tbl)) {
  shown <- summary_tbl |>
    dplyr::filter(!is.na(.data$htmlReason), !is.na(.data$page)) |>
    dplyr::mutate(htmlReason = factor(.data$htmlReason, levels = c(
      "largest error", "random", "close agreement",
      "random (no manual comparison)", "requested"
    ))) |>
    dplyr::arrange(.data$pop, .data$stim, .data$cyt, .data$htmlReason)
  pct <- function(x) if (is.finite(x)) sprintf("%.2f%%", x) else "not available"
  for (pop in unique(shown$pop)) {
    .analysis_heading(paste("Population:", pop), 4)
    rows_pop <- shown[shown$pop == pop, ]
    for (combo in unique(paste(rows_pop$stim, rows_pop$cyt))) {
      rows <- rows_pop[paste(rows_pop$stim, rows_pop$cyt) == combo, ]
      .analysis_heading(paste0(
        "Stimulation ", rows$stim[[1L]], ", ", rows$cyt[[1L]]
      ), 5)
      for (i in seq_len(nrow(rows))) {
        row <- rows[i, ]
        cat(
          "\n\n**", row$SampleID, "** (", as.character(row$htmlReason),
          "): StimGate ", pct(row$freqBsAnalysis9), ", manual ",
          pct(row$freqBsManual), ". Page ", row$page, " of `",
          paste("per_sample", pop, row$pdf, sep = "/"), "`.\n\n",
          sep = ""
        )
        x <- readRDS(file.path(
          paths_debug[[pop]]$html, paste0(row$ind, "_", row$chnl, ".rds")
        ))
        .analysis_print_fig(
          .acsCytofDebugFigure(x$rec, x$row),
          width = 36, height = 48
        )
      }
    }
  }
}
```
