---
title: Compare StimGate, Tailgate and F-beta on OMIP-111
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_preprocessing: null
  run_methods: null
  run_plots: true
  run_diagnostics: true
---

## Suitability and scope

OMIP-111 is suitable for an external **agreement comparison** of cytokine
gating methods. The deposit includes ten matched mice (five C57BL/6, five
BALB/c), each with a Brefeldin A control and a mitogen plus Brefeldin A tube,
two author-supplied FlowJo workspaces, and a sample-key spreadsheet. The 18
single-colour/unstained reference files are excluded from the biological
comparison. The full-panel FCS files are already spectrally unmixed.

The [OMIP-111 publication](https://pmc.ncbi.nlm.nih.gov/articles/PMC12758606/)
describes 6-hour PMA/ionomycin stimulation and gating cytokines in conventional
non-naïve CD4/CD8 T cells. We reconstruct those parent gate definitions, retaining
the authors' lymphocyte, singlet, live CD45, lineage and non-naïve gates. All
methods receive the same reconstructed parent cells. CD4 is evaluated for IFNγ, IL-2, TNF,
**combined IL-4/5**, and IL-17A; CD8 is evaluated for IFNγ, IL-2 and TNF,
matching the cytokine reference gates supplied in the workspaces.

Manual gating is an imperfect reference, not biological ground truth. This
small, strongly stimulated mouse dataset cannot establish performance on rare
antigen-specific human responses. IL-17A and IL-4/5 are particularly sparse;
percentage-point differences remain interpretable when relative errors would
be unstable or undefined. The non-naïve denominator changes after stimulation;
frequencies below refer to each tube's own non-naïve parent population.
Strains are analysed separately, avoiding shared gates across different panels
(NK1.1 versus NKp46) and potentially different display transformations.

## Setup and scientific settings

Run from the package root:

```bash
RUN_SIMULATIONS=true RUN_PLOTS=true quarto render analysis/14-real-compare-omip111.qmd
```

`RUN_PREPROCESSING` and `RUN_METHODS` override the umbrella computation flag.
To redraw validated saved results, use `RUN_SIMULATIONS=false RUN_PLOTS=true`.
The workspace importer requires Python with `flowkit==1.3.2`. On this workspace
it is installed in the isolated `cache/omip111-python` environment. Override its
Python path with `OMIP111_PYTHON` when needed. Create an equivalent environment
with `python3 -m venv --system-site-packages <path>` and
`<path>/bin/python -m pip install flowkit==1.3.2`.
F-beta requires Python with NumPy, selected through `RETICULATE_PYTHON` where
needed. Missing method dependencies fail before any method runs.

```{r}
#| label: setup
#| include: false
root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)
scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
isolated_library <- .analysis_project_dir("cache", "omip111-R-library", root_dir, create = FALSE)
if (dir.exists(isolated_library)) .libPaths(c(isolated_library, .libPaths()))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))
source(file.path(scripts_r_dir, "omip016-methods.R"))
source(file.path(scripts_r_dir, "omip111.R"))
source(file.path(scripts_r_dir, "sim-debug-loc.R"))
source(file.path(scripts_r_dir, "acs_cytof-debug.R"))
source(file.path(scripts_r_dir, "omip111-debug.R"))
run_simulations <- .as_flag(.get_qmd_param_env("run_simulations", "RUN_SIMULATIONS", FALSE))
run_preprocessing <- .as_flag(.get_qmd_param_env("run_preprocessing", "RUN_PREPROCESSING", run_simulations))
run_methods <- .as_flag(.get_qmd_param_env("run_methods", "RUN_METHODS", run_simulations))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
run_diagnostics <- .as_flag(.get_qmd_param_env("run_diagnostics", "RUN_DIAGNOSTICS", TRUE))
fig_key <- .analysis_mode_key("14-real-compare-omip111", sized = FALSE)
path_raw <- .analysis_project_dir("raw-data-large", "OMIP-111", root_dir, create = FALSE)
path_prepared <- .analysis_project_dir("cache", c("omip111", "prepared"), root_dir, create = FALSE)
path_results <- .analysis_project_dir("cache", c("omip111", "current"), root_dir, create = FALSE)
path_fbeta <- file.path(root_dir, "scripts", "python", "fbeta.py")
path_python <- Sys.getenv("OMIP111_PYTHON", unset = file.path(
  .analysis_project_dir("cache", "omip111-python", root_dir, create = FALSE), "bin", "python"))
path_importer <- file.path(root_dir, "scripts", "python", "omip111_masks.py")
if (run_preprocessing && !file.exists(path_python)) stop("FlowKit Python environment is missing.")
if (run_methods) .acsCytofEnsureCurrentCheckout(root_dir)
```

```{r}
#| label: scientific-settings
scientific_settings <- list(
  semantics = "omip111-v4", seed = 20261009L,
  preprocessing = list(reference = "raw cytokine lower cutoff on common non-naive parent",
    cofactor = 150, parentRelativeTolerance = 0.02,
    frequencyTolerancePp = 0.1),
  control = list(bwMtd = "nrd0", bwScope = "cytokine", bwNcellMax = 1e4,
    bwNcellMin = 1e4, bwFallback = "auto", bwMin = "none", bwMax = "none",
    biasUnsFactor = 1, gateCombn = "min", clusterGates = TRUE,
    calcCytPosGates = TRUE, minCell = 100, locThresholdMethod = "cap"),
  comparators = list(
    fbeta = .acsCytofComparatorSettings("fbeta_default")$params,
    tailgate = .acsCytofComparatorSettings("tailgate_default")$params))
```

Before fitting any method, raw unmixed cytokine intensities are transformed
with `asinh(x / 150)`, retaining negative values and casting to the same float32
precision used by the package. Cofactor 150 is a conventional starting point
for flow cytometry ([workflow reference](https://pmc.ncbi.nlm.nih.gov/articles/PMC5473464/));
it is fixed here, not tuned against the projected reference. The authors' display
transformations are used only for reconstructing their gates. Both examined
importers differ from the stored FlowJo counts, so exact event reproduction is
not claimed. Diagnostic guardrails require every non-naïve parent count to
agree within 2%. Per-tube discrepancies above 0.1 percentage points flag the
reconstructed **two-dimensional** author gates in the contextual audit. The
primary one-dimensional reference is calculated directly from raw expression
and is not subject to that display-coordinate reconstruction discrepancy.
These are import checks, not statements of biological accuracy. Actual differences
are reported below. The CSV retains the original FlowJo two-dimensional net
frequencies as context, clearly separate from the projected cytokine-cut reference.

StimGate uses shared per-cytokine bandwidths, bandwidth-based unstimulated bias,
threshold sharing and cytokine-positive refinement, with the same control
settings as the ACS real-data analysis. We score its final `loc_minClust`
**combination counts**, marginalised to each cytokine; this preserves the
conditional refinement actually applied by the package.

F-beta uses β = 0.8, θ = 2, smoothing width 10 and automatic bin count.
The existing Richards implementation uses common histogram edges spanning
both tubes, rather than the original unstimulated-only range. Tailgate uses
the stimulated distribution, automatic bandwidth and automatic derivative
tolerance, without an added bias. We retain zero and negative fluorescence
values, with no ACS-specific CyTOF zero-spike adjustments. Every comparator
threshold is applied to both matched tubes using strict `x > threshold`.
Genuine no-cutpoint outcomes retain an explicitly labelled empty-gate fallback;
runtime errors retain missing estimates and counts.

The main estimand is **unclipped** stimulated minus unstimulated frequency.
Negative estimates are retained. The primary reference uses the cytokine lower boundary from each tube's author
rectangle, applied alone to all cells in the common non-naïve parent. The author
rectangles also require a CD44 interval; applying their complete 2-D counts as a
reference for unrestricted cytokine positivity would change the estimand. We
therefore project only the cytokine lower cutoff, omit the CD44 restriction and
the finite upper display extent, and retain counts beyond that upper extent in
the audit CSV. This projected reference is **derived from**, rather than
identical to, the published manual frequencies. It is an imperfect one-dimensional
reference, not biological truth. Primary reference subtraction uses each tube's
own projected cutoff; automated subtraction uses the stimulated sample's
final gating rule in both tubes. Errors are expressed in percentage points.
Each mouse is the independent biological unit; markers and cell populations
from one mouse are repeated outcomes, not extra independent replicates.

## Preparation and method execution

```{r}
#| label: run-analysis
#| include: false
if (run_preprocessing) .omip111Preprocess(path_raw, path_prepared, scientific_settings$preprocessing, path_python, path_importer)
if (run_methods) {
  if (!file.exists(file.path(path_prepared, "COMPLETE"))) stop("Validated OMIP-111 inputs are missing.")
  prepared_inputs <- readRDS(file.path(path_prepared, "inputs.rds"))
  .omip111Run(prepared_inputs, path_results, scientific_settings, path_fbeta)
}
if (!file.exists(file.path(path_results, "COMPLETE"))) stop("Complete OMIP-111 results are missing; run the analysis first.")
comparison_object <- readRDS(file.path(path_results, "result.rds"))
if (!identical(comparison_object$settings, scientific_settings)) stop("Saved scientific settings differ; rerun methods.")
comparison_results <- comparison_object$results
comparison_summary <- .omip111Summary(comparison_results)
```

## Validation and method coverage

```{r}
#| label: validation
#| results: asis
if (isTRUE(run_plots)) {
  prepared_inputs <- readRDS(file.path(path_prepared, "inputs.rds"))
  if (!identical(prepared_inputs$manifest, comparison_object$preprocessing)) {
    stop("Prepared inputs differ from the method comparison manifest.")
  }
  cat("FlowKit import audited **", nrow(prepared_inputs$validation),
    "** population counts against the saved FlowJo counts. Maximum absolute count difference across the full tree: **",
    max(abs(prepared_inputs$validation$difference)), "**.\n\n", sep = "")
  ref_audit <- prepared_inputs$references
  cat("Maximum non-naïve parent-count difference: **",
    round(100 * max(abs(ref_audit$importedParent / ref_audit$flowjoParent - 1)), 3),
    "%**. Maximum per-tube manual cytokine-frequency difference: **",
    round(max(abs(ref_audit$frequencyDifferencePp)), 4), " percentage points**.\n\n", sep = "")
  .analysis_report_table(ref_audit, c(fig_key, "reference-audit.csv"),
    "Full 2-D author-gate reproduction and projected 1-D reference counts.", root_dir)
  .analysis_report_table(prepared_inputs$validation, c(fig_key, "gate-validation.csv"), "Gate-count validation.", root_dir)
  .analysis_report_table(prepared_inputs$populations, c(fig_key, "parent-populations.csv"), "Parent-population cell counts.", root_dir)
  coverage <- comparison_results |>
    dplyr::group_by(.data$method) |>
    dplyr::summarise(intended = dplyr::n(), finite = sum(is.finite(.data$propBs)),
      reference_eligible = sum(.data$referenceAccepted),
      errors = sum(!is.na(.data$error)), fallbacks = sum(.data$thresholdFallbackUsed, na.rm = TRUE),
      .groups = "drop")
  print(knitr::kable(coverage))
  .analysis_report_table(comparison_results, c(fig_key, "comparison.csv"), "Every individual comparison, including failure and threshold provenance.", root_dir)
  .analysis_report_table(comparison_summary, c(fig_key, "summary.csv"), "Per-strain, population and marker errors with finite-mouse coverage.", root_dir)
}
```

## Cell classification and frequency correlation

Classification uses only stimulated tubes; unstimulated tubes have no
responders. The primary manual reference is the raw cytokine lower cutoff
on the common parent cells. The CD44-restricted author masks remain context.
Manual gates are an imperfect reference. The mouse table retains every
TP, FP, FN and TN count and each metric's denominator. The summary pools
counts within strain, population, marker and method, and reports medians over
mice with the number of defined mouse metrics. Failed comparator gates remain
missing and contribute only to intended coverage.

```{r}
#| label: classification-tables
#| results: asis
if (isTRUE(run_plots) && exists("comparison_results") && nrow(comparison_results) > 0L) {
  report_results <- comparison_results
  classification <- .omip111Classification(prepared_inputs, report_results)
  .analysis_report_table(classification, c(fig_key, "classification-mice.csv"),
    "Stimulated-cell classification against the imperfect manual reference for every mouse.", root_dir)
  .analysis_report_table(.omip111ClassificationSummary(classification),
    c(fig_key, "classification.csv"),
    "Pooled cell counts and metric medians over mice, with defined coverage and denominators.", root_dir)
}
```

Pearson correlation measures association and Lin concordance correlation
measures agreement with the imperfect manual reference. Both use net
frequencies, retaining negative values, across mice within strain, population
and marker, and pooled across strains and markers within each population.
The pooled rows combine different markers and strains and are descriptive.
The table records intended and finite pairs; correlations are missing with
fewer than three finite pairs or zero variance in either frequency.

```{r}
#| label: frequency-correlations
#| results: asis
if (isTRUE(run_plots) && exists("report_results") && nrow(report_results) > 0L) {
  correlations <- .omipFrequencyCorrelationTable(report_results,
    c("strain", "population", "marker"), "propBs", "manualBs", "population")
  .analysis_report_table(correlations, c(fig_key, "frequency-correlations.csv"),
    "Pearson and population-moment Lin concordance correlations against manual net frequencies.", root_dir)
}
```

## Frequency agreement and signed errors

Each point represents one matched mouse and cytokine. The identity line marks
agreement with the projected cytokine-cut reference. The signed-error plots retain exact zeros
and negative net frequencies. No confidence intervals are inferred from the
80 marker/population outcomes.

```{r}
#| label: agreement-figures
#| results: asis
#| fig-width: 10
#| fig-height: 7
if (isTRUE(run_plots)) {
  figure_dir <- .analysis_fig_dir(c(fig_key, "agreement"), root_dir)
  method_colours <- c(StimGate = "#0072B2", `F-beta` = "#D55E00", Tailgate = "#009E73")
  for (population in c("CD4", "CD8")) {
    cat("\n\n### Non-naïve ", population, "\n\n", sep = "")
    plot_data <- comparison_results[comparison_results$population == population & comparison_results$referenceAccepted, ]
    plot_data$method <- factor(plot_data$method, levels = names(method_colours))
    p <- ggplot2::ggplot(plot_data, ggplot2::aes(100 * manualBs, 100 * propBs,
      colour = method, shape = method)) +
      ggplot2::geom_abline(slope = 1, intercept = 0, colour = "grey60") +
      ggplot2::geom_point(size = 1.7, alpha = 0.8, na.rm = TRUE) +
      ggplot2::facet_wrap(ggplot2::vars(strain, marker), scales = "free", ncol = 5) +
      ggplot2::scale_colour_manual(values = method_colours) +
      ggplot2::scale_y_continuous(labels = .analysis_label_number) +
      ggplot2::scale_x_continuous(labels = .analysis_label_number) +
      ggplot2::labs(x = "Projected cytokine-cut net frequency (%)", y = "Automated net frequency (%)") +
      .analysis_theme()
    print(p)
    ggplot2::ggsave(file.path(figure_dir, paste0(population, "-frequency-agreement.png")), p,
      width = 10, height = 7, units = "in", dpi = 150)
    p_error <- ggplot2::ggplot(plot_data, ggplot2::aes(method, errorPp, colour = method)) +
      ggplot2::geom_hline(yintercept = 0, colour = "grey60") +
      ggplot2::geom_point(ggplot2::aes(group = mouse), position = ggplot2::position_dodge(width = 0.4),
        size = 1.7, alpha = 0.8, na.rm = TRUE) +
      ggplot2::facet_wrap(ggplot2::vars(strain, marker), scales = "free_y", ncol = 5) +
      .analysis_y_floor(c(-10, 10)) +
      ggplot2::scale_colour_manual(values = method_colours) +
      ggplot2::scale_y_continuous(labels = .analysis_label_number) +
      ggplot2::labs(x = NULL, y = "Automated − projected reference (percentage points)") +
      .analysis_theme() + ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
    print(p_error)
    ggplot2::ggsave(file.path(figure_dir, paste0(population, "-signed-error.png")), p_error,
      width = 10, height = 7, units = "in", dpi = 150)
  }
}
```

## Investigate individual gates

These are the same diagnostic panels used for ACS in Analysis 13, for **every
mouse, population and marker**. Expand a mouse below to inspect the control and
stimulated densities, response probability and its derivative, expected responding
cells per bin, and taut-string/filter boundaries. The lines overlay the initial
gate, the final StimGate gate and conditional cytokine-positive gate, F-beta,
Tailgate, and the actual author cytokine lower cutoffs transformed to `asinh(raw / 150)`.
Overlapping stimulated/control author cutoffs can appear as one line.

The purple author lines are the projected single-channel reference, rather than
the full CD44/cytokine rectangles. The result block uses the saved Analysis 14
frequencies: StimGate's combination-count result may differ from applying just
its base gate to every cell. All samples are rerun together within each strain
and population, with the original settings and seed, to preserve gate sharing.
Both final base and conditional gates must exactly reproduce the saved comparison
before any pages are published. Each PDF contains all five mice for its marker.

`RUN_DIAGNOSTICS=true` regenerates these panels without requiring preprocessing
or comparator refitting. Set `RUN_DIAGNOSTICS=false` to reuse validated pages.

```{r}
#| label: individual-gate-diagnostics
#| results: asis
#| echo: false
#| out-width: 100%
if (isTRUE(run_plots)) {
  diagnostic_dir <- .analysis_fig_dir(c(fig_key, "gate-diagnostics"), root_dir)
  if (run_diagnostics) {
    .acsCytofEnsureCurrentCheckout(root_dir)
    .omip111DebugRun(prepared_inputs, comparison_results, scientific_settings,
      diagnostic_dir, path_results)
  }
  if (!file.exists(file.path(diagnostic_dir, "COMPLETE"))) stop("Gate diagnostics are missing; use RUN_DIAGNOSTICS=true.")
  diagnostic_object <- readRDS(file.path(diagnostic_dir, "diagnostics.rds"))
  if (!identical(diagnostic_object$settings, scientific_settings) ||
    !identical(diagnostic_object$context, prepared_inputs$manifest)) {
    stop("Saved gate diagnostics are stale; use RUN_DIAGNOSTICS=true.")
  }
  diagnostic_index <- diagnostic_object$index
  .analysis_report_table(diagnostic_index, c(fig_key, "gate-diagnostic-index.csv"),
    "All 80 diagnostic pages, with PDF page numbers and saved-gate parity.", root_dir)
  for (strain in c("C57", "BALB")) {
    for (population in c("CD4", "CD8")) {
      cat("\n\n### ", strain, " non-naïve ", population, "\n\n", sep = "")
      subset <- diagnostic_index[diagnostic_index$strain == strain & diagnostic_index$population == population, ]
      for (marker in unique(subset$marker)) {
        cat("\n\n#### ", marker, "\n\n", sep = "")
        pages <- subset[subset$marker == marker, ]
        for (i in seq_len(nrow(pages))) {
          cat("\n<details><summary>", pages$mouse[[i]], " — ", marker,
            " (PDF page ", pages$page[[i]], ")</summary>\n\n", sep = "")
          cat("![Gate diagnostics for ", pages$mouse[[i]], " ", population,
            " ", marker, "](", knitr::image_uri(file.path(diagnostic_dir, pages$png[[i]])),
            ")\n", sep = "")
          cat("\n\n</details>\n\n")
        }
      }
    }
  }
}
```

## Observed agreement

All three methods completed 80 outcomes without runtime errors. One StimGate
CD4 IL-17A outcome used a recorded shared-gate fallback. The full 2-D author
reference remains contextual: its CD44 restriction can change the net frequency
by almost 80 percentage points relative to the projected cytokine cutoff.
No events in these parent populations exceed the authors' cytokine upper extents.


```{r}
#| label: observed-agreement
#| results: asis
if (isTRUE(run_plots)) {
  descriptive_errors <- comparison_results |>
    dplyr::group_by(.data$method) |>
    dplyr::summarise(eligible_outcomes = sum(is.finite(.data$errorPp)),
      mean_absolute_error_pp = mean(.data$absoluteErrorPp, na.rm = TRUE),
      median_absolute_error_pp = stats::median(.data$absoluteErrorPp, na.rm = TRUE),
      .groups = "drop")
  cat("Descriptive errors across retained marker/population outcomes, in percentage points. ",
    "Outcomes from one mouse are repeated measurements; these averages do not provide independent-replicate confidence intervals.\n\n", sep = "")
  print(knitr::kable(descriptive_errors, digits = 3))
}
```

The full two-dimensional author gates require CD44-high expression as well as
cytokine positivity. Their frequencies can be much lower than cytokine-cut
positivity across the whole non-naïve compartment, especially for IL-2 and TNF.
Those contextual differences must not be attributed to cytokine-threshold
performance. The plots and error summaries above use the projected single-channel
reference to compare like estimands.

OMIP-111 is useful as an external agreement dataset and a stress test of
background matching. A lower pooled error on this small panel should not be
interpreted as general method superiority. The control and stimulated non-naïve
populations have markedly different compositions, and the projected thresholds
were originally drawn in a CD44-high context. Transformation sensitivity and
expert review remain appropriate before claims about biological accuracy.

## Interpretation limits

The full CSV companions retain per-mouse, per-marker results and all missing
outcomes. Agreement with manual gating supports external plausibility; it
cannot establish biological sensitivity or false discovery rate. The controls
have many fewer non-naïve cells than the stimulated tubes, so background
estimation precision differs by condition. The two mouse strains, five mice
per strain, one acquisition setting and mitogen stimulus provide a bounded
external validation rather than a representative survey of use cases.
