---
title: OMIP-111 comparison with the shifted-peak rule
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_methods: null
  run_plots: true
  run_diagnostics: true
---

## What this document is

This is the **shifted-peak rule on** version of Analysis 14
(`14-real-compare-omip111.qmd`). Analysis 14 stays the current-default version.
Everything is the same as Analysis 14 (prepared inputs, settings, seed,
comparators, scoring) except one StimGate setting:
`stimControl(locShiftedPeakRef = TRUE)`.

StimGate normally starts looking for responding cells one third of a
left-window width above the higher of the stimulated and unstimulated main
peaks. After PMA/ionomycin, most cells make TNF, so the stimulated main peak
is made of responding cells and the search starts above most of them. With
the rule on, if the stimulated main peak lies more than 2 bandwidths to the
right of the unstimulated main peak, the search starts above the
unstimulated peak instead, using only the unstimulated tube's left-window
width. The bandwidth is the shared per-marker bandwidth estimated on 10,000
cells, without the widening for smaller tubes.

The results here are written to their own cache (`omip111/shifted-peak`) and
figure folders, and the Analysis 14 results are only read. Read Analysis 14
for the dataset, the reference and the limits of the comparison; they all
apply here.

```{r}
#| label: setup
#| include: false
root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)
scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))
source(file.path(scripts_r_dir, "omip016-methods.R"))
source(file.path(scripts_r_dir, "omip111.R"))
source(file.path(scripts_r_dir, "sim-debug-loc.R"))
source(file.path(scripts_r_dir, "acs_cytof-debug.R"))
source(file.path(scripts_r_dir, "omip111-debug.R"))
source(file.path(scripts_r_dir, "omip-shifted-peak.R"))
run_simulations <- .as_flag(.get_qmd_param_env("run_simulations", "RUN_SIMULATIONS", FALSE))
run_methods <- .as_flag(.get_qmd_param_env("run_methods", "RUN_METHODS", run_simulations))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
run_diagnostics <- .as_flag(.get_qmd_param_env("run_diagnostics", "RUN_DIAGNOSTICS", TRUE))
fig_key <- .analysis_mode_key("14b-real-compare-omip111-shifted-peak", sized = FALSE)
path_prepared <- .analysis_project_dir("cache", c("omip111", "prepared"), root_dir, create = FALSE)
path_default <- .analysis_project_dir("cache", c("omip111", "current"), root_dir, create = FALSE)
path_results <- .analysis_project_dir("cache", c("omip111", "shifted-peak", "current"), root_dir, create = FALSE)
path_fbeta <- file.path(root_dir, "scripts", "python", "fbeta.py")
if (run_methods) .acsCytofEnsureCurrentCheckout(root_dir)
```

```{r}
#| label: scientific-settings
# Analysis 14's settings, with only the shifted-peak rule switched on.
default_settings <- list(
  semantics = "omip111-v4", seed = 20261009L,
  preprocessing = list(reference = "raw cytokine lower cutoff on common non-naive parent",
    cofactor = 150, parentRelativeTolerance = 0.02,
    frequencyTolerancePp = 0.1),
  control = list(bwMtd = "nrd0", bwScope = "cytokine", bwNcellMax = 1e4,
    bwNcellMin = 1e4, bwFallback = "auto", bwMin = "none", bwMax = "none",
    biasUnsFactor = 1, gateCombn = "min", clusterGates = TRUE,
    calcCytPosGates = TRUE, minCell = 100, locThresholdMethod = "cap"),
  comparators = list(
    fbeta = .acsCytofComparatorSettings("fbeta_default")$params,
    tailgate = .acsCytofComparatorSettings("tailgate_default")$params))
scientific_settings <- .omipShiftedPeakSettings(default_settings)
```

```{r}
#| label: run-analysis
#| include: false
if (!file.exists(file.path(path_prepared, "COMPLETE"))) {
  stop("Analysis 14's prepared OMIP-111 inputs are missing; render Analysis 14 first.")
}
prepared_inputs <- readRDS(file.path(path_prepared, "inputs.rds"))
if (run_methods) .omip111Run(prepared_inputs, path_results, scientific_settings, path_fbeta)
if (!file.exists(file.path(path_results, "COMPLETE"))) stop("Shifted-peak OMIP-111 results are missing; run with RUN_SIMULATIONS=true.")
comparison_object <- readRDS(file.path(path_results, "result.rds"))
if (!identical(comparison_object$settings, scientific_settings)) stop("Saved shifted-peak settings differ; rerun methods.")
if (!identical(comparison_object$preprocessing, prepared_inputs$manifest)) stop("Prepared inputs differ from the shifted-peak results.")
comparison_results <- comparison_object$results
if (!file.exists(file.path(path_default, "COMPLETE"))) stop("Analysis 14 default results are missing.")
default_object <- readRDS(file.path(path_default, "result.rds"))
if (!identical(default_object$settings, default_settings)) stop("Analysis 14 default results use other settings.")
default_results <- default_object$results
fired <- .omipShiftedPeakFired(
  file.path(path_results, rep(c("C57", "BALB"), each = 2L), c("CD4", "CD8"), "stimgate"),
  labels = data.frame(strain = rep(c("C57", "BALB"), each = 2L), population = c("CD4", "CD8")))
paired <- .omip111ShiftedPeakPaired(default_results, comparison_results, fired, prepared_inputs$samples)
```

## How often the rule applied

The table counts the stimulated tubes (mice) where the rule applied, for each
strain, population and marker. It reads the `locShiftedPeakRef` flag that
StimGate records for each tube's own local-FDR gate. Threshold sharing and
cytokine-positive refinement can still move a tube's final gate.

```{r}
#| label: fired
#| results: asis
fired_summary <- paired |>
  dplyr::filter(.data$method == "StimGate") |>
  dplyr::group_by(.data$strain, .data$population, .data$marker) |>
  dplyr::summarise(mice = dplyr::n(), rule_applied = sum(.data$shiftedPeakRef),
    .groups = "drop")
cat("The rule applied for **", sum(fired_summary$rule_applied), "** of **",
  sum(fired_summary$mice), "** StimGate tube/marker outcomes.\n\n", sep = "")
print(knitr::kable(fired_summary))
comparators_unchanged <- .omipShiftedPeakComparatorsUnchanged(paired)
cat("\nF-beta and Tailgate results are identical to Analysis 14: **",
  if (comparators_unchanged) "yes" else "no", "**.\n\n", sep = "")
```

## Cell classification and frequency correlation

Classification uses only stimulated tubes; unstimulated tubes have no
responders. The primary manual reference is the raw cytokine lower cutoff
on the common parent cells. The CD44-restricted author masks remain context.
Manual gates are an imperfect reference. The mouse table retains every
TP, FP, FN and TN count and each metric's denominator. The summary pools
counts within strain, population, marker and method, and reports medians over
mice with the number of defined mouse metrics. Failed comparator gates remain
missing and contribute only to intended coverage.

```{r}
#| label: classification-tables
#| results: asis
if (isTRUE(run_plots) && exists("comparison_results") && nrow(comparison_results) > 0L) {
  if (!identical(default_object$preprocessing, prepared_inputs$manifest)) {
    stop("Prepared inputs differ from the default OMIP-111 results.")
  }
  report_results <- comparison_results
  report_results$method[report_results$method == "StimGate"] <- "StimGate shifted peak"
  default_sg <- default_results[default_results$method == "StimGate", ]
  default_sg$method <- "StimGate default"
  report_results <- dplyr::bind_rows(report_results, default_sg)
  classification <- .omip111Classification(prepared_inputs, report_results)
  .analysis_report_table(classification, c(fig_key, "classification-mice.csv"),
    "Stimulated-cell classification against the imperfect manual reference for every mouse.", root_dir)
  .analysis_report_table(.omip111ClassificationSummary(classification),
    c(fig_key, "classification.csv"),
    "Pooled cell counts and metric medians over mice, with defined coverage and denominators.", root_dir)
}
```

Pearson correlation measures association and Lin concordance correlation
measures agreement with the imperfect manual reference. Both use net
frequencies, retaining negative values, across mice within strain, population
and marker, and pooled across strains and markers within each population.
The pooled rows combine different markers and strains and are descriptive.
The table records intended and finite pairs; correlations are missing with
fewer than three finite pairs or zero variance in either frequency.

**Summary by population.** Counts are summed over strains, markers and mice
within each population before the proportions are calculated; Pearson and
concordance correlations (CCC) are the pooled rows of the CSV. `n_defined`
counts outcomes with a usable gate, and `n_pairs` the finite frequency pairs.
The summary table follows the link to the full CSV.

```{r}
#| label: frequency-correlations
#| results: asis
if (isTRUE(run_plots) && exists("report_results") && nrow(report_results) > 0L) {
  correlations <- .omipFrequencyCorrelationTable(report_results,
    c("strain", "population", "marker"), "propBs", "manualBs", "population")
  .analysis_report_table(correlations, c(fig_key, "frequency-correlations.csv"),
    "Pearson and population-moment Lin concordance correlations against manual net frequencies.", root_dir)
  .omipAgreementKable(.omipAgreementOverview(classification, correlations, "population"))
}
```

## StimGate errors with and without the rule

Errors are StimGate's net frequency minus the projected author cytokine-cut
reference, in percentage points, as in Analysis 14. Each row averages the five
mice of one strain, population and marker. "Default" is Analysis 14's saved
result.

```{r}
#| label: error-table
#| results: asis
error_summary <- .omipShiftedPeakErrorSummary(paired, c("strain", "population", "marker"))
print(knitr::kable(error_summary, digits = 2))
if (isTRUE(run_plots)) {
  .analysis_report_table(paired, c(fig_key, "paired-comparison.csv"),
    "Every outcome under the default and shifted-peak settings.", root_dir)
  .analysis_report_table(error_summary, c(fig_key, "error-summary.csv"),
    "Mean StimGate error per strain, population and marker.", root_dir)
}
```

Each point below is one mouse. Lines join the same mouse's default and
shifted-peak StimGate errors; filled points mark tubes where the rule applied.

```{r}
#| label: error-figure
#| fig-width: 10
#| fig-height: 7
if (isTRUE(run_plots)) {
  figure_dir <- .analysis_fig_dir(c(fig_key, "stimgate-error"), root_dir)
  p <- .omipShiftedPeakErrorPlot(paired, facet = c("strain", "population"))
  print(p)
  ggplot2::ggsave(file.path(figure_dir, "stimgate-error-default-vs-shifted-peak.png"), p,
    width = 10, height = 7, units = "in", dpi = 150)
}
```

## All three methods with the rule on

```{r}
#| label: method-errors
#| results: asis
descriptive_errors <- comparison_results |>
  dplyr::group_by(.data$method) |>
  dplyr::summarise(eligible_outcomes = sum(is.finite(.data$errorPp)),
    mean_absolute_error_pp = mean(.data$absoluteErrorPp, na.rm = TRUE),
    median_absolute_error_pp = stats::median(.data$absoluteErrorPp, na.rm = TRUE),
    .groups = "drop")
print(knitr::kable(descriptive_errors, digits = 3))
```

## Gate diagnostics

These are Analysis 14's per-mouse diagnostic pages, rerun with the rule on,
for every mouse, population and marker. The rerun must reproduce every saved
final gate exactly before any page is shown. Each summary gives StimGate's
error with and without the rule and marks the outcomes where the rule applied.

```{r}
#| label: individual-gate-diagnostics
#| results: asis
#| echo: false
#| out-width: 100%
if (isTRUE(run_plots)) {
  diagnostic_dir <- .analysis_fig_dir(c(fig_key, "gate-diagnostics"), root_dir)
  if (run_diagnostics) {
    .acsCytofEnsureCurrentCheckout(root_dir)
    .omip111DebugRun(prepared_inputs, comparison_results, scientific_settings,
      diagnostic_dir, path_results)
  }
  if (!file.exists(file.path(diagnostic_dir, "COMPLETE"))) stop("Gate diagnostics are missing; use RUN_DIAGNOSTICS=true.")
  diagnostic_object <- readRDS(file.path(diagnostic_dir, "diagnostics.rds"))
  if (!identical(diagnostic_object$settings, scientific_settings) ||
    !identical(diagnostic_object$context, prepared_inputs$manifest)) {
    stop("Saved gate diagnostics are stale; use RUN_DIAGNOSTICS=true.")
  }
  diagnostic_index <- diagnostic_object$index
  .analysis_report_table(diagnostic_index, c(fig_key, "gate-diagnostic-index.csv"),
    "All diagnostic pages, with PDF page numbers and saved-gate parity.", root_dir)
  sg_rows <- paired[paired$method == "StimGate", ]
  k <- match(
    paste(diagnostic_index$strain, diagnostic_index$population, diagnostic_index$marker, diagnostic_index$mouse),
    paste(sg_rows$strain, sg_rows$population, sg_rows$marker, sg_rows$mouse)
  )
  if (anyNA(k) || nrow(diagnostic_index) != nrow(sg_rows)) {
    stop("Diagnostic pages do not match the StimGate outcomes.")
  }
  diagnostic_index$note <- sprintf(": error %+.1f pp; default %+.1f pp%s",
    sg_rows$errorPp[k], sg_rows$errorPpDefault[k],
    ifelse(sg_rows$shiftedPeakRef[k], "; rule applied", ""))
  .omip111DiagnosticPages(diagnostic_index, diagnostic_dir)
}
```
