---
title: Compare StimGate, Tailgate and F-beta on OMIP-016
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_preprocessing: null
  run_stimgate: null
  run_comparators: null
  run_plots: true
---

Validate StimGate on the OMIP-016 intracellular cytokine staining data
(FlowRepository FR-FCM-ZZ2T; one donor, unstimulated, Gag-peptide and SEB
tubes acquired on an LSRII) and compare its gates with Tailgate, F-beta and
the authors' manual gates in CD4 T cells.

## Set-up

`RUN_SIMULATIONS=true` prepares the data and runs all three methods;
`RUN_SIMULATIONS=false RUN_PLOTS=true` reads the saved results. The
individual `RUN_PREPROCESSING`, `RUN_STIMGATE` and `RUN_COMPARATORS`
controls override those stages when supplied. Preparation needs the
FlowRepository files under the `raw-data-large` projr path `OMIP-016`
(only read), `python3` for the workspace decoder, and the curated maps
under `raw-data-small/comparison_data/omip016`. Prepared data are written to
the `cache` projr path `omip016/prepared` and method results to
`omip016/methods`.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("15-real-compare-omip016", sized = FALSE)
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-misc.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))
source(file.path(scripts_r_dir, "sim-debug-loc.R"))
source(file.path(scripts_r_dir, "acs_cytof-debug.R"))
source(file.path(scripts_r_dir, "omip016-prepare.R"))
source(file.path(scripts_r_dir, "omip016-methods.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations", "RUN_SIMULATIONS", FALSE
))
run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing", "RUN_PREPROCESSING", run_simulations
))
run_stimgate <- .as_flag(.get_qmd_param_env(
  "run_stimgate", "RUN_STIMGATE", run_simulations
))
run_comparators <- .as_flag(.get_qmd_param_env(
  "run_comparators", "RUN_COMPARATORS", run_simulations
))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))

.acsCytofEnsureCurrentCheckout(root_dir)

# Setup starts in the QMD directory; keep configured paths absolute.
path_raw <- projr::projr_path_get("raw-data-large", "OMIP-016", format = "absolute")
path_small <- projr::projr_path_get(
  "raw-data-small", "comparison_data", "omip016", format = "absolute"
)
path_prepared <- projr::projr_path_get("cache", "omip016", "prepared", format = "absolute")
path_methods <- projr::projr_path_get("cache", "omip016", "methods", format = "absolute")
path_fbeta <- file.path(root_dir, "scripts", "python", "fbeta.py")
if (isTRUE(run_preprocessing) && !dir.exists(path_raw)) {
  stop("OMIP-016 raw directory does not exist: ", path_raw)
}
if (isTRUE(run_comparators)) {
  if (!requireNamespace("cytoUtils", quietly = TRUE) ||
      !requireNamespace("reticulate", quietly = TRUE) || !file.exists(path_fbeta)) {
    stop("OMIP-016 comparisons require cytoUtils, reticulate and scripts/python/fbeta.py.")
  }
  if (!reticulate::py_module_available("numpy")) {
    stop("F-beta requires numpy in the Python environment selected by reticulate.")
  }
}
```

## Scientific settings

StimGate uses Analysis 9's settings: `bwMtd = "nrd0"` with
`bwNcellMax = 1e4`, shared per-marker bandwidths, `gateCombn = "min"`,
`minCell = 100`, threshold clustering and cytokine-positive refinement on,
`locThresholdMethod = "cap"` and `biasUns` equal to the shared bandwidth. The
comparison scores the final `loc_minClust` gates. Tailgate (stimulated tube,
automatic tolerance) and F-beta (`beta = 0.8`, `theta = 2`, `width = 10`)
keep their published defaults: these are conventional flow data without the
CyTOF zero spike that motivated Analysis 9's tuned versions, and tuning them
against these manual gates would make the comparison circular.

```{r}
#| label: scientific-settings
analysis_seed <- 20261009L
stimgate_settings <- .omip016StimGateSettings(
  locThresholdMethod = "cap", clusterGates = TRUE, calcCytPosGates = TRUE
)
```

## Data preparation

The FCS files are uncompensated (`$SPILL` is the identity). The authors'
legacy binary FlowJo for Mac workspace holds the compensation matrix
(`mtx 20110118 PYY1002 SG`) and the gating tree; neither CytoML nor FlowKit
reads this format, so `scripts/python/omip016_flowjo_jo.py` decodes them and
the gates are re-applied to the raw events. Everything is rebuilt in a
temporary sibling and swapped in only on success.

```{r}
#| label: prepare
#| include: false
if (isTRUE(run_preprocessing)) {
  .omip016Prepare(
    pathRaw = path_raw, pathSmall = path_small,
    pathOut = path_prepared, pathRoot = root_dir
  )
}
prepared_paths <- .omip016Paths(path_prepared)
if (!file.exists(prepared_paths$manifest)) {
  stop("Prepared OMIP-016 data not found at ", path_prepared,
       "; render with RUN_PREPROCESSING=true.")
}
prepared_manifest <- readRDS(prepared_paths$manifest)
read_qc <- function(...) utils::read.csv(file.path(path_prepared, ...), check.names = FALSE)
raw_validation <- read_qc("qc", "raw_validation.csv")
population_counts <- read_qc("qc", "population_counts.csv")
convention_sensitivity <- read_qc("qc", "convention_sensitivity.csv")
gate_vs_threshold <- read_qc("qc", "manual_gate_vs_threshold.csv")
bead_spill <- read_qc("qc", "compensation_beads_vs_workspace.csv")
manual_thresholds <- read_qc("reference", "manual_thresholds.csv")
manual_frequencies <- read_qc("reference", "manual_frequencies.csv")
sample_map <- .omip016SampleMap(path_small)
```

### Files

All `r nrow(raw_validation)` deposited files match the recorded sizes and MD5
checksums (`r sum(raw_validation$md5_ok)` of `r nrow(raw_validation)`). Cell
tubes and controls:

```{r}
#| label: sample-map
knitr::kable(sample_map[, c("file", "role", "SampleID", "stim", "well_id", "acquired")])
```

The workspace also lists `CTRL_CELL SEULES_G01.fcs`, `CTRL_BLUEVID_G02.fcs`
and `13170 17012011_NS_A01.fcs`, which were not deposited; none is needed
to re-apply the gates. The viability dye (Blue Vid) has no deposited single-stain
control, and the IL-2 (APC) channel's control is a CD8-APC bead acquired the
next day.

### Compensation

The bead single stains give an independent estimate of the spillover matrix.
Agreement with the decoded workspace matrix (off-diagonal entries):

```{r}
#| label: compensation
off <- bead_spill[bead_spill$fluorochrome_channel != bead_spill$detector, ]
off$difference <- off$spill_beads - off$spill_workspace
knitr::kable(data.frame(
  entries = nrow(off),
  max_abs_difference = max(abs(off$difference)),
  largest_entry = with(off[which.max(abs(off$difference)), ],
                       paste0(fluorochrome_channel, " -> ", detector, ": ",
                              signif(spill_beads, 3), " (beads) vs ",
                              signif(spill_workspace, 3), " (workspace)"))
), digits = 3)
```

### Re-applied manual gating

Counts for the populations leading to CD4 T cells and their response gates:

```{r}
#| label: population-counts
main_paths <- c(
  "Single cells", "Single cells/Lymphocytes", "Single cells/Lymphocytes/Lives",
  "Single cells/Lymphocytes/Lives/CD3+", "Single cells/Lymphocytes/Lives/CD3+/CD4+"
)
pc <- population_counts[population_counts$path %in% main_paths, ]
pc$stim <- sample_map$stim[match(pc$file, sample_map$file)]
pc$population <- basename(pc$path)
knitr::kable(
  stats::reshape(pc[, c("population", "stim", "pct_parent")], idvar = "population",
                 timevar = "stim", direction = "wide"),
  digits = 1, row.names = FALSE, col.names = c("Population", paste0("% parent, ", unique(pc$stim)))
)
```

The workspace stores no statistics, so the reconstruction cannot be checked
against FlowJo's own counts. Two FlowJo conventions were assumed. Extending
gate vertices drawn below zero on compensated axes to the bottom of the data
(CytoML's default) changes the CD4 population by at most
`r max(abs(with(convention_sensitivity[convention_sensitivity$variant == "no_edge_extension" & convention_sensitivity$path == main_paths[5], ], count_variant - count_primary)))`
cells per tube. The CD4 gate is the Boolean `G1 | G4 | G5 | G2 | G3 & ! G0`
(the union of five CD4 gates drawn against each response marker, excluding
`CD8+ (to excl)`); it is evaluated from left to right, so the CD8 exclusion
applies to the whole union, as its name intends. Binding `&` before `|`
instead would exclude CD8 cells only from the TNF-alpha CD4 gate:

```{r}
#| label: boolean-sensitivity
bs <- convention_sensitivity[convention_sensitivity$variant == "boolean_precedence" &
  startsWith(convention_sensitivity$path, main_paths[5]), ]
bs$stim <- sample_map$stim[match(bs$file, sample_map$file)]
bs$population <- sub(paste0(main_paths[5], "/?"), "CD4+ ", bs$path)
bs <- bs[!grepl("CD45RA", bs$path), ]
knitr::kable(bs[, c("stim", "population", "count_primary", "count_variant")],
             row.names = FALSE,
             col.names = c("Tube", "Population", "Left to right", "& before |"))
```

The difference is small for the unstimulated and Gag tubes but large for SEB,
where cytokine-bright CD8 T cells fall inside the CD4 gates' sloped lower
boundaries. The left-to-right reading is used throughout.

### Manual reference

Each manual response gate is a rectangle on the response marker against a
second parameter, used identically in all three tubes. Its lower bound on the
response marker is the manual threshold; on the GatingSet scale
(`asinh(x / 150)` of compensated values):

```{r}
#| label: manual-thresholds
knitr::kable(manual_thresholds[, c("marker", "channel", "threshold_raw", "threshold_trans", "y_param")],
             digits = 3, row.names = FALSE)
```

The rectangle's other bounds exclude at most
`r max(gate_vs_threshold$n_disagree)` CD4 T cells above a threshold in any
tube. The reference labels are the rectangles themselves. Manual frequencies
in CD4 T cells (%; background-subtracted values are clipped at zero, as in
Analysis 9):

```{r}
#| label: manual-frequencies
knitr::kable(manual_frequencies[, c("stim", "cyt", "n_cell", "count_man",
                                    "freq_stim_man", "freq_bs_man")],
             digits = 3, row.names = FALSE)
```

Histograms of every response channel in CD4 T cells, with the manual
thresholds, are saved with the prepared data as
`omip016/prepared/qc/cd4_response_channels.png` in the projr cache.

## Methods

```{r}
#| label: run-stimgate
#| include: false
omip016_prep <- .omip016ReadPrepared(path_prepared)
path_stimgate <- file.path(path_methods, "stimgate")
if (isTRUE(run_stimgate)) {
  set.seed(analysis_seed)
  dir.create(path_methods, recursive = TRUE, showWarnings = FALSE)
  .acsCytofReplaceDir(path_stimgate, function(path_tmp) {
    .omip016RunStimGate(omip016_prep, path_tmp, stimgate_settings)
  })
}
```

```{r}
#| label: run-comparators
#| include: false
path_comparators <- file.path(path_methods, "comparators.rds")
if (isTRUE(run_comparators)) {
  comparator_results <- lapply(c("tailgate", "fbeta"), function(method) {
    .omip016RunComparator(omip016_prep, method, pathFbeta = path_fbeta)
  })
  names(comparator_results) <- c("tailgate", "fbeta")
  comparator_results$preprocessing <- omip016_prep$pre[c("version", "inputContentHash", "settings")]
  .write_rds_atomic(comparator_results, path_comparators)
}
```

```{r}
#| label: collate
#| include: false
if (!dir.exists(path_stimgate) || !file.exists(path_comparators)) {
  stop("OMIP-016 method results are missing; render with RUN_SIMULATIONS=true.")
}
if (!file.exists(file.path(path_stimgate, "omip016-debug-records.rds"))) {
  stop("Saved StimGate result lacks its local-FDR records; re-run with RUN_STIMGATE=true.")
}
stimgate_manifest <- readRDS(file.path(path_stimgate, "omip016-manifest.rds"))
comparator_results <- readRDS(path_comparators)
current_pre <- omip016_prep$pre[c("version", "inputContentHash", "settings")]
if (!identical(stimgate_manifest$preprocessing, current_pre) ||
    !identical(comparator_results$preprocessing, current_pre)) {
  stop("OMIP-016 method results were made from other prepared data; re-run all methods.")
}
if (!identical(stimgate_manifest$settings, stimgate_settings)) {
  stop("Saved StimGate result used other settings; re-run with RUN_STIMGATE=true.")
}
for (method in c("tailgate", "fbeta")) {
  if (!identical(comparator_results[[method]]$settings, .omip016ComparatorSettings(method))) {
    stop("Saved ", method, " result used other settings; re-run with RUN_COMPARATORS=true.")
  }
}
gates <- rbind(
  .omip016StimGateGates(omip016_prep, path_stimgate),
  comparator_results$tailgate$gates,
  comparator_results$fbeta$gates,
  .omip016ManualGates(omip016_prep, manual_thresholds)
)
scored <- .omip016Score(omip016_prep, gates)
.omip016CheckStimGateCounts(omip016_prep, path_stimgate, scored)
scored$method <- factor(scored$method, levels = c("manual", .omip016Methods()))
scored$stim <- factor(scored$stim, levels = c("gag", "seb"))
scored$cyt <- factor(scored$cyt, levels = unname(omip016_prep$pre$responseChannels))
scored <- scored[order(scored$stim, scored$cyt, scored$method), ]
```

Cells are called positive as the package calls them: above the gate
(strict `x > gate`), or, for StimGate, above its lower cytokine-positive gate
(`gateCyt`) when ordinarily positive for another response marker. The
unstimulated tube is classified with the stimulated tube's gates. StimGate's
counts from `getStimStats()` equal these recomputed counts. Gates that are not
a method's own direct estimate:

```{r}
#| label: provenance
prov <- scored[scored$method != "manual", c("method", "stim", "cyt", "thresholdOrigin", "thresholdFallbackUsed")]
knitr::kable(as.data.frame(table(
  method = droplevels(prov$method), origin = prov$thresholdOrigin,
  fallback = prov$thresholdFallbackUsed
)) |> subset(Freq > 0), row.names = FALSE)
```

## Results

With one donor and two stimulated tubes, these are case-level comparisons,
not estimates of average performance; no intervals are reported.

### Gates

Gates on the GatingSet scale, `asinh(x / 150)`:

```{r}
#| label: gates
gate_wide <- stats::reshape(
  scored[, c("stim", "cyt", "method", "gate")],
  idvar = c("stim", "cyt"), timevar = "method", direction = "wide"
)
names(gate_wide) <- sub("^gate\\.", "", names(gate_wide))
knitr::kable(gate_wide, digits = 2, row.names = FALSE)
```

### Response frequencies

Background-subtracted frequencies (%, clipped at zero), each method's gate
applied to its stimulated tube and to the unstimulated tube:

```{r}
#| label: frequencies
freq_tbl <- scored[, c("method", "stim", "cyt", "freq_stim", "freq_uns", "freq_bs",
                       "freq_bs_man")]
freq_tbl$error_bs <- freq_tbl$freq_bs - freq_tbl$freq_bs_man
if (isTRUE(run_plots)) {
  .analysis_report_table(freq_tbl, c(fig_key, "frequencies.csv"),
                         "Frequencies by method, tube and marker.")
}
freq_wide <- stats::reshape(
  freq_tbl[freq_tbl$method != "manual", c("stim", "cyt", "method", "freq_bs")],
  idvar = c("stim", "cyt"), timevar = "method", direction = "wide"
)
names(freq_wide) <- sub("^freq_bs\\.", "", names(freq_wide))
freq_wide$manual <- freq_tbl$freq_bs_man[match(
  paste(freq_wide$stim, freq_wide$cyt),
  paste(freq_tbl$stim, freq_tbl$cyt)
)]
knitr::kable(freq_wide, digits = 3, row.names = FALSE)
```

```{r}
#| label: fig-frequencies
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- freq_tbl[freq_tbl$method != "manual", ]
  p <- ggplot(plot_tbl, aes(x = freq_bs_man, y = freq_bs, colour = method, shape = method)) +
    geom_abline(slope = 1, intercept = 0, colour = "grey50", linetype = "dashed") +
    geom_point(size = 2.5) +
    facet_wrap(~stim, scales = "free") +
    .analysis_theme() +
    labs(x = "Manual background-subtracted frequency (%)",
         y = "Method background-subtracted frequency (%)",
         colour = "Method", shape = "Method")
  .analysis_print_save_fig(p, file.path(.analysis_fig_dir(c(fig_key, "frequency")),
                                        "freq_bs-vs-manual.png"))
}
```

### Agreement with the manual labels

Per-cell agreement in the stimulated tube with the manual labels: false
discovery proportion (FDP), precision, sensitivity, specificity and F1.
Manual gates are an imperfect reference. The CSV retains TP, FP, FN, TN,
metric denominators and defined cell coverage. Undefined proportions are missing.

```{r}
#| label: classification
#| results: asis
if (isTRUE(run_plots) && exists("scored") && nrow(scored) > 0L) {
  cls <- scored[, c("method", "stim", "cyt", "n", "n_defined", "tp", "fp", "fn", "tn",
    "n_selected", "n_manual_positive", "n_manual_negative", "n_f1",
    "fdp", "precision", "sensitivity", "specificity", "f1")]
  .analysis_report_table(cls, c(fig_key, "classification.csv"),
    "Stimulated-cell agreement with the imperfect manual labels.")
}
```

The manual threshold rows (`method = "manual"`) differ from perfect agreement
only through the rectangles' other bounds.

Frequency correlations compare each method's background-subtracted frequency
with the imperfect manual reference across tubes, within stimulation and
cytokine and pooled over all tubes and cytokines. Frequencies retain the
existing clipping at zero. Pooled rows combine different cytokines and
stimulations and are descriptive. The CSV gives Pearson and Lin concordance
correlations and intended and finite pair counts. Fewer than three finite
pairs or zero variance gives missing correlations; a stratum with only one
tube therefore has no defined correlation.

**Summary by method.** Counts are summed over stimulations and cytokines
before the proportions are calculated; Pearson and concordance correlations
(CCC) are the pooled rows of the CSV. `n_defined` counts outcomes with a usable
gate, and `n_pairs` the finite frequency pairs.
The summary table follows the link to the full CSV.

```{r}
#| label: frequency-correlations
#| results: asis
if (isTRUE(run_plots) && exists("scored") && nrow(scored) > 0L) {
  report_scored <- scored[scored$method != "manual", ]
  correlations <- .omipFrequencyCorrelationTable(report_scored,
    c("stim", "cyt"), "freq_bs", "freq_bs_man")
  .analysis_report_table(correlations, c(fig_key, "frequency-correlations.csv"),
    "Frequency association and population-moment Lin concordance against the manual reference.")
  .omipAgreementKable(.omipAgreementOverview(report_scored, correlations, character()))
}
```

### Gates on the data

```{r}
#| label: fig-gates
#| results: asis
if (isTRUE(run_plots)) {
  channels <- names(omip016_prep$pre$responseChannels)
  ex_tbl <- do.call(rbind, lapply(seq_len(nrow(omip016_prep$pre$sampleMap)), function(ind) {
    ex <- .omip016Expr(omip016_prep$gs, ind, channels)
    do.call(rbind, lapply(channels, function(chnl) data.frame(
      tube = omip016_prep$pre$sampleMap$stim[[ind]],
      cyt = omip016_prep$pre$responseChannels[[chnl]],
      x = ex[, chnl]
    )))
  }))
  ex_tbl$cyt <- factor(ex_tbl$cyt, levels = levels(scored$cyt))
  for (st in levels(scored$stim)) {
    cat("\n\n#### ", toupper(st), " and unstimulated tubes\n\n", sep = "")
    g <- scored[scored$stim == st, ]
    p <- ggplot(ex_tbl[ex_tbl$tube %in% c("uns", st), ], aes(x = x)) +
      geom_density(aes(linetype = tube), adjust = 0.5) +
      geom_vline(data = g, aes(xintercept = gate, colour = method, group = method)) +
      facet_wrap(~cyt, scales = "free") +
      scale_y_sqrt() +
      .analysis_theme() +
      labs(x = "asinh(x / 150), compensated", y = "Density (square-root scale)",
           colour = "Gate", linetype = "Tube")
    .analysis_print_save_fig(p, file.path(.analysis_fig_dir(c(fig_key, "gates")),
                                          paste0("gates-", st, ".png")))
  }
}
```

### StimGate's gate for SEB MIP-1beta

```{r}
#| label: mip1b-debug
mip1b_chnl <- names(omip016_prep$pre$responseChannels)[omip016_prep$pre$responseChannels == "MIP1b"]
mip1b_ind <- omip016_prep$pre$sampleMap$ind[omip016_prep$pre$sampleMap$stim == "seb"]
mip1b_debug <- .omip016DebugGate(path_stimgate, mip1b_ind, mip1b_chnl, scored)
mip1b_info <- mip1b_debug$summary
mip1b_rows <- scored[scored$stim == "seb" & scored$cyt == "MIP1b", ]
mip1b_sg <- mip1b_rows[mip1b_rows$method == "stimgate", ]
knitr::kable(data.frame(
  quantity = c("Own local-FDR gate", "Own response estimate (%)", "Probability smoothing",
               "Final ordinary gate", "Final cytokine-positive gate", "Final gate reason",
               "Manual threshold", "Manual-positive cells missed"),
  value = c(signif(mip1b_info$ownGate, 3), signif(100 * mip1b_info$ownPropRespEst, 2),
            mip1b_info$smoothing, signif(mip1b_info$finalGate, 3),
            signif(mip1b_info$finalGateCyt, 3), mip1b_info$finalReason,
            signif(mip1b_info$manualGate, 3),
            paste0(mip1b_sg$fn, " of ", mip1b_sg$nManualStim))
), row.names = FALSE)
```

The SEB MIP-1beta page under "StimGate's local-FDR gating for every tube"
below shows StimGate's local-FDR gating of this tube: the stimulated and
unstimulated densities (panel A), the response probability (panel B) and the
expected responding cells (panel C). The response
probability rises once, at the lower edge of the MIP-1beta-positive bump, and
the gate sits there, next to the manual threshold.

Before stimgate 0.99.30 this gate sat above essentially the whole response.
The probability smoother could not fit channels whose names are not valid R
names, such as `PE-A`, so every OMIP-016 tube used the unsmoothed
probabilities. A handful of cells in the sparse upper tail then gave a spiky
probability curve that set the gate near the top of the data. The table above
reports the smoothing method used (`scam_mpi` is the monotone fit).

### StimGate's local-FDR gating for every tube

The recorded initial local-FDR gating of every stimulated tube and marker,
with StimGate's final gates, the F-beta and Tailgate gates and the manual
threshold.

```{r}
#| label: stimgate-debug-pages
#| results: asis
if (isTRUE(run_plots)) {
  .omip016DebugPages(path_stimgate, scored,
    .analysis_fig_dir(c(fig_key, "stimgate_debug")), heading = "####")
}
```

## Limitations

- One donor gives one unstimulated and two stimulated tubes, so StimGate's
  shared bandwidths and threshold clusters rest on two tubes per marker, and
  differences between methods cannot be generalised.
- The manual reference is the authors' gating re-applied from a decoded
  binary workspace without FlowJo-reported counts to check against. The CD4
  definition depends on the Boolean evaluation order (see above); the SEB
  reference changes materially under the alternative.
- Only CD4 T cells carry manual response gates in these samples; the
  workspace's CD8 tables refer to populations absent from the deposited
  samples.
- Manual gates were drawn once and applied to every tube; manual gating is an
  imperfect reference.
