---
title: OMIP-016 comparison with the shifted-peak rule
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_stimgate: null
  run_plots: true
---

## What this document is

This is the **shifted-peak rule on** version of Analysis 15
(`15-real-compare-omip016.qmd`). Analysis 15 stays the current-default version.
Everything is the same as Analysis 15 (prepared data, StimGate settings, seed,
comparators, manual reference and scoring) except one StimGate setting:
`stimControl(locShiftedPeakRef = TRUE)`.

StimGate normally starts looking for responding cells one third of a
left-window width above the higher of the stimulated and unstimulated main
peaks. With the rule on, if the stimulated main peak lies more than 2
bandwidths to the right of the unstimulated main peak, the search starts
above the unstimulated peak instead, using only the unstimulated tube's
left-window width. The bandwidth is the shared per-marker bandwidth estimated
on 10,000 cells, without the widening for smaller tubes.

**Smoothing fix.** stimgate 0.99.30 fixed the smoothing of the response
probability for channels whose names are not valid R names, such as `PE-A`.
Before that, smoothing failed for every OMIP-016 tube. 15b includes the
smoothing fix; compare it with Analysis 15 re-gated on the same fixed version.
The comparison below therefore isolates the shifted-peak rule.

StimGate results are written to their own cache (`omip016/shifted-peak`) and
figure folders. Analysis 15's prepared data, StimGate result and comparator
results are only read: Tailgate and F-beta do not depend on StimGate, so their
saved results are reused. Read Analysis 15 for the data, the manual reference
and the limits of the comparison; they all apply here.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("15b-real-compare-omip016-shifted-peak", sized = FALSE)
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-misc.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "sim-debug-loc.R"))
source(file.path(scripts_r_dir, "acs_cytof-debug.R"))
source(file.path(scripts_r_dir, "omip016-prepare.R"))
source(file.path(scripts_r_dir, "omip016-methods.R"))
source(file.path(scripts_r_dir, "omip-shifted-peak.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations", "RUN_SIMULATIONS", FALSE
))
run_stimgate <- .as_flag(.get_qmd_param_env(
  "run_stimgate", "RUN_STIMGATE", run_simulations
))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))

.acsCytofEnsureCurrentCheckout(root_dir)

# Setup starts in the QMD directory; keep configured paths absolute.
path_small <- projr::projr_path_get(
  "raw-data-small", "comparison_data", "omip016", format = "absolute"
)
path_prepared <- .analysis_project_dir("cache", c("omip016", "prepared"), root_dir, create = FALSE)
path_methods_default <- .analysis_project_dir("cache", c("omip016", "methods"), root_dir, create = FALSE)
path_methods <- .analysis_project_dir("cache", c("omip016", "shifted-peak", "methods"), root_dir, create = FALSE)
```

```{r}
#| label: scientific-settings
analysis_seed <- 20261009L
default_settings <- .omip016StimGateSettings(
  locThresholdMethod = "cap", clusterGates = TRUE, calcCytPosGates = TRUE
)
stimgate_settings <- .omip016StimGateSettings(
  locThresholdMethod = "cap", clusterGates = TRUE, calcCytPosGates = TRUE,
  locShiftedPeakRef = TRUE
)
```

```{r}
#| label: run-stimgate
#| include: false
prepared <- .omip016ReadPrepared(path_prepared)
path_stimgate <- file.path(path_methods, "stimgate")
if (isTRUE(run_stimgate)) {
  set.seed(analysis_seed)
  dir.create(path_methods, recursive = TRUE, showWarnings = FALSE)
  .acsCytofReplaceDir(path_stimgate, function(path_tmp) {
    .omip016RunStimGate(prepared, path_tmp, stimgate_settings)
  })
}
```

```{r}
#| label: collate
#| include: false
path_stimgate_default <- file.path(path_methods_default, "stimgate")
path_comparators <- file.path(path_methods_default, "comparators.rds")
if (!dir.exists(path_stimgate)) {
  stop("Shifted-peak StimGate result is missing; render with RUN_SIMULATIONS=true.")
}
if (!dir.exists(path_stimgate_default) || !file.exists(path_comparators)) {
  stop("Analysis 15 results are missing; render Analysis 15 first.")
}
current_pre <- prepared$pre[c("version", "inputContentHash", "settings")]
manifest <- readRDS(file.path(path_stimgate, "omip016-manifest.rds"))
manifest_default <- readRDS(file.path(path_stimgate_default, "omip016-manifest.rds"))
comparator_results <- readRDS(path_comparators)
if (!identical(manifest$preprocessing, current_pre) ||
    !identical(manifest_default$preprocessing, current_pre) ||
    !identical(comparator_results$preprocessing, current_pre)) {
  stop("OMIP-016 results were made from other prepared data.")
}
if (!identical(manifest$settings, stimgate_settings)) {
  stop("Saved shifted-peak StimGate result used other settings; re-run with RUN_STIMGATE=true.")
}
if (!identical(manifest_default$settings, default_settings)) {
  stop("Analysis 15's StimGate result used other settings.")
}
for (method in c("tailgate", "fbeta")) {
  if (!identical(comparator_results[[method]]$settings, .omip016ComparatorSettings(method))) {
    stop("Analysis 15's ", method, " result used other settings.")
  }
}
manual_thresholds <- utils::read.csv(file.path(path_prepared, "reference", "manual_thresholds.csv"),
                                     check.names = FALSE)
comparators <- rbind(comparator_results$tailgate$gates, comparator_results$fbeta$gates,
                     .omip016ManualGates(prepared, manual_thresholds))
scored <- .omip016Score(prepared, rbind(.omip016StimGateGates(prepared, path_stimgate), comparators))
.omip016CheckStimGateCounts(prepared, path_stimgate, scored)
scored_default <- .omip016Score(prepared, rbind(.omip016StimGateGates(prepared, path_stimgate_default), comparators))
fired <- .omipShiftedPeakFired(path_stimgate, labels = data.frame(project = "OMIP-016"))
sg <- scored[scored$method == "stimgate", ]
sg_default <- scored_default[scored_default$method == "stimgate", ]
sg_default <- sg_default[match(paste(sg$ind, sg$chnl), paste(sg_default$ind, sg_default$chnl)), ]
paired <- data.frame(
  method = "stimgate", stim = sg$stim, marker = sg$cyt, unit = sg$stim,
  threshold = sg$gate, thresholdDefault = sg_default$gate,
  gateCyt = sg$gateCyt, gateCytDefault = sg_default$gateCyt,
  freqBsManual = sg$freq_bs_man, freqBs = sg$freq_bs, freqBsDefault = sg_default$freq_bs,
  errorPp = sg$freq_bs - sg$freq_bs_man,
  errorPpDefault = sg_default$freq_bs - sg_default$freq_bs_man,
  shiftedPeakRef = fired$shiftedPeakRef[match(paste(sg$ind, sg$chnl), paste(fired$ind, fired$chnl))]
)
if (anyNA(paired$shiftedPeakRef)) stop("Missing shifted-peak flags.")
```

## How often the rule applied

The rule applied for **`r sum(paired$shiftedPeakRef)`** of
**`r nrow(paired)`** stimulated tube/marker outcomes. The flag is StimGate's
record for each tube's own local-FDR gate; threshold sharing and
cytokine-positive refinement can still move the final gate.

```{r}
#| label: fired
knitr::kable(paired[, c("stim", "marker", "shiftedPeakRef")], row.names = FALSE)
```

## StimGate with and without the rule

Gates on the GatingSet scale (`asinh(x / 150)`) and background-subtracted
frequencies (%, clipped at zero, as in Analysis 15). Errors are StimGate minus
the manual frequency, in percentage points. "Default" is Analysis 15's saved
StimGate result.

```{r}
#| label: error-table
show <- paired[, c("stim", "marker", "shiftedPeakRef", "thresholdDefault", "threshold",
                   "freqBsManual", "freqBsDefault", "freqBs", "errorPpDefault", "errorPp")]
knitr::kable(show, digits = 3, row.names = FALSE,
             col.names = c("Tube", "Marker", "Rule applied", "Gate (default)",
                           "Gate (shifted peak)", "Manual (%)", "Default (%)",
                           "Shifted peak (%)", "Error default (pp)", "Error shifted peak (pp)"))
if (isTRUE(run_plots)) {
  .analysis_report_table(paired, c(fig_key, "stimgate-default-vs-shifted-peak.csv"),
                         "StimGate gates, frequencies and errors with and without the rule.")
}
```

The rule changed StimGate's final gate for **`r sum(paired$threshold != paired$thresholdDefault)`**
of `r nrow(paired)` outcomes.

```{r}
#| label: error-figure
#| results: asis
#| fig-width: 10
#| fig-height: 5
if (isTRUE(run_plots)) {
  p <- .omipShiftedPeakErrorPlot(paired, facet = "stim")
  .analysis_print_save_fig(p, file.path(.analysis_fig_dir(c(fig_key, "stimgate-error")),
                                        "stimgate-error-default-vs-shifted-peak.png"))
}
```

## All methods with the rule on

Background-subtracted frequencies (%) with StimGate's shifted-peak gates; the
other methods are unchanged from Analysis 15.

```{r}
#| label: frequencies
freq_wide <- stats::reshape(
  scored[scored$method != "manual", c("stim", "cyt", "method", "freq_bs")],
  idvar = c("stim", "cyt"), timevar = "method", direction = "wide"
)
names(freq_wide) <- sub("^freq_bs\\.", "", names(freq_wide))
freq_wide$manual <- scored$freq_bs_man[match(paste(freq_wide$stim, freq_wide$cyt),
                                             paste(scored$stim, scored$cyt))]
knitr::kable(freq_wide, digits = 3, row.names = FALSE)
cls <- scored[scored$method != "manual", c("method", "stim", "cyt", "fdp", "sensitivity", "f1")]
if (isTRUE(run_plots)) {
  .analysis_report_table(cls, c(fig_key, "classification.csv"),
                         "Per-cell agreement with the manual labels, rule on.")
}
```

## Local-FDR gating where the rule applied

The recorded initial local-FDR gating of each outcome where the rule applied,
with StimGate's final gates and the manual threshold.

```{r}
#| label: fired-debug
#| results: asis
#| fig-width: 12
#| fig-height: 10
if (isTRUE(run_plots)) {
  fired_rows <- which(paired$shiftedPeakRef)
  if (length(fired_rows) == 0L) cat("The rule did not apply to any outcome.\n\n")
  for (i in fired_rows) {
    cat("\n\n### ", toupper(paired$stim[[i]]), " ", paired$marker[[i]], "\n\n", sep = "")
    dbg <- .omip016DebugGate(path_stimgate, sg$ind[[i]], sg$chnl[[i]], scored)
    .analysis_print_save_fig(
      dbg$plot,
      file.path(.analysis_fig_dir(c(fig_key, "stimgate_debug")),
                paste0(paired$stim[[i]], "-", paired$marker[[i]], "-local_fdr.png")),
      width = 30, height = 25
    )
  }
}
```

## Limitations

With one donor and two stimulated tubes, these are case-level comparisons;
see Analysis 15's limitations.
