---
title: Cytokine distributions among cells positive for another cytokine (ACS CyTOF)
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: false
  run_plots: true
  pops: all
  stims: all
---

## What this document is

StimGate's cytokine-positive gates lower a cytokine's gate for cells that are
already positive for another cytokine, because a much larger share of those
cells respond. Choosing a better rule for that lower gate needs a view of how
each cytokine is actually distributed among those cells, and how that differs
from all cells.

This document shows, for every ACS CyTOF donor with manual gating (12 donors),
each population and each stimulation:

1. hexagon plots of IFNγ against TNF in the stimulated and unstimulated tubes,
   with StimGate's ordinary gates (orange lines);
2. TNF among IFNγ+ cells and IFNγ among TNF+ cells (the partner in the
   hexagon plots), and each among cells positive for any other cytokine
   (what StimGate's cytokine-positive rule uses), in the stimulated (red)
   and unstimulated (blue) tubes. The conditional cells are drawn as
   histograms (bins 0.25 wide, scaled to density so tubes of different sizes
   compare), since tens of cells are too few for a reliable kernel density;
   all cells are drawn as dotted density curves for reference.

"Positive for another cytokine" uses StimGate's ordinary gates for the other
five cytokines (IL-2, IL-6, IL-17, IL-22 and TNF or IFNγ), counting a cell
when its value is strictly above the gate, as StimGate does. In the density
plots the dashed dark line is StimGate's ordinary gate and the solid orange
line its cytokine-positive gate (drawn on top when the two are equal). Exact
zeros are left out of the densities, as StimGate leaves out each tube's
minimum; each panel's label gives the number of stimulated cells in the
conditional subset (before zeros are dropped). Histograms and curves are
each scaled to their own cells, so they compare shapes, not cell numbers.
The all-cell curves use their own Sheather-Jones bandwidth (at least 0.05).
The y axis is on a square-root scale so the tails remain visible next to the
large peak near zero.

Everything is read from Analysis 9's most recent run (its cached GatingSets
and StimGate gates); nothing is gated again. The gates are therefore those of
the version Analysis 9 last ran with.

## Set-up

Parameters (or environment variables `ACS_EXPLORE_POPS`, `ACS_EXPLORE_STIMS`)
choose the populations and stimulations; lists are comma separated and `all`
selects everything.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("16-explore-acs-cytof-cytpos", sized = FALSE)
for (fn in c(
  "acs_cytof-helper.R", "acs_cytof-gate.R", "acs_cytof-methods.R",
  "acs_cytof-debug.R", "acs_cytof-explore-cytpos.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
sel_pops <- .acsCytofDebugList(.get_qmd_param_env("pops", "ACS_EXPLORE_POPS", "all"))
sel_stims <- .acsCytofDebugList(.get_qmd_param_env("stims", "ACS_EXPLORE_STIMS", "all"))
```

```{r}
#| label: paths
#| include: false
# Analysis 9's caches are only read: never create or change them here.
path_gs_base <- projr::projr_path_get(
  "cache", "acs_cytof", "gs",
  create = FALSE, format = "absolute"
)
path_scratch_base <- projr::projr_path_get(
  "cache", "acs_cytof", "scratch",
  create = FALSE, format = "absolute"
)
path_comparison <- projr::projr_path_get(
  "cache", "acs_cytof", "manual-comparison", "manual-comparison.rds",
  create = FALSE, format = "absolute"
)
if (!file.exists(path_comparison)) {
  stop("Analysis 9's manual comparison is missing; render Analysis 9 first.")
}
comparison <- readRDS(path_comparison)
manual_donors <- sort(unique(as.character(comparison$SampleID)))
settings <- .acsCytofDebugSettings(
  file.path(root_dir, "analysis", "9-real-compare-acs-cytof.qmd")
)
pops <- intersect(
  settings$popVec,
  intersect(sel_pops %||% settings$popVec, unique(as.character(comparison$popCode)))
)
channel_map <- .acsCytofChannelMap()
```

Donors with manual gating: `r paste(manual_donors, collapse = ", ")`.
Populations shown: `r paste(pops, collapse = ", ")`.

```{r}
#| label: plots
#| results: asis
#| fig-width: 14
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (pop in pops) {
    paths <- .acsCytofPopulationPaths(pop, "unused", path_gs_base, path_scratch_base)
    gs <- flowWorkspace::load_gs(paths$gs, backend_readonly = TRUE)
    sample_map <- .acsCytofReadPreprocessing(paths$gs, gs)$sampleMap
    gates <- stimgate::getStimGates(paths$stimgate)
    stims <- intersect(sel_stims %||% sort(setdiff(unique(sample_map$stim), "uns")),
      setdiff(unique(sample_map$stim), "uns"))
    cat("\n\n## ", pop, "\n\n", sep = "")
    for (stim in stims) {
      cat("\n\n### ", pop, ": ", stim, "\n\n", sep = "")
      donors <- intersect(manual_donors, sample_map$SampleID[sample_map$stim == stim])
      dat <- stats::setNames(lapply(donors, function(donor) {
        .acsCytposTubeData(gs, sample_map, gates, donor, stim, channel_map)
      }), donors)
      donor_gates <- lapply(dat, function(d) d[c("gate", "gateCyt")])
      fig_dir <- .analysis_fig_dir(c(fig_key, pop), root_dir)

      cat("\n\n#### IFNγ against TNF\n\n")
      hex_tbl <- dplyr::bind_rows(lapply(donors, function(d) .acsCytposHexTbl(dat[[d]], d)))
      .analysis_print_save_fig(
        .acsCytposHexPlot(hex_tbl, donor_gates),
        file.path(fig_dir, paste0(stim, "-hex-IFNg-TNF.png")),
        width = 36, height = 26
      )
      views <- list(
        list(cyt = "TNF", given = "IFNg", heading = "TNF among IFNγ+ cells"),
        list(cyt = "TNF", given = NULL, heading = "TNF among cells positive for another cytokine"),
        list(cyt = "IFNg", given = "TNF", heading = "IFNγ among TNF+ cells"),
        list(cyt = "IFNg", given = NULL, heading = "IFNγ among cells positive for another cytokine")
      )
      for (view in views) {
        cat("\n\n#### ", view$heading, "\n\n", sep = "")
        dens_tbl <- dplyr::bind_rows(lapply(donors, function(d) {
          .acsCytposDensityTbl(dat[[d]], d, view$cyt, given = view$given)
        }))
        given_key <- if (is.null(view$given)) "any_other" else paste0(view$given, "pos")
        .analysis_print_save_fig(
          .acsCytposDensityPlot(dens_tbl, donor_gates, view$cyt),
          file.path(fig_dir, paste0(stim, "-", view$cyt, "-given-", given_key, ".png")),
          width = 30, height = 24
        )
      }
    }
  }
}
```

```{r}
```
