---
title: Cytokine distributions among cells positive for another cytokine (ACS CyTOF)
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: false
  run_plots: true
  pops: all
  stims: all
---

## What this document is

StimGate's cytokine-positive gates lower a cytokine's gate for cells that are
already positive for another cytokine, because a much larger share of those
cells respond. Choosing a better rule for that lower gate needs a view of how
each cytokine is actually distributed among those cells, and how that differs
from all cells.

This document shows, for every ACS CyTOF donor with manual gating (12 donors),
each population and each stimulation:

1. hexagon plots of IFNγ against TNF in the stimulated and unstimulated tubes,
   with StimGate's ordinary gates (orange lines);
2. the IFNγ distribution among all cells (dotted) and among cells positive
   for at least one other cytokine (solid), in the stimulated (red) and
   unstimulated (blue) tubes;
3. the same for TNF.

"Positive for another cytokine" uses StimGate's ordinary gates for the other
five cytokines (IL-2, IL-6, IL-17, IL-22 and TNF or IFNγ), counting a cell
when its value is strictly above the gate, as StimGate does. In the density
plots the dashed dark line is StimGate's ordinary gate and the solid orange
line its cytokine-positive gate (drawn on top when the two are equal). Exact
zeros are left out of the densities, as StimGate leaves out each tube's
minimum; each panel's label gives the number of stimulated cells positive for
another cytokine. Each density is scaled to its own cells, so the curves
compare shapes, not cell numbers. Each curve uses its own Sheather-Jones
bandwidth (at least 0.05), which follows two modes in a small subset without
being as narrow as a bandwidth chosen for tens of thousands of cells. Curves
are drawn only for groups with at least five non-zero values. The y axis
is on a square-root scale so the tails remain visible next to the large peak
near zero.

Everything is read from Analysis 9's most recent run (its cached GatingSets
and StimGate gates); nothing is gated again. The gates are therefore those of
the version Analysis 9 last ran with.

## Set-up

Parameters (or environment variables `ACS_EXPLORE_POPS`, `ACS_EXPLORE_STIMS`)
choose the populations and stimulations; lists are comma separated and `all`
selects everything.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("16-explore-acs-cytof-cytpos", sized = FALSE)
for (fn in c(
  "acs_cytof-helper.R", "acs_cytof-gate.R", "acs_cytof-methods.R",
  "acs_cytof-debug.R", "acs_cytof-explore-cytpos.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
sel_pops <- .acsCytofDebugList(.get_qmd_param_env("pops", "ACS_EXPLORE_POPS", "all"))
sel_stims <- .acsCytofDebugList(.get_qmd_param_env("stims", "ACS_EXPLORE_STIMS", "all"))
```

```{r}
#| label: paths
#| include: false
# Analysis 9's caches are only read: never create or change them here.
path_gs_base <- projr::projr_path_get(
  "cache", "acs_cytof", "gs",
  create = FALSE, format = "absolute"
)
path_scratch_base <- projr::projr_path_get(
  "cache", "acs_cytof", "scratch",
  create = FALSE, format = "absolute"
)
path_comparison <- projr::projr_path_get(
  "cache", "acs_cytof", "manual-comparison", "manual-comparison.rds",
  create = FALSE, format = "absolute"
)
if (!file.exists(path_comparison)) {
  stop("Analysis 9's manual comparison is missing; render Analysis 9 first.")
}
comparison <- readRDS(path_comparison)
manual_donors <- sort(unique(as.character(comparison$SampleID)))
settings <- .acsCytofDebugSettings(
  file.path(root_dir, "analysis", "9-real-compare-acs-cytof.qmd")
)
pops <- intersect(
  settings$popVec,
  intersect(sel_pops %||% settings$popVec, unique(as.character(comparison$popCode)))
)
channel_map <- .acsCytofChannelMap()
```

Donors with manual gating: `r paste(manual_donors, collapse = ", ")`.
Populations shown: `r paste(pops, collapse = ", ")`.

```{r}
#| label: plots
#| results: asis
#| fig-width: 14
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (pop in pops) {
    paths <- .acsCytofPopulationPaths(pop, "unused", path_gs_base, path_scratch_base)
    gs <- flowWorkspace::load_gs(paths$gs, backend_readonly = TRUE)
    sample_map <- .acsCytofReadPreprocessing(paths$gs, gs)$sampleMap
    gates <- stimgate::getStimGates(paths$stimgate)
    stims <- intersect(sel_stims %||% sort(setdiff(unique(sample_map$stim), "uns")),
      setdiff(unique(sample_map$stim), "uns"))
    cat("\n\n## ", pop, "\n\n", sep = "")
    for (stim in stims) {
      cat("\n\n### ", pop, ": ", stim, "\n\n", sep = "")
      donors <- intersect(manual_donors, sample_map$SampleID[sample_map$stim == stim])
      dat <- stats::setNames(lapply(donors, function(donor) {
        .acsCytposTubeData(gs, sample_map, gates, donor, stim, channel_map)
      }), donors)
      donor_gates <- lapply(dat, function(d) d[c("gate", "gateCyt")])
      fig_dir <- .analysis_fig_dir(c(fig_key, pop), root_dir)

      cat("\n\n#### IFNγ against TNF\n\n")
      hex_tbl <- dplyr::bind_rows(lapply(donors, function(d) .acsCytposHexTbl(dat[[d]], d)))
      .analysis_print_save_fig(
        .acsCytposHexPlot(hex_tbl, donor_gates),
        file.path(fig_dir, paste0(stim, "-hex-IFNg-TNF.png")),
        width = 36, height = 26
      )
      for (cyt in c("IFNg", "TNF")) {
        cat("\n\n#### ", cyt, " among cells positive for another cytokine\n\n", sep = "")
        dens_tbl <- dplyr::bind_rows(lapply(donors, function(d) {
          .acsCytposDensityTbl(dat[[d]], d, cyt)
        }))
        .analysis_print_save_fig(
          .acsCytposDensityPlot(dens_tbl, donor_gates, cyt),
          file.path(fig_dir, paste0(stim, "-density-", cyt, ".png")),
          width = 30, height = 24
        )
      }
    }
  }
}
```

```{r}
```
