---
title: A co-expression score for ACS CyTOF cells
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: false
  run_plots: true
---

## What this document is

Analysis 16 showed that a cytokine's gate set on all cells can sit too high
for cells that also make another cytokine (TNF among IFNγ+ CD4 cells), and that
in NK cells the unstimulated tube co-expresses too. This document explores a
co-expression score that judges several markers together.

For a set of markers A, each cell gets

  score = −log10 ∏ S_j(x_j), with S_j(x) = P(control cell ≥ x on marker j).

Each S_j is a tail p-value against the cell's own control (unstimulated)
tube: (control cells ≥ x + 1) / (control cells + 1) for stimulated cells, and
control cells ≥ x / control cells for the control cells themselves. Exact zeros
get S_j = 1. A high score means a cell is jointly unusual for every marker in
A, even when it clears no single gate. The product would be the probability of
such a cell at baseline only if the markers were independent there; they need
not be (NK cells), so the score is not read as a probability. Instead the
control tube's own scores are the reference: where the control tube's score
distribution runs out is where co-expression becomes very unlikely at baseline.

The populations, stimulations and markers are:

- CD4 T cells, p1 and mtb;
- NK cells, ebv and p1 (their strongest manual IFNγ and TNF responses);
- each pair of IFNγ, TNF and IL-2.

For each, per manually gated donor (12):

1. the score histogram of both tubes (fraction of each tube's cells per bin,
   log scale; cells scoring 0 are omitted from the plot but counted);
2. the fraction of cells at or above each score in both tubes (log scale);
3. the net (stimulated minus control) frequency at or above each score, with
   the net frequency of cells above both ordinary gates (orange dashed line);
4. hexagon plots with the score's contours at 2, 3 and 4 (products of 10⁻²,
   10⁻³, 10⁻⁴; white) and StimGate's ordinary gates (orange).

For IFNγ and TNF, a second view asks where the two markers co-occur more than
their marginals explain. In each tube, on bins 0.2 wide, two Poisson GAMs of
the bin counts are fitted: an independence model, log rate = s(IFNγ) +
s(TNF) (the product of non-linear marginals), and an interaction model that
adds ti(IFNγ, TNF). Their difference is the co-expression beyond
independence. Only bins where cells are positive for at least one marker by
its ordinary gate are considered, so the question is which single-positive
cells also carry the other marker. Per donor this section shows:

5. the Pearson residuals (O − E) / √E of each tube against independence
   (E from that tube's own marginals; red = more cells than independence
   predicts), as a raw view;
6. log2(interaction / independence) in the single-positive region of both
   tubes, with the ordinary gates;
7. the difference between the tubes: stimulated minus control excess over
   independence per bin (percentage points), the co-expression the
   stimulation adds;
8. a table with the net (stimulated minus control) excess over independence
   in the single-positive region and the cells called co-expressing
   (single-positive cells in bins where the stimulated tube has at least twice
   the independence expectation and more excess than the control tube);
9. lowered cytokine-positive gates (next section) on the hexagon plots, with
   histograms of each marker among cells positive for the other and a table of
   the net double-positive frequencies.

## Lowering a gate for cells positive for the other cytokine

The rectangle of the two ordinary gates misses co-expressing responders that
sit below one gate, such as the diagonal IFNγ/TNF band in CD4 cells. Taking
TNF among IFNγ+ cells (cells above IFNγ's ordinary gate):

1. **Is there a double-positive response?** Nothing moves unless the cells
   above both ordinary gates show one: z = (s − u·nS/nU) / √(s + u·(nS/nU)²) > 2,
   with s and u the stimulated and control double positives and nS, nU the tube
   sizes.
2. **Candidate bins.** The distance from TNF's ordinary gate down to its floor
   is cut into 20 bins. The floor is the control tube's main negative TNF peak
   plus 1.5 robust SDs (the peak's left half-width at half maximum / 1.177),
   from the non-zero values only: exact zeros are a spike, not part of the
   negative peak's shape.
3. **When a set of bins passes.** Its IFNγ+ stimulated cells must
   - depart from independence: (observed − expected) / √expected ≥ 3.5, with
     expected = stimulated tube size × IFNγ+ share × share of cells in the
     bins (both shares over all the tube's cells, zeros included); and
   - be pure enough: purity = 1 − (control share) / (stimulated share), the
     estimated fraction of responders, at least 0.75 times the double
     positives' purity. Shares are over each tube's cells, zeros included.
4. **The lowered gate.** Working back from the floor, the gate is the first bin
   edge where that bin passes on its own and all the bins from the ordinary
   gate down to it pass together. Bins that fail can lie in between, as long
   as the whole set still passes.
5. **Trim of the added band.** IFNγ+ cells only just above IFNγ's gate may be
   background. The distance from IFNγ's gate down to its floor, in 20 steps,
   sets slices above IFNγ's gate. Working up from the gate, IFNγ's cut for the
   added band is the first slice edge where that slice passes on purity alone
   and the band above it does too. If the band runs out, TNF's gate is not
   lowered.

A cell is then TNF+ if above TNF's ordinary gate, or above the raised IFNγ cut
and the lowered TNF gate. The same is done for IFNγ among TNF+ cells. Cells
called IFNγ+TNF+ are those above both ordinary gates or added by either lowered
gate, in both tubes; the net frequency subtracts the control tube's.

On the hexagon plots the ordinary gates are orange and the lowered gates blue,
each drawn from the raised cut on the other marker. The histograms show each
marker among cells positive for the other, as cells per 100,000 cells of each
tube, so the bars compare directly: where the control bar is a quarter of the
stimulated bar, purity there is 0.75. Exact zeros are left out of the bars. The
lines are the ordinary gate (black dashed), the lowered gate (blue) and the
floor (grey dotted). CSV tables (paths printed under each heading) give each
donor's z, floors, gates and cuts, and the net IFNγ+TNF+ frequencies and
control cells called, with the lowered gates and with the rectangle.

A CSV gives each donor's KL divergence of the stimulated from the control
score distribution (bits; bins 0.25 wide, add-half smoothing), a summary of
how differently the two tubes score, not a gate.

Everything is read from Analysis 9's most recent run; nothing is gated again.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("17-explore-acs-cytof-coexpression", sized = FALSE)
for (fn in c(
  "acs_cytof-helper.R", "acs_cytof-gate.R", "acs_cytof-methods.R",
  "acs_cytof-debug.R", "acs_cytof-explore-cytpos.R", "acs_cytof-explore-score.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
```

```{r}
#| label: paths
#| include: false
# Analysis 9's caches are only read: never create or change them here.
path_gs_base <- projr::projr_path_get(
  "cache", "acs_cytof", "gs",
  create = FALSE, format = "absolute"
)
path_scratch_base <- projr::projr_path_get(
  "cache", "acs_cytof", "scratch",
  create = FALSE, format = "absolute"
)
path_comparison <- projr::projr_path_get(
  "cache", "acs_cytof", "manual-comparison", "manual-comparison.rds",
  create = FALSE, format = "absolute"
)
if (!file.exists(path_comparison)) {
  stop("Analysis 9's manual comparison is missing; render Analysis 9 first.")
}
manual_donors <- sort(unique(as.character(readRDS(path_comparison)$SampleID)))
channel_map <- .acsCytofChannelMap()
selection <- list(cd4 = c("p1", "mtb"), nk = c("ebv", "p1"))
pairs <- list(c("IFNg", "TNF"), c("IFNg", "IL2"), c("TNF", "IL2"))
```

```{r}
#| label: plots
#| results: asis
#| fig-width: 14
#| fig-height: 9
if (isTRUE(run_plots)) {
  kl_all <- list()
  mono_all <- list()
  low_all <- list()
  for (pop in names(selection)) {
    paths <- .acsCytofPopulationPaths(pop, "unused", path_gs_base, path_scratch_base)
    gs <- flowWorkspace::load_gs(paths$gs, backend_readonly = TRUE)
    sample_map <- .acsCytofReadPreprocessing(paths$gs, gs)$sampleMap
    gates <- stimgate::getStimGates(paths$stimgate)
    fig_dir <- .analysis_fig_dir(c(fig_key, pop), root_dir)
    cat("\n\n## ", pop, "\n\n", sep = "")
    for (stim in selection[[pop]]) {
      donors <- intersect(manual_donors, sample_map$SampleID[sample_map$stim == stim])
      dat <- stats::setNames(lapply(donors, function(donor) {
        .acsCytposTubeData(gs, sample_map, gates, donor, stim, channel_map)
      }), donors)
      donor_gates <- lapply(dat, `[[`, "gate")
      for (pair in pairs) {
        pair_key <- paste(pair, collapse = "-")
        cat("\n\n### ", pop, ", ", stim, ": ", pair[[1]], " and ", pair[[2]], "\n\n", sep = "")
        score_tbl <- do.call(rbind, lapply(donors, function(d) .acsScoreTbl(dat[[d]], pair, d)))
        tail <- .acsScoreTail(score_tbl)
        rect <- do.call(rbind, lapply(donors, function(d) .acsScoreRectNet(dat[[d]], pair, d)))
        kl_all[[length(kl_all) + 1L]] <- cbind(
          pop = pop, stim = stim, markers = pair_key, .acsScoreKl(score_tbl)
        )

        cat("\n\n#### Score histograms\n\n")
        .analysis_print_save_fig(.acsScoreHistPlot(score_tbl),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-hist.png")), width = 30, height = 22)
        cat("\n\n#### Cells at or above each score\n\n")
        .analysis_print_save_fig(.acsScoreTailPlot(tail),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-tail.png")), width = 30, height = 22)
        cat("\n\n#### Net frequency at or above each score\n\n")
        .analysis_print_save_fig(.acsScoreNetPlot(tail, rect),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-net.png")), width = 30, height = 22)
        cat("\n\n#### ", pair[[1]], " against ", pair[[2]], " with score contours\n\n", sep = "")
        hex_tbl <- do.call(rbind, lapply(donors, function(d) {
          .acsCytposHexTbl(dat[[d]], d, x = pair[[1]], y = pair[[2]])
        }))
        grid <- do.call(rbind, lapply(donors, function(d) {
          .acsScoreGrid(dat[[d]], pair[[1]], pair[[2]], d)
        }))
        .analysis_print_save_fig(
          .acsScoreHexPlot(hex_tbl, grid, donor_gates, pair[[1]], pair[[2]]),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-hex.png")), width = 36, height = 26
        )
        if (identical(pair_key, "IFNg-TNF")) {
          cat("\n\n#### Departure from independence\n\n")
          res <- do.call(rbind, lapply(donors, function(d) {
            .acsCoexResiduals(.acsCoexBins(dat[[d]], pair[[1]], pair[[2]], d, width = 0.25))
          }))
          .analysis_print_save_fig(.acsCoexResidualPlot(res, donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-residuals.png")), width = 36, height = 26)
          indep <- lapply(donors, function(d) .acsCoexIndependence(dat[[d]], pair[[1]], pair[[2]], d))
          cat("\n\n#### Co-expression beyond independence\n\n")
          .analysis_print_save_fig(
            .acsCoexIndependencePlot(do.call(rbind, indep), donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-interaction.png")), width = 36, height = 26
          )
          cat("\n\n#### Stimulated minus control departure from independence\n\n")
          diff_tbl <- do.call(rbind, lapply(indep, .acsCoexDifference))
          .analysis_print_save_fig(
            .acsCoexDifferencePlot(diff_tbl, donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-interaction-difference.png")), width = 30, height = 22
          )
          mono_tbl <- do.call(rbind, lapply(seq_along(donors), function(i) {
            .acsCoexIndependenceSummary(indep[[i]], dat[[donors[[i]]]], pair[[1]], pair[[2]])
          }))
          is_num <- vapply(mono_tbl, is.numeric, logical(1))
          mono_tbl[is_num] <- lapply(mono_tbl[is_num], round, 3)
          mono_all[[length(mono_all) + 1L]] <- cbind(pop = pop, stim = stim, mono_tbl)
          cat("\n\n#### Co-expression beyond independence against the rectangle\n\n")
          .analysis_report_table(mono_tbl,
            c(fig_key, paste0(pop, "-", stim, "-coexpression-beyond-independence.csv")),
            "Net excess over independence and cells called, against the rectangle.", root_dir)

          lower <- stats::setNames(lapply(donors, function(d) {
            .acsCoexLowerSummary(dat[[d]], pair[[1]], pair[[2]], d)
          }), donors)
          cat("\n\n#### Lowered gates\n\n")
          .analysis_print_save_fig(
            .acsCoexLowerHexPlot(hex_tbl, lower, donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-lowered-hex.png")), width = 36, height = 26
          )
          for (cond_pair in list(pair, rev(pair))) {
            cat("\n\n#### ", cond_pair[[2]], " among ", cond_pair[[1]], "+ cells\n\n", sep = "")
            .analysis_print_save_fig(
              .acsCoexCondHistPlot(dat, lower, a = cond_pair[[1]], b = cond_pair[[2]]),
              file.path(fig_dir, paste0(stim, "-", cond_pair[[2]], "-given-", cond_pair[[1]], "-hist.png")),
              width = 30, height = 22
            )
          }
          low_tbl <- do.call(rbind, lapply(lower, `[[`, "summary"))
          low_show <- low_tbl
          is_num <- vapply(low_show, is.numeric, logical(1))
          low_show[is_num] <- lapply(low_show[is_num], round, 3)
          low_all[[length(low_all) + 1L]] <- cbind(pop = pop, stim = stim, low_tbl)
          .analysis_report_table(low_show,
            c(fig_key, paste0(pop, "-", stim, "-lowered-gates.csv")),
            "Lowered gates and net double-positive frequencies per donor.", root_dir)
        }
      }
    }
  }
  .analysis_report_table(do.call(rbind, mono_all),
    c(fig_key, "coexpression-beyond-independence.csv"),
    "IFNg-TNF co-expression beyond independence against the rectangle of ordinary gates.",
    root_dir, quiet = TRUE)
  low_all <- do.call(rbind, low_all)
  .analysis_report_table(low_all, c(fig_key, "lowered-gates.csv"),
    "Lowered IFNg and TNF gates and net double-positive frequencies per donor.", root_dir,
    quiet = TRUE)
  cat("\n\n## Lowered gates: summary\n\n")
  cat("Per population and stimulation (12 donors): donors with a clear double-positive",
    "response (z > 2), donors whose TNF gate (for IFNγ+ cells) and IFNγ gate (for TNF+",
    "cells) were lowered, and the median net IFNγ+TNF+ frequency (%) and control cells",
    "called, with the lowered gates and with the rectangle of ordinary gates.\n\n")
  low_summary <- do.call(rbind, lapply(split(low_all, list(low_all$pop, low_all$stim), drop = TRUE), function(d) {
    data.frame(pop = d$pop[[1]], stim = d$stim[[1]],
      clearResponse = sum(d$zDp > 2),
      tnfLowered = sum(d$cutY < d$gateY - 1e-9),
      ifngLowered = sum(d$cutX < d$gateX - 1e-9),
      medianNetLowered = round(stats::median(d$netPct), 3),
      medianNetRectangle = round(stats::median(d$netRectanglePct), 3),
      medianControlLowered = stats::median(d$controlCalled),
      medianControlRectangle = stats::median(d$controlRectangle))
  }))
  print(knitr::kable(low_summary, row.names = FALSE))
  kl_tbl <- do.call(rbind, kl_all)
  kl_tbl$klBits <- round(kl_tbl$klBits, 3)
  cat("\n\n## KL divergence of stimulated from control scores\n\n")
  .analysis_report_table(kl_tbl, c(fig_key, "kl-divergence.csv"),
    "KL divergence (bits) of each donor's stimulated from control score distribution.",
    root_dir)

}
```

```{r}
```
