---
title: A co-expression score for ACS CyTOF cells
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: false
  run_plots: true
---

## What this document is

Analysis 16 showed that a cytokine's gate set on all cells can sit too high
for cells that also make another cytokine (TNF among IFNγ+ CD4 cells), and that
in NK cells the unstimulated tube co-expresses too. This document explores a
co-expression score that judges several markers together.

For a set of markers A, each cell gets

  score = −log10 ∏ S_j(x_j), with S_j(x) = P(control cell ≥ x on marker j).

Each S_j is a tail p-value against the cell's own control (unstimulated)
tube: (control cells ≥ x + 1) / (control cells + 1) for stimulated cells, and
control cells ≥ x / control cells for the control cells themselves. Exact zeros
get S_j = 1. A high score means a cell is jointly unusual for every marker in
A, even when it clears no single gate. The product would be the probability of
such a cell at baseline only if the markers were independent there; they need
not be (NK cells), so the score is not read as a probability. Instead the
control tube's own scores are the reference: where the control tube's score
distribution runs out is where co-expression becomes very unlikely at baseline.

The populations, stimulations and markers are:

- CD4 T cells, p1 and mtb;
- NK cells, ebv and p1 (their strongest manual IFNγ and TNF responses);
- each pair of IFNγ, TNF and IL-2.

For each, per manually gated donor (12):

1. the score histogram of both tubes (fraction of each tube's cells per bin,
   log scale; cells scoring 0 are omitted from the plot but counted);
2. the fraction of cells at or above each score in both tubes (log scale);
3. the net (stimulated minus control) frequency at or above each score, with
   the net frequency of cells above both ordinary gates (orange dashed line);
4. hexagon plots with the score's contours at 2, 3 and 4 (products of 10⁻²,
   10⁻³, 10⁻⁴; white) and StimGate's ordinary gates (orange).

For IFNγ and TNF, a second view asks where the two markers co-occur more than
their marginals explain. In each tube, on bins 0.2 wide, two Poisson GAMs of
the bin counts are fitted: an independence model, log rate = s(IFNγ) +
s(TNF) (the product of non-linear marginals), and an interaction model that
adds ti(IFNγ, TNF). Their difference is the co-expression beyond
independence. Only bins where cells are positive for at least one marker by
its ordinary gate are considered, so the question is which single-positive
cells also carry the other marker. Per donor this section shows:

5. the Pearson residuals (O − E) / √E of each tube against independence
   (E from that tube's own marginals; red = more cells than independence
   predicts), as a raw view;
6. log2(interaction / independence) in the single-positive region of both
   tubes, with the ordinary gates;
7. the difference between the tubes: stimulated minus control excess over
   independence per bin (percentage points), the co-expression the
   stimulation adds;
8. a table with the net (stimulated minus control) excess over independence
   in the single-positive region and the cells called co-expressing
   (single-positive cells in bins where the stimulated tube has at least twice
   the independence expectation and more excess than the control tube);
9. the co-expression region grown from the rectangle (next section) and the
   F-beta cytokine-positive gates, drawn on the hexagon plots, with tables of
   the net double-positive frequencies.

## Growing the co-expression region

The rectangle of the two ordinary gates misses co-expressing responders that
sit below one gate, such as the diagonal IFNγ/TNF band in CD4 cells. The
region is grown from the rectangle, bin by bin, wherever the data show
co-expression the stimulation adds, with no assumed direction:

1. The two markers are binned in squares 0.2 wide. Bins above both ordinary
   gates (the rectangle) start the region.
2. Candidate bins lie above both markers' floors (each marker's main negative
   peak in the control tube plus 1.5 robust SDs, the peak's left half-width at
   half maximum / 1.177) and above at least one ordinary gate, so the region
   never grows into the bulk of the negatives.
3. A candidate bin is acceptable when
   - (departure from independence) the stimulated tube holds clearly more
     cells there than independence of the two markers predicts:
     (observed − expected) / √expected ≥ 2, with expected from the stimulated
     tube's own marginals; and
   - (low background) the stimulated tube's share of cells there is at least
     twice the control tube's ((control count + 0.5) / control tube size), and
     the bin holds at most two control cells.
4. An acceptable bin joins the region only if it can be reached from the
   rectangle through neighbouring bins (including diagonal neighbours)
   crossing at most two rejected bins. Empty bins and acceptable bins cost
   nothing to cross, so a sparse band of responders stays connected, but an
   island cut off by bins that fail the checks is not added.
5. Cells in the rectangle and in the added bins are called IFNγ+TNF+, in both
   tubes; the net frequency subtracts the control tube's.

For comparison, the F-beta cytokine-positive gates lower one marker's gate for
cells positive for the other. Taking TNF among IFNγ+ cells: every TNF
threshold from TNF's ordinary gate down to its floor (steps of 0.05) is
evaluated. At each, the co-expression the stimulation adds is estimated as the
stimulated tube's IFNγ+ cells above it in excess of independence, minus the
same excess in the control tube (scaled to the stimulated tube's size); TNF
responders whose IFNγ is independent of their TNF add nothing. With that
estimate as the true positives, precision is its share of the stimulated IFNγ+
cells above the threshold and recall its share of the largest estimate. The
threshold with the largest F1 is chosen (the highest on ties) and kept only if
the estimate there is clear (z > 2); otherwise TNF's ordinary gate is kept.
The same is done for IFNγ among TNF+ cells. A cell is then TNF+ if above TNF's
ordinary gate, or IFNγ+ and above the lowered gate.

On the hexagon plots the added bins are outlined in black and the F-beta
gates are grey dashed lines, drawn only where they apply. CSV tables (paths printed under each heading) give the
floors, bins added and cells called by region growing, the F-beta gates with
their z, and both net IFNγ+TNF+ frequencies against the rectangle. A donor
without a response can still gain an isolated bin of a few cells (two
stimulated, one control for CD4 p1 030697_D0).

A CSV gives each donor's KL divergence of the stimulated from the control
score distribution (bits; bins 0.25 wide, add-half smoothing), a summary of
how differently the two tubes score, not a gate.

Everything is read from Analysis 9's most recent run; nothing is gated again.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("17-explore-acs-cytof-coexpression", sized = FALSE)
for (fn in c(
  "acs_cytof-helper.R", "acs_cytof-gate.R", "acs_cytof-methods.R",
  "acs_cytof-debug.R", "acs_cytof-explore-cytpos.R", "acs_cytof-explore-score.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
```

```{r}
#| label: paths
#| include: false
# Analysis 9's caches are only read: never create or change them here.
path_gs_base <- projr::projr_path_get(
  "cache", "acs_cytof", "gs",
  create = FALSE, format = "absolute"
)
path_scratch_base <- projr::projr_path_get(
  "cache", "acs_cytof", "scratch",
  create = FALSE, format = "absolute"
)
path_comparison <- projr::projr_path_get(
  "cache", "acs_cytof", "manual-comparison", "manual-comparison.rds",
  create = FALSE, format = "absolute"
)
if (!file.exists(path_comparison)) {
  stop("Analysis 9's manual comparison is missing; render Analysis 9 first.")
}
manual_donors <- sort(unique(as.character(readRDS(path_comparison)$SampleID)))
channel_map <- .acsCytofChannelMap()
selection <- list(cd4 = c("p1", "mtb"), nk = c("ebv", "p1"))
pairs <- list(c("IFNg", "TNF"), c("IFNg", "IL2"), c("TNF", "IL2"))
```

```{r}
#| label: plots
#| results: asis
#| fig-width: 14
#| fig-height: 9
if (isTRUE(run_plots)) {
  kl_all <- list()
  mono_all <- list()
  just_all <- list()
  for (pop in names(selection)) {
    paths <- .acsCytofPopulationPaths(pop, "unused", path_gs_base, path_scratch_base)
    gs <- flowWorkspace::load_gs(paths$gs, backend_readonly = TRUE)
    sample_map <- .acsCytofReadPreprocessing(paths$gs, gs)$sampleMap
    gates <- stimgate::getStimGates(paths$stimgate)
    fig_dir <- .analysis_fig_dir(c(fig_key, pop), root_dir)
    cat("\n\n## ", pop, "\n\n", sep = "")
    for (stim in selection[[pop]]) {
      donors <- intersect(manual_donors, sample_map$SampleID[sample_map$stim == stim])
      dat <- stats::setNames(lapply(donors, function(donor) {
        .acsCytposTubeData(gs, sample_map, gates, donor, stim, channel_map)
      }), donors)
      donor_gates <- lapply(dat, `[[`, "gate")
      for (pair in pairs) {
        pair_key <- paste(pair, collapse = "-")
        cat("\n\n### ", pop, ", ", stim, ": ", pair[[1]], " and ", pair[[2]], "\n\n", sep = "")
        score_tbl <- do.call(rbind, lapply(donors, function(d) .acsScoreTbl(dat[[d]], pair, d)))
        tail <- .acsScoreTail(score_tbl)
        rect <- do.call(rbind, lapply(donors, function(d) .acsScoreRectNet(dat[[d]], pair, d)))
        kl_all[[length(kl_all) + 1L]] <- cbind(
          pop = pop, stim = stim, markers = pair_key, .acsScoreKl(score_tbl)
        )

        cat("\n\n#### Score histograms\n\n")
        .analysis_print_save_fig(.acsScoreHistPlot(score_tbl),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-hist.png")), width = 30, height = 22)
        cat("\n\n#### Cells at or above each score\n\n")
        .analysis_print_save_fig(.acsScoreTailPlot(tail),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-tail.png")), width = 30, height = 22)
        cat("\n\n#### Net frequency at or above each score\n\n")
        .analysis_print_save_fig(.acsScoreNetPlot(tail, rect),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-net.png")), width = 30, height = 22)
        cat("\n\n#### ", pair[[1]], " against ", pair[[2]], " with score contours\n\n", sep = "")
        hex_tbl <- do.call(rbind, lapply(donors, function(d) {
          .acsCytposHexTbl(dat[[d]], d, x = pair[[1]], y = pair[[2]])
        }))
        grid <- do.call(rbind, lapply(donors, function(d) {
          .acsScoreGrid(dat[[d]], pair[[1]], pair[[2]], d)
        }))
        .analysis_print_save_fig(
          .acsScoreHexPlot(hex_tbl, grid, donor_gates, pair[[1]], pair[[2]]),
          file.path(fig_dir, paste0(stim, "-", pair_key, "-hex.png")), width = 36, height = 26
        )
        if (identical(pair_key, "IFNg-TNF")) {
          cat("\n\n#### Departure from independence\n\n")
          res <- do.call(rbind, lapply(donors, function(d) {
            .acsCoexResiduals(.acsCoexBins(dat[[d]], pair[[1]], pair[[2]], d, width = 0.25))
          }))
          .analysis_print_save_fig(.acsCoexResidualPlot(res, donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-residuals.png")), width = 36, height = 26)
          indep <- lapply(donors, function(d) .acsCoexIndependence(dat[[d]], pair[[1]], pair[[2]], d))
          cat("\n\n#### Co-expression beyond independence\n\n")
          .analysis_print_save_fig(
            .acsCoexIndependencePlot(do.call(rbind, indep), donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-interaction.png")), width = 36, height = 26
          )
          cat("\n\n#### Stimulated minus control departure from independence\n\n")
          diff_tbl <- do.call(rbind, lapply(indep, .acsCoexDifference))
          .analysis_print_save_fig(
            .acsCoexDifferencePlot(diff_tbl, donor_gates, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-interaction-difference.png")), width = 30, height = 22
          )
          mono_tbl <- do.call(rbind, lapply(seq_along(donors), function(i) {
            .acsCoexIndependenceSummary(indep[[i]], dat[[donors[[i]]]], pair[[1]], pair[[2]])
          }))
          is_num <- vapply(mono_tbl, is.numeric, logical(1))
          mono_tbl[is_num] <- lapply(mono_tbl[is_num], round, 3)
          mono_all[[length(mono_all) + 1L]] <- cbind(pop = pop, stim = stim, mono_tbl)
          cat("\n\n#### Co-expression beyond independence against the rectangle\n\n")
          .analysis_report_table(mono_tbl,
            c(fig_key, paste0(pop, "-", stim, "-coexpression-beyond-independence.csv")),
            "Net excess over independence and cells called, against the rectangle.", root_dir)

          grow <- lapply(donors, function(d) .acsCoexGrow(dat[[d]], pair[[1]], pair[[2]], d))
          fbeta_tbl <- do.call(rbind, lapply(donors, function(d) {
            .acsCoexFbetaSummary(dat[[d]], pair[[1]], pair[[2]], d)
          }))
          cat("\n\n#### Grown co-expression region and F-beta gates\n\n")
          .analysis_print_save_fig(
            .acsCoexGrowHexPlot(hex_tbl, grow, donor_gates, fbeta_tbl, pair[[1]], pair[[2]]),
            file.path(fig_dir, paste0(stim, "-", pair_key, "-grown-region-hex.png")), width = 36, height = 26
          )
          grow_tbl <- do.call(rbind, lapply(grow, `[[`, "summary"))
          just_tbl <- data.frame(
            grow_tbl[c("donor", "floorX", "floorY", "binsAdded")],
            grownStim = grow_tbl$stimCalled, grownControl = grow_tbl$controlCalled,
            grownNetPct = grow_tbl$netCalledPct,
            fbetaGateX = fbeta_tbl$gateCytX, fbetaZX = fbeta_tbl$zX,
            fbetaGateY = fbeta_tbl$gateCytY, fbetaZY = fbeta_tbl$zY,
            fbetaStim = fbeta_tbl$stimCalled, fbetaControl = fbeta_tbl$controlCalled,
            fbetaNetPct = fbeta_tbl$netCalledPct,
            gateX = fbeta_tbl$gateX, gateY = fbeta_tbl$gateY,
            rectangleNetPct = grow_tbl$netRectanglePct
          )
          just_show <- just_tbl
          is_num <- vapply(just_show, is.numeric, logical(1))
          just_show[is_num] <- lapply(just_show[is_num], round, 3)
          just_all[[length(just_all) + 1L]] <- cbind(pop = pop, stim = stim, just_tbl)
          .analysis_report_table(just_show,
            c(fig_key, paste0(pop, "-", stim, "-grown-region-and-fbeta.csv")),
            "Grown co-expression region and F-beta gates per donor.", root_dir)
        }
      }
    }
  }
  .analysis_report_table(do.call(rbind, mono_all),
    c(fig_key, "coexpression-beyond-independence.csv"),
    "IFNg-TNF co-expression beyond independence against the rectangle of ordinary gates.",
    root_dir, quiet = TRUE)
  just_all <- do.call(rbind, just_all)
  .analysis_report_table(just_all, c(fig_key, "grown-region-and-fbeta.csv"),
    "Grown IFNg-TNF co-expression region and F-beta gates per donor.", root_dir, quiet = TRUE)
  cat("\n\n## Grown region and F-beta gates: summary\n\n")
  cat("Per population and stimulation (12 donors): median bins added by region growing,",
    "donors whose F-beta TNF gate (for IFNγ+ cells) and IFNγ gate (for TNF+ cells) were",
    "lowered, and the median net IFNγ+TNF+ frequency (%) and control cells called by each",
    "rule, against the rectangle of ordinary gates.\n\n")
  just_summary <- do.call(rbind, lapply(split(just_all, list(just_all$pop, just_all$stim), drop = TRUE), function(d) {
    data.frame(pop = d$pop[[1]], stim = d$stim[[1]],
      medianBinsAdded = stats::median(d$binsAdded),
      fbetaTnfLowered = sum(d$fbetaGateY < d$gateY - 1e-9),
      fbetaIfngLowered = sum(d$fbetaGateX < d$gateX - 1e-9),
      medianNetGrown = round(stats::median(d$grownNetPct), 3),
      medianNetFbeta = round(stats::median(d$fbetaNetPct), 3),
      medianNetRectangle = round(stats::median(d$rectangleNetPct), 3),
      medianControlGrown = stats::median(d$grownControl),
      medianControlFbeta = stats::median(d$fbetaControl))
  }))
  print(knitr::kable(just_summary, row.names = FALSE))
  kl_tbl <- do.call(rbind, kl_all)
  kl_tbl$klBits <- round(kl_tbl$klBits, 3)
  cat("\n\n## KL divergence of stimulated from control scores\n\n")
  .analysis_report_table(kl_tbl, c(fig_key, "kl-divergence.csv"),
    "KL divergence (bits) of each donor's stimulated from control score distribution.",
    root_dir)

}
```

```{r}
```
