---
title: Co-expression gates on every ACS sample
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
  run_compass: null
---

## What this document is

Analysis 17 explored a rule that lowers a cytokine's gate for cells positive
for another cytokine where the two are co-expressed beyond independence (the
"co-expression gates"; `scripts/r/coexpression-gates.R`, also available in
StimGate as `stimControl(cytPosMethod = "coexpression")`). Here the rule is
applied to every ACS population, stimulation and donor, starting from Analysis
9's ordinary gates, to see what it does at scale and to catch anything
pathological.

For each stimulated tube and every ordered pair of cytokines (a, b), b's gate
is lowered for cells above a's gate when:

1. the cells above both ordinary gates show a clear net response (z > 2);
2. working back from a floor (the control tube's main negative peak, from
   non-zero values, plus 1.5 robust SDs) in 20 bins, the bins from the ordinary
   gate down to the new gate depart from independence (Pearson residual ≥ 3.5,
   expected count from the stimulated tube's own marginal shares) and are at
   least 0.75 times as pure as the double positives (purity = 1 − control
   share / stimulated share), both for the lowest bin on its own and for all
   the bins together;
3. the added band is trimmed along a: a's cut rises (in 20ths of a's
   gate-to-floor distance) until the slice above it, and the band above that,
   are pure enough.

A cell is b-positive if above b's ordinary gate, or above the raised a cut and
the lowered b gate for any a. Control tubes are classified with their
stimulated tube's gates. Cytokines without a gate are negative.

For each population this document shows:

- background-subtracted frequencies (stimulated minus control) of
  single-positive cells (exactly one cytokine) and multi-positive cells (two or
  more), ordinary gates against co-expression gates, one point per donor;
- for each cytokine, how its single-positive frequency (positive for it alone)
  and multi-positive frequency (positive for it and another) change;
- a table of possible pathologies per tube (below);
- hexagon plots of the tubes whose multi-positive frequency changed most, for
  the pair of cytokines that added the most cells;
- COMPASS functionality and polyfunctionality scores fitted separately on the
  ordinary and the co-expression combination counts.

Possible pathologies flagged per tube:

- **bottom bin**: a gate lowered into the bin next to its floor, so the floor,
  not the data, stopped it;
- **control**: control cells called multi-positive at least doubled and up by
  ten or more;
- **multi jump**: the net multi-positive frequency more than tripled (from at
  least 0.01%);
- **single negative**: the net single-positive frequency driven below −0.05%.

Everything is read from Analysis 9's most recent run; Analysis 9's caches are
never changed. Results are cached under the projr cache folder
`acs_cytof_coexpression/`.

```{r}
#| label: setup
#| include: false
library(ggplot2)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
fig_key <- .analysis_mode_key("18-explore-acs-cytof-coexpression-gates", sized = FALSE)
for (fn in c(
  "acs_cytof-helper.R", "acs_cytof-gate.R", "acs_cytof-methods.R",
  "acs_cytof-debug.R", "acs_cytof-explore-cytpos.R", "coexpression-gates.R",
  "acs_cytof-coexpression-gates.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
run_simulations <- .as_flag(.get_qmd_param_env("run_simulations", "RUN_SIMULATIONS", FALSE))
run_plots <- .as_flag(.get_qmd_param_env("run_plots", "RUN_PLOTS", TRUE))
run_compass <- .get_qmd_param_env("run_compass", "RUN_COMPASS", NULL)
run_compass <- if (is.null(run_compass)) run_simulations else .as_flag(run_compass)
n_workers <- as.integer(Sys.getenv("ACS_COEX_WORKERS", "1"))
compass_iterations <- as.integer(Sys.getenv("ACS_COEX_COMPASS_ITERATIONS", "10000"))
```

```{r}
#| label: settings
#| include: false
pop_vec <- c("tcrgd", "cd4", "cd8", "nk", "nk_pre", "b")
rule_settings <- .acsCoexGatesSettings()
run_settings <- list(
  semantics = .acsCoexGatesSemantics, rule = rule_settings, pops = pop_vec
)
compass_settings <- c(run_settings, list(iterations = compass_iterations, replications = 8L, seed = 100L))
```

```{r}
#| label: paths
#| include: false
# Analysis 9's caches are only read: never create or change them here.
path_gs_base <- projr::projr_path_get("cache", "acs_cytof", "gs", create = FALSE, format = "absolute")
path_scratch_base <- projr::projr_path_get("cache", "acs_cytof", "scratch", create = FALSE, format = "absolute")
cache_dir <- projr::projr_path_get_dir(
  "cache", "acs_cytof_coexpression", create = run_simulations || run_compass, format = "absolute"
)
channel_map <- .acsCytofChannelMap()
path_pop <- function(pop) file.path(cache_dir, paste0(pop, ".rds"))
path_compass <- file.path(cache_dir, "compass.rds")
```

```{r}
#| label: run
#| include: false
# One population at a time; each is written atomically once complete.
if (isTRUE(run_simulations)) {
  for (pop in pop_vec) {
    message("Co-expression gates: ", pop)
    paths <- .acsCytofPopulationPaths(pop, "unused", path_gs_base, path_scratch_base)
    res <- .acsCoexGatesPopulation(pop, paths, channel_map, rule_settings, workers = n_workers)
    .write_rds_atomic(list(settings = run_settings, result = res), path_pop(pop))
  }
}
```

```{r}
#| label: collate
#| include: false
read_pop <- function(pop) {
  if (!file.exists(path_pop(pop))) {
    stop("No co-expression gate results for population '", pop, "' at ", path_pop(pop),
      ". Render with RUN_SIMULATIONS=true first.")
  }
  x <- readRDS(path_pop(pop))
  if (!identical(x$settings, run_settings)) {
    stop("Cached co-expression gate results for '", pop, "' were made with other settings.")
  }
  x$result
}
res_all <- lapply(stats::setNames(pop_vec, pop_vec), read_pop)
low_all <- do.call(rbind, lapply(res_all, `[[`, "low"))
counts_all <- do.call(rbind, lapply(res_all, `[[`, "counts"))
freq_all <- .acsCoexGatesFreq(counts_all)
paired_all <- .acsCoexGatesPaired(.acsCoexGatesSingleMulti(freq_all))
flags_all <- .acsCoexGatesFlags(low_all, paired_all, rule_settings)
```

```{r}
#| label: compass
#| include: false
if (isTRUE(run_compass)) {
  # One fit per population, stimulation and rule, forked over the workers.
  # Both rules use the same categories (`.acsCoexGatesCompassKeep()`).
  fit_jobs <- do.call(rbind, lapply(pop_vec, function(pop) {
    stims <- sort(unique(counts_all$stim[counts_all$pop == pop]))
    expand.grid(pop = pop, stim = stims, rule = c("ordinary", "coexpr"), stringsAsFactors = FALSE)
  }))
  fit_one <- function(i) {
    job <- fit_jobs[i, ]
    counts_curr <- counts_all[counts_all$pop == job$pop & counts_all$stim == job$stim, ]
    fit <- .acsCoexGatesCompass(counts_curr, job$rule, keep = .acsCoexGatesCompassKeep(counts_curr),
      iterations = compass_settings$iterations, replications = compass_settings$replications,
      seed = compass_settings$seed)
    fit$scores$pop <- job$pop
    fit$scores$stim <- job$stim
    fit$subsets$pop <- job$pop
    fit$subsets$stim <- job$stim
    fit
  }
  fits <- if (n_workers > 1L) {
    oplan <- future::plan(future::multicore, workers = n_workers)
    fits_out <- future.apply::future_lapply(seq_len(nrow(fit_jobs)), fit_one, future.seed = TRUE)
    future::plan(oplan)
    fits_out
  } else {
    lapply(seq_len(nrow(fit_jobs)), fit_one)
  }
  .write_rds_atomic(list(
    settings = compass_settings,
    scores = do.call(rbind, lapply(fits, `[[`, "scores")),
    subsets = do.call(rbind, lapply(fits, `[[`, "subsets"))
  ), path_compass)
}
compass <- if (file.exists(path_compass)) readRDS(path_compass)
if (!is.null(compass) && !identical(compass$settings, compass_settings)) compass <- NULL
```

## Overview

```{r}
#| label: overview
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  cat("Tubes per population (stimulated tubes with a control tube), median net single- and",
    "multi-positive frequencies (%) with the ordinary and co-expression gates, the number of",
    "ordered pairs lowered per tube (median) and tubes with each flag.\n\n")
  any_tbl <- paired_all[paired_all$cytokine == "any", ]
  overview <- do.call(rbind, lapply(pop_vec, function(pop) {
    s <- any_tbl[any_tbl$pop == pop & any_tbl$type == "single", ]
    m <- any_tbl[any_tbl$pop == pop & any_tbl$type == "multi", ]
    f <- flags_all[flags_all$pop == pop, ]
    data.frame(pop = pop, tubes = nrow(f),
      singleOrd = round(100 * stats::median(s$propBsOrdinary), 3),
      singleCoex = round(100 * stats::median(s$propBsCoexpr), 3),
      multiOrd = round(100 * stats::median(m$propBsOrdinary), 3),
      multiCoex = round(100 * stats::median(m$propBsCoexpr), 3),
      pairsLowered = stats::median(f$nLowered),
      bottomBin = sum(f$flagBottomBin), control = sum(f$flagControl),
      multiJump = sum(f$flagMultiJump), singleNegative = sum(f$flagSingleNegative))
  }))
  print(knitr::kable(overview, row.names = FALSE))
  .analysis_report_table(flags_all, c(fig_key, "tube-flags.csv"),
    "Per tube: pairs lowered, cells added and pathology flags.", root_dir)
  .analysis_report_table(low_all, c(fig_key, "lowered-gates.csv"),
    "Per tube and ordered cytokine pair: ordinary, lowered and raised gates, floors, z and purity.",
    root_dir, quiet = TRUE)
  .analysis_report_table(paired_all, c(fig_key, "single-multi-frequencies.csv"),
    "Per tube: single- and multi-positive background-subtracted frequencies, ordinary and co-expression gates.",
    root_dir, quiet = TRUE)
}
```

```{r}
#| label: populations
#| echo: false
#| results: asis
#| fig-width: 12
#| fig-height: 7
if (isTRUE(run_plots)) {
  for (pop in pop_vec) {
    fig_dir <- .analysis_fig_dir(c(fig_key, pop), root_dir)
    pop_paired <- paired_all[paired_all$pop == pop, ]
    cat("\n\n## ", pop, "\n\n", sep = "")
    cat("\n\n### Single- and multi-positive frequencies\n\n")
    .analysis_print_save_fig(.acsCoexGatesScatterPlot(pop_paired),
      file.path(fig_dir, "single-multi-scatter.png"), width = 30, height = 16)
    cat("\n\n### Change for each cytokine\n\n")
    .analysis_print_save_fig(.acsCoexGatesDiffPlot(pop_paired),
      file.path(fig_dir, "cytokine-change.png"), width = 30, height = 16)

    cat("\n\n### Tubes that changed most\n\n")
    cat("The eight tubes with the largest change in net multi-positive frequency; for each,",
      "the lowered pair that added the most stimulated cells (x: the cytokine the cells are",
      "positive for; y: the cytokine whose gate is lowered).\n\n")
    f <- flags_all[flags_all$pop == pop, ]
    f <- f[order(-abs(f$multiBsCoexpr - f$multiBsOrdinary)), ]
    f <- utils::head(f[f$nLowered > 0, ], 8L)
    if (nrow(f)) {
      paths <- .acsCytofPopulationPaths(pop, "unused", path_gs_base, path_scratch_base)
      gs <- flowWorkspace::load_gs(paths$gs, backend_readonly = TRUE)
      sample_map <- .acsCytofReadPreprocessing(paths$gs, gs)$sampleMap
      gates <- stimgate::getStimGates(paths$stimgate)
      tubes <- lapply(seq_len(nrow(f)), function(i) {
        lw <- low_all[low_all$pop == pop & low_all$stim == f$stim[[i]] &
          low_all$donor == f$donor[[i]] & low_all$lowered, ]
        lw <- lw[which.max(lw$addedStim), ]
        list(
          dat = .acsCytposTubeData(gs, sample_map, gates, f$donor[[i]], f$stim[[i]], channel_map),
          a = lw$a, b = lw$b, cut = lw$cut, condCut = lw$condCut,
          label = paste0(f$donor[[i]], " ", f$stim[[i]], ": ", lw$b, " among ", lw$a, "+")
        )
      })
      .analysis_print_save_fig(.acsCoexGatesTubeHexPlot(tubes),
        file.path(fig_dir, "most-changed-hex.png"), width = 30, height = 34)
    }
  }
}
```

## COMPASS

COMPASS (`SimpleCOMPASS`, `r compass_iterations` iterations, 8 replications,
seed 100) is fitted separately for each population and stimulation on the
combination counts from the ordinary gates and from the co-expression gates.
Both fits use the same categories: the all-negative one and those with at
least 5 stimulated cells in at least 3 donors under either rule (the rare
ones carry no information and slow the sampler). Each point is a donor's score
with the two sets of counts.

```{r}
#| label: compass-plots
#| echo: false
#| results: asis
#| fig-width: 14
#| fig-height: 12
if (isTRUE(run_plots)) {
  if (is.null(compass)) {
    cat("No COMPASS results for these settings; render with RUN_COMPASS=true.\n\n")
  } else {
    fig_dir <- .analysis_fig_dir(c(fig_key, "compass"), root_dir)
    cat("\n\n### Functionality score\n\n")
    .analysis_print_save_fig(.acsCoexGatesCompassPlot(compass$scores, "fs"),
      file.path(fig_dir, "functionality.png"), width = 36, height = 30)
    cat("\n\n### Polyfunctionality score\n\n")
    .analysis_print_save_fig(.acsCoexGatesCompassPlot(compass$scores, "pfs"),
      file.path(fig_dir, "polyfunctionality.png"), width = 36, height = 30)
    .analysis_report_table(compass$scores, c(fig_key, "compass-scores.csv"),
      "COMPASS functionality and polyfunctionality scores per donor, ordinary and co-expression gates.",
      root_dir)
    .analysis_report_table(compass$subsets, c(fig_key, "compass-subsets.csv"),
      "COMPASS mean posterior response probability per subset, ordinary and co-expression gates.",
      root_dir)
  }
}
```

```{r}
```
