---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Checking what the estimators actually estimate

Get the parameters over which we simulate.

So, as with checking what the actual estimated bandwidths are, we are going to have to vary over the number of cells, the probability of response, the mean position of the response, and the transformation applied to the data.

We will need to vary over more response probabilities, however, and of course need to vary over the bandwidths themselves.

Regarding the bandwidths themselves, as the `gamma` scenario is estimating such low bandwidths, we're going to have to start low.

We should probably allow varying amounts of intra-sample variation between the stimulated and unstimulated conditions.
But for now, let's just have two levels - no variation, slight variation.

Let's keep moderate background at 0.2.

Let's use a low and medium amount of bias.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()
transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1:2]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# fix the bandwidth
bandwidth_value_vec <- c(0.01, 0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 2, 5)[c(1, 3, 5, 6, 7)]

# can vary these later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- c(0, 0.05, 0.1, seq(0.2, 2, by = 0.2))[c(2, 4)]

if ("quick" %in% projr::projr_profile_get()) {
  n_cell_stim_vec <- c(1e3, 5e4)
  prob_response_vec <- c(0.002, 0.02)
  bandwidth_value_vec <- c(0.01, 0.1, 0.5, 0.5, 1)
  bias_uns_vec <- 0.1
  bandwidth_value_vec <- c(0.05, 0.1, 0.25, 0.5, 0.75)
  condition_perturbation_sd_vec <- 0
}
```

### Varying bandwidth manually

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
So, we can vary the bandwidth from a small number (say 0.1) to a large number (say 10), and see how estimated background-subtracted frequencies compare to actual.

It should be doable with just the `bw` parameter, but ah, we haven't set that yet.

Okay, we've got the code working above. Now we'll vary the fixed bandwidth from 0.01 to 4, along with making the bias_unstim some reasonably large value (we'll have to fix it).

And we'll do this for varying numbers of cells. 

Generate the grid of parameters to simulate.

```{r}
#| label: bw-manual-grid
# 2. Flatten your nested loops into a single combination grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw             = bandwidth_value_vec,
  bias_uns       = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw, bias_uns,
    sample_perturbation_sd, condition_perturbation_sd,
    cluster_perturbation_sd, background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))


if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw == 0.5,
      bias_uns == 0.05,
      condition_perturbation_sd == 0
    )
}
sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-manual-loop
#| eval: false
bw_list_raw <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))
  sim_res <- stimgate:::.simBandwidthBsFreq(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 1,
    biasUns = sim_grid$bias_uns[i],
    bw = sim_grid$bw[i],
    bwFallback = sim_grid$bw[i],
    bwMin = "none",
    bwMax = "none",
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = sim_grid$sample_perturbation_sd[i],
    conditionPerturbationSd = sim_grid$condition_perturbation_sd[i],
    clusterPerturbationSd = sim_grid$cluster_perturbation_sd[i],
    backgroundRelativeToResponse =
      sim_grid$background_relative_to_response[i],
    ncellUnsRelativeToStim =
      sim_grid$n_cell_uns_relative_to_stim[i],
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw = sim_grid$bw[i],
      bias_uns = sim_grid$bias_uns[i],
      sample_perturbation_sd = sim_grid$sample_perturbation_sd[i],
      condition_perturbation_sd =
        sim_grid$condition_perturbation_sd[i],
      cluster_perturbation_sd = sim_grid$cluster_perturbation_sd[i],
      background_relative_to_response =
        sim_grid$background_relative_to_response[i],
      n_cell_uns_relative_to_stim =
        sim_grid$n_cell_uns_relative_to_stim[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation, prob_response, n_cell,
      mean_pos_setting, mean_pos, bw, bias_uns,
      sample_perturbation_sd, condition_perturbation_sd,
      cluster_perturbation_sd, background_relative_to_response,
      n_cell_uns_relative_to_stim, sim_id,
      dplyr::everything()
    )

  bw_list_raw[[i]] <- sim_res
}
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-manual-parallel
#| results: hide
# Activate global progress bars in your console
if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache", "sim", "bw", "freq_bs", "progress.txt"
)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

# Set up the parallel backend
future::plan(
  future::multisession,
  workers = max(1, future::availableCores() - 1)
)

# Wrap the parallel execution in with_progress()
bw_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | bw = ", bw,
        " | bias = ", bias_uns,
        " | sample_sd = ", sample_perturbation_sd,
        " | condition_sd = ", condition_perturbation_sd,
        " | cluster_sd = ", cluster_perturbation_sd,
        " | bg_rel = ", background_relative_to_response,
        " | n_uns_rel = ", n_cell_uns_relative_to_stim,
        " | sim_id = ", sim_id
      )

      log_msg <- paste0("Running: ", settings_row, "
")
      cat(log_msg, file = path_progress_file, append = TRUE)

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthBsFreq(
            nSample = 5,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 5,
            biasUns = bias_uns,
            bw = bw,
            bwMin = "none",
            bwMax = "none",
            bwFallback = bw,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            samplePerturbationSd = sample_perturbation_sd,
            conditionPerturbationSd = condition_perturbation_sd,
            clusterPerturbationSd = cluster_perturbation_sd,
            backgroundRelativeToResponse = background_relative_to_response,
            ncellUnsRelativeToStim = n_cell_uns_relative_to_stim,
            covEvMin = 1.5,
            covEvMax = 1.5,
            tolClust = NULL
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw = bw,
              bias_uns = bias_uns,
              sample_perturbation_sd = sample_perturbation_sd,
              condition_perturbation_sd = condition_perturbation_sd,
              cluster_perturbation_sd = cluster_perturbation_sd,
              background_relative_to_response =
                background_relative_to_response,
              n_cell_uns_relative_to_stim =
                n_cell_uns_relative_to_stim,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos_setting,
              mean_pos,
              bw,
              bias_uns,
              sample_perturbation_sd,
              condition_perturbation_sd,
              cluster_perturbation_sd,
              background_relative_to_response,
              n_cell_uns_relative_to_stim,
              sim_id,
              dplyr::everything()
            )

          cat(
            "Completed: ",
            settings_row,
            "
",
            file = path_progress_file,
            append = TRUE
          )

          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          err_msg <- paste0("Error [", settings_row, "]: ", e$message, "
")
          cat(err_msg, file = path_progress_file, append = TRUE)
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw = bw,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response =
              background_relative_to_response,
            n_cell_uns_relative_to_stim =
              n_cell_uns_relative_to_stim,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            method = NA_character_,
            propRespTruth = NA_real_,
            propRespEst = NA_real_,
            propBsEst = NA_real_,
            threshold = NA_real_,
            thresholdOrigin = NA_character_,
            gateReturnPoint = NA_character_,
            locGenerated = NA,
            locGeneratedDirect = NA,
            locSource = NA_character_,
            locReason = NA_character_,
            detailLevel = NA_character_,
            nCellStim = NA_real_,
            nCellUns = NA_real_,
            nPosStim = NA_integer_,
            nPosUns = NA_integer_
          )
        }
      )

      return(out)
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

# Close the parallel workers when done
future::plan(future::sequential)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "output", "sim", "bw", "freq_bs", "bw_list_raw.rds"
  )
)
```

Collate the results and calculate the metrics of interest.

```{r}
#| label: bw-manual-collate
.add_missing_cols <- function(.data, cols) {
  for (nm in names(cols)) {
    if (!nm %in% names(.data)) {
      .data[[nm]] <- cols[[nm]]
    }
  }
  .data
}

# Prefer the object just created in this render. If this chunk is re-run alone,
# fall back to the cached/output RDS. Do not allow a failed collation to be
# silently swallowed, otherwise later chunks fail with an unhelpful
# "object not found" error.
if (exists("bw_list_raw") && length(bw_list_raw) > 0L) {
  bw_list_raw_collate <- bw_list_raw
} else {
  path_vec <- c(
    projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "bw_list_raw.rds",
      safe = TRUE
    ),
    projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "bw_list_raw.rds",
      safe = FALSE
    ),
    projr::projr_path_get(
      "raw-data-archive-manual",
      "v0.103.0",
      "output",
      "sim",
      "bw",
      "freq_bs",
      "bw_list_raw.rds",
      safe = FALSE
    )
  )
  path_vec <- path_vec[file.exists(path_vec)]
  if (length(path_vec) == 0L) {
    stop("Could not find bw_list_raw.rds, and bw_list_raw is not available in memory.")
  }
  bw_list_raw_collate <- readRDS(file = path_vec[[1]])
}

bw_tbl <- bw_list_raw_collate |>
  purrr::list_rbind()

bw_tbl <- .add_missing_cols(
  bw_tbl,
  list(
    iter = NA_integer_,
    sample = NA_character_,
    ind = NA_character_,
    propBsEst = NA_real_,
    threshold = NA_real_
  )
)

bw_tbl_results_raw <- bw_tbl |>
  dplyr::select(
    dplyr::any_of(colnames(sim_grid)),
    iter,
    sample,
    ind,
    propBsEst,
    threshold
  ) |>
  dplyr::filter(is.finite(threshold) & is.finite(propBsEst)) |>
  dplyr::group_by(sim_id, iter, ind) |>
  dplyr::mutate(
    threshold = unique(threshold[is.finite(threshold)])[[1]],
    propBsEst = unique(propBsEst[is.finite(propBsEst)])[[1]]
  ) |>
  dplyr::ungroup() |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid))),
    iter,
    ind
  ) |>
  dplyr::slice(1) |>
  dplyr::ungroup()

if (nrow(bw_tbl_results_raw) == 0L) {
  warning(
    "No valid bandwidth-frequency results were collated. ",
    "Check the progress log for simulation-level errors."
  )
}

bw_tbl_results_summary <- bw_tbl_results_raw |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    dplyr::across(
      c(threshold, propBsEst),
      list(
        min = ~ min(.x, na.rm = TRUE),
        max = ~ max(.x, na.rm = TRUE),
        mean = ~ mean(.x, na.rm = TRUE),
        median = ~ stats::median(.x, na.rm = TRUE)
      )
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    propBsEst_median_diff = propBsEst_median - prob_response,
    propBsEst_median_diff_rel = (propBsEst_median - prob_response) /
      prob_response
  )
```

Okay, now we must plot which bandwidths are best for which scenarios.

```{r}
if (requireNamespace("UtilsDataRSV", quietly = TRUE)) {
  sim_grid |>
    UtilsDataRSV::view_cols()
} else {
  dplyr::glimpse(sim_grid)
}
```

Okay, so we varied the transformation, the prob_response, the n_cell, the mean_pos_setting, the bw.

So, let's plot it perhaps how we did it for the bandwidth estimators? We have one line for each bandwidth, and then the x-axis as the number of cells, and the y-axis as the estimated background-subtracted frequency.
We'll do one plot for each of the mean_pos_settings.
Then, we have transformation along the x-axis, as usual, and the response probability along the y-axis (ito the facet, i.e. response probability gives you the row, transformation gives you the column).
Okay, and we can plot the median estimated background-subtracted frequency.
Let's leave error bars out for now.

```{r}
col_vec_bw <- c(
  "#e0ecf4",
  "#bfd3e6",
  "#9ebcda",
  "#8c96c6",
  "#8c6bb1",
  "#88419d",
  "#810f7c"
)
for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
  plot_tbl <- bw_tbl_results_summary |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

  col_vec_bw_curr <- col_vec_bw[seq(length(col_vec_bw) - length(unique(plot_tbl$bw)) + 1, length(col_vec_bw))]

  p <- ggplot(
    plot_tbl,
    aes(x = n_cell, y = propBsEst_median, color = factor(bw))
  ) +
    geom_line() +
    geom_hline(
      mapping = aes(yintercept = prob_response),
      linetype = "dashed",
      color = "black"
    ) +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation, scales = "free_y") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = paste0("Estimated background-subtracted frequency vs number of cells (mean_pos_setting = ", mean_pos_setting_curr, ")"),
      x = "Number of cells (log scale)",
      y = "Estimated background-subtracted frequency (median)",
      color = "Bandwidth"
    ) +
    scale_colour_manual(
      values = col_vec_bw_curr
    )

  path_p <- projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "freq_bs",
    "est",
    "median",
    paste0("bw_est_vs_n_cell_mean_pos_setting_", mean_pos_setting_curr, ".png")
  )
  dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
  ggsave(
    filename = path_p,
    plot = p,
    width = 18,
    height = 30,
    units = "cm"
  )
}
```

Okay, now, let's do the same plot but where we show the actual frequencies. We'll have to treat the number of cells as a categorical variable, and use the bandwidth as positoin = "dodge".

```{r}
for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
  plot_tbl <- bw_tbl_results_raw |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)
  
  n_bw <- length(unique(plot_tbl$bw))
  bw_col_vec_curr <- col_vec_bw[seq(length(col_vec_bw) - n_bw + 1, length(col_vec_bw))]
  p <- ggplot(
    plot_tbl,
    aes(x = factor(n_cell), y = propBsEst, colour = factor(bw))
  ) +
    ggforce::geom_sina(position = "dodge") +
    geom_hline(
      mapping = aes(yintercept = prob_response),
      linetype = "dashed",
      color = "black"
    ) +
    facet_grid(prob_response ~ transformation, scales = "free_y") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = paste0("Actual background-subtracted frequency vs number of cells (mean_pos_setting = ", mean_pos_setting_curr, ")"),
      x = "Number of cells",
      y = "Actual background-subtracted frequency",
      colour = "Bandwidth"
    ) +
    scale_colour_manual(
      values = col_vec_bw
    )

  path_p <- projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "freq_bs",
    "est",
    "raw",
    paste0("actual_freq_vs_n_cell_mean_pos_setting_", mean_pos_setting_curr, ".png")
  )
  dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
  ggsave(
    filename = path_p,
    plot = p,
    width = 18,
    height = 30,
    units = "cm"
  )
}
```

Okay, now we wanto plot the actual thresholds.
We should take the same plot as used to display the simulations, and plot those.
Except now we'll facet with the rows being probability of a response, and a file per mean_pos_setting. Column transformation again.
We'll just keep the number of cells high.

```{r}
#| label: main-settings
main_settings <- tibble::tibble(
  n_cell = 1e5,
  background_relative_to_response = 0.2,
  prob_exact = TRUE,
  mixture_type = "gaussianOnly",
  cluster_perturbation_sd = 0,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5
)

mean_pos_settings_tbl <- purrr::map_df(
  unique(sim_grid$prob_response), function(prob_response) {
    .simMiscGetMeanPosTbl() |>
      dplyr::mutate(
        prob_response = prob_response
      )
  }
) 
transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

main_settings
```

```{r}
make_bw_linetype_scale <- function(bw_vec, digits = 2) {
  bw_num <- sort(unique(as.numeric(bw_vec)))
  bw_lab <- formatC(bw_num, format = "f", digits = digits)

  n_bw <- length(bw_num)

  if (n_bw == 1L) {
    lty <- "solid"
  } else {
    # solid, then progressively shorter dash lengths
    dash_len <- round(seq(8, 2, length.out = n_bw - 1L))
    gap_len <- 4L

    lty <- c(
      "solid",
      paste0(
        as.character(as.hexmode(dash_len)),
        as.character(as.hexmode(gap_len))
      )
    )
  }

  list(
    levels = bw_lab,
    values = stats::setNames(lty, bw_lab)
  )
}
```

```{r}
# These objects are only needed for the optional density/rug overlay in the
# threshold plots. Keep the document renderable if they have not been prepared
# in this run.
if (!exists("dens_tbl")) {
  dens_tbl <- tibble::tibble(
    F1 = numeric(),
    density = numeric(),
    condition = character(),
    prob_response = numeric(),
    transformation = character(),
    mean_pos_setting = character()
  )
}

if (!exists("rug_tbl")) {
  rug_tbl <- tibble::tibble(
    F1 = numeric(),
    condition = character(),
    prob_response = numeric(),
    transformation = character(),
    mean_pos_setting = character()
  )
}

if (!exists("mean_line_tbl")) {
  mean_line_tbl <- tibble::tibble(
    mean_pos_setting = character()
  )
}
```

```{r}
for (mean_pos_setting in unique(bw_tbl_results_summary$mean_pos_setting)) {
  threshold_tbl_curr <- bw_tbl_results_summary |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting)
  mean_line_tbl_curr <- mean_line_tbl |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting)
  rug_tbl_curr <- rug_tbl |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting)
  dens_tbl_curr <- dens_tbl |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting)

  bw_scale <- make_bw_linetype_scale(threshold_tbl_curr$bw)

  threshold_tbl_curr <- threshold_tbl_curr |>
    dplyr::mutate(
      bw_lab = factor(
        formatC(as.numeric(.data$bw), format = "f", digits = 2),
        levels = bw_scale$levels
      )
    )


  p <- ggplot() +
    geom_area(
      data = dens_tbl_curr,
      mapping = aes(x = F1, y=  density, colour = condition, fill = condition),
      alpha = 0.18,
      position = "identity"
    ) +
    geom_line(
      data = dens_tbl_curr,
      mapping = aes(x = F1, y=  density, colour = condition),
      linewidth = 0.25
    ) +
      geom_rug(
    data = rug_tbl_curr,
      aes(
        x = F1,
        colour = condition
      ),
      sides = "b",
      alpha = 0.25,
      linewidth = 0.15
    ) +
      facet_grid(
        rows = vars(.data$prob_response),
        cols = vars(.data$transformation),
        scales = "free_x"
      ) +
      scale_y_sqrt() +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom",
        strip.text.y = element_text(angle = 0)
      ) +
      coord_cartesian(ylim = c(0, 1)) +
      scale_colour_manual(
        values = .simMiscGetStimColVec()
      ) +
      labs(
        x = "Marker expression",
        y = "Density, square-root scale",
        colour = NULL,
        fill = NULL,
        linetype = "Bandwidth"
      ) +
      geom_vline(
        data = threshold_tbl_curr,
        mapping = aes(xintercept = threshold_median, linetype = bw_lab),
        linewidth = 0.45
      ) +
      scale_linetype_manual(
        values = bw_scale$values,
        breaks = bw_scale$levels,
        labels = bw_scale$levels
      )
    path_p <- projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "thresholds",
      "median",
      paste0("threshold_vs_n_cell_mean_pos_setting_", mean_pos_setting, ".png")
    )
    dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
    ggsave(
      filename = path_p,
      plot = p,
      width = 18,
      height = 30,
      units = "cm"
    ) 
}
```

Okay, now let's do the same, but we will make also one plot for each bandwidth setting, and show the actual thresholds all the samples and iterations.

So, now, we loop not just over mean_pos_settings, but also over the bandwidths.

```{r}
for (mean_pos_setting in unique(bw_tbl_results_raw$mean_pos_setting)) {
  for (bw_curr in unique(bw_tbl_results_raw$bw)) {
    if (!is.finite(bw_curr)) {
      next
    }
    threshold_tbl_curr <- bw_tbl_results_raw |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting,
        bw == !!bw_curr
      )
    threshold_tbl_curr_med_iqr <- threshold_tbl_curr |>
      dplyr::group_by(
        dplyr::pick(dplyr::any_of(colnames(sim_grid)))
      ) |>
      dplyr::summarise(
        threshold_median = stats::median(threshold, na.rm = TRUE),
        threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
        threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
        .groups = "drop"
      ) |>
      tidyr::pivot_longer(
        threshold_median:threshold_iqr_upper,
        names_to = "threshold_stat",
        values_to = "threshold"
      ) |>
      dplyr::mutate(
        threshold_stat = gsub("threshold_", "", threshold_stat),
        threshold_stat = factor(
          threshold_stat,
          levels = c(
            "iqr_lower",
            "median",
            "iqr_upper"
          )
        )
      )
    rug_tbl_curr <- rug_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting)
    dens_tbl_curr <- dens_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting)

    p <- ggplot() +
      geom_area(
        data = dens_tbl_curr,
        mapping = aes(x = F1, y=  density, colour = condition, fill = condition),
        alpha = 0.18,
        position = "identity"
      ) +
      geom_line(
        data = dens_tbl_curr,
        mapping = aes(x = F1, y=  density, colour = condition),
        linewidth = 0.25
      ) +
      geom_rug(
        data = rug_tbl_curr,
        aes(
          x = F1,
          colour = condition
        ),
        sides = "b",
        alpha = 0.25,
        linewidth = 0.15
      ) +
      facet_grid(
        rows = vars(.data$prob_response),
        cols = vars(.data$transformation),
        scales = "free_x"
      ) +
      scale_y_sqrt() +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom",
        strip.text.y = element_text(angle = 0)
      ) +
      coord_cartesian(ylim = c(0, 1)) +
      scale_colour_manual(
        values = .simMiscGetStimColVec()
      ) +
      labs(
        x = "Marker expression",
        y = "Density, square-root scale",
        colour = NULL,
        fill = NULL
      ) +
      geom_vline(
        data = threshold_tbl_curr,
        mapping = aes(xintercept = threshold, group = interaction(iter, sample)),
        linewidth = 0.25,
        alpha = 0.25
      ) +
      geom_vline(
        data = threshold_tbl_curr_med_iqr,
        aes(xintercept = threshold, linetype = threshold_stat),
        linewidth = 0.45
      ) +
      scale_linetype_manual(
        values = c(
          "iqr_lower" = "dashed",
          "median" = "solid",
          "iqr_upper" = "dashed"
        ),
        breaks = c("iqr_lower", "median", "iqr_upper"),
        labels = c("IQR lower", "Median", "IQR upper")
      )

    path_p <- projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "thresholds",
      "raw",
      formatC(bw_curr, format = "f", digits = 2),
      paste0(
        "threshold_vs_n_cell_mean_pos_setting_bw_",
        mean_pos_setting,
        "_bw_",
        formatC(bw_curr, format = "f", digits = 2),
        ".png"
      )
    )
    dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
    ggsave(
      filename = path_p,
      plot = p,
      width = 18,
      height = 30,
      units = "cm"
    )
  }
}
```

Okay, let's get a summary table, where for each setting above, for each bandwidth we have the min, max, median, iqr lower, iqr upper, sd, mad, iqr length.

```{r}
summary_tbl <- bw_tbl_results_raw |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid))),
    bw
  ) |>
  dplyr::summarise(
    threshold_min = min(threshold, na.rm = TRUE),
    threshold_max = max(threshold, na.rm = TRUE),
    threshold_median = stats::median(threshold, na.rm = TRUE),
    threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
    threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
    threshold_sd = stats::sd(threshold, na.rm = TRUE),
    threshold_mad = stats::mad(threshold, na.rm = TRUE),
    threshold_iqr_length = threshold_iqr_upper - threshold_iqr_lower,
    .groups = "drop"
  )
```

```{r}
summary_tbl |>
  dplyr::select(transformation:bw, threshold_sd) |>
  dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 2))) |>
  dplyr::arrange(threshold_sd) |>
  dplyr::mutate(
    threshold_sd_rank = dplyr::dense_rank(threshold_sd)
  ) |>
  dplyr::group_by(transformation, mean_pos_setting, bw) |>
  dplyr::summarise(
    mean_rank = mean(threshold_sd_rank, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(mean_rank) |>
  dplyr::filter(!transformation == "gamma") |>
  knitr::kable()
```