---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```


### Checking what the estimators actually estimate

Get the parameters over which we simulate.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_vec <- seq_len(8)[c(2, 4, 6, 8)]
transformation_vec <- c("gaussian", "gamma", "skew")
sample_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
condition_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
cluster_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
background_relative_to_response <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)
n_cell_uns_relative_to_stim <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)

# bandwidth settings
bandwidth_method_vec <- c("hpi0", "hpi1", "hpi2", "hpi3", "sj", "nrd0")
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- seq(0, 2, by = 0.2)

# fix the bandwidth
bandwidth_value_vec <- c(0.01, 0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 2, 5)[c(3:8)]

# number of iterations
length(n_cell_stim_vec) * length(prob_response_vec) * length(mean_pos_vec) * length(transformation_vec) * length(bandwidth_value_vec)
```


### Varying bandwidth manually

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
So, we can vary the bandwidth from a small number (say 0.1) to a large number (say 10), and see how estimated background-subtracted frequencies compare to actual.

It should be doable with just the `bw` parameter, but ah, we haven't set that yet.

Okay, we've got the code working above. Now we'll vary the fixed bandwidth from 0.01 to 4, along with making the bias_unstim some reasonably large value (we'll have to fix it).

And we'll do this for varying numbers of cells. 

Generate the grid of parameters to simulate.

```{r}
#| label: bw-manual-grid
# 2. Flatten your nested loops into a single combination grid
sim_grid <- tidyr::expand_grid(
  transformation = transformation_vec,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  mean_pos       = mean_pos_vec,
  bw             = bandwidth_value_vec
) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))


if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos == 8,
      transformation %in% transformation_vec,
      bw == 0.5
    )
}
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-manual-loop
#| eval: false
for (i in seq_len(nrow(sim_grid))) {
  sim_res <- stimgate:::.simBandwidthBsFreq(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 10,
    biasUns = 0.05,
    bw = sim_grid$bw[i],
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = 0,
    conditionPerturbationSd = 0,
    clusterPerturbationSd = 0,
    backgroundRelativeToResponse = 0.2, 
    ncellUnsRelativeToStim = 1,
    covEvMin = 2, 
    covEvMax = 2,
    tolClust = NULL
  ) |>
    dplyr::mutate(
      transformation = transformation,
      prob_response = prob_response,
      n_cell = n_cell,
      mean_pos = mean_pos,
      bw = bw
    ) |>
    dplyr::select(
      transformation, prob_response, n_cell,
      mean_pos, bw, dplyr::everything()
    )
}
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-manual-parallel
# Activate global progress bars in your console
if (interactive()) {
  handlers(global = TRUE)
}


path_progress_file <- projr::projr_path_get(
  "cache", "sim-bw", "bs-freq", "progress.txt"
)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

# Set up the parallel backend
plan(multisession, workers = availableCores() - 1)

# 3. Wrap your parallel execution in with_progress()
bw_list_raw <- with_progress({
  
  # Create a progressor function with the total number of steps
  p <- progressor(steps = nrow(sim_grid))
  
  future_pmap(sim_grid, function(transformation, prob_response, n_cell, mean_pos, bw, sim_id) {

    settings_row <- paste0("trans = ", transformation, 
                           " | prob = ", prob_response, 
                           " | n_cell = ", n_cell, 
                           " | mean = ", mean_pos, 
                           " | bw = ", bw,
                           " | sim_id = ", sim_id)

    log_msg <- paste0("Running: ", settings_row, "\n")
    cat(log_msg, file = path_progress_file, append = TRUE)
    
    # Run the simulation
    out <- tryCatch({
      sim_res <- stimgate:::.simBandwidthBsFreq(
        nSample = 10,
        nMarker = 1,
        nCondition = 2,
        nCluster = 2,
        nIter = 10,
        biasUns = 0.05,
        bw = bw,
        nCellStim = n_cell,
        probResponse = prob_response,
        probExact = TRUE,
        meanPos = mean_pos,
        transformation = transformation,
        samplePerturbationSd = 0,
        conditionPerturbationSd = 0,
        clusterPerturbationSd = 0,
        backgroundRelativeToResponse = 0.2, 
        ncellUnsRelativeToStim = 1,
        covEvMin = 2, 
        covEvMax = 2,
        tolClust = NULL
      ) |>
        dplyr::mutate(
          transformation = transformation,
          prob_response = prob_response,
          n_cell = n_cell,
          mean_pos = mean_pos,
          bw = bw
        ) |>
        dplyr::select(
          transformation, prob_response, n_cell,
          mean_pos, bw, dplyr::everything()
        )

      cat("Completed: ", settings_row, "\n", file = path_progress_file, append = TRUE)
      
      # Signal successful step to the progress bar with a custom message
      p(sprintf("Completed sim_id: %s", sim_id))
      
      sim_res
    }, error = function(e) {
      
      # Signal failed step to the progress bar so it keeps moving
      err_msg <- paste0("Error [", settings_row, "]: ", e$message, "\n")
      cat(err_msg, file = path_progress_file, append = TRUE)
      p(sprintf("ERROR on sim_id: %s", sim_id))
      
      tibble::tibble(
        transformation = transformation,
        prob_response = prob_response,
        n_cell = n_cell,
        mean_pos = mean_pos,
        bw = bw,
        iter = NA,
        chnl = NA,
        sample = NA
      )
    })
    
    return(out)
  }, .options = furrr_options(seed = TRUE))
})

# Close the parallel workers when done
plan(sequential)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "output", "sim-bw", "bs-freq", "bw_list_raw.rds"
  )
)
```

Collate the results and calculate the metrics of interest.

```{r}
#| label: bw-manual-collate
#| error: true
bw_list_raw <- readRDS(
  file = projr::projr_path_get(
    "output", "sim-bw", "bs-freq", "bw_list_raw.rds"
  )
)

bw_tbl <- bw_list_raw |>
  dplyr::bind_rows() |>
  dplyr::filter(!is.na(iter))

# Extract gates
bw_gate_detail_tbl <- bw_tbl |>
  dplyr::filter(grepl("^loc_", .data$method)) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos,
    bw,
    iter,
    chnl,
    sample,
    method,
    propRespTruth,
    propRespEst,
    threshold,
    thresholdOrigin,
    locGenerated,
    locGeneratedDirect,
    locSource,
    locReason,
    detailLevel
  )

bw_gate_detail_tbl

bw_tbl |>
  dplyr::group_by(
    transformation, prob_response, n_cell, mean_pos,
    bw, method, chnl
    ) |>
  dplyr::summarise(
    mean_diff = mean(abs(propRespEst - propRespTruth)),
    rel_diff = mean(abs(propRespEst - propRespTruth)) / mean(propRespTruth),
    ccc = DescTools::CCC(propRespTruth, propRespEst)$rho.c$est,
    pearson = cor(propRespTruth, propRespEst, method = "pearson"),
    spearman = cor(propRespTruth, propRespEst, method = "spearman"),
    .groups = "drop"
  )  |>
  dplyr::arrange(transformation, prob_response, n_cell, mean_pos, bw)
```

```{r}
#| label: bw-manual-collate-1
#| error: true
bw_1 <- bw_list_raw[[1]] |>
  dplyr::filter(
    transformation == transformation[[1]],
    prob_response == max(prob_response),
    n_cell == n_cell[[1]],
    mean_pos == max(mean_pos),
    chnl == chnl[[1]],
    method == method[[1]]
  )
bw_1
bw_1 |>
  dplyr::group_by(
    transformation, prob_response, n_cell, mean_pos,
    bw, method, chnl
    ) |>
  dplyr::summarise(
    ccc = DescTools::CCC(propRespTruth, propRespEst)$rho.c$est,
    pearson = cor(propRespTruth, propRespEst, method = "pearson"),
    spearman = cor(propRespTruth, propRespEst, method = "spearman"),
    .groups = "drop"
  ) |>
  dplyr::arrange(transformation, prob_response, n_cell, mean_pos, bw)
```