---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Checking what the estimators actually estimate

Get the parameters over which we simulate.

So, as with checking what the actual estimated bandwidths are, we are going to have to vary over the number of cells, the probability of response, the mean position of the response, and the transformation applied to the data.

We will need to vary over more response probabilities, however, and of course need to vary over the bandwidths themselves.

Regarding the bandwidths themselves, as the `gamma` scenario is estimating such low bandwidths, we're going to have to start low.

We should probably allow varying amounts of intra-sample variation between the stimulated and unstimulated conditions.
But for now, let's just have two levels - no variation, slight variation.

Let's keep moderate background at 0.2.

Let's use a low and medium amount of bias.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_vec <- seq_len(8)[c(4, 8)]
transformation_vec <- c("gaussian", "gamma", "skew")
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1:2]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# fix the bandwidth
bandwidth_value_vec <- c(0.01, 0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 2, 5)[c(1, 3, 5, 6, 7)]

# can vary these later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- c(0, 0.05, 0.1, seq(0.2, 2, by = 0.2))[c(2, 4)]
```

### Varying bandwidth manually

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
So, we can vary the bandwidth from a small number (say 0.1) to a large number (say 10), and see how estimated background-subtracted frequencies compare to actual.

It should be doable with just the `bw` parameter, but ah, we haven't set that yet.

Okay, we've got the code working above. Now we'll vary the fixed bandwidth from 0.01 to 4, along with making the bias_unstim some reasonably large value (we'll have to fix it).

And we'll do this for varying numbers of cells. 

Generate the grid of parameters to simulate.

```{r}
#| label: bw-manual-grid
# 2. Flatten your nested loops into a single combination grid
sim_grid <- tidyr::expand_grid(
  transformation = transformation_vec,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  mean_pos       = mean_pos_vec,
  bw             = bandwidth_value_vec,
  bias_uns       = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))


if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos == 8,
      transformation %in% transformation_vec,
      bw == 0.5,
      bias_uns == 0.05,
      condition_perturbation_sd == 0
    )
}
sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-manual-loop
#| eval: false
bw_list_raw <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))
  sim_res <- stimgate:::.simBandwidthBsFreq(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 1,
    biasUns = sim_grid$bias_uns[i],
    bw = sim_grid$bw[i],
    bwFallback = sim_grid$bw[i],
    bwMin = "none",
    bwMax = "none",
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = sim_grid$sample_perturbation_sd[i],
    conditionPerturbationSd = sim_grid$condition_perturbation_sd[i],
    clusterPerturbationSd = sim_grid$cluster_perturbation_sd[i],
    backgroundRelativeToResponse =
      sim_grid$background_relative_to_response[i],
    ncellUnsRelativeToStim =
      sim_grid$n_cell_uns_relative_to_stim[i],
    covEvMin = 2,
    covEvMax = 2,
    tolClust = NULL
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos = sim_grid$mean_pos[i],
      bw = sim_grid$bw[i],
      bias_uns = sim_grid$bias_uns[i],
      sample_perturbation_sd = sim_grid$sample_perturbation_sd[i],
      condition_perturbation_sd =
        sim_grid$condition_perturbation_sd[i],
      cluster_perturbation_sd = sim_grid$cluster_perturbation_sd[i],
      background_relative_to_response =
        sim_grid$background_relative_to_response[i],
      n_cell_uns_relative_to_stim =
        sim_grid$n_cell_uns_relative_to_stim[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation, prob_response, n_cell,
      mean_pos, bw, bias_uns,
      sample_perturbation_sd, condition_perturbation_sd,
      cluster_perturbation_sd, background_relative_to_response,
      n_cell_uns_relative_to_stim, sim_id,
      dplyr::everything()
    )

  bw_list_raw[[i]] <- sim_res
}
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-manual-parallel
#| results: hide
# Activate global progress bars in your console
if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache", "sim", "bw", "freq_bs", "progress.txt"
)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

# Set up the parallel backend
future::plan(
  future::multisession,
  workers = max(1, future::availableCores() - 1)
)

# Wrap the parallel execution in with_progress()
bw_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(
      transformation,
      prob_response,
      n_cell,
      mean_pos,
      bw,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | bw = ", bw,
        " | bias = ", bias_uns,
        " | sample_sd = ", sample_perturbation_sd,
        " | condition_sd = ", condition_perturbation_sd,
        " | cluster_sd = ", cluster_perturbation_sd,
        " | bg_rel = ", background_relative_to_response,
        " | n_uns_rel = ", n_cell_uns_relative_to_stim,
        " | sim_id = ", sim_id
      )

      log_msg <- paste0("Running: ", settings_row, "
")
      cat(log_msg, file = path_progress_file, append = TRUE)

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthBsFreq(
            nSample = 5,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 5,
            biasUns = bias_uns,
            bw = bw,
            bwMin = "none",
            bwMax = "none",
            bwFallback = bw,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            samplePerturbationSd = sample_perturbation_sd,
            conditionPerturbationSd = condition_perturbation_sd,
            clusterPerturbationSd = cluster_perturbation_sd,
            backgroundRelativeToResponse = background_relative_to_response,
            ncellUnsRelativeToStim = n_cell_uns_relative_to_stim,
            covEvMin = 2,
            covEvMax = 2,
            tolClust = NULL
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos = mean_pos,
              bw = bw,
              bias_uns = bias_uns,
              sample_perturbation_sd = sample_perturbation_sd,
              condition_perturbation_sd = condition_perturbation_sd,
              cluster_perturbation_sd = cluster_perturbation_sd,
              background_relative_to_response =
                background_relative_to_response,
              n_cell_uns_relative_to_stim =
                n_cell_uns_relative_to_stim,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos,
              bw,
              bias_uns,
              sample_perturbation_sd,
              condition_perturbation_sd,
              cluster_perturbation_sd,
              background_relative_to_response,
              n_cell_uns_relative_to_stim,
              sim_id,
              dplyr::everything()
            )

          cat(
            "Completed: ",
            settings_row,
            "
",
            file = path_progress_file,
            append = TRUE
          )

          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          err_msg <- paste0("Error [", settings_row, "]: ", e$message, "
")
          cat(err_msg, file = path_progress_file, append = TRUE)
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos = mean_pos,
            bw = bw,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response =
              background_relative_to_response,
            n_cell_uns_relative_to_stim =
              n_cell_uns_relative_to_stim,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            method = NA_character_,
            propRespTruth = NA_real_,
            propRespEst = NA_real_,
            threshold = NA_real_,
            thresholdOrigin = NA_character_,
            gateReturnPoint = NA_character_,
            locGenerated = NA,
            locGeneratedDirect = NA,
            locSource = NA_character_,
            locReason = NA_character_,
            detailLevel = NA_character_,
            nCellStim = NA_real_,
            nCellUns = NA_real_,
            nPosStim = NA_integer_,
            nPosUns = NA_integer_
          )
        }
      )

      return(out)
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

# Close the parallel workers when done
future::plan(future::sequential)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "output", "sim", "bw", "freq_bs", "bw_list_raw.rds"
  )
)
```

Collate the results and calculate the metrics of interest.

```{r}
#| label: bw-manual-collate
#| error: true
bw_list_raw <- readRDS(
  file = projr::projr_path_get(
    "output", "sim", "bw", "freq_bs", "bw_list_raw.rds"
  )
)

bw_tbl <- bw_list_raw |>
  purrr::list_rbind()

bw_tbl |>
  dplyr::select(
    any_of(colnames(sim_grid)),
    iter,
    sample,
    ind,
    propBsEst,
    threshold
  ) |>
  dplyr::group_by(sim_id, iter, ind) |>
  dplyr::mutate(
    threshold = setdiff(unique(threshold), NA)[[1]],
    propBsEst = setdiff(unique(propBsEst), NA)[[1]]
  ) |>
  dplyr::ungroup() |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid))),
    iter,
    ind
  ) |>
  dplyr::slice(1) |>
  dplyr::ungroup() |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    across(
      c(threshold, propBsEst),
      list(
        min = ~ min(.x, na.rm = TRUE),
        max = ~ max(.x, na.rm = TRUE),
        mean = ~ mean(.x, na.rm = TRUE),
        median = ~ median(.x, na.rm = TRUE)
      )
    ),
    .groups = "drop"
  )
```

