---
title: Simulate best bandwidth (2a)
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
  sim_size: final
  sim_retry_errors: true
---

This analysis asks how fixed bandwidth affects the estimated background-subtracted response frequency. Matched stimulated and unstimulated samples are simulated across cell counts, response probabilities and marker-expression distributions. The estimates and expression thresholds are compared across bandwidths and unstimulated-sample bias settings.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
analysis_qmd <- "analysis/2a-sim-bw-freq_bs-global.qmd"

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
# On resume, rerun simulations whose saved output recorded an error.
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "freq_bs", "global")
analysis_semantics_version <- "global-bw-freq-v7"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "final" (default) or "draft" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("2a-sim-bw-freq_bs-global")

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source analysis helpers in dependency order after the generic QMD runtime
# helper has been loaded.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

Get the parameters over which we simulate.

So, as with checking what the actual estimated bandwidths are, we are going to have to vary over the number of cells, the probability of response, the mean position of the response, and the transformation applied to the data.

We will need to vary over more response probabilities, however, and of course need to vary over the bandwidths themselves.

Regarding the bandwidths themselves, as the `gamma` scenario is estimating such low bandwidths, we're going to have to start low.

We could allow varying amounts of intra-sample variation between the stimulated and unstimulated conditions.
For now, the stimulated and unstimulated conditions are simulated without any such variation (no condition perturbation).

Let's keep moderate background at 0.2.

Let's include no, low, and high unstimulated-sample bias settings, with transformation-specific magnitudes.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()
transformation_vec <- c("gaussian", "skew", "gamma")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# fix the bandwidth
bandwidth_value_vec <- c(
  # gamma
  0.001, 0.0025, 0.005, 0.0075, 0.01,
  0.0125, 0.015, 0.0175, 0.02, 0.03,
  0.04,
  # mixed
  0.05, 0.1, 0.15,
  # skew and gaussian
  0.2, 0.25, 0.5, 0.75, 1, 1.25, 1.5
)

# StimGate method settings for bandwidth-performance analysis:
# Threshold clustering is intentionally OFF to isolate and
# measure the effect of the bandwidth selector itself without multi-sample
# cluster adjustments modifying thresholds post-local-FDR.
# Share thresholds across clusters of similar samples via stimControl(clusterGates = ).
cluster_gates <- FALSE
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

# can vary these later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.0025,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.01,
  "high", "skew", 0.25
)
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))

# Include quick-mode scenarios before assigning stable IDs and seeds. The
# execution filters below keep the original full and quick scientific settings.
n_cell_stim_grid_vec <- unique(c(n_cell_stim_vec, 5e4))
prob_response_grid_vec <- unique(c(prob_response_vec, 0.02))

```

## Simulation

Method and tuning settings share the same biological-scenario seed, assigned on the full grid before execution filters or chunking. Each replicate has a seed drawn before any method runs, so later replicates remain paired despite method RNG use.

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
We vary the fixed bandwidth over a transformation-dependent range: gamma uses the low end of the common grid (0.001 to 0.15), while Gaussian and skew scenarios use 0.05 to 1.5. We then compare the estimated background-subtracted frequencies with the simulated truth across the other scenario settings.

And we'll do this for varying numbers of cells.

Generate the grid of parameters to simulate.

```{r}
#| label: bw-manual-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_grid_vec,
  n_cell = n_cell_stim_grid_vec,
  bw = bandwidth_value_vec,
  bias_uns = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw,
    bias_uns_setting,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04),
    (transformation == "gamma" & bw <= 0.15) |
      (transformation != "gamma" & bw >= 0.05)
  ) |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::mutate(
    base_scenario_id = dplyr::cur_group_id(),
    sim_seed = as.integer(simulation_seed + base_scenario_id - 1L)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(sim_id = dplyr::row_number())

# sim_id and sim_seed are fixed on the full grid, before dev/quick filters,
# shuffling and chunking, so each row keeps its seed in every run.
sim_grid_full <- sim_grid

# Quick mode keeps cells, response, bandwidth and transformation.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(1e3, 5e3),
      prob_response %in% c(0.002, 0.02),
      bw %in% c(0.05, 0.1, 0.25, 0.5, 0.75)
    )
} else {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% n_cell_stim_vec,
      prob_response %in% prob_response_vec
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      # Select the largest cell count remaining in either execution mode.
      n_cell == max(n_cell),
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw == 0.1,
      bias_uns_setting == "low",
      condition_perturbation_sd == 0
    )
}

# Keep the complete grid for later collation.
sim_grid_all <- sim_grid
if (nrow(sim_grid_all) == 0L) {
  stop("The analysis 2a simulation grid is empty; check the dev/quick filters.")
}
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

# Randomise execution order, then distribute rows evenly across external chunks.
sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(
    .data$.sim_grid_chunk == sim_grid_chunk_index
  ) |>
  dplyr::select(
    -.sim_grid_chunk
  )

sim_grid

print(
  paste0(
    "Performing ",
    nrow(sim_grid),
    " of ",
    nrow(sim_grid_all),
    " simulations for sim_grid chunk ",
    sim_grid_chunk_index,
    " of ",
    sim_grid_n_chunks,
    " using shuffle seed ",
    sim_grid_shuffle_seed
  )
)
```

Fixed settings passed to `.simBandwidthBsFreq()` for every grid row, and the
manifest parameters that canonical results must match. Each scenario simulates
200 samples in final runs and 50 in draft runs (`SIM_SIZE=draft`); draft results
are kept separately and are for iterating, not for reporting.

```{r}
#| label: bw-manual-settings
#| include: false
n_sample_sim <- if (analysis_quick) 1L else c(final = 200L, draft = 50L)[[sim_size]]
scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = 1,
  bwMin = "none",
  bwMax = "none",
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  clusterGates = cluster_gates,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates
)

analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

Rerun one simulation interactively, for example to see why it failed.
Each row is seeded from its biological scenario’s `sim_seed` with a fixed RNG kind inside
`.simBandwidthRunRow()`, which the parallel run also calls, so this
reproduces the random numbers of that row exactly, whatever the chunking,
shuffling or worker scheduling. Call `debugonce()` on any helper first to
step through it.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L

row_target <- sim_grid_full |>
  dplyr::filter(.data$sim_id == sim_id_target)
rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthFreqBsGlobalScenario,
  settings = scenario_settings
)

# Optionally compare with the stored output for this sim_id. Point `dir_run`
# at a staged run directory instead to check a run that was not promoted.
dir_run <- .analysis_cache_dir(
  c(analysis_key, "current"),
  path_root = root_dir,
  create = FALSE
)
path_stored <- .simBandwidthFindSimOutput(dir_run, sim_id_target)
if (length(path_stored) == 1L) {
  all.equal(rerun_res, readRDS(path_stored))
}
```

Run the simulations in parallel, with progress reporting, then validate this
chunk and promote the run once every chunk is complete.

```{r}
#| label: bw-manual-parallel
#| include: false
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthFreqBsGlobalScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    heading = "GLOBAL BANDWIDTH FREQUENCY PROGRESS",
    path_root = root_dir
  )

  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    collate_fn = function(tbl) {
      .simBandwidthFreqBsGlobalCollate(tbl, colnames(sim_grid_all))
    },
    label = "analysis 2a"
  )

  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  # Read-only: plots use the canonical promoted results under current/.
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}
```

## Collation

Read the canonical promoted results for plotting (both run contexts point
at the same `current/` directory).

```{r}
#| label: bw-manual-collate
#| include: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results_raw.rds"),
    required_params = analysis_required_params
  )
  bw_tbl_results_summary <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results_summary.rds"),
    required_params = analysis_required_params
  )
  bias_uns_setting_vec_plot <- bias_uns_settings_tbl |>
    dplyr::distinct(.data$bias_uns_setting) |>
    dplyr::pull(.data$bias_uns_setting)
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

## Results

Okay, now we must plot which bandwidths are best for which scenarios.

```{r}
#| include: false
if (isTRUE(run_plots)) {
  if (requireNamespace("UtilsDataRSV", quietly = TRUE)) {
    sim_grid |>
      UtilsDataRSV::view_cols()
  } else {
    dplyr::glimpse(sim_grid)
  }
}
```

### Median response frequency

Okay, so we varied the transformation, the response probability, the number of cells, the mean position and the bandwidth, as well as the unstimulated-sample bias.

So, let's plot it perhaps how we did it for the bandwidth estimators? We have one line for each bandwidth, and then the x-axis as the number of cells, and the y-axis as the estimated background-subtracted frequency.
There is one figure for each mean position and bias setting, under its own heading.
Columns of panels separate transformations (Gaussian, Skew, Gamma) and rows the response probability.
Okay, and we can plot the median estimated background-subtracted frequency.
Let's leave error bars out for now.

Dashed lines show the simulated truth; compare the curves with those lines.

```{r}
#| label: frequency-median
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
      plot_tbl <- bw_tbl_results_summary |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting_curr,
          bias_uns_setting == !!bias_uns_setting_curr,
          condition_perturbation_sd == 0
        )

      if (nrow(plot_tbl) == 0L) {
        warning(
          "No median estimate results for mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        )
        next
      }

      col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw)
      plot_tbl <- plot_tbl |>
        dplyr::mutate(
          transformation = .analysis_trans_factor(transformation),
          bw_lab = factor(format_bw_lab(bw), levels = names(col_vec_bw_curr))
        )

      p <- ggplot(
        plot_tbl,
        aes(x = n_cell, y = propRespEst_median, color = bw_lab)
      ) +
        geom_line(alpha = 0.75) +
        geom_hline(
          mapping = aes(yintercept = propRespTruth_median),
          linetype = "dashed",
          color = "black"
        ) +
        geom_point(alpha = 0.75) +
        scale_x_log10(
          breaks = sort(unique(plot_tbl$n_cell)),
          labels = .analysis_label_number
        ) +
        scale_y_continuous(labels = .analysis_label_percent) +
        facet_grid(
          rows = vars(prob_response),
          cols = vars(transformation), scales = "free_y",
          labeller = labeller(
            prob_response = .analysis_labeller_percent("Response: ")
          )
        ) +
        .analysis_theme() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        labs(
          x = "Number of cells (log scale)",
          y = "Estimated background-subtracted frequency (median)",
          color = "Bandwidth"
        ) +
        scale_colour_manual(
          values = col_vec_bw_curr
        ) +
        expand_limits(y = c(0))

      path_p <- file.path(
        .analysis_fig_dir(c(fig_key, "est", "median"), root_dir),
        paste0(
          "est_vs_n_cell_mean_pos_setting_",
          mean_pos_setting_curr,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".png"
        )
      )
      .analysis_save_fig(p, path_p, height = 22)
      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 5)
      .analysis_print_fig(p)
    }
  }
}
```

### Sample response frequencies

Okay, now, let's do the same plot using the individual final sample-level frequency estimates rather than the scenario median. We'll treat the number of cells as a categorical variable and place the bandwidths side by side at each cell count.

Each bandwidth's estimates are drawn as a sina plot (points spread sideways in proportion to how many estimates lie at that height). The black point and vertical bar on top mark the median and the interquartile range (25th to 75th percentile). Compare the spread of estimates with the dashed truth lines.

```{r}
#| label: frequency-samples
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
      plot_tbl <- bw_tbl_results_raw |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting_curr,
          bias_uns_setting == !!bias_uns_setting_curr,
          condition_perturbation_sd == 0
        )

      if (nrow(plot_tbl) == 0L) {
        warning(
          "No raw estimate results for mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        )
        next
      }

      col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw)
      n_cell_levels <- sort(unique(plot_tbl$n_cell))
      # Place the bandwidths side by side within each cell count: each
      # transformation's bandwidths share 90% of the space between cell counts.
      plot_tbl <- plot_tbl |>
        dplyr::group_by(transformation) |>
        dplyr::mutate(
          bw_idx = match(bw, sort(unique(bw))),
          n_bw = dplyr::n_distinct(bw)
        ) |>
        dplyr::ungroup() |>
        dplyr::mutate(
          transformation = .analysis_trans_factor(transformation),
          bw_lab = factor(format_bw_lab(bw), levels = names(col_vec_bw_curr)),
          x_pos = match(n_cell, n_cell_levels) +
            0.9 * ((bw_idx - 0.5) / n_bw - 0.5),
          x_group = interaction(n_cell, bw)
        )
      truth_tbl <- plot_tbl |>
        dplyr::distinct(transformation, prob_response, propRespTruth)

      p <- ggplot(
        plot_tbl,
        aes(x = x_pos, y = propRespEst, color = bw_lab, group = x_group)
      ) +
        # Each sina stays within its bandwidth's slot.
        ggforce::geom_sina(
          maxwidth = 0.8 * 0.9 / max(plot_tbl$n_bw),
          size = 0.4,
          alpha = 0.5
        ) +
        stat_summary(
          fun = function(z) stats::median(z, na.rm = TRUE),
          fun.min = function(z) stats::quantile(z, 0.25, na.rm = TRUE),
          fun.max = function(z) stats::quantile(z, 0.75, na.rm = TRUE),
          geom = "pointrange",
          colour = "black",
          size = 0.1,
          linewidth = 0.4
        ) +
        geom_hline(
          data = truth_tbl,
          mapping = aes(yintercept = propRespTruth),
          linetype = "dashed",
          color = "black"
        ) +
        scale_x_continuous(
          breaks = seq_along(n_cell_levels),
          labels = .analysis_label_number(n_cell_levels),
          minor_breaks = NULL
        ) +
        scale_y_continuous(labels = .analysis_label_percent) +
        facet_grid(
          rows = vars(prob_response),
          cols = vars(transformation),
          scales = "free_y",
          labeller = labeller(
            prob_response = .analysis_labeller_percent("Response: ")
          )
        ) +
        .analysis_theme() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        expand_limits(y = 0) +
        labs(
          x = "Number of cells",
          y = "Estimated background-subtracted frequency",
          colour = "Bandwidth"
        ) +
        scale_colour_manual(
          values = col_vec_bw_curr
        ) +
        guides(colour = guide_legend(override.aes = list(size = 2, alpha = 1)))

      path_p <- file.path(
        .analysis_fig_dir(
          # Only condition perturbation 0 is simulated; the folder name is kept.
          c(fig_key, "est", "raw", "condition_perturbation_sd_0"),
          root_dir
        ),
        paste0(
          "actual_freq_vs_n_cell_mean_pos_setting_",
          mean_pos_setting_curr,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".png"
        )
      )
      .analysis_save_fig(p, path_p, height = 22)
      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 5)
      .analysis_print_fig(p)
    }
  }
}
```

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw |>
    dplyr::filter(
      mean_pos_setting == "high",
      bias_uns_setting == "low",
      condition_perturbation_sd == 0
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    dplyr::group_by(
      transformation, n_cell, bw, prob_response
    ) |>
    dplyr::summarise(
      perc_g_005 = sum(propRespEst > 0.005, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_01 = sum(propRespEst > 0.01, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_02 = sum(propRespEst > 0.02, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_05 = sum(propRespEst > 0.05, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_1 = sum(propRespEst > 0.1, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_2 = sum(propRespEst > 0.2, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_5 = sum(propRespEst > 0.5, na.rm = TRUE) / dplyr::n() * 1e2,
      .groups = "drop"
    ) |>
    dplyr::arrange(transformation, n_cell, bw) |>
    print(width = Inf, n = Inf)
}
```

### Relative error averaged over cell counts

The table above is hard to read.
Can't we, for example, compare the average error to the mean?

Relative error is plotted against bandwidth, with one figure per mean position. Scenario-level median, 95th-percentile and maximum errors are averaged equally over cell counts and response probabilities, using low bias; panels separate transformations.

```{r}
#| label: relative-error-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 4.7
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    err_rel_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        mean_pos_setting == mean_pos_setting_curr,
        bias_uns_setting == "low",
        condition_perturbation_sd == 0
      ) |>
      dplyr::group_by(transformation, n_cell, bw, prob_response) |>
      dplyr::summarise(
        err_rel_median = median(abs(propRespEst - propRespTruth) / propRespTruth, na.rm = TRUE),
        err_rel_95 = quantile(abs(propRespEst - propRespTruth) / propRespTruth, probs = 0.95, na.rm = TRUE),
        err_rel_max = max(abs(propRespEst - propRespTruth) / propRespTruth, na.rm = TRUE),
        .groups = "drop"
      )
    if (nrow(err_rel_tbl) == 0L) {
      next
    }
    err_tbl_rel_mean_trans_bw <- err_rel_tbl |>
      dplyr::group_by(transformation, bw) |>
      dplyr::summarise(
        err_rel_median_avg = mean(err_rel_median, na.rm = TRUE),
        err_rel_95_avg = mean(err_rel_95, na.rm = TRUE),
        err_rel_max_avg = mean(err_rel_max, na.rm = TRUE),
        .groups = "drop"
      ) |>
      dplyr::mutate(
        dplyr::across(
          dplyr::starts_with("err_rel_"),
          ~ signif(.x, digits = 3) * 1e2
        )
      ) |>
      dplyr::arrange(transformation, bw)
    plot_tbl_error_rel_mean_trans_bw <- err_tbl_rel_mean_trans_bw |>
      tidyr::pivot_longer(
        cols = dplyr::starts_with("err_rel_"),
        names_to = "err_type",
        values_to = "err_value"
      ) |>
      dplyr::mutate(
        err_type = factor(
          err_type,
          levels = c(
            "err_rel_median_avg",
            "err_rel_95_avg",
            "err_rel_max_avg"
          )
        ),
        transformation = .analysis_trans_factor(transformation),
        bw_fct = factor(
          .analysis_label_number(bw),
          levels = .analysis_label_number(sort(unique(bw)))
        )
      )
    p <- ggplot(
      plot_tbl_error_rel_mean_trans_bw,
      aes(x = bw_fct, y = err_value, color = err_type, group = err_type)
    ) +
      geom_line(alpha = 0.75) +
      geom_point(alpha = 0.75) +
      .analysis_theme() +
      labs(
        x = "Bandwidth",
        y = "Relative error (%)",
        color = "Error type"
      ) +
      expand_limits(y = 0) +
      facet_wrap(~transformation, scales = "free") +
      scale_colour_manual(
        values = c(
          "err_rel_median_avg" = .analysis_stat_colours[["median"]],
          "err_rel_95_avg" = .analysis_stat_colours[["upper"]],
          "err_rel_max_avg" = .analysis_stat_colours[["max"]]
        ),
        labels = c(
          "err_rel_median_avg" = "Median",
          "err_rel_95_avg" = "95th percentile",
          "err_rel_max_avg" = "Maximum"
        )
      ) +
      theme(
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
      )
    path_p <- file.path(
      .analysis_fig_dir(c(fig_key, "relative_error_averaged"), root_dir),
      paste0(
        "avg_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr, ".pdf"
      )
    )
    .analysis_save_fig(p, path_p, height = 12)
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    .analysis_print_fig(p)
  }
}
```

### Relative error at each cell count

The following plots retain each cell count separately. As above, median,
95th-percentile and maximum errors are calculated within each response-probability
scenario before averaging those summaries over response probabilities.

```{r}
#| label: relative-error-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 4.7
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::group_by(mean_pos_setting, transformation, n_cell, bw, prob_response) |>
    dplyr::summarise(
      err_rel_median_avg = median(
        abs(propRespEst - propRespTruth) / propRespTruth, na.rm = TRUE
      ),
      err_rel_95_avg = quantile(
        abs(propRespEst - propRespTruth) / propRespTruth,
        probs = 0.95, na.rm = TRUE
      ),
      err_rel_max_avg = max(
        abs(propRespEst - propRespTruth) / propRespTruth, na.rm = TRUE
      ),
      .groups = "drop"
    ) |>
    dplyr::group_by(mean_pos_setting, transformation, n_cell, bw) |>
    dplyr::summarise(
      dplyr::across(
        dplyr::starts_with("err_rel_"),
        ~ signif(mean(.x, na.rm = TRUE), digits = 3) * 1e2
      ),
      .groups = "drop"
    ) |>
    tidyr::pivot_longer(
      cols = dplyr::starts_with("err_rel_"),
      names_to = "err_type", values_to = "err_value"
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation))

  for (mean_pos_setting_curr in unique(err_rel_by_cell$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell$n_cell))) {
      plot_tbl <- err_rel_by_cell |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::mutate(bw_fct = factor(
          .analysis_label_number(bw),
          levels = .analysis_label_number(sort(unique(bw)))
        ))
      if (nrow(plot_tbl) == 0L) next
      p <- ggplot(
        plot_tbl,
        aes(x = bw_fct, y = err_value, colour = err_type, group = err_type)
      ) +
        geom_line(alpha = 0.75) +
        geom_point(alpha = 0.75) +
        facet_wrap(~transformation, scales = "free") +
        .analysis_theme() +
        expand_limits(y = 0) +
        scale_colour_manual(
          values = c(
            err_rel_median_avg = .analysis_stat_colours[["median"]],
            err_rel_95_avg = .analysis_stat_colours[["upper"]],
            err_rel_max_avg = .analysis_stat_colours[["max"]]
          ),
          labels = c(
            err_rel_median_avg = "Median", err_rel_95_avg = "95th percentile",
            err_rel_max_avg = "Maximum"
          )
        ) +
        labs(
          x = "Bandwidth", y = "Relative error (%)", colour = "Error type"
        ) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "relative_error_by_n_cell",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, ".pdf"
        )
      )
      .analysis_save_fig(p, path_p, height = 12)
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_fig(p)
    }
  }
}
```

### Over- and under-estimates averaged over cell counts

The overall plots above use the size of the error only. Here relative error, (estimate - truth) / truth, is split by direction: over-estimates (teal) sit above zero and under-estimates (brown) below, with darker shades for larger statistics. Within each scenario, the median, 95th-percentile and maximum error sizes are calculated separately for over- and under-estimates; these are then averaged equally over cell counts and response probabilities, using low bias. Line weight shows the share of estimates in that direction. Under-estimates are on a linear scale down to -100% (nothing gated), and over-estimates on a fold scale, so -100% and +100% (twice the truth) are the same distance from zero; both are always shown. Over-estimates larger than +1500% (16 times the truth) are drawn at +1500%, and that tick is then labelled "≥ +1500% (16x)". Panels separate transformations, and there is one figure per mean position.

```{r}
#| label: signed-error-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    err_rel_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        mean_pos_setting == mean_pos_setting_curr,
        bias_uns_setting == "low",
        condition_perturbation_sd == 0
      ) |>
      dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth)
    if (nrow(err_rel_tbl) == 0L) {
      next
    }
    # Summarise each scenario, then average equally over cell counts and
    # response probabilities.
    err_rel_avg_tbl <- err_rel_tbl |>
      .simBandwidthSignedErrorSummary(
        c("transformation", "n_cell", "bw", "prob_response")
      ) |>
      .simBandwidthSignedErrorAverage(c("transformation", "bw"))
    p <- .simBandwidthGlobalSignedErrorPlot(err_rel_avg_tbl)
    path_p <- file.path(
      .analysis_fig_dir(c(fig_key, "signed_error_averaged"), root_dir),
      paste0(
        "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr, ".pdf"
      )
    )
    .analysis_save_fig(p, path_p, height = 14)
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    .analysis_print_fig(p)
  }
}
```

### Over- and under-estimates at each cell count

As above, but keeping each cell count separate and averaging only over response probabilities.

```{r}
#| label: signed-error-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth) |>
    .simBandwidthSignedErrorSummary(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response")
    ) |>
    .simBandwidthSignedErrorAverage(
      c("mean_pos_setting", "transformation", "n_cell", "bw")
    )

  for (mean_pos_setting_curr in unique(err_rel_by_cell$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell$n_cell))) {
      plot_tbl <- err_rel_by_cell |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr)
      if (nrow(plot_tbl) == 0L) next
      p <- .simBandwidthGlobalSignedErrorPlot(plot_tbl)
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "signed_error_by_n_cell",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, ".pdf"
        )
      )
      .analysis_save_fig(p, path_p, height = 14)
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_fig(p)
    }
  }
}
```

### Over- and under-estimates by cell count and response probability

As above, but without averaging over response probabilities: rows of panels
separate response probabilities, so large over-estimates can be traced to
particular (for example, low) response probabilities.

```{r}
#| label: signed-error-by-n-cell-prob
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell_prob <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth) |>
    .simBandwidthSignedErrorSummary(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response")
    )

  for (mean_pos_setting_curr in unique(err_rel_by_cell_prob$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell_prob$n_cell))) {
      plot_tbl <- err_rel_by_cell_prob |>
        dplyr::filter(
          mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr
        )
      if (nrow(plot_tbl) == 0L) next
      p <- .simBandwidthGlobalSignedErrorPlot(plot_tbl, by_prob = TRUE)
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "signed_error_by_n_cell_prob",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, "_by_prob.pdf"
        )
      )
      n_prob <- length(unique(plot_tbl$prob_response))
      .analysis_save_fig(p, path_p, height = min(22, 6 + 3 * n_prob))
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_fig(p)
    }
  }
}
```


```{r}
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw |>
    dplyr::filter(
      mean_pos_setting == "high",
      bias_uns_setting == "low"
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    dplyr::group_by(transformation, n_cell, bw, sample_perturbation_sd, condition_perturbation_sd, cluster_perturbation_sd) |>
    dplyr::summarise(
      perc_g_005 = sum(propRespEst > 0.005, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_01 = sum(propRespEst > 0.01, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_02 = sum(propRespEst > 0.02, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_05 = sum(propRespEst > 0.05, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_1 = sum(propRespEst > 0.1, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_2 = sum(propRespEst > 0.2, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_5 = sum(propRespEst > 0.5, na.rm = TRUE) / dplyr::n() * 1e2,
      .groups = "drop"
    ) |>
    dplyr::arrange(transformation, condition_perturbation_sd, n_cell, bw) |>
    print(width = Inf, n = Inf)
}
```

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw |>
    dplyr::filter(
      mean_pos_setting == "high",
      bias_uns_setting == "low"
    ) |>
    dplyr::filter(prob_response == 0.002, transformation == "gaussian") |>
    dplyr::group_by(n_cell, bw, sample_perturbation_sd, condition_perturbation_sd, cluster_perturbation_sd) |>
    dplyr::summarise(
      perc_g_005 = sum(propRespEst > 0.005, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_01 = sum(propRespEst > 0.01, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_02 = sum(propRespEst > 0.02, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_05 = sum(propRespEst > 0.05, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_1 = sum(propRespEst > 0.1, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_2 = sum(propRespEst > 0.2, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_5 = sum(propRespEst > 0.5, na.rm = TRUE) / dplyr::n() * 1e2,
      .groups = "drop"
    ) |>
    dplyr::arrange(condition_perturbation_sd, sample_perturbation_sd, cluster_perturbation_sd, n_cell, bw)
}
```

### Median expression thresholds

Now we plot the median thresholds themselves, as vertical lines with one
line type per bandwidth, with one figure per mean position and bias setting.
Rows of panels are the response probability and number of cells, and
columns the transformation.

```{r}
#| label: threshold-median
#| results: asis
#| fig-width: 6.3
#| fig-height: 11.8
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting in unique(bw_tbl_results_summary$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting), 4)
    for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
      threshold_tbl_curr <- bw_tbl_results_summary |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting,
          bias_uns_setting == !!bias_uns_setting_curr,
          condition_perturbation_sd == 0
        )

      if (nrow(threshold_tbl_curr) == 0L) {
        warning(
          "No median threshold results for mean_pos_setting = ",
          mean_pos_setting,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        )
        next
      }

      bw_scale <- make_bw_linetype_scale(threshold_tbl_curr$bw)

      threshold_tbl_curr <- threshold_tbl_curr |>
        dplyr::mutate(
          bw_lab = factor(
            format_bw_lab(.data$bw),
            levels = bw_scale$levels
          ),
          transformation = .analysis_trans_factor(.data$transformation)
        )

      p <- ggplot() +
        facet_grid(
          rows = vars(.data$prob_response, .data$n_cell),
          cols = vars(.data$transformation),
          scales = "free_x",
          labeller = labeller(
            prob_response = .analysis_labeller_percent("Response: "),
            n_cell = function(x) {
              paste0("Cells: ", .analysis_label_number(as.numeric(x)))
            }
          )
        ) +
        .analysis_theme() +
        theme(strip.text.y = element_text(angle = 0)) +
        labs(
          x = "Marker expression",
          y = NULL,
          linetype = "Bandwidth"
        ) +
        geom_vline(
          data = threshold_tbl_curr,
          mapping = aes(
            xintercept = threshold_median, linetype = bw_lab,
            group = seq_len(nrow(threshold_tbl_curr))
          ),
          linewidth = 0.45,
          alpha = 0.75
        ) +
        scale_linetype_manual(
          values = bw_scale$values,
          breaks = bw_scale$levels,
          labels = bw_scale$levels
        )
      path_p <- file.path(
        .analysis_fig_dir(c(fig_key, "thresholds", "median"), root_dir),
        paste0(
          "threshold_vs_n_cell_mean_pos_setting_",
          mean_pos_setting,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".png"
        )
      )
      # One row of panels per response probability and cell count.
      .analysis_save_fig(p, path_p, height = 30, allow_tall = TRUE)
      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 5)
      .analysis_print_fig(p)
    }
  }
}
```

### Sample expression thresholds

Okay, now let's do the same, but we will make also one plot for each bandwidth setting, and show the actual thresholds all the samples and iterations.

So, now, we loop not just over mean positions, but also over the bandwidths; headings give the mean position, bandwidth and bias setting.

Solid and dashed lines mark the median and interquartile limits within each scenario.

```{r}
#| label: threshold-samples
#| results: asis
#| fig-width: 6.3
#| fig-height: 11.8
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting in unique(bw_tbl_results_raw$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting), 4)
    for (bw_curr in sort(unique(bw_tbl_results_raw$bw))) {
      if (!is.finite(bw_curr)) {
        next
      }
      .analysis_heading(paste0("Bandwidth: ", .analysis_label_number(bw_curr)), 5)
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        threshold_tbl_curr <- bw_tbl_results_raw |>
          dplyr::filter(
            mean_pos_setting == !!mean_pos_setting,
            bw == !!bw_curr,
            bias_uns_setting == !!bias_uns_setting_curr,
            condition_perturbation_sd == 0
          )

        if (nrow(threshold_tbl_curr) == 0L) {
          warning(
            "No raw threshold results for mean_pos_setting = ",
            mean_pos_setting,
            ", bw = ", bw_curr,
            ", bias_uns_setting = ",
            bias_uns_setting_curr
          )
          next
        }

        threshold_tbl_curr_med_iqr <- threshold_tbl_curr |>
          dplyr::group_by(
            dplyr::pick(dplyr::any_of(colnames(sim_grid)))
          ) |>
          dplyr::summarise(
            threshold_median = stats::median(threshold, na.rm = TRUE),
            threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
            threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
            .groups = "drop"
          ) |>
          tidyr::pivot_longer(
            threshold_median:threshold_iqr_upper,
            names_to = "threshold_stat",
            values_to = "threshold"
          ) |>
          dplyr::mutate(
            threshold_stat = gsub("threshold_", "", threshold_stat),
            threshold_stat = factor(
              threshold_stat,
              levels = c(
                "iqr_lower",
                "median",
                "iqr_upper"
              )
            ),
            transformation = .analysis_trans_factor(transformation)
          )
        threshold_tbl_curr <- threshold_tbl_curr |>
          dplyr::mutate(transformation = .analysis_trans_factor(transformation))
        p <- ggplot() +
          facet_grid(
            rows = vars(.data$prob_response, .data$n_cell),
            cols = vars(.data$transformation),
            scales = "free_x",
            labeller = labeller(
              prob_response = .analysis_labeller_percent("Response: "),
              n_cell = function(x) {
                paste0("Cells: ", .analysis_label_number(as.numeric(x)))
              }
            )
          ) +
          .analysis_theme() +
          theme(strip.text.y = element_text(angle = 0)) +
          labs(
            x = "Marker expression",
            y = NULL
          ) +
          geom_vline(
            data = threshold_tbl_curr,
            mapping = aes(xintercept = threshold, group = seq_len(nrow(threshold_tbl_curr))),
            linewidth = 0.25,
            alpha = 0.25
          ) +
          geom_vline(
            data = threshold_tbl_curr_med_iqr,
            aes(
              xintercept = threshold, linetype = threshold_stat,
              group = seq_len(nrow(threshold_tbl_curr_med_iqr))
            ),
            linewidth = 0.45
          ) +
          scale_linetype_manual(
            values = c(
              "iqr_lower" = "dashed",
              "median" = "solid",
              "iqr_upper" = "dashed"
            ),
            breaks = c("iqr_lower", "median", "iqr_upper"),
            labels = c("IQR lower", "Median", "IQR upper")
          )

        path_p <- file.path(
          .analysis_fig_dir(
            c(
              fig_key, "thresholds", "raw",
              paste0("bias_uns_", bias_uns_setting_curr),
              signif(bw_curr)
            ),
            root_dir
          ),
          paste0(
            "threshold_vs_n_cell_mean_pos_setting_bw_",
            mean_pos_setting,
            "_bw_",
            signif(bw_curr),
            "_bias_uns_",
            bias_uns_setting_curr,
            ".png"
          )
        )
        # One row of panels per response probability and cell count.
        .analysis_save_fig(p, path_p, height = 30, allow_tall = TRUE)
        .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 6)
        .analysis_print_fig(p)
      }
    }
  }
}
```

### Threshold summaries

Okay, let's get a summary table, where for each setting above, for each bandwidth we have the min, max, median, iqr lower, iqr upper, sd, mad, iqr length.

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  summary_tbl <- bw_tbl_results_raw |>
    dplyr::group_by(
      dplyr::pick(dplyr::any_of(colnames(sim_grid))),
      bw
    ) |>
    dplyr::summarise(
      threshold_min = min(threshold, na.rm = TRUE),
      threshold_max = max(threshold, na.rm = TRUE),
      threshold_median = stats::median(threshold, na.rm = TRUE),
      threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
      threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
      threshold_sd = stats::sd(threshold, na.rm = TRUE),
      threshold_mad = stats::mad(threshold, na.rm = TRUE),
      threshold_iqr_length = threshold_iqr_upper - threshold_iqr_lower,
      .groups = "drop"
    )
}
```

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  summary_tbl |>
    dplyr::select(
      transformation, bias_uns_setting, bias_uns,
      mean_pos_setting, bw, threshold_sd
    ) |>
    dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 2))) |>
    dplyr::arrange(threshold_sd) |>
    dplyr::mutate(
      threshold_sd_rank = dplyr::dense_rank(threshold_sd)
    ) |>
    dplyr::group_by(transformation, bias_uns_setting, bias_uns, mean_pos_setting, bw) |>
    dplyr::summarise(
      mean_rank = mean(threshold_sd_rank, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::arrange(bias_uns_setting, mean_rank) |>
    dplyr::filter(!transformation == "gamma") |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    knitr::kable()
}
```
