---
title: Simulate best bandwidth (2a)
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
  show_mcse: "on"
  sim_size: final
  sim_retry_errors: true
---

This analysis asks how fixed bandwidth affects the estimated background-subtracted response frequency. Matched stimulated and unstimulated samples are simulated across cell counts, response probabilities and marker-expression distributions. The estimates and expression thresholds are compared across bandwidths and unstimulated-sample bias settings.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
analysis_qmd <- "analysis/2a-sim-bw-freq_bs-global.qmd"

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
# Monte Carlo error bars on the summary plots (see analysis-mcse.R).
mcse_mode <- .analysis_mcse_mode(.get_qmd_param_env("show_mcse", "SHOW_MCSE", "on"))
show_mcse <- mcse_mode != "off"
# On resume, rerun simulations whose saved output recorded an error.
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "freq_bs", "global")
analysis_semantics_version <- "global-bw-freq-v9"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "final" (default) or "draft" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("2a-sim-bw-freq_bs-global")

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source analysis helpers in dependency order after the generic QMD runtime
# helper has been loaded.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

Get the parameters over which we simulate.

So, as with checking what the actual estimated bandwidths are, we are going to have to vary over the number of cells, the probability of response, the mean position of the response, and the transformation applied to the data.

We will need to vary over more response probabilities, however, and of course need to vary over the bandwidths themselves.

Regarding the bandwidths themselves, as the `gamma` scenario is estimating such low bandwidths, we're going to have to start low.

We could allow varying amounts of intra-sample variation between the stimulated and unstimulated conditions.
For now, the stimulated and unstimulated conditions are simulated without any such variation (no condition perturbation).

Let's keep moderate background at 0.2.

Let's include no, low, and high unstimulated-sample bias settings, with transformation-specific magnitudes.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()
transformation_vec <- c("gaussian", "skew", "gamma")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# fix the bandwidth
bandwidth_value_vec <- c(
  # gamma
  0.001, 0.0025, 0.005, 0.0075, 0.01,
  0.0125, 0.015, 0.0175, 0.02, 0.03,
  0.04,
  # mixed
  0.05, 0.1, 0.15,
  # skew and gaussian
  0.2, 0.25, 0.5, 0.75, 1, 1.25, 1.5
)

# StimGate method settings for bandwidth-performance analysis:
# Threshold clustering is intentionally OFF to isolate and
# measure the effect of the bandwidth selector itself without multi-sample
# cluster adjustments modifying thresholds post-local-FDR.
# Share thresholds across clusters of similar samples via stimControl(clusterGates = ).
cluster_gates <- FALSE
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

# can vary these later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.0025,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.01,
  "high", "skew", 0.25
)
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))

# Include quick-mode scenarios before assigning stable IDs and seeds. The
# execution filters below keep the original full and quick scientific settings.
n_cell_stim_grid_vec <- unique(c(n_cell_stim_vec, 5e4))
prob_response_grid_vec <- unique(c(prob_response_vec, 0.02))

```

## Simulation

Method and tuning settings share the same biological-scenario seed, assigned on the full grid before execution filters or chunking. Each replicate has a seed drawn before any method runs, so later replicates remain paired despite method RNG use.

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
We vary the fixed bandwidth over a transformation-dependent range: gamma uses the low end of the common grid (0.001 to 0.15), while Gaussian and skew scenarios use 0.05 to 1.5. We then compare the estimated background-subtracted frequencies with the simulated truth across the other scenario settings.

And we'll do this for varying numbers of cells.

Generate the grid of parameters to simulate.

```{r}
#| label: bw-manual-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_grid_vec,
  n_cell = n_cell_stim_grid_vec,
  bw = bandwidth_value_vec,
  bias_uns = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw,
    bias_uns_setting,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04),
    (transformation == "gamma" & bw <= 0.15) |
      (transformation != "gamma" & bw >= 0.05)
  ) |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::mutate(
    base_scenario_id = dplyr::cur_group_id(),
    sim_seed = as.integer(simulation_seed + base_scenario_id - 1L)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(sim_id = dplyr::row_number())

# sim_id and sim_seed are fixed on the full grid, before dev/quick filters,
# shuffling and chunking, so each row keeps its seed in every run.
sim_grid_full <- sim_grid

# Quick mode keeps cells, response, bandwidth and transformation.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(1e3, 5e3),
      prob_response %in% c(0.002, 0.02),
      bw %in% c(0.05, 0.1, 0.25, 0.5, 0.75)
    )
} else {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% n_cell_stim_vec,
      prob_response %in% prob_response_vec
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      # Select the largest cell count remaining in either execution mode.
      n_cell == max(n_cell),
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw == 0.1,
      bias_uns_setting == "low",
      condition_perturbation_sd == 0
    )
}

# Keep the complete grid for later collation.
sim_grid_all <- sim_grid
if (nrow(sim_grid_all) == 0L) {
  stop("The analysis 2a simulation grid is empty; check the dev/quick filters.")
}
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

# Randomise execution order, then distribute rows evenly across external chunks.
sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(
    .data$.sim_grid_chunk == sim_grid_chunk_index
  ) |>
  dplyr::select(
    -.sim_grid_chunk
  )

sim_grid

print(
  paste0(
    "Performing ",
    nrow(sim_grid),
    " of ",
    nrow(sim_grid_all),
    " simulations for sim_grid chunk ",
    sim_grid_chunk_index,
    " of ",
    sim_grid_n_chunks,
    " using shuffle seed ",
    sim_grid_shuffle_seed
  )
)
```

Fixed settings passed to `.simBandwidthBsFreq()` for every grid row, and the
manifest parameters that canonical results must match. Each scenario simulates
200 samples in final runs and 50 in draft runs (`SIM_SIZE=draft`); draft results
are kept separately and are for iterating, not for reporting.

```{r}
#| label: bw-manual-settings
#| include: false
n_sample_sim <- if (analysis_quick) 1L else c(final = 200L, draft = 50L)[[sim_size]]
scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = 1,
  bwMin = "none",
  bwMax = "none",
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  clusterGates = cluster_gates,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates
)

analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

Rerun one simulation interactively, for example to see why it failed.
Each row is seeded from its biological scenario’s `sim_seed` with a fixed RNG kind inside
`.simBandwidthRunRow()`, which the parallel run also calls, so this
reproduces the random numbers of that row exactly, whatever the chunking,
shuffling or worker scheduling. Call `debugonce()` on any helper first to
step through it.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L

row_target <- sim_grid_full |>
  dplyr::filter(.data$sim_id == sim_id_target)
rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthFreqBsGlobalScenario,
  settings = scenario_settings
)

# Optionally compare with the stored output for this sim_id. Point `dir_run`
# at a staged run directory instead to check a run that was not promoted.
dir_run <- .analysis_cache_dir(
  c(analysis_key, "current"),
  path_root = root_dir,
  create = FALSE
)
path_stored <- .simBandwidthFindSimOutput(dir_run, sim_id_target)
if (length(path_stored) == 1L) {
  all.equal(rerun_res, readRDS(path_stored))
}
```

Run the simulations in parallel, with progress reporting, then validate this
chunk and promote the run once every chunk is complete.

```{r}
#| label: bw-manual-parallel
#| include: false
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthFreqBsGlobalScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    heading = "GLOBAL BANDWIDTH FREQUENCY PROGRESS",
    path_root = root_dir
  )

  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    collate_fn = function(tbl) {
      .simBandwidthFreqBsGlobalCollate(
        tbl, colnames(sim_grid_all),
        n_sample_expected = scenario_settings$nSample,
        n_iter_expected = scenario_settings$nIter
      )
    },
    label = "analysis 2a"
  )

  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  # Read-only: plots use the canonical promoted results under current/.
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}
```

## Collation

Read the canonical promoted results for plotting (both run contexts point
at the same `current/` directory).

```{r}
#| label: bw-manual-collate
#| include: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results_raw.rds"),
    required_params = analysis_required_params
  )
  bw_tbl_results_summary <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results_summary.rds"),
    required_params = analysis_required_params
  )
  # Persisted raw rows retain failures; plots and errors use valid estimates.
  bw_tbl_results_raw <- bw_tbl_results_raw |>
    dplyr::mutate(
      propRespEst = dplyr::if_else(valid_estimate, propRespEst, NA_real_),
      threshold = dplyr::if_else(valid_estimate, threshold, NA_real_)
    )
  bias_uns_setting_vec_plot <- bias_uns_settings_tbl |>
    dplyr::distinct(.data$bias_uns_setting) |>
    dplyr::pull(.data$bias_uns_setting)
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

## Results

Error and frequency plots use valid final sample estimates only. Companion
coverage tables retain every expected sample: failure fractions use all samples,
and fallback fractions use samples with known `locGenerated` provenance
(including failed estimates). `n_scenario_valid` counts scenarios with at least
one valid estimate; all-failed scenarios contribute no error statistic.

Averaged curves show **Mean of scenario medians**, **Mean of scenario 95th
percentiles** and **Mean of scenario maxima**, rather than pooled statistics.
The companion tables report the number of contributing scenarios alongside
the total selected scenarios.

**Monte Carlo error bars.** Set `show_mcse` or `SHOW_MCSE` to `off` or `on` (the default).
Each render prints one version and saves it under `mcse_off/` or `mcse_on/`,
with the same points and axes. Historical false/true values select
off/on. With error bars on, the summary plots show how much each plotted value could move
if the simulation were repeated. The error bars use only the samples already
simulated. They treat the simulated samples within a scenario as independent
draws, which holds here: the bandwidth and the unstimulated-sample bias are
fixed for each scenario and gates are not clustered across samples. For a
median or other percentile, the bar is a 95% interval taken from the sorted
sample values themselves (it needs no assumption about the shape of the
distribution); it is left out when there are too few samples for such an
interval. For an average over scenarios (cell counts and response
probabilities), the bar is the average plus or minus 1.96 combined standard
errors, treating the scenarios as independent, which holds across biological
scenarios. Maxima have no error bar: the samples say little about how much
larger the maximum could be. Bars are thin, in the colour of their line, and
never extend past the +1500% cap of the relative-error plots.

```{r}
#| include: false
if (isTRUE(run_plots)) {
  if (requireNamespace("UtilsDataRSV", quietly = TRUE)) {
    sim_grid |>
      UtilsDataRSV::view_cols()
  } else {
    dplyr::glimpse(sim_grid)
  }
}
```

### Median response frequency

Okay, so we varied the transformation, the response probability, the number of cells, the mean position and the bandwidth, as well as the unstimulated-sample bias.

So, let's plot it perhaps how we did it for the bandwidth estimators? We have one line for each bandwidth, and then the x-axis as the number of cells, and the y-axis as the estimated background-subtracted frequency.
There is one figure for each mean position and bias setting, under its own heading.
Columns of panels separate transformations (Gaussian, Skew, Gamma) and rows the response probability.
Okay, and we can plot the median estimated background-subtracted frequency.
With `show_mcse` on, each median has a thin 95% error bar from the sorted
sample estimates of that scenario.

Dashed lines show the simulated truth; compare the curves with those lines.

```{r}
#| label: frequency-median
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  # 95% order-statistic interval for each scenario's median estimate.
  freq_median_mcse <- bw_tbl_results_raw |>
    dplyr::group_by(sim_id) |>
    dplyr::reframe(.analysis_mcse_quantile(propRespEst, 0.5)) |>
    dplyr::select(
      sim_id,
      propRespEst_median_lower = lower,
      propRespEst_median_upper = upper
    )
  for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
      plot_tbl <- bw_tbl_results_summary |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting_curr,
          bias_uns_setting == !!bias_uns_setting_curr,
          condition_perturbation_sd == 0
        )

      if (nrow(plot_tbl) == 0L) {
        warning(
          "No median estimate results for mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        )
        next
      }

      col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw)
      plot_tbl <- plot_tbl |>
        dplyr::left_join(freq_median_mcse, by = "sim_id") |>
        dplyr::mutate(
          transformation = .analysis_trans_factor(transformation),
          bw_lab = factor(format_bw_lab(bw), levels = names(col_vec_bw_curr))
        )

      p <- ggplot(
        plot_tbl,
        aes(x = n_cell, y = propRespEst_median, color = bw_lab)
      ) +
        geom_line(alpha = 0.75) +
        (if (isTRUE(show_mcse)) {
          .analysis_mcse_errorbar(
            plot_tbl, "propRespEst_median_lower", "propRespEst_median_upper"
          )
        }) +
        geom_hline(
          mapping = aes(yintercept = propRespTruth_median),
          linetype = "dashed",
          color = "black"
        ) +
        geom_point(alpha = 0.75) +
        scale_x_log10(
          breaks = sort(unique(plot_tbl$n_cell)),
          labels = .analysis_label_number
        ) +
        scale_y_continuous(labels = .analysis_label_percent) +
        facet_grid(
          rows = vars(prob_response),
          cols = vars(transformation), scales = "free_y",
          labeller = labeller(
            prob_response = .analysis_labeller_percent("Response: ")
          )
        ) +
        .analysis_theme() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        labs(
          x = "Number of cells (log scale)",
          y = "Estimated background-subtracted frequency (median)",
          color = "Bandwidth"
        ) +
        scale_colour_manual(
          values = col_vec_bw_curr
        ) +
        expand_limits(y = c(0))

      path_p <- file.path(
        .analysis_fig_dir(c(fig_key, "est", "median"), root_dir),
        paste0(
          "est_vs_n_cell_mean_pos_setting_",
          mean_pos_setting_curr,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".png"
        )
      )
      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 5)
      .analysis_print_save_fig(p, path_p, height = 22, mcse_mode = mcse_mode)
      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, bias_uns_setting == bias_uns_setting_curr))
    }
  }
}
```

### Sample response frequencies

Okay, now, let's do the same plot using the individual final sample-level frequency estimates rather than the scenario median. We'll treat the number of cells as a categorical variable and place the bandwidths side by side at each cell count.

Each bandwidth's estimates are drawn as a sina plot (points spread sideways in proportion to how many estimates lie at that height). The black point and vertical bar on top mark the median and the interquartile range (25th to 75th percentile). Compare the spread of estimates with the dashed truth lines.

```{r}
#| label: frequency-samples
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
      plot_tbl <- bw_tbl_results_raw |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting_curr,
          bias_uns_setting == !!bias_uns_setting_curr,
          condition_perturbation_sd == 0
        )

      if (nrow(plot_tbl) == 0L) {
        warning(
          "No raw estimate results for mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        )
        next
      }

      col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw)
      n_cell_levels <- sort(unique(plot_tbl$n_cell))
      # Place the bandwidths side by side within each cell count: each
      # transformation's bandwidths share 90% of the space between cell counts.
      plot_tbl <- plot_tbl |>
        dplyr::group_by(transformation) |>
        dplyr::mutate(
          bw_idx = match(bw, sort(unique(bw))),
          n_bw = dplyr::n_distinct(bw)
        ) |>
        dplyr::ungroup() |>
        dplyr::mutate(
          transformation = .analysis_trans_factor(transformation),
          bw_lab = factor(format_bw_lab(bw), levels = names(col_vec_bw_curr)),
          x_pos = match(n_cell, n_cell_levels) +
            0.9 * ((bw_idx - 0.5) / n_bw - 0.5),
          x_group = interaction(n_cell, bw)
        )
      truth_tbl <- plot_tbl |>
        dplyr::distinct(transformation, prob_response, propRespTruth)

      p <- ggplot(
        plot_tbl,
        aes(x = x_pos, y = propRespEst, color = bw_lab, group = x_group)
      ) +
        # Each sina stays within its bandwidth's slot.
        ggforce::geom_sina(
          maxwidth = 0.8 * 0.9 / max(plot_tbl$n_bw),
          size = 0.4,
          alpha = 0.5
        ) +
        stat_summary(
          fun = function(z) stats::median(z, na.rm = TRUE),
          fun.min = function(z) stats::quantile(z, 0.25, na.rm = TRUE),
          fun.max = function(z) stats::quantile(z, 0.75, na.rm = TRUE),
          geom = "pointrange",
          colour = "black",
          size = 0.1,
          linewidth = 0.4
        ) +
        geom_hline(
          data = truth_tbl,
          mapping = aes(yintercept = propRespTruth),
          linetype = "dashed",
          color = "black"
        ) +
        scale_x_continuous(
          breaks = seq_along(n_cell_levels),
          labels = .analysis_label_number(n_cell_levels),
          minor_breaks = NULL
        ) +
        scale_y_continuous(labels = .analysis_label_percent) +
        facet_grid(
          rows = vars(prob_response),
          cols = vars(transformation),
          scales = "free_y",
          labeller = labeller(
            prob_response = .analysis_labeller_percent("Response: ")
          )
        ) +
        .analysis_theme() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        expand_limits(y = 0) +
        labs(
          x = "Number of cells",
          y = "Estimated background-subtracted frequency",
          colour = "Bandwidth"
        ) +
        scale_colour_manual(
          values = col_vec_bw_curr
        ) +
        guides(colour = guide_legend(override.aes = list(size = 2, alpha = 1)))

      path_p <- file.path(
        .analysis_fig_dir(
          # Only condition perturbation 0 is simulated; the folder name is kept.
          c(fig_key, "est", "raw", "condition_perturbation_sd_0"),
          root_dir
        ),
        paste0(
          "actual_freq_vs_n_cell_mean_pos_setting_",
          mean_pos_setting_curr,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".png"
        )
      )
      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 5)
      .analysis_print_save_fig(p, path_p, height = 22)
      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, bias_uns_setting == bias_uns_setting_curr))
    }
  }
}
```

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw |>
    dplyr::filter(
      mean_pos_setting == "high",
      bias_uns_setting == "low",
      condition_perturbation_sd == 0
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    dplyr::group_by(
      transformation, n_cell, bw, prob_response
    ) |>
    dplyr::summarise(
      perc_g_005 = sum(propRespEst > 0.005, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_01 = sum(propRespEst > 0.01, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_02 = sum(propRespEst > 0.02, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_05 = sum(propRespEst > 0.05, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_1 = sum(propRespEst > 0.1, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_2 = sum(propRespEst > 0.2, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_5 = sum(propRespEst > 0.5, na.rm = TRUE) / dplyr::n() * 1e2,
      .groups = "drop"
    ) |>
    dplyr::arrange(transformation, n_cell, bw) |>
    print(width = Inf, n = Inf)
}
```

### Relative error averaged over cell counts

The table above is hard to read.
Can't we, for example, compare the average error to the mean?

Relative error is plotted against bandwidth, with one figure per mean position. Scenario-level median, 95th-percentile and maximum errors are averaged equally over cell counts and response probabilities, using low bias; panels separate transformations. With `show_mcse` on, the median and 95th percentile have error bars: each scenario's standard error comes from the width of its order-statistic interval, and the averaged value has plus or minus 1.96 combined standard errors (treating scenarios as independent). The maximum has none.

```{r}
#| label: relative-error-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 4.7
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    err_rel_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        mean_pos_setting == mean_pos_setting_curr,
        bias_uns_setting == "low",
        condition_perturbation_sd == 0
      )
    if (nrow(err_rel_tbl) == 0L) {
      next
    }
    # Scenario statistics, averaged equally over cell counts and response
    # probabilities, in percent.
    err_tbl_rel_mean_trans_bw <- err_rel_tbl |>
      dplyr::mutate(
        abs_rel_error = abs(propRespEst - propRespTruth) / propRespTruth
      ) |>
      .simBandwidthGlobalAbsErrorAverage(
        c("transformation", "n_cell", "bw", "prob_response"),
        c("transformation", "bw")
      ) |>
      dplyr::arrange(transformation, bw)
    plot_tbl_error_rel_mean_trans_bw <- err_tbl_rel_mean_trans_bw |>
      .simBandwidthStatLongBounds(
        c("err_rel_median_avg", "err_rel_95_avg", "err_rel_max_avg"),
        "err_type", "err_value"
      ) |>
      dplyr::mutate(
        err_type = factor(
          err_type,
          levels = c(
            "err_rel_median_avg",
            "err_rel_95_avg",
            "err_rel_max_avg"
          )
        ),
        transformation = .analysis_trans_factor(transformation),
        bw_fct = factor(
          .analysis_label_number(bw),
          levels = .analysis_label_number(sort(unique(bw)))
        )
      )
    p <- ggplot(
      plot_tbl_error_rel_mean_trans_bw,
      aes(x = bw_fct, y = err_value, color = err_type, group = err_type)
    ) +
      geom_line(alpha = 0.75) +
      geom_point(alpha = 0.75) +
      (if (isTRUE(show_mcse)) {
        .analysis_mcse_errorbar(plot_tbl_error_rel_mean_trans_bw, width = 0.3)
      }) +
      .analysis_theme() +
      labs(
        x = "Bandwidth",
        y = "Relative error (%)",
        color = "Error type",
        caption = .simBandwidthScenarioCaption(plot_tbl_error_rel_mean_trans_bw)
      ) +
      expand_limits(y = 0) +
      facet_wrap(~transformation, scales = "free") +
      scale_colour_manual(
        values = c(
          "err_rel_median_avg" = .analysis_stat_colours[["median"]],
          "err_rel_95_avg" = .analysis_stat_colours[["upper"]],
          "err_rel_max_avg" = .analysis_stat_colours[["max"]]
        ),
        labels = c(
          "err_rel_median_avg" = "Mean of scenario medians",
          "err_rel_95_avg" = "Mean of scenario 95th percentiles",
          "err_rel_max_avg" = "Mean of scenario maxima"
        )
      ) +
      theme(
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
      )
    path_p <- file.path(
      .analysis_fig_dir(c(fig_key, "relative_error_averaged"), root_dir),
      paste0(
        "avg_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr, ".pdf"
      )
    )
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    .analysis_print_save_fig(p, path_p, height = 12, mcse_mode = mcse_mode)
    .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr) |>
      dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
  }
}
```

### Relative error at each cell count

The following plots retain each cell count separately. As above, median,
95th-percentile and maximum errors are calculated within each response-probability
scenario before averaging those summaries over response probabilities. Error
bars (with `show_mcse` on) are as above, combining only over response
probabilities.

```{r}
#| label: relative-error-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 4.7
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(
      abs_rel_error = abs(propRespEst - propRespTruth) / propRespTruth
    ) |>
    .simBandwidthGlobalAbsErrorAverage(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response"),
      c("mean_pos_setting", "transformation", "n_cell", "bw")
    ) |>
    .simBandwidthStatLongBounds(
      c("err_rel_median_avg", "err_rel_95_avg", "err_rel_max_avg"),
      "err_type", "err_value"
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation))

  for (mean_pos_setting_curr in unique(err_rel_by_cell$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell$n_cell))) {
      plot_tbl <- err_rel_by_cell |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::mutate(bw_fct = factor(
          .analysis_label_number(bw),
          levels = .analysis_label_number(sort(unique(bw)))
        ))
      if (nrow(plot_tbl) == 0L) next
      p <- ggplot(
        plot_tbl,
        aes(x = bw_fct, y = err_value, colour = err_type, group = err_type)
      ) +
        geom_line(alpha = 0.75) +
        geom_point(alpha = 0.75) +
        (if (isTRUE(show_mcse)) {
          .analysis_mcse_errorbar(plot_tbl, width = 0.3)
        }) +
        facet_wrap(~transformation, scales = "free") +
        .analysis_theme() +
        expand_limits(y = 0) +
        scale_colour_manual(
          values = c(
            err_rel_median_avg = .analysis_stat_colours[["median"]],
            err_rel_95_avg = .analysis_stat_colours[["upper"]],
            err_rel_max_avg = .analysis_stat_colours[["max"]]
          ),
          labels = c(
            err_rel_median_avg = "Mean of scenario medians", err_rel_95_avg = "Mean of scenario 95th percentiles",
            err_rel_max_avg = "Mean of scenario maxima"
          )
        ) +
        labs(
          x = "Bandwidth", y = "Relative error (%)", colour = "Error type",
          caption = .simBandwidthScenarioCaption(plot_tbl)
        ) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "relative_error_by_n_cell",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, ".pdf"
        )
      )
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_save_fig(p, path_p, height = 12, mcse_mode = mcse_mode)
      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
    }
  }
}
```

Each signed-error figure also has an estimate/reference ratio companion in a
sibling `ratio*` folder. The curves and intervals are identical: 1x is exact,
0x is an estimate of zero and 2x is twice the reference. Negative ratios remain
visible; a capped upper tick reads “≥ 16x”.

### Over- and under-estimates averaged over cell counts

The overall plots above use the size of the error only. Here relative error, (estimate - truth) / truth, is split by direction: over-estimates (teal) sit above zero and under-estimates (brown) below, with darker shades for larger statistics. Within each scenario, the median, 95th-percentile and maximum error sizes are calculated separately for over- and under-estimates; these are then averaged equally over cell counts and response probabilities, using low bias. Line weight shows the share of estimates in that direction. Under-estimates are on a linear scale, including errors below -100% for negative background-subtracted estimates, with -100% (0x, an estimate of zero), and over-estimates on a fold scale, so -100% and +100% (twice the truth) are the same distance from zero; both are always shown. Over-estimates larger than +1500% (16 times the truth) are drawn at +1500%, and that tick is then labelled "≥ +1500% (16x)". Panels separate transformations, and there is one figure per mean position.

With `show_mcse` on, the median and 95th percentile of each direction have error bars. Within each scenario they come from the sorted errors in that direction; the number of errors in each direction is itself random, so the bars for a rare direction understate the uncertainty a little (the line weight shows how rare it is). Averaged values have plus or minus 1.96 combined standard errors, treating scenarios as independent. Maxima have none.

```{r}
#| label: signed-error-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    err_rel_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        mean_pos_setting == mean_pos_setting_curr,
        bias_uns_setting == "low",
        condition_perturbation_sd == 0
      ) |>
      dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth)
    if (nrow(err_rel_tbl) == 0L) {
      next
    }
    # Summarise each scenario, then average equally over cell counts and
    # response probabilities.
    err_rel_avg_tbl <- err_rel_tbl |>
      .simBandwidthSignedErrorSummary(
        c("transformation", "n_cell", "bw", "prob_response"),
        mcse = show_mcse
      ) |>
      .simBandwidthSignedErrorAverage(c("transformation", "bw"))
    p <- .simBandwidthGlobalSignedErrorPlot(err_rel_avg_tbl, mcse = show_mcse)
    path_p <- file.path(
      .analysis_fig_dir(c(fig_key, "signed_error_averaged"), root_dir),
      paste0(
        "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr, ".pdf"
      )
    )
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    .analysis_print_save_fig(p, path_p, height = 14, mcse_mode = mcse_mode)
    .simBandwidthPrintRatioTwin(p, path_p, height = 14, mcse_mode = mcse_mode)
    .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr) |>
      dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
  }
}
```

### Percentiles of signed relative error averaged over cell counts

The seven curves summarise the full signed relative error distribution, including zero errors, rather than conditioning on over- or under-estimation. They show the 2.5th, 10th, 25th, 50th, 75th, 90th and 97.5th percentiles together. If either outer percentile lacks an interval at any finite point, the whole figure uses the 5th/95th pair and labels that fallback; unavailable intervals are omitted without dropping points. The median is darkest and thickest; symmetric percentile pairs share progressively lighter, thinner tiers. Percentiles are calculated before applying the signed display scale.

Independent-sample intervals use order statistics within scenarios; equal-weight scenario averages combine their standard errors as sqrt(sum(se^2))/k. These views retain the fixed bandwidth/bias subset and its scientific exclusions.

These figures use low bias and no condition perturbation, averaging scenario percentiles equally over cell counts and response probabilities.

```{r}
#| label: signed-percentiles-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
    err_rel_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        mean_pos_setting == mean_pos_setting_curr,
        bias_uns_setting == "low",
        condition_perturbation_sd == 0
      ) |>
      dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth)
    if (nrow(err_rel_tbl) == 0L) {
      next
    }
    # Summarise each scenario, then average equally over cell counts and
    # response probabilities.
    err_rel_avg_tbl <- err_rel_tbl |>
      .simBandwidthSignedPercentileSummary(
        c("transformation", "n_cell", "bw", "prob_response"),
        mcse = show_mcse
      ) |>
      .simBandwidthSignedPercentileAverage(c("transformation", "bw"))
    p <- .simBandwidthSignedPercentilePlot(err_rel_avg_tbl, mcse = show_mcse)
    path_p <- file.path(
      .analysis_fig_dir(c(fig_key, "signed_percentiles_averaged"), root_dir),
      paste0(
        "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr, ".pdf"
      )
    )
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    .analysis_print_save_fig(p, path_p, height = 14, mcse_mode = mcse_mode)

    .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr) |>
      dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
  }
}
```

### Over- and under-estimates at each cell count

As above, but keeping each cell count separate and averaging only over response probabilities.

```{r}
#| label: signed-error-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth) |>
    .simBandwidthSignedErrorSummary(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response"),
      mcse = show_mcse
    ) |>
    .simBandwidthSignedErrorAverage(
      c("mean_pos_setting", "transformation", "n_cell", "bw")
    )

  for (mean_pos_setting_curr in unique(err_rel_by_cell$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell$n_cell))) {
      plot_tbl <- err_rel_by_cell |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr)
      if (nrow(plot_tbl) == 0L) next
      p <- .simBandwidthGlobalSignedErrorPlot(plot_tbl, mcse = show_mcse)
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "signed_error_by_n_cell",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, ".pdf"
        )
      )
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_save_fig(p, path_p, height = 14, mcse_mode = mcse_mode)
      .simBandwidthPrintRatioTwin(p, path_p, height = 14, mcse_mode = mcse_mode)
      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
    }
  }
}
```

### Percentiles of signed relative error at each cell count

The seven curves summarise the full signed relative error distribution, including zero errors, rather than conditioning on over- or under-estimation. They show the 2.5th, 10th, 25th, 50th, 75th, 90th and 97.5th percentiles together. If either outer percentile lacks an interval at any finite point, the whole figure uses the 5th/95th pair and labels that fallback; unavailable intervals are omitted without dropping points. The median is darkest and thickest; symmetric percentile pairs share progressively lighter, thinner tiers. Percentiles are calculated before applying the signed display scale.

Independent-sample intervals use order statistics within scenarios; equal-weight scenario averages combine their standard errors as sqrt(sum(se^2))/k. These views retain the fixed bandwidth/bias subset and its scientific exclusions.

These figures use low bias and no condition perturbation, averaging scenario percentiles equally over response probabilities at each cell count.

```{r}
#| label: signed-percentiles-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth) |>
    .simBandwidthSignedPercentileSummary(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response"),
      mcse = show_mcse
    ) |>
    .simBandwidthSignedPercentileAverage(
      c("mean_pos_setting", "transformation", "n_cell", "bw")
    )

  for (mean_pos_setting_curr in unique(err_rel_by_cell$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell$n_cell))) {
      plot_tbl <- err_rel_by_cell |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr)
      if (nrow(plot_tbl) == 0L) next
      p <- .simBandwidthSignedPercentilePlot(plot_tbl, mcse = show_mcse)
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "signed_percentiles_by_n_cell",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, ".pdf"
        )
      )
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_save_fig(p, path_p, height = 14, mcse_mode = mcse_mode)

      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
    }
  }
}
```

### Over- and under-estimates by cell count and response probability

As above, but without averaging over response probabilities: rows of panels
separate response probabilities, so large over-estimates can be traced to
particular (for example, low) response probabilities. Error bars are the
per-scenario order-statistic intervals.

```{r}
#| label: signed-error-by-n-cell-prob
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell_prob <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth) |>
    .simBandwidthSignedErrorSummary(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response"),
      mcse = show_mcse
    )

  for (mean_pos_setting_curr in unique(err_rel_by_cell_prob$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell_prob$n_cell))) {
      plot_tbl <- err_rel_by_cell_prob |>
        dplyr::filter(
          mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr
        )
      if (nrow(plot_tbl) == 0L) next
      p <- .simBandwidthGlobalSignedErrorPlot(
        plot_tbl,
        by_prob = TRUE, mcse = show_mcse
      )
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "signed_error_by_n_cell_prob",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, "_by_prob.pdf"
        )
      )
      n_prob <- length(unique(plot_tbl$prob_response))
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_save_fig(p, path_p, height = min(22, 6 + 3 * n_prob), mcse_mode = mcse_mode)
      .simBandwidthPrintRatioTwin(p, path_p, height = min(22, 6 + 3 * n_prob), mcse_mode = mcse_mode)
      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
    }
  }
}
```

### Percentiles of signed relative error by cell count and response probability

The seven curves summarise the full signed relative error distribution, including zero errors, rather than conditioning on over- or under-estimation. They show the 2.5th, 10th, 25th, 50th, 75th, 90th and 97.5th percentiles together. If either outer percentile lacks an interval at any finite point, the whole figure uses the 5th/95th pair and labels that fallback; unavailable intervals are omitted without dropping points. The median is darkest and thickest; symmetric percentile pairs share progressively lighter, thinner tiers. Percentiles are calculated before applying the signed display scale.

Independent-sample intervals use order statistics within scenarios; equal-weight scenario averages combine their standard errors as sqrt(sum(se^2))/k. These views retain the fixed bandwidth/bias subset and its scientific exclusions.

These figures use low bias and no condition perturbation; panels keep response probabilities separate at each cell count.

```{r}
#| label: signed-percentiles-by-n-cell-prob
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  err_rel_by_cell_prob <- bw_tbl_results_raw |>
    dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0) |>
    dplyr::mutate(rel_error = (propRespEst - propRespTruth) / propRespTruth) |>
    .simBandwidthSignedPercentileSummary(
      c("mean_pos_setting", "transformation", "n_cell", "bw", "prob_response"),
      mcse = show_mcse
    )

  for (mean_pos_setting_curr in unique(err_rel_by_cell_prob$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 4)
    for (n_cell_curr in sort(unique(err_rel_by_cell_prob$n_cell))) {
      plot_tbl <- err_rel_by_cell_prob |>
        dplyr::filter(
          mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr
        )
      if (nrow(plot_tbl) == 0L) next
      p <- .simBandwidthSignedPercentilePlot(
        plot_tbl,
        by_prob = TRUE, mcse = show_mcse
      )
      path_p <- file.path(
        .analysis_fig_dir(
          c(
            fig_key, "signed_percentiles_by_n_cell_prob",
            paste0("mean_pos_setting_", mean_pos_setting_curr)
          ),
          root_dir
        ),
        paste0(
          "avg_signed_rel_error_vs_bw_mean_pos_setting_", mean_pos_setting_curr,
          "_n_cell_", n_cell_curr, "_by_prob.pdf"
        )
      )
      n_prob <- length(unique(plot_tbl$prob_response))
      .analysis_heading(paste0("Cells: ", .analysis_label_number(n_cell_curr)), 5)
      .analysis_print_save_fig(p, path_p, height = min(22, 6 + 3 * n_prob), mcse_mode = mcse_mode)

      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == mean_pos_setting_curr, n_cell == n_cell_curr) |>
        dplyr::filter(bias_uns_setting == "low", condition_perturbation_sd == 0))
    }
  }
}
```


```{r}
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw |>
    dplyr::filter(
      mean_pos_setting == "high",
      bias_uns_setting == "low"
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    dplyr::group_by(transformation, n_cell, bw, sample_perturbation_sd, condition_perturbation_sd, cluster_perturbation_sd) |>
    dplyr::summarise(
      perc_g_005 = sum(propRespEst > 0.005, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_01 = sum(propRespEst > 0.01, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_02 = sum(propRespEst > 0.02, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_05 = sum(propRespEst > 0.05, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_1 = sum(propRespEst > 0.1, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_2 = sum(propRespEst > 0.2, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_5 = sum(propRespEst > 0.5, na.rm = TRUE) / dplyr::n() * 1e2,
      .groups = "drop"
    ) |>
    dplyr::arrange(transformation, condition_perturbation_sd, n_cell, bw) |>
    print(width = Inf, n = Inf)
}
```

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw |>
    dplyr::filter(
      mean_pos_setting == "high",
      bias_uns_setting == "low"
    ) |>
    dplyr::filter(prob_response == 0.002, transformation == "gaussian") |>
    dplyr::group_by(n_cell, bw, sample_perturbation_sd, condition_perturbation_sd, cluster_perturbation_sd) |>
    dplyr::summarise(
      perc_g_005 = sum(propRespEst > 0.005, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_01 = sum(propRespEst > 0.01, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_02 = sum(propRespEst > 0.02, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_05 = sum(propRespEst > 0.05, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_1 = sum(propRespEst > 0.1, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_2 = sum(propRespEst > 0.2, na.rm = TRUE) / dplyr::n() * 1e2,
      perc_g_5 = sum(propRespEst > 0.5, na.rm = TRUE) / dplyr::n() * 1e2,
      .groups = "drop"
    ) |>
    dplyr::arrange(condition_perturbation_sd, sample_perturbation_sd, cluster_perturbation_sd, n_cell, bw)
}
```

### Reference densities for threshold plots

The “With densities” companions below show the true generating distributions
approximated by a large reference simulation (100,000 cells per tube), rather
than the individual simulated samples. The blue curve uses only the
unstimulated tube's negative component; the vermillion curve uses all stimulated
cells. The reference uses each panel's transformation, mean position, response
probability and perturbation settings, with the same covariance and exact
mixture settings as the scenarios. Unstimulated background probability is the
response probability multiplied by `background_relative_to_response`, as in
the simulations. Curves are cached in memory once per distinct biological
setting and reused across cell counts, bandwidths and bias settings. They do
not include the gating bias applied to the unstimulated sample.

```{r}
#| label: threshold-reference-densities
#| include: false
if (isTRUE(run_plots)) {
  threshold_reference_densities <- .simBandwidthThresholdDensities(
    sim_grid |> dplyr::filter(condition_perturbation_sd == 0),
    scenario_settings
  )
}
```

### Median expression thresholds

Now we plot the median thresholds themselves, as vertical lines with one
line type per bandwidth, with one figure per mean position and bias setting.
Rows of panels are the response probability and number of cells, and
columns the transformation.

```{r}
#| label: threshold-median
#| results: asis
#| fig-width: 6.3
#| fig-height: 11.8
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting in unique(bw_tbl_results_summary$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting), 4)
    for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
      threshold_tbl_curr <- bw_tbl_results_summary |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting,
          bias_uns_setting == !!bias_uns_setting_curr,
          condition_perturbation_sd == 0
        )

      if (nrow(threshold_tbl_curr) == 0L) {
        warning(
          "No median threshold results for mean_pos_setting = ",
          mean_pos_setting,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        )
        next
      }

      bw_scale <- make_bw_linetype_scale(threshold_tbl_curr$bw)

      threshold_tbl_curr <- threshold_tbl_curr |>
        dplyr::mutate(
          bw_lab = factor(
            format_bw_lab(.data$bw),
            levels = bw_scale$levels
          ),
          transformation = .analysis_trans_factor(.data$transformation)
        )

      p <- ggplot() +
        facet_grid(
          rows = vars(.data$prob_response, .data$n_cell),
          cols = vars(.data$transformation),
          scales = "free_x",
          labeller = labeller(
            prob_response = .analysis_labeller_percent("Response: "),
            n_cell = function(x) {
              paste0("Cells: ", .analysis_label_number(as.numeric(x)))
            }
          )
        ) +
        .analysis_theme() +
        theme(strip.text.y = element_text(angle = 0)) +
        labs(
          x = "Marker expression",
          y = NULL,
          linetype = "Bandwidth"
        ) +
        geom_vline(
          data = threshold_tbl_curr,
          mapping = aes(
            xintercept = threshold_median, linetype = bw_lab,
            group = seq_len(nrow(threshold_tbl_curr))
          ),
          linewidth = 0.45,
          alpha = 0.75
        ) +
        scale_linetype_manual(
          values = bw_scale$values,
          breaks = bw_scale$levels,
          labels = bw_scale$levels
        )
      path_p <- file.path(
        .analysis_fig_dir(c(fig_key, "thresholds", "median"), root_dir),
        paste0(
          "threshold_vs_n_cell_mean_pos_setting_",
          mean_pos_setting,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".png"
        )
      )
      # One row of panels per response probability and cell count.
      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 5)
      .analysis_print_save_fig(p, path_p, height = 30, allow_tall = TRUE)
      p_density <- .simBandwidthThresholdDensityPlot(
        p, threshold_tbl_curr, threshold_reference_densities
      )
      path_density <- file.path(
        .analysis_fig_dir(c(fig_key, "thresholds_density", "median"), root_dir),
        basename(path_p)
      )
      .analysis_heading("With densities", 6)
      .analysis_print_save_fig(p_density, path_density, height = 30, allow_tall = TRUE)
      .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
        dplyr::filter(mean_pos_setting == !!mean_pos_setting, bias_uns_setting == bias_uns_setting_curr))
    }
  }
}
```

### Sample expression thresholds

Okay, now let's do the same, but we will make also one plot for each bandwidth setting, and show the actual thresholds all the samples and iterations.

So, now, we loop not just over mean positions, but also over the bandwidths; headings give the mean position, bandwidth and bias setting.

Solid and dashed lines mark the median and interquartile limits within each scenario.

```{r}
#| label: threshold-samples
#| results: asis
#| fig-width: 6.3
#| fig-height: 11.8
#| echo: false
if (isTRUE(run_plots)) {
  for (mean_pos_setting in unique(bw_tbl_results_raw$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting), 4)
    for (bw_curr in sort(unique(bw_tbl_results_raw$bw))) {
      if (!is.finite(bw_curr)) {
        next
      }
      .analysis_heading(paste0("Bandwidth: ", .analysis_label_number(bw_curr)), 5)
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        threshold_tbl_curr <- bw_tbl_results_raw |>
          dplyr::filter(
            mean_pos_setting == !!mean_pos_setting,
            bw == !!bw_curr,
            bias_uns_setting == !!bias_uns_setting_curr,
            condition_perturbation_sd == 0
          )

        if (nrow(threshold_tbl_curr) == 0L) {
          warning(
            "No raw threshold results for mean_pos_setting = ",
            mean_pos_setting,
            ", bw = ", bw_curr,
            ", bias_uns_setting = ",
            bias_uns_setting_curr
          )
          next
        }

        threshold_tbl_curr_med_iqr <- threshold_tbl_curr |>
          dplyr::group_by(
            dplyr::pick(dplyr::any_of(colnames(sim_grid)))
          ) |>
          dplyr::summarise(
            threshold_median = stats::median(threshold, na.rm = TRUE),
            threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
            threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
            .groups = "drop"
          ) |>
          tidyr::pivot_longer(
            threshold_median:threshold_iqr_upper,
            names_to = "threshold_stat",
            values_to = "threshold"
          ) |>
          dplyr::mutate(
            threshold_stat = gsub("threshold_", "", threshold_stat),
            threshold_stat = factor(
              threshold_stat,
              levels = c(
                "iqr_lower",
                "median",
                "iqr_upper"
              )
            ),
            transformation = .analysis_trans_factor(transformation)
          )
        threshold_tbl_curr <- threshold_tbl_curr |>
          dplyr::mutate(transformation = .analysis_trans_factor(transformation))
        p <- ggplot() +
          facet_grid(
            rows = vars(.data$prob_response, .data$n_cell),
            cols = vars(.data$transformation),
            scales = "free_x",
            labeller = labeller(
              prob_response = .analysis_labeller_percent("Response: "),
              n_cell = function(x) {
                paste0("Cells: ", .analysis_label_number(as.numeric(x)))
              }
            )
          ) +
          .analysis_theme() +
          theme(strip.text.y = element_text(angle = 0)) +
          labs(
            x = "Marker expression",
            y = NULL
          ) +
          geom_vline(
            data = threshold_tbl_curr,
            mapping = aes(xintercept = threshold, group = seq_len(nrow(threshold_tbl_curr))),
            linewidth = 0.25,
            alpha = 0.25
          ) +
          geom_vline(
            data = threshold_tbl_curr_med_iqr,
            aes(
              xintercept = threshold, linetype = threshold_stat,
              group = seq_len(nrow(threshold_tbl_curr_med_iqr))
            ),
            linewidth = 0.45
          ) +
          scale_linetype_manual(
            values = c(
              "iqr_lower" = "dashed",
              "median" = "solid",
              "iqr_upper" = "dashed"
            ),
            breaks = c("iqr_lower", "median", "iqr_upper"),
            labels = c("IQR lower", "Median", "IQR upper")
          )

        path_p <- file.path(
          .analysis_fig_dir(
            c(
              fig_key, "thresholds", "raw",
              paste0("bias_uns_", bias_uns_setting_curr),
              signif(bw_curr)
            ),
            root_dir
          ),
          paste0(
            "threshold_vs_n_cell_mean_pos_setting_bw_",
            mean_pos_setting,
            "_bw_",
            signif(bw_curr),
            "_bias_uns_",
            bias_uns_setting_curr,
            ".png"
          )
        )
        # One row of panels per response probability and cell count.
        .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 6)
        .analysis_print_save_fig(p, path_p, height = 30, allow_tall = TRUE)
        p_density <- .simBandwidthThresholdDensityPlot(
          p, threshold_tbl_curr, threshold_reference_densities
        )
        path_density <- file.path(
          .analysis_fig_dir(
            c(fig_key, "thresholds_density", "raw",
              paste0("bias_uns_", bias_uns_setting_curr), signif(bw_curr)),
            root_dir
          ),
          basename(path_p)
        )
        .analysis_heading("With densities", 6)
        .analysis_print_save_fig(p_density, path_density, height = 30, allow_tall = TRUE)
        .simBandwidthPrintCoverage(p, bw_tbl_results_summary |>
          dplyr::filter(mean_pos_setting == !!mean_pos_setting, bias_uns_setting == bias_uns_setting_curr))
      }
    }
  }
}
```

### Threshold summaries

Okay, let's get a summary table, where for each setting above, for each bandwidth we have the min, max, median, iqr lower, iqr upper, sd, mad, iqr length.

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  summary_tbl <- bw_tbl_results_raw |>
    dplyr::group_by(
      dplyr::pick(dplyr::any_of(colnames(sim_grid))),
      bw
    ) |>
    dplyr::summarise(
      threshold_min = min(threshold, na.rm = TRUE),
      threshold_max = max(threshold, na.rm = TRUE),
      threshold_median = stats::median(threshold, na.rm = TRUE),
      threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
      threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
      threshold_sd = stats::sd(threshold, na.rm = TRUE),
      threshold_mad = stats::mad(threshold, na.rm = TRUE),
      threshold_iqr_length = threshold_iqr_upper - threshold_iqr_lower,
      .groups = "drop"
    )
}
```

```{r}
#| echo: false
if (isTRUE(run_plots)) {
  summary_tbl |>
    dplyr::select(
      transformation, bias_uns_setting, bias_uns,
      mean_pos_setting, bw, threshold_sd
    ) |>
    dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 2))) |>
    dplyr::arrange(threshold_sd) |>
    dplyr::mutate(
      threshold_sd_rank = dplyr::dense_rank(threshold_sd)
    ) |>
    dplyr::group_by(transformation, bias_uns_setting, bias_uns, mean_pos_setting, bw) |>
    dplyr::summarise(
      mean_rank = mean(threshold_sd_rank, na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::arrange(bias_uns_setting, mean_rank) |>
    dplyr::filter(!transformation == "gamma") |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    knitr::kable()
}
```
