---
title: Tune unstimulated-sample bias (2b)
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20261003
  run_simulations: true
  run_plots: false
  sim_size: final
  sim_retry_errors: true
---

Analysis 2a identifies a plausible bandwidth range while keeping `bias_uns`
deliberately simple. This analysis asks the complementary question: across
those plausible bandwidths, how should `bias_uns` be set when the stimulated
negative population is displaced relative to its matched unstimulated
population?

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
analysis_qmd <- "analysis/2b-sim-bias_uns-freq_bs.qmd"

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20261003L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bias_uns", "freq_bs")
analysis_semantics_version <- "bias-uns-freq-v6"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "final" (default) or "draft" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("2b-sim-bias_uns-freq_bs")

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

The batch mismatch follows Analysis 8 and is applied only to the stimulated
negative/background component. Random sample, condition and cluster
perturbations are disabled so that the controlled mismatch is the source of
stimulated-versus-unstimulated disagreement.

```{r}
#| label: bias-uns-settings
n_cell_vec <- c(1e4, 5e4)
prob_response_vec <- c(0.002, 0.05)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

bandwidth_settings_tbl <- tibble::tribble(
  ~transformation, ~bw,
  "gaussian", 0.05,
  "gaussian", 0.10,
  "gaussian", 0.25,
  "gaussian", 0.50,
  "skew",     0.05,
  "skew",     0.10,
  "skew",     0.25,
  "skew",     0.50,
  "gamma",    0.001,
  "gamma",    0.0025,
  "gamma",    0.005,
  "gamma",    0.010
)

bias_uns_settings_tbl <- dplyr::bind_rows(
  tibble::tibble(
    bias_uns_basis = "bandwidth",
    bias_uns_multiplier = c(
      0, 0.1, 0.25, 0.33, 0.5, 0.75, 1, 1.25, 1.5, 2
    )
  ),
  tibble::tibble(
    bias_uns_basis = "negative_width",
    bias_uns_multiplier = c(0.1, 0.25, 0.5, 0.75)
  )
) |>
  dplyr::mutate(
    bias_uns_rule = paste0(
      as.character(.data$bias_uns_multiplier),
      " x ",
      dplyr::if_else(
        .data$bias_uns_basis == "bandwidth",
        "bandwidth",
        "negative width"
      )
    )
  )

mean_shift_gaussian_skew <- c(0, 0.025, 0.05, 0.2)
mean_shift_gamma <- c(0, 0.005, 0.01, 0.05)

mean_shift_tbl <- dplyr::bind_rows(
  tidyr::expand_grid(
    transformation = c("gaussian", "skew"),
    mismatch_val = mean_shift_gaussian_skew
  ),
  tidyr::expand_grid(
    transformation = "gamma",
    mismatch_val = mean_shift_gamma
  )
) |>
  dplyr::mutate(
    mismatch_type = "mean_shift",
    mismatch_label = paste0("mean shift ", .data$mismatch_val),
    stim_mean_shift = .data$mismatch_val,
    stim_mean_shift_clusters = "gn",
    stim_sd_multiplier = 1,
    stim_sd_multiplier_clusters = NA_character_,
    sd_increase = 0
  )

sd_inflation_tbl <- tibble::tibble(
  transformation = c("gaussian", "gamma", "skew"),
  mismatch_val = 0.10,
  mismatch_type = "sd_inflation",
  mismatch_label = "SD inflation 10%",
  stim_mean_shift = 0,
  stim_mean_shift_clusters = NA_character_,
  stim_sd_multiplier = 1.10,
  stim_sd_multiplier_clusters = "gn",
  sd_increase = 0.10
)

mismatch_settings_tbl <- dplyr::bind_rows(
  mean_shift_tbl,
  sd_inflation_tbl
)

background_relative_to_response <- 0.2
n_cell_uns_relative_to_stim <- 1

# Share thresholds across clusters of similar samples via stimControl(clusterGates = ).
cluster_gates <- FALSE
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
```

For the negative-width rules, the width is calculated from the pooled
unstimulated cells in the simulation row, using the same fixed KDE bandwidth
being evaluated. Starting at the dominant negative peak, the right endpoint is
the first point at which the density reaches 15% of the peak height or the next
antimode, whichever comes first.

## Simulation

Method and tuning settings share the same biological-scenario seed, assigned on the full grid before execution filters or chunking. Each replicate has a seed drawn before any method runs, so later replicates remain paired despite method RNG use.

```{r}
#| label: bias-uns-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  n_cell = n_cell_vec,
  prob_response = prob_response_vec,
  bias_uns_settings_tbl
) |>
  dplyr::inner_join(
    bandwidth_settings_tbl,
    by = "transformation"
  ) |>
  dplyr::inner_join(
    mismatch_settings_tbl,
    by = "transformation"
  ) |>
  dplyr::mutate(
    background_relative_to_response = background_relative_to_response,
    n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim
  ) |>
  dplyr::arrange(
    .data$transformation,
    .data$n_cell,
    .data$prob_response,
    .data$mean_pos_setting,
    .data$mismatch_type,
    .data$mismatch_val,
    .data$bw,
    .data$bias_uns_basis,
    .data$bias_uns_multiplier
  )

base_scenario_tbl <- sim_grid |>
  dplyr::distinct(
    .data$transformation,
    .data$n_cell,
    .data$prob_response,
    .data$mean_pos_setting,
    .data$mean_pos
  ) |>
  dplyr::mutate(base_scenario_id = dplyr::row_number())

sim_grid <- sim_grid |>
  dplyr::left_join(
    base_scenario_tbl,
    by = c(
      "transformation",
      "n_cell",
      "prob_response",
      "mean_pos_setting",
      "mean_pos"
    )
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number(),
    # Bandwidth, bias rule and deterministic mismatch settings all use the
    # same underlying simulated data within each biological scenario.
    sim_seed = as.integer(simulation_seed + .data$base_scenario_id - 1L)
  )

if (nrow(sim_grid) != 6720L) {
  stop(
    "Expected 6,720 Analysis 2b grid rows, found ",
    nrow(sim_grid),
    "."
  )
}

sim_grid_full <- sim_grid

# Quick mode keeps cells, mismatch, bandwidth, bias multiplier and bias basis.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(1e4, 5e4),
      prob_response == 0.002,
      mean_pos_setting == "high",
      mismatch_label %in% c("mean shift 0", "SD inflation 10%"),
      bw %in% c(0.05, 0.1, 0.0025, 0.005),
      bias_uns_multiplier %in% c(0, 0.25, 0.5)
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      transformation == "gaussian",
      n_cell == 1e4,
      prob_response == 0.002,
      mean_pos_setting == "high",
      mismatch_label == "mean shift 0.05",
      bw == 0.1,
      (
        (bias_uns_basis == "bandwidth" & bias_uns_multiplier == 0.25) |
          (bias_uns_basis == "negative_width" & bias_uns_multiplier == 0.5)
      )
    )
}

sim_grid_all <- sim_grid
if (nrow(sim_grid_all) == 0L) {
  stop("The Analysis 2b simulation grid is empty; check the dev/quick filters.")
}

analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
  dplyr::select(-.sim_grid_chunk)

print(
  paste0(
    "Performing ",
    nrow(sim_grid),
    " of ",
    nrow(sim_grid_all),
    " Analysis 2b simulations for chunk ",
    sim_grid_chunk_index,
    " of ",
    sim_grid_n_chunks
  )
)
```

Each scenario simulates 25 samples in final runs and 10 in draft runs
(`SIM_SIZE=draft`); draft results are kept separately and are for iterating,
not for reporting.

```{r}
#| label: bias-uns-run-settings
#| include: false
n_sample_sim <- if (analysis_quick) 1L else c(final = 25L, draft = 10L)[[sim_size]]
scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = 1,
  biasUnsWidthHeightFrac = 0.15,
  bwMin = "none",
  bwMax = "none",
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  clusterGates = cluster_gates,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates
)

analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

A single row can be rerun interactively with exactly the same random numbers
used by the parallel run.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L
row_target <- sim_grid_full |>
  dplyr::filter(.data$sim_id == sim_id_target)

rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthBiasUnsScenario,
  settings = scenario_settings
)
```

```{r}
#| label: bias-uns-parallel
#| include: false
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthBiasUnsScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    heading = "BIAS_UNS FREQUENCY PROGRESS",
    path_root = root_dir
  )

  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    collate_fn = function(tbl) {
      .simBandwidthBiasUnsCollate(
        tbl, colnames(sim_grid_all), n_sample_expected = scenario_settings$nSample
      )
    },
    label = "analysis 2b"
  )

  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}
```

## Collation

```{r}
#| label: bias-uns-collate
#| include: false
if (isTRUE(run_plots)) {
  bias_uns_results_raw <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bias_uns_results_raw.rds"),
    required_params = analysis_required_params
  )
  bias_uns_results_summary <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bias_uns_results_summary.rds"),
    required_params = analysis_required_params
  )
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

## Results

The primary summaries are relative error and the estimated
background-subtracted frequencies themselves. Plot each response probability
separately so that every remaining varying scenario dimension is either
faceted or included in the grouping. Invalid final estimates remain in the raw
results; summaries record `n_sample`, `n_valid`, `n_failed` and `failure_fraction`.
Frequency and error summaries use valid estimates only; settings with no valid
estimates remain in the summary with missing frequency/error values.

The negative-width bias rule is still simulated but is left out of every
figure and table below. Its multipliers give far larger biases than the
bandwidth rule, so the two are not comparable until it is rescaled
(issue #500). Bias multiplier 0 (no bias) is also left out of every figure
and table below. With only the bandwidth rule shown, colour (bandwidth) is
the only legend for the curves.

```{r}
#| label: bias-uns-hide-negative-width
#| include: false
if (isTRUE(run_plots)) {
  bias_uns_results_raw <- bias_uns_results_raw |>
    dplyr::filter(
      bias_uns_basis != "negative_width",
      bias_uns_multiplier != 0
    )
  bias_uns_results_summary <- bias_uns_results_summary |>
    dplyr::filter(
      bias_uns_basis != "negative_width",
      bias_uns_multiplier != 0
    )
}
```

### Relative error across bias settings

Bias multiplier is on the horizontal axis and absolute relative error on the vertical axis. Rows of panels show the median, 90th-percentile and maximum error over the simulated samples in each setting, so a rare large miss is visible even when the typical error does not change. Colour identifies bandwidth, with the legend underneath; columns of panels separate mismatch settings, and rows the statistic and cell count. Headings give the transformation, mean position and response probability. Lower values indicate closer agreement with the simulated truth.

```{r}
#| label: bias-uns-error-summary
#| include: false
if (isTRUE(run_plots)) {
  # Error statistics for each grid setting, from valid final estimates only.
  bias_uns_error_cols <- c(
    "transformation", "mean_pos_setting", "prob_response", "n_cell",
    "mismatch_label", "mismatch_type", "mismatch_val", "bw",
    "bias_uns_basis", "bias_uns_multiplier"
  )
  bias_uns_abs_error <- bias_uns_results_raw |>
    dplyr::group_by(dplyr::across(dplyr::all_of(bias_uns_error_cols))) |>
    dplyr::summarise(
      median_abs_rel_error = stats::median(abs_rel_error, na.rm = TRUE),
      q90_abs_rel_error = stats::quantile(
        abs_rel_error,
        probs = 0.9, na.rm = TRUE, names = FALSE
      ),
      max_abs_rel_error = if (any(is.finite(abs_rel_error))) {
        max(abs_rel_error, na.rm = TRUE)
      } else {
        NA_real_
      },
      .groups = "drop"
    )
  # Over- and under-estimates summarised separately.
  bias_uns_signed_error <- .simBandwidthSignedErrorSummary(
    bias_uns_results_raw, bias_uns_error_cols
  )
}
```

```{r}
#| label: relative-error
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "relative_error"),
    path_root = root_dir
  )

  for (transformation_curr in .analysis_trans_order(bias_uns_abs_error$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_abs_error$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_abs_error$prob_response)) {
        plot_tbl <- bias_uns_abs_error |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          )

        p <- .simBandwidthBiasRelativeErrorPlot(
          plot_tbl,
          facet = ggplot2::facet_grid(
            statistic + n_cell ~ mismatch_label,
            scales = "free_y",
            labeller = ggplot2::labeller(n_cell = function(x) {
              paste0("Cells: ", .analysis_label_number(as.numeric(x)))
            })
          )
        )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        .analysis_save_fig(p, path_p, height = 22)
        .analysis_heading(
          paste0(
            "Response probability: ", .analysis_label_percent(prob_response_curr)
          ),
          6
        )
        .analysis_print_fig(p)
        .simBandwidthPrintCoverage(p, bias_uns_results_summary)
      }
    }
  }
}
```

### Relative error averaged over cell counts

Only valid estimates contribute to error statistics. Coverage tables beside each
figure report all expected samples and the failure fraction; the fallback
fraction uses all samples with known gate provenance, including failures.
`n_scenario_valid` gives the number of contributing scenarios. Averaged panels
show **Mean of scenario medians**, **Mean of scenario 90th percentiles** and
**Mean of scenario maxima**; signed-direction contributions exclude scenarios
with no estimates in that direction (the counts are shown separately).

For an overview, average the scenario-level statistics equally over cell
counts, keeping response probabilities and mismatch settings separate.

Colour identifies bandwidth; columns of panels separate mismatch settings and rows the statistic.

```{r}
#| label: relative-error-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "relative_error_averaged_n_cell"),
    path_root = root_dir
  )

  # Average scenario statistics equally across cell counts; do not pool samples.
  bias_uns_error_avg <- bias_uns_abs_error |>
    dplyr::group_by(dplyr::across(dplyr::all_of(c(
      "transformation", "mean_pos_setting", "prob_response", "mismatch_label",
      "mismatch_type", "mismatch_val", "bw", "bias_uns_basis",
      "bias_uns_multiplier"
    )))) |>
    dplyr::summarise(
      n_scenario_median = sum(is.finite(median_abs_rel_error)),
      n_scenario_q90 = sum(is.finite(q90_abs_rel_error)),
      n_scenario_max = sum(is.finite(max_abs_rel_error)),
      dplyr::across(
        c(median_abs_rel_error, q90_abs_rel_error, max_abs_rel_error),
        ~ mean(.x, na.rm = TRUE)
      ),
      .groups = "drop"
    )

  for (transformation_curr in .analysis_trans_order(bias_uns_error_avg$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_error_avg$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_error_avg$prob_response)) {
        plot_tbl <- bias_uns_error_avg |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          )

        p <- .simBandwidthBiasRelativeErrorPlot(
          plot_tbl,
          y_label = "Mean of scenario medians / 90th percentiles / maxima (absolute relative error)"
        )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        .analysis_save_fig(p, path_p, height = 14)
        .analysis_heading(
          paste0(
            "Response probability: ", .analysis_label_percent(prob_response_curr)
          ),
          6
        )
        .analysis_print_fig(p)
        .simBandwidthPrintCoverage(p, bias_uns_results_summary)
      }
    }
  }
}
```

### Relative error at each cell count

Compare the curves across cell counts while holding the biological setting fixed.

```{r}
#| label: relative-error-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "relative_error_by_n_cell"),
    path_root = root_dir
  )

  for (transformation_curr in .analysis_trans_order(bias_uns_abs_error$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_abs_error$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_abs_error$prob_response)) {
        for (n_cell_curr in sort(unique(bias_uns_abs_error$n_cell))) {
          plot_tbl <- bias_uns_abs_error |>
            dplyr::filter(
              transformation == transformation_curr,
              mean_pos_setting == mean_pos_setting_curr,
              prob_response == prob_response_curr,
              n_cell == n_cell_curr
            )
          if (nrow(plot_tbl) == 0L) next

          p <- .simBandwidthBiasRelativeErrorPlot(plot_tbl)

          path_p <- file.path(
            plot_dir,
            paste0(
              transformation_curr,
              "_", mean_pos_setting_curr,
              "_p_", format(prob_response_curr, scientific = TRUE),
              "_n_cell_", n_cell_curr,
              ".pdf"
            )
          )
          .analysis_save_fig(p, path_p, height = 14)
          # Response probability and cells share one heading level, so the
          # headings stay within six levels.
          .analysis_heading(
            paste0(
              "Response probability: ", .analysis_label_percent(prob_response_curr),
              "; cells: ", .analysis_label_number(n_cell_curr)
            ),
            6
          )
          .analysis_print_fig(p)
          .simBandwidthPrintCoverage(p, bias_uns_results_summary)
        }
      }
    }
  }
}
```

### Over- and under-estimates across bias settings

The plots above use the size of the error only. Here relative error, (estimate - truth) / truth, is split by direction: each statistic is calculated separately over over-estimates (curves above zero) and under-estimates (curves below zero), and line weight shows the share of estimates in that direction. Under-estimates are on a linear scale, including errors below -100% for negative background-subtracted estimates, with -100% (an estimate of zero, zero background-subtracted response), and over-estimates on a fold scale, so -100% and +100% (twice the truth) are the same distance from zero; both are always shown. Over-estimates larger than +1500% (16 times the truth) are drawn at +1500%, and that tick is then labelled "≥ +1500% (16x)". Rows of panels show the median, 90th-percentile and maximum error over the simulated samples in each setting, so a rare large miss is visible even when the typical error does not change. Colour identifies bandwidth; lines and points are slightly transparent so overlapping curves stay visible. The bandwidth and line-weight legends sit underneath each plot.

```{r}
#| label: signed-error
#| results: asis
#| fig-width: 6.3
#| fig-height: 8.7
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "signed_error"),
    path_root = root_dir
  )

  for (transformation_curr in .analysis_trans_order(bias_uns_signed_error$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_signed_error$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_signed_error$prob_response)) {
        plot_tbl <- bias_uns_signed_error |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          )

        p <- .simBandwidthBiasSignedErrorPlot(
          plot_tbl,
          facet = ggplot2::facet_grid(
            statistic + n_cell ~ mismatch_label,
            scales = "free_y",
            labeller = ggplot2::labeller(n_cell = function(x) {
              paste0("Cells: ", .analysis_label_number(as.numeric(x)))
            })
          )
        )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        .analysis_save_fig(p, path_p, height = 22)
        .analysis_heading(
          paste0(
            "Response probability: ", .analysis_label_percent(prob_response_curr)
          ),
          6
        )
        .analysis_print_fig(p)
        .simBandwidthPrintCoverage(p, bias_uns_results_summary)
      }
    }
  }
}
```

### Over- and under-estimates averaged over cell counts

The scenario-level statistics and shares for each direction are averaged equally over cell counts, keeping response probabilities and mismatch settings separate.

```{r}
#| label: signed-error-averaged-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "signed_error_averaged_n_cell"),
    path_root = root_dir
  )

  # Average scenario statistics equally across cell counts; do not pool samples.
  bias_uns_signed_error_avg <- bias_uns_signed_error |>
    .simBandwidthSignedErrorAverage(c(
      "transformation", "mean_pos_setting", "prob_response", "mismatch_label",
      "mismatch_type", "mismatch_val", "bw", "bias_uns_basis",
      "bias_uns_multiplier"
    ))

  for (transformation_curr in .analysis_trans_order(bias_uns_signed_error_avg$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_signed_error_avg$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_signed_error_avg$prob_response)) {
        plot_tbl <- bias_uns_signed_error_avg |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          )


        p <- .simBandwidthBiasSignedErrorPlot(
          plot_tbl,
          y_label = "Mean of scenario medians / 90th percentiles / maxima (relative error)"
        )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        .analysis_save_fig(p, path_p, height = 14)
        .analysis_heading(
          paste0(
            "Response probability: ", .analysis_label_percent(prob_response_curr)
          ),
          6
        )
        .analysis_print_fig(p)
        .simBandwidthPrintCoverage(p, bias_uns_results_summary)
      }
    }
  }
}
```

### Over- and under-estimates at each cell count

Each cell count is shown on its own.

```{r}
#| label: signed-error-by-n-cell
#| results: asis
#| fig-width: 6.3
#| fig-height: 5.5
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "signed_error_by_n_cell"),
    path_root = root_dir
  )

  for (transformation_curr in .analysis_trans_order(bias_uns_signed_error$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_signed_error$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_signed_error$prob_response)) {
        for (n_cell_curr in sort(unique(bias_uns_signed_error$n_cell))) {
          plot_tbl <- bias_uns_signed_error |>
            dplyr::filter(
              transformation == transformation_curr,
              mean_pos_setting == mean_pos_setting_curr,
              prob_response == prob_response_curr,
              n_cell == n_cell_curr
            )
          if (nrow(plot_tbl) == 0L) next

          p <- .simBandwidthBiasSignedErrorPlot(plot_tbl)

          path_p <- file.path(
            plot_dir,
            paste0(
              transformation_curr,
              "_", mean_pos_setting_curr,
              "_p_", format(prob_response_curr, scientific = TRUE),
              "_n_cell_", n_cell_curr,
              ".pdf"
            )
          )
          .analysis_save_fig(p, path_p, height = 14)
          # Response probability and cells share one heading level, so the
          # headings stay within six levels.
          .analysis_heading(
            paste0(
              "Response probability: ", .analysis_label_percent(prob_response_curr),
              "; cells: ", .analysis_label_number(n_cell_curr)
            ),
            6
          )
          .analysis_print_fig(p)
          .simBandwidthPrintCoverage(p, bias_uns_results_summary)
        }
      }
    }
  }
}
```


### Estimated response frequency

Median estimated frequency is plotted against bias multiplier, with colour identifying bandwidth. Rows of panels separate cell counts and columns mismatch settings; dashed lines show the simulated truth. Curves closer to those lines indicate better agreement.

```{r}
#| label: estimated-frequency
#| results: asis
#| fig-width: 6.3
#| fig-height: 4.3
#| echo: false
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(
    c(fig_key, "estimated_frequency"),
    path_root = root_dir
  )

  for (transformation_curr in .analysis_trans_order(bias_uns_results_summary$transformation)) {
    .analysis_heading(
      paste0("Transformation: ", .analysis_trans_factor(transformation_curr)), 4
    )
    for (mean_pos_setting_curr in unique(bias_uns_results_summary$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 5)
      for (prob_response_curr in unique(bias_uns_results_summary$prob_response)) {
        plot_tbl <- bias_uns_results_summary |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          ) |>
          dplyr::mutate(bw_lab = .simBandwidthBwLabFactor(bw))

        p <- ggplot(
          plot_tbl,
          aes(
            x = bias_uns_multiplier,
            y = propRespEst_median,
            colour = bw_lab,
            group = interaction(bw, bias_uns_basis)
          )
        ) +
          geom_hline(
            aes(yintercept = propRespTruth),
            linetype = "dashed",
            colour = "black"
          ) +
          geom_line(alpha = 0.75) +
          geom_point(alpha = 0.75) +
          facet_grid(
            n_cell ~ mismatch_label,
            scales = "free_y",
            labeller = labeller(n_cell = function(x) {
              paste0("Cells: ", .analysis_label_number(as.numeric(x)))
            })
          ) +
          .simBandwidthBwColourScale(plot_tbl$bw) +
          scale_x_continuous(labels = .analysis_label_number) +
          scale_y_continuous(labels = .analysis_label_percent) +
          .analysis_theme() +
          labs(
            x = "Bias multiplier",
            y = "Estimated background-subtracted frequency (median)",
            colour = "Bandwidth"
          )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        .analysis_save_fig(p, path_p, height = 11)
        .analysis_heading(
          paste0(
            "Response probability: ", .analysis_label_percent(prob_response_curr)
          ),
          6
        )
        .analysis_print_fig(p)
        .simBandwidthPrintCoverage(p, bias_uns_results_summary)
      }
    }
  }
}
```
