---
title: Tune unstimulated-sample bias (2b)
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20261003
  run_simulations: true
  run_plots: false
  sim_retry_errors: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20261003L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bias_uns", "freq_bs")
analysis_semantics_version <- "bias-uns-freq-v1"
analysis_quick <- isTRUE(.isQuick())
analysis_dev <- isTRUE(.isDev())

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

Analysis 2a identifies a plausible bandwidth range while keeping `bias_uns`
deliberately simple. This analysis asks the complementary question: across
those plausible bandwidths, how should `bias_uns` be set when the stimulated
negative population is displaced relative to its matched unstimulated
population?

The batch mismatch follows Analysis 8 and is applied only to the stimulated
negative/background component. Random sample, condition and cluster
perturbations are disabled so that the controlled mismatch is the source of
stimulated-versus-unstimulated disagreement.

### Simulation grid

```{r}
#| label: bias-uns-settings
n_cell_vec <- c(1e4, 5e4)
prob_response_vec <- c(0.002, 0.05)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

bandwidth_settings_tbl <- tibble::tribble(
  ~transformation, ~bw,
  "gaussian", 0.05,
  "gaussian", 0.10,
  "gaussian", 0.25,
  "gaussian", 0.50,
  "skew",     0.05,
  "skew",     0.10,
  "skew",     0.25,
  "skew",     0.50,
  "gamma",    0.001,
  "gamma",    0.0025,
  "gamma",    0.005,
  "gamma",    0.010
)

bias_uns_settings_tbl <- dplyr::bind_rows(
  tibble::tibble(
    bias_uns_basis = "bandwidth",
    bias_uns_multiplier = c(
      0, 0.1, 0.25, 0.33, 0.5, 0.75, 1, 1.25, 1.5, 2
    )
  ),
  tibble::tibble(
    bias_uns_basis = "negative_width",
    bias_uns_multiplier = c(0.1, 0.25, 0.5, 0.75)
  )
) |>
  dplyr::mutate(
    bias_uns_rule = paste0(
      format(.data$bias_uns_multiplier, trim = TRUE),
      " x ",
      dplyr::if_else(
        .data$bias_uns_basis == "bandwidth",
        "bandwidth",
        "negative width"
      )
    )
  )

mean_shift_gaussian_skew <- c(0, 0.025, 0.05, 0.2)
mean_shift_gamma <- c(0, 0.005, 0.01, 0.05)

mean_shift_tbl <- dplyr::bind_rows(
  tidyr::expand_grid(
    transformation = c("gaussian", "skew"),
    mismatch_val = mean_shift_gaussian_skew
  ),
  tidyr::expand_grid(
    transformation = "gamma",
    mismatch_val = mean_shift_gamma
  )
) |>
  dplyr::mutate(
    mismatch_type = "mean_shift",
    mismatch_label = paste0("mean shift ", .data$mismatch_val),
    stim_mean_shift = .data$mismatch_val,
    stim_mean_shift_clusters = "gn",
    stim_sd_multiplier = 1,
    stim_sd_multiplier_clusters = NA_character_,
    sd_increase = 0
  )

sd_inflation_tbl <- tibble::tibble(
  transformation = c("gaussian", "gamma", "skew"),
  mismatch_val = 0.10,
  mismatch_type = "sd_inflation",
  mismatch_label = "SD inflation 10%",
  stim_mean_shift = 0,
  stim_mean_shift_clusters = NA_character_,
  stim_sd_multiplier = 1.10,
  stim_sd_multiplier_clusters = "gn",
  sd_increase = 0.10
)

mismatch_settings_tbl <- dplyr::bind_rows(
  mean_shift_tbl,
  sd_inflation_tbl
)

background_relative_to_response <- 0.2
n_cell_uns_relative_to_stim <- 1

tol_clust <- NULL
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
```

For the negative-width rules, the width is calculated from the pooled
unstimulated cells in the simulation row, using the same fixed KDE bandwidth
being evaluated. Starting at the dominant negative peak, the right endpoint is
the first point at which the density reaches 15% of the peak height or the next
antimode, whichever comes first.

```{r}
#| label: bias-uns-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  n_cell = n_cell_vec,
  prob_response = prob_response_vec,
  bias_uns_settings_tbl
) |>
  dplyr::inner_join(
    bandwidth_settings_tbl,
    by = "transformation"
  ) |>
  dplyr::inner_join(
    mismatch_settings_tbl,
    by = "transformation"
  ) |>
  dplyr::mutate(
    background_relative_to_response = background_relative_to_response,
    n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim
  ) |>
  dplyr::arrange(
    .data$transformation,
    .data$n_cell,
    .data$prob_response,
    .data$mean_pos_setting,
    .data$mismatch_type,
    .data$mismatch_val,
    .data$bw,
    .data$bias_uns_basis,
    .data$bias_uns_multiplier
  )

base_scenario_tbl <- sim_grid |>
  dplyr::distinct(
    .data$transformation,
    .data$n_cell,
    .data$prob_response,
    .data$mean_pos_setting,
    .data$mean_pos,
    .data$mismatch_type,
    .data$mismatch_val,
    .data$stim_mean_shift,
    .data$stim_mean_shift_clusters,
    .data$stim_sd_multiplier,
    .data$stim_sd_multiplier_clusters
  ) |>
  dplyr::mutate(base_scenario_id = dplyr::row_number())

sim_grid <- sim_grid |>
  dplyr::left_join(
    base_scenario_tbl,
    by = c(
      "transformation",
      "n_cell",
      "prob_response",
      "mean_pos_setting",
      "mean_pos",
      "mismatch_type",
      "mismatch_val",
      "stim_mean_shift",
      "stim_mean_shift_clusters",
      "stim_sd_multiplier",
      "stim_sd_multiplier_clusters"
    )
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number(),
    # Rows that differ only by bandwidth or bias rule use identical simulated
    # data, so tuning comparisons use common random numbers.
    sim_seed = as.integer(simulation_seed + .data$base_scenario_id - 1L)
  )

if (nrow(sim_grid) != 6720L) {
  stop(
    "Expected 6,720 Analysis 2b grid rows, found ",
    nrow(sim_grid),
    "."
  )
}

sim_grid_full <- sim_grid

if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e4,
      prob_response == 0.002,
      mean_pos_setting == "high",
      mismatch_label %in% c("mean shift 0", "SD inflation 10%"),
      bw %in% c(0.05, 0.1, 0.005),
      bias_uns_multiplier %in% c(0, 0.25, 0.5)
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      transformation == "gaussian",
      n_cell == 1e4,
      prob_response == 0.002,
      mean_pos_setting == "high",
      mismatch_label == "mean shift 0.05",
      bw == 0.1,
      bias_uns_rule %in% c(
        "0.25 x bandwidth",
        "0.5 x negative width"
      )
    )
}

sim_grid_all <- sim_grid
if (nrow(sim_grid_all) == 0L) {
  stop("The Analysis 2b simulation grid is empty; check the dev/quick filters.")
}

analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
  dplyr::select(-.data$.sim_grid_chunk)

print(
  paste0(
    "Performing ",
    nrow(sim_grid),
    " of ",
    nrow(sim_grid_all),
    " Analysis 2b simulations for chunk ",
    sim_grid_chunk_index,
    " of ",
    sim_grid_n_chunks
  )
)
```

```{r}
#| label: bias-uns-run-settings
scenario_settings <- list(
  nSample = 25,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = 1,
  biasUnsWidthHeightFrac = 0.15,
  bwMin = "none",
  bwMax = "none",
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = tol_clust,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates
)

analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

A single row can be rerun interactively with exactly the same random numbers
used by the parallel run.

```{r}
#| label: rerun-one-simulation
#| eval: false
sim_id_target <- 1L
row_target <- sim_grid_full |>
  dplyr::filter(.data$sim_id == sim_id_target)

rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthBiasUnsScenario,
  settings = scenario_settings
)
```

```{r}
#| label: bias-uns-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthBiasUnsScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    heading = "BIAS_UNS FREQUENCY PROGRESS",
    path_root = root_dir
  )

  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    collate_fn = function(tbl) {
      .simBandwidthBiasUnsCollate(tbl, colnames(sim_grid_all))
    },
    label = "analysis 2b"
  )

  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  run_ctx <- .analysis_results_context(analysis_key, path_root = root_dir)
}
```

### Results

```{r}
#| label: bias-uns-collate
if (isTRUE(run_plots)) {
  read_current <- function(nm) {
    readRDS(.analysis_current_file(
      run_ctx = run_ctx,
      relative_path = c("collated", paste0(nm, ".rds")),
      required_params = analysis_required_params
    ))
  }
  bias_uns_results_raw <- read_current("bias_uns_results_raw")
  bias_uns_results_summary <- read_current("bias_uns_results_summary")
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

The primary summaries are relative error and the estimated
background-subtracted frequencies themselves. Plot each response probability
separately so that every remaining varying scenario dimension is either
faceted or included in the grouping.

```{r}
#| label: plot-relative-error
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_cache_dir(
    c(analysis_key, "relative_error"),
    path_root = root_dir
  )
  dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

  for (transformation_curr in unique(bias_uns_results_summary$transformation)) {
    for (mean_pos_setting_curr in unique(bias_uns_results_summary$mean_pos_setting)) {
      for (prob_response_curr in unique(bias_uns_results_summary$prob_response)) {
        plot_tbl <- bias_uns_results_summary |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          )

        p <- ggplot(
          plot_tbl,
          aes(
            x = bias_uns_multiplier,
            y = median_abs_rel_error,
            colour = factor(bw),
            linetype = bias_uns_basis,
            group = interaction(bw, bias_uns_basis)
          )
        ) +
          geom_line() +
          geom_point() +
          facet_grid(n_cell ~ mismatch_label, scales = "free_y") +
          cowplot::theme_cowplot() +
          cowplot::background_grid(major = "xy") +
          theme(legend.position = "bottom") +
          labs(
            title = paste0(
              transformation_curr,
              "; mean position = ", mean_pos_setting_curr,
              "; response probability = ", prob_response_curr
            ),
            x = "Bias multiplier",
            y = "Median absolute relative error",
            colour = "Bandwidth",
            linetype = "Bias scale"
          )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        ggsave(path_p, p, width = 28, height = 13, units = "cm")
        print(p)
      }
    }
  }
}
```

```{r}
#| label: plot-estimated-frequency
if (isTRUE(run_plots)) {
  plot_dir <- .analysis_cache_dir(
    c(analysis_key, "estimated_frequency"),
    path_root = root_dir
  )
  dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

  for (transformation_curr in unique(bias_uns_results_summary$transformation)) {
    for (mean_pos_setting_curr in unique(bias_uns_results_summary$mean_pos_setting)) {
      for (prob_response_curr in unique(bias_uns_results_summary$prob_response)) {
        plot_tbl <- bias_uns_results_summary |>
          dplyr::filter(
            transformation == transformation_curr,
            mean_pos_setting == mean_pos_setting_curr,
            prob_response == prob_response_curr
          )

        p <- ggplot(
          plot_tbl,
          aes(
            x = bias_uns_multiplier,
            y = propRespEst_median,
            colour = factor(bw),
            linetype = bias_uns_basis,
            group = interaction(bw, bias_uns_basis)
          )
        ) +
          geom_hline(
            aes(yintercept = propRespTruth),
            linetype = "dashed",
            colour = "black"
          ) +
          geom_line() +
          geom_point() +
          facet_grid(n_cell ~ mismatch_label, scales = "free_y") +
          cowplot::theme_cowplot() +
          cowplot::background_grid(major = "xy") +
          theme(legend.position = "bottom") +
          labs(
            title = paste0(
              transformation_curr,
              "; mean position = ", mean_pos_setting_curr,
              "; response probability = ", prob_response_curr
            ),
            x = "Bias multiplier",
            y = "Estimated background-subtracted frequency (median)",
            colour = "Bandwidth",
            linetype = "Bias scale"
          )

        path_p <- file.path(
          plot_dir,
          paste0(
            transformation_curr,
            "_", mean_pos_setting_curr,
            "_p_", format(prob_response_curr, scientific = TRUE),
            ".pdf"
          )
        )
        ggsave(path_p, p, width = 28, height = 13, units = "cm")
        print(p)
      }
    }
  }
}
```
