---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```


### Checking what the estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider two variations - whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator.
We can also vary the amount of separation, I suppose.

We'll need to vary over a range of parent population sizes, however, as that will tell us how the estimator behaves over the most crucial dimension (apart from the actual transformation, of course).

Initially, we will ignore setting maximum/minimum number of cells.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# bandwidth settings
bandwidth_method_vec <- c("hpi0", "hpi1", "hpi2", "hpi3", "sj", "nrd0")

# set this later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    bias_uns_setting == "low"
  )
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw_mtd         = bandwidth_method_vec,
  bias_uns       = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd,
    bias_uns_setting, bias_uns
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw_mtd == "hpi1"
    )
}
sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))

bw_fallback <- 0.23482348792138919129198282389
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
bw_list_raw_mtd_loop <- lapply(seq_len(nrow(sim_grid)), function(i) {
  NULL
})

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | bias_uns_setting = ",
    sim_grid$bias_uns_setting[i],
    " | bias_uns = ",
    sim_grid$bias_uns[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)
  bw_list_raw_mtd_loop[[i]] <- stimgate:::.simBandwidthEstBwDirect(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 10,
    biasUns = sim_grid$bias_uns[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwFallback = bw_fallback,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bias_uns_setting = sim_grid$bias_uns_setting[i],
      bias_uns = sim_grid$bias_uns[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bias_uns_setting,
      bias_uns,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

dir.create(
  dirname(projr::projr_path_get(
    "cache", "sim", "bw", "est", "base", "bw_list_raw_mtd_loop.rds"
  )),
  recursive = TRUE,
  showWarnings = FALSE
)

saveRDS(
  bw_list_raw_mtd_loop,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "base",
    "bw_list_raw_mtd_loop.rds"
  )
)

bw_list_raw_mtd_loop |>
  dplyr::bind_rows() |>
  dplyr::group_by(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd,
    bias_uns_setting, bias_uns, chnl,
    n_cell_stim, n_cell_uns
  ) |>
  dplyr::summarise(
    mean_bw_stim = mean(bw_stim, na.rm = TRUE),
    mean_bw_uns = mean(bw_uns, na.rm = TRUE),
    mean_bw = mean(bw, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation, mean_bw
  )
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel

if (interactive()) {
  progressr::handlers(global = TRUE)
}

# 1. Setup paths
path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "bw",
  "est",
  "base",
  "progress.txt"
)
dir_progress <- dirname(path_progress_file)
dir_jobs <- file.path(dir_progress, "jobs")

# 2. Clean up previous progress file and the jobs subdirectory
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}
if (dir.exists(dir_jobs)) {
  unlink(dir_jobs, recursive = TRUE, force = TRUE)
}
dir.create(dir_jobs, recursive = TRUE, showWarnings = FALSE)

# 3. Setup parallel plan
future::plan(
  future::multisession,
  workers = max(1, .simGetCores())
)

# Total number of iterations
total_sims <- nrow(sim_grid)

bw_list_raw_mtd <- progressr::with_progress({
  p <- progressr::progressor(steps = total_sims)

  furrr::future_pmap(
    sim_grid,
    function(transformation, prob_response, n_cell, mean_pos_setting, mean_pos, bw_mtd, bias_uns_setting, bias_uns, sim_id) {
      settings_row <- paste0(
        "trans = ",
        transformation,
        " | prob = ",
        prob_response,
        " | n_cell = ",
        n_cell,
        " | mean = ",
        mean_pos,
        " | bw_mtd = ",
        bw_mtd,
        " | bias_uns_setting = ",
        bias_uns_setting,
        " | bias_uns = ",
        bias_uns,
        " | sim_id = ",
        sim_id
      )

      # Define file markers for this specific simulation
      file_running <- file.path(dir_jobs, paste0("running-", sim_id))
      file_completed <- file.path(dir_jobs, paste0("completed-", sim_id))
      file_error <- file.path(dir_jobs, paste0("error-", sim_id))

      # Helper function to recalculate state and update progress.txt
      update_progress_summary <- function() {
        # List all generated status files
        all_job_files <- list.files(dir_jobs, full.names = FALSE)

        n_running <- sum(grepl("^running-", all_job_files))
        n_completed <- sum(grepl("^completed-", all_job_files))
        n_error <- sum(grepl("^error-", all_job_files))
        n_done <- n_completed + n_error

        # Extract specifically which ones are currently running
        running_files <- all_job_files[grepl("^running-", all_job_files)]
        running_ids <- sub("^running-", "", running_files)
        running_list_str <- if (length(running_ids) > 0) {
          paste(running_ids, collapse = ", ")
        } else {
          "None"
        }

        # Build dashboard content
        summary_text <- paste0(
          "========================================\n",
          " SIMULATION PROGRESS DASHBOARD          \n",
          "========================================\n",
          "Total Simulations  : ", total_sims, "\n",
          "In Progress        : ", n_running, "\n",
          "Completed (Success): ", n_completed, "\n",
          "Completed (Errors) : ", n_error, "\n",
          "Total Done          : ", n_done, "\n",
          "----------------------------------------\n",
          "Currently Running IDs:\n", running_list_str, "\n",
          "========================================\n"
        )

        # Overwrite the summary file safely
        cat(summary_text, file = path_progress_file, append = FALSE)
      }

      # --- START WORKER PROCESS ---
      # Mark as running
      file.create(file_running)
      update_progress_summary()

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthEstBwDirect(
            nSample = 50,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 1,
            biasUns = bias_uns,
            bwMtd = bw_mtd,
            bwFallback = bw_fallback,
            bwMin = -Inf,
            bwMax = Inf,
            bwCluster = 0.5,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            backgroundRelativeToResponse = 0.2,
            ncellUnsRelativeToStim = 1,
            covEvMin = 1.5,
            covEvMax = 1.5,
            tolClust = NULL,
            summarise = FALSE
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              bias_uns_setting = bias_uns_setting,
              bias_uns = bias_uns,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos_setting,
              mean_pos,
              bw_mtd,
              bias_uns_setting,
              bias_uns,
              sim_id,
              dplyr::everything()
            )

          # Move marker from running to completed
          file.create(file_completed)
          if (file.exists(file_running)) file.remove(file_running)

          update_progress_summary()
          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          # Move marker from running to error
          file.create(file_error)
          if (file.exists(file_running)) file.remove(file_running)

          update_progress_summary()
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw_mtd = bw_mtd,
            bias_uns_setting = bias_uns_setting,
            bias_uns = bias_uns,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            bw_uns = NA_real_,
            bw_stim = NA_real_,
            bw = NA_real_,
            bw_source = NA_character_,
            error = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

dir.create(
  dirname(projr::projr_path_get(
    "cache", "sim", "bw", "est", "base", "bw_list_raw_mtd.rds"
  )),
  recursive = TRUE,
  showWarnings = FALSE
)

saveRDS(
  bw_list_raw_mtd,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "base",
    "bw_list_raw_mtd.rds"
  )
)
```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true
path_vec <- c(
  projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "base",
    "bw_list_raw_mtd.rds",
    safe = TRUE
  ),
  projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est", "base",
    "bw_list_raw_mtd.rds",
    safe = FALSE
  )
)
path_vec <- path_vec[file.exists(path_vec)]
if (length(path_vec) == 0L) {
  stop("Could not find bw_list_raw_mtd.rds. Run the simulation chunk first.")
}

bw_list_raw_mtd <- readRDS(
  file = path_vec[[1]]
)

bw_tbl <- bw_list_raw_mtd |>
  purrr::list_rbind()

if (!"bias_uns" %in% colnames(bw_tbl)) {
  bw_tbl$bias_uns <- NA_real_
}
if (!"bias_uns_setting" %in% colnames(bw_tbl)) {
  bw_tbl$bias_uns_setting <- NA_character_
}

if (all(is.na(bw_tbl$bias_uns_setting)) && any(!is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-bias_uns_setting) |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns")
    )
}

bw_tbl <- bw_tbl |>
  dplyr::filter(!is.na(.data$bias_uns_setting))

if (nrow(bw_tbl) == 0L) {
  stop("No cached rows contain the current bias_uns settings. Rerun the simulation chunk.")
}

sim_grid_key_cols <- setdiff(
  intersect(colnames(sim_grid), colnames(bw_tbl)),
  "sim_id"
)

bw_tbl <- bw_tbl |>
  dplyr::semi_join(
    sim_grid |>
      dplyr::select(dplyr::all_of(sim_grid_key_cols)) |>
      dplyr::distinct(),
    by = sim_grid_key_cols
  )

if (nrow(bw_tbl) == 0L) {
  stop("Cached rows do not match the current simulation grid. Rerun the simulation chunk.")
}

bw_tbl_results <- bw_tbl |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    mean_bw_stim = mean(bw_stim, na.rm = TRUE),
    mean_bw_uns = mean(bw_uns, na.rm = TRUE),
    mean_bw = mean(bw, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::select(
    dplyr::any_of(colnames(sim_grid)), mean_bw
  )

for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
  for (bias_uns_setting_curr in unique(bw_tbl_results$bias_uns_setting)) {
    plot_tbl <- bw_tbl_results |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr,
        bias_uns_setting == !!bias_uns_setting_curr
      )

    if (nrow(plot_tbl) == 0L) next

    p <- ggplot(
      plot_tbl,
      aes(x = n_cell, y = mean_bw, color = bw_mtd)
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(
        transformation ~ prob_response,
        scales = "free_y"
      ) +
      labs(
        title = paste0(
          "Mean position setting: ", mean_pos_setting_curr,
          ", bias_uns_setting: ", bias_uns_setting_curr
        ),
        x = "Number of Cells",
        y = "Estimated Bandwidth"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy") +
      theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        legend.position = "bottom"
      ) +
      scale_colour_manual(
        values = .simBwMtdColVec
      )

    path_plot <- projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est", "base",
      paste0(
        "bw_estimate_meanpos_setting_", mean_pos_setting_curr,
        "_bias_uns_", bias_uns_setting_curr,
        ".png"
      )
    )

    ggsave(
      filename = path_plot,
      plot = p,
      width = 15,
      height = 20,
      units = "cm"
    )
  }
}
```


