---
title: Simulate base bandwidth estimators
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
  show_mcse: "on"
  sim_retry_errors: true
---

This analysis asks how ordinary bandwidth estimates depend on cell count, response frequency and separation. Matched stimulated and unstimulated samples are simulated with Gaussian, gamma and skew transformations, then bandwidths are estimated separately for each condition.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
analysis_qmd <- "analysis/3-sim-bw-est-base.qmd"

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(sim_grid_shuffle_seed)) {
  stop("sim_grid_shuffle_seed must be coercible to an integer.")
}
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "est", "base")
analysis_semantics_version <- "bandwidth-est-base-v7"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "draft" (current default) or "final" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("3-sim-bw-est-base")
n_sample_sim <- if (analysis_quick) 1L else c(final = 50L, draft = 15L)[[sim_size]]
n_iter_sim <- 1L
bw_fallback <- NA_real_
bw_min <- -Inf
bw_max <- Inf
prob_exact_sim <- TRUE
background_relative_to_response <- 0.2
ncell_uns_relative_to_stim <- 1
cov_ev_min <- 1.5
cov_ev_max <- 1.5
exc_min <- TRUE
cap_stim_range <- FALSE
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
# Monte Carlo error bars on the summary plots (see analysis-mcse.R).
mcse_mode <- .analysis_mcse_mode(.get_qmd_param_env("show_mcse", "SHOW_MCSE", "on"))
show_mcse <- mcse_mode != "off"

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-bandwidth.R for simulation utilities.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

Get the parameters over which we simulate.

We mainly need to consider two variations - whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator.
We can also vary the amount of separation, I suppose.

We'll need to vary over a range of parent population sizes, however, as that will tell us how the estimator behaves over the most crucial dimension (apart from the actual transformation, of course).

Only low unstimulated-sample bias is studied, with transformation-specific
magnitudes. Bandwidth estimates are unbounded (`bw_min = -Inf`,
`bw_max = Inf`); cell-count limits use the wrapper defaults (`NULL`).

External chunks share one run ID and promote results only after all selected
scenarios validate. Failed simulations are retried on resume. Plots require
matching promoted results; render with simulations disabled after a multi-chunk
run completes.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 5e3, 2e4, 1e5)
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
# bandwidth settings
bandwidth_method_vec <- c("hpi0", "hpi1", "hpi2", "hpi3", "sj", "nrd0")

bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05
)
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw_mtd         = bandwidth_method_vec,
  bias_uns       = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd,
    bias_uns_setting, bias_uns
  ) |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos
  ) |>
  dplyr::mutate(
    base_scenario_id = dplyr::cur_group_id(),
    sim_seed = as.integer(simulation_seed + base_scenario_id - 1L)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(sim_id = dplyr::row_number())

sim_grid_full <- sim_grid

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

# Quick mode keeps cells, method, transformation and response.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(1e3, 5e3),
      prob_response %in% c(0.002, 0.2),
      transformation %in% c("gaussian", "gamma"),
      mean_pos_setting == "high",
      bw_mtd %in% c("hpi1", "nrd0")
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw_mtd == "hpi1"
    )
}

sim_grid_all <- sim_grid
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

if (nrow(sim_grid_all) == 0L) {
  stop("The analysis 3 simulation grid is empty; check the dev filters.")
}

sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk = ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
  dplyr::arrange(.data$n_cell, dplyr::desc(.data$prob_response)) |>
  dplyr::select(-.sim_grid_chunk)

sim_grid
print(paste0(
  "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
  " simulations for sim_grid chunk ", sim_grid_chunk_index,
  " of ", sim_grid_n_chunks,
  " using shuffle seed ", sim_grid_shuffle_seed
))

```

Fixed settings and the complete selected cross-chunk grid are recorded in the
manifest and required when reading canonical results. Each scenario simulates
50 samples in final runs and 15 in draft runs (`SIM_SIZE=draft`); draft results
are kept separately and are for iterating, not for reporting.

```{r}
#| label: bw-estimate-settings
#| include: false
scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = n_iter_sim,
  bwFallback = bw_fallback,
  bwMin = bw_min,
  bwMax = bw_max,
  probExact = prob_exact_sim,
  backgroundRelativeToResponse = background_relative_to_response,
  ncellUnsRelativeToStim = ncell_uns_relative_to_stim,
  covEvMin = cov_ev_min,
  covEvMax = cov_ev_max,
  excMin = exc_min,
  capStimRange = cap_stim_range,
  summarise = FALSE
)
analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

## Simulation

Method and tuning settings share the same biological-scenario seed, assigned on the full grid before execution filters or chunking. Each replicate has a seed drawn before any method runs, so later replicates remain paired despite method RNG use.

Rerun one full-grid simulation with the same random numbers as its worker.
Call `debugonce()` on a helper first to investigate a failure.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L
row_target <- sim_grid_full[sim_grid_full$sim_id == sim_id_target, ]
rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthEstBaseScenario,
  settings = scenario_settings,
  error_col = "error"
)
dir_run <- .analysis_cache_dir(
  c(analysis_key, "current"), path_root = root_dir, create = FALSE
)
path_stored <- .simBandwidthFindSimOutput(dir_run, sim_id_target)
if (length(path_stored) == 1L) {
  all.equal(rerun_res, readRDS(path_stored))
}
```

Run the simulations and validate this chunk. Promotion waits for all chunks.

```{r}
#| label: bw-estimate-parallel
#| include: false
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthEstBaseScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    error_col = "error",
    heading = "BASE BANDWIDTH ESTIMATION PROGRESS",
    path_root = root_dir
  )
  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    collate_fn = function(tbl) {
      .simBandwidthEstBaseCollate(tbl, colnames(sim_grid_all))
    },
    validate_fn = function(tbl) {
      .simBandwidthEstBaseValidate(tbl, scenario_settings)
    },
    error_col = "error",
    label = "analysis 3"
  )
  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
}
```

## Collation

Read matching canonical results and plot the scenario means.

The plotted mean uses the smaller stimulated and unstimulated bandwidth for each complete finite pair; `prop_bw_finite` records the proportion of such pairs.

```{r}
#| label: bw-estimate-collate
#| include: false
if (isTRUE(run_plots)) {
  path_raw <- .analysis_current_file(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_list_raw_mtd.rds"),
    required_params = analysis_required_params
  )
  path_results <- .analysis_current_file(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results.rds"),
    required_params = analysis_required_params
  )
  bw_tbl <- .analysis_read_rds(path_raw, analysis_key, analysis_qmd)
  bw_tbl_results <- .analysis_read_rds(path_results, analysis_key, analysis_qmd)
  # Derive diagnostics from saved raw provenance, including older summaries.
  bw_tbl_results <- .simBandwidthEstBaseSummary(bw_tbl, colnames(sim_grid_all))
  if (any(bw_tbl_results$n_bw_finite < bw_tbl_results$n_bw_total)) {
    message(
      "Some bandwidth estimates are non-finite; prop_bw_finite records ",
      "coverage and mean_bw uses complete estimator pairs."
    )
  }
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

## Results

Each figure has a companion table showing finite paired bandwidth estimates
out of the expected sample count and the selected bandwidth's normalisation
fallback share. Means use complete finite stim/unstim pairs. Analysis 3 disables
numeric bandwidth fallbacks; an estimator failure remains a coverage failure,
rather than a fallback estimate.

Mean bandwidth estimates are plotted against cell count on a logarithmic axis, with colour identifying the estimator. Facets separate transformations (Gaussian, Skew, Gamma) and response frequencies, shown as percentages; each figure sits under a heading naming its response-separation and bias setting. Compare the methods within each panel and follow how their estimates change with cell count.

Set `show_mcse` or `SHOW_MCSE` to `off` or `on` (the default). Each render prints one version and saves it under `mcse_off/` or `mcse_on/`, with the same points and axes; historical false/true values select off/on. With error bars on, each mean has a thin error bar in its line's colour: the mean plus or minus 1.96 Monte Carlo standard errors (the standard deviation of the scenario's bandwidth estimates divided by the square root of their number). The error bars use only the estimates already simulated and treat them as independent draws, which holds here: each simulated sample is drawn and its bandwidth estimated separately, with no sharing across samples.

```{r}
#| label: bw-estimate-plots
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  # Monte Carlo interval for each scenario's mean bandwidth: mean +/- 1.96
  # sd / sqrt(n) over its complete finite stim/unstim pairs.
  bw_mean_mcse <- bw_tbl |>
    dplyr::mutate(
      bw_pair = dplyr::if_else(
        is.finite(bw_stim) & is.finite(bw_uns),
        pmin(bw_stim, bw_uns),
        NA_real_
      )
    ) |>
    dplyr::group_by(sim_id) |>
    dplyr::reframe(.analysis_mcse_mean(bw_pair)) |>
    dplyr::select(sim_id, mean_bw_lower = lower, mean_bw_upper = upper)
  for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
    .analysis_heading(paste0("Mean position: ", mean_pos_setting_curr), 3)
    for (bias_uns_setting_curr in unique(bw_tbl_results$bias_uns_setting)) {
      plot_tbl <- bw_tbl_results |>
        dplyr::filter(
          .data$mean_pos_setting == !!mean_pos_setting_curr,
          .data$bias_uns_setting == !!bias_uns_setting_curr
        )

      if (nrow(plot_tbl) == 0L) next

      .analysis_heading(paste0("Bias: ", bias_uns_setting_curr), 4)

      plot_tbl <- plot_tbl |>
        dplyr::left_join(bw_mean_mcse, by = "sim_id") |>
        dplyr::mutate(
          bw_mtd = factor(
            .data$bw_mtd,
            levels = c("nrd0", "sj", "hpi0", "hpi1", "hpi2", "hpi3")
          ),
          transformation = .analysis_trans_factor(.data$transformation)
        )

      p <- ggplot(
        plot_tbl,
        aes(x = .data$n_cell, y = .data$mean_bw, color = .data$bw_mtd, shape = .data$bw_mtd)
      ) +
        (if (isTRUE(show_mcse)) {
          .analysis_mcse_errorbar(plot_tbl, "mean_bw_lower", "mean_bw_upper")
        }) +
        geom_line(alpha = 0.75) +
        geom_point(alpha = 0.75) +
        facet_grid(
          transformation ~ prob_response,
          scales = "free_y",
          labeller = labeller(prob_response = .analysis_labeller_percent())
        ) +
        labs(
          x = "Number of cells",
          y = "Estimated bandwidth",
          colour = "Bandwidth method"
        ) +
        .analysis_theme() +
        scale_colour_manual(
          values = .simBwMtdColVec,
          labels = .simBwMtdLabVec
        ) +
        scale_shape_manual(
          name = "Bandwidth method",
          values = c(nrd0 = 16, sj = 17, hpi0 = 15, hpi1 = 18, hpi2 = 3, hpi3 = 7),
          labels = .simBwMtdLabVec
        ) +
        scale_x_log10(
          breaks = c(1e3, 1e4, 1e5),
          labels = .analysis_label_number,
          guide = guide_axis(angle = 45)
        ) +
        scale_y_continuous(labels = .analysis_label_number)

      path_plot <- file.path(
        .analysis_fig_dir(c(fig_key, "est_bw_base"), root_dir),
        paste0(
          "est_bw_base",
          "_meanpos_",
          mean_pos_setting_curr,
          "_bias_uns_",
          bias_uns_setting_curr,
          ".pdf"
        )
      )
      .analysis_print_save_fig(p, path_plot, height = 15, mcse_mode = mcse_mode)
      .analysis_report_table(plot_tbl |> dplyr::select(
        transformation, prob_response, n_cell, bw_mtd,
        n_bw_finite, n_bw_total, prop_bw_finite, norm_fallback_rate
      ),
        c(fig_key, "bandwidth-coverage", paste0(tools::file_path_sans_ext(basename(path_plot)), ".csv")),
        "Finite bandwidth-estimate coverage and normalisation fallback rates for this figure.", path_root = root_dir)
    }
  }
}
```
