---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```


### Checking what the estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider two variations - whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator.
We can also vary the amount of separation, I suppose.

We'll need to vary over a range of parent population sizes, however, as that will tell us how the estimator behaves over the most crucial dimension (apart from the actual transformation, of course).

Initially, we will ignore setting maximum/minimum number of cells.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_vec <- c(4, 8)
transformation_vec <- c("gaussian", "gamma", "skew")

# bandwidth settings
bandwidth_method_vec <- c("hpi0", "hpi1", "hpi2", "hpi3", "sj", "nrd0")

# set this later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- seq(0, 2, by = 0.2)
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  transformation = transformation_vec,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  mean_pos       = mean_pos_vec,
  bw_mtd         = bandwidth_method_vec
) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos == 8,
      transformation %in% transformation_vec,
      bw_mtd == "hpi1"
    )
}
sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
bw_list_raw_mtd_loop <- lapply(seq_len(nrow(sim_grid)), function(i) {
  NULL
})

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)

  bw_list_raw_mtd_loop[[i]] <- stimgate:::.simBandwidthEstBwDirect(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 10,
    biasUns = 0.05,
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 2,
    covEvMax = 2,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos,
      bw_mtd,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

saveRDS(
  bw_list_raw_mtd_loop,
  file = projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "est",
    "bw_list_raw_mtd_loop.rds"
  )
)

bw_list_raw_mtd_loop |>
  dplyr::bind_rows() |>
  dplyr::group_by(
    transformation, prob_response, n_cell,
    mean_pos, bw_mtd, chnl,
    n_cell_stim, n_cell_uns
  ) |>
  dplyr::summarise(
    mean_bw_stim = mean(bw_stim),
    mean_bw_uns = mean(bw_uns),
    mean_bw = mean(bw),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation, mean_bw
  )
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel

if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache",
  "sim",
  "bw",
  "est",
  "progress.txt"
)

if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

future::plan(
  future::multisession,
  workers = max(1, future::availableCores() - 1)
)

bw_list_raw_mtd <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(transformation, prob_response, n_cell, mean_pos, bw_mtd, sim_id) {
      settings_row <- paste0(
        "trans = ",
        transformation,
        " | prob = ",
        prob_response,
        " | n_cell = ",
        n_cell,
        " | mean = ",
        mean_pos,
        " | bw_mtd = ",
        bw_mtd,
        " | sim_id = ",
        sim_id
      )

      log_msg <- paste0("Running: ", settings_row, "\n")
      cat(log_msg, file = path_progress_file, append = TRUE)

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthEstBwDirect(
            nSample = 50,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 1,
            biasUns = 0.05,
            bwMtd = bw_mtd,
            bwMin = -Inf,
            bwMax = Inf,
            bwCluster = 0.5,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            backgroundRelativeToResponse = 0.2,
            ncellUnsRelativeToStim = 1,
            covEvMin = 2,
            covEvMax = 2,
            tolClust = NULL,
            summarise = FALSE
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos,
              bw_mtd,
              sim_id,
              dplyr::everything()
            )

          cat(
            "Completed: ",
            settings_row,
            "\n",
            file = path_progress_file,
            append = TRUE
          )

          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          err_msg <- paste0("Error [", settings_row, "]: ", e$message, "\n")
          cat(err_msg, file = path_progress_file, append = TRUE)

          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos = mean_pos,
            bw_mtd = bw_mtd,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            bw_uns = NA_real_,
            bw_stim = NA_real_,
            bw = NA_real_,
            bw_source = NA_character_,
            error = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

saveRDS(
  bw_list_raw_mtd,
  file = projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "est",
    "bw_list_raw_mtd.rds"
  )
)
```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true
bw_list_raw_mtd <- readRDS(
  file = projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "est",
    "bw_list_raw_mtd.rds"
  )
)
bw_list_raw_mtd |>
  dplyr::bind_rows() |>
  dplyr::group_by(
    transformation, prob_response, n_cell,
    mean_pos, bw_mtd, chnl,
    n_cell_stim, n_cell_uns
  ) |>
  print(width )
  dplyr::summarise(
    mean_bw_stim = mean(bw_stim),
    mean_bw_uns = mean(bw_uns),
    mean_bw = mean(bw),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation, mean_bw
  )
```


