---
title: Simulate normalised bandwidth estimators
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
  sim_retry_errors: true
---

This analysis compares ordinary and normalised HPI bandwidth estimators as cell count, response frequency and separation vary. Matched stimulated and unstimulated samples are simulated under three transformations. Methods and cell-count caps within each biological scenario use the same simulated dataset.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
analysis_qmd <- "analysis/4-sim-bw-est-norm.qmd"

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
# On resume, rerun simulations whose saved output recorded an error.
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "est", "norm")
analysis_semantics_version <- "bandwidth-est-norm-v4"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
analysis_key <- .analysis_mode_key(analysis_key)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-bandwidth.R for simulation utilities.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

Get the parameters over which we simulate.

We compare the ordinary and normalised HPI estimators across parent-population size, response frequency, response separation and transformation.

We also vary the maximum number of cells used by the bandwidth estimator. For a fair comparison, estimator methods and cell caps within the same biological scenario use the same simulated dataset.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 5e3, 2e4, 1e5)
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")

# Compare each HPI derivative estimator with its normalised counterpart.
bandwidth_method_vec <- c(
  "hpi1", "hpi1Norm",
  "hpi2", "hpi2Norm",
  "hpi3", "hpi3Norm"
)

bandwidth_ncell_upper_vec <- c(
  500,
  2500,
  10000,
  50000,
  Inf
)

# Only the "low" unstimulated-sample bias setting is studied here, with
# transformation-specific magnitudes.
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05
)
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = bandwidth_method_vec,
  bw_ncell_upper = bandwidth_ncell_upper_vec,
  bias_uns = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns
  ) |>
  # A finite cap is only informative when it is smaller than the
  # available sample size. Always retain the uncapped estimate.
  dplyr::filter(
    is.infinite(.data$bw_ncell_upper) |
      .data$bw_ncell_upper < .data$n_cell
  ) |>
  dplyr::arrange(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bias_uns_setting,
    bias_uns,
    bw_mtd,
    bw_ncell_upper
  ) |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bias_uns_setting,
    bias_uns
  ) |>
  dplyr::mutate(
    # Estimator/cap comparisons must use exactly the same simulated data.
    sim_seed = as.integer(simulation_seed + dplyr::cur_group_id() - 1L)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

# sim_id and sim_seed are fixed on the full grid, before dev/quick filters,
# shuffling and chunking, so each row keeps its seed in every run.
sim_grid_full <- sim_grid

# Quick mode keeps cells, HPI method, normalisation, transformation, response and cap.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(1e3, 5e3),
      prob_response %in% c(0.002, 0.2),
      transformation %in% c("gaussian", "gamma"),
      mean_pos_setting == "high",
      bw_mtd %in% c("hpi1", "hpi1Norm", "hpi2", "hpi2Norm"),
      bw_ncell_upper %in% c(500, Inf)
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 5e3,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw_mtd %in% c("hpi1", "hpi1Norm"),
      bw_ncell_upper %in% c(2500, Inf)
    )
}

# Keep the complete selected grid for validation and collation.
sim_grid_all <- sim_grid
if (nrow(sim_grid_all) == 0L) {
  stop("The analysis 4 simulation grid is empty; check the dev filter.")
}
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

# Randomise execution order, then distribute rows evenly across external chunks.
sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
  dplyr::select(-.sim_grid_chunk)

sim_grid

print(
  paste0(
    "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
    " simulations for sim_grid chunk ", sim_grid_chunk_index,
    " of ", sim_grid_n_chunks,
    " using shuffle seed ", sim_grid_shuffle_seed
  )
)
```

Fixed settings passed to `.simBandwidthEstBwDirect()` for every grid row, and
the manifest parameters that canonical results must match.

```{r}
#| label: bw-estimate-settings
#| include: false
n_sample_sim <- if (analysis_quick) 1L else 10L
n_iter_sim <- 1L

scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = n_iter_sim,
  bwFallback = NA_real_,
  bwMin = -Inf,
  bwMax = Inf,
  bwCluster = 0.5,
  capStimRange = FALSE,
  probExact = TRUE,
  backgroundRelativeToResponse = 0.2,
  ncellUnsRelativeToStim = 1,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = NULL,
  summarise = FALSE
)

analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

## Simulation

Rerun one simulation interactively, for example to see why it failed.
Each row is seeded from its own `sim_seed` with a fixed RNG kind inside
`.simBandwidthRunRow()`, which the parallel run also calls, so this
reproduces the random numbers of that row exactly, whatever the chunking,
shuffling or worker scheduling. Call `debugonce()` on any helper first to
step through it.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L

row_target <- sim_grid_full[sim_grid_full$sim_id == sim_id_target, ]
rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthEstNormScenario,
  settings = scenario_settings
)

# Optionally compare with the stored output for this sim_id. Point `dir_run`
# at a staged run directory instead to check a run that was not promoted.
dir_run <- .analysis_cache_dir(
  c(analysis_key, "current"),
  path_root = root_dir,
  create = FALSE
)
path_stored <- .simBandwidthFindSimOutput(dir_run, sim_id_target)
if (length(path_stored) == 1L) {
  all.equal(rerun_res, readRDS(path_stored))
}
```

Run the simulations in parallel, with progress reporting, then validate this
chunk and promote the run once every chunk is complete.

```{r}
#| label: bw-estimate-parallel
#| include: false
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthEstNormScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS",
    path_root = root_dir
  )

  promoted <- .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    validate_fn = .simBandwidthEstNormValidator(n_sample_sim * n_iter_sim),
    collate_fn = function(tbl) {
      .simBandwidthEstNormCollate(tbl, colnames(sim_grid_all))
    },
    label = "analysis 4"
  )

  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    run_plots <- FALSE
  } else if (isTRUE(run_plots) && !isTRUE(promoted)) {
    message("Results were not promoted, so there is nothing to plot.")
    knitr::knit_exit()
  }
} else if (isTRUE(run_plots)) {
  # Read-only: plots use the canonical promoted results under current/.
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}
```

## Collation

Read the canonical promoted results for plotting (both run contexts point
at the same `current/` directory). In the collated `bw_tbl_results`,
`mean_bw` is the mean over complete finite stim/unstim pairs of
`pmin(bw_stim, bw_uns)`.

```{r}
#| label: bw-estimate-collate
#| echo: false
if (isTRUE(run_plots)) {
  bw_tbl_results <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results.rds"),
    required_params = analysis_required_params
  )
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

## Results

Plot the mean bandwidth against the number of cells.

The cell-count axis is logarithmic. Colours distinguish the HPI methods, while line types and point shapes distinguish requested normalisation; facets separate transformations and response frequencies. Each figure fixes response separation, bias and cell-count cap, so compare the ordinary and normalised curves within each panel.

```{r}
#| label: fig-bw-estimate-plots
#| echo: false
#| fig-cap: "Mean ordinary and normalised HPI bandwidth estimates at each cell-count cap."
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  # bw_mtd_norm describes the requested estimator. Normalised estimators can
  # fall back internally to the corresponding ordinary HPI estimator;
  # norm_fallback_rate records how often the selected final bandwidth did so.
  for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
    for (bias_uns_setting_curr in unique(bw_tbl_results$bias_uns_setting)) {
      for (bw_ncell_upper_curr in unique(bw_tbl_results$bw_ncell_upper)) {
        plot_tbl <- bw_tbl_results |>
          dplyr::filter(
            mean_pos_setting == !!mean_pos_setting_curr,
            bias_uns_setting == !!bias_uns_setting_curr,
            bw_ncell_upper == !!bw_ncell_upper_curr
          )

        if (nrow(plot_tbl) == 0L) next

        p <- ggplot(
          plot_tbl,
          aes(
            x = n_cell,
            y = mean_bw,
            colour = bw_mtd_base
          )
        ) +
          geom_line(
            aes(linetype = bw_mtd_norm),
            linewidth = 0.7,
            alpha = 0.75
          ) +
          geom_point(
            aes(shape = bw_mtd_norm),
            size = 1.7,
            alpha = 0.75
          ) +
          scale_x_log10() +
          facet_grid(
            transformation ~ prob_response,
            scales = "free_y"
          ) +
          labs(
            title = paste0(
              "Mean position setting: ", mean_pos_setting_curr,
              ", bias_uns_setting: ", bias_uns_setting_curr,
              ", bw_ncell_upper: ", bw_ncell_upper_curr
            ),
            x = "Number of cells",
            y = "Estimated bandwidth",
            colour = "Bandwidth method",
            linetype = "Requested normalisation",
            shape = "Requested normalisation"
          ) +
          cowplot::theme_cowplot() +
          cowplot::background_grid(major = "xy") +
          theme(
            panel.background = element_rect(fill = "white", colour = NA),
            plot.background = element_rect(fill = "white", colour = NA),
            legend.position = "bottom",
            legend.box = "vertical"
          )

        path_plot <- file.path(
          .analysis_cache_dir(c(analysis_key, "plots"), root_dir),
          paste0(
            "bw_estimate_meanpos_setting_", mean_pos_setting_curr,
            "_bias_uns_", bias_uns_setting_curr,
            "_ncell_upper_", bw_ncell_upper_curr,
            ".pdf"
          )
        )

        ggsave(
          filename = path_plot,
          plot = p,
          width = 25,
          height = 20,
          units = "cm"
        )
        print(p)
      }
    }
  }
}
```
