---
title: Simulate normalised bandwidth estimators
format:
  html:
    embed-resources: true
    toc: true
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "est", "norm")
analysis_dev <- isTRUE(.isDev())
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-bandwidth.R for simulation utilities.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Checking what the estimators actually estimate

Get the parameters over which we simulate.

We compare the ordinary and normalised HPI estimators across parent-population size, response frequency, response separation and transformation.

We also vary the maximum number of cells used by the bandwidth estimator. For a fair comparison, estimator methods and cell caps within the same biological scenario use the same simulated dataset.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 5e3, 2e4, 1e5)
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")

# Compare each HPI derivative estimator with its normalised counterpart.
bandwidth_method_vec <- c(
  "hpi1", "hpi1Norm",
  "hpi2", "hpi2Norm",
  "hpi3", "hpi3Norm"
)

bandwidth_ncell_upper_vec <- c(
  500,
  2500,
  10000,
  50000,
  Inf
)
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    bias_uns_setting == "low"
  )
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = bandwidth_method_vec,
  bw_ncell_upper = bandwidth_ncell_upper_vec,
  bias_uns = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns
  ) |>
  # A finite cap is only informative when it is smaller than the
  # available sample size. Always retain the uncapped estimate.
  dplyr::filter(
    is.infinite(.data$bw_ncell_upper) |
      .data$bw_ncell_upper < .data$n_cell
  ) |>
  dplyr::arrange(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bias_uns_setting,
    bias_uns,
    bw_mtd,
    bw_ncell_upper
  ) |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bias_uns_setting,
    bias_uns
  ) |>
  dplyr::mutate(
    # Estimator/cap comparisons must use exactly the same simulated data.
    sim_seed = as.integer(simulation_seed + dplyr::cur_group_id() - 1L)
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 5e3,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw_mtd %in% c("hpi1", "hpi1Norm"),
      bw_ncell_upper %in% c(2500, Inf)
    )
}

sim_grid_all <- sim_grid
sim_grid_definition <- sim_grid_all |>
  dplyr::select(-.data$sim_id, -.data$sim_seed)

if (nrow(sim_grid_all) > 0L) {
  set.seed(sim_grid_shuffle_seed)
  sim_grid <- sim_grid_all |>
    dplyr::mutate(
      .shuffle_order = sample.int(dplyr::n())
    ) |>
    dplyr::arrange(.data$.shuffle_order) |>
    dplyr::mutate(
      .sim_grid_chunk = ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
    ) |>
    dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
    dplyr::arrange(
      .data$n_cell,
      dplyr::desc(.data$prob_response),
      .data$transformation,
      .data$mean_pos_setting,
      .data$bw_mtd,
      .data$bw_ncell_upper
    ) |>
    dplyr::select(-.shuffle_order, -.sim_grid_chunk)
}

sim_grid

print(
  paste0(
    "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
    " simulations for sim_grid chunk ", sim_grid_chunk_index,
    " of ", sim_grid_n_chunks,
    " using shuffle seed ", sim_grid_shuffle_seed
  )
)

bw_fallback <- 0.23482348792138919129198282389
analysis_semantics_version <- "bandwidth-est-norm-v2"
n_sample_sim <- 10L
n_iter_sim <- 1L
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
devtools::load_all()
bw_list_raw_mtd_loop <- lapply(seq_len(nrow(sim_grid)), function(i) {
  NULL
})

for (i in seq_len(nrow(sim_grid))) {
  set.seed(as.integer(sim_grid$sim_seed[i]))
  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | bw_ncell_upper = ",
    sim_grid$bw_ncell_upper[i],
    " | bias_uns_setting = ",
    sim_grid$bias_uns_setting[i],
    " | bias_uns = ",
    sim_grid$bias_uns[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)
  bw_list_raw_mtd_loop[[i]] <- .simBandwidthEstBwDirect(
    nSample = n_sample_sim,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = n_iter_sim,
    biasUns = sim_grid$bias_uns[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwNcellMax = sim_grid$bw_ncell_upper[i],
    bwFallback = bw_fallback,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_upper = sim_grid$bw_ncell_upper[i],
      bias_uns_setting = sim_grid$bias_uns_setting[i],
      bias_uns = sim_grid$bias_uns[i],
      sim_seed = sim_grid$sim_seed[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_upper,
      bias_uns_setting,
      bias_uns,
      sim_seed,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

.write_rds_atomic(
  bw_list_raw_mtd_loop,
  projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "norm",
    "bw_list_raw_mtd_loop.rds"
  )
)

bw_list_raw_mtd_loop |>
  dplyr::bind_rows() |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    chnl,
    n_cell_stim,
    n_cell_uns
  ) |>
  dplyr::summarise(
    n_est = sum(is.finite(bw)),
    mean_bw_stim = .simBandwidthFiniteMean(bw_stim),
    mean_bw_uns = .simBandwidthFiniteMean(bw_uns),
    mean_bw = .simBandwidthFiniteMean(bw),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    n_est,
    mean_bw_stim,
    mean_bw_uns,
    mean_bw
  ) |>
  print(width = Inf)
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel

if (interactive()) {
  progressr::handlers(global = TRUE)
}

analysis_params <- list(
  run_simulations = run_simulations,
  run_plots = run_plots,
  simulation_seed = simulation_seed,
  analysis_dev = analysis_dev,
  sim_grid_definition = sim_grid_definition,
  sim_grid_chunk_index = sim_grid_chunk_index,
  sim_grid_n_chunks = sim_grid_n_chunks,
  sim_grid_shuffle_seed = sim_grid_shuffle_seed,
  analysis_semantics_version = analysis_semantics_version,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  bw_fallback = bw_fallback
)

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    path_root = root_dir,
    params = analysis_params,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_jobs_chunk <- run_ctx$chunk_jobs_dir
  dir_output <- run_ctx$chunk_output_dir

  dir.create(dir_jobs_chunk, recursive = TRUE, showWarnings = FALSE)
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir
  )
  message("Skipping simulations because run_simulations is FALSE.")
}

bw_list_raw_mtd <- list()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  old_plan <- future::plan()
  future::plan(
    future::multisession,
    workers = max(1, .simGetCores())
  )

  total_sims <- nrow(sim_grid)

  bw_list_raw_mtd <- tryCatch(
    progressr::with_progress({
      p <- progressr::progressor(steps = total_sims)

      furrr::future_pmap(
        sim_grid,
        function(transformation, prob_response, n_cell, mean_pos_setting,
                 mean_pos, bw_mtd, bw_ncell_upper, bias_uns_setting,
                 bias_uns, sim_seed, sim_id) {
          .simBandwidthEnsureCurrentCheckout(root_dir)
          set.seed(as.integer(sim_seed))

          settings_row <- paste0(
            "trans = ", transformation,
            " | prob = ", prob_response,
            " | n_cell = ", n_cell,
            " | mean = ", mean_pos,
            " | bw_mtd = ", bw_mtd,
            " | bw_ncell_upper = ", bw_ncell_upper,
            " | bias_uns_setting = ", bias_uns_setting,
            " | bias_uns = ", bias_uns,
            " | sim_seed = ", sim_seed,
            " | sim_id = ", sim_id,
            " | chunk = ", sim_grid_chunk_index,
            " of ", sim_grid_n_chunks
          )

          file_running <- file.path(dir_jobs_chunk, paste0("running-", sim_id))
          file_completed <- file.path(dir_jobs_chunk, paste0("completed-", sim_id))
          file_error <- file.path(dir_jobs_chunk, paste0("error-", sim_id))
          file_output <- .path_sim_output(
            sim_id,
            dir_output = dir_output,
            sim_grid_chunk_index = sim_grid_chunk_index,
            sim_grid_n_chunks = sim_grid_n_chunks
          )

          existing_output <- if (file.exists(file_output)) {
            tryCatch(readRDS(file_output), error = function(e) NULL)
          } else {
            NULL
          }

          if (!is.null(existing_output)) {
            .analysis_reconcile_resume_markers(
              existing_output = existing_output,
              file_completed = file_completed,
              file_error = file_error,
              file_running = file_running,
              error_col = "error_message"
            )
            .update_progress_summary(
              path_progress_file = path_progress_file,
              dir_jobs_chunk = dir_jobs_chunk,
              total_sims = total_sims,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              dir_output = dir_output,
              heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
            )
            p(sprintf("Skipped existing sim_id: %s", sim_id))
            return(existing_output)
          }

          file.create(file_running)
          .update_progress_summary(
            path_progress_file = path_progress_file,
            dir_jobs_chunk = dir_jobs_chunk,
            total_sims = total_sims,
            sim_grid_chunk_index = sim_grid_chunk_index,
            sim_grid_n_chunks = sim_grid_n_chunks,
            dir_output = dir_output,
            heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
          )

          out <- tryCatch(
            {
              sim_res <- .simBandwidthEstBwDirect(
                nSample = n_sample_sim,
                nMarker = 1,
                nCondition = 2,
                nCluster = 2,
                nIter = n_iter_sim,
                biasUns = bias_uns,
                bwMtd = bw_mtd,
                bwFallback = bw_fallback,
                bwMin = -Inf,
                bwMax = Inf,
                bwNcellMax = bw_ncell_upper,
                bwCluster = 0.5,
                nCellStim = n_cell,
                probResponse = prob_response,
                probExact = TRUE,
                meanPos = mean_pos,
                transformation = transformation,
                backgroundRelativeToResponse = 0.2,
                ncellUnsRelativeToStim = 1,
                covEvMin = 1.5,
                covEvMax = 1.5,
                tolClust = NULL,
                summarise = FALSE
              ) |>
                dplyr::mutate(
                  transformation = transformation,
                  prob_response = prob_response,
                  n_cell = n_cell,
                  mean_pos_setting = mean_pos_setting,
                  mean_pos = mean_pos,
                  bw_mtd = bw_mtd,
                  bw_ncell_upper = bw_ncell_upper,
                  bias_uns_setting = bias_uns_setting,
                  bias_uns = bias_uns,
                  sim_seed = sim_seed,
                  sim_id = sim_id,
                  sim_grid_chunk_index = sim_grid_chunk_index,
                  sim_grid_n_chunks = sim_grid_n_chunks,
                  error_message = NA_character_
                ) |>
                dplyr::select(
                  transformation,
                  prob_response,
                  n_cell,
                  mean_pos_setting,
                  mean_pos,
                  bw_mtd,
                  bw_ncell_upper,
                  bias_uns_setting,
                  bias_uns,
                  sim_seed,
                  sim_id,
                  sim_grid_chunk_index,
                  sim_grid_n_chunks,
                  dplyr::everything()
                )

              .write_rds_atomic(sim_res, file_output)
              file.create(file_completed)
              if (file.exists(file_running)) {
                file.remove(file_running)
              }

              .update_progress_summary(
                path_progress_file = path_progress_file,
                dir_jobs_chunk = dir_jobs_chunk,
                total_sims = total_sims,
                sim_grid_chunk_index = sim_grid_chunk_index,
                sim_grid_n_chunks = sim_grid_n_chunks,
                dir_output = dir_output,
                heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
              )
              p(sprintf("Completed sim_id: %s", sim_id))

              sim_res
            },
            error = function(e) {
              err_res <- tibble::tibble(
                transformation = transformation,
                prob_response = prob_response,
                n_cell = n_cell,
                mean_pos_setting = mean_pos_setting,
                mean_pos = mean_pos,
                bw_mtd = bw_mtd,
                bw_ncell_upper = bw_ncell_upper,
                bias_uns_setting = bias_uns_setting,
                bias_uns = bias_uns,
                sim_seed = sim_seed,
                sim_id = sim_id,
                sim_grid_chunk_index = sim_grid_chunk_index,
                sim_grid_n_chunks = sim_grid_n_chunks,
                iter = NA_character_,
                sample = NA_character_,
                ind = NA_character_,
                chnl = NA_character_,
                n_cell_uns = NA_integer_,
                n_cell_stim = NA_integer_,
                n_uns_bw = NA_integer_,
                n_stim_bw = NA_integer_,
                max_dens_x = NA_real_,
                bw_uns = NA_real_,
                bw_stim = NA_real_,
                bw = NA_real_,
                bw_source = NA_character_,
                error_message = conditionMessage(e)
              )

              .write_rds_atomic(err_res, file_output)
              file.create(file_error)
              if (file.exists(file_running)) {
                file.remove(file_running)
              }
              .update_progress_summary(
                path_progress_file = path_progress_file,
                dir_jobs_chunk = dir_jobs_chunk,
                total_sims = total_sims,
                sim_grid_chunk_index = sim_grid_chunk_index,
                sim_grid_n_chunks = sim_grid_n_chunks,
                dir_output = dir_output,
                heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
              )
              p(sprintf("ERROR on sim_id: %s", sim_id))

              err_res
            }
          )

          out
        },
        .options = furrr::furrr_options(
          seed = TRUE,
          chunk_size = 10L
        )
      )
    }),
    finally = future::plan(old_plan)
  )

  .write_rds_atomic(
    bw_list_raw_mtd,
    file.path(
      run_ctx$staging_collated_dir,
      sprintf(
        "bw_list_raw_mtd_chunk_%03d-of_%03d.rds",
        sim_grid_chunk_index,
        sim_grid_n_chunks
      )
    )
  )
} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
}

```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true

summarise_bw_results <- function(bw_tbl) {
  bw_tbl_results <- bw_tbl |>
    dplyr::group_by(
      dplyr::pick(dplyr::any_of(colnames(sim_grid_all)))
    ) |>
    dplyr::summarise(
      n_total = dplyr::n(),
      n_est = sum(is.finite(.data$bw)),
      mean_bw_stim = .simBandwidthFiniteMean(.data$bw_stim),
      mean_bw_uns = .simBandwidthFiniteMean(.data$bw_uns),
      mean_bw = .simBandwidthFiniteMean(.data$bw),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      estimate_rate = .data$n_est / .data$n_total,
      bw_mtd_base = gsub("Norm$", "", .data$bw_mtd),
      bw_mtd_norm = ifelse(
        grepl("Norm$", .data$bw_mtd),
        "norm",
        "non-norm"
      )
    )

  if (any(bw_tbl_results$n_est == 0L)) {
    stop(
      "At least one requested bandwidth estimator has zero finite values. ",
      "Review the corresponding raw rows before plotting."
    )
  }

  if (any(bw_tbl_results$n_est < bw_tbl_results$n_total)) {
    warning(
      "Some estimator/scenario combinations have non-finite bandwidths; ",
      "estimate_rate records the fraction retained in each mean."
    )
  }

  bw_tbl_results
}

promote_analysis4_if_ready <- function(
    run_ctx,
    sim_grid_all,
    total_sims,
    completed_sims,
    failed_sims) {
  if (!.analysis_can_promote(run_ctx)) {
    return(NULL)
  }

  full_output_paths <- .find_bw_list_output_files(
    output_dir = run_ctx$staging_run_dir,
    allow_cache_fallback = FALSE,
    merge_dirs = FALSE
  )
  full_output_ids <- suppressWarnings(as.integer(sub(
    "^.*-sim_id_([0-9]+)[.]rds$",
    "\\1",
    basename(full_output_paths)
  )))
  expected_sim_ids <- sort(unique(as.integer(sim_grid_all$sim_id)))
  full_outputs_ok <-
    length(full_output_paths) == length(expected_sim_ids) &&
    identical(sort(unique(full_output_ids)), expected_sim_ids)

  if (!isTRUE(full_outputs_ok)) {
    error_message <- paste0(
      "Refusing to promote analysis 4: staged outputs do not contain ",
      "exactly the complete simulation grid."
    )
    .analysis_mark_chunk(
      run_ctx = run_ctx,
      total_sims = total_sims,
      completed_sims = completed_sims,
      failed_sims = failed_sims,
      collate_ok = FALSE,
      validation_ok = FALSE,
      error_message = error_message
    )
    stop(error_message)
  }

  full_bw_tbl <- purrr::map(full_output_paths, readRDS) |>
    purrr::list_rbind()

  full_error_ids <- sort(unique(as.integer(
    full_bw_tbl$sim_id[
      !is.na(full_bw_tbl$error_message) &
        nzchar(as.character(full_bw_tbl$error_message))
    ]
  )))

  expected_rows_per_sim <- as.integer(n_sample_sim * n_iter_sim)
  full_row_counts <- full_bw_tbl |>
    dplyr::count(.data$sim_id, name = "n_rows")
  full_rows_ok <-
    nrow(full_row_counts) == length(expected_sim_ids) &&
    all(full_row_counts$n_rows == expected_rows_per_sim)

  if (length(full_error_ids) > 0L || !isTRUE(full_rows_ok)) {
    error_message <- if (length(full_error_ids) > 0L) {
      paste0(
        "Refusing to promote analysis 4: simulation errors remain for sim_id ",
        paste(full_error_ids, collapse = ", "),
        "."
      )
    } else {
      paste0(
        "Refusing to promote analysis 4: one or more scenarios do not have ",
        expected_rows_per_sim,
        " sample-level estimator rows."
      )
    }
    .analysis_mark_chunk(
      run_ctx = run_ctx,
      total_sims = total_sims,
      completed_sims = completed_sims,
      failed_sims = failed_sims,
      collate_ok = TRUE,
      validation_ok = FALSE,
      error_message = error_message
    )
    stop(error_message)
  }

  full_bw_tbl_results <- summarise_bw_results(full_bw_tbl)

  .write_rds_atomic(
    full_bw_tbl,
    file.path(run_ctx$staging_collated_dir, "bw_list_raw_mtd.rds")
  )
  .write_rds_atomic(
    full_bw_tbl_results,
    file.path(run_ctx$staging_collated_dir, "bw_tbl_results.rds")
  )
  .analysis_promote_run(run_ctx)

  full_bw_tbl_results
}

results_read_only <- isTRUE(run_ctx$read_only)

if (results_read_only) {
  path_results <- .analysis_current_file(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results.rds"),
    required_params = list(
      analysis_semantics_version = analysis_semantics_version,
      simulation_seed = simulation_seed,
      analysis_dev = analysis_dev,
      sim_grid_definition = sim_grid_definition,
      n_sample_sim = n_sample_sim,
      n_iter_sim = n_iter_sim
    )
  )
  bw_tbl_results <- readRDS(path_results)
} else if (nrow(sim_grid) == 0L) {
  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = 0L,
    completed_sims = 0L,
    failed_sims = 0L,
    collate_ok = TRUE,
    validation_ok = TRUE
  )

  invisible(promote_analysis4_if_ready(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = 0L,
    completed_sims = 0L,
    failed_sims = 0L
  ))

  message("No simulations were assigned to this chunk; marked it complete.")
  knitr::knit_exit()
} else {
  bw_list_path_vec <- .find_bw_list_output_files(
    output_dir = run_ctx$chunk_dir,
    allow_cache_fallback = FALSE,
    merge_dirs = FALSE
  )

  if (length(bw_list_path_vec) == 0L) {
    stop("No saved simulation results were available for the current chunk.")
  }

  bw_tbl <- purrr::map(
    bw_list_path_vec,
    function(path) {
      tryCatch(
        readRDS(path),
        error = function(e) {
          warning("Could not read saved simulation result: ", path)
          NULL
        }
      )
    }
  ) |>
    purrr::compact() |>
    purrr::list_rbind()

  expected_chunk_ids <- sort(unique(as.integer(sim_grid$sim_id)))
  collated_chunk_ids <- sort(unique(as.integer(bw_tbl$sim_id)))
  output_error_ids <- sort(unique(as.integer(
    bw_tbl$sim_id[
      !is.na(bw_tbl$error_message) &
        nzchar(as.character(bw_tbl$error_message))
    ]
  )))

  expected_rows_per_sim <- as.integer(n_sample_sim * n_iter_sim)
  chunk_row_counts <- bw_tbl |>
    dplyr::count(.data$sim_id, name = "n_rows")
  chunk_rows_ok <-
    nrow(chunk_row_counts) == length(expected_chunk_ids) &&
    all(chunk_row_counts$n_rows == expected_rows_per_sim)

  chunk_collate_ok <-
    length(bw_list_path_vec) == length(expected_chunk_ids) &&
    identical(collated_chunk_ids, expected_chunk_ids)
  chunk_validation_ok <-
    chunk_collate_ok &&
    chunk_rows_ok &&
    length(output_error_ids) == 0L

  validation_error <- if (!chunk_collate_ok) {
    "Chunk collation did not contain exactly the assigned sim_id set."
  } else if (length(output_error_ids) > 0L) {
    paste0(
      "Chunk contains simulation errors for sim_id: ",
      paste(output_error_ids, collapse = ", "),
      "."
    )
  } else if (!chunk_rows_ok) {
    paste0(
      "At least one scenario did not produce exactly ",
      expected_rows_per_sim,
      " sample-level estimator rows."
    )
  } else {
    NULL
  }

  bw_tbl_results <- summarise_bw_results(bw_tbl)

  collate_suffix <- if (sim_grid_n_chunks > 1L) {
    sprintf("_chunk_%03d-of_%03d", sim_grid_chunk_index, sim_grid_n_chunks)
  } else {
    ""
  }
  .write_rds_atomic(
    bw_tbl_results,
    file.path(
      run_ctx$staging_collated_dir,
      paste0("bw_tbl_results", collate_suffix, ".rds")
    )
  )

  chunk_job_files <- list.files(dir_jobs_chunk, full.names = FALSE)
  chunk_completed_ids <- unique(sub(
    "^completed-",
    "",
    chunk_job_files[grepl("^completed-", chunk_job_files)]
  ))
  chunk_failed_ids <- unique(sub(
    "^error-",
    "",
    chunk_job_files[grepl("^error-", chunk_job_files)]
  ))
  chunk_completed_ids <- setdiff(chunk_completed_ids, chunk_failed_ids)

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = length(chunk_completed_ids),
    failed_sims = length(chunk_failed_ids),
    collate_ok = chunk_collate_ok,
    validation_ok = chunk_validation_ok,
    error_message = validation_error
  )

  if (!isTRUE(chunk_validation_ok) || length(chunk_failed_ids) > 0L) {
    stop(
      if (!is.null(validation_error)) {
        validation_error
      } else {
        paste0(
          "Chunk contains ",
          length(chunk_failed_ids),
          " failed simulation(s)."
        )
      }
    )
  }

  promoted_results <- promote_analysis4_if_ready(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = length(chunk_completed_ids),
    failed_sims = length(chunk_failed_ids)
  )
  if (!is.null(promoted_results)) {
    bw_tbl_results <- promoted_results
  }
}

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after simulation/collation.")
  knitr::knit_exit()
}

if (!results_read_only && sim_grid_n_chunks > 1L) {
  message(
    "Skipping plots during a multi-chunk simulation render. ",
    "Render again with run_simulations = FALSE after promotion."
  )
  knitr::knit_exit()
}

# bw_mtd_norm describes the requested estimator. The current normalised
# bandwidth implementation may fall back internally to the corresponding
# ordinary HPI estimator; that fallback is not yet exposed in row-level
# provenance, so it cannot be separated in these plots.
for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
  for (bias_uns_setting_curr in unique(bw_tbl_results$bias_uns_setting)) {
    for (bw_ncell_upper_curr in unique(bw_tbl_results$bw_ncell_upper)) {
      plot_tbl <- bw_tbl_results |>
        dplyr::filter(
          mean_pos_setting == !!mean_pos_setting_curr,
          bias_uns_setting == !!bias_uns_setting_curr,
          bw_ncell_upper == !!bw_ncell_upper_curr
        )

      if (nrow(plot_tbl) == 0L) next

      p <- ggplot(
        plot_tbl,
        aes(
          x = n_cell,
          y = mean_bw,
          colour = bw_mtd_base
        )
      ) +
        geom_line(
          aes(linetype = bw_mtd_norm),
          linewidth = 0.7,
          alpha = 0.75
        ) +
        geom_point(
          aes(shape = bw_mtd_norm),
          size = 1.7,
          alpha = 0.75
        ) +
        scale_x_log10() +
        facet_grid(
          transformation ~ prob_response,
          scales = "free_y"
        ) +
        labs(
          title = paste0(
            "Mean position setting: ", mean_pos_setting_curr,
            ", bias_uns_setting: ", bias_uns_setting_curr,
            ", bw_ncell_upper: ", bw_ncell_upper_curr
          ),
          x = "Number of cells",
          y = "Estimated bandwidth",
          colour = "Bandwidth method",
          linetype = "Requested normalisation",
          shape = "Requested normalisation"
        ) +
        cowplot::theme_cowplot() +
        cowplot::background_grid(major = "xy") +
        theme(
          panel.background = element_rect(fill = "white", colour = NA),
          plot.background = element_rect(fill = "white", colour = NA),
          legend.position = "bottom",
          legend.box = "vertical"
        )

      path_plot <- projr::projr_path_get(
        "cache",
        "sim",
        "bw",
        "est",
        "norm",
        paste0(
          "bw_estimate_meanpos_setting_", mean_pos_setting_curr,
          "_bias_uns_", bias_uns_setting_curr,
          "_ncell_upper_", bw_ncell_upper_curr,
          ".pdf"
        )
      )
      dir.create(dirname(path_plot), recursive = TRUE, showWarnings = FALSE)

      ggsave(
        filename = path_plot,
        plot = p,
        width = 25,
        height = 20,
        units = "cm"
      )
      print(p)
    }
  }
}
```
