---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
params:
  analysis_run_id: null
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source simulation helpers in dependency order:
# sim-misc.R must be sourced before sim-bandwidth.R for simulation utilities.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Checking what the estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider two variations - whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator.
We can also vary the amount of separation, I suppose.

We'll need to vary over a range of parent population sizes, however, as that will tell us how the estimator behaves over the most crucial dimension (apart from the actual transformation, of course).

Initially, we will ignore setting maximum/minimum number of cells.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# bandwidth settings
bandwidth_method_vec <- c("hpi0Norm", "hpi1Norm", "hpi2Norm", "hpi3Norm", "sjNorm", "nrd0Norm", "hpi1", "hpi2", "hpi3", "sj", "nrd0")
bandwidth_method_vec <- bandwidth_method_vec[grepl("hpi", bandwidth_method_vec)]
bandwidth_method_vec <- bandwidth_method_vec[!grepl("hpi0", bandwidth_method_vec)]

# set this later
bandwidth_ncell_upper_vec <- c(
  500,
  2500,
  10000,
  50000,
  Inf
)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    bias_uns_setting == "low"
  )
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = bandwidth_method_vec,
  bw_ncell_upper = bandwidth_ncell_upper_vec,
  bias_uns = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns
  ) |>
  # A finite cap is only informative when it is smaller than the
  # available sample size. Always retain the uncapped estimate.
  dplyr::filter(
    is.infinite(.data$bw_ncell_upper) |
      .data$bw_ncell_upper < .data$n_cell
  ) |>
  dplyr::arrange(
    n_cell,
    dplyr::desc(prob_response)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

if (.isDev()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(5e3, 1e4, 2e4),
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      grepl("hpi1", bw_mtd),
      grepl("hpi1Norm", bw_mtd)
    ) |>
    dplyr::sample_n(
      min(5L, dplyr::n())
    )
}

sim_grid_all <- sim_grid

if (nrow(sim_grid_all) > 0L) {
  set.seed(sim_grid_shuffle_seed)
  sim_grid <- sim_grid_all |>
    dplyr::mutate(
      .shuffle_order = sample.int(dplyr::n())
    ) |>
    dplyr::arrange(.data$.shuffle_order) |>
    dplyr::mutate(
      .sim_grid_chunk = ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
    ) |>
    dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
    dplyr::arrange(
      .data$n_cell,
      dplyr::desc(.data$prob_response),
      .data$transformation,
      .data$mean_pos_setting,
      .data$bw_mtd,
      .data$bw_ncell_upper
    ) |>
    dplyr::select(-.shuffle_order, -.sim_grid_chunk)
}

sim_grid

print(
  paste0(
    "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
    " simulations for sim_grid chunk ", sim_grid_chunk_index,
    " of ", sim_grid_n_chunks,
    " using shuffle seed ", sim_grid_shuffle_seed
  )
)

bw_fallback <- 0.23482348792138919129198282389
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
devtools::load_all()
bw_list_raw_mtd_loop <- lapply(seq_len(nrow(sim_grid)), function(i) {
  NULL
})

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | bw_ncell_upper = ",
    sim_grid$bw_ncell_upper[i],
    " | bias_uns_setting = ",
    sim_grid$bias_uns_setting[i],
    " | bias_uns = ",
    sim_grid$bias_uns[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)
  bw_list_raw_mtd_loop[[i]] <- .simBandwidthEstBwDirect(
    nSample = if (.isDev()) 1 else 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 1,
    biasUns = sim_grid$bias_uns[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwNcellMax = sim_grid$bw_ncell_upper[i],
    bwFallback = bw_fallback,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_upper = sim_grid$bw_ncell_upper[i],
      bias_uns_setting = sim_grid$bias_uns_setting[i],
      bias_uns = sim_grid$bias_uns[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_upper,
      bias_uns_setting,
      bias_uns,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

saveRDS(
  bw_list_raw_mtd_loop,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "norm",
    "bw_list_raw_mtd_loop.rds"
  )
)

bw_list_raw_mtd_loop |>
  dplyr::bind_rows() |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    chnl,
    n_cell_stim,
    n_cell_uns
  ) |>
  dplyr::summarise(
    n_est = sum(is.finite(bw)),
    mean_bw_stim = .simBandwidthFiniteMean(bw_stim),
    mean_bw_uns = .simBandwidthFiniteMean(bw_uns),
    mean_bw = .simBandwidthFiniteMean(bw),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    n_est,
    mean_bw_stim,
    mean_bw_uns,
    mean_bw
  ) |>
  print(width = Inf)
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel

if (interactive()) {
  progressr::handlers(global = TRUE)
}

analysis_semantics_version <- "bandwidth-est-norm-v1"
n_sample_sim <- 10L
n_iter_sim <- 1L

run_ctx <- .analysis_run_context(
  analysis_key = c("sim", "bw", "est", "norm"),
  run_id = analysis_run_id,
  path_root = root_dir,
  params = list(
    run_simulations = run_simulations,
    run_plots = run_plots,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    sim_grid_shuffle_seed = sim_grid_shuffle_seed,
    analysis_semantics_version = analysis_semantics_version,
    n_sample_sim = n_sample_sim,
    n_iter_sim = n_iter_sim,
    bw_fallback = bw_fallback
  ),
  sim_grid_chunk_index = sim_grid_chunk_index,
  sim_grid_n_chunks = sim_grid_n_chunks
)

path_progress_file <- run_ctx$progress_file
dir_jobs_chunk <- run_ctx$chunk_jobs_dir
dir_output <- run_ctx$chunk_output_dir

dir.create(dir_jobs_chunk, recursive = TRUE, showWarnings = FALSE)
dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)

bw_list_raw_mtd <- list()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  future::plan(
    future::multisession,
    workers = max(1, .simGetCores())
  )

  total_sims <- nrow(sim_grid)

  bw_list_raw_mtd <- progressr::with_progress({
    p <- progressr::progressor(steps = total_sims)

    furrr::future_pmap(
      sim_grid,
      function(transformation, prob_response, n_cell, mean_pos_setting,
               mean_pos, bw_mtd, bw_ncell_upper, bias_uns_setting,
               bias_uns, sim_id) {
        .simBandwidthEnsureCurrentCheckout(root_dir)
        settings_row <- paste0(
          "trans = ",
          transformation,
          " | prob = ",
          prob_response,
          " | n_cell = ",
          n_cell,
          " | mean = ",
          mean_pos,
          " | bw_mtd = ",
          bw_mtd,
          " | bw_ncell_upper = ",
          bw_ncell_upper,
          " | bias_uns_setting = ",
          bias_uns_setting,
          " | bias_uns = ",
          bias_uns,
          " | sim_id = ",
          sim_id,
          " | chunk = ",
          sim_grid_chunk_index,
          " of ",
          sim_grid_n_chunks
        )

        file_running <- file.path(dir_jobs_chunk, paste0("running-", sim_id))
        file_completed <- file.path(dir_jobs_chunk, paste0("completed-", sim_id))
        file_error <- file.path(dir_jobs_chunk, paste0("error-", sim_id))
        file_output <- .path_sim_output(
          sim_id,
          dir_output = dir_output,
          sim_grid_chunk_index = sim_grid_chunk_index,
          sim_grid_n_chunks = sim_grid_n_chunks
        )

        existing_output <- if (file.exists(file_output)) {
          tryCatch(readRDS(file_output), error = function(e) NULL)
        } else {
          NULL
        }

        if (!is.null(existing_output)) {
          .analysis_reconcile_resume_markers(
            existing_output = existing_output,
            file_completed = file_completed,
            file_error = file_error,
            file_running = file_running,
            error_col = "error_message"
          )
          .update_progress_summary(
            path_progress_file = path_progress_file,
            dir_jobs_chunk = dir_jobs_chunk,
            total_sims = total_sims,
            sim_grid_chunk_index = sim_grid_chunk_index,
            sim_grid_n_chunks = sim_grid_n_chunks,
            dir_output = dir_output,
            heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
          )
          p(sprintf("Skipped existing sim_id: %s", sim_id))
          return(existing_output)
        }

        file.create(file_running)
        .update_progress_summary(
          path_progress_file = path_progress_file,
          dir_jobs_chunk = dir_jobs_chunk,
          total_sims = total_sims,
          sim_grid_chunk_index = sim_grid_chunk_index,
          sim_grid_n_chunks = sim_grid_n_chunks,
          dir_output = dir_output,
          heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
        )

        out <- tryCatch(
          {
            sim_res <- .simBandwidthEstBwDirect(
              nSample = n_sample_sim,
              nMarker = 1,
              nCondition = 2,
              nCluster = 2,
              nIter = n_iter_sim,
              biasUns = bias_uns,
              bwMtd = bw_mtd,
              bwFallback = bw_fallback,
              bwMin = -Inf,
              bwMax = Inf,
              bwNcellMax = bw_ncell_upper,
              bwCluster = 0.5,
              nCellStim = n_cell,
              probResponse = prob_response,
              probExact = TRUE,
              meanPos = mean_pos,
              transformation = transformation,
              backgroundRelativeToResponse = 0.2,
              ncellUnsRelativeToStim = 1,
              covEvMin = 1.5,
              covEvMax = 1.5,
              tolClust = NULL,
              summarise = FALSE
            ) |>
              dplyr::mutate(
                transformation = transformation,
                prob_response = prob_response,
                n_cell = n_cell,
                mean_pos_setting = mean_pos_setting,
                mean_pos = mean_pos,
                bw_mtd = bw_mtd,
                bw_ncell_upper = bw_ncell_upper,
                bias_uns_setting = bias_uns_setting,
                bias_uns = bias_uns,
                sim_id = sim_id,
                sim_grid_chunk_index = sim_grid_chunk_index,
                sim_grid_n_chunks = sim_grid_n_chunks,
                error_message = NA_character_
              ) |>
              dplyr::select(
                transformation,
                prob_response,
                n_cell,
                mean_pos_setting,
                mean_pos,
                bw_mtd,
                bw_ncell_upper,
                bias_uns_setting,
                bias_uns,
                sim_id,
                sim_grid_chunk_index,
                sim_grid_n_chunks,
                dplyr::everything()
              )

            .write_rds_atomic(sim_res, file_output)
            file.create(file_completed)
            if (file.exists(file_running)) {
              file.remove(file_running)
            }

            .update_progress_summary(
              path_progress_file = path_progress_file,
              dir_jobs_chunk = dir_jobs_chunk,
              total_sims = total_sims,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              dir_output = dir_output,
              heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
            )
            p(sprintf("Completed sim_id: %s", sim_id))

            sim_res
          },
          error = function(e) {
            file.create(file_error)
            if (file.exists(file_running)) {
              file.remove(file_running)
            }

            err_res <- tibble::tibble(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              bw_ncell_upper = bw_ncell_upper,
              bias_uns_setting = bias_uns_setting,
              bias_uns = bias_uns,
              sim_id = sim_id,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              iter = NA_character_,
              sample = NA_character_,
              ind = NA_character_,
              chnl = NA_character_,
              n_cell_uns = NA_integer_,
              n_cell_stim = NA_integer_,
              n_uns_bw = NA_integer_,
              n_stim_bw = NA_integer_,
              max_dens_x = NA_real_,
              bw_uns = NA_real_,
              bw_stim = NA_real_,
              bw = NA_real_,
              bw_source = NA_character_,
              error_message = e$message
            )

            .write_rds_atomic(err_res, file_output)
            .update_progress_summary(
              path_progress_file = path_progress_file,
              dir_jobs_chunk = dir_jobs_chunk,
              total_sims = total_sims,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              dir_output = dir_output,
              heading = "NORMALISED BANDWIDTH ESTIMATION PROGRESS"
            )
            p(sprintf("ERROR on sim_id: %s", sim_id))

            err_res
          }
        )

        out
      },
      .options = furrr::furrr_options(
        seed = TRUE,
        chunk_size = 10L
      )
    )
  })

  future::plan(future::sequential)

  .write_rds_atomic(
    bw_list_raw_mtd,
    file.path(
      run_ctx$staging_collated_dir,
      sprintf(
        "bw_list_raw_mtd_chunk_%03d-of_%03d.rds",
        sim_grid_chunk_index,
        sim_grid_n_chunks
      )
    )
  )
} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}

if (isTRUE(run_simulations)) {
  chunk_job_files <- list.files(dir_jobs_chunk, full.names = FALSE)
  chunk_completed_ids <- unique(sub(
    "^completed-",
    "",
    chunk_job_files[grepl("^completed-", chunk_job_files)]
  ))
  chunk_failed_ids <- unique(sub(
    "^error-",
    "",
    chunk_job_files[grepl("^error-", chunk_job_files)]
  ))
  chunk_completed_ids <- setdiff(chunk_completed_ids, chunk_failed_ids)
  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = length(chunk_completed_ids),
    failed_sims = length(chunk_failed_ids)
  )
}
```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true
bw_list_path_vec <- .find_bw_list_output_files(
  output_dir = run_ctx$staging_run_dir,
  cache_dir = run_ctx$current_dir,
  allow_cache_fallback = !isTRUE(run_simulations)
)

if (length(bw_list_path_vec) > 0L) {
  bw_list_raw_mtd_collate <- purrr::map(
    bw_list_path_vec,
    function(path) {
      tryCatch(
        readRDS(path),
        error = function(e) {
          warning("Could not read saved simulation result: ", path)
          NULL
        }
      )
    }
  ) |>
    purrr::compact()
} else if (length(bw_list_raw_mtd) > 0L) {
  bw_list_raw_mtd_collate <- bw_list_raw_mtd
} else {
  collated_dir <- if (isTRUE(run_simulations)) {
    run_ctx$staging_collated_dir
  } else {
    file.path(run_ctx$current_dir, "collated")
  }
  path_vec <- if (dir.exists(collated_dir)) {
    list.files(
      collated_dir,
      pattern = "^bw_list_raw_mtd(_chunk_[0-9]+-of_[0-9]+)?[.]rds$",
      full.names = TRUE
    )
  } else {
    character(0)
  }
  if (length(path_vec) == 0L) {
    stop("Could not find canonical normalised bandwidth-estimation results.")
  }
  bw_list_raw_mtd_collate <- readRDS(path_vec[[1]])
}

bw_tbl <- if (is.data.frame(bw_list_raw_mtd_collate)) {
  bw_list_raw_mtd_collate
} else {
  purrr::list_rbind(bw_list_raw_mtd_collate)
}

if (!"bias_uns" %in% colnames(bw_tbl)) {
  bw_tbl$bias_uns <- NA_real_
}
if (!"bias_uns_setting" %in% colnames(bw_tbl)) {
  bw_tbl$bias_uns_setting <- NA_character_
}

if (all(is.na(bw_tbl$bias_uns_setting)) && any(!is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-bias_uns_setting) |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns")
    )
}

bw_tbl <- bw_tbl |>
  dplyr::filter(!is.na(.data$bias_uns_setting))

if (nrow(bw_tbl) == 0L) {
  stop("No cached rows contain the current bias_uns settings. Rerun the simulation chunk.")
}

sim_grid_key_cols <- setdiff(
  intersect(colnames(sim_grid_all), colnames(bw_tbl)),
  "sim_id"
)

bw_tbl <- bw_tbl |>
  dplyr::semi_join(
    sim_grid_all |>
      dplyr::select(dplyr::all_of(sim_grid_key_cols)) |>
      dplyr::distinct(),
    by = sim_grid_key_cols
  )

if (nrow(bw_tbl) == 0L) {
  stop("Cached rows do not match the current simulation grid. Rerun the simulation chunk.")
}

bw_tbl_results <- bw_tbl |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid_all)))
  ) |>
  dplyr::summarise(
    n_est = sum(is.finite(bw)),
    mean_bw_stim = .simBandwidthFiniteMean(bw_stim),
    mean_bw_uns = .simBandwidthFiniteMean(bw_uns),
    mean_bw = .simBandwidthFiniteMean(bw),
    .groups = "drop"
  ) |>
  dplyr::select(
    dplyr::any_of(colnames(sim_grid_all)),
    n_est,
    mean_bw_stim,
    mean_bw_uns,
    mean_bw
  )

if (any(bw_tbl_results$n_est == 0L)) {
  stop(
    "At least one requested bandwidth estimator has zero finite values.",
    " Review the corresponding raw rows before plotting."
  )
}
bw_tbl_results <- bw_tbl_results |>
  dplyr::mutate(
    bw_mtd_base = gsub("Norm$", "", bw_mtd),
    bw_mtd_norm = grepl("Norm$", bw_mtd),
    bw_mtd_norm = ifelse(bw_mtd_norm, "norm", "non-norm")
  )

collated_sim_ids <- sort(unique(as.integer(bw_tbl$sim_id)))
chunk_collate_ok <- all(as.integer(sim_grid$sim_id) %in% collated_sim_ids)

collate_suffix <- if (sim_grid_n_chunks > 1L) {
  sprintf("_chunk_%03d-of_%03d", sim_grid_chunk_index, sim_grid_n_chunks)
} else {
  ""
}
.write_rds_atomic(
  bw_tbl_results,
  file.path(
    run_ctx$staging_collated_dir,
    paste0("bw_tbl_results", collate_suffix, ".rds")
  )
)

if (isTRUE(run_simulations)) {
  chunk_job_files <- list.files(dir_jobs_chunk, full.names = FALSE)
  chunk_completed_ids <- unique(sub(
    "^completed-",
    "",
    chunk_job_files[grepl("^completed-", chunk_job_files)]
  ))
  chunk_failed_ids <- unique(sub(
    "^error-",
    "",
    chunk_job_files[grepl("^error-", chunk_job_files)]
  ))
  chunk_completed_ids <- setdiff(chunk_completed_ids, chunk_failed_ids)
  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = length(chunk_completed_ids),
    failed_sims = length(chunk_failed_ids),
    collate_ok = chunk_collate_ok,
    validation_ok = chunk_collate_ok && nrow(bw_tbl_results) > 0L
  )
}

if (.analysis_can_promote(run_ctx)) {
  expected_sim_ids <- sort(unique(as.integer(sim_grid_all$sim_id)))
  full_validation_ok <- identical(collated_sim_ids, expected_sim_ids)
  if (!isTRUE(full_validation_ok)) {
    error_message <- paste0(
      "Refusing to promote analysis 4: collated output does not contain ",
      "exactly the complete simulation grid."
    )
    .analysis_mark_chunk(
      run_ctx = run_ctx,
      total_sims = nrow(sim_grid),
      completed_sims = length(chunk_completed_ids),
      failed_sims = length(chunk_failed_ids),
      collate_ok = FALSE,
      validation_ok = FALSE,
      error_message = error_message
    )
    stop(error_message)
  }

  .write_rds_atomic(
    bw_tbl,
    file.path(run_ctx$staging_collated_dir, "bw_list_raw_mtd.rds")
  )
  .write_rds_atomic(
    bw_tbl_results,
    file.path(run_ctx$staging_collated_dir, "bw_tbl_results.rds")
  )
  .analysis_promote_run(run_ctx)
}

if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
    for (bias_uns_setting_curr in unique(bw_tbl_results$bias_uns_setting)) {
      for (bw_ncell_upper_curr in unique(bw_tbl_results$bw_ncell_upper)) {
        plot_tbl <- bw_tbl_results |>
          dplyr::filter(
            mean_pos_setting == !!mean_pos_setting_curr,
            bias_uns_setting == !!bias_uns_setting_curr,
            bw_ncell_upper == !!bw_ncell_upper_curr
          )

        if (nrow(plot_tbl) == 0L) next

        p <- ggplot(
          plot_tbl,
          aes(
            x = n_cell,
            y = mean_bw,
            colour = bw_mtd_base
          )
        ) +
          geom_line(
            aes(linetype = bw_mtd_norm),
            linewidth = 0.7,
            alpha = 0.75
          ) +
          geom_point(
            aes(shape = bw_mtd_norm),
            size = 1.7,
            alpha = 0.75
          ) +
          scale_x_log10() +
          facet_grid(
            transformation ~ prob_response,
            scales = "free_y"
          ) +
          labs(
            title = paste0(
              "Mean position setting: ", mean_pos_setting_curr,
              ", bias_uns_setting: ", bias_uns_setting_curr,
              ", bw_ncell_upper: ", bw_ncell_upper_curr
            ),
            x = "Number of Cells",
            y = "Estimated Bandwidth",
            colour = "Bandwidth method",
            linetype = "Normalisation",
            shape = "Normalisation"
          ) +
          cowplot::theme_cowplot() +
          cowplot::background_grid(major = "xy") +
          theme(
            panel.background = element_rect(fill = "white", colour = NA),
            plot.background = element_rect(fill = "white", colour = NA),
            legend.position = "bottom",
            legend.box = "vertical"
          )

        path_plot <- projr::projr_path_get(
          "cache",
          "sim",
          "bw",
          "est",
          "norm",
          paste0(
            "bw_estimate_meanpos_setting_", mean_pos_setting_curr,
            "_bias_uns_", bias_uns_setting_curr,
            "_ncell_upper_", bw_ncell_upper_curr,
            ".pdf"
          )
        )
        dir.create(dirname(path_plot), recursive = TRUE, showWarnings = FALSE)

        ggsave(
          filename = path_plot,
          plot = p,
          width = 25,
          height = 20,
          units = "cm"
        )
        print(p)
      }
    }
  }
}
```
