---
title: Simulate best adaptive bandwidth
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source analysis helpers in dependency order after the generic QMD runtime
# helper has been loaded.
for (fn in c(
  "sim-misc.R",
  "functionsForBenchmarking-Cyt.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Checking what the adaptive estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider two variations: whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator. We can also vary the amount of separation.

We'll vary over a range of parent population sizes, as that tells us how the estimator behaves over the most important dimension apart from the transformation.

For the adaptive estimator, the output has separate bandwidths for the background-core and high-expression-extra components, and those are estimated separately for the stimulated and unstimulated samples. The collation therefore works with four bandwidth columns:

- `bw_uns_core`
- `bw_stim_core`
- `bw_uns_extra`
- `bw_stim_extra`

```{r}
#| label: actual-settings
# Simulation settings.
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# Adaptive core/extra bandwidths are only produced for the normalised methods.
# Keep the hpi methods used in the norm QMD, but exclude hpi0 as there.
bandwidth_method_vec <- c("hpi1Norm", "hpi2Norm", "hpi3Norm")

bandwidth_ncell_upper_vec <- c(
  5e2, 1e3, 2.5e3, 5e3, 1e4, 2e4, 5e4, Inf
)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)

bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    .data$bias_uns_setting == "low"
  )

bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw_mtd         = bandwidth_method_vec,
  bw_ncell_upper = bandwidth_ncell_upper_vec,
  bias_uns       = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd,
    bw_ncell_upper, bias_uns_setting, bias_uns
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    .data$n_cell,
    dplyr::desc(.data$prob_response),
    .data$transformation,
    .data$mean_pos_setting,
    .data$bw_mtd,
    .data$bw_ncell_upper
  )

if (.isDev()) {
  set.seed(3)
  sim_grid <- sim_grid |>
    dplyr::filter(
      .data$n_cell %in% c(5e3, 1e4, 2e4),
      .data$prob_response == 0.002,
      .data$mean_pos_setting == "high",
      .data$transformation %in% transformation_vec,
      grepl("hpi1", .data$bw_mtd)
    ) |>
    dplyr::sample_n(5)
}

sim_grid_all <- sim_grid

if (nrow(sim_grid_all) > 0L) {
  set.seed(sim_grid_shuffle_seed)

  sim_grid <- sim_grid_all |>
    dplyr::mutate(
      .shuffle_order = sample.int(dplyr::n())
    ) |>
    dplyr::arrange(.data$.shuffle_order) |>
    dplyr::mutate(
      .sim_grid_chunk = ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
    ) |>
    dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
    dplyr::arrange(
      .data$n_cell,
      dplyr::desc(.data$prob_response),
      .data$transformation,
      .data$mean_pos_setting,
      .data$bw_mtd,
      .data$bw_ncell_upper
    ) |>
    dplyr::select(-.shuffle_order, -.sim_grid_chunk)
} else {
  sim_grid <- sim_grid_all
}

sim_grid
print(paste0(
  "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
  " simulations for sim_grid chunk ", sim_grid_chunk_index,
  " of ", sim_grid_n_chunks,
  " using shuffle seed ", sim_grid_shuffle_seed
))

bw_fallback <- 0.23482348792138919129198282389
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
devtools::load_all()
bw_list_raw_loop <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ", sim_grid$transformation[i],
    " | prob = ", sim_grid$prob_response[i],
    " | n_cell = ", sim_grid$n_cell[i],
    " | mean = ", sim_grid$mean_pos[i],
    " | bw_mtd = ", sim_grid$bw_mtd[i],
    " | bw_ncell_upper = ", sim_grid$bw_ncell_upper[i],
    " | bias_uns_setting = ", sim_grid$bias_uns_setting[i],
    " | bias_uns = ", sim_grid$bias_uns[i],
    " | sim_id = ", sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)
  bw_list_raw_loop[[i]] <- .simBandwidthEstBwDirectAdaptive(
    nSample = if (.isDev()) 1L else 5L,
    nMarker = 1L,
    nCondition = 2L,
    nCluster = 2L,
    nIter = if (.isDev()) 1L else 5L,
    biasUns = sim_grid$bias_uns[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwNcellMax = sim_grid$bw_ncell_upper[i],
    bwFallback = bw_fallback,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_upper = sim_grid$bw_ncell_upper[i],
      bias_uns_setting = sim_grid$bias_uns_setting[i],
      bias_uns = sim_grid$bias_uns[i],
      sim_id = sim_grid$sim_id[i],
      error_message = NA_character_
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_upper,
      bias_uns_setting,
      bias_uns,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

dir.create(
  dirname(projr::projr_path_get(
    "cache", "sim", "bw", "est", "adaptive", "bw_list_raw_loop.rds"
  )),
  recursive = TRUE,
  showWarnings = FALSE
)

saveRDS(
  bw_list_raw_loop,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "adaptive",
    "bw_list_raw_loop.rds"
  )
)

bw_list_raw_loop |>
  purrr::list_rbind() |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    chnl,
    n_cell_stim,
    n_cell_uns
  ) |>
  dplyr::summarise(
    n_est_uns_core = sum(is.finite(.data$bw_uns_core)),
    n_est_stim_core = sum(is.finite(.data$bw_stim_core)),
    n_est_uns_extra = sum(is.finite(.data$bw_uns_extra)),
    n_est_stim_extra = sum(is.finite(.data$bw_stim_extra)),
    mean_bw_uns_core = mean(.data$bw_uns_core, na.rm = TRUE),
    mean_bw_stim_core = mean(.data$bw_stim_core, na.rm = TRUE),
    mean_bw_uns_extra = mean(.data$bw_uns_extra, na.rm = TRUE),
    mean_bw_stim_extra = mean(.data$bw_stim_extra, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    dplyr::starts_with("n_est"),
    dplyr::starts_with("mean_bw")
  ) |>
  print(width = Inf)
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

run_ctx <- .analysis_run_context(
  analysis_key = c("sim", "bw", "est", "adaptive"),
  run_id = analysis_run_id,
  path_root = root_dir,
  params = list(
    run_simulations = run_simulations,
    run_plots = run_plots,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    sim_grid_shuffle_seed = sim_grid_shuffle_seed
  ),
  sim_grid_chunk_index = sim_grid_chunk_index,
  sim_grid_n_chunks = sim_grid_n_chunks
)

path_progress_file <- run_ctx$progress_file
dir_progress <- run_ctx$progress_run_dir
dir_jobs_chunk <- run_ctx$chunk_jobs_dir
dir_output <- run_ctx$chunk_output_dir

dir.create(dir_jobs_chunk, recursive = TRUE, showWarnings = FALSE)
dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)

bw_list_raw <- list()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  future::plan(
    future::multisession,
    workers = max(1, .simGetCores())
  )

  total_sims <- nrow(sim_grid)

  bw_list_raw <- progressr::with_progress({
    p <- progressr::progressor(steps = total_sims)

    furrr::future_pmap(
      sim_grid,
      function(transformation, prob_response, n_cell, mean_pos_setting, mean_pos,
               bw_mtd, bw_ncell_upper, bias_uns_setting, bias_uns, sim_id) {
        settings_row <- paste0(
          "trans = ", transformation,
          " | prob = ", prob_response,
          " | n_cell = ", n_cell,
          " | mean = ", mean_pos,
          " | bw_mtd = ", bw_mtd,
          " | bw_ncell_upper = ", bw_ncell_upper,
          " | bias_uns_setting = ", bias_uns_setting,
          " | bias_uns = ", bias_uns,
          " | sim_id = ", sim_id,
          " | chunk = ", sim_grid_chunk_index,
          " of ", sim_grid_n_chunks
        )

        file_running <- file.path(dir_jobs_chunk, paste0("running-", sim_id))
        file_completed <- file.path(dir_jobs_chunk, paste0("completed-", sim_id))
        file_error <- file.path(dir_jobs_chunk, paste0("error-", sim_id))
        file_output <- .path_sim_output(
          sim_id,
          dir_output = dir_output,
          sim_grid_chunk_index = sim_grid_chunk_index,
          sim_grid_n_chunks = sim_grid_n_chunks
        )

        existing_output <- if (file.exists(file_output)) {
          tryCatch(readRDS(file_output), error = function(e) NULL)
        } else {
          NULL
        }

        if (!is.null(existing_output)) {
          p(sprintf("Skipped existing sim_id: %s", sim_id))
          return(existing_output)
        }

        file.create(file_running)
        .update_progress_summary(
          path_progress_file = path_progress_file,
          dir_jobs_chunk = dir_jobs_chunk,
          total_sims = total_sims,
          sim_grid_chunk_index = sim_grid_chunk_index,
          sim_grid_n_chunks = sim_grid_n_chunks,
          dir_output = dir_output
        )

        out <- tryCatch(
          {
            sim_res <- .simBandwidthEstBwDirectAdaptive(
              nSample = if (.isDev()) 1L else 5L,
              nMarker = 1L,
              nCondition = 2L,
              nCluster = 2L,
              nIter = if (.isDev()) 1L else 5L,
              biasUns = bias_uns,
              bwMtd = bw_mtd,
              bwMin = -Inf,
              bwMax = Inf,
              bwNcellMax = bw_ncell_upper,
              bwFallback = bw_fallback,
              bwCluster = 0.5,
              nCellStim = n_cell,
              probResponse = prob_response,
              probExact = TRUE,
              meanPos = mean_pos,
              transformation = transformation,
              backgroundRelativeToResponse = 0.2,
              ncellUnsRelativeToStim = 1,
              covEvMin = 1.5,
              covEvMax = 1.5,
              tolClust = NULL,
              summarise = FALSE
            ) |>
              dplyr::mutate(
                transformation = transformation,
                prob_response = prob_response,
                n_cell = n_cell,
                mean_pos_setting = mean_pos_setting,
                mean_pos = mean_pos,
                bw_mtd = bw_mtd,
                bw_ncell_upper = bw_ncell_upper,
                bias_uns_setting = bias_uns_setting,
                bias_uns = bias_uns,
                sim_id = sim_id,
                sim_grid_chunk_index = sim_grid_chunk_index,
                sim_grid_n_chunks = sim_grid_n_chunks,
                error_message = NA_character_
              ) |>
              dplyr::select(
                transformation,
                prob_response,
                n_cell,
                mean_pos_setting,
                mean_pos,
                bw_mtd,
                bw_ncell_upper,
                bias_uns_setting,
                bias_uns,
                sim_id,
                sim_grid_chunk_index,
                sim_grid_n_chunks,
                dplyr::everything()
              )

            .write_rds_atomic(sim_res, file_output)

            file.create(file_completed)
            if (file.exists(file_running)) {
              file.remove(file_running)
            }

            .update_progress_summary(
              path_progress_file = path_progress_file,
              dir_jobs_chunk = dir_jobs_chunk,
              total_sims = total_sims,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              dir_output = dir_output
            )
            p(sprintf("Completed sim_id: %s", sim_id))

            sim_res
          },
          error = function(e) {
            file.create(file_error)
            if (file.exists(file_running)) {
              file.remove(file_running)
            }

            calculated_n_cell_uns <- round(n_cell * 1)

            err_res <- tibble::tibble(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              bw_ncell_upper = bw_ncell_upper,
              bias_uns_setting = bias_uns_setting,
              bias_uns = bias_uns,
              sim_id = sim_id,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              iter = NA_integer_,
              sample = NA_character_,
              ind = NA_character_,
              chnl = "F1",
              n_cell_uns = calculated_n_cell_uns,
              n_cell_stim = n_cell,
              n_uns_bw_core = NA_integer_,
              n_stim_bw_core = NA_integer_,
              max_dens_x = NA_real_,
              bw_uns_core = NA_real_,
              bw_stim_core = NA_real_,
              bw_uns_extra = NA_real_,
              bw_stim_extra = NA_real_,
              threshold_uns = NA_real_,
              threshold_stim = NA_real_,
              shared_grid = list(NA_real_),
              dens_uns_weight = list(NA_real_),
              dens_stim_weight = list(NA_real_),
              error_message = e$message
            )

            .write_rds_atomic(err_res, file_output)
            .update_progress_summary(
              path_progress_file = path_progress_file,
              dir_jobs_chunk = dir_jobs_chunk,
              total_sims = total_sims,
              sim_grid_chunk_index = sim_grid_chunk_index,
              sim_grid_n_chunks = sim_grid_n_chunks,
              dir_output = dir_output
            )
            p(sprintf("ERROR on sim_id: %s", sim_id))

            err_res
          }
        )

        out
      },
      .options = furrr::furrr_options(seed = TRUE, scheduling = Inf)
    )
  })

  future::plan(future::sequential)

  .write_rds_atomic(
    bw_list_raw,
    file.path(
      run_ctx$staging_collated_dir,
      sprintf(
        "bw_list_raw_chunk_%03d-of_%03d.rds",
        sim_grid_chunk_index,
        sim_grid_n_chunks
      )
    )
  )
} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}

if (isTRUE(run_simulations)) {
  chunk_job_files <- list.files(dir_jobs_chunk, full.names = FALSE)
  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = sum(grepl("^completed-", chunk_job_files)),
    failed_sims = sum(grepl("^error-", chunk_job_files))
  )
}
```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true
#| fig-width: 7
#| fig-height: 10
bw_list_path_vec <- .find_bw_list_output_files(
  output_dir = run_ctx$staging_run_dir,
  cache_dir = run_ctx$current_dir,
  cache_path = c("cache", "log", "analysis", "sim", "bw", "est", "adaptive")
)

if (length(bw_list_path_vec) > 0L) {
  message("Collating ", length(bw_list_path_vec), " saved simulation result files from output.")

  bw_list_raw_collate <- purrr::map(
    bw_list_path_vec,
    function(path) {
      tryCatch(
        readRDS(file = path),
        error = function(e) {
          warning("Could not read saved simulation result: ", path)
          NULL
        }
      )
    }
  ) |>
    purrr::compact()
} else if (exists("bw_list_raw") && length(bw_list_raw) > 0L) {
  message("No saved simulation result files found in output; using bw_list_raw from memory.")
  bw_list_raw_collate <- bw_list_raw
} else {
  path_vec <- c(
    file.path(run_ctx$staging_collated_dir, "bw_list_raw.rds"),
    file.path(run_ctx$current_dir, "collated", "bw_list_raw.rds")
  )
  path_vec <- path_vec[file.exists(path_vec)]
  if (length(path_vec) == 0L) {
    stop(
      "Could not find saved output result files, bw_list_raw.rds, ",
      "or bw_list_raw in memory."
    )
  }
  bw_list_raw_collate <- readRDS(file = path_vec[[1]])
}

bw_tbl <- if (is.data.frame(bw_list_raw_collate)) {
  bw_list_raw_collate
} else {
  purrr::list_rbind(bw_list_raw_collate)
}

bw_tbl <- .simBandwidthAddMissingColumns(
  bw_tbl,
  list(
    mean_pos_setting = NA_character_,
    bw_ncell_upper = NA_real_,
    bias_uns = NA_real_,
    bias_uns_setting = NA_character_,
    sim_id = NA_integer_,
    sim_grid_chunk_index = NA_integer_,
    sim_grid_n_chunks = NA_integer_,
    error_message = NA_character_,
    bw_uns_core = NA_real_,
    bw_stim_core = NA_real_,
    bw_uns_extra = NA_real_,
    bw_stim_extra = NA_real_
  )
)

if (all(is.na(bw_tbl$mean_pos_setting)) && any(!is.na(bw_tbl$mean_pos))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-mean_pos_setting) |>
    dplyr::left_join(
      mean_pos_settings_tbl,
      by = c("transformation", "mean_pos")
    )
}

bw_tbl |>
  dplyr::select(transformation, mean_pos_setting, mean_pos) |>
  dplyr::distinct()

if (all(is.na(bw_tbl$bias_uns_setting)) && all(is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::mutate(bias_uns_setting = "low") |>
    dplyr::select(-bias_uns) |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns_setting")
    )
}

bw_tbl |>
  dplyr::select(transformation, bias_uns_setting, bias_uns) |>
  dplyr::distinct()

if (all(is.na(bw_tbl$bias_uns_setting)) && any(!is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns")
    )
}

bw_tbl <- bw_tbl |>
  dplyr::filter(!is.na(.data$bias_uns_setting))

if (nrow(bw_tbl) == 0L) {
  stop("No cached rows contain the current bias_uns settings. Rerun the simulation chunk.")
}

sim_grid_key_cols <- setdiff(
  intersect(colnames(sim_grid_all), colnames(bw_tbl)),
  "sim_id"
)

if (nrow(bw_tbl) == 0L) {
  stop("Cached rows do not match the current simulation grid. Rerun the simulation chunk.")
}

bw_measure_tbl <- tibble::tribble(
  ~bw_component, ~bw_condition, ~bw_col,
  "core", "unstim", "bw_uns_core",
  "core", "stim", "bw_stim_core",
  "extra", "unstim", "bw_uns_extra",
  "extra", "stim", "bw_stim_extra"
)

bw_tbl_long <- purrr::map_dfr(seq_len(nrow(bw_measure_tbl)), function(i) {
  bw_component_curr <- bw_measure_tbl$bw_component[[i]]
  bw_condition_curr <- bw_measure_tbl$bw_condition[[i]]
  bw_col_curr <- bw_measure_tbl$bw_col[[i]]

  bw_tbl |>
    dplyr::mutate(
      bw_component = bw_component_curr,
      bw_condition = bw_condition_curr,
      bw_est = .data[[bw_col_curr]]
    )
}) |>
  dplyr::mutate(
    bw_component = factor(.data$bw_component, levels = c("core", "extra")),
    bw_condition = factor(.data$bw_condition, levels = c("unstim", "stim"))
  )

bw_tbl_results <- bw_tbl_long |>
  dplyr::group_by(
    dplyr::pick(
      dplyr::any_of(c(colnames(sim_grid_all), "bw_component", "bw_condition"))
    )
  ) |>
  dplyr::summarise(
    n_est = sum(is.finite(.data$bw_est)),
    mean_bw = dplyr::if_else(
      any(is.finite(.data$bw_est)),
      mean(.data$bw_est, na.rm = TRUE),
      NA_real_
    ),
    .groups = "drop"
  ) |>
  dplyr::select(
    dplyr::any_of(c(colnames(sim_grid_all), "bw_component", "bw_condition")),
    n_est,
    mean_bw
  ) |>
  dplyr::mutate(
    bw_mtd_base = gsub("Norm$", "", .data$bw_mtd),
    bw_mtd_norm = grepl("Norm$", .data$bw_mtd),
    bw_mtd_norm = ifelse(.data$bw_mtd_norm, "norm", "non-norm")
  )

bw_tbl_results |>
  dplyr::group_by(.data$bw_component, .data$bw_condition) |>
  dplyr::summarise(
    n_grid_rows = dplyr::n(),
    n_grid_rows_with_estimates = sum(is.finite(.data$mean_bw)),
    .groups = "drop"
  ) |>
  print(width = Inf)

dir_results <- run_ctx$staging_collated_dir

collate_suffix <- if (sim_grid_n_chunks > 1L) {
  sprintf("_chunk_%03d-of_%03d", sim_grid_chunk_index, sim_grid_n_chunks)
} else {
  ""
}

saveRDS(
  bw_tbl_long,
  file = file.path(dir_results, paste0("bw_tbl_long", collate_suffix, ".rds"))
)
saveRDS(
  bw_tbl_results,
  file = file.path(dir_results, paste0("bw_tbl_results", collate_suffix, ".rds"))
)

chunk_job_files <- list.files(dir_jobs_chunk, full.names = FALSE)
.analysis_mark_chunk(
  run_ctx = run_ctx,
  total_sims = nrow(sim_grid),
  completed_sims = sum(grepl("^completed-", chunk_job_files)),
  failed_sims = sum(grepl("^error-", chunk_job_files)),
  collate_ok = TRUE,
  validation_ok = nrow(bw_tbl_results) > 0L
)

if (.analysis_can_promote(run_ctx)) {
  .analysis_promote_run(run_ctx)
}

if (isTRUE(run_plots)) {
  mean_pos_setting_curr <- unique(bw_tbl_results$mean_pos_setting)[[1]]
  bias_uns_setting_curr <- unique(bw_tbl_results$bias_uns_setting)[[1]]
  bw_ncell_upper_curr <- NA
  bw_component_curr <- "extra"
  bw_condition_curr <- "stim"
  for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
    bw_tbl_mean_pos <- bw_tbl_results |>
      dplyr::filter(.data$mean_pos_setting == !!mean_pos_setting_curr)

    if (nrow(bw_tbl_mean_pos) == 0L) {
      print(paste0("No results for mean_pos_setting: ", mean_pos_setting_curr))
      next
    }

    for (bias_uns_setting_curr in unique(bw_tbl_mean_pos$bias_uns_setting)) {
      bw_tbl_bias <- bw_tbl_mean_pos |>
        dplyr::filter(.data$bias_uns_setting == !!bias_uns_setting_curr)

      if (nrow(bw_tbl_bias) == 0L) {
        print(paste0(
          "No results for mean_pos_setting: ", mean_pos_setting_curr,
          ", bias_uns_setting: ", bias_uns_setting_curr
        ))
        next
      }

      for (bw_ncell_upper_curr in unique(bw_tbl_bias$bw_ncell_upper)) {
        bw_tbl_ncell <- bw_tbl_bias
        if (!is.na(bw_ncell_upper_curr)) {
          bw_tbl_ncell <- bw_tbl_ncell |>
            dplyr::filter(.data$bw_ncell_upper == !!bw_ncell_upper_curr)
        }

        if (nrow(bw_tbl_ncell) == 0L) {
          print(paste0(
            "No results for mean_pos_setting: ", mean_pos_setting_curr,
            ", bias_uns_setting: ", bias_uns_setting_curr,
            ", bw_ncell_upper: ", bw_ncell_upper_curr
          ))
          next
        }

        for (bw_component_curr in levels(bw_tbl_ncell$bw_component)) {
          bw_tbl_comp <- bw_tbl_ncell |>
            dplyr::filter(.data$bw_component == !!bw_component_curr)

          if (nrow(bw_tbl_comp) == 0L) {
            print(paste0(
              "No results for mean_pos_setting: ", mean_pos_setting_curr,
              ", bias_uns_setting: ", bias_uns_setting_curr,
              ", bw_ncell_upper: ", bw_ncell_upper_curr,
              ", bw_component: ", bw_component_curr
            ))
            next
          }

          for (bw_condition_curr in levels(bw_tbl_comp$bw_condition)) {
            plot_tbl <- bw_tbl_comp |>
              dplyr::filter(.data$bw_condition == !!bw_condition_curr)

            if (nrow(plot_tbl) == 0L) {
              print(paste0(
                "No results for plotting for mean_pos_setting: ", mean_pos_setting_curr,
                ", bias_uns_setting: ", bias_uns_setting_curr,
                ", bw_ncell_upper: ", bw_ncell_upper_curr,
                ", bw_component: ", bw_component_curr,
                ", bw_condition: ", bw_condition_curr
              ))
              next
            }

            print(paste0(
              "Plotting for mean_pos_setting: ", mean_pos_setting_curr,
              ", bias_uns_setting: ", bias_uns_setting_curr,
              ", bw_ncell_upper: ", bw_ncell_upper_curr,
              ", bw_component: ", bw_component_curr,
              ", bw_condition: ", bw_condition_curr
            ))

            # --- GGPLOT AND GGSAVE CODE CONTINUES UNCHANGED ---
            p <- ggplot(
              plot_tbl,
              aes(x = .data$n_cell, y = .data$mean_bw, colour = .data$bw_mtd_base)
            ) +
              geom_line(linetype = "solid") +
              geom_point() +
              scale_x_log10() +
              facet_grid(transformation ~ prob_response, scales = "free_y") +
              labs(
                title = paste0(
                  "Mean position setting: ", mean_pos_setting_curr,
                  ",\nbias_uns_setting: ", bias_uns_setting_curr,
                  ",\nbw_ncell_upper: ", bw_ncell_upper_curr,
                  ",\ncomponent: ", bw_component_curr,
                  ",\ncondition: ", bw_condition_curr
                ),
                x = "Number of cells", y = "Estimated bandwidth", colour = "Method"
              ) +
              cowplot::theme_cowplot() +
              cowplot::background_grid(major = "xy") +
              theme(
                panel.background = element_rect(fill = "white", colour = NA),
                plot.background = element_rect(fill = "white", colour = NA),
                legend.position = "bottom"
              ) +
              scale_colour_manual(values = .simBwMtdColVec, drop = FALSE)

            path_plot <- file.path(
              dir_results, "plots",
              paste0(
                "bw_estimate_meanpos_setting_", .file_safe(mean_pos_setting_curr),
                "_bias_uns_", .file_safe(bias_uns_setting_curr),
                "_ncell_upper_", .file_safe(bw_ncell_upper_curr),
                "_component_", .file_safe(bw_component_curr),
                "_condition_", .file_safe(bw_condition_curr), ".png"
              )
            )
            dir.create(dirname(path_plot), recursive = TRUE, showWarnings = FALSE)
            ggsave(filename = path_plot, plot = p, width = 12, height = 20, units = "cm")
            print(p)
          }
        }
      }
    }
  }
}
```
