---
title: Simulate best adaptive bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)

# Find the build directory root.
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Checking what the adaptive estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider two variations: whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator. We can also vary the amount of separation.

We'll vary over a range of parent population sizes, as that tells us how the estimator behaves over the most important dimension apart from the transformation.

For the adaptive estimator, the output has separate bandwidths for the background-core and high-expression-extra components, and those are estimated separately for the stimulated and unstimulated samples. The collation therefore works with four bandwidth columns:

- `bw_uns_core`
- `bw_stim_core`
- `bw_uns_extra`
- `bw_stim_extra`

```{r}
#| label: actual-settings
# Simulation settings.
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# Adaptive core/extra bandwidths are only produced for the normalised methods.
# Keep the hpi methods used in the norm QMD, but exclude hpi0 as there.
bandwidth_method_vec <- c("hpi1Norm", "hpi2Norm", "hpi3Norm")

bandwidth_ncell_upper_vec <- c(
  5e2, 1e3, 2.5e3, 5e3, 1e4, 2e4, 5e4, Inf
)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)

bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    .data$bias_uns_setting == "low"
  )

bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw_mtd         = bandwidth_method_vec,
  bw_ncell_upper = bandwidth_ncell_upper_vec,
  bias_uns       = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd,
    bw_ncell_upper, bias_uns_setting, bias_uns
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    .data$n_cell,
    dplyr::desc(.data$prob_response),
    .data$transformation,
    .data$mean_pos_setting,
    .data$bw_mtd,
    .data$bw_ncell_upper
  )

if (.isDev()) {
  set.seed(3)
  sim_grid <- sim_grid |>
    dplyr::filter(
      .data$n_cell %in% c(5e3, 1e4, 2e4),
      .data$prob_response == 0.002,
      .data$mean_pos_setting == "high",
      .data$transformation %in% transformation_vec,
      grepl("hpi1", .data$bw_mtd)
    ) |>
    dplyr::sample_n(5)
}

sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))

bw_fallback <- 0.23482348792138919129198282389
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
devtools::load_all()
bw_list_raw_loop <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ", sim_grid$transformation[i],
    " | prob = ", sim_grid$prob_response[i],
    " | n_cell = ", sim_grid$n_cell[i],
    " | mean = ", sim_grid$mean_pos[i],
    " | bw_mtd = ", sim_grid$bw_mtd[i],
    " | bw_ncell_upper = ", sim_grid$bw_ncell_upper[i],
    " | bias_uns_setting = ", sim_grid$bias_uns_setting[i],
    " | bias_uns = ", sim_grid$bias_uns[i],
    " | sim_id = ", sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)
  bw_list_raw_loop[[i]] <- .simBandwidthEstBwDirectAdaptive(
    nSample = if (.isDev()) 1L else 5L,
    nMarker = 1L,
    nCondition = 2L,
    nCluster = 2L,
    nIter = if (.isDev()) 1L else 5L,
    biasUns = sim_grid$bias_uns[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwNcellMax = sim_grid$bw_ncell_upper[i],
    bwFallback = bw_fallback,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_upper = sim_grid$bw_ncell_upper[i],
      bias_uns_setting = sim_grid$bias_uns_setting[i],
      bias_uns = sim_grid$bias_uns[i],
      sim_id = sim_grid$sim_id[i],
      error_message = NA_character_
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_upper,
      bias_uns_setting,
      bias_uns,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

dir.create(
  dirname(projr::projr_path_get(
    "cache", "sim", "bw", "est", "adaptive", "bw_list_raw_loop.rds"
  )),
  recursive = TRUE,
  showWarnings = FALSE
)

saveRDS(
  bw_list_raw_loop,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "adaptive",
    "bw_list_raw_loop.rds"
  )
)

bw_list_raw_loop |>
  purrr::list_rbind() |>
  dplyr::group_by(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    chnl,
    n_cell_stim,
    n_cell_uns
  ) |>
  dplyr::summarise(
    n_est_uns_core = sum(is.finite(.data$bw_uns_core)),
    n_est_stim_core = sum(is.finite(.data$bw_stim_core)),
    n_est_uns_extra = sum(is.finite(.data$bw_uns_extra)),
    n_est_stim_extra = sum(is.finite(.data$bw_stim_extra)),
    mean_bw_uns_core = mean(.data$bw_uns_core, na.rm = TRUE),
    mean_bw_stim_core = mean(.data$bw_stim_core, na.rm = TRUE),
    mean_bw_uns_extra = mean(.data$bw_uns_extra, na.rm = TRUE),
    mean_bw_stim_extra = mean(.data$bw_stim_extra, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation,
    bw_mtd,
    bw_ncell_upper,
    bias_uns_setting,
    bias_uns,
    dplyr::starts_with("n_est"),
    dplyr::starts_with("mean_bw")
  ) |>
  print(width = Inf)
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

# 1. Setup paths.
path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "bw",
  "est",
  "adaptive",
  "progress.txt"
)
dir_progress <- dirname(path_progress_file)
dir_jobs <- file.path(dir_progress, "jobs")

# 2. Clean up previous progress file and the jobs subdirectory.
dir.create(dir_progress, recursive = TRUE, showWarnings = FALSE)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}
if (dir.exists(dir_jobs)) {
  unlink(dir_jobs, recursive = TRUE, force = TRUE)
}
dir.create(dir_jobs, recursive = TRUE, showWarnings = FALSE)

# 3. Setup parallel plan.
future::plan(
  future::multisession,
  workers = max(1, .simGetCores())
)

total_sims <- nrow(sim_grid)

bw_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = total_sims)

  furrr::future_pmap(
    sim_grid,
    function(
      transformation, prob_response, n_cell, mean_pos_setting, mean_pos,
      bw_mtd, bw_ncell_upper, bias_uns_setting, bias_uns, sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | bw_mtd = ", bw_mtd,
        " | bw_ncell_upper = ", bw_ncell_upper,
        " | bias_uns_setting = ", bias_uns_setting,
        " | bias_uns = ", bias_uns,
        " | sim_id = ", sim_id
      )

      file_running <- file.path(dir_jobs, paste0("running-", sim_id))
      file_completed <- file.path(dir_jobs, paste0("completed-", sim_id))
      file_error <- file.path(dir_jobs, paste0("error-", sim_id))

      update_progress_summary <- function() {
        all_job_files <- list.files(dir_jobs, full.names = FALSE)

        n_running <- sum(grepl("^running-", all_job_files))
        n_completed <- sum(grepl("^completed-", all_job_files))
        n_error <- sum(grepl("^error-", all_job_files))
        n_done <- n_completed + n_error

        running_files <- all_job_files[grepl("^running-", all_job_files)]
        running_ids <- sub("^running-", "", running_files)
        running_list_str <- if (length(running_ids) > 0) {
          paste(running_ids, collapse = ", ")
        } else {
          "None"
        }

        summary_text <- paste0(
          "========================================\n",
          " SIMULATION PROGRESS DASHBOARD          \n",
          "========================================\n",
          "Total Simulations  : ", total_sims, "\n",
          "In Progress        : ", n_running, "\n",
          "Completed (Success): ", n_completed, "\n",
          "Completed (Errors) : ", n_error, "\n",
          "Total Done         : ", n_done, "\n",
          "----------------------------------------\n",
          "Currently Running IDs:\n", running_list_str, "\n",
          "========================================\n"
        )

        cat(summary_text, file = path_progress_file, append = FALSE)
      }

      file.create(file_running)
      update_progress_summary()

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthEstBwDirectAdaptive(
            nSample = if (.isDev()) 1L else 5L,
            nMarker = 1L,
            nCondition = 2L,
            nCluster = 2L,
            nIter = if (.isDev()) 1L else 5L,
            biasUns = bias_uns,
            bwMtd = bw_mtd,
            bwMin = -Inf,
            bwMax = Inf,
            bwNcellMax = bw_ncell_upper,
            bwFallback = bw_fallback,
            bwCluster = 0.5,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            backgroundRelativeToResponse = 0.2,
            ncellUnsRelativeToStim = 1,
            covEvMin = 1.5,
            covEvMax = 1.5,
            tolClust = NULL,
            summarise = FALSE
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              bw_ncell_upper = bw_ncell_upper,
              bias_uns_setting = bias_uns_setting,
              bias_uns = bias_uns,
              sim_id = sim_id,
              error_message = NA_character_
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos_setting,
              mean_pos,
              bw_mtd,
              bw_ncell_upper,
              bias_uns_setting,
              bias_uns,
              sim_id,
              dplyr::everything()
            )

          file.create(file_completed)
          if (file.exists(file_running)) {
            file.remove(file_running)
          }

          update_progress_summary()
          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          file.create(file_error)
          if (file.exists(file_running)) {
            file.remove(file_running)
          }

          update_progress_summary()
          p(sprintf("ERROR on sim_id: %s", sim_id))

          calculated_n_cell_uns <- round(n_cell * 1)

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw_mtd = bw_mtd,
            bw_ncell_upper = bw_ncell_upper,
            bias_uns_setting = bias_uns_setting,
            bias_uns = bias_uns,
            sim_id = sim_id,
            iter = NA_integer_,
            sample = NA_character_,
            ind = NA_character_,
            chnl = "F1",
            n_cell_uns = calculated_n_cell_uns,
            n_cell_stim = n_cell,
            n_uns_bw_core = NA_integer_,
            n_stim_bw_core = NA_integer_,
            max_dens_x = NA_real_,
            bw_uns_core = NA_real_,
            bw_stim_core = NA_real_,
            bw_uns_extra = NA_real_,
            bw_stim_extra = NA_real_,
            threshold_uns = NA_real_,
            threshold_stim = NA_real_,
            shared_grid = list(NA_real_),
            dens_uns_weight = list(NA_real_),
            dens_stim_weight = list(NA_real_),
            error_message = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

dir.create(
  dirname(projr::projr_path_get(
    "cache", "sim", "bw", "est", "adaptive", "bw_list_raw.rds"
  )),
  recursive = TRUE,
  showWarnings = FALSE
)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "adaptive",
    "bw_list_raw.rds"
  )
)
```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true
#| fig-width: 7
#| fig-height: 10
.add_missing_cols <- function(.data, cols) {
  for (nm in names(cols)) {
    if (!nm %in% names(.data)) {
      .data[[nm]] <- cols[[nm]]
    }
  }
  .data
}

.file_safe <- function(x) {
  x <- as.character(x)
  x <- gsub("[^A-Za-z0-9]+", "_", x)
  x <- gsub("^_+|_+$", "", x)
  x
}

if (exists("bw_list_raw") && length(bw_list_raw) > 0L) {
  bw_list_raw_collate <- bw_list_raw
} else {
  path_vec <- c(
    projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "bw_list_raw.rds",
      safe = TRUE
    ),
    projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "bw_list_raw.rds",
      safe = FALSE
    )
  )
  path_vec <- path_vec[file.exists(path_vec)]
  if (length(path_vec) == 0L) {
    stop("Could not find bw_list_raw.rds. Run the simulation chunk first.")
  }
  bw_list_raw_collate <- readRDS(file = path_vec[[1]])
}

bw_tbl <- if (is.data.frame(bw_list_raw_collate)) {
  bw_list_raw_collate
} else {
  purrr::list_rbind(bw_list_raw_collate)
}

bw_tbl <- .add_missing_cols(
  bw_tbl,
  list(
    mean_pos_setting = NA_character_,
    bw_ncell_upper = NA_real_,
    bias_uns = NA_real_,
    bias_uns_setting = NA_character_,
    sim_id = NA_integer_,
    error_message = NA_character_,
    bw_uns_core = NA_real_,
    bw_stim_core = NA_real_,
    bw_uns_extra = NA_real_,
    bw_stim_extra = NA_real_
  )
)

if (all(is.na(bw_tbl$bias_uns_setting)) && any(!is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-bias_uns_setting) |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns")
    )
}

if (all(is.na(bw_tbl$mean_pos_setting)) && any(!is.na(bw_tbl$mean_pos))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-mean_pos_setting) |>
    dplyr::left_join(
      mean_pos_settings_tbl,
      by = c("transformation", "mean_pos")
    )
}

bw_tbl <- bw_tbl |>
  dplyr::filter(!is.na(.data$bias_uns_setting))

if (nrow(bw_tbl) == 0L) {
  stop("No cached rows contain the current bias_uns settings. Rerun the simulation chunk.")
}

sim_grid_key_cols <- setdiff(
  intersect(colnames(sim_grid), colnames(bw_tbl)),
  "sim_id"
)

bw_tbl <- bw_tbl |>
  dplyr::semi_join(
    sim_grid |>
      dplyr::select(dplyr::all_of(sim_grid_key_cols)) |>
      dplyr::distinct(),
    by = sim_grid_key_cols
  )

if (nrow(bw_tbl) == 0L) {
  stop("Cached rows do not match the current simulation grid. Rerun the simulation chunk.")
}

bw_measure_tbl <- tibble::tribble(
  ~bw_component, ~bw_condition, ~bw_col,
  "core", "unstim", "bw_uns_core",
  "core", "stim", "bw_stim_core",
  "extra", "unstim", "bw_uns_extra",
  "extra", "stim", "bw_stim_extra"
)

bw_tbl_long <- purrr::map_dfr(seq_len(nrow(bw_measure_tbl)), function(i) {
  bw_component_curr <- bw_measure_tbl$bw_component[[i]]
  bw_condition_curr <- bw_measure_tbl$bw_condition[[i]]
  bw_col_curr <- bw_measure_tbl$bw_col[[i]]

  bw_tbl |>
    dplyr::mutate(
      bw_component = bw_component_curr,
      bw_condition = bw_condition_curr,
      bw_est = .data[[bw_col_curr]]
    )
}) |>
  dplyr::mutate(
    bw_component = factor(.data$bw_component, levels = c("core", "extra")),
    bw_condition = factor(.data$bw_condition, levels = c("unstim", "stim"))
  )

bw_tbl_results <- bw_tbl_long |>
  dplyr::group_by(
    dplyr::pick(
      dplyr::any_of(c(colnames(sim_grid), "bw_component", "bw_condition"))
    )
  ) |>
  dplyr::summarise(
    n_est = sum(is.finite(.data$bw_est)),
    mean_bw = dplyr::if_else(
      any(is.finite(.data$bw_est)),
      mean(.data$bw_est, na.rm = TRUE),
      NA_real_
    ),
    .groups = "drop"
  ) |>
  dplyr::select(
    dplyr::any_of(c(colnames(sim_grid), "bw_component", "bw_condition")),
    n_est,
    mean_bw
  ) |>
  dplyr::mutate(
    bw_mtd_base = gsub("Norm$", "", .data$bw_mtd),
    bw_mtd_norm = grepl("Norm$", .data$bw_mtd),
    bw_mtd_norm = ifelse(.data$bw_mtd_norm, "norm", "non-norm")
  )

bw_tbl_results |>
  dplyr::group_by(.data$bw_component, .data$bw_condition) |>
  dplyr::summarise(
    n_grid_rows = dplyr::n(),
    n_grid_rows_with_estimates = sum(is.finite(.data$mean_bw)),
    .groups = "drop"
  ) |>
  print(width = Inf)

dir_results <- projr::projr_path_get(
  "cache",
  "sim",
  "bw",
  "est",
  "adaptive"
)
dir.create(dir_results, recursive = TRUE, showWarnings = FALSE)

saveRDS(
  bw_tbl_long,
  file = file.path(dir_results, "bw_tbl_long.rds")
)
saveRDS(
  bw_tbl_results,
  file = file.path(dir_results, "bw_tbl_results.rds")
)

for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
  for (bias_uns_setting_curr in unique(bw_tbl_results$bias_uns_setting)) {
    for (bw_ncell_upper_curr in unique(bw_tbl_results$bw_ncell_upper)) {
      for (bw_component_curr in levels(bw_tbl_results$bw_component)) {
        for (bw_condition_curr in levels(bw_tbl_results$bw_condition)) {
          plot_tbl <- bw_tbl_results |>
            dplyr::filter(
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_ncell_upper == !!bw_ncell_upper_curr,
              .data$bw_component == !!bw_component_curr,
              .data$bw_condition == !!bw_condition_curr
            )

          if (nrow(plot_tbl) == 0L) {
            next
          }

          p <- ggplot(
            plot_tbl,
            aes(x = .data$n_cell, y = .data$mean_bw, colour = .data$bw_mtd_base)
          ) +
            geom_line(mapping = aes(linetype = .data$bw_mtd_norm)) +
            geom_point() +
            scale_x_log10() +
            facet_grid(
              transformation ~ prob_response,
              scales = "free_y"
            ) +
            labs(
              title = paste0(
                "Mean position setting: ", mean_pos_setting_curr,
                ", bias_uns_setting: ", bias_uns_setting_curr,
                ", bw_ncell_upper: ", bw_ncell_upper_curr,
                ", component: ", bw_component_curr,
                ", condition: ", bw_condition_curr
              ),
              x = "Number of cells",
              y = "Estimated bandwidth",
              colour = "Bandwidth method",
              linetype = "Bandwidth method type"
            ) +
            cowplot::theme_cowplot() +
            cowplot::background_grid(major = "xy") +
            theme(
              panel.background = element_rect(fill = "white", colour = NA),
              plot.background = element_rect(fill = "white", colour = NA),
              legend.position = "bottom"
            ) +
            scale_colour_manual(
              values = .simBwMtdColVec,
              drop = FALSE
            )

          path_plot <- file.path(
            dir_results,
            "plots",
            paste0(
              "bw_estimate_meanpos_setting_", .file_safe(mean_pos_setting_curr),
              "_bias_uns_", .file_safe(bias_uns_setting_curr),
              "_ncell_upper_", .file_safe(bw_ncell_upper_curr),
              "_component_", .file_safe(bw_component_curr),
              "_condition_", .file_safe(bw_condition_curr),
              ".png"
            )
          )
          dir.create(dirname(path_plot), recursive = TRUE, showWarnings = FALSE)

          ggsave(
            filename = path_plot,
            plot = p,
            width = 12,
            height = 20,
            units = "cm"
          )

          print(p)
        }
      }
    }
  }
}
```
