---
title: Adaptive and constant bandwidth diagnostics
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---


```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
path_out <- projr::projr_path_get_dir(
  "cache",
  "sim-bw-adaptive"
)
```

### Checking what the estimators actually estimate

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# bandwidth settings
bandwidth_method_vec <- c("hpi0Norm", "hpi1Norm", "hpi2Norm", "hpi3Norm", "sjNorm", "nrd0Norm")

# set this later
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- seq(0, 2, by = 0.2)
```


```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw_mtd         = bandwidth_method_vec
) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw_mtd == "hpi1Norm"
    )
}
sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))

bw_fallback <- 0.23482348792138919129198282389
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
bw_list_raw_mtd_loop <- lapply(seq_len(nrow(sim_grid)), function(i) {
  NULL
})

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)
  bw_list_raw_mtd_loop[[i]] <- .simBandwidthEstBwDirectAdaptive(
    nSample = 2,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 1,
    biasUns = if (sim_grid$transformation[i] == "gamma") 0.005 else 0.05,
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = -Inf,
    bwMax = Inf,
    bwFallback = bw_fallback,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    summarise = FALSE
  ) |>
     dplyr::mutate(
       mean_pos_setting = sim_grid$mean_pos_setting[i],
       sim_id = sim_grid$sim_id[i]
     ) |>
     dplyr::select(
       transformation:bw_mtd,
       mean_pos_setting,
       sim_id,
       dplyr::everything()
     )

  message("Completed: ", settings_row)
}

saveRDS(
  bw_list_raw_mtd_loop,
  file = projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "est",
    "adaptive",
    "bw_list_raw_mtd_loop.rds"
  )
)

bw_list_raw_mtd_loop |>
  dplyr::bind_rows() |>
  dplyr::filter(transformation == "gamma") |>
  print(width = Inf) |>
  dplyr::group_by(
    transformation, prob_response, n_cell,
    mean_pos, bw_mtd, chnl,
    n_cell_stim, n_cell_uns
  ) |>
  dplyr::summarise(
    mean_bw_stim = mean(bw_stim, na.rm = TRUE),
    mean_bw_uns = mean(bw_uns, na.rm = TRUE),
    mean_bw = mean(bw, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation, mean_bw
  )
```








---------------------------

## Purpose

This notebook is for inspecting the new adaptive normalised bandwidth behaviour in one simulated stim/unstim pair.
It reports the component bandwidths for the background core and extra component, plots the location-specific bandwidth curves, and plots the densities produced from those bandwidths. It also runs the same diagnostics for non-adaptive bandwidths so the adaptive and constant versions can be compared directly.

## Settings

Change these settings first. The default grid deliberately includes both adaptive `*Norm` methods and ordinary non-adaptive methods.

```{r}
#| label: diagnostic-settings
set.seed(1)

# Simulation settings ---------------------------------------------------------
n_sample <- 1L
n_marker <- 1L
n_condition <- 2L
n_cluster <- 2L
n_cell_stim <- 20000L
ncell_uns_relative_to_stim <- 1
prob_response <- 0.002
background_relative_to_response <- 0.2
mean_pos <- 8
transformation <- "gaussian"
prob_exact <- TRUE
bias_uns <- 0.05
exc_min <- TRUE
cap_stim_range <- TRUE
cov_ev_min <- 1.5
cov_ev_max <- 1.5

# Bandwidth settings ---------------------------------------------------------
bw_method_vec <- c("hpi1Norm", "hpi2Norm", "hpi3Norm", "hpi1", "hpi2")
bw_adaptive_vec <- c(TRUE, FALSE)

# Adaptive mode only applies to *Norm methods. Ordinary methods will be treated
# as constant bandwidths even when bw_adaptive is TRUE.

bw_adj <- 1
bw_ncell_min <- NULL
bw_ncell_max <- NULL
bw_min <- -Inf
bw_max <- Inf
bw_fallback <- 0.25

# Normalised/adaptive internals ---------------------------------------------
norm_mtd <- "moments"
norm_peak_frac <- 0.1
norm_peak_min_rel <- 0.75
norm_extra_frac <- 0.2
norm_extra_max <- Inf
norm_extra_jitter_frac <- 0.25
norm_density_n <- 512L
norm_excess_bw_mtd <- "hpi3"
norm_excess_ncell <- 10000L
norm_adaptive_ncell <- 2500L
bw_adaptive_density_n <- 512L
bw_adaptive_pad_frac <- 0.15
```

## Simulate one stim/unstim pair

```{r}
#| label: simulate-one
sim_obj <- .bwDiagSimulateOne(
  n_sample = n_sample,
  n_marker = n_marker,
  n_condition = n_condition,
  n_cluster = n_cluster,
  n_cell_stim = n_cell_stim,
  ncell_uns_relative_to_stim = ncell_uns_relative_to_stim,
  prob_response = prob_response,
  background_relative_to_response = background_relative_to_response,
  mean_pos = mean_pos,
  transformation = transformation,
  prob_exact = prob_exact,
  bias_uns = bias_uns,
  exc_min = exc_min,
  cap_stim_range = cap_stim_range,
  cov_ev_min = cov_ev_min,
  cov_ev_max = cov_ev_max
)

tibble::tibble(
  arm = c("stim", "unstim"),
  n_cell_after_exc_min_and_cap = c(length(sim_obj$x_stim), length(sim_obj$x_uns)),
  prob_g_min = c(sim_obj$prob_g_min_stim, sim_obj$prob_g_min_uns),
  max_dens_x = sim_obj$max_dens_x
)
```

## Run adaptive and constant bandwidth diagnostics

```{r}
#| label: run-diagnostics
diag_grid <- tidyr::expand_grid(
  bw_mtd = bw_method_vec,
  bw_adaptive = bw_adaptive_vec
) |>
  dplyr::mutate(
    # Ordinary non-Norm methods cannot return adaptive objects in the current
    # helper, so remove duplicate constant calculations.
    bw_adaptive = dplyr::if_else(grepl("Norm$", .data$bw_mtd), .data$bw_adaptive, FALSE)
  ) |>
  dplyr::distinct()

diag_list <- purrr::pmap(
  diag_grid,
  function(bw_mtd, bw_adaptive) {
    .bwDiagRunOne(
      x_stim = sim_obj$x_stim,
      x_uns = sim_obj$x_uns,
      prob_g_min_stim = sim_obj$prob_g_min_stim,
      prob_g_min_uns = sim_obj$prob_g_min_uns,
      bw_mtd = bw_mtd,
      bw_adaptive = bw_adaptive,
      bw_adj = bw_adj,
      bw_ncell_min = bw_ncell_min,
      bw_ncell_max = bw_ncell_max,
      bw_min = bw_min,
      bw_max = bw_max,
      bw_fallback = bw_fallback,
      norm_peak_frac = norm_peak_frac,
      norm_peak_min_rel = norm_peak_min_rel,
      norm_extra_frac = norm_extra_frac,
      norm_extra_max = norm_extra_max,
      norm_extra_jitter_frac = norm_extra_jitter_frac,
      norm_density_n = norm_density_n,
      norm_excess_bw_mtd = norm_excess_bw_mtd,
      norm_excess_ncell = norm_excess_ncell,
      norm_adaptive_ncell = norm_adaptive_ncell,
      norm_mtd = norm_mtd,
      bw_adaptive_density_n = bw_adaptive_density_n,
      bw_adaptive_pad_frac = bw_adaptive_pad_frac
    )
  }
)

component_bw_tbl <- purrr::map_dfr(diag_list, "component")
bw_grid_tbl <- purrr::map_dfr(diag_list, "bw_grid")
density_tbl <- purrr::map_dfr(diag_list, "density")

component_bw_tbl
```

## Core and extra bandwidths by arm

For adaptive `*Norm` methods, `bw_core` and `bw_extra` are the separate normalised component bandwidths. For constant methods, inspect `bw_scalar` instead.

```{r}
#| label: plot-component-bandwidths
component_long_tbl <- component_bw_tbl |>
  tidyr::pivot_longer(
    cols = c("bw_scalar", "bw_core", "bw_extra"),
    names_to = "bw_component",
    values_to = "bw"
  ) |>
  dplyr::filter(is.finite(.data$bw))

p_component <- ggplot(
  component_long_tbl,
  aes(x = arm, y = bw, fill = bw_component)
) +
  geom_col(position = "dodge") +
  facet_wrap(~ config, scales = "free_y") +
  labs(
    x = NULL,
    y = "Bandwidth",
    fill = "Component",
    title = "Core, extra, and scalar bandwidths by stim/unstim arm"
  ) +
  cowplot::theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
  ) +
  theme(legend.position = "bottom")

p_component

ggsave(
  filename = file.path(path_out, "component_bandwidths.png"),
  plot = p_component,
  width = 12,
  height = 8,
  units = "in"
)
```

## Bandwidth curve over expression bins

The adaptive rows show the stim-specific, unstim-specific, and final shared bandwidth curves. The constant rows show flat curves.

```{r}
#| label: plot-bandwidth-curves
p_bw_curve <- ggplot(
  bw_grid_tbl,
  aes(x = x, y = bw, colour = bw_curve)
) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ config, scales = "free_y") +
  labs(
    x = "Expression",
    y = "Bandwidth",
    colour = "Curve",
    title = "Actual bandwidth used at each expression bin"
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
  ) +
  theme(legend.position = "bottom")

p_bw_curve

ggsave(
  filename = file.path(path_out, "bandwidth_curves_by_bin.png"),
  plot = p_bw_curve,
  width = 12,
  height = 8,
  units = "in"
)
```

## Final densities calculated from the shared bandwidth vector

Each final density is calculated on the padded shared grid, area-normalised, then multiplied by the relevant `probGMin`.

```{r}
#| label: plot-final-densities
p_density <- ggplot(
  density_tbl,
  aes(x = x, y = dens, colour = arm)
) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ config, scales = "free_y") +
  labs(
    x = "Expression",
    y = "Density",
    colour = "Arm",
    title = "Stim and unstim densities from adaptive or constant bandwidths"
  ) +
  cowplot::theme_cowplot() +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),
    panel.background = element_rect(fill = "white", colour = NA),
  ) +
  cowplot::background_grid(major = "xy") +
  theme(legend.position = "bottom")

p_density

ggsave(
  filename = file.path(path_out, "densities_from_bandwidths.png"),
  plot = p_density,
  width = 12,
  height = 8,
  units = "in"
)
```

## Area checks

The final density area should be close to `probGMin` for each arm because the density is normalised first and then scaled by `probGMin`.

```{r}
#| label: area-checks
area_check_tbl <- density_tbl |>
  dplyr::group_by(config, bw_mtd, bw_mode, arm) |>
  dplyr::summarise(
    area = .bwDiagTrapz(x, dens),
    median_bw = stats::median(bw, na.rm = TRUE),
    min_bw = min(bw, na.rm = TRUE),
    max_bw = max(bw, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    expected_area = dplyr::if_else(
      .data$arm == "stim",
      sim_obj$prob_g_min_stim,
      sim_obj$prob_g_min_uns
    ),
    area_error = .data$area - .data$expected_area
  )

area_check_tbl

utils::write.csv(
  area_check_tbl,
  file = file.path(path_out, "density_area_checks.csv"),
  row.names = FALSE
)
```

## Zoomed comparison

Use this when the far tail stretches the y-axis too much.

```{r}
#| label: plot-zoomed-densities
x_zoom <- stats::quantile(
  c(sim_obj$x_stim, sim_obj$x_uns),
  probs = c(0.001, 0.999),
  na.rm = TRUE,
  names = FALSE
)

p_density_zoom <- p_density +
  coord_cartesian(xlim = x_zoom)

p_density_zoom

ggsave(
  filename = file.path(path_out, "densities_from_bandwidths_zoomed.png"),
  plot = p_density_zoom,
  width = 12,
  height = 8,
  units = "in"
)
```

## Optional: inspect one method in detail

Change `method_to_inspect` and `mode_to_inspect` to focus on one curve.

```{r}
#| label: inspect-one-method
method_to_inspect <- "hpi1Norm"
mode_to_inspect <- "adaptive"

bw_grid_tbl |>
  dplyr::filter(
    .data$bw_mtd == method_to_inspect,
    .data$bw_mode == mode_to_inspect
  ) |>
  ggplot(aes(x = x, y = bw, colour = bw_curve)) +
  geom_line(linewidth = 0.8) +
  labs(
    x = "Expression",
    y = "Bandwidth",
    colour = "Curve",
    title = paste("Bandwidth curves:", method_to_inspect, mode_to_inspect)
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(legend.position = "bottom")

density_tbl |>
  dplyr::filter(
    .data$bw_mtd == method_to_inspect,
    .data$bw_mode == mode_to_inspect
  ) |>
  ggplot(aes(x = x, y = dens, colour = arm)) +
  geom_line(linewidth = 0.8) +
  coord_cartesian(xlim = x_zoom) +
  labs(
    x = "Expression",
    y = "Density",
    colour = "Arm",
    title = paste("Densities:", method_to_inspect, mode_to_inspect)
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(legend.position = "bottom")
```
