---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Checking what the estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider two variations - whether the number of parent population cells is large or small, and whether the size of the responding population is large enough to impact the estimator.
We can also vary the amount of separation, I suppose.

We'll need to vary over a range of parent population sizes, however, as that will tell us how the estimator behaves over the most crucial dimension (apart from the actual transformation, of course).

Initially, we will ignore setting maximum/minimum number of cells.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# bandwidth settings
bandwidth_method_vec <- c(
  "hpi0Norm", "hpi1Norm", "hpi2Norm", "hpi3Norm"
  )

# set this later
bandwidth_ncell_upper_vec <- c(5e2, 1e3, 2.5e3, 5e3, 1e4, 2e4, 5e4, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    bias_uns_setting == "low"
  )
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  bw_mtd         = bandwidth_method_vec,
  bw_ncell_upper = bandwidth_ncell_upper_vec,
  bias_uns       = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw_mtd,
    bw_ncell_upper, bias_uns_setting, bias_uns
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if (.isDev()) {
  set.seed(3)
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(5e3, 1e4, 2e4),
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      grepl("hpi1", bw_mtd)
    ) |>
    dplyr::sample_n(5)
}
sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))

bw_fallback <- 0.23482348792138919129198282389
```

### Manual loop for debugging

```{r}
#| label: bw-manual-loop
#| eval: false
devtools::load_all()
bw_list_raw_loop <- vector("list", nrow(sim_grid))

sim_grid <- sim_grid |> dplyr::filter(sim_id == 301)
set.seed(4)
for (i in seq_len(nrow(sim_grid))) {

  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | bw_ncell_upper = ",
    sim_grid$bw_ncell_upper[i],
    " | bias_setting = ",
    sim_grid$bias_uns_setting[i],
    " | bias = ",
    sim_grid$bias_uns[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))
  message("Running: ", settings_row)
  bw_list_raw_loop[[i]] <- .simBandwidthEstBwDirectAdaptive(
    nSample = 5L,
    nIter = 5L,
    transformation = sim_grid$transformation[i],
    probResponse = sim_grid$prob_response[i],
    nCellStim = sim_grid$n_cell[i],
    meanPos = sim_grid$mean_pos[i],
    bwMtd = sim_grid$bw_mtd[i],
    biasUns = sim_grid$bias_uns[i]
  )
}

saveRDS(
  bw_list_raw_loop,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "adaptive",
    "bw_list_raw_loop.rds"
  )
)

bw_list_raw_loop |>
  purrr::list_rbind() |>
  print(width = Inf)
```

### Parallel processing

```{r}
#| label: bw-manual-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

# 1. Setup paths
path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "bw",
  "adaptive",
  "progress.txt"
)
dir_progress <- dirname(path_progress_file)
dir_jobs <- file.path(dir_progress, "jobs")

# 2. Clean up previous progress file and the jobs subdirectory
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}
if (dir.exists(dir_jobs)) {
  unlink(dir_jobs, recursive = TRUE, force = TRUE)
}
dir.create(dir_jobs, recursive = TRUE, showWarnings = FALSE)

# 3. Setup parallel plan
future::plan(
  future::multisession,
  workers = max(1, future::availableCores() - 1)
)

# Total number of iterations
total_sims <- nrow(sim_grid)

bw_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = total_sims)

  furrr::future_pmap(
    sim_grid,
    function(
      transformation, prob_response, n_cell, mean_pos_setting, mean_pos,
      bw_mtd, bw_ncell_upper, bias_uns_setting, bias_uns, sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | bw_mtd = ", bw_mtd,
        " | bw_ncell_upper = ", bw_ncell_upper,
        " | bias_setting = ", bias_uns_setting,
        " | bias = ", bias_uns,
        " | sim_id = ", sim_id
      )

      # Define file markers for this specific simulation
      file_running   <- file.path(dir_jobs, paste0("running-", sim_id))
      file_completed <- file.path(dir_jobs, paste0("completed-", sim_id))
      file_error     <- file.path(dir_jobs, paste0("error-", sim_id))

      # Helper function to recalculate state and update progress.txt
      update_progress_summary <- function() {
        # List all generated status files
        all_job_files <- list.files(dir_jobs, full.names = FALSE)
        
        n_running   <- sum(grepl("^running-", all_job_files))
        n_completed <- sum(grepl("^completed-", all_job_files))
        n_error     <- sum(grepl("^error-", all_job_files))
        n_done      <- n_completed + n_error
        
        # Extract specifically which ones are currently running
        running_files <- all_job_files[grepl("^running-", all_job_files)]
        running_ids   <- sub("^running-", "", running_files)
        running_list_str <- if (length(running_ids) > 0) {
          paste(running_ids, collapse = ", ")
        } else {
          "None"
        }

        # Build dashboard content
        summary_text <- paste0(
          "========================================\n",
          " SIMULATION PROGRESS DASHBOARD          \n",
          "========================================\n",
          "Total Simulations  : ", total_sims, "\n",
          "In Progress        : ", n_running, "\n",
          "Completed (Success): ", n_completed, "\n",
          "Completed (Errors) : ", n_error, "\n",
          "Total Done          : ", n_done, "\n",
          "----------------------------------------\n",
          "Currently Running IDs:\n", running_list_str, "\n",
          "========================================\n"
        )
        
        # Overwrite the summary file safely
        cat(summary_text, file = path_progress_file, append = FALSE)
      }

      # --- START WORKER PROCESS ---
      # Mark as running
      file.create(file_running)
      update_progress_summary()

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthEstBwDirectAdaptive(
            nSample = 5L,
            nIter = 5L,
            transformation = transformation,
            probResponse = prob_response,
            nCellStim = n_cell,
            meanPos = mean_pos,
            bwMtd = bw_mtd,
            biasUns = bias_uns
          )

          # Move marker from running to completed
          file.create(file_completed)
          if (file.exists(file_running)) file.remove(file_running)
          
          update_progress_summary()
          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          # Move marker from running to error
          file.create(file_error)
          if (file.exists(file_running)) file.remove(file_running)
          
          update_progress_summary()
          p(sprintf("ERROR on sim_id: %s", sim_id))

          # Infer standard derived metadata variable matching function internals
          calculated_n_cell_uns <- round(n_cell * 1) # matches default ncellUnsRelativeToStim = 1

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos = mean_pos,
            bw_mtd = bw_mtd,
            iter = NA_integer_,
            sample = NA_character_,
            ind = NA_character_,
            chnl = "F1",
            n_cell_uns = calculated_n_cell_uns,
            n_cell_stim = n_cell,
            n_uns_bw_core = NA_integer_,
            n_stim_bw_core = NA_integer_,
            max_dens_x = NA_real_,
            bw_uns_core = NA_real_,
            bw_stim_core = NA_real_,
            bw_uns_extra = NA_real_,
            bw_stim_extra = NA_real_,
            threshold_uns = NA_real_,
            threshold_stim = NA_real_,
            shared_grid = list(NA_real_),
            dens_uns_weight = list(NA_real_),
            dens_stim_weight = list(NA_real_),
            # Added meta-columns helpful for hunting down pipeline failures
            sim_id = sim_id,
            error_message = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "adaptive",
    "bw_list_raw.rds"
  )
)
```

### Collate results

```{r}
#| label: bw-manual-collate
.add_missing_cols <- function(.data, cols) {
  for (nm in names(cols)) {
    if (!nm %in% names(.data)) {
      .data[[nm]] <- cols[[nm]]
    }
  }
  .data
}

if (exists("bw_list_raw") && length(bw_list_raw) > 0L) {
  bw_list_raw_collate <- bw_list_raw
} else {
  path_vec <- c(
    projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "bw_list_raw.rds",
      safe = TRUE
    ),
    projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "bw_list_raw.rds",
      safe = FALSE
    )
  )
  path_vec <- path_vec[file.exists(path_vec)]
  if (length(path_vec) == 0L) {
    stop("Could not find bw_list_raw.rds, and bw_list_raw is not available in memory.")
  }
  bw_list_raw_collate <- readRDS(file = path_vec[[1]])
}

bw_tbl <- bw_list_raw_collate |>
  purrr::list_rbind()

bw_tbl <- .add_missing_cols(
  bw_tbl,
  list(
    iter = NA_integer_,
    sample = NA_character_,
    ind = NA_character_,
    propBsEst = NA_real_,
    propRespEst = NA_real_,
    threshold = NA_real_,
    bias_uns_setting = NA_character_,
    error_message = NA_character_
  )
)

if (all(is.na(bw_tbl$bias_uns_setting)) && any(!is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-bias_uns_setting) |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns")
    )
}

bw_tbl <- bw_tbl |>
  dplyr::filter(!is.na(.data$bias_uns_setting))

sim_grid_key_cols <- setdiff(
  intersect(colnames(sim_grid), colnames(bw_tbl)),
  "sim_id"
)

bw_tbl <- bw_tbl |>
  dplyr::semi_join(
    sim_grid |>
      dplyr::select(dplyr::all_of(sim_grid_key_cols)) |>
      dplyr::distinct(),
    by = sim_grid_key_cols
  )

if (nrow(bw_tbl) == 0L) {
  stop("Cached rows do not match the current simulation grid. Rerun the simulation chunk.")
}

bw_tbl_results_raw <- bw_tbl |>
  dplyr::mutate(
    propBsEst = dplyr::coalesce(
      suppressWarnings(as.numeric(.data$propBsEst)),
      suppressWarnings(as.numeric(.data$propRespEst))
    )
  ) |>
  dplyr::select(
    dplyr::any_of(colnames(sim_grid)),
    iter,
    sample,
    ind,
    method,
    propRespTruth,
    propBsEst,
    threshold,
    error_message
  ) |>
  dplyr::filter(is.finite(.data$threshold) & is.finite(.data$propBsEst)) |>
  dplyr::group_by(sim_id, iter, ind) |>
  dplyr::mutate(
    threshold = unique(threshold[is.finite(threshold)])[[1]],
    propBsEst = unique(propBsEst[is.finite(propBsEst)])[[1]]
  ) |>
  dplyr::ungroup() |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid))),
    iter,
    ind
  ) |>
  dplyr::slice(1) |>
  dplyr::ungroup()

if (nrow(bw_tbl_results_raw) == 0L) {
  warning(
    "No valid manual adaptive frequency results were collated. ",
    "Check the progress log for simulation-level errors."
  )
}

bw_tbl_results_summary <- bw_tbl_results_raw |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    dplyr::across(
      c(threshold, propBsEst),
      list(
        min = ~ min(.x, na.rm = TRUE),
        max = ~ max(.x, na.rm = TRUE),
        mean = ~ mean(.x, na.rm = TRUE),
        median = ~ stats::median(.x, na.rm = TRUE),
        sd = ~ stats::sd(.x, na.rm = TRUE),
        mad = ~ stats::mad(.x, na.rm = TRUE)
      )
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    propBsEst_median_diff = .data$propBsEst_median - .data$prob_response,
    propBsEst_median_diff_rel = (.data$propBsEst_median - .data$prob_response) /
      .data$prob_response,
    propBsEst_median_abs_diff = abs(.data$propBsEst_median_diff),
    propBsEst_median_abs_diff_rel = abs(.data$propBsEst_median_diff_rel)
  )
```

### Estimate plots

```{r}
#| label: plot-estimates
bias_uns_setting_vec_plot <- bias_uns_settings_tbl |>
  dplyr::distinct(.data$bias_uns_setting) |>
  dplyr::pull(.data$bias_uns_setting)

for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
  for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
    plot_tbl <- bw_tbl_results_summary |>
      dplyr::filter(
        .data$mean_pos_setting == !!mean_pos_setting_curr,
        .data$bias_uns_setting == !!bias_uns_setting_curr
      )

    if (nrow(plot_tbl) == 0L) {
      warning(
        "No median estimate results for mean_pos_setting = ",
        mean_pos_setting_curr,
        ", bias_uns_setting = ",
        bias_uns_setting_curr
      )
      next
    }

    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = propBsEst_median,
        colour = factor(bw_adaptive_core),
        linetype = factor(bw_adaptive_extra)
      )
    ) +
      geom_line() +
      geom_point() +
      geom_hline(
        mapping = aes(yintercept = prob_response),
        linetype = "dashed",
        colour = "black"
      ) +
      scale_x_log10() +
      facet_grid(transformation ~ prob_response, scales = "free_y") +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy") +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom"
      ) +
      labs(
        title = paste0(
          "Manual adaptive estimates: mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        ),
        x = "Number of cells",
        y = "Estimated background-subtracted frequency (median)",
        colour = "Core bandwidth",
        linetype = "Extra bandwidth"
      )

    path_p <- projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "est_median",
      paste0(
        "est_vs_n_cell_mean_pos_setting_",
        mean_pos_setting_curr,
        "_bias_uns_",
        bias_uns_setting_curr,
        ".png"
      )
    )
    ggsave(
      filename = path_p,
      plot = p,
      width = 24,
      height = 30,
      units = "cm"
    )
  }
}
```

```{r}
#| label: plot-raw-estimates
for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
  for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
    plot_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        .data$mean_pos_setting == !!mean_pos_setting_curr,
        .data$bias_uns_setting == !!bias_uns_setting_curr
      )

    if (nrow(plot_tbl) == 0L) {
      warning(
        "No raw estimate results for mean_pos_setting = ",
        mean_pos_setting_curr,
        ", bias_uns_setting = ",
        bias_uns_setting_curr
      )
      next
    }

    p <- ggplot(
      plot_tbl,
      aes(
        x = factor(n_cell),
        y = propBsEst,
        colour = factor(bw_adaptive_core)
      )
    ) +
      ggforce::geom_sina(position = "dodge", alpha = 0.45, size = 0.8) +
      geom_hline(
        mapping = aes(yintercept = prob_response),
        linetype = "dashed",
        colour = "black"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy") +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom"
      ) +
      labs(
        title = paste0(
          "Manual adaptive raw estimates: mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        ),
        x = "Number of cells",
        y = "Estimated background-subtracted frequency",
        colour = "Core bandwidth"
      ) +
      facet_grid(
        transformation + bw_adaptive_extra ~ prob_response,
        scales = "free_y"
      )

    path_p <- projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "est_raw",
      paste0(
        "raw_est_vs_n_cell_mean_pos_setting_",
        mean_pos_setting_curr,
        "_bias_uns_",
        bias_uns_setting_curr,
        ".png"
      )
    )
    dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
    ggsave(
      filename = path_p,
      plot = p,
      width = 24,
      height = 36,
      units = "cm"
    )
  }
}
```

### Core by extra heatmap

```{r}
#| label: manual-adaptive-heatmap
heat_tbl <- bw_tbl_results_summary |>
  dplyr::group_by(
    .data$transformation,
    .data$mean_pos_setting,
    .data$bias_uns_setting,
    .data$bw_adaptive_core,
    .data$bw_adaptive_extra
  ) |>
  dplyr::summarise(
    median_abs_rel_error = stats::median(
      .data$propBsEst_median_abs_diff_rel,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
  heat_tbl_curr <- heat_tbl |>
    dplyr::filter(.data$bias_uns_setting == !!bias_uns_setting_curr)

  if (nrow(heat_tbl_curr) == 0L) {
    next
  }

  p_heat <- ggplot(
    heat_tbl_curr,
    aes(
      x = factor(bw_adaptive_core),
      y = factor(bw_adaptive_extra),
      fill = median_abs_rel_error
    )
  ) +
    geom_tile() +
    facet_grid(mean_pos_setting ~ transformation) +
    scale_fill_viridis_c(trans = "log10", option = "magma", direction = -1) +
    cowplot::theme_cowplot() +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA),
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      title = paste0(
        "Manual adaptive bandwidth grid: bias_uns_setting = ",
        bias_uns_setting_curr
      ),
      x = "Core bandwidth",
      y = "Extra bandwidth",
      fill = "Median absolute\nrelative error"
    )

  path_p_heat <- projr::projr_path_get(
    "cache",
    "sim",
    "bw",
    "est",
    "adaptive",
    "heatmap",
    paste0("core_extra_heatmap_bias_uns_", bias_uns_setting_curr, ".png")
  )
  dir.create(dirname(path_p_heat), recursive = TRUE, showWarnings = FALSE)
  ggsave(path_p_heat, p_heat, width = 24, height = 18, units = "cm")
}
```

### Threshold distributions

```{r}
#| label: threshold-distributions
threshold_summary_tbl <- bw_tbl_results_raw |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    threshold_median = stats::median(.data$threshold, na.rm = TRUE),
    threshold_iqr_lower = stats::quantile(.data$threshold, 0.25, na.rm = TRUE),
    threshold_iqr_upper = stats::quantile(.data$threshold, 0.75, na.rm = TRUE),
    threshold_sd = stats::sd(.data$threshold, na.rm = TRUE),
    threshold_mad = stats::mad(.data$threshold, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    threshold_iqr_length = .data$threshold_iqr_upper - .data$threshold_iqr_lower
  )

for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
  for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
    plot_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        .data$mean_pos_setting == !!mean_pos_setting_curr,
        .data$bias_uns_setting == !!bias_uns_setting_curr,
        is.finite(.data$threshold)
      ) |>
      dplyr::mutate(
        bw_pair = paste0(
          "core ", .data$bw_adaptive_core,
          ", extra ", .data$bw_adaptive_extra
        )
      )

    if (nrow(plot_tbl) == 0L) {
      next
    }

    p_thresh <- ggplot(
      plot_tbl,
      aes(x = bw_pair, y = threshold, colour = factor(bw_adaptive_core))
    ) +
      ggforce::geom_sina(alpha = 0.25, size = 0.6) +
      facet_grid(transformation ~ prob_response, scales = "free_y") +
      coord_flip() +
      cowplot::theme_cowplot() +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom",
        axis.text.y = element_text(size = 5)
      ) +
      labs(
        title = paste0(
          "Threshold distributions: mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        ),
        x = "Core and extra bandwidth",
        y = "Threshold",
        colour = "Core bandwidth"
      )

    path_p_thresh <- projr::projr_path_get(
      "cache",
      "sim",
      "bw",
      "est",
      "adaptive",
      "thresholds",
      paste0(
        "thresholds_mean_pos_setting_",
        mean_pos_setting_curr,
        "_bias_uns_",
        bias_uns_setting_curr,
        ".png"
      )
    )
    dir.create(dirname(path_p_thresh), recursive = TRUE, showWarnings = FALSE)
    ggsave(path_p_thresh, p_thresh, width = 24, height = 30, units = "cm")
  }
}
```

### Summary table

```{r}
#| label: summary-table
method_rank_tbl <- bw_tbl_results_summary |>
  dplyr::group_by(
    .data$bias_uns_setting,
    .data$bias_uns,
    .data$transformation,
    .data$mean_pos_setting,
    .data$bw_adaptive_core,
    .data$bw_adaptive_extra
  ) |>
  dplyr::summarise(
    n_scenarios = dplyr::n(),
    median_abs_error = stats::median(.data$propBsEst_median_abs_diff, na.rm = TRUE),
    median_abs_rel_error = stats::median(.data$propBsEst_median_abs_diff_rel, na.rm = TRUE),
    median_threshold_sd = stats::median(.data$threshold_sd, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(
    .data$bias_uns_setting,
    .data$transformation,
    .data$mean_pos_setting,
    .data$median_abs_rel_error
  )

method_rank_tbl |>
  dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))) |>
  dplyr::group_by(
    .data$bias_uns_setting,
    .data$transformation,
    .data$mean_pos_setting
  ) |>
  dplyr::slice_head(n = 20) |>
  dplyr::ungroup() |>
  knitr::kable()
```
