---
title: Manual adaptive bandwidth frequency simulation
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)

`%||%` <- function(x, y) {
  if (is.null(x)) y else x
}

.is_dev_profile <- function() {
  "dev" %in% projr::projr_profile_get()
}

.is_quick_profile <- function() {
  "quick" %in% projr::projr_profile_get()
}

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Checking what the fixed adaptive bandwidths do

This is the manual adaptive analogue of the fixed-bandwidth frequency simulation.
Instead of fixing one scalar bandwidth, we fix two numeric bandwidths: one for the background core and one for the high-expression extra component.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()
transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# Fix the adaptive bandwidths. These are deliberately numeric settings, not labels.
bw_adaptive_core_vec <- c(0.0025, 0.005, 0.01, 0.02, 0.05, 0.1, 0.25)
bw_adaptive_extra_vec <- c(0.01, 0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1)

# A missing crossover uses the adaptive component-density weighting. Keep this
# fixed here so the grid varies only core and extra bandwidths.
bw_adaptive_crossover <- NULL
bw_adaptive_transition_width <- 0

# Bias settings. The numeric values are transformation-specific but the setting
# label is carried through to plotting and summaries.
bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  )
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))

# Gating/bandwidth settings
bw_mtd <- "hpi1Norm"
bw_adj <- 1
bw_ncell_min <- 100
bw_ncell_max <- 1e5
bw_cluster <- NULL
bw_min <- "none"
bw_max <- "none"
bw_fallback <- "auto"
min_cell <- 100
max_pos_prob_x <- Inf
gate_combn <- "min"
tol_clust <- NULL

bw_adaptive_density_n <- NULL
bw_adaptive_pad_frac <- 0.15
norm_peak_frac <- 0.1
norm_peak_min_rel <- 0.75
norm_extra_frac <- 0.2
norm_extra_max <- Inf
norm_extra_jitter_frac <- 0.25
norm_lambda <- seq(-2, 2, length.out = 81)
norm_density_n <- 512L
norm_excess_bw_mtd <- "hpi3"
norm_excess_ncell <- 10000L
norm_adaptive_ncell <- 2500L
norm_mtd <- "moments"

# Similar to the earlier fixed-bandwidth frequency simulation.
n_sample_run <- 5L
n_iter_run <- 5L
n_marker_run <- 1L
n_condition_run <- 2L
n_cluster_run <- 2L
prob_exact <- TRUE
cov_ev_min <- 1.5
cov_ev_max <- 1.5
calc_cyt_pos_gates <- FALSE
calc_single_pos_gates <- FALSE

if (.is_quick_profile()) {
  n_cell_stim_vec <- c(1e3, 5e4)
  prob_response_vec <- c(0.002, 0.02)
  bw_adaptive_core_vec <- c(0.02, 0.05, 0.1)
  bw_adaptive_extra_vec <- c(0.1, 0.25, 0.5)
  condition_perturbation_sd_vec <- 0
  n_sample_run <- 2L
  n_iter_run <- 2L
}
```

### Varying manual adaptive bandwidths

```{r}
#| label: bw-manual-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_adaptive_core = bw_adaptive_core_vec,
  bw_adaptive_extra = bw_adaptive_extra_vec,
  bias_uns = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_adaptive_core,
    bw_adaptive_extra,
    bias_uns_setting,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    .data$bw_adaptive_extra >= .data$bw_adaptive_core
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::filter(
    (
      .data$transformation == "gamma" &
        .data$bw_adaptive_core <= 0.05 &
        .data$bw_adaptive_extra <= 0.25
    ) |
      (
        .data$transformation != "gamma" &
          .data$bw_adaptive_core >= 0.02 &
          .data$bw_adaptive_extra >= 0.05
      )
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

bad_mean_pos_grid <- sim_grid |>
  dplyr::anti_join(
    mean_pos_settings_tbl,
    by = c("transformation", "mean_pos_setting", "mean_pos")
  )

if (nrow(bad_mean_pos_grid) > 0L) {
  stop(
    "sim_grid contains transformation/mean_pos_setting/mean_pos combinations ",
    "not returned by .simMiscGetMeanPosTbl()."
  )
}

bad_bias_grid <- sim_grid |>
  dplyr::anti_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns_setting", "bias_uns")
  )

if (nrow(bad_bias_grid) > 0L) {
  stop(
    "sim_grid contains transformation/bias_uns_setting/bias_uns combinations ",
    "not returned by bias_uns_settings_tbl."
  )
}

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if (.is_dev_profile()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bw_adaptive_core == 0.05,
      bw_adaptive_extra == 0.25,
      bias_uns_setting == "low",
      condition_perturbation_sd == 0
    )
  n_sample_run <- 2L
  n_iter_run <- 1L
}

sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))
```

### Local simulation helper

```{r}
#| label: adaptive-simulation-helper
.null_if_not_finite <- function(x) {
  x <- suppressWarnings(as.numeric(x)[1])
  if (is.finite(x)) x else NULL
}

.read_rds_or_null <- function(path) {
  if (!file.exists(path)) {
    return(NULL)
  }
  tryCatch(
    readRDS(path),
    error = function(e) NULL
  )
}

.sample_from_ind <- function(ind, n_condition) {
  ind_num <- suppressWarnings(as.numeric(ind))
  as.character(((ind_num - 1) %/% n_condition) + 1)
}

.add_missing_cols <- function(.data, cols) {
  for (nm in names(cols)) {
    if (!nm %in% names(.data)) {
      .data[[nm]] <- cols[[nm]]
    }
  }
  .data
}

.loc_detail_gate_point <- function(
  detail_level,
  loc_generated,
  loc_generated_direct,
  loc_source
) {
  dplyr::case_when(
    detail_level %in% "condition" &
      loc_generated %in% TRUE &
      loc_generated_direct %in% TRUE ~ "condition_direct_local_fdr",
    detail_level %in% "condition" &
      !(loc_generated %in% TRUE) ~ "condition_fallback_high_value",
    detail_level %in% "sample" &
      loc_generated %in% TRUE &
      loc_source %in% "combined" ~ "sample_combined_from_other_stim_conditions",
    detail_level %in% "sample" &
      loc_generated %in% TRUE &
      loc_source %in% "prejoin" ~ "sample_prejoin_from_joined_stim_conditions",
    detail_level %in% "sample" &
      loc_generated %in% TRUE ~ "sample_final_local_fdr",
    detail_level %in% "sample" &
      !(loc_generated %in% TRUE) ~ "sample_fallback_high_value",
    TRUE ~ NA_character_
  )
}

.read_loc_details <- function(
  path_project,
  n_sample,
  n_condition
) {
  path_dir_int_init <- file.path(path_project, "intermediateData", "init")
  if (!dir.exists(path_dir_int_init)) {
    return(tibble::tibble())
  }

  chnl_vec <- list.dirs(path_dir_int_init, full.names = FALSE, recursive = FALSE)
  if (length(chnl_vec) == 0L) {
    return(tibble::tibble())
  }

  detail_tbl <- purrr::map_df(chnl_vec, function(chnl) {
    stage_chnl <- file.path(path_dir_int_init, chnl)

    purrr::map_df(seq_len(n_sample), function(sample_curr) {
      ind_batch <- seq(
        (sample_curr - 1L) * n_condition + 1L,
        sample_curr * n_condition
      )
      ind_stim <- ind_batch[-1L]
      ind_combined <- paste0(ind_stim, collapse = "_")

      condition_tbl <- purrr::map_df(ind_stim, function(ind) {
        path_ind <- file.path(stage_chnl, "ind", as.character(ind))
        out <- .read_rds_or_null(file.path(path_ind, "locDetailCondition.rds"))
        if (!is.data.frame(out) || nrow(out) == 0L) {
          return(tibble::tibble())
        }
        out
      })

      sample_tbl <- .read_rds_or_null(
        file.path(stage_chnl, "ind", ind_combined, "locDetailSample.rds")
      )
      if (!is.data.frame(sample_tbl) || nrow(sample_tbl) == 0L) {
        sample_tbl <- tibble::tibble()
      }

      dplyr::bind_rows(condition_tbl, sample_tbl)
    })
  })

  if (!is.data.frame(detail_tbl) || nrow(detail_tbl) == 0L) {
    return(tibble::tibble())
  }

  detail_tbl <- .add_missing_cols(
    detail_tbl,
    list(
      detailLevel = NA_character_,
      stage = NA_character_,
      chnl = NA_character_,
      ind = NA_character_,
      threshold = NA_real_,
      thresholdOrigin = NA_character_,
      locGenerated = NA,
      locGeneratedDirect = NA,
      locSource = NA_character_,
      locReason = NA_character_,
      bias = NA_real_,
      propBsEst = NA_real_,
      propBsDiff = NA_real_,
      nCellStim = NA_integer_,
      nCellUns = NA_integer_,
      propStim = NA_real_,
      propUns = NA_real_,
      propBs = NA_real_
    )
  )

  condition_threshold_tbl <- detail_tbl |>
    dplyr::filter(.data$detailLevel %in% "condition") |>
    dplyr::transmute(
      chnl = .data$chnl,
      ind = as.character(.data$ind),
      thresholdCondition = suppressWarnings(as.numeric(.data$threshold)),
      thresholdOriginCondition = .data$thresholdOrigin,
      locGeneratedCondition = .data$locGenerated %in% TRUE,
      locSourceCondition = .data$locSource,
      locReasonCondition = .data$locReason
    )

  detail_tbl |>
    dplyr::mutate(
      ind = as.character(.data$ind),
      sample = .sample_from_ind(.data$ind, n_condition),
      method = paste0("loc_", .data$detailLevel),
      propRespEst = .data$propBs,
      propBsEst = dplyr::coalesce(
        suppressWarnings(as.numeric(.data$propBsEst)),
        suppressWarnings(as.numeric(.data$propBs))
      ),
      nCellStim = suppressWarnings(as.numeric(.data$nCellStim)),
      nCellUns = suppressWarnings(as.numeric(.data$nCellUns)),
      propStim = suppressWarnings(as.numeric(.data$propStim)),
      propUns = suppressWarnings(as.numeric(.data$propUns)),
      nPosStim = dplyr::if_else(
        is.finite(.data$propStim) & is.finite(.data$nCellStim),
        as.integer(round(.data$propStim * .data$nCellStim)),
        NA_integer_
      ),
      nPosUns = dplyr::if_else(
        is.finite(.data$propUns) & is.finite(.data$nCellUns),
        as.integer(round(.data$propUns * .data$nCellUns)),
        NA_integer_
      ),
      gateReturnPoint = .loc_detail_gate_point(
        detail_level = .data$detailLevel,
        loc_generated = .data$locGenerated,
        loc_generated_direct = .data$locGeneratedDirect,
        loc_source = .data$locSource
      )
    ) |>
    dplyr::left_join(
      condition_threshold_tbl,
      by = c("chnl", "ind")
    ) |>
    dplyr::mutate(
      thresholdBeforeSampleCombining = dplyr::if_else(
        .data$detailLevel %in% "sample",
        .data$thresholdCondition,
        NA_real_
      ),
      thresholdChangedBySampleCombining = dplyr::case_when(
        .data$detailLevel %in% "sample" &
          is.finite(.data$threshold) &
          is.finite(.data$thresholdCondition) ~
          abs(.data$threshold - .data$thresholdCondition) >
            sqrt(.Machine$double.eps),
        .data$detailLevel %in% "sample" ~ NA,
        TRUE ~ FALSE
      )
    ) |>
    dplyr::select(
      sample,
      ind,
      chnl,
      method,
      propRespEst,
      propBsEst,
      detailLevel,
      stage,
      threshold,
      thresholdOrigin,
      gateReturnPoint,
      locGenerated,
      locGeneratedDirect,
      locSource,
      locReason,
      thresholdBeforeSampleCombining,
      thresholdChangedBySampleCombining,
      thresholdCondition,
      thresholdOriginCondition,
      locGeneratedCondition,
      locSourceCondition,
      locReasonCondition,
      nCellStim,
      nCellUns,
      nPosStim,
      nPosUns,
      propStim,
      propUns,
      propBs,
      propBsDiff,
      bias
    )
}

.run_manual_adaptive_freq_bs <- function(
  transformation,
  prob_response,
  n_cell,
  mean_pos_setting,
  mean_pos,
  bw_adaptive_core,
  bw_adaptive_extra,
  bias_uns_setting,
  bias_uns,
  sample_perturbation_sd,
  condition_perturbation_sd,
  cluster_perturbation_sd,
  background_relative_to_response,
  n_cell_uns_relative_to_stim,
  sim_id,
  nSample = n_sample_run,
  nMarker = n_marker_run,
  nCondition = n_condition_run,
  nCluster = n_cluster_run,
  nIter = n_iter_run,
  bwAdaptiveCrossover = bw_adaptive_crossover,
  bwAdaptiveTransitionWidth = bw_adaptive_transition_width
) {
  purrr::map_df(seq_len(nIter), function(iter_num) {
    n_cell_uns <- round(n_cell * n_cell_uns_relative_to_stim)
    n_cell_by_condition <- c(n_cell_uns, n_cell)
    transformation_func <- .simMiscGetTrans(transformation)

    mean_expr_mat <- matrix(
      c(0, mean_pos),
      byrow = TRUE,
      ncol = 1
    )
    cluster_label_vec <- c("gn", "gp")
    prob_response_uns <- prob_response * background_relative_to_response
    prob_vec_uns <- c(1 - prob_response_uns, prob_response_uns)
    prob_response_vec_by_stim_condition <- list(c(-prob_response, prob_response))

    out_list_experiment <- simCytExperiment(
      nSample = nSample,
      nMarker = nMarker,
      nCondition = nCondition,
      nCluster = nCluster,
      nCellByCondition = n_cell_by_condition,
      transformationFunc = transformation_func,
      mixtureType = "gaussianOnly",
      meanExprMat = mean_expr_mat,
      clusterLabelVec = cluster_label_vec,
      probVecUns = prob_vec_uns,
      probExact = prob_exact,
      probResponseVecByStimCondition = prob_response_vec_by_stim_condition,
      conditionPerturbationSd = condition_perturbation_sd,
      clusterPerturbationSd = cluster_perturbation_sd,
      samplePerturbationSd = sample_perturbation_sd,
      covEvMin = cov_ev_min,
      covEvMax = cov_ev_max
    )

    fs <- as(out_list_experiment[["flowFrameList"]], "flowSet")
    gs <- flowWorkspace::GatingSet(fs)
    labels_list <- out_list_experiment[["labelsList"]]

    path_project <- file.path(
      tempdir(),
      "stimgate",
      "sim-bw",
      "manual-adaptive",
      paste0(
        "sim-", sim_id,
        "-iter-", iter_num,
        "-pid-", Sys.getpid(),
        "-", format(Sys.time(), "%Y%m%d%H%M%S"),
        "-", sample.int(1e7, 1)
      )
    )

    on.exit(
      {
        if (dir.exists(path_project)) {
          unlink(path_project, recursive = TRUE)
        }
      },
      add = TRUE
    )

    if (dir.exists(path_project)) {
      unlink(path_project, recursive = TRUE)
    }
    dir.create(path_project, recursive = TRUE, showWarnings = FALSE)

    batch_list <- lapply(seq_len(nSample), function(i) {
      seq((i - 1L) * nCondition + 1L, i * nCondition)
    })

    Sys.setenv("STIMGATE_INTERMEDIATE" = "TRUE")
    invisible(gateStim(
      .data = gs,
      pathProject = path_project,
      popGate = "root",
      batchList = batch_list,
      marker = paste0("MarkerF", seq_len(nMarker)),
      calcCytPosGates = calc_cyt_pos_gates,
      calcSinglePosGates = calc_single_pos_gates,
      biasUns = bias_uns,
      bw = NULL,
      bwFallback = bw_fallback,
      bwMin = bw_min,
      bwMax = bw_max,
      bwMtd = bw_mtd,
      bwAdj = bw_adj,
      bwNcellMin = bw_ncell_min,
      bwNcellMax = bw_ncell_max,
      bwCluster = bw_cluster,
      bwAdaptive = TRUE,
      bwAdaptiveDensityN = bw_adaptive_density_n,
      bwAdaptivePadFrac = bw_adaptive_pad_frac,
      bwAdaptiveCore = bw_adaptive_core,
      bwAdaptiveExtra = bw_adaptive_extra,
      bwAdaptiveCrossover = .null_if_not_finite(bwAdaptiveCrossover),
      bwAdaptiveTransitionWidth = bwAdaptiveTransitionWidth,
      normPeakFrac = norm_peak_frac,
      normPeakMinRel = norm_peak_min_rel,
      normExtraFrac = norm_extra_frac,
      normExtraMax = norm_extra_max,
      normExtraJitterFrac = norm_extra_jitter_frac,
      normLambda = norm_lambda,
      normDensityN = norm_density_n,
      normExcessBwMtd = norm_excess_bw_mtd,
      normExcessNcell = norm_excess_ncell,
      normAdaptiveNcell = norm_adaptive_ncell,
      normMtd = norm_mtd,
      minCell = min_cell,
      maxPosProbX = max_pos_prob_x,
      tolClust = tol_clust,
      gateCombn = gate_combn
    ))

    stopifnot(file.exists(file.path(path_project, "gateStats.rds")))

    prop_bs_tbl_truth <- purrr::map_df(seq_len(nSample), function(sample_curr) {
      ind_uns <- (sample_curr - 1L) * nCondition + 1L
      ind_stim <- seq.int(ind_uns + 1L, sample_curr * nCondition)
      label_vec_uns <- labels_list[[ind_uns]]
      prop_uns_truth <- sum(grepl("^gp$", label_vec_uns)) / length(label_vec_uns)

      purrr::map_df(ind_stim, function(ind) {
        label_vec_stim <- labels_list[[ind]]
        prop_stim_truth <- sum(grepl("^gp$", label_vec_stim)) /
          length(label_vec_stim)
        tibble::tibble(
          sample = as.character(sample_curr),
          ind = as.character(ind),
          chnl = "F1",
          propStimTruth = prop_stim_truth,
          propUnsTruth = prop_uns_truth,
          propRespTruth = prop_stim_truth - prop_uns_truth
        )
      })
    })

    path_dir_int_init <- file.path(path_project, "intermediateData", "init")
    chnl_vec <- list.dirs(path_dir_int_init, full.names = FALSE, recursive = FALSE)

    prop_bs_tbl_est_smooth <- purrr::map_df(chnl_vec, function(chnl) {
      ind_vec_stim <- unlist(lapply(seq_len(nSample), function(sample_curr) {
        ind_uns <- (sample_curr - 1L) * nCondition + 1L
        seq.int(ind_uns + 1L, sample_curr * nCondition)
      }))

      purrr::map_df(ind_vec_stim, function(ind) {
        path_ind <- file.path(path_dir_int_init, chnl, "ind", as.character(ind))
        prob_smooth <- .read_rds_or_null(file.path(path_ind, "dataMod.rds"))
        if (!inherits(prob_smooth, "data.frame")) {
          return(tibble::tibble(
            sample = .sample_from_ind(ind, nCondition),
            ind = as.character(ind),
            chnl = chnl,
            method = c("propRespSmooth", "propRespPred"),
            propRespEst = c(0, 0),
            propBsEst = c(0, 0),
            nCellStim = n_cell,
            nCellUns = n_cell_uns,
            nPosStim = NA_integer_,
            nPosUns = NA_integer_,
            threshold = NA_real_,
            thresholdOrigin = NA_character_,
            gateReturnPoint = NA_character_,
            locGenerated = NA,
            locGeneratedDirect = NA,
            locSource = NA_character_,
            locReason = NA_character_,
            detailLevel = NA_character_,
            stage = NA_character_,
            thresholdBeforeSampleCombining = NA_real_,
            thresholdChangedBySampleCombining = NA
          ))
        }
        ncell_resp_smooth <- if (nrow(prob_smooth) > 0L) {
          sum(prob_smooth$probSmooth)
        } else {
          0
        }
        ncell_resp_pred <- if (nrow(prob_smooth) > 0L) {
          sum(prob_smooth$pred)
        } else {
          0
        }
        tibble::tibble(
          sample = .sample_from_ind(ind, nCondition),
          ind = as.character(ind),
          chnl = chnl,
          method = c("propRespSmooth", "propRespPred"),
          propRespEst = c(
            ncell_resp_smooth / n_cell,
            ncell_resp_pred / n_cell
          ),
          propBsEst = c(
            ncell_resp_smooth / n_cell,
            ncell_resp_pred / n_cell
          ),
          nCellStim = n_cell,
          nCellUns = n_cell_uns,
          nPosStim = NA_integer_,
          nPosUns = NA_integer_,
          threshold = NA_real_,
          thresholdOrigin = NA_character_,
          gateReturnPoint = NA_character_,
          locGenerated = NA,
          locGeneratedDirect = NA,
          locSource = NA_character_,
          locReason = NA_character_,
          detailLevel = NA_character_,
          stage = NA_character_,
          thresholdBeforeSampleCombining = NA_real_,
          thresholdChangedBySampleCombining = NA
        )
      })
    })

    prop_bs_tbl_detailed <- .read_loc_details(
      path_project = path_project,
      n_sample = nSample,
      n_condition = nCondition
    ) |>
      dplyr::filter(.data$detailLevel %in% c("condition", "sample"))

    prop_bs_tbl_est <- dplyr::bind_rows(
      prop_bs_tbl_est_smooth,
      prop_bs_tbl_detailed
    )

    comparison_tbl <- prop_bs_tbl_truth |>
      dplyr::left_join(
        prop_bs_tbl_est,
        by = c("sample", "ind", "chnl")
      )

    comparison_tbl |>
      dplyr::mutate(
        transformation = transformation,
        prob_response = prob_response,
        n_cell = n_cell,
        mean_pos_setting = mean_pos_setting,
        mean_pos = mean_pos,
        bw_adaptive_core = bw_adaptive_core,
        bw_adaptive_extra = bw_adaptive_extra,
        bias_uns_setting = bias_uns_setting,
        bias_uns = bias_uns,
        sample_perturbation_sd = sample_perturbation_sd,
        condition_perturbation_sd = condition_perturbation_sd,
        cluster_perturbation_sd = cluster_perturbation_sd,
        background_relative_to_response = background_relative_to_response,
        n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim,
        sim_id = sim_id,
        iter = iter_num,
        bwMtd = bw_mtd,
        bwAdaptive = TRUE,
        bwAdaptiveCore = bw_adaptive_core,
        bwAdaptiveExtra = bw_adaptive_extra,
        bwAdaptiveCrossover = .null_if_not_finite(bwAdaptiveCrossover) %||% NA_real_,
        bwAdaptiveTransitionWidth = bwAdaptiveTransitionWidth,
        error_message = NA_character_
      ) |>
      dplyr::select(
        transformation,
        prob_response,
        n_cell,
        mean_pos_setting,
        mean_pos,
        bw_adaptive_core,
        bw_adaptive_extra,
        bias_uns_setting,
        bias_uns,
        sample_perturbation_sd,
        condition_perturbation_sd,
        cluster_perturbation_sd,
        background_relative_to_response,
        n_cell_uns_relative_to_stim,
        sim_id,
        iter,
        chnl,
        sample,
        ind,
        method,
        propRespTruth,
        propRespEst,
        propBsEst,
        threshold,
        dplyr::everything()
      )
  })
}
```

### Manual loop for debugging

```{r}
#| label: bw-manual-loop
#| eval: false
bw_list_raw <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))

  row <- sim_grid[i, , drop = FALSE]
  sim_res <- do.call(
    .run_manual_adaptive_freq_bs,
    as.list(row)
  )

  bw_list_raw[[i]] <- sim_res
}

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "freq_bs",
    "adaptive",
    "bw_list_raw_loop.rds"
  )
)

bw_list_raw |>
  purrr::list_rbind() |>
  print(width = Inf)
```

### Parallel processing

```{r}
#| label: bw-manual-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "bw",
  "adaptive",
  "progress.txt"
)
dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

future::plan(
  future::multisession,
  workers = max(1, future::availableCores() - 1)
)

bw_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_adaptive_core,
      bw_adaptive_extra,
      bias_uns_setting,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | core = ", bw_adaptive_core,
        " | extra = ", bw_adaptive_extra,
        " | bias_setting = ", bias_uns_setting,
        " | bias = ", bias_uns,
        " | sample_sd = ", sample_perturbation_sd,
        " | condition_sd = ", condition_perturbation_sd,
        " | cluster_sd = ", cluster_perturbation_sd,
        " | bg_rel = ", background_relative_to_response,
        " | n_uns_rel = ", n_cell_uns_relative_to_stim,
        " | sim_id = ", sim_id
      )

      cat(paste0("Running: ", settings_row, "\n"), file = path_progress_file, append = TRUE)

      out <- tryCatch(
        {
          sim_res <- .run_manual_adaptive_freq_bs(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw_adaptive_core = bw_adaptive_core,
            bw_adaptive_extra = bw_adaptive_extra,
            bias_uns_setting = bias_uns_setting,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response = background_relative_to_response,
            n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim,
            sim_id = sim_id
          )

          cat(paste0("Completed: ", settings_row, "\n"), file = path_progress_file, append = TRUE)
          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          err_msg <- paste0("Error [", settings_row, "]: ", e$message, "\n")
          cat(err_msg, file = path_progress_file, append = TRUE)
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw_adaptive_core = bw_adaptive_core,
            bw_adaptive_extra = bw_adaptive_extra,
            bias_uns_setting = bias_uns_setting,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response = background_relative_to_response,
            n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            method = NA_character_,
            propRespTruth = NA_real_,
            propRespEst = NA_real_,
            propBsEst = NA_real_,
            threshold = NA_real_,
            thresholdOrigin = NA_character_,
            gateReturnPoint = NA_character_,
            locGenerated = NA,
            locGeneratedDirect = NA,
            locSource = NA_character_,
            locReason = NA_character_,
            detailLevel = NA_character_,
            nCellStim = NA_real_,
            nCellUns = NA_real_,
            nPosStim = NA_integer_,
            nPosUns = NA_integer_,
            error_message = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "freq_bs",
    "adaptive",
    "bw_list_raw.rds"
  )
)
```

### Collate results

```{r}
#| label: bw-manual-collate
.add_missing_cols <- function(.data, cols) {
  for (nm in names(cols)) {
    if (!nm %in% names(.data)) {
      .data[[nm]] <- cols[[nm]]
    }
  }
  .data
}

if (exists("bw_list_raw") && length(bw_list_raw) > 0L) {
  bw_list_raw_collate <- bw_list_raw
} else {
  path_vec <- c(
    projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "adaptive",
      "bw_list_raw.rds",
      safe = TRUE
    ),
    projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "adaptive",
      "bw_list_raw.rds",
      safe = FALSE
    )
  )
  path_vec <- path_vec[file.exists(path_vec)]
  if (length(path_vec) == 0L) {
    stop("Could not find bw_list_raw.rds, and bw_list_raw is not available in memory.")
  }
  bw_list_raw_collate <- readRDS(file = path_vec[[1]])
}

bw_tbl <- bw_list_raw_collate |>
  purrr::list_rbind()

bw_tbl <- .add_missing_cols(
  bw_tbl,
  list(
    iter = NA_integer_,
    sample = NA_character_,
    ind = NA_character_,
    propBsEst = NA_real_,
    propRespEst = NA_real_,
    threshold = NA_real_,
    bias_uns_setting = NA_character_,
    error_message = NA_character_
  )
)

if (all(is.na(bw_tbl$bias_uns_setting)) && any(!is.na(bw_tbl$bias_uns))) {
  bw_tbl <- bw_tbl |>
    dplyr::select(-bias_uns_setting) |>
    dplyr::left_join(
      bias_uns_settings_tbl,
      by = c("transformation", "bias_uns")
    )
}

bw_tbl <- bw_tbl |>
  dplyr::filter(!is.na(.data$bias_uns_setting))

sim_grid_key_cols <- setdiff(
  intersect(colnames(sim_grid), colnames(bw_tbl)),
  "sim_id"
)

bw_tbl <- bw_tbl |>
  dplyr::semi_join(
    sim_grid |>
      dplyr::select(dplyr::all_of(sim_grid_key_cols)) |>
      dplyr::distinct(),
    by = sim_grid_key_cols
  )

if (nrow(bw_tbl) == 0L) {
  stop("Cached rows do not match the current simulation grid. Rerun the simulation chunk.")
}

bw_tbl_results_raw <- bw_tbl |>
  dplyr::mutate(
    propBsEst = dplyr::coalesce(
      suppressWarnings(as.numeric(.data$propBsEst)),
      suppressWarnings(as.numeric(.data$propRespEst))
    )
  ) |>
  dplyr::select(
    dplyr::any_of(colnames(sim_grid)),
    iter,
    sample,
    ind,
    method,
    propRespTruth,
    propBsEst,
    threshold,
    error_message
  ) |>
  dplyr::filter(is.finite(.data$threshold) & is.finite(.data$propBsEst)) |>
  dplyr::group_by(sim_id, iter, ind) |>
  dplyr::mutate(
    threshold = unique(threshold[is.finite(threshold)])[[1]],
    propBsEst = unique(propBsEst[is.finite(propBsEst)])[[1]]
  ) |>
  dplyr::ungroup() |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid))),
    iter,
    ind
  ) |>
  dplyr::slice(1) |>
  dplyr::ungroup()

if (nrow(bw_tbl_results_raw) == 0L) {
  warning(
    "No valid manual adaptive frequency results were collated. ",
    "Check the progress log for simulation-level errors."
  )
}

bw_tbl_results_summary <- bw_tbl_results_raw |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    dplyr::across(
      c(threshold, propBsEst),
      list(
        min = ~ min(.x, na.rm = TRUE),
        max = ~ max(.x, na.rm = TRUE),
        mean = ~ mean(.x, na.rm = TRUE),
        median = ~ stats::median(.x, na.rm = TRUE),
        sd = ~ stats::sd(.x, na.rm = TRUE),
        mad = ~ stats::mad(.x, na.rm = TRUE)
      )
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    propBsEst_median_diff = .data$propBsEst_median - .data$prob_response,
    propBsEst_median_diff_rel = (.data$propBsEst_median - .data$prob_response) /
      .data$prob_response,
    propBsEst_median_abs_diff = abs(.data$propBsEst_median_diff),
    propBsEst_median_abs_diff_rel = abs(.data$propBsEst_median_diff_rel)
  )
```

### Estimate plots

```{r}
#| label: plot-estimates
bias_uns_setting_vec_plot <- bias_uns_settings_tbl |>
  dplyr::distinct(.data$bias_uns_setting) |>
  dplyr::pull(.data$bias_uns_setting)

for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
  for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
    plot_tbl <- bw_tbl_results_summary |>
      dplyr::filter(
        .data$mean_pos_setting == !!mean_pos_setting_curr,
        .data$bias_uns_setting == !!bias_uns_setting_curr
      )

    if (nrow(plot_tbl) == 0L) {
      warning(
        "No median estimate results for mean_pos_setting = ",
        mean_pos_setting_curr,
        ", bias_uns_setting = ",
        bias_uns_setting_curr
      )
      next
    }

    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = propBsEst_median,
        colour = factor(bw_adaptive_core),
        linetype = factor(bw_adaptive_extra)
      )
    ) +
      geom_line() +
      geom_point() +
      geom_hline(
        mapping = aes(yintercept = prob_response),
        linetype = "dashed",
        colour = "black"
      ) +
      scale_x_log10() +
      facet_grid(transformation ~ prob_response, scales = "free_y") +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy") +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom"
      ) +
      labs(
        title = paste0(
          "Manual adaptive estimates: mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        ),
        x = "Number of cells",
        y = "Estimated background-subtracted frequency (median)",
        colour = "Core bandwidth",
        linetype = "Extra bandwidth"
      )

    path_p <- projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "adaptive",
      "est_median",
      paste0(
        "est_vs_n_cell_mean_pos_setting_",
        mean_pos_setting_curr,
        "_bias_uns_",
        bias_uns_setting_curr,
        ".png"
      )
    )
    ggsave(
      filename = path_p,
      plot = p,
      width = 24,
      height = 30,
      units = "cm"
    )
  }
}
```

```{r}
#| label: plot-raw-estimates
for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
  for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
    plot_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        .data$mean_pos_setting == !!mean_pos_setting_curr,
        .data$bias_uns_setting == !!bias_uns_setting_curr
      )

    if (nrow(plot_tbl) == 0L) {
      warning(
        "No raw estimate results for mean_pos_setting = ",
        mean_pos_setting_curr,
        ", bias_uns_setting = ",
        bias_uns_setting_curr
      )
      next
    }

    p <- ggplot(
      plot_tbl,
      aes(
        x = factor(n_cell),
        y = propBsEst,
        colour = factor(bw_adaptive_core)
      )
    ) +
      ggforce::geom_sina(position = "dodge", alpha = 0.45, size = 0.8) +
      geom_hline(
        mapping = aes(yintercept = prob_response),
        linetype = "dashed",
        colour = "black"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy") +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom"
      ) +
      labs(
        title = paste0(
          "Manual adaptive raw estimates: mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        ),
        x = "Number of cells",
        y = "Estimated background-subtracted frequency",
        colour = "Core bandwidth"
      ) +
      facet_grid(
        transformation + bw_adaptive_extra ~ prob_response,
        scales = "free_y"
      )

    path_p <- projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "adaptive",
      "est_raw",
      paste0(
        "raw_est_vs_n_cell_mean_pos_setting_",
        mean_pos_setting_curr,
        "_bias_uns_",
        bias_uns_setting_curr,
        ".png"
      )
    )
    dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
    ggsave(
      filename = path_p,
      plot = p,
      width = 24,
      height = 36,
      units = "cm"
    )
  }
}
```

### Core by extra heatmap

```{r}
#| label: manual-adaptive-heatmap
heat_tbl <- bw_tbl_results_summary |>
  dplyr::group_by(
    .data$transformation,
    .data$mean_pos_setting,
    .data$bias_uns_setting,
    .data$bw_adaptive_core,
    .data$bw_adaptive_extra
  ) |>
  dplyr::summarise(
    median_abs_rel_error = stats::median(
      .data$propBsEst_median_abs_diff_rel,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
  heat_tbl_curr <- heat_tbl |>
    dplyr::filter(.data$bias_uns_setting == !!bias_uns_setting_curr)

  if (nrow(heat_tbl_curr) == 0L) {
    next
  }

  p_heat <- ggplot(
    heat_tbl_curr,
    aes(
      x = factor(bw_adaptive_core),
      y = factor(bw_adaptive_extra),
      fill = median_abs_rel_error
    )
  ) +
    geom_tile() +
    facet_grid(mean_pos_setting ~ transformation) +
    scale_fill_viridis_c(trans = "log10", option = "magma", direction = -1) +
    cowplot::theme_cowplot() +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA),
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      title = paste0(
        "Manual adaptive bandwidth grid: bias_uns_setting = ",
        bias_uns_setting_curr
      ),
      x = "Core bandwidth",
      y = "Extra bandwidth",
      fill = "Median absolute\nrelative error"
    )

  path_p_heat <- projr::projr_path_get(
    "output",
    "sim",
    "bw",
    "freq_bs",
    "adaptive",
    "heatmap",
    paste0("core_extra_heatmap_bias_uns_", bias_uns_setting_curr, ".png")
  )
  dir.create(dirname(path_p_heat), recursive = TRUE, showWarnings = FALSE)
  ggsave(path_p_heat, p_heat, width = 24, height = 18, units = "cm")
}
```

### Threshold distributions

```{r}
#| label: threshold-distributions
threshold_summary_tbl <- bw_tbl_results_raw |>
  dplyr::group_by(
    dplyr::pick(dplyr::any_of(colnames(sim_grid)))
  ) |>
  dplyr::summarise(
    threshold_median = stats::median(.data$threshold, na.rm = TRUE),
    threshold_iqr_lower = stats::quantile(.data$threshold, 0.25, na.rm = TRUE),
    threshold_iqr_upper = stats::quantile(.data$threshold, 0.75, na.rm = TRUE),
    threshold_sd = stats::sd(.data$threshold, na.rm = TRUE),
    threshold_mad = stats::mad(.data$threshold, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    threshold_iqr_length = .data$threshold_iqr_upper - .data$threshold_iqr_lower
  )

for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
  for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
    plot_tbl <- bw_tbl_results_raw |>
      dplyr::filter(
        .data$mean_pos_setting == !!mean_pos_setting_curr,
        .data$bias_uns_setting == !!bias_uns_setting_curr,
        is.finite(.data$threshold)
      ) |>
      dplyr::mutate(
        bw_pair = paste0(
          "core ", .data$bw_adaptive_core,
          ", extra ", .data$bw_adaptive_extra
        )
      )

    if (nrow(plot_tbl) == 0L) {
      next
    }

    p_thresh <- ggplot(
      plot_tbl,
      aes(x = bw_pair, y = threshold, colour = factor(bw_adaptive_core))
    ) +
      ggforce::geom_sina(alpha = 0.25, size = 0.6) +
      facet_grid(transformation ~ prob_response, scales = "free_y") +
      coord_flip() +
      cowplot::theme_cowplot() +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA),
        legend.position = "bottom",
        axis.text.y = element_text(size = 5)
      ) +
      labs(
        title = paste0(
          "Threshold distributions: mean_pos_setting = ",
          mean_pos_setting_curr,
          ", bias_uns_setting = ",
          bias_uns_setting_curr
        ),
        x = "Core and extra bandwidth",
        y = "Threshold",
        colour = "Core bandwidth"
      )

    path_p_thresh <- projr::projr_path_get(
      "output",
      "sim",
      "bw",
      "freq_bs",
      "adaptive",
      "thresholds",
      paste0(
        "thresholds_mean_pos_setting_",
        mean_pos_setting_curr,
        "_bias_uns_",
        bias_uns_setting_curr,
        ".png"
      )
    )
    dir.create(dirname(path_p_thresh), recursive = TRUE, showWarnings = FALSE)
    ggsave(path_p_thresh, p_thresh, width = 24, height = 30, units = "cm")
  }
}
```

### Summary table

```{r}
#| label: summary-table
method_rank_tbl <- bw_tbl_results_summary |>
  dplyr::group_by(
    .data$bias_uns_setting,
    .data$bias_uns,
    .data$transformation,
    .data$mean_pos_setting,
    .data$bw_adaptive_core,
    .data$bw_adaptive_extra
  ) |>
  dplyr::summarise(
    n_scenarios = dplyr::n(),
    median_abs_error = stats::median(.data$propBsEst_median_abs_diff, na.rm = TRUE),
    median_abs_rel_error = stats::median(.data$propBsEst_median_abs_diff_rel, na.rm = TRUE),
    median_threshold_sd = stats::median(.data$threshold_sd, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(
    .data$bias_uns_setting,
    .data$transformation,
    .data$mean_pos_setting,
    .data$median_abs_rel_error
  )

method_rank_tbl |>
  dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))) |>
  dplyr::group_by(
    .data$bias_uns_setting,
    .data$transformation,
    .data$mean_pos_setting
  ) |>
  dplyr::slice_head(n = 20) |>
  dplyr::ungroup() |>
  knitr::kable()
```
