---
title: Simulate adaptive bandwidth estimates
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
  sim_retry_errors: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
# On resume, rerun simulations whose saved output recorded an error.
sim_retry_errors <- .as_flag(.get_qmd_param_env(
  "sim_retry_errors",
  "SIM_RETRY_ERRORS",
  TRUE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "est", "adaptive")
analysis_semantics_version <- "adaptive-bw-est-v3"
analysis_dev <- isTRUE(.isDev())
n_sample_sim <- if (analysis_dev) 1L else 5L
n_iter_sim <- if (analysis_dev) 1L else 5L
norm_adaptive_ncell <- 2500L

run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source analysis helpers in dependency order after the generic QMD runtime
# helper has been loaded.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Checking what the adaptive estimators actually estimate

Get the parameters over which we simulate.

We mainly need to consider whether the number of parent-population cells is large or small, whether the responding population is large enough to affect the estimator, and the amount of separation.

We'll vary over a range of parent-population sizes, as that tells us how the estimator behaves over the most important dimension apart from the transformation.

The current adaptive normalised estimator calculates the core and extra bandwidths from fixed-size synthetic normal components. That internal sample size is controlled by `normAdaptiveNcell`, not by `bwNcellMax`. We therefore keep `normAdaptiveNcell = 2500` fixed here rather than varying the non-adaptive cell-count cap.

For the adaptive estimator, the output has separate bandwidths for the background-core and high-expression-extra components, and those are estimated separately for the stimulated and unstimulated samples. The collation therefore works with four bandwidth columns:

- `bw_uns_core`
- `bw_stim_core`
- `bw_uns_extra`
- `bw_stim_extra`

```{r}
#| label: actual-settings
# Simulation settings.
n_cell_stim_vec <- c(1e3, 5e3, 2e4, 1e5)
prob_response_vec <- c(0.002, 0.2)
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))

# Adaptive core/extra bandwidths are only produced for the normalised methods.
# Keep the hpi methods used in the norm QMD, but exclude hpi0 as there.
bandwidth_method_vec <- c("hpi1Norm", "hpi2Norm", "hpi3Norm")

bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.00125,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.005,
  "high", "skew", 0.25
) |>
  dplyr::mutate(
    bias_uns_setting = factor(
      .data$bias_uns_setting,
      levels = c("none", "low", "high")
    )
  ) |>
  dplyr::filter(
    .data$bias_uns_setting == "low"
  )

bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))
```

```{r}
#| label: bw-estimate-grid
data_scenario_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bias_uns = bias_uns_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bias_uns_setting, bias_uns
  ) |>
  dplyr::arrange(
    .data$n_cell,
    dplyr::desc(.data$prob_response),
    .data$transformation,
    .data$mean_pos_setting
  ) |>
  dplyr::mutate(
    data_scenario_id = dplyr::row_number(),
    sim_seed = as.integer(simulation_seed + .data$data_scenario_id - 1L)
  )

# Candidate bandwidth methods are compared on the same simulated data.
# sim_seed therefore belongs to the data-generating scenario, not the method.
sim_grid <- tidyr::crossing(
  data_scenario_grid,
  bw_mtd = bandwidth_method_vec
) |>
  dplyr::arrange(.data$data_scenario_id, .data$bw_mtd) |>
  dplyr::mutate(sim_id = dplyr::row_number())

# sim_id and sim_seed are fixed on the full grid, before the dev filter,
# shuffling and chunking, so each row keeps its seed in every run.
sim_grid_full <- sim_grid

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      .data$n_cell %in% c(5e3, 2e4),
      .data$prob_response == 0.002,
      .data$mean_pos_setting == "high",
      grepl("hpi1", .data$bw_mtd)
    )
}

# Keep the complete selected grid for validation and collation.
sim_grid_all <- sim_grid
if (nrow(sim_grid_all) == 0L) {
  stop("The analysis 5 simulation grid is empty; check the dev filter.")
}
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

# Randomise execution order, then distribute rows evenly across external chunks.
sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
  dplyr::select(-".sim_grid_chunk")

sim_grid
print(paste0(
  "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
  " simulations for sim_grid chunk ", sim_grid_chunk_index,
  " of ", sim_grid_n_chunks,
  " using shuffle seed ", sim_grid_shuffle_seed
))
```


Fixed settings passed to `.simBandwidthEstBwDirectAdaptive()` for every grid
row, and the manifest parameters that canonical results must match.

```{r}
#| label: bw-estimate-settings
scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1L,
  nCondition = 2L,
  nCluster = 2L,
  nIter = n_iter_sim,
  bwMin = -Inf,
  bwMax = Inf,
  bwFallback = NA_real_,
  bwCluster = 0.5,
  normAdaptiveNcell = norm_adaptive_ncell,
  probExact = TRUE,
  backgroundRelativeToResponse = 0.2,
  ncellUnsRelativeToStim = 1,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = NULL,
  summarise = FALSE
)

analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_dev = analysis_dev,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

Rerun one simulation interactively, for example to see why it failed.
Each row is seeded from its own `sim_seed` with a fixed RNG kind inside
`.simBandwidthRunRow()`, which the parallel run also calls, so this
reproduces the random numbers of that row exactly, whatever the chunking,
shuffling or worker scheduling. Call `debugonce()` on any helper first to
step through it.

```{r}
#| label: rerun-one-simulation
#| eval: false
sim_id_target <- 1L

row_target <- sim_grid_full |>
  dplyr::filter(.data$sim_id == sim_id_target)
rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthEstAdaptiveScenario,
  settings = scenario_settings
)

# Optionally compare with the stored output for this sim_id. Point `dir_run`
# at a staged run directory instead to check a run that was not promoted.
dir_run <- .analysis_cache_dir(
  c(analysis_key, "current"),
  path_root = root_dir,
  create = FALSE
)
path_stored <- .simBandwidthFindSimOutput(dir_run, sim_id_target)
if (length(path_stored) == 1L) {
  all.equal(rerun_res, readRDS(path_stored))
}
```

Run the simulations in parallel, with progress reporting, then validate this
chunk and promote the run once every chunk is complete.

```{r}
#| label: bw-estimate-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks,
        sim_grid_shuffle_seed = sim_grid_shuffle_seed
      ),
      analysis_required_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthEstAdaptiveScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = sim_retry_errors,
    heading = "ADAPTIVE BANDWIDTH ESTIMATE PROGRESS",
    path_root = root_dir
  )

  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    validate_fn = function(tbl) {
      .simBandwidthEstAdaptiveValidate(tbl, n_sample_sim * n_iter_sim)
    },
    collate_fn = function(tbl) {
      .simBandwidthEstAdaptiveCollate(tbl, colnames(sim_grid_all))
    },
    label = "analysis 5"
  )

  if (isTRUE(run_plots)) {
    message(
      "Skipping plots during a simulation render. ",
      "Render again with run_simulations = FALSE and run_plots = TRUE."
    )
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  # Read-only: plots use the canonical promoted results under current/.
  run_ctx <- .analysis_results_context(analysis_key, path_root = root_dir)
}
```

Read the canonical promoted results for plotting.

```{r}
#| label: bw-estimate-collate
if (isTRUE(run_plots)) {
  bw_tbl_results <- readRDS(.analysis_current_file(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results.rds"),
    required_params = analysis_required_params
  ))

  bw_tbl_results |>
    dplyr::group_by(.data$bw_mtd, .data$bw_component, .data$bw_condition) |>
    dplyr::summarise(
      n_grid_rows = dplyr::n(),
      n_grid_rows_with_estimates = sum(.data$n_est > 0L),
      mean_prop_est = mean(.data$prop_est),
      min_prop_est = min(.data$prop_est),
      .groups = "drop"
    ) |>
    print(width = Inf)

  if (any(bw_tbl_results$prop_est < 1)) {
    message(
      "Some adaptive bandwidths could not be estimated. ",
      "See prop_est in bw_tbl_results; means are conditional on finite ",
      "estimates."
    )
  }
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

```{r}
#| label: bw-estimate-plots
#| fig-width: 7
#| fig-height: 10
if (isTRUE(run_plots)) {
  dir_results <- .analysis_cache_dir(c(analysis_key, "plots"), root_dir)

  for (mean_pos_setting_curr in unique(bw_tbl_results$mean_pos_setting)) {
    bw_tbl_mean_pos <- bw_tbl_results |>
      dplyr::filter(.data$mean_pos_setting == !!mean_pos_setting_curr)

    if (nrow(bw_tbl_mean_pos) == 0L) {
      next
    }

    for (bias_uns_setting_curr in unique(bw_tbl_mean_pos$bias_uns_setting)) {
      bw_tbl_bias <- bw_tbl_mean_pos |>
        dplyr::filter(.data$bias_uns_setting == !!bias_uns_setting_curr)

      if (nrow(bw_tbl_bias) == 0L) {
        next
      }

      for (bw_component_curr in levels(bw_tbl_bias$bw_component)) {
        bw_tbl_comp <- bw_tbl_bias |>
          dplyr::filter(.data$bw_component == !!bw_component_curr)

        if (nrow(bw_tbl_comp) == 0L) {
          next
        }

        for (bw_condition_curr in levels(bw_tbl_comp$bw_condition)) {
          plot_tbl <- bw_tbl_comp |>
            dplyr::filter(.data$bw_condition == !!bw_condition_curr)

          if (nrow(plot_tbl) == 0L) {
            next
          }

          p <- ggplot(
            plot_tbl,
            aes(
              x = .data$n_cell,
              y = .data$mean_bw,
              colour = .data$bw_mtd_base
            )
          ) +
            geom_line(linetype = "solid") +
            geom_point() +
            scale_x_log10() +
            facet_grid(transformation ~ prob_response, scales = "free_y") +
            labs(
              title = paste0(
                "Mean position setting: ", mean_pos_setting_curr,
                ",\nbias_uns_setting: ", bias_uns_setting_curr,
                ",\ncomponent: ", bw_component_curr,
                ",\ncondition: ", bw_condition_curr
              ),
              x = "Number of cells",
              y = "Estimated bandwidth",
              colour = "Method"
            ) +
            cowplot::theme_cowplot() +
            cowplot::background_grid(major = "xy") +
            theme(
              panel.background = element_rect(fill = "white", colour = NA),
              plot.background = element_rect(fill = "white", colour = NA),
              legend.position = "bottom"
            ) +
            scale_colour_manual(values = .simBwMtdColVec, drop = FALSE)

          path_plot <- file.path(
            dir_results,
            paste0(
              "bw_estimate_meanpos_setting_",
              .file_safe(mean_pos_setting_curr),
              "_bias_uns_",
              .file_safe(bias_uns_setting_curr),
              "_component_",
              .file_safe(bw_component_curr),
              "_condition_",
              .file_safe(bw_condition_curr),
              ".png"
            )
          )
          dir.create(dirname(path_plot), recursive = TRUE, showWarnings = FALSE)
          ggsave(
            filename = path_plot,
            plot = p,
            width = 12,
            height = 20,
            units = "cm"
          )
          print(p)
        }
      }
    }
  }
}
```
