---
title: Manual adaptive bandwidth performance simulation
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  sim_grid_shuffle_seed: 20260707
  run_simulations: true
  run_plots: false
---

This analysis asks how the core and extra bandwidths of an adaptive density affect the estimated background-subtracted response frequency. Matched unstimulated and stimulated samples are simulated with Gaussian, gamma and skewed expression distributions. The analysis varies both bandwidths, cell counts, response frequencies and unstimulated-sample bias, then compares final sample estimates with the simulated truth.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
analysis_qmd <- "analysis/6-sim-bw-freq_bs-adaptive.qmd"

sim_grid_chunk_index <- as.integer(.get_qmd_param_env(
  "sim_grid_chunk_index",
  "SIM_GRID_CHUNK_INDEX",
  1L
))
sim_grid_n_chunks <- as.integer(.get_qmd_param_env(
  "sim_grid_n_chunks",
  "SIM_GRID_N_CHUNKS",
  1L
))
sim_grid_shuffle_seed <- as.integer(.get_qmd_param_env(
  "sim_grid_shuffle_seed",
  "SIM_GRID_SHUFFLE_SEED",
  20260707L
))
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "bw", "freq_bs", "adaptive")
analysis_semantics_version <- "adaptive-bw-freq-v3"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("6-sim-bw-freq_bs-adaptive")
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
# Source analysis helpers in dependency order after the generic QMD runtime
# helper has been loaded.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-run.R",
  "sim-bandwidth-analysis-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

The core bandwidth describes the background density; the extra bandwidth describes the high-expression component. Cluster refinement and cytokine-positive refinement are disabled to isolate the bandwidth settings.

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# Adaptive bandwidth settings. Gamma uses the low end of this range, while
# gaussian and skew use the wider-bandwidth part of the same vector.
bandwidth_value_vec <- c(
  # gamma
  0.001, 0.0025, 0.005, 0.0075, 0.01,
  0.0125, 0.015, 0.0175, 0.02, 0.03,
  0.04,
  # mixed
  0.05, 0.1, 0.15,
  # skew and gaussian
  0.2, 0.25, 0.5, 0.75, 1
)
bw_core_vec <- bandwidth_value_vec
bw_extra_vec <- bandwidth_value_vec

# NULL/NA crossover means use the density-weighted component blend. Finite
# values force a core-to-extra crossover at that expression value.
bw_crossover_vec <- c(NA_real_, 4.5, 5.5, 6.5)[[1]]
bw_transition_width_vec <- c(0, 0.25)[[1]]

# StimGate method settings for bandwidth-performance analysis:
# Cluster refinement (tolClust) is intentionally set to NULL (OFF) to isolate and
# measure the effect of the bandwidth selector itself without multi-sample
# cluster adjustments modifying thresholds post-local-FDR.
tol_clust <- NULL
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

bias_uns_settings_tbl <- tibble::tribble(
  ~bias_uns_setting, ~transformation, ~bias_uns,
  "none", "gaussian", 0,
  "none", "gamma", 0,
  "none", "skew", 0,
  "low", "gaussian", 0.05,
  "low", "gamma", 0.0025,
  "low", "skew", 0.05,
  "high", "gaussian", 0.25,
  "high", "gamma", 0.01,
  "high", "skew", 0.25
)
bias_uns_vec <- sort(unique(bias_uns_settings_tbl$bias_uns))

n_sample_sim <- 5L
n_iter_sim <- 5L

```

### Varying the adaptive core and extra bandwidths manually

The adaptive density has two manually fixed bandwidths: one for the background core and one for the high-expression extra component. The grid therefore contains every valid `bw_core` and `bw_extra` combination for the corresponding transformation family. Gamma is kept on the low-bandwidth range, while gaussian and skew are kept on the wider range.

```{r}
#| label: bw-manual-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = sort(unique(c(prob_response_vec, 0.002, 0.02))),
  n_cell = sort(unique(c(n_cell_stim_vec, 1e3, 5e4))),
  bw_core = bw_core_vec,
  bw_extra = bw_extra_vec,
  bias_uns = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::inner_join(
    bias_uns_settings_tbl,
    by = c("transformation", "bias_uns")
  ) |>
  dplyr::mutate(
    bw_fallback = dplyr::if_else(.data$transformation == "gamma", 0.01, 0.5),
    bw_crossover = bw_crossover_vec,
    bw_transition_width = bw_transition_width_vec
  ) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos,
    bw_core, bw_extra, bw_fallback,
    bw_crossover, bw_transition_width,
    bias_uns_setting, bias_uns,
    sample_perturbation_sd, condition_perturbation_sd,
    cluster_perturbation_sd, background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::filter(
    (
      .data$transformation == "gamma" &
        .data$bw_core <= 0.15 &
        .data$bw_extra <= 0.15
    ) |
      (
        .data$transformation != "gamma" &
          .data$bw_core >= 0.05 &
          .data$bw_extra >= 0.05
      )
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number(),
    sim_seed = as.integer(simulation_seed + sim_id - 1L)
  )

# Fix identities on the full grid before selecting execution subsets.
sim_grid_full <- sim_grid

if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(1e3, 5e4),
      prob_response %in% c(0.002, 0.02),
      bw_core %in% c(0.01, 0.05, 0.15, 0.5),
      bw_extra %in% c(0.01, 0.05, 0.15, 0.5),
      condition_perturbation_sd == 0
    )
} else {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% n_cell_stim_vec,
      prob_response %in% prob_response_vec
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      bw_core == 0.15,
      bw_extra == 0.15,
      bias_uns_setting == "low",
      condition_perturbation_sd == 0
    ) |>
    dplyr::slice_head(n = 5)
}

if (analysis_quick && nrow(sim_grid) > 100L) {
  sim_grid <- withr::with_seed(
    42L,
    dplyr::slice_sample(sim_grid, n = 100L),
    .rng_kind = .simBandwidthRngKind[["kind"]],
    .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
    .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
  )
}

sim_grid_all <- sim_grid

if (nrow(sim_grid_all) == 0L) {
  stop("The analysis 6 simulation grid is empty; check dev/quick filters.")
}
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-sim_seed)

sim_grid <- withr::with_seed(
  sim_grid_shuffle_seed,
  dplyr::slice_sample(sim_grid_all, prop = 1),
  .rng_kind = .simBandwidthRngKind[["kind"]],
  .rng_normal_kind = .simBandwidthRngKind[["normal.kind"]],
  .rng_sample_kind = .simBandwidthRngKind[["sample.kind"]]
) |>
  dplyr::mutate(
    .sim_grid_chunk =
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L
  ) |>
  dplyr::filter(.data$.sim_grid_chunk == sim_grid_chunk_index) |>
  dplyr::arrange(.data$n_cell, dplyr::desc(.data$prob_response)) |>
  dplyr::select(-.sim_grid_chunk)

sim_grid
print(paste0(
  "Performing ", nrow(sim_grid), " of ", nrow(sim_grid_all),
  " simulations for sim_grid chunk ", sim_grid_chunk_index,
  " of ", sim_grid_n_chunks,
  " using shuffle seed ", sim_grid_shuffle_seed
))
```

## Simulation

Fixed settings are shared by the parallel worker and interactive rerun.
The manifest includes the complete selected cross-chunk scientific grid.

```{r}
#| label: bw-manual-settings
#| include: false
scenario_settings <- list(
  nSample = n_sample_sim,
  nMarker = 1L,
  nCondition = 2L,
  nCluster = 2L,
  nIter = n_iter_sim,
  bw = NULL,
  bwAdaptive = TRUE,
  bwMin = "none",
  bwMax = "none",
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = tol_clust,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates
)
analysis_required_params <- list(
  analysis_semantics_version = analysis_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  scenario_settings = scenario_settings,
  sim_grid_spec = analysis_grid_spec
)
```

Rerun one full-grid simulation with the same settings and explicit RNG kinds
used by the parallel runner, regardless of chunk or worker scheduling.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 1L
row_target <- sim_grid_full[sim_grid_full$sim_id == sim_id_target, ]
rerun_res <- .simBandwidthRunRow(
  row_target,
  scenario_fn = .simBandwidthFreqBsAdaptiveScenario,
  settings = scenario_settings
)
dir_run <- .analysis_cache_dir(
  c(analysis_key, "current"), path_root = root_dir, create = FALSE
)
path_stored <- .simBandwidthFindSimOutput(dir_run, sim_id_target)
if (length(path_stored) == 1L) {
  all.equal(rerun_res, readRDS(path_stored))
}
```

Run this chunk, retry previously failed scenarios, and promote only after
all chunks and final sample results validate.

```{r}
#| label: bw-manual-parallel
#| include: false
if (interactive()) {
  progressr::handlers(global = TRUE)
}
if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(list(
      run_simulations = run_simulations,
      run_plots = run_plots,
      sim_grid_chunk_index = sim_grid_chunk_index,
      sim_grid_n_chunks = sim_grid_n_chunks,
      sim_grid_shuffle_seed = sim_grid_shuffle_seed
    ), analysis_required_params),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )
  .simBandwidthRunGrid(
    sim_grid,
    scenario_fn = .simBandwidthFreqBsAdaptiveScenario,
    settings = scenario_settings,
    run_ctx = run_ctx,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks,
    retry_errors = TRUE,
    heading = "ADAPTIVE BANDWIDTH FREQUENCY PROGRESS",
    path_root = root_dir
  )
  .simBandwidthFinishChunk(
    run_ctx = run_ctx,
    sim_grid = sim_grid,
    sim_grid_all = sim_grid_all,
    collate_fn = function(tbl) {
      .simBandwidthFreqBsAdaptiveCollate(tbl, colnames(sim_grid_all))
    },
    validate_fn = .simBandwidthFreqBsAdaptiveValidate,
    label = "analysis 6"
  )
  if (isTRUE(run_plots) && sim_grid_n_chunks > 1L) {
    message("Skipping plots during a multi-chunk simulation render.")
    run_plots <- FALSE
  }
} else if (isTRUE(run_plots)) {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}
```

## Collation

Only complete, validated results are promoted to the canonical cache. Plot-only renders read that cache with the same scientific settings and dev/quick profile.

```{r}
#| label: bw-manual-collate
#| include: false
if (isTRUE(run_plots)) {
  bw_tbl_results_raw <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results_raw.rds"),
    required_params = analysis_required_params
  )
  bw_tbl_results_summary <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "bw_tbl_results_summary.rds"),
    required_params = analysis_required_params
  )
  summary_tbl <- .analysis_read_current(
    run_ctx = run_ctx,
    relative_path = c("collated", "summary_tbl.rds"),
    required_params = analysis_required_params
  )
} else {
  message("run_plots is false, so stopping after simulation/collation.")
}
```

```{r}
#| label: plot-helpers
#| include: false
if (isTRUE(run_plots)) {
  bias_uns_setting_vec_plot <- bias_uns_settings_tbl |>
    dplyr::distinct(.data$bias_uns_setting) |>
    dplyr::pull(.data$bias_uns_setting)
}
```

## Results

### Estimated frequency: fixed core bandwidth, varying extra bandwidth

Cell count is on the logarithmic horizontal axis and median estimated frequency is on the vertical axis. Colour distinguishes extra bandwidths, panels separate response frequencies, and dashed lines mark the simulated truth; estimates nearer these lines indicate closer agreement.

```{r}
#| label: est-fixed-core
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_summary$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_core_curr in unique(bw_tbl_results_summary$bw_core)) {
          plot_tbl <- bw_tbl_results_summary |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_core == !!bw_core_curr
            ) |>
            add_bw_labs()

          if (nrow(plot_tbl) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Core bandwidth: ", format_bw_lab(bw_core_curr)),
            6
          )

          col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw_extra)
          plot_tbl <- plot_tbl |>
            dplyr::mutate(
              bw_extra_lab = factor(.data$bw_extra_lab, levels = names(col_vec_bw_curr))
            )

          p <- ggplot(
            plot_tbl,
            aes(x = n_cell, y = propRespEst_median, colour = bw_extra_lab)
          ) +
            geom_line(alpha = 0.75) +
            geom_hline(
              mapping = aes(yintercept = propRespTruth_median),
              linetype = "dashed",
              colour = "black"
            ) +
            geom_point(alpha = 0.75) +
            scale_x_log10(
              labels = .analysis_label_number,
              guide = guide_axis(angle = 45)
            ) +
            scale_y_continuous(labels = .analysis_label_percent) +
            facet_grid(
              . ~ prob_response,
              scales = "free_y",
              labeller = labeller(prob_response = .analysis_labeller_percent())
            ) +
            .analysis_theme() +
            labs(
              x = "Number of cells (log scale)",
              y = "Estimated background-subtracted frequency (median)",
              colour = "Extra bandwidth"
            ) +
            scale_colour_manual(values = col_vec_bw_curr)

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "est",
            "median",
            "fixed_core_vary_extra",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "est_vs_n_cell_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_core_",
              format_bw_file(bw_core_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 8)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Estimated frequency: fixed extra bandwidth, varying core bandwidth

Cell count is on the logarithmic horizontal axis and median estimated frequency is on the vertical axis. Colour distinguishes core bandwidths and panels separate response frequencies. Compare estimates with the dashed lines marking the simulated truth.

```{r}
#| label: est-fixed-extra
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_summary$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_extra_curr in unique(bw_tbl_results_summary$bw_extra)) {
          plot_tbl <- bw_tbl_results_summary |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_extra == !!bw_extra_curr
            ) |>
            add_bw_labs()

          if (nrow(plot_tbl) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Extra bandwidth: ", format_bw_lab(bw_extra_curr)),
            6
          )

          col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw_core)
          plot_tbl <- plot_tbl |>
            dplyr::mutate(
              bw_core_lab = factor(.data$bw_core_lab, levels = names(col_vec_bw_curr))
            )

          p <- ggplot(
            plot_tbl,
            aes(x = n_cell, y = propRespEst_median, colour = bw_core_lab)
          ) +
            geom_line(alpha = 0.75) +
            geom_hline(
              mapping = aes(yintercept = propRespTruth_median),
              linetype = "dashed",
              colour = "black"
            ) +
            geom_point(alpha = 0.75) +
            scale_x_log10(
              labels = .analysis_label_number,
              guide = guide_axis(angle = 45)
            ) +
            scale_y_continuous(labels = .analysis_label_percent) +
            facet_grid(
              . ~ prob_response,
              scales = "free_y",
              labeller = labeller(prob_response = .analysis_labeller_percent())
            ) +
            .analysis_theme() +
            labs(
              x = "Number of cells (log scale)",
              y = "Estimated background-subtracted frequency (median)",
              colour = "Core bandwidth"
            ) +
            scale_colour_manual(values = col_vec_bw_curr)

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "est",
            "median",
            "fixed_extra_vary_core",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "est_vs_n_cell_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_extra_",
              format_bw_file(bw_extra_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 8)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Raw frequency estimates: fixed core bandwidth, varying extra bandwidth

Each point is a final sample estimate, with cell count on the horizontal axis and estimated frequency on the vertical axis. Colour distinguishes extra bandwidths and panels separate response frequencies. The spread around the dashed truth lines shows variation across samples and simulation iterations.

```{r}
#| label: raw-fixed-core
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_raw$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_core_curr in unique(bw_tbl_results_raw$bw_core)) {
          plot_tbl <- bw_tbl_results_raw |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_core == !!bw_core_curr
            ) |>
            add_bw_labs()

          if (nrow(plot_tbl) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Core bandwidth: ", format_bw_lab(bw_core_curr)),
            6
          )

          col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw_extra)
          plot_tbl <- plot_tbl |>
            dplyr::mutate(
              bw_extra_lab = factor(.data$bw_extra_lab, levels = names(col_vec_bw_curr))
            )

          p <- ggplot(
            plot_tbl,
            aes(x = factor(n_cell), y = propRespEst, colour = bw_extra_lab)
          ) +
            ggforce::geom_sina(position = "dodge", alpha = 0.75) +
            geom_hline(
              mapping = aes(yintercept = propRespTruth),
              linetype = "dashed",
              colour = "black"
            ) +
            scale_x_discrete(
              labels = function(x) .analysis_label_number(as.numeric(x)),
              guide = guide_axis(angle = 45)
            ) +
            scale_y_continuous(labels = .analysis_label_percent) +
            facet_grid(
              . ~ prob_response,
              scales = "free_y",
              labeller = labeller(prob_response = .analysis_labeller_percent())
            ) +
            .analysis_theme() +
            labs(
              x = "Number of cells",
              y = "Estimated background-subtracted frequency",
              colour = "Extra bandwidth"
            ) +
            scale_colour_manual(values = col_vec_bw_curr)

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "est",
            "raw",
            "fixed_core_vary_extra",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "actual_freq_vs_n_cell_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_core_",
              format_bw_file(bw_core_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 8)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Raw frequency estimates: fixed extra bandwidth, varying core bandwidth

Each point is a final sample estimate, with cell count on the horizontal axis and estimated frequency on the vertical axis. Colour distinguishes core bandwidths and panels separate response frequencies. Compare the spread of points with the dashed lines marking the simulated truth.

```{r}
#| label: raw-fixed-extra
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_raw$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_extra_curr in unique(bw_tbl_results_raw$bw_extra)) {
          plot_tbl <- bw_tbl_results_raw |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_extra == !!bw_extra_curr
            ) |>
            add_bw_labs()

          if (nrow(plot_tbl) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Extra bandwidth: ", format_bw_lab(bw_extra_curr)),
            6
          )

          col_vec_bw_curr <- make_bw_colour_values(plot_tbl$bw_core)
          plot_tbl <- plot_tbl |>
            dplyr::mutate(
              bw_core_lab = factor(.data$bw_core_lab, levels = names(col_vec_bw_curr))
            )

          p <- ggplot(
            plot_tbl,
            aes(x = factor(n_cell), y = propRespEst, colour = bw_core_lab)
          ) +
            ggforce::geom_sina(position = "dodge", alpha = 0.75) +
            geom_hline(
              mapping = aes(yintercept = propRespTruth),
              linetype = "dashed",
              colour = "black"
            ) +
            scale_x_discrete(
              labels = function(x) .analysis_label_number(as.numeric(x)),
              guide = guide_axis(angle = 45)
            ) +
            scale_y_continuous(labels = .analysis_label_percent) +
            facet_grid(
              . ~ prob_response,
              scales = "free_y",
              labeller = labeller(prob_response = .analysis_labeller_percent())
            ) +
            .analysis_theme() +
            labs(
              x = "Number of cells",
              y = "Estimated background-subtracted frequency",
              colour = "Core bandwidth"
            ) +
            scale_colour_manual(values = col_vec_bw_curr)

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "est",
            "raw",
            "fixed_extra_vary_core",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "actual_freq_vs_n_cell_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_extra_",
              format_bw_file(bw_extra_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 8)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Threshold plots

Threshold plots show final thresholds as vertical lines. Median plots
vary one bandwidth component at a time; raw plots also show sample thresholds
and their interquartile range.

### Median thresholds: fixed core bandwidth, varying extra bandwidth

Vertical lines mark median thresholds on the marker-expression axis; line type distinguishes extra bandwidths. Rows separate response frequencies and columns separate cell counts. Compare the positions of the lines to see how the extra bandwidth changes the gate.

```{r}
#| label: threshold-median-fixed-core
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_summary$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_core_curr in unique(bw_tbl_results_summary$bw_core)) {
          threshold_tbl_curr <- bw_tbl_results_summary |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_core == !!bw_core_curr
            ) |>
            add_bw_labs()

          if (nrow(threshold_tbl_curr) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Core bandwidth: ", format_bw_lab(bw_core_curr)),
            6
          )

          bw_scale <- make_bw_linetype_scale(threshold_tbl_curr$bw_extra)
          threshold_tbl_curr <- threshold_tbl_curr |>
            dplyr::mutate(
              bw_extra_lab = factor(.data$bw_extra_lab, levels = bw_scale$levels)
            )

          p <- ggplot() +
            facet_grid(
              rows = vars(.data$prob_response),
              cols = vars(.data$n_cell),
              scales = "free_x",
              labeller = labeller(
                prob_response = .analysis_labeller_percent(),
                n_cell = function(x) .analysis_label_number(as.numeric(x))
              )
            ) +
            .analysis_theme() +
            theme(strip.text.y = element_text(angle = 0)) +
            labs(
              x = "Marker expression",
              y = NULL,
              colour = NULL,
              fill = NULL,
              linetype = "Extra bandwidth"
            ) +
            geom_vline(
              data = threshold_tbl_curr,
              mapping = aes(
                xintercept = threshold_median, linetype = bw_extra_lab,
                group = bw_extra_lab
              ),
              linewidth = 0.45
            ) +
            scale_linetype_manual(
              values = bw_scale$values,
              breaks = bw_scale$levels,
              labels = bw_scale$levels
            )

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "thresholds",
            "median",
            "fixed_core_vary_extra",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "threshold_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_core_",
              format_bw_file(bw_core_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 22)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Median thresholds: fixed extra bandwidth, varying core bandwidth

Vertical lines mark median thresholds on the marker-expression axis; line type distinguishes core bandwidths. Rows separate response frequencies and columns separate cell counts. Compare the positions of the lines to see how the core bandwidth changes the gate.

```{r}
#| label: threshold-median-fixed-extra
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_summary$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_summary$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_extra_curr in unique(bw_tbl_results_summary$bw_extra)) {
          threshold_tbl_curr <- bw_tbl_results_summary |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_extra == !!bw_extra_curr
            ) |>
            add_bw_labs()

          if (nrow(threshold_tbl_curr) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Extra bandwidth: ", format_bw_lab(bw_extra_curr)),
            6
          )

          bw_scale <- make_bw_linetype_scale(threshold_tbl_curr$bw_core)
          threshold_tbl_curr <- threshold_tbl_curr |>
            dplyr::mutate(
              bw_core_lab = factor(.data$bw_core_lab, levels = bw_scale$levels)
            )

          p <- ggplot() +
            facet_grid(
              rows = vars(.data$prob_response),
              cols = vars(.data$n_cell),
              scales = "free_x",
              labeller = labeller(
                prob_response = .analysis_labeller_percent(),
                n_cell = function(x) .analysis_label_number(as.numeric(x))
              )
            ) +
            .analysis_theme() +
            theme(strip.text.y = element_text(angle = 0)) +
            labs(
              x = "Marker expression",
              y = NULL,
              colour = NULL,
              fill = NULL,
              linetype = "Core bandwidth"
            ) +
            geom_vline(
              data = threshold_tbl_curr,
              mapping = aes(
                xintercept = threshold_median, linetype = bw_core_lab,
                group = bw_core_lab
              ),
              linewidth = 0.45
            ) +
            scale_linetype_manual(
              values = bw_scale$values,
              breaks = bw_scale$levels,
              labels = bw_scale$levels
            )

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "thresholds",
            "median",
            "fixed_extra_vary_core",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "threshold_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_extra_",
              format_bw_file(bw_extra_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 22)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Raw thresholds: fixed core bandwidth, varying extra bandwidth

Faint vertical lines mark individual sample thresholds on the marker-expression axis, coloured by extra bandwidth. Solid lines mark medians and dashed lines mark the lower and upper quartiles. Rows separate response frequencies and columns separate cell counts, allowing the threshold spread to be compared within each setting.

```{r}
#| label: threshold-raw-fixed-core
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_raw$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_core_curr in unique(bw_tbl_results_raw$bw_core)) {
          threshold_tbl_curr <- bw_tbl_results_raw |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_core == !!bw_core_curr
            ) |>
            add_bw_labs()

          if (nrow(threshold_tbl_curr) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Core bandwidth: ", format_bw_lab(bw_core_curr)),
            6
          )

          col_vec_bw_curr <- make_bw_colour_values(threshold_tbl_curr$bw_extra)
          threshold_tbl_curr <- threshold_tbl_curr |>
            dplyr::mutate(
              bw_extra_lab = factor(.data$bw_extra_lab, levels = names(col_vec_bw_curr))
            )

          threshold_tbl_curr_med_iqr <- threshold_tbl_curr |>
            dplyr::group_by(
              dplyr::pick(dplyr::any_of(colnames(sim_grid_all)))
            ) |>
            dplyr::summarise(
              threshold_median = stats::median(threshold, na.rm = TRUE),
              threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
              threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
              .groups = "drop"
            ) |>
            tidyr::pivot_longer(
              threshold_median:threshold_iqr_upper,
              names_to = "threshold_stat",
              values_to = "threshold"
            ) |>
            dplyr::mutate(
              threshold_stat = gsub("threshold_", "", .data$threshold_stat),
              threshold_stat = factor(
                .data$threshold_stat,
                levels = c("iqr_lower", "median", "iqr_upper")
              ),
              bw_extra_lab = factor(format_bw_lab(.data$bw_extra), levels = names(col_vec_bw_curr))
            )

          p <- ggplot() +
            facet_grid(
              rows = vars(.data$prob_response),
              cols = vars(.data$n_cell),
              scales = "free_x",
              labeller = labeller(
                prob_response = .analysis_labeller_percent(),
                n_cell = function(x) .analysis_label_number(as.numeric(x))
              )
            ) +
            .analysis_theme() +
            theme(strip.text.y = element_text(angle = 0)) +
            scale_colour_manual(
              values = col_vec_bw_curr
            ) +
            labs(
              x = "Marker expression",
              y = NULL,
              colour = "Extra bandwidth",
              fill = NULL,
              linetype = "Threshold summary"
            ) +
            geom_vline(
              data = threshold_tbl_curr,
              mapping = aes(
                xintercept = threshold,
                colour = bw_extra_lab,
                group = interaction(bw_extra_lab, iter, sample)
              ),
              linewidth = 0.25,
              alpha = 0.22
            ) +
            geom_vline(
              data = threshold_tbl_curr_med_iqr,
              aes(
                xintercept = threshold,
                colour = bw_extra_lab,
                linetype = threshold_stat,
                group = interaction(bw_extra_lab, threshold_stat)
              ),
              linewidth = 0.45
            ) +
            scale_linetype_manual(
              values = c(
                "iqr_lower" = "dashed",
                "median" = "solid",
                "iqr_upper" = "dashed"
              ),
              breaks = c("iqr_lower", "median", "iqr_upper"),
              labels = c("IQR lower", "Median", "IQR upper")
            )

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "thresholds",
            "raw",
            "fixed_core_vary_extra",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "threshold_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_core_",
              format_bw_file(bw_core_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 22)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Raw thresholds: fixed extra bandwidth, varying core bandwidth

Faint vertical lines mark individual sample thresholds on the marker-expression axis, coloured by core bandwidth. Solid lines mark medians and dashed lines mark the lower and upper quartiles. Rows separate response frequencies and columns separate cell counts, allowing the threshold spread to be compared within each setting.

```{r}
#| label: threshold-raw-fixed-extra
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (transformation_curr in .analysis_trans_order(bw_tbl_results_raw$transformation)) {
    .analysis_heading(
      paste0(
        "Transformation: ",
        as.character(.analysis_trans_factor(transformation_curr))
      ),
      4
    )
    for (mean_pos_setting_curr in unique(bw_tbl_results_raw$mean_pos_setting)) {
      for (bias_uns_setting_curr in bias_uns_setting_vec_plot) {
        .analysis_heading(
          paste0(
            "Mean position: ", mean_pos_setting_curr,
            "; bias: ", bias_uns_setting_curr
          ),
          5
        )
        for (bw_extra_curr in unique(bw_tbl_results_raw$bw_extra)) {
          threshold_tbl_curr <- bw_tbl_results_raw |>
            dplyr::filter(
              .data$transformation == !!transformation_curr,
              .data$mean_pos_setting == !!mean_pos_setting_curr,
              .data$bias_uns_setting == !!bias_uns_setting_curr,
              .data$bw_extra == !!bw_extra_curr
            ) |>
            add_bw_labs()

          if (nrow(threshold_tbl_curr) == 0L) {
            next
          }

          .analysis_heading(
            paste0("Extra bandwidth: ", format_bw_lab(bw_extra_curr)),
            6
          )

          col_vec_bw_curr <- make_bw_colour_values(threshold_tbl_curr$bw_core)
          threshold_tbl_curr <- threshold_tbl_curr |>
            dplyr::mutate(
              bw_core_lab = factor(.data$bw_core_lab, levels = names(col_vec_bw_curr))
            )

          threshold_tbl_curr_med_iqr <- threshold_tbl_curr |>
            dplyr::group_by(
              dplyr::pick(dplyr::any_of(colnames(sim_grid_all)))
            ) |>
            dplyr::summarise(
              threshold_median = stats::median(threshold, na.rm = TRUE),
              threshold_iqr_lower = stats::quantile(threshold, 0.25, na.rm = TRUE),
              threshold_iqr_upper = stats::quantile(threshold, 0.75, na.rm = TRUE),
              .groups = "drop"
            ) |>
            tidyr::pivot_longer(
              threshold_median:threshold_iqr_upper,
              names_to = "threshold_stat",
              values_to = "threshold"
            ) |>
            dplyr::mutate(
              threshold_stat = gsub("threshold_", "", .data$threshold_stat),
              threshold_stat = factor(
                .data$threshold_stat,
                levels = c("iqr_lower", "median", "iqr_upper")
              ),
              bw_core_lab = factor(format_bw_lab(.data$bw_core), levels = names(col_vec_bw_curr))
            )

          p <- ggplot() +
            facet_grid(
              rows = vars(.data$prob_response),
              cols = vars(.data$n_cell),
              scales = "free_x",
              labeller = labeller(
                prob_response = .analysis_labeller_percent(),
                n_cell = function(x) .analysis_label_number(as.numeric(x))
              )
            ) +
            .analysis_theme() +
            theme(strip.text.y = element_text(angle = 0)) +
            scale_colour_manual(
              values = col_vec_bw_curr
            ) +
            labs(
              x = "Marker expression",
              y = NULL,
              colour = "Core bandwidth",
              fill = NULL,
              linetype = "Threshold summary"
            ) +
            geom_vline(
              data = threshold_tbl_curr,
              mapping = aes(
                xintercept = threshold,
                colour = bw_core_lab,
                group = interaction(bw_core_lab, iter, sample)
              ),
              linewidth = 0.25,
              alpha = 0.22
            ) +
            geom_vline(
              data = threshold_tbl_curr_med_iqr,
              aes(
                xintercept = threshold,
                colour = bw_core_lab,
                linetype = threshold_stat,
                group = interaction(bw_core_lab, threshold_stat)
              ),
              linewidth = 0.45
            ) +
            scale_linetype_manual(
              values = c(
                "iqr_lower" = "dashed",
                "median" = "solid",
                "iqr_upper" = "dashed"
              ),
              breaks = c("iqr_lower", "median", "iqr_upper"),
              labels = c("IQR lower", "Median", "IQR upper")
            )

          path_p <- file.path(
            .analysis_fig_dir(fig_key, root_dir),
            "thresholds",
            "raw",
            "fixed_extra_vary_core",
            safe_file_lab(transformation_curr),
            paste0("bias_uns_", safe_file_lab(bias_uns_setting_curr)),
            paste0(
              "threshold_mean_pos_setting_",
              safe_file_lab(mean_pos_setting_curr),
              "_bw_extra_",
              format_bw_file(bw_extra_curr),
              ".png"
            )
          )
          dir.create(dirname(path_p), recursive = TRUE, showWarnings = FALSE)
          .analysis_save_fig(p, path_p, height = 22)
          .analysis_print_fig(p)
        }
      }
    }
  }
}
```

### Threshold summary table

The tables summarise threshold variation and relative frequency error for each bandwidth pair. The ranked table averages threshold-variation ranks and absolute relative error over cell counts and response frequencies within each transformation, bias and mean-expression setting.

```{r}
#| label: threshold-summary
#| echo: false
if (isTRUE(run_plots)) {
  # summary_tbl was collated and validated during promotion.
  summary_tbl |>
    dplyr::mutate(
      transformation = .analysis_trans_factor(.data$transformation)
    ) |>
    dplyr::arrange(.data$transformation) |>
    dplyr::mutate(dplyr::across(where(is.numeric), .analysis_label_number)) |>
    print(width = Inf)
}
```

```{r}
#| label: threshold-ranks
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  summary_tbl |>
    dplyr::select(
      transformation, bias_uns_setting, bias_uns,
      mean_pos_setting, bw_core, bw_extra, threshold_sd,
      propRespEst_median_diff_rel
    ) |>
    dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))) |>
    dplyr::arrange(threshold_sd) |>
    dplyr::mutate(
      threshold_sd_rank = dplyr::dense_rank(threshold_sd)
    ) |>
    dplyr::group_by(
      transformation, bias_uns_setting, bias_uns,
      mean_pos_setting, bw_core, bw_extra
    ) |>
    dplyr::summarise(
      mean_threshold_sd_rank = mean(threshold_sd_rank, na.rm = TRUE),
      mean_abs_rel_error = mean(abs(propRespEst_median_diff_rel), na.rm = TRUE),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      transformation = .analysis_trans_factor(.data$transformation)
    ) |>
    dplyr::arrange(transformation, bias_uns_setting, mean_threshold_sd_rank) |>
    dplyr::mutate(dplyr::across(where(is.numeric), .analysis_label_number)) |>
    knitr::kable() |>
    print()
}
```
