---
title: Manual adaptive bandwidth performance simulation
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

## Purpose

This notebook assesses how well scalar, automatic adaptive, and manually
specified adaptive bandwidth methods recover the known background-subtracted
response frequency in simulation.

It follows the same broad structure as the fixed-bandwidth frequency simulation:

1. define the simulation and bandwidth grid;
2. build `sim_grid`;
3. provide a manual loop for debugging individual failures;
4. run the same grid in parallel;
5. collate and rank methods by error.

The manual adaptive settings tested here are the package-level settings passed
through `gateStim()`:

```{r}
#| label: manual-adaptive-args-note
#| echo: true
# bwAdaptive = TRUE
# bwAdaptiveCore = <manual core bandwidth>
# bwAdaptiveExtra = <manual extra bandwidth>
# bwAdaptiveCrossover = <optional crossover expression value, or NULL>
# bwAdaptiveTransitionWidth = <0 for hard switch; >0 for smooth switch>
```

## Simulation and method settings

```{r}
#| label: actual-settings
set.seed(1)

# Simulation settings ---------------------------------------------------------
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl() |>
  dplyr::mutate(
    bias_uns = ifelse(transformation == "gamma", 0.01, 0.05)
  )

sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

# Bandwidth ranges ------------------------------------------------------------
# This is the same scalar bandwidth range as in the fixed-bandwidth simulation.
bandwidth_value_vec <- c(
  0.001, 0.0025, 0.005, 0.0075, 0.01, 0.0125,
  0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1
)

bw_core_vec <- bandwidth_value_vec
bw_extra_vec <- bandwidth_value_vec

# NULL/NA crossover means use the density-weighted component blend. Finite
# values force a core-to-extra crossover at that expression value.
bw_crossover_vec <- c(NA_real_, 4.5, 5.5, 6.5)
bw_transition_width_vec <- c(0, 0.25)

# Constant-bandwidth comparators use the same range as the existing simulation.
bw_constant_vec <- bandwidth_value_vec

# bias_uns_vec <- c(0, bandwidth_value_vec / 2)

# Gating/bandwidth settings ---------------------------------------------------
bw_mtd <- "hpi1Norm"
bw_adj <- 1
bw_ncell_min <- 100
bw_ncell_max <- 1e5
bw_cluster <- NULL
bw_min <- "none"
bw_max <- "none"
min_cell <- 100
max_pos_prob_x <- Inf
gate_combn <- "min"
tol_clust <- NULL

bw_adaptive_density_n <- NULL
bw_adaptive_pad_frac <- 0.15

norm_peak_frac <- 0.1
norm_peak_min_rel <- 0.75
norm_extra_frac <- 0.2
norm_extra_max <- Inf
norm_extra_jitter_frac <- 0.25
norm_lambda <- seq(-2, 2, length.out = 81)
norm_density_n <- 512L
norm_excess_bw_mtd <- "hpi3"
norm_excess_ncell <- 10000L
norm_adaptive_ncell <- 2500L
norm_mtd <- "moments"

# Simulation execution settings ---------------------------------------------
n_sample_run <- 5L
n_iter_run <- 5L
n_marker_run <- 1L
n_condition_run <- 2L
n_cluster_run <- 2L
prob_exact <- TRUE
cov_ev_min <- 1.5
cov_ev_max <- 1.5
calc_cyt_pos_gates <- FALSE
calc_single_pos_gates <- FALSE

if ("quick" %in% projr::projr_profile_get()) {
  n_cell_stim_vec <- c(1e3, 5e4)
  prob_response_vec <- c(0.002, 0.02)
  bw_core_vec <- c(0.02, 0.05, 0.1)
  bw_extra_vec <- c(0.1, 0.25, 0.5)
  bw_crossover_vec <- c(NA_real_, 5.5)
  bw_transition_width_vec <- c(0)
  bw_constant_vec <- c(0.02, 0.05, 0.1, 0.25, 0.5)
  bias_uns_vec <- c(0, 0.05)
  condition_perturbation_sd_vec <- 0
  n_sample_run <- 2L
  n_iter_run <- 2L
}
```

## Build method grid

```{r}
#| label: method-grid
manual_adaptive_grid <- tidyr::expand_grid(
  method_family = "manual_adaptive",
  bwAdaptive = TRUE,
  bw = NA_real_,
  bwAdaptiveCore = bw_core_vec,
  bwAdaptiveExtra = bw_extra_vec,
  bwAdaptiveCrossover = bw_crossover_vec,
  bwAdaptiveTransitionWidth = bw_transition_width_vec
) |>
  dplyr::filter(
    .data$bwAdaptiveExtra >= .data$bwAdaptiveCore
  )

manual_constant_grid <- tibble::tibble(
  method_family = "manual_constant",
  bwAdaptive = FALSE,
  bw = bw_constant_vec,
  bwAdaptiveCore = NA_real_,
  bwAdaptiveExtra = NA_real_,
  bwAdaptiveCrossover = NA_real_,
  bwAdaptiveTransitionWidth = NA_real_
)

auto_grid <- tibble::tibble(
  method_family = c("auto_constant", "auto_adaptive"),
  bwAdaptive = c(FALSE, TRUE),
  bw = NA_real_,
  bwAdaptiveCore = NA_real_,
  bwAdaptiveExtra = NA_real_,
  bwAdaptiveCrossover = NA_real_,
  bwAdaptiveTransitionWidth = NA_real_
)

method_grid <- dplyr::bind_rows(
  manual_adaptive_grid,
  manual_constant_grid,
  auto_grid
) |>
  dplyr::mutate(
    crossover_label = dplyr::if_else(
      is.na(.data$bwAdaptiveCrossover),
      "auto",
      formatC(.data$bwAdaptiveCrossover, format = "f", digits = 2)
    ),
    method_id = dplyr::case_when(
      .data$method_family == "manual_constant" ~ paste0(
        "constant_bw", formatC(.data$bw, format = "f", digits = 4)
      ),
      .data$method_family == "manual_adaptive" ~ paste0(
        "adapt_core", formatC(.data$bwAdaptiveCore, format = "f", digits = 4),
        "_extra", formatC(.data$bwAdaptiveExtra, format = "f", digits = 4),
        "_cross", .data$crossover_label,
        "_tw", formatC(.data$bwAdaptiveTransitionWidth, format = "f", digits = 3)
      ),
      TRUE ~ .data$method_family
    )
  ) |>
  dplyr::select(-crossover_label)

method_grid
```

## Build simulation grid

```{r}
#| label: sim-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bias_uns = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec,
  method_grid
) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim,
    method_family,
    method_id,
    bwAdaptive,
    bw,
    bwAdaptiveCore,
    bwAdaptiveExtra,
    bwAdaptiveCrossover,
    bwAdaptiveTransitionWidth
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::filter(
    (bias_uns <= 0.01 & transformation == "gamma") |
      (bias_uns >= 0.01 & transformation != "gamma")
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    .data$n_cell,
    dplyr::desc(.data$prob_response),
    .data$transformation,
    .data$mean_pos_setting,
    .data$method_family,
    .data$method_id
  )

if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::slice(1:3)
}

sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulation settings"))
```

## Manual loop for debugging

Run this chunk when a particular combination fails and you want an ordinary
sequential traceback/log.

```{r}
#| label: bw-manual-loop
#| eval: false
bw_list_raw <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))

  row <- sim_grid[i, , drop = FALSE]

 # debugonce(.verifyChnlSettingsChnl)

  .run_manual_adaptive_freq_bs(
    transformation = row$transformation,
    prob_response = row$prob_response,
    n_cell = row$n_cell,
    mean_pos_setting = row$mean_pos_setting,
    mean_pos = row$mean_pos,
    bias_uns = row$bias_uns,
    sample_perturbation_sd = row$sample_perturbation_sd,
    condition_perturbation_sd = row$condition_perturbation_sd,
    cluster_perturbation_sd = row$cluster_perturbation_sd,
    background_relative_to_response = row$background_relative_to_response,
    n_cell_uns_relative_to_stim = row$n_cell_uns_relative_to_stim,
    method_family = row$method_family,
    method_id = row$method_id,
    bwAdaptive = row$bwAdaptive,
    bw = row$bw,
    bwAdaptiveCore = row$bwAdaptiveCore,
    bwAdaptiveExtra = row$bwAdaptiveExtra,
    bwAdaptiveCrossover = row$bwAdaptiveCrossover,
    bwAdaptiveTransitionWidth = row$bwAdaptiveTransitionWidth,
    sim_id = row$sim_id,
    nSample = 1,
    nIter = 1,
  )

  bw_list_raw[[i]] <- sim_res
}
```

## Parallel processing

```{r}
#| label: bw-manual-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache", "log", "analysis", "sim", "bw", "manual_adaptive_perf", "progress.txt"
)
dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

future::plan(
  future::multisession,
  workers = max(1, .simGetCores())
)

bw_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      method_family,
      method_id,
      bwAdaptive,
      bw,
      bwAdaptiveCore,
      bwAdaptiveExtra,
      bwAdaptiveCrossover,
      bwAdaptiveTransitionWidth,
      sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | method = ", method_id,
        " | bias = ", bias_uns,
        " | condition_sd = ", condition_perturbation_sd,
        " | sim_id = ", sim_id
      )

      cat(paste0("Running: ", settings_row, "\n"), file = path_progress_file, append = TRUE)

      out <- tryCatch(
        {
          sim_res <- .run_manual_adaptive_freq_bs(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response = background_relative_to_response,
            n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim,
            method_family = method_family,
            method_id = method_id,
            bwAdaptive = bwAdaptive,
            bw = bw,
            bwAdaptiveCore = bwAdaptiveCore,
            bwAdaptiveExtra = bwAdaptiveExtra,
            bwAdaptiveCrossover = bwAdaptiveCrossover,
            bwAdaptiveTransitionWidth = bwAdaptiveTransitionWidth,
            sim_id = sim_id
          )

          cat(paste0("Completed: ", settings_row, "\n"), file = path_progress_file, append = TRUE)
          p(sprintf("Completed sim_id: %s", sim_id))
          sim_res
        },
        error = function(e) {
          cat(
            paste0("Error [", settings_row, "]: ", e$message, "\n"),
            file = path_progress_file,
            append = TRUE
          )
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response = background_relative_to_response,
            n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim,
            method_family = method_family,
            method_id = method_id,
            bwAdaptive = bwAdaptive,
            bw = bw,
            bwAdaptiveCore = bwAdaptiveCore,
            bwAdaptiveExtra = bwAdaptiveExtra,
            bwAdaptiveCrossover = bwAdaptiveCrossover,
            bwAdaptiveTransitionWidth = bwAdaptiveTransitionWidth,
            sim_id = sim_id,
            iter = NA_integer_,
            sample = NA_character_,
            ind = NA_character_,
            chnl = "F1",
            propRespTruth = NA_real_,
            propRespEst = NA_real_,
            threshold = NA_real_,
            error = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

path_raw_list <- projr::projr_path_get(
  "cache", "sim", "bw", "manual_adaptive_perf", "bw_list_raw.rds"
)
dir.create(dirname(path_raw_list), recursive = TRUE, showWarnings = FALSE)
saveRDS(bw_list_raw, file = path_raw_list)
```

## Collate results

```{r}
#| label: bw-manual-collate
if (exists("bw_list_raw") && length(bw_list_raw) > 0L) {
  bw_list_raw_collate <- bw_list_raw
} else {
  path_vec <- c(
    projr::projr_path_get(
      "cache", "sim", "bw", "manual_adaptive_perf", "bw_list_raw.rds", safe = TRUE
    ),
    projr::projr_path_get(
      "cache", "sim", "bw", "manual_adaptive_perf", "bw_list_raw.rds", safe = FALSE
    )
  )
  path_vec <- path_vec[file.exists(path_vec)]
  if (length(path_vec) == 0L) {
    stop("Could not find bw_list_raw.rds, and bw_list_raw is not available in memory.")
  }
  bw_list_raw_collate <- readRDS(file = path_vec[[1]])
}

bw_tbl_raw <- bw_list_raw_collate |>
  purrr::list_rbind()

bw_tbl_raw <- .add_missing_cols(
  bw_tbl_raw,
  list(
    error = NA_character_,
    propRespTruth = NA_real_,
    propRespEst = NA_real_,
    threshold = NA_real_
  )
)

bw_tbl_results_raw <- bw_tbl_raw |>
  dplyr::mutate(
    propRespError = .data$propRespEst - .data$propRespTruth,
    propRespAbsError = abs(.data$propRespError),
    propRespRelError = dplyr::if_else(
      is.finite(.data$propRespTruth) & .data$propRespTruth != 0,
      .data$propRespError / .data$propRespTruth,
      NA_real_
    ),
    propRespAbsRelError = abs(.data$propRespRelError)
  )

summary_group_cols <- c(
  colnames(sim_grid),
  "chnl"
)
summary_group_cols <- intersect(summary_group_cols, names(bw_tbl_results_raw))

bw_tbl_results_summary <- bw_tbl_results_raw |>
  dplyr::group_by(dplyr::across(dplyr::all_of(summary_group_cols))) |>
  dplyr::summarise(
    n_est = sum(is.finite(.data$propRespEst)),
    n_error = sum(!is.na(.data$error)),
    propRespTruth_median = stats::median(.data$propRespTruth, na.rm = TRUE),
    propRespEst_median = stats::median(.data$propRespEst, na.rm = TRUE),
    propRespError_mean = mean(.data$propRespError, na.rm = TRUE),
    propRespError_median = stats::median(.data$propRespError, na.rm = TRUE),
    propRespAbsError_median = stats::median(.data$propRespAbsError, na.rm = TRUE),
    propRespAbsRelError_median = stats::median(.data$propRespAbsRelError, na.rm = TRUE),
    propRespRmse = sqrt(mean(.data$propRespError^2, na.rm = TRUE)),
    threshold_median = stats::median(.data$threshold, na.rm = TRUE),
    threshold_q25 = stats::quantile(.data$threshold, 0.25, na.rm = TRUE),
    threshold_q75 = stats::quantile(.data$threshold, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

method_rank_tbl <- bw_tbl_results_summary |>
  dplyr::group_by(
    .data$method_family,
    .data$method_id,
    .data$bwAdaptive,
    .data$bw,
    .data$bwAdaptiveCore,
    .data$bwAdaptiveExtra,
    .data$bwAdaptiveCrossover,
    .data$bwAdaptiveTransitionWidth
  ) |>
  dplyr::summarise(
    n_scenarios = dplyr::n(),
    n_est_total = sum(.data$n_est, na.rm = TRUE),
    median_abs_error = stats::median(.data$propRespAbsError_median, na.rm = TRUE),
    median_abs_rel_error = stats::median(.data$propRespAbsRelError_median, na.rm = TRUE),
    median_rmse = stats::median(.data$propRespRmse, na.rm = TRUE),
    mean_bias = mean(.data$propRespError_mean, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(.data$median_abs_rel_error, .data$median_abs_error)

path_out_dir <- projr::projr_path_get("cache", "sim", "bw", "manual_adaptive_perf")
dir.create(path_out_dir, recursive = TRUE, showWarnings = FALSE)

saveRDS(bw_tbl_results_raw, file = file.path(path_out_dir, "bw_tbl_results_raw.rds"))
saveRDS(bw_tbl_results_summary, file = file.path(path_out_dir, "bw_tbl_results_summary.rds"))
saveRDS(method_rank_tbl, file = file.path(path_out_dir, "method_rank_tbl.rds"))
readr::write_csv(bw_tbl_results_summary, file.path(path_out_dir, "bw_tbl_results_summary.csv"))
readr::write_csv(method_rank_tbl, file.path(path_out_dir, "method_rank_tbl.csv"))

method_rank_tbl
```

## Best methods overall

```{r}
#| label: best-method-table
method_rank_tbl |>
  dplyr::slice_head(n = 25)
```

## Plot performance by number of cells

```{r}
#| label: plot-performance-by-ncell
top_method_tbl <- method_rank_tbl |>
  dplyr::slice_head(n = 12) |>
  dplyr::select(method_id)

plot_tbl <- bw_tbl_results_summary |>
  dplyr::semi_join(top_method_tbl, by = "method_id")

p <- ggplot(
  plot_tbl,
  aes(
    x = n_cell,
    y = propRespAbsRelError_median,
    colour = method_id,
    group = method_id
  )
) +
  geom_line(alpha = 0.8) +
  geom_point(size = 0.8) +
  scale_x_log10() +
  scale_y_log10() +
  facet_grid(transformation ~ prob_response, scales = "free_y") +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "bottom",
    legend.text = element_text(size = 6)
  ) +
  guides(colour = guide_legend(ncol = 1)) +
  labs(
    title = "Median absolute relative error by method",
    x = "Stimulated cells",
    y = "Median absolute relative error",
    colour = "Method"
  )

path_p <- file.path(path_out_dir, "median_abs_rel_error_by_n_cell.png")
ggsave(path_p, p, width = 24, height = 30, units = "cm")
p
```

## Manual adaptive heatmap

This is most useful for the density-weighted crossover-free adaptive methods.

```{r}
#| label: manual-adaptive-heatmap
heat_tbl <- bw_tbl_results_summary |>
  dplyr::filter(
    .data$method_family == "manual_adaptive",
    is.na(.data$bwAdaptiveCrossover),
    .data$bwAdaptiveTransitionWidth == 0
  ) |>
  dplyr::group_by(
    .data$transformation,
    .data$mean_pos_setting,
    .data$bwAdaptiveCore,
    .data$bwAdaptiveExtra
  ) |>
  dplyr::summarise(
    median_abs_rel_error = stats::median(.data$propRespAbsRelError_median, na.rm = TRUE),
    .groups = "drop"
  )

if (nrow(heat_tbl) > 0L) {
  p_heat <- ggplot(
    heat_tbl,
    aes(
      x = factor(bwAdaptiveCore),
      y = factor(bwAdaptiveExtra),
      fill = median_abs_rel_error
    )
  ) +
    geom_tile() +
    facet_grid(mean_pos_setting ~ transformation) +
    scale_fill_viridis_c(trans = "log10", option = "magma", direction = -1) +
    cowplot::theme_cowplot() +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA),
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      title = "Manual adaptive bandwidth grid: density-weighted crossover",
      x = "Core bandwidth",
      y = "Extra bandwidth",
      fill = "Median absolute\nrelative error"
    )

  path_p_heat <- file.path(path_out_dir, "manual_adaptive_core_extra_heatmap.png")
  ggsave(path_p_heat, p_heat, width = 24, height = 18, units = "cm")
  p_heat
}
```

## Bias versus variance summary

```{r}
#| label: bias-variance-summary
method_rank_tbl |>
  dplyr::mutate(
    bw = dplyr::if_else(is.finite(.data$bw), .data$bw, NA_real_),
    bwAdaptiveCore = dplyr::if_else(is.finite(.data$bwAdaptiveCore), .data$bwAdaptiveCore, NA_real_),
    bwAdaptiveExtra = dplyr::if_else(is.finite(.data$bwAdaptiveExtra), .data$bwAdaptiveExtra, NA_real_)
  ) |>
  dplyr::select(
    method_family,
    method_id,
    bw,
    bwAdaptiveCore,
    bwAdaptiveExtra,
    bwAdaptiveCrossover,
    bwAdaptiveTransitionWidth,
    n_scenarios,
    n_est_total,
    median_abs_error,
    median_abs_rel_error,
    median_rmse,
    mean_bias
  ) |>
  dplyr::arrange(.data$median_abs_rel_error) |>
  dplyr::slice_head(n = 50)
```

## Threshold distributions for top methods

```{r}
#| label: threshold-distributions
threshold_plot_tbl <- bw_tbl_results_raw |>
  dplyr::semi_join(top_method_tbl, by = "method_id") |>
  dplyr::filter(is.finite(.data$threshold))

if (nrow(threshold_plot_tbl) > 0L) {
  p_thresh <- ggplot(
    threshold_plot_tbl,
    aes(x = method_id, y = threshold, colour = method_family)
  ) +
    ggforce::geom_sina(alpha = 0.25, size = 0.6) +
    facet_grid(transformation ~ mean_pos_setting, scales = "free_y") +
    coord_flip() +
    cowplot::theme_cowplot() +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA),
      legend.position = "bottom",
      axis.text.y = element_text(size = 5)
    ) +
    labs(
      title = "Threshold distributions for top-ranked methods",
      x = "Method",
      y = "Threshold",
      colour = "Method family"
    )

  path_p_thresh <- file.path(path_out_dir, "threshold_distributions_top_methods.png")
  ggsave(path_p_thresh, p_thresh, width = 24, height = 24, units = "cm")
  p_thresh
}
```
