---
title: Choosing the F-beta and Tailgate settings
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_plots: false
  simulation_seed: 60600
---

Analyses 7 and 8 compare StimGate with two other ways of setting a gate,
F-beta and Tailgate. This page chooses the settings those two methods use. It
looks at a response probability of 0.2%, all three transformations, both
degrees of separation, and parent populations of 5,000 and 100,000 cells.

For Tailgate, the settings are the derivative tolerance and the bias added to
its cutpoint. For F-beta, the setting is beta. Each choice is compared with
the published default settings and with the settings Analyses 7 and 8 use now.

## Set-up

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(ggplot2)
  library(tibble)
  library(dplyr)
  library(purrr)
  library(tidyr)
})

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev

simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  60600L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}

devtools::load_all(root_dir, quiet = TRUE)

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "flowStats", "simcyto", "ks", "reticulate"))
}

# Source simulation helpers in dependency order.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R",
  "sim-compare-tune.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

## Scientific settings

The data are simulated as in Analysis 7: one marker, negative cells centred at
zero and responding cells centred at `mean_pos`, with variance 1.5, then
transformed (Gaussian, skewed or gamma). The low and high separations are
Analysis 7's. In the stimulated tube 0.2% of the cells respond. The
unstimulated tube has background responders at 0.2 times that rate (0.04%) and
the same number of cells. The seeds differ from Analysis 7's, so the settings
are chosen on different simulated data from the data used to report results.

Only F-beta and Tailgate are run. Both use the same inputs as in Analysis 7:
F-beta compares the stimulated tube with its unstimulated tube, and Tailgate
gates the stimulated tube on its own. As in Analysis 7, when a method finds
no cutpoint (without an error), the gate is placed just above the highest
cell. Errors are kept as missing results.

Tailgate settings:

- **Tolerance.** Tailgate's cutpoint is the first point to the right of the
  main peak, after the steepest descent, where the slope of the density is
  flatter than the tolerance. cytoUtils' default tolerance is an absolute
  value, 0.01, which depends on the scale of the data. Analysis 7 uses
  `autoTol = TRUE`: a tolerance of 1e-2 times the steepest slope. The grid
  varies that fraction in half steps on a log10 scale, from 1e-6 to 1e-1, and
  keeps the absolute 0.01 for comparison. A smaller tolerance moves the
  cutpoint further right; if the density never becomes that flat, Tailgate
  finds no cutpoint and the gate falls back to just above the highest cell.
- **Bias.** The cutpoint is then moved up by the bias, as StimGate moves its
  gates up by its own bias. Analysis 7 uses a bias of 0.1.

F-beta setting:

- **beta.** F-beta picks the gate that maximises the F-beta score of the
  stimulated tube against the unstimulated tube. Beta below 1 weights
  precision more than recall, which pushes the gate up; beta above 1 pushes
  it down. The published value is 0.8. The other F-beta settings stay at their
  published values (theta 2, moving-average width 10).

```{r}
#| label: scientific-settings
tune_settings <- .simTuneSettings()
sim_grid_full <- .simTuneGrid(simulation_seed, prob_response = 0.002, n_cell = c(5e3, 1e5))

# Datasets of 20 samples, each sample one unstimulated/stimulated pair. The
# dataset is the unit for uncertainty; draft runs use fewer datasets.
sim_size <- .analysis_sim_size()
n_sample <- 20L
n_iter <- c(final = 10L, draft = 3L)[[sim_size]]
# Samples of the first dataset whose gates are drawn.
n_hist_sample <- 4L
# Settings whose mean median F1 is within this of the best are compared on
# their estimates.
f1_tolerance <- 0.02

sim_grid <- sim_grid_full
# Dev keeps one debugging scenario; quick keeps the smallest grid that still
# draws every figure.
if (analysis_dev) {
  sim_grid <- dplyr::filter(
    sim_grid, .data$transformation == "gamma", .data$mean_pos_setting == "low",
    .data$n_cell == 1e5
  )
  n_sample <- 2L
  n_iter <- 1L
  n_hist_sample <- 2L
} else if (analysis_quick) {
  sim_grid <- dplyr::filter(
    sim_grid, .data$transformation %in% c("gaussian", "gamma"),
    .data$mean_pos_setting == "low", .data$n_cell == 5e3
  )
  n_sample <- 3L
  n_iter <- 1L
  n_hist_sample <- 2L
}

path_fbeta <- file.path(root_dir, "scripts", "python", "fbeta.py")
if (isTRUE(run_simulations) && !file.exists(path_fbeta)) {
  stop("F-beta comparison script not found: ", path_fbeta)
}
```

Each scenario has `r n_iter` datasets of `r n_sample` samples, so
`r n_iter * n_sample` stimulated tubes. Tailgate is run with
`r length(tune_settings$tailgate$tol_rel) + 1L` tolerances and
`r length(tune_settings$tailgate$bias)` biases, and F-beta with
`r length(tune_settings$fbeta$beta)` values of beta, all on the same tubes.

## Simulation

```{r}
#| label: simulate
#| include: false
analysis_key <- .analysis_mode_key(c("sim", "tune-comparators"))
fig_key <- .analysis_mode_key("6-sim-tune-comparators")
dir_cache <- .analysis_cache_dir(analysis_key, root_dir, create = run_simulations)
path_results <- file.path(dir_cache, "results.rds")

cache_settings <- .simTuneCacheSettings(
  grid = sim_grid,
  n_sample = n_sample,
  n_iter = n_iter,
  settings = tune_settings,
  simulation_seed = simulation_seed,
  profile = list(
    analysis_dev = analysis_dev, analysis_quick = analysis_quick,
    sim_size = sim_size, n_hist_sample = n_hist_sample
  )
)

if (isTRUE(run_simulations)) {
  # Datasets run in parallel; each has its own seed, so the number of
  # workers does not change the results.
  results <- .simTuneRunGrid(
    sim_grid, n_sample, n_iter, path_fbeta, tune_settings, n_hist_sample,
    workers = max(1L, .simGetCores()), root_dir = root_dir
  )
  .simTuneWriteCache(results, cache_settings, path_results)
}
```

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_target <- 10L
rerun <- .simTuneRunScenario(
  dplyr::filter(sim_grid_full, .data$sim_id == sim_id_target),
  n_sample, n_iter,
  fbeta_env = .simCompareFbetaEnvironment(pathFbeta = path_fbeta),
  settings = tune_settings, n_hist_sample = n_hist_sample
)
```

## Collation

```{r}
#| label: collate
#| include: false
if (!isTRUE(run_simulations)) {
  results <- .simTuneReadCache(path_results, cache_settings, analysis_key)
}
metrics <- .simTuneMetrics(results$scores)
summary_tbl <- .simTuneSummary(metrics)
rank_tbl <- .simTuneRank(summary_tbl, f1_tolerance = f1_tolerance)
compare_tbl <- .simTuneCompareSettings(rank_tbl)
selected_tbl <- dplyr::filter(rank_tbl, .data$selected)
selected_lab <- function(m) {
  lab <- selected_tbl$setting_lab[selected_tbl$method == m]
  if (length(lab) == 0L) "none (no setting had a finite F1 in every scenario)" else lab
}
n_cell_vec <- sort(unique(sim_grid$n_cell))
```

## How the settings are chosen

Two things are judged, each over the stimulated tubes of a scenario:

- **F1 score**, the median and the 10th and 90th percentiles. F1 combines
  sensitivity and precision of the gate against the simulated labels:
  2TP / (2TP + FP + FN).
- **Estimated frequency**, the 10th and 90th percentiles of the
  background-subtracted frequency (stimulated minus unstimulated cells above
  the gate). The truth is 0.2%. The estimate error of a setting is the larger
  of the relative errors of these two percentiles, so it is small only when the
  middle 80% of estimates lies close to the truth.

The percentiles pool the tubes of all datasets in a scenario. The twelve
scenarios count equally. For each method, the settings whose mean median F1 is
within `r f1_tolerance` of the best are kept, and of those the one with the
smallest mean estimate error is chosen (ties go to the higher mean 10th
percentile of F1).

```{r}
#| label: selection
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  cat("\n\n**Selected F-beta setting:** ", selected_lab("fbeta"), ".\n\n", sep = "")
  cat("**Selected Tailgate setting:** ", selected_lab("tailgate"), ".\n\n", sep = "")
  rank_display <- rank_tbl |>
    dplyr::group_by(.data$method) |>
    dplyr::slice_head(n = 5L) |>
    dplyr::ungroup() |>
    dplyr::transmute(
      setting = .data$setting_lab,
      `mean median F1` = round(.data$mean_f1_median, 3),
      `mean 10th pct F1` = round(.data$mean_f1_p10, 3),
      `mean estimate error` = round(.data$mean_est_band_error, 3),
      `truth in 10th-90th` = paste0(.data$n_truth_in_band, " / ", .data$n_scenarios)
    )
  print(knitr::kable(rank_display, caption = paste(
    "The five best settings of each method by the rule above (mean over",
    "scenarios). Estimate error is the larger relative error of the 10th and",
    "90th percentiles of the estimates."
  )))
  .analysis_report_table(rank_tbl, c(fig_key, "setting_rank.csv"),
    "Every setting of each method, averaged over scenarios, in selection order.",
    root_dir)
}
```

## Selected against default settings

Each point is the median over the stimulated tubes of one scenario, and the
bar runs from the 10th to the 90th percentile. The reference settings are
F-beta's published beta (0.8); Tailgate with cytoUtils' default absolute
tolerance and no bias; Tailgate with Analysis 7's automatic tolerance and no
bias ("Tailgate (default settings)" in Analysis 7); and Tailgate as Analysis 7
runs it now (automatic tolerance 1%, bias 0.1).

### F1

```{r}
#| label: fig-compare-f1
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotCompare(summary_tbl, compare_tbl, "f1")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "compare"), root_dir), "compare-f1.png"
  ), height = 22)
}
```

### Estimated frequency

Estimates are shown relative to the truth, so 1 (dashed) is correct.

```{r}
#| label: fig-compare-est
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotCompare(summary_tbl, compare_tbl, "est")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "compare"), root_dir), "compare-estimate.png"
  ), height = 22)
}
```

### Paired differences within datasets

For each dataset, the median F1 and the median estimate under the selected
setting minus those under each reference setting. The table gives the mean
over datasets and its standard error (datasets are independent; the tubes of a
dataset are not separate units here).

```{r}
#| label: dataset-differences
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  diff_tbl <- .simTuneDatasetDifferences(metrics, compare_tbl)
  .analysis_report_table(diff_tbl, c(fig_key, "dataset_differences.csv"),
    "Selected minus reference setting, per scenario: mean and standard error over datasets of the dataset median F1 and median estimate.",
    root_dir)
}
```

## Gates

### On a subset of samples

The first `r n_hist_sample` samples of the first dataset in each scenario
(chosen in advance). Bars are the stimulated tube, with simulated responders
(orange) stacked on the negative cells (grey); the dark outline is the
unstimulated tube. Counts use a pseudo-log scale so that the few responding
cells stay visible. Vertical lines are the gates of the compared settings;
selected settings are dashed, so they remain visible where they give the same
gate as a reference setting.

```{r}
#| label: fig-gates
#| echo: false
#| results: asis
#| fig-width: 6.3
#| fig-height: 6.5
if (isTRUE(run_plots)) {
  scen_tbl <- dplyr::distinct(sim_grid, .data$sim_id, .data$transformation,
    .data$mean_pos_setting, .data$mean_pos, .data$n_cell)
  for (trans in unique(scen_tbl$transformation)) {
    .analysis_heading(trans, 4L)
    for (i in which(scen_tbl$transformation == trans)) {
      row <- scen_tbl[i, ]
      .analysis_heading(paste0(
        row$mean_pos_setting, " separation (", row$mean_pos, "), ",
        .simTuneCellLab(row$n_cell)
      ), 5L)
      p <- .simTunePlotGates(
        dplyr::filter(results$hist, .data$sim_id == row$sim_id),
        dplyr::filter(metrics, .data$sim_id == row$sim_id),
        compare_tbl
      )
      .analysis_print_save_fig(p, file.path(
        .analysis_fig_dir(c(fig_key, "gates"), root_dir),
        paste0("gates-", row$transformation, "-", row$mean_pos_setting,
          "-n_cell_", format(row$n_cell, scientific = FALSE), ".png")
      ), height = 17)
    }
  }
}
```

### Across all tubes

The gates of every stimulated tube under the compared settings.

```{r}
#| label: fig-gate-spread
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotGateSpread(metrics, compare_tbl)
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "gates"), root_dir), "gate-spread.png"
  ), height = 22)
}
```

## Tailgate: tolerance and bias

Each cell is one tolerance and bias in one scenario. The red box marks the
selected setting. Tolerances are fractions of the steepest slope, except
"abs 0.01", cytoUtils' default absolute tolerance.

### Median F1

```{r}
#| label: fig-tailgate-heat-f1
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateHeat(summary_tbl, "f1_median", "Median F1",
    dplyr::filter(selected_tbl, .data$method == "tailgate"))
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "heat-f1-median.png"
  ), height = 22)
}
```

### 10th percentile of F1

```{r}
#| label: fig-tailgate-heat-f1-p10
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateHeat(summary_tbl, "f1_p10", "10th pct F1",
    dplyr::filter(selected_tbl, .data$method == "tailgate"))
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "heat-f1-p10.png"
  ), height = 22)
}
```

### Estimate error

The larger of the relative errors of the 10th and 90th percentiles of the
estimates (0 is best).

```{r}
#| label: fig-tailgate-heat-est
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateHeat(summary_tbl, "est_band_error", "Estimate error",
    dplyr::filter(selected_tbl, .data$method == "tailgate"), direction = -1)
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "heat-estimate-error.png"
  ), height = 22)
}
```

### F1 against tolerance

The median F1 against the tolerance (log10 scale), one line per bias; the
bands run from the 10th to the 90th percentile. The dotted line is Analysis
7's tolerance (1e-2). cytoUtils' absolute tolerance is not on this scale and
appears only in the heat maps.

```{r}
#| label: fig-tailgate-tol-f1
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateTol(summary_tbl, "f1")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "tol-f1.png"
  ), height = 22)
}
```

### Estimated frequency against tolerance

```{r}
#| label: fig-tailgate-tol-est
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateTol(summary_tbl, "est")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "tol-estimate.png"
  ), height = 22)
}
```

### F1 against bias

The median F1 against bias, one line per tolerance; the bands run from the
10th to the 90th percentile.

```{r}
#| label: fig-tailgate-lines-f1
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateLines(summary_tbl, "f1")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "lines-f1.png"
  ), height = 22)
}
```

### Estimated frequency against bias

```{r}
#| label: fig-tailgate-lines-est
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotTailgateLines(summary_tbl, "est")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "tailgate"), root_dir), "lines-estimate.png"
  ), height = 22)
}
```

## F-beta: beta

### F1

```{r}
#| label: fig-fbeta-f1
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotFbeta(summary_tbl, "f1")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "fbeta"), root_dir), "beta-f1.png"
  ), height = 22)
}
```

### Estimated frequency

```{r}
#| label: fig-fbeta-est
#| echo: false
#| fig-width: 6.3
#| fig-height: 9
if (isTRUE(run_plots)) {
  p <- .simTunePlotFbeta(summary_tbl, "est")
  .analysis_print_save_fig(p, file.path(
    .analysis_fig_dir(c(fig_key, "fbeta"), root_dir), "beta-estimate.png"
  ), height = 22)
}
```

## Coverage

Errors and fallback gates, by scenario and setting. A fallback gate sits just
above the highest cell, because the method found no cutpoint.

```{r}
#| label: coverage
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  coverage_tbl <- summary_tbl |>
    dplyr::select(dplyr::all_of(c(.simTuneScenarioCols, "method", "setting_lab")),
      "n", "n_datasets", "n_error", "n_fallback", "n_f1_finite", "n_est_finite")
  .analysis_report_table(coverage_tbl, c(fig_key, "coverage.csv"),
    "Tubes, errors, fallback gates and finite F1 and estimates for every scenario and setting.",
    root_dir)
  .analysis_report_table(summary_tbl, c(fig_key, "summary.csv"),
    "Median and 10th/90th percentiles of F1 and of the estimates for every scenario and setting.",
    root_dir, quiet = TRUE)
}
```

```{r}

```
