---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
---

This analysis compares background-subtracted response frequencies from StimGate,
F-beta and Tailgate. Paired stimulated and unstimulated samples are simulated
with different response frequencies, cell counts and expression distributions.
The estimates are compared with the known response frequency and their threshold
fallbacks are recorded.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
analysis_qmd <- "analysis/7-sim-compare-freq_bs.qmd"

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "corrected-comparison-v4"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
analysis_key <- .analysis_mode_key(analysis_key)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  stop("Expected the stimgate DESCRIPTION at the repository root: ", root_dir)
}

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "simcyto"))
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

```

## Scientific settings

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at
   `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the `cytoUtils` package.

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but uses common histogram support spanning the paired
stimulated and unstimulated distributions. The published code derives the bin
range from the negative distribution alone, which can omit a simulated response
tail beyond the observed negative range.

For StimGate, bandwidth is estimated automatically using the ordinary `hpi1`
selector, without background normalisation, with the bandwidth-estimation sample
capped at 10,000 cells. Tailgate is deliberately left to use its usual automatic
bandwidth estimation so that it is evaluated as it would normally run.

### Simulation settings

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
condition_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1:2]
cluster_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
background_relative_to_response_vec <- c(
  2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001
)[4]
n_cell_uns_relative_to_stim_vec <- c(
  0.1, 0.2, 0.5, 1, 2, 5, 10
)[4]

# StimGate bandwidth settings.
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# StimGate method settings for the full method comparison.
# Cluster refinement is enabled to represent the full one-marker procedure.
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
n_sample_sim <- if (analysis_quick) 1L else 10
n_iter_sim <- if (analysis_quick) 1L else 2

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- file.path(
  root_dir,
  "scripts",
  "python",
  "fbeta.py"
)
if (isTRUE(run_simulations)) {
  if (!file.exists(path_fbeta)) {
    stop("F-beta comparison script not found: ", path_fbeta)
  }
  .simCompareFbetaEnvironment(pathFbeta = path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"
```

## Simulation

### Simulation grid

```{r}
#| label: compare-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = stimgate_bw_mtd,
  bw_ncell_max = stimgate_bw_ncell_max,
  bw_fallback = stimgate_bw_fallback,
  bw_min = stimgate_bw_min,
  bw_max = stimgate_bw_max,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::mutate(
    bias_uns = dplyr::case_when(
      transformation == "gamma" ~ 0.01,
      transformation %in% c("gaussian", "skew") ~ 0.15,
      TRUE ~ NA_real_
    )
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_max,
    bw_fallback,
    bw_min,
    bw_max,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    n_cell,
    dplyr::desc(prob_response)
  ) |>
  dplyr::mutate(
    sim_seed = as.integer(simulation_seed + sim_id - 1L)
  )

# Keep every seeded row available for an exact interactive rerun.
sim_grid_full <- sim_grid

# Quick mode keeps cells, response and transformation; all comparison methods remain.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(5e3, 2e4),
      prob_response %in% c(0.002, 0.2),
      transformation %in% c("gaussian", "gamma"),
      mean_pos_setting == "high",
      condition_perturbation_sd == 0
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      condition_perturbation_sd == 0
    )
}

sim_grid_all <- sim_grid

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

sim_grid
print(paste0("Total simulation scenarios: ", nrow(sim_grid)))
```

Run the simulations in parallel, with progress reporting. Each scenario has a
deterministic seed `simulation_seed + sim_id - 1`, so its simulation is stable
across worker scheduling and grid chunking.

```{r}
#| label: compare-parallel
#| include: false
n_workers <- max(1, .simGetCores())

analysis_result_params <- list(
  comparison_semantics_version = comparison_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  n_cell_stim_vec = n_cell_stim_vec,
  prob_response_vec = prob_response_vec,
  transformation_vec = transformation_vec,
  mean_pos_settings_tbl = mean_pos_settings_tbl,
  sample_perturbation_sd_vec = sample_perturbation_sd_vec,
  condition_perturbation_sd_vec = condition_perturbation_sd_vec,
  cluster_perturbation_sd_vec = cluster_perturbation_sd_vec,
  background_relative_to_response_vec = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim_vec = n_cell_uns_relative_to_stim_vec,
  bias_uns_gamma = 0.01,
  bias_uns_gaussian_skew = 0.15,
  prob_exact = TRUE,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  stimgate_bw_mtd = stimgate_bw_mtd,
  stimgate_bw_ncell_max = stimgate_bw_ncell_max,
  stimgate_bw_fallback = stimgate_bw_fallback,
  stimgate_bw_min = stimgate_bw_min,
  stimgate_bw_max = stimgate_bw_max,
  tol_clust = tol_clust,
  loc_enforce_shape_threshold = loc_enforce_shape_threshold,
  calc_cyt_pos_gates = calc_cyt_pos_gates,
  fbeta_beta = fbeta_beta,
  fbeta_theta = fbeta_theta,
  fbeta_width = fbeta_width,
  tailgate_adjust = tailgate_adjust,
  tailgate_method = tailgate_method,
  tailgate_tol = tailgate_tol,
  tailgate_auto_tol = tailgate_auto_tol,
  tailgate_x = tailgate_x
)


compare_raw <- tibble::tibble()

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks
      ),
      analysis_result_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )

  if (interactive()) {
    progressr::handlers(global = TRUE)
  }

  if (nrow(sim_grid) > 0L) {
    compare_raw <- .simCompareFreqBsGrid(
      sim_grid = sim_grid,
      nSample = n_sample_sim,
      nIter = n_iter_sim,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      probExact = TRUE,
      covEvMin = 1.5,
      covEvMax = 1.5,
      tolClust = tol_clust,
      locEnforceShapeThreshold = loc_enforce_shape_threshold,
      calcCytPosGates = calc_cyt_pos_gates,
      pathFbeta = path_fbeta,
      fbetaBeta = fbeta_beta,
      fbetaTheta = fbeta_theta,
      fbetaWidth = fbeta_width,
      tailgateAdjust = tailgate_adjust,
      tailgateAutoTol = tailgate_auto_tol,
      tailgateMethod = tailgate_method,
      tailgateTol = tailgate_tol,
      tailgateX = tailgate_x,
      parallel = (n_workers > 1L),
      workers = n_workers,
      dirCache = dir_output,
      pathProgress = path_progress_file,
      resume = TRUE,
      retryErrors = TRUE,
      sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
        sim_grid_chunk_index
      } else {
        NULL
      },
      sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
        sim_grid_n_chunks
      } else {
        NULL
      }
    )

    .write_rds_atomic(compare_raw, path_rds)
  } else {
    message("No simulations assigned to this chunk.")
  }

  chunk_check <- .simCompareGridOutputStatus(
    compare_raw,
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )
  chunk_collate_ok <- isTRUE(chunk_check$collate_ok)
  chunk_validation_ok <- isTRUE(chunk_check$validation_ok)
  n_failed <- length(chunk_check$failed_ids)
  n_completed <- length(chunk_check$completed_ids)

  validation_error <- if (!chunk_collate_ok) {
    paste0(
      "Chunk collation does not match the assigned simulation grid. Missing sim_id: ",
      paste(chunk_check$missing_ids, collapse = ", "),
      "; extra sim_id: ",
      paste(chunk_check$extra_ids, collapse = ", "),
      "."
    )
  } else if (length(chunk_check$failed_ids) > 0L) {
    paste0(
      "Chunk contains incomplete, non-finite or errored primary method results for sim_id: ",
      paste(chunk_check$failed_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_collate_ok,
    validation_ok = chunk_validation_ok,
    error_message = validation_error
  )

  if (!isTRUE(chunk_validation_ok)) {
    stop(validation_error)
  }

  .simComparePromoteIfReady(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )

  if (sim_grid_n_chunks > 1L) {
    message(
      "Completed this simulation chunk. ",
      "Render with run_simulations = FALSE after all chunks are promoted ",
      "to inspect summaries or write plots."
    )
    knitr::knit_exit()
  }
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
  message("Skipping simulations; using canonical promoted results.")
}
```

## Collation

```{r}
#| label: compare-collate
#| echo: false
# Without `run_ctx` (simulation chunk not run in this session), read the
# canonical promoted results under `current/`.
if (!exists("run_ctx")) {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}

path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = analysis_result_params
)
compare_raw <- .analysis_read_rds(path_rds, analysis_key, analysis_qmd)

compare_tbl <- compare_raw

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(
    method %in% c(
      "stimgate",
      "fbeta",
      "tailgate"
    )
  ) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

scenario_cols <- c(
  setdiff(colnames(sim_grid), "sim_id"),
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_tbl,
  scenarioCols = scenario_cols
)

if (isTRUE(run_plots)) {
  compare_summary |>
    dplyr::arrange(rmse) |>
    dplyr::select(
      any_of(colnames(sim_grid)),
      approach,
      method,
      n_est,
      n_threshold_fallback,
      propRespTruth_mean,
      propRespEst_mean,
      bias,
      mean_abs_error,
      rmse,
      threshold_median
    )
}

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after the canonical comparison summary.")
  knitr::knit_exit()
}

if (isTRUE(run_plots)) {
  plot_dir <- .analysis_cache_dir(analysis_key, path_root = root_dir)
}
```

## Results

### Mean absolute error by cell count

Mean absolute frequency error is plotted against stimulated cell count on a
logarithmic axis, with colour identifying the method. Facets separate response
frequencies and expression distributions; each plot uses one mean position
setting and no random condition perturbation. Lower values indicate closer
agreement with the true frequency.

```{r}
#| label: fig-mae-cell-count
#| fig-cap: "Mean absolute frequency error by stimulated cell count."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr,
        condition_perturbation_sd == 0
      )

    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = mean_abs_error,
        color = method
      )
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Mean absolute error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_mae_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Median relative error by cell count

Median absolute relative error is plotted against stimulated cell count, with
colour identifying the method and facets separating response frequencies and
expression distributions. Each plot uses one mean position setting and no random
condition perturbation. Values nearer the dashed zero line indicate smaller
errors relative to the true frequency.

```{r}
#| label: fig-med-rel-error-cell-count
#| fig-cap: "Median absolute relative frequency error by stimulated cell count."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    print(paste0("Plotting median relative error for mean position setting: ", mean_pos_setting_curr))
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, med_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = med_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::mutate(
        max_error = pmax(fbeta, stimgate, tailgate)
      ) |>
      # dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = med_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Median absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_med_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### 90th percentile relative error by cell count

The 90th percentile of absolute relative error is plotted against stimulated
cell count, with colour identifying the method. Facets separate response
frequencies and expression distributions, with one mean position setting per
plot and no random condition perturbation. Compare the upper-tail errors with
the dashed zero line.

```{r}
#| label: fig-90th-rel-error-cell-count
#| fig-cap: "90th percentile absolute relative frequency error by cell count."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, q90_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = q90_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = q90_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "90th percentile absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_q90_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### 95th percentile relative error by cell count

The 95th percentile of absolute relative error is plotted against stimulated
cell count, with colour identifying the method. Facets separate response
frequencies and expression distributions, with one mean position setting per
plot and no random condition perturbation. This shows errors further into the
upper tail; the dashed line marks zero error.

```{r}
#| label: fig-95th-rel-error-cell-count
#| fig-cap: "95th percentile absolute relative frequency error by cell count."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, q95_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = q95_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = q95_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "95th percentile absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_q95_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

```{r}
#| label: compare-relative-error-summary
#| echo: false
if (isTRUE(run_plots)) {
  compare_summary |>
    dplyr::filter(
      condition_perturbation_sd == 0
    ) |>
    dplyr::select(
      approach, max_abs_rel_error,
      transformation, prob_response, n_cell, mean_pos_setting,
      max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error
    ) |>
    dplyr::group_by(
      mean_pos_setting,
      transformation,
      prob_response,
      n_cell
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = c(max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error)
    ) |>
    dplyr::arrange(
      dplyr::desc(q95_abs_rel_error_fbeta)
    ) |>
    print(n = Inf, width = Inf)
}
```

### Max relative error by cell count

Maximum absolute relative error is plotted against stimulated cell count,
with colour identifying the method. Facets separate response frequencies and
expression distributions, with one mean position setting per plot and no random
condition perturbation. The dashed zero line provides a reference for the
largest observed errors.

```{r}
#| label: fig-max-rel-error-cell-count
#| fig-cap: "Maximum absolute relative frequency error by stimulated cell count."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, max_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = max_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = max_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Maximum absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_max_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Estimated versus true response frequency

Estimated frequency is plotted against the true background-subtracted frequency,
with colour identifying the method and facets separating expression distributions.
Each plot uses one mean position setting and pools the simulated cell counts
without random condition perturbation. Points closer to the dashed identity
line indicate better agreement.

```{r}
#| label: fig-est-vs-truth
#| fig-cap: "Estimated and true background-subtracted response frequencies."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl_focus |>
    dplyr::filter(
      is.finite(propRespTruth),
      is.finite(propRespEst)
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )

  for (mean_pos_setting_curr in unique(plot_tbl$mean_pos_setting)) {
    plot_tbl_curr <- plot_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p_est_truth <- ggplot(
      plot_tbl_curr,
      aes(
        x = propRespTruth,
        y = propRespEst,
        color = method
      )
    ) +
      geom_point(alpha = 0.35, size = 1) +
      expand_limits(
        y = c(
          0,
          max(plot_tbl_curr$propRespTruth, na.rm = TRUE) * 2
        ),
        x = c(
          0,
          max(plot_tbl_curr$propRespTruth, na.rm = TRUE) * 2
        )
      ) +
      geom_abline(
        slope = 1,
        intercept = 0,
        linetype = "dashed"
      ) +
      facet_wrap(~transformation) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "True background-subtracted response frequency",
        y = "Estimated background-subtracted response frequency",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_est_vs_truth_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p_est_truth,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Threshold fallback rate

The proportion of estimates using a threshold fallback is plotted against
stimulated cell count, with colour identifying the method. Facets separate
response frequencies and expression distributions, with one mean position
setting per plot and no random condition perturbation. Read this alongside
frequency error to distinguish calculated gates from fallbacks.

```{r}
#| label: fig-threshold-fallback
#| fig-cap: "Threshold fallback rates by stimulated cell count."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  fallback_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0) |>
    dplyr::mutate(
      fallback_rate = n_threshold_fallback / n
    )

  for (mean_pos_setting_curr in unique(fallback_tbl$mean_pos_setting)) {
    fallback_tbl_curr <- fallback_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p_fallback <- ggplot(
      fallback_tbl_curr,
      aes(
        x = n_cell,
        y = fallback_rate,
        color = method
      )
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(
        prob_response ~ transformation
      ) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Threshold fallback rate",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    .analysis_save_plot(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_threshold_fallback_rate_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p_fallback,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Threshold distributions

Histograms show the density of finite thresholds, with colour and fill
identifying the approach. Facets separate response frequencies and expression
distributions; each plot uses one mean position setting and pools cell counts
without random condition perturbation. Compare the spread and overlap of the
threshold distributions, including finite fallback thresholds.

```{r}
#| label: fig-threshold-density
#| fig-cap: "Distributions of finite thresholds from the comparison methods."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl |>
    dplyr::filter(
      condition_perturbation_sd == 0,
      is.finite(threshold)
    ) |>
    dplyr::select(
      approach, threshold,
      transformation, prob_response, n_cell, mean_pos_setting
    )

  for (mean_pos_setting_curr in unique(plot_tbl$mean_pos_setting)) {
    plot_tbl_curr <- plot_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p <- ggplot(
      plot_tbl_curr,
      aes(
        x = threshold,
        y = after_stat(density),
        fill = approach,
        colour = approach
      )
    ) +
      geom_histogram(alpha = 0.5, position = "identity") +
      facet_grid(prob_response ~ transformation, scales = "free") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Threshold",
        y = "Density",
        fill = "Method",
        colour = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    path_p <- file.path(
      plot_dir,
      paste0(
        "freq_bs_threshold_density_",
        mean_pos_setting_curr,
        ".png"
      )
    )
    .analysis_save_plot(
      filename = path_p,
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

## Rerun one scenario

A single scenario can be rerun with the production settings and RNG.
Run the setup, settings and grid chunks first.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_rerun <- 1L
row_rerun <- sim_grid_full[sim_grid_full$sim_id == sim_id_rerun, , drop = FALSE]
stopifnot(nrow(row_rerun) == 1L)
compare_one <- .simCompareRunScenario(
  row = row_rerun,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = tol_clust,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  tailgateX = tailgate_x,
  resume = FALSE
)
```
