---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
params:
  analysis_run_id: null
  simulation_seed: 1
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  1L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "corrected-comparison-v1"
analysis_dev <- isTRUE(.isDev())
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

# Tailgate failures are otherwise converted to high-threshold fallbacks by the
# comparison helper, so missing cytoUtils must fail before a comparison run.
if (
  isTRUE(run_simulations) &&
    !requireNamespace("cytoUtils", quietly = TRUE)
) {
  stop("Package 'cytoUtils' is required for tailgate comparisons.")
}
if (
  isTRUE(run_simulations) &&
    !requireNamespace("reticulate", quietly = TRUE)
) {
  stop("Package 'reticulate' is required for F-beta comparisons.")
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

# Keep scenario-level RNG stable across worker scheduling and grid chunking.
.simCompareRunScenarioUnseeded <- .simCompareRunScenario
.simCompareRunScenario <- function(row, ...) {
  if (
    "sim_seed" %in% names(row) &&
      length(row$sim_seed) > 0L &&
      !is.na(row$sim_seed[[1]])
  ) {
    set.seed(as.integer(row$sim_seed[[1]]))
  }
  .simCompareRunScenarioUnseeded(row = row, ...)
}
```

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at
   `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the `cytoUtils` package.

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but uses common histogram support spanning the paired
stimulated and unstimulated distributions. The published code derives the bin
range from the negative distribution alone, which can omit a simulated response
tail beyond the observed negative range.

For StimGate, bandwidth is estimated automatically using the ordinary `hpi1`
selector, without background normalisation, with the bandwidth-estimation sample
capped at 10,000 cells. Tailgate is deliberately left to use its usual automatic
bandwidth estimation so that it is evaluated as it would normally run.

### Simulation settings

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
condition_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1:2]
cluster_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
background_relative_to_response_vec <- c(
  2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001
)[4]
n_cell_uns_relative_to_stim_vec <- c(
  0.1, 0.2, 0.5, 1, 2, 5, 10
)[4]

# StimGate bandwidth settings.
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"
stimgate_gate_combn <- "min"

# StimGate method settings for the full method comparison.
# Cluster refinement is enabled to represent the full one-marker procedure.
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)
if (isTRUE(run_simulations) && !file.exists(path_fbeta)) {
  stop("F-beta comparison script not found: ", path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"
```

### Simulation grid

```{r}
#| label: compare-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = stimgate_bw_mtd,
  bw_ncell_max = stimgate_bw_ncell_max,
  bw_fallback = stimgate_bw_fallback,
  bw_min = stimgate_bw_min,
  bw_max = stimgate_bw_max,
  gate_combn = stimgate_gate_combn,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::mutate(
    bias_uns = dplyr::case_when(
      transformation == "gamma" ~ 0.01,
      transformation %in% c("gaussian", "skew") ~ 0.15,
      TRUE ~ NA_real_
    )
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_max,
    bw_fallback,
    bw_min,
    bw_max,
    gate_combn,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    n_cell,
    dplyr::desc(prob_response)
  ) |>
  dplyr::mutate(
    sim_seed = as.integer(simulation_seed + sim_id - 1L)
  )

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      condition_perturbation_sd == 0
    )
}

sim_grid_all <- sim_grid

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

sim_grid
print(paste0("Total simulation scenarios: ", nrow(sim_grid)))
```

A loop for debugging a single failure or inspecting intermediate objects.

```{r}
#| label: compare-loop
#| eval: false
compare_list_raw_loop <- vector("list", nrow(sim_grid))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))
  set.seed(sim_grid$sim_seed[i])

  compare_list_raw_loop[[i]] <- .simCompareFreqBs(
    nSample = 5,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 2,
    biasUns = sim_grid$bias_uns[i],
    bw = NULL,
    bwFallback = sim_grid$bw_fallback[i],
    bwMin = sim_grid$bw_min[i],
    bwMax = sim_grid$bw_max[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwNcellMax = sim_grid$bw_ncell_max[i],
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = sim_grid$sample_perturbation_sd[i],
    conditionPerturbationSd = sim_grid$condition_perturbation_sd[i],
    clusterPerturbationSd = sim_grid$cluster_perturbation_sd[i],
    backgroundRelativeToResponse =
      sim_grid$background_relative_to_response[i],
    ncellUnsRelativeToStim =
      sim_grid$n_cell_uns_relative_to_stim[i],
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = tol_clust,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol,
    tailgateX = tailgate_x
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_max = sim_grid$bw_ncell_max[i],
      bw_fallback = sim_grid$bw_fallback[i],
      bw_min = sim_grid$bw_min[i],
      bw_max = sim_grid$bw_max[i],
      bias_uns = sim_grid$bias_uns[i],
      sample_perturbation_sd = sim_grid$sample_perturbation_sd[i],
      condition_perturbation_sd =
        sim_grid$condition_perturbation_sd[i],
      cluster_perturbation_sd =
        sim_grid$cluster_perturbation_sd[i],
      background_relative_to_response =
        sim_grid$background_relative_to_response[i],
      n_cell_uns_relative_to_stim =
        sim_grid$n_cell_uns_relative_to_stim[i],
      sim_id = sim_grid$sim_id[i],
      sim_seed = sim_grid$sim_seed[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_max,
      bw_fallback,
      bw_min,
      bw_max,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id,
      sim_seed,
      dplyr::everything()
    )
}
```

Run the simulations in parallel, with progress reporting. Each scenario has a
deterministic seed derived from `sim_id`, so its simulation is stable across
worker scheduling and grid chunking.

```{r}
#| label: compare-parallel
#| results: hide
n_sample_sim <- 10
n_iter_sim <- 2
n_workers <- max(1, .simGetCores())

analysis_result_params <- list(
  comparison_semantics_version = comparison_semantics_version,
  simulation_seed = simulation_seed,
  analysis_dev = analysis_dev,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  stimgate_bw_mtd = stimgate_bw_mtd,
  stimgate_bw_ncell_max = stimgate_bw_ncell_max,
  stimgate_bw_fallback = stimgate_bw_fallback,
  stimgate_bw_min = stimgate_bw_min,
  stimgate_bw_max = stimgate_bw_max,
  stimgate_gate_combn = stimgate_gate_combn,
  tol_clust = tol_clust,
  loc_enforce_shape_threshold = loc_enforce_shape_threshold,
  calc_cyt_pos_gates = calc_cyt_pos_gates,
  fbeta_beta = fbeta_beta,
  fbeta_theta = fbeta_theta,
  fbeta_width = fbeta_width,
  tailgate_adjust = tailgate_adjust,
  tailgate_method = tailgate_method,
  tailgate_tol = tailgate_tol,
  tailgate_auto_tol = tailgate_auto_tol,
  tailgate_x = tailgate_x
)

promote_analysis7_if_ready <- function(
    run_ctx,
    sim_grid_all,
    total_sims,
    completed_sims,
    failed_sims) {
  if (!.analysis_can_promote(run_ctx)) {
    return(invisible(FALSE))
  }

  scenario_paths <- list.files(
    run_ctx$staging_run_dir,
    pattern = "^(compare_raw.*|sim_scenario.*|sim_raw.*)sim_id_[0-9]+[.]rds$",
    recursive = TRUE,
    full.names = TRUE
  )
  compare_raw_full <- .simCompareCollateScenarioOutputs(
    pathList = scenario_paths,
    sim_grid = sim_grid_all
  )
  expected_sim_ids <- sort(unique(as.integer(sim_grid_all$sim_id)))
  collated_sim_ids <- if (
    nrow(compare_raw_full) > 0L && "sim_id" %in% names(compare_raw_full)
  ) {
    sort(unique(as.integer(compare_raw_full$sim_id)))
  } else {
    integer()
  }

  full_collate_ok <-
    length(scenario_paths) == length(expected_sim_ids) &&
    identical(collated_sim_ids, expected_sim_ids)
  full_validation_ok <- full_collate_ok &&
    (!"error" %in% names(compare_raw_full) ||
      !any(!is.na(compare_raw_full$error) & compare_raw_full$error != ""))

  if (!isTRUE(full_validation_ok)) {
    error_message <- paste0(
      "Refusing to promote analysis 7: canonical collation did not contain ",
      "exactly the complete error-free simulation grid."
    )
    .analysis_mark_chunk(
      run_ctx = run_ctx,
      total_sims = total_sims,
      completed_sims = completed_sims,
      failed_sims = failed_sims,
      collate_ok = full_collate_ok,
      validation_ok = FALSE,
      error_message = error_message
    )
    stop(error_message)
  }

  path_rds_full <- file.path(run_ctx$staging_collated_dir, "compare_raw.rds")
  .write_rds_atomic(compare_raw_full, path_rds_full)
  .analysis_promote_run(run_ctx)
  invisible(TRUE)
}

compare_raw <- tibble::tibble()

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_shuffle_seed = simulation_seed,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks
      ),
      analysis_result_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )

  if (interactive()) {
    progressr::handlers(global = TRUE)
  }

  if (nrow(sim_grid) > 0L) {
    compare_raw <- .simCompareFreqBsGrid(
      sim_grid = sim_grid,
      nSample = n_sample_sim,
      nIter = n_iter_sim,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      probExact = TRUE,
      covEvMin = 1.5,
      covEvMax = 1.5,
      tolClust = tol_clust,
      locEnforceShapeThreshold = loc_enforce_shape_threshold,
      calcCytPosGates = calc_cyt_pos_gates,
      gateCombn = stimgate_gate_combn,
      pathFbeta = path_fbeta,
      fbetaBeta = fbeta_beta,
      fbetaTheta = fbeta_theta,
      fbetaWidth = fbeta_width,
      tailgateAdjust = tailgate_adjust,
      tailgateAutoTol = tailgate_auto_tol,
      tailgateMethod = tailgate_method,
      tailgateTol = tailgate_tol,
      tailgateX = tailgate_x,
      parallel = (n_workers > 1L),
      workers = n_workers,
      dirCache = dir_output,
      pathProgress = path_progress_file,
      resume = TRUE,
      sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
        sim_grid_chunk_index
      } else {
        NULL
      },
      sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
        sim_grid_n_chunks
      } else {
        NULL
      }
    )

    .write_rds_atomic(compare_raw, path_rds)
  } else {
    message("No simulations assigned to this chunk.")
  }

  expected_chunk_ids <- sort(unique(as.integer(sim_grid$sim_id)))
  collated_chunk_ids <- if (
    nrow(compare_raw) > 0L && "sim_id" %in% names(compare_raw)
  ) {
    sort(unique(as.integer(compare_raw$sim_id)))
  } else {
    integer()
  }

  error_sim_ids <- if (
    nrow(compare_raw) > 0L &&
      "sim_id" %in% names(compare_raw) &&
      "error" %in% names(compare_raw)
  ) {
    sort(unique(as.integer(
      compare_raw$sim_id[
        !is.na(compare_raw$error) &
          nzchar(as.character(compare_raw$error))
      ]
    )))
  } else {
    integer()
  }

  chunk_collate_ok <- identical(collated_chunk_ids, expected_chunk_ids)
  chunk_validation_ok <- chunk_collate_ok && length(error_sim_ids) == 0L
  n_failed <- length(intersect(expected_chunk_ids, error_sim_ids))
  n_completed <- length(setdiff(collated_chunk_ids, error_sim_ids))

  validation_error <- if (!chunk_collate_ok) {
    "Chunk collation did not contain exactly the assigned sim_id set."
  } else if (length(error_sim_ids) > 0L) {
    paste0(
      "Chunk contains simulation errors for sim_id: ",
      paste(error_sim_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_collate_ok,
    validation_ok = chunk_validation_ok,
    error_message = validation_error
  )

  if (!isTRUE(chunk_validation_ok)) {
    stop(validation_error)
  }

  promote_analysis7_if_ready(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed
  )

  if (sim_grid_n_chunks > 1L) {
    message(
      "Completed this simulation chunk. ",
      "Render with run_simulations = FALSE after all chunks are promoted ",
      "to inspect summaries or write plots."
    )
    knitr::knit_exit()
  }
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir
  )
  message("Skipping simulations; using canonical promoted results.")
}
```

### Collate results

```{r}
#| label: compare-collate
#| error: true
# Without `run_ctx` (simulation chunk not run in this session), read the
# canonical promoted results under `current/`.
if (!exists("run_ctx")) {
  run_ctx <- .analysis_results_context(analysis_key, path_root = root_dir)
}

path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = analysis_result_params
)
compare_raw <- readRDS(path_rds)

compare_tbl <- compare_raw

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(
    method %in% c(
      "stimgate",
      "fbeta",
      "tailgate"
    )
  ) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

scenario_cols <- c(
  setdiff(colnames(sim_grid), "sim_id"),
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_tbl,
  scenarioCols = scenario_cols
)

compare_summary |>
  dplyr::arrange(rmse) |>
  dplyr::select(
    any_of(colnames(sim_grid)),
    approach,
    method,
    n_est,
    n_threshold_fallback,
    propRespTruth_mean,
    propRespEst_mean,
    bias,
    mean_abs_error,
    rmse,
    threshold_median
  )

plot_dir <- projr::projr_path_get(
  "cache",
  "sim",
  "compare",
  "freq_bs"
)
if (isTRUE(run_plots)) {
  dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)
}
```

### Mean absolute error by cell count

```{r}
#| label: plot-mae-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr,
        condition_perturbation_sd == 0
      )

    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = mean_abs_error,
        color = method
      )
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Mean absolute error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_mae_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Median relative error by cell count

```{r}
#| label: plot-med-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    print(paste0("Plotting median relative error for mean position setting: ", mean_pos_setting_curr))
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, med_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = med_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::mutate(
        max_error = pmax(fbeta, stimgate, tailgate)
      ) |>
      # dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = med_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Median absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_med_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### 90th percentile relative error by cell count

```{r}
#| label: plot-90th-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, q90_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = q90_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = q90_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "90th percentile absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_q90_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### 95th percentile relative error by cell count

```{r}
#| label: plot-95th-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, q95_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = q95_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = q95_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "95th percentile absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_q95_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

```{r}
compare_summary |>
  dplyr::filter(
    condition_perturbation_sd == 0
  ) |>
  dplyr::select(
    approach, max_abs_rel_error,
    transformation, prob_response, n_cell, mean_pos_setting,
    max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error
  ) |>
  dplyr::group_by(
    mean_pos_setting,
    transformation,
    prob_response,
    n_cell
  ) |>
  tidyr::pivot_wider(
    names_from = approach,
    values_from = c(max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error)
  ) |>
  dplyr::arrange(
    dplyr::desc(q95_abs_rel_error_fbeta)
  ) |>
  print(n = Inf, width = Inf)
```

### Max relative error by cell count

```{r}
#| label: plot-max-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, max_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = max_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = max_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Maximum absolute relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_max_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Estimated versus true response frequency

```{r}
#| label: plot-est-vs-truth
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl_focus |>
    dplyr::filter(
      is.finite(propRespTruth),
      is.finite(propRespEst)
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )

  for (mean_pos_setting_curr in unique(plot_tbl$mean_pos_setting)) {
    plot_tbl_curr <- plot_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p_est_truth <- ggplot(
      plot_tbl_curr,
      aes(
        x = propRespTruth,
        y = propRespEst,
        color = method
      )
    ) +
      geom_point(alpha = 0.35, size = 1) +
      expand_limits(
        y = c(
          0,
          max(plot_tbl_curr$propRespTruth, na.rm = TRUE) * 2
        ),
        x = c(
          0,
          max(plot_tbl_curr$propRespTruth, na.rm = TRUE) * 2
        )
      ) +
      geom_abline(
        slope = 1,
        intercept = 0,
        linetype = "dashed"
      ) +
      facet_wrap(~transformation) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "True background-subtracted response frequency",
        y = "Estimated background-subtracted response frequency",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p_est_truth)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_est_vs_truth_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p_est_truth,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Threshold fallback rate

```{r}
#| label: plot-threshold-fallback
if (isTRUE(run_plots)) {
  fallback_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0) |>
    dplyr::mutate(
      fallback_rate = n_threshold_fallback / n
    )

  for (mean_pos_setting_curr in unique(fallback_tbl$mean_pos_setting)) {
    fallback_tbl_curr <- fallback_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p_fallback <- ggplot(
      fallback_tbl_curr,
      aes(
        x = n_cell,
        y = fallback_rate,
        color = method
      )
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(
        prob_response ~ transformation
      ) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Threshold fallback rate",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p_fallback)

    ggsave(
      filename = file.path(
        plot_dir,
        paste0(
          "freq_bs_threshold_fallback_rate_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p_fallback,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

## Plot the thresholds

```{r}
#| label: plot-threshold-density
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl |>
    dplyr::filter(
      condition_perturbation_sd == 0,
      is.finite(threshold)
    ) |>
    dplyr::select(
      approach, threshold,
      transformation, prob_response, n_cell, mean_pos_setting
    )

  for (mean_pos_setting_curr in unique(plot_tbl$mean_pos_setting)) {
    plot_tbl_curr <- plot_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p <- ggplot(
      plot_tbl_curr,
      aes(
        x = threshold,
        y = after_stat(density),
        fill = approach,
        colour = approach
      )
    ) +
      geom_histogram(alpha = 0.5, position = "identity") +
      facet_grid(prob_response ~ transformation, scales = "free") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Threshold",
        y = "Density",
        fill = "Method",
        colour = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    path_p <- file.path(
      plot_dir,
      paste0(
        "freq_bs_threshold_density_",
        mean_pos_setting_curr,
        ".png"
      )
    )
    ggplot2::ggsave(
      filename = path_p,
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```
