---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)

root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

path_compare_helpers <- file.path(root_dir, "R", "sim-compare-freq_bs.R")
if (file.exists(path_compare_helpers)) {
  source(path_compare_helpers)
}
```

### Aim

Compare the fixed-bandwidth StimGate background-subtracted frequency estimate against two alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the cytoUtils/openCyto helper scripts.

For now, the bandwidth is fixed only for StimGate. Tailgate is deliberately left to use its usual automatic bandwidth estimation, because the aim here is to compare StimGate with the alternative methods as they normally run.

### Simulation settings

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
condition_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1:2]
cluster_perturbation_sd_vec <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)[1]
background_relative_to_response_vec <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)[4]
n_cell_uns_relative_to_stim_vec <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)[4]

bw_fixed <- 0.5
bias_uns_vec <- c(0, 0.05, 0.1, seq(0.2, 2, by = 0.2))[c(2, 4)]

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

path_tailgate_sources <- c(
  file.path(root_dir, "scripts", "R", "openCyto-find_peaks_and_valleys.R"),
  file.path(root_dir, "scripts", "R", "cytoUtils-cytokine_cutpoint.R")
)
path_tailgate_sources <- path_tailgate_sources[file.exists(path_tailgate_sources)]
if (length(path_tailgate_sources) == 0L) {
  path_tailgate_sources <- NULL
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
```

### Simulation grid

```{r}
#| label: compare-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw = bw_fixed,
  bias_uns = bias_uns_vec,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::select(
    transformation, prob_response, n_cell,
    mean_pos_setting, mean_pos, bw, bias_uns,
    sample_perturbation_sd, condition_perturbation_sd,
    cluster_perturbation_sd, background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if (.isDev()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bias_uns == 0.05,
      condition_perturbation_sd == 0
    )
}

sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))
```

A loop for debugging a single failure or inspecting intermediate objects.

```{r}
#| label: compare-loop
#| eval: false
compare_list_raw_loop <- vector("list", nrow(sim_grid))

for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))

  compare_list_raw_loop[[i]] <- stimgate:::.simCompareFreqBs(
    nSample = 5,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 2,
    biasUns = sim_grid$bias_uns[i],
    bw = sim_grid$bw[i],
    bwFallback = sim_grid$bw[i],
    bwMin = "none",
    bwMax = "none",
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = sim_grid$sample_perturbation_sd[i],
    conditionPerturbationSd = sim_grid$condition_perturbation_sd[i],
    clusterPerturbationSd = sim_grid$cluster_perturbation_sd[i],
    backgroundRelativeToResponse = sim_grid$background_relative_to_response[i],
    ncellUnsRelativeToStim = sim_grid$n_cell_uns_relative_to_stim[i],
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = NULL,
    includeLocCondition = TRUE,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateSourceFiles = path_tailgate_sources,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = TRUE,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw = sim_grid$bw[i],
      bias_uns = sim_grid$bias_uns[i],
      sample_perturbation_sd = sim_grid$sample_perturbation_sd[i],
      condition_perturbation_sd = sim_grid$condition_perturbation_sd[i],
      cluster_perturbation_sd = sim_grid$cluster_perturbation_sd[i],
      background_relative_to_response =
        sim_grid$background_relative_to_response[i],
      n_cell_uns_relative_to_stim =
        sim_grid$n_cell_uns_relative_to_stim[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation, prob_response, n_cell,
      mean_pos_setting, mean_pos, bw, bias_uns,
      sample_perturbation_sd, condition_perturbation_sd,
      cluster_perturbation_sd, background_relative_to_response,
      n_cell_uns_relative_to_stim, sim_id,
      dplyr::everything()
    )
}
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: compare-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "compare",
  "freq_bs",
  "progress.txt"
)
dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

future::plan(
  future::multisession,
  workers = max(1, .simGetCores())
)

compare_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id
    ) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | bw = ", bw,
        " | bias = ", bias_uns,
        " | sample_sd = ", sample_perturbation_sd,
        " | condition_sd = ", condition_perturbation_sd,
        " | cluster_sd = ", cluster_perturbation_sd,
        " | bg_rel = ", background_relative_to_response,
        " | n_uns_rel = ", n_cell_uns_relative_to_stim,
        " | sim_id = ", sim_id
      )

      cat(
        paste0("Running: ", settings_row, "\n"),
        file = path_progress_file,
        append = TRUE
      )

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simCompareFreqBs(
            nSample = 5,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 5,
            biasUns = bias_uns,
            bw = bw,
            bwFallback = bw,
            bwMin = "none",
            bwMax = "none",
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            samplePerturbationSd = sample_perturbation_sd,
            conditionPerturbationSd = condition_perturbation_sd,
            clusterPerturbationSd = cluster_perturbation_sd,
            backgroundRelativeToResponse = background_relative_to_response,
            ncellUnsRelativeToStim = n_cell_uns_relative_to_stim,
            covEvMin = 1.5,
            covEvMax = 1.5,
            tolClust = NULL,
            includeLocCondition = TRUE,
            pathFbeta = path_fbeta,
            fbetaBeta = fbeta_beta,
            fbetaTheta = fbeta_theta,
            fbetaWidth = fbeta_width,
            tailgateSourceFiles = path_tailgate_sources,
            tailgateAdjust = tailgate_adjust,
            tailgateMethod = tailgate_method,
            tailgateTol = tailgate_tol
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw = bw,
              bias_uns = bias_uns,
              sample_perturbation_sd = sample_perturbation_sd,
              condition_perturbation_sd = condition_perturbation_sd,
              cluster_perturbation_sd = cluster_perturbation_sd,
              background_relative_to_response =
                background_relative_to_response,
              n_cell_uns_relative_to_stim =
                n_cell_uns_relative_to_stim,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos_setting,
              mean_pos,
              bw,
              bias_uns,
              sample_perturbation_sd,
              condition_perturbation_sd,
              cluster_perturbation_sd,
              background_relative_to_response,
              n_cell_uns_relative_to_stim,
              sim_id,
              dplyr::everything()
            )

          cat(
            paste0("Completed: ", settings_row, "\n"),
            file = path_progress_file,
            append = TRUE
          )
          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          cat(
            paste0("Error [", settings_row, "]: ", e$message, "\n"),
            file = path_progress_file,
            append = TRUE
          )
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw = bw,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response = background_relative_to_response,
            n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            approach = NA_character_,
            method = NA_character_,
            propRespTruth = NA_real_,
            propRespEst = NA_real_,
            threshold = NA_real_,
            thresholdOrigin = NA_character_,
            gateReturnPoint = NA_character_,
            nCellStim = NA_real_,
            nCellUns = NA_real_,
            nPosStim = NA_integer_,
            nPosUns = NA_integer_,
            error = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

path_compare_out <- projr::projr_path_get(
  "cache", "sim", "compare", "freq_bs", "compare_list_raw.rds"
)
dir.create(dirname(path_compare_out), recursive = TRUE, showWarnings = FALSE)
saveRDS(compare_list_raw, file = path_compare_out)
```

### Collate results

```{r}
#| label: compare-collate
#| error: true
path_vec <- c(
  projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "compare_list_raw.rds",
    safe = TRUE
  ),
  projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "compare_list_raw.rds",
    safe = FALSE
  ),
  projr::projr_path_get(
    "raw-data-archive-manual",
    "v0.103.0",
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "compare_list_raw.rds",
    safe = FALSE
  )
)
path_vec <- path_vec[file.exists(path_vec)][[1]]

compare_list_raw <- readRDS(file = path_vec)

compare_tbl <- compare_list_raw |>
  purrr::list_rbind()

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(method %in% c("stimgate_loc_sample", "fbeta", "tailgate")) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

compare_summary <- stimgate:::.simCompareSummariseFreqBs(compare_tbl)

compare_summary |>
  dplyr::arrange(rmse) |>
  dplyr::select(
    any_of(colnames(sim_grid)),
    approach,
    method,
    n_est,
    n_threshold_fallback,
    propRespTruth_mean,
    propRespEst_mean,
    bias,
    mean_abs_error,
    rmse,
    threshold_median
  )
```

### Mean absolute error by cell count

```{r}
#| label: plot-mae-cell-count
for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

  p <- ggplot(
    plot_tbl,
    aes(x = n_cell, y = mean_abs_error, color = method)
  ) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation) +
    labs(
      title = paste0("Mean position setting: ", mean_pos_setting_curr),
      x = "Number of stimulated cells",
      y = "Mean absolute error",
      color = "Method"
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    )

  print(p)

  ggsave(
    filename = projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0("freq_bs_mae_meanpos_setting_", mean_pos_setting_curr, ".png")
    ),
    plot = p,
    width = 18,
    height = 15,
    units = "cm"
  )
}
```

### Estimated versus true response frequency

```{r}
#| label: plot-est-vs-truth
plot_tbl <- compare_tbl_focus |>
  dplyr::filter(is.finite(propRespTruth), is.finite(propRespEst))

p_est_truth <- ggplot(
  plot_tbl, #|>
    #dplyr::mutate(
    #  propRespEst = pmax(0, propRespEst),
   # ),
  aes(x = propRespTruth, y = propRespEst, color = method)
) +
  geom_point(alpha = 0.35, size = 1)  +
  expand_limits(
    y = c(0, max(plot_tbl$propRespTruth, na.rm = TRUE) * 2),
    x = c(0, max(plot_tbl$propRespTruth, na.rm = TRUE) * 2)
    ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  facet_grid(mean_pos_setting ~ transformation) +
  labs(
    x = "True background-subtracted response frequency",
    y = "Estimated background-subtracted response frequency",
    color = "Method"
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "bottom"
  )

print(p_est_truth)

ggsave(
  filename = projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "freq_bs_est_vs_truth.png"
  ),
  plot = p_est_truth,
  width = 18,
  height = 15,
  units = "cm"
)
```

### Threshold fallback rate

```{r}
#| label: plot-threshold-fallback
fallback_tbl <- compare_summary |>
  dplyr::mutate(
    fallback_rate = n_threshold_fallback / n
  )

p_fallback <- ggplot(
  fallback_tbl,
  aes(x = n_cell, y = fallback_rate, color = method)
) +
  geom_line() +
  geom_point() +
  scale_x_log10() +
  facet_grid(prob_response + mean_pos_setting ~ transformation) +
  labs(
    x = "Number of stimulated cells",
    y = "Threshold fallback rate",
    color = "Method"
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "bottom"
  )

print(p_fallback)

ggsave(
  filename = projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "freq_bs_threshold_fallback_rate.png"
  ),
  plot = p_fallback,
  width = 18,
  height = 15,
  units = "cm"
)
```
