---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)

root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

# Tailgate comparisons use the cytoUtils package directly.
if (!requireNamespace("cytoUtils", quietly = TRUE)) {
  warning(
    "Package 'cytoUtils' is not installed. Tailgate comparisons will fail if run."
  )
}

scripts_r_dir <- file.path(root_dir, "scripts", "r")

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at
   `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the `cytoUtils` package.

For StimGate, bandwidth is estimated automatically using the ordinary `hpi1`
selector, without background normalisation, with the bandwidth-estimation sample
capped at 10,000 cells. Tailgate is deliberately left to use its usual automatic
bandwidth estimation so that it is evaluated as it would normally run.

### Simulation settings

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
condition_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1:2]
cluster_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
background_relative_to_response_vec <- c(
  2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001
)[4]
n_cell_uns_relative_to_stim_vec <- c(
  0.1, 0.2, 0.5, 1, 2, 5, 10
)[4]

# StimGate bandwidth settings.
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# StimGate method settings for the full method comparison.
# Cluster refinement is enabled to represent the full one-marker procedure.
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
```

### Simulation grid

```{r}
#| label: compare-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = stimgate_bw_mtd,
  bw_ncell_max = stimgate_bw_ncell_max,
  bw_fallback = stimgate_bw_fallback,
  bw_min = stimgate_bw_min,
  bw_max = stimgate_bw_max,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::mutate(
    bias_uns = dplyr::case_when(
      transformation == "gamma" ~ 0.01,
      transformation %in% c("gaussian", "skew") ~ 0.15,
      TRUE ~ NA_real_
    )
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_max,
    bw_fallback,
    bw_min,
    bw_max,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    n_cell,
    dplyr::desc(prob_response)
  )

if (.isDev()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bias_uns == 0.05,
      condition_perturbation_sd == 0
    )
}

sim_grid
print(paste0("Performing ", nrow(sim_grid), " simulations"))

set.seed(1)

sim_grid_run <- sim_grid |>
  dplyr::slice_sample(prop = 1)
```

A loop for debugging a single failure or inspecting intermediate objects.

```{r}
#| label: compare-loop
#| eval: false
compare_list_raw_loop <- vector("list", nrow(sim_grid))
source("scripts/r/sim-compare-freq_bs.R")
for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))

  compare_list_raw_loop[[i]] <- .simCompareFreqBs(
    nSample = 5,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 2,
    biasUns = sim_grid$bias_uns[i],
    bw = NULL,
    bwFallback = sim_grid$bw_fallback[i],
    bwMin = sim_grid$bw_min[i],
    bwMax = sim_grid$bw_max[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwNcellMax = sim_grid$bw_ncell_max[i],
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = sim_grid$sample_perturbation_sd[i],
    conditionPerturbationSd = sim_grid$condition_perturbation_sd[i],
    clusterPerturbationSd = sim_grid$cluster_perturbation_sd[i],
    backgroundRelativeToResponse =
      sim_grid$background_relative_to_response[i],
    ncellUnsRelativeToStim =
      sim_grid$n_cell_uns_relative_to_stim[i],
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = tol_clust,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    includeLocCondition = TRUE,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_max = sim_grid$bw_ncell_max[i],
      bw_fallback = sim_grid$bw_fallback[i],
      bw_min = sim_grid$bw_min[i],
      bw_max = sim_grid$bw_max[i],
      bias_uns = sim_grid$bias_uns[i],
      sample_perturbation_sd = sim_grid$sample_perturbation_sd[i],
      condition_perturbation_sd =
        sim_grid$condition_perturbation_sd[i],
      cluster_perturbation_sd =
        sim_grid$cluster_perturbation_sd[i],
      background_relative_to_response =
        sim_grid$background_relative_to_response[i],
      n_cell_uns_relative_to_stim =
        sim_grid$n_cell_uns_relative_to_stim[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_max,
      bw_fallback,
      bw_min,
      bw_max,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id,
      dplyr::everything()
    )
}
```

Run the simulations in parallel, with progress reporting. The execution grid is
shuffled first so that the larger and smaller cell-count scenarios are mixed
across workers. `sim_id` remains unchanged.

```{r}
#| label: compare-parallel
#| results: hide
if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "compare",
  "freq_bs",
  "progress.txt"
)
dir.create(
  dirname(path_progress_file),
  recursive = TRUE,
  showWarnings = FALSE
)

if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

future::plan(
  future::multisession,
  workers = max(1, .simGetCores())
)

compare_list_raw <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid_run))

  furrr::future_pmap(
    sim_grid_run,
    function(transformation,
             prob_response,
             n_cell,
             mean_pos_setting,
             mean_pos,
             bw_mtd,
             bw_ncell_max,
             bw_fallback,
             bw_min,
             bw_max,
             bias_uns,
             sample_perturbation_sd,
             condition_perturbation_sd,
             cluster_perturbation_sd,
             background_relative_to_response,
             n_cell_uns_relative_to_stim,
             sim_id) {
      settings_row <- paste0(
        "trans = ", transformation,
        " | prob = ", prob_response,
        " | n_cell = ", n_cell,
        " | mean = ", mean_pos,
        " | bw_mtd = ", bw_mtd,
        " | bw_ncell_max = ", bw_ncell_max,
        " | bias = ", bias_uns,
        " | sample_sd = ", sample_perturbation_sd,
        " | condition_sd = ", condition_perturbation_sd,
        " | cluster_sd = ", cluster_perturbation_sd,
        " | bg_rel = ", background_relative_to_response,
        " | n_uns_rel = ", n_cell_uns_relative_to_stim,
        " | sim_id = ", sim_id
      )

      cat(
        paste0("Running: ", settings_row, "\n"),
        file = path_progress_file,
        append = TRUE
      )

      out <- tryCatch(
        {
          sim_res <- .simCompareFreqBs(
            nSample = 10,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 2,
            biasUns = bias_uns,
            bw = NULL,
            bwFallback = bw_fallback,
            bwMin = bw_min,
            bwMax = bw_max,
            bwMtd = bw_mtd,
            bwNcellMax = bw_ncell_max,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            samplePerturbationSd = sample_perturbation_sd,
            conditionPerturbationSd = condition_perturbation_sd,
            clusterPerturbationSd = cluster_perturbation_sd,
            backgroundRelativeToResponse =
              background_relative_to_response,
            ncellUnsRelativeToStim =
              n_cell_uns_relative_to_stim,
            covEvMin = 1.5,
            covEvMax = 1.5,
            tolClust = tol_clust,
            locEnforceShapeThreshold = loc_enforce_shape_threshold,
            calcCytPosGates = calc_cyt_pos_gates,
            includeLocCondition = TRUE,
            pathFbeta = path_fbeta,
            fbetaBeta = fbeta_beta,
            fbetaTheta = fbeta_theta,
            fbetaWidth = fbeta_width,
            tailgateAdjust = tailgate_adjust,
            tailgateAutoTol = tailgate_auto_tol,
            tailgateMethod = tailgate_method,
            tailgateTol = tailgate_tol
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos_setting = mean_pos_setting,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              bw_ncell_max = bw_ncell_max,
              bw_fallback = bw_fallback,
              bw_min = bw_min,
              bw_max = bw_max,
              bias_uns = bias_uns,
              sample_perturbation_sd = sample_perturbation_sd,
              condition_perturbation_sd =
                condition_perturbation_sd,
              cluster_perturbation_sd =
                cluster_perturbation_sd,
              background_relative_to_response =
                background_relative_to_response,
              n_cell_uns_relative_to_stim =
                n_cell_uns_relative_to_stim,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos_setting,
              mean_pos,
              bw_mtd,
              bw_ncell_max,
              bw_fallback,
              bw_min,
              bw_max,
              bias_uns,
              sample_perturbation_sd,
              condition_perturbation_sd,
              cluster_perturbation_sd,
              background_relative_to_response,
              n_cell_uns_relative_to_stim,
              sim_id,
              dplyr::everything()
            )

          cat(
            paste0("Completed: ", settings_row, "\n"),
            file = path_progress_file,
            append = TRUE
          )
          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          cat(
            paste0(
              "Error [",
              settings_row,
              "]: ",
              e$message,
              "\n"
            ),
            file = path_progress_file,
            append = TRUE
          )
          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos_setting = mean_pos_setting,
            mean_pos = mean_pos,
            bw_mtd = bw_mtd,
            bw_ncell_max = bw_ncell_max,
            bw_fallback = bw_fallback,
            bw_min = bw_min,
            bw_max = bw_max,
            bias_uns = bias_uns,
            sample_perturbation_sd = sample_perturbation_sd,
            condition_perturbation_sd = condition_perturbation_sd,
            cluster_perturbation_sd = cluster_perturbation_sd,
            background_relative_to_response =
              background_relative_to_response,
            n_cell_uns_relative_to_stim =
              n_cell_uns_relative_to_stim,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            approach = NA_character_,
            method = NA_character_,
            propRespTruth = NA_real_,
            propRespEst = NA_real_,
            threshold = NA_real_,
            thresholdOrigin = NA_character_,
            gateReturnPoint = NA_character_,
            nCellStim = NA_real_,
            nCellUns = NA_real_,
            nPosStim = NA_integer_,
            nPosUns = NA_integer_,
            error = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE, chunk_size = 10L)
  )
})

future::plan(future::sequential)

path_compare_out <- projr::projr_path_get(
  "cache",
  "sim",
  "compare",
  "freq_bs",
  "compare_list_raw.rds"
)
dir.create(
  dirname(path_compare_out),
  recursive = TRUE,
  showWarnings = FALSE
)
saveRDS(compare_list_raw, file = path_compare_out)
```

### Collate results

```{r}
#| label: compare-collate
#| error: true
path_vec <- c(
  projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "compare_list_raw.rds",
    safe = TRUE
  ),
  projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "compare_list_raw.rds",
    safe = FALSE
  )
)

path_vec <- path_vec[file.exists(path_vec)][[1]]

compare_list_raw <- readRDS(file = path_vec)

compare_tbl <- compare_list_raw |>
  purrr::list_rbind()

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(
    method %in% c(
      "stimgate_loc_sample",
      "fbeta",
      "tailgate"
    )
  ) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

scenario_cols <- c(
  setdiff(colnames(sim_grid), "sim_id"),
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_tbl,
  scenarioCols = scenario_cols
)

compare_summary |>
  dplyr::arrange(rmse) |>
  dplyr::select(
    any_of(colnames(sim_grid)),
    approach,
    method,
    n_est,
    n_threshold_fallback,
    propRespTruth_mean,
    propRespEst_mean,
    bias,
    mean_abs_error,
    rmse,
    threshold_median
  )
```

### Mean absolute error by cell count

```{r}
#| label: plot-mae-cell-count
for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(
      mean_pos_setting == !!mean_pos_setting_curr
    )

  p <- ggplot(
    plot_tbl,
    aes(
      x = n_cell,
      y = mean_abs_error,
      color = method
    )
  ) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation) +
    labs(
      title = paste0(
        "Mean position setting: ",
        mean_pos_setting_curr
      ),
      x = "Number of stimulated cells",
      y = "Mean absolute error",
      color = "Method"
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(
        fill = "white",
        color = NA
      ),
      plot.background = element_rect(
        fill = "white",
        color = NA
      ),
      legend.position = "bottom"
    )

  print(p)

  ggsave(
    filename = projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0(
        "freq_bs_mae_meanpos_setting_",
        mean_pos_setting_curr,
        ".png"
      )
    ),
    plot = p,
    width = 18,
    height = 15,
    units = "cm"
  )
}
```

### Median relative error by cell count

```{r}
#| label: plot-med-rel-error-cell-count
for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(
      mean_pos_setting == !!mean_pos_setting_curr
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )
  plot_tbl |>
    dplyr::select(
      approach, med_abs_rel_error,
      transformation, prob_response, n_cell, mean_pos_setting,
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = med_abs_rel_error
    ) |>
    dplyr::select(-mean_pos_setting) |>
    dplyr::mutate(
      winner = dplyr::case_when(
        stimgate < fbeta &
          stimgate < tailgate ~ "stimgate",
        fbeta < stimgate &
          fbeta < tailgate ~ "fbeta",
        tailgate < stimgate &
          tailgate < fbeta ~ "tailgate",
        TRUE ~ "tie"
      )
    ) |>
    dplyr::mutate(
      big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
    ) |>
    dplyr::arrange(dplyr::desc(big_winner), winner) |>
    dplyr::arrange(dplyr::desc(fbeta)) |>
    print(n = Inf, width = Inf)


  p <- ggplot(
    plot_tbl,
    aes(
      x = n_cell,
      y = med_abs_rel_error,
      color = method
    )
  ) +
    geom_hline(
      yintercept = 0,
      colour = "gray25",
      linetype = "dashed"
    ) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation, scales = "free_y") +
    labs(
      title = paste0(
        "Mean position setting: ",
        mean_pos_setting_curr
      ),
      x = "Number of stimulated cells",
      y = "Median relative error",
      color = "Method"
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(
        fill = "white",
        color = NA
      ),
      plot.background = element_rect(
        fill = "white",
        color = NA
      ),
      legend.position = "bottom"
    )

  print(p)

  ggsave(
    filename = projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0(
        "freq_bs_med_rel_error_meanpos_setting_",
        mean_pos_setting_curr,
        ".png"
      )
    ),
    plot = p,
    width = 18,
    height = 15,
    units = "cm"
  )
}
```

### 90th percentile relative error by cell count

```{r}
#| label: plot-90th-rel-error-cell-count
for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(
      mean_pos_setting == !!mean_pos_setting_curr
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )
  plot_tbl |>
    dplyr::select(
      approach, q90_abs_rel_error,
      transformation, prob_response, n_cell, mean_pos_setting,
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = q90_abs_rel_error
    ) |>
    dplyr::select(-mean_pos_setting) |>
    dplyr::mutate(
      winner = dplyr::case_when(
        stimgate < fbeta &
          stimgate < tailgate ~ "stimgate",
        fbeta < stimgate &
          fbeta < tailgate ~ "fbeta",
        tailgate < stimgate &
          tailgate < fbeta ~ "tailgate",
        TRUE ~ "tie"
      )
    ) |>
    dplyr::mutate(
      big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
    ) |>
    dplyr::arrange(dplyr::desc(big_winner), winner) |>
    dplyr::arrange(dplyr::desc(fbeta)) |>
    print(n = Inf, width = Inf)


  p <- ggplot(
    plot_tbl,
    aes(
      x = n_cell,
      y = q90_abs_rel_error,
      color = method
    )
  ) +
    geom_hline(
      yintercept = 0,
      colour = "gray25",
      linetype = "dashed"
    ) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation, scales = "free_y") +
    labs(
      title = paste0(
        "Mean position setting: ",
        mean_pos_setting_curr
      ),
      x = "Number of stimulated cells",
      y = "90th percentile relative error",
      color = "Method"
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(
        fill = "white",
        color = NA
      ),
      plot.background = element_rect(
        fill = "white",
        color = NA
      ),
      legend.position = "bottom"
    )

  print(p)

  ggsave(
    filename = projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0(
        "freq_bs_q90_rel_error_meanpos_setting_",
        mean_pos_setting_curr,
        ".png"
      )
    ),
    plot = p,
    width = 18,
    height = 15,
    units = "cm"
  )
}
```

### 95th percentile relative error by cell count

```{r}
#| label: plot-95th-rel-error-cell-count
for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(
      mean_pos_setting == !!mean_pos_setting_curr
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )
  plot_tbl |>
    dplyr::select(
      approach, q95_abs_rel_error,
      transformation, prob_response, n_cell, mean_pos_setting,
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = q95_abs_rel_error
    ) |>
    dplyr::select(-mean_pos_setting) |>
    dplyr::mutate(
      winner = dplyr::case_when(
        stimgate < fbeta &
          stimgate < tailgate ~ "stimgate",
        fbeta < stimgate &
          fbeta < tailgate ~ "fbeta",
        tailgate < stimgate &
          tailgate < fbeta ~ "tailgate",
        TRUE ~ "tie"
      )
    ) |>
    dplyr::mutate(
      big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
    ) |>
    dplyr::arrange(dplyr::desc(big_winner), winner) |>
    dplyr::arrange(dplyr::desc(fbeta)) |>
    print(n = Inf, width = Inf)


  p <- ggplot(
    plot_tbl,
    aes(
      x = n_cell,
      y = q95_abs_rel_error,
      color = method
    )
  ) +
    geom_hline(
      yintercept = 0,
      colour = "gray25",
      linetype = "dashed"
    ) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation, scales = "free_y") +
    labs(
      title = paste0(
        "Mean position setting: ",
        mean_pos_setting_curr
      ),
      x = "Number of stimulated cells",
      y = "95th percentile relative error",
      color = "Method"
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(
        fill = "white",
        color = NA
      ),
      plot.background = element_rect(
        fill = "white",
        color = NA
      ),
      legend.position = "bottom"
    )

  print(p)

  ggsave(
    filename = projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0(
        "freq_bs_q95_rel_error_meanpos_setting_",
        mean_pos_setting_curr,
        ".png"
      )
    ),
    plot = p,
    width = 18,
    height = 15,
    units = "cm"
  )
}
```

```{r}
compare_summary |>
  dplyr::filter(
    condition_perturbation_sd == 0
  ) |>
  dplyr::select(
    approach, max_abs_rel_error,
    transformation, prob_response, n_cell, mean_pos_setting,
    max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error
  ) |>
  dplyr::group_by(
    mean_pos_setting,
    transformation,
    prob_response,
    n_cell
  ) |>
  tidyr::pivot_wider(
    names_from = approach,
    values_from = c(max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error)
  ) |>
  dplyr::arrange(
    dplyr::desc(q95_abs_rel_error_fbeta)
  ) |>
  print(n = Inf, width = Inf)
```

### Max relative error by cell count

```{r}
#| label: plot-max-rel-error-cell-count
for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(
      mean_pos_setting == !!mean_pos_setting_curr
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )
  plot_tbl |>
    dplyr::select(
      approach, max_abs_rel_error,
      transformation, prob_response, n_cell, mean_pos_setting,
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = max_abs_rel_error
    ) |>
    dplyr::select(-mean_pos_setting) |>
    dplyr::mutate(
      winner = dplyr::case_when(
        stimgate < fbeta &
          stimgate < tailgate ~ "stimgate",
        fbeta < stimgate &
          fbeta < tailgate ~ "fbeta",
        tailgate < stimgate &
          tailgate < fbeta ~ "tailgate",
        TRUE ~ "tie"
      )
    ) |>
    dplyr::mutate(
      big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
    ) |>
    dplyr::arrange(dplyr::desc(big_winner), winner) |>
    dplyr::arrange(dplyr::desc(fbeta)) |>
    print(n = Inf, width = Inf)


  p <- ggplot(
    plot_tbl,
    aes(
      x = n_cell,
      y = max_abs_rel_error,
      color = method
    )
  ) +
    geom_hline(
      yintercept = 0,
      colour = "gray25",
      linetype = "dashed"
    ) +
    geom_line() +
    geom_point() +
    scale_x_log10() +
    facet_grid(prob_response ~ transformation, scales = "free_y") +
    labs(
      title = paste0(
        "Mean position setting: ",
        mean_pos_setting_curr
      ),
      x = "Number of stimulated cells",
      y = "Maximum relative error",
      color = "Method"
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy") +
    theme(
      panel.background = element_rect(
        fill = "white",
        color = NA
      ),
      plot.background = element_rect(
        fill = "white",
        color = NA
      ),
      legend.position = "bottom"
    )

  print(p)

  ggsave(
    filename = projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0(
        "freq_bs_max_rel_error_meanpos_setting_",
        mean_pos_setting_curr,
        ".png"
      )
    ),
    plot = p,
    width = 18,
    height = 15,
    units = "cm"
  )
}
```

### Estimated versus true response frequency

```{r}
#| label: plot-est-vs-truth
plot_tbl <- compare_tbl_focus |>
  dplyr::filter(
    is.finite(propRespTruth),
    is.finite(propRespEst)
  ) |>
  dplyr::filter(
    condition_perturbation_sd == 0
  )

p_est_truth <- ggplot(
  plot_tbl,
  aes(
    x = propRespTruth,
    y = propRespEst,
    color = method
  )
) +
  geom_point(alpha = 0.35, size = 1) +
  expand_limits(
    y = c(
      0,
      max(plot_tbl$propRespTruth, na.rm = TRUE) * 2
    ),
    x = c(
      0,
      max(plot_tbl$propRespTruth, na.rm = TRUE) * 2
    )
  ) +
  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed"
  ) +
  facet_grid(mean_pos_setting ~ transformation) +
  labs(
    x = "True background-subtracted response frequency",
    y = "Estimated background-subtracted response frequency",
    color = "Method"
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    panel.background = element_rect(
      fill = "white",
      color = NA
    ),
    plot.background = element_rect(
      fill = "white",
      color = NA
    ),
    legend.position = "bottom"
  )

print(p_est_truth)

ggsave(
  filename = projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "freq_bs_est_vs_truth.png"
  ),
  plot = p_est_truth,
  width = 18,
  height = 15,
  units = "cm"
)
```

### Threshold fallback rate

```{r}
#| label: plot-threshold-fallback
fallback_tbl <- compare_summary |>
  dplyr::mutate(
    fallback_rate = n_threshold_fallback / n
  )

p_fallback <- ggplot(
  fallback_tbl,
  aes(
    x = n_cell,
    y = fallback_rate,
    color = method
  )
) +
  geom_line() +
  geom_point() +
  scale_x_log10() +
  facet_grid(
    prob_response + mean_pos_setting ~ transformation
  ) +
  labs(
    x = "Number of stimulated cells",
    y = "Threshold fallback rate",
    color = "Method"
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    panel.background = element_rect(
      fill = "white",
      color = NA
    ),
    plot.background = element_rect(
      fill = "white",
      color = NA
    ),
    legend.position = "bottom"
  )

print(p_fallback)

ggsave(
  filename = projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs",
    "freq_bs_threshold_fallback_rate.png"
  ),
  plot = p_fallback,
  width = 18,
  height = 15,
  units = "cm"
)
```

## Plot the thresholds

```{r}
plot_tbl <- compare_tbl |>
  dplyr::filter(
    condition_perturbation_sd == 0,
    is.finite(threshold)
  ) |>
  dplyr::select(
    approach, threshold,
    transformation, prob_response, n_cell, mean_pos_setting,
  )

p <- ggplot(
  plot_tbl,
  aes(
    x = threshold,
    fill = approach,
    colour = approach
  )
) +
  geom_histogram(alpha = 0.5) +
  facet_grid(prob_response + mean_pos_setting ~ transformation, scales = "free") +
  labs(
    x = "Threshold",
    y = "Density",
    fill = "Method",
    colour = "Method"
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy") +
  theme(
    panel.background = element_rect(
      fill = "white",
      color = NA
    ),
    plot.background = element_rect(
      fill = "white",
      color = NA
    ),
    legend.position = "bottom"
  )
path_p <- projr::projr_path_get(
  "cache",
  "sim",
  "compare",
  "freq_bs",
  "freq_bs_threshold_density.png"
)
ggplot2::ggsave(
  filename = path_p,
  plot = p,
  width = 18,
  height = 15,
  units = "cm"
)
```
