---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
  show_mcse: "on"
---

This analysis compares background-subtracted response frequencies from StimGate,
F-beta and Tailgate. Paired stimulated and unstimulated samples are simulated
with different response frequencies, cell counts and expression distributions.
The estimates are compared with the known response frequency and their threshold
fallbacks are recorded.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
analysis_qmd <- "analysis/7-sim-compare-freq_bs.qmd"

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "corrected-comparison-v17"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "draft" (current default) or "final" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("7-sim-compare-freq_bs")
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
# Monte Carlo error bars on the summary plots (see analysis-mcse.R).
mcse_mode <- .analysis_mcse_mode(.get_qmd_param_env("show_mcse", "SHOW_MCSE", "on"))
show_mcse <- mcse_mode != "off"
sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  stop("Expected the stimgate DESCRIPTION at the repository root: ", root_dir)
}

if (isTRUE(run_plots)) {
  .analysis_require_packages(c("simcyto", "ggforce"))
}

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "flowStats", "simcyto"))
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R",
  "sim-compare-freq_bs.R",
  "sim-compare-performance-plot.R",
  "sim-compare-qmd7-presentation.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

```

## Scientific settings

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at
   `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the `cytoUtils` package.

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but uses common histogram support spanning the paired
stimulated and unstimulated distributions. The published code derives the bin
range from the negative distribution alone, which can omit a simulated response
tail beyond the observed negative range.

For StimGate, bandwidth is estimated automatically using the `nrd0` selector
(Silverman's rule of thumb), without background normalisation, with the
bandwidth-estimation sample capped at 10,000 cells. Bandwidths are shared per
marker (`bwScope = "cytokine"`). StimGate's
unstimulated-sample bias is not fixed: it equals the bandwidth StimGate
estimates at the start (`biasUns = NULL` with `biasUnsFactor = 1`). Tailgate is deliberately left to use its usual automatic
bandwidth estimation so that it is evaluated as it would normally run.

### Simulation settings

Each scenario simulates 20 datasets of 20 samples in final runs and 5 datasets
of 20 samples in draft runs (`SIM_SIZE=draft`). The samples in a dataset are
gated together, with shared settings and cluster gates, so the dataset is the
unit for uncertainty. Draft results are kept separately and are for iterating,
not for reporting.

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 12, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
condition_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1:2]
cluster_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
background_relative_to_response_vec <- c(
  2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001
)[4]
n_cell_uns_relative_to_stim_vec <- c(
  0.1, 0.2, 0.5, 1, 2, 5, 10
)[4]

# StimGate bandwidth settings.
stimgate_bw_mtd <- "nrd0"
stimgate_bw_scope <- "cytokine"
# StimGate's unstimulated-sample bias equals its initial bandwidth estimate:
# with biasUns = NULL, gateStim() uses bwFallback * biasUnsFactor.
stimgate_bias_uns_factor <- 1
stimgate_bw_ncell_max <- 1e4
# Tubes with fewer cells are upsampled to this many before bandwidth selection.
stimgate_bw_ncell_min <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# StimGate method settings for the full method comparison.
# Cluster refinement is enabled to represent the full one-marker procedure.
# Share thresholds across clusters of similar samples via stimControl(clusterGates = ).
cluster_gates <- TRUE
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
# Each dataset of `n_sample_sim` samples is gated together (shared settings
# and cluster gates), so the dataset is the unit for uncertainty. Draft keeps
# the samples per dataset and uses a quarter of the datasets (at least 5).
n_sample_sim <- if (analysis_quick) 1L else 20L
n_iter_sim <- if (analysis_quick) 1L else c(final = 20L, draft = 5L)[[sim_size]]

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- file.path(
  root_dir,
  "scripts",
  "python",
  "fbeta.py"
)
if (isTRUE(run_simulations)) {
  if (!file.exists(path_fbeta)) {
    stop("F-beta comparison script not found: ", path_fbeta)
  }
  .simCompareFbetaEnvironment(pathFbeta = path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"
```

## Simulation

Method and tuning settings share the same biological-scenario seed, assigned on the full grid before execution filters or chunking. Each replicate has a seed drawn before any method runs, so later replicates remain paired despite method RNG use.

### Simulation grid

```{r}
#| label: compare-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = stimgate_bw_mtd,
  bw_scope = stimgate_bw_scope,
  bw_ncell_max = stimgate_bw_ncell_max,
  bw_ncell_min = stimgate_bw_ncell_min,
  bw_fallback = stimgate_bw_fallback,
  bw_min = stimgate_bw_min,
  bw_max = stimgate_bw_max,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::mutate(
    bias_uns = NA_real_,
    bias_uns_factor = stimgate_bias_uns_factor
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_scope,
    bw_ncell_max,
    bw_ncell_min,
    bw_fallback,
    bw_min,
    bw_max,
    bias_uns,
    bias_uns_factor,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  .simCompareQmd7SeedGrid(simulation_seed)

# Keep every seeded row available for an exact interactive rerun.
sim_grid_full <- sim_grid

# Quick mode keeps cells, response and transformation; all comparison methods remain.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(5e3, 2e4),
      prob_response %in% c(0.002, 0.2),
      transformation %in% c("gaussian", "gamma"),
      mean_pos_setting == "high",
      condition_perturbation_sd == 0
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      condition_perturbation_sd == 0
    )
}

sim_grid_all <- sim_grid
analysis_grid_spec <- sim_grid_all

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

sim_grid
print(paste0("Total simulation scenarios: ", nrow(sim_grid)))
```

Run the simulations in parallel, with progress reporting. Each scenario has a
deterministic seed recorded in `sim_seed`, so its simulation is stable
across worker scheduling and grid chunking.

```{r}
#| label: compare-parallel
#| include: false
n_workers <- max(1, .simGetCores())

analysis_result_params <- list(
  comparison_semantics_version = comparison_semantics_version,
  simulation_seed = simulation_seed,
  analysis_grid_spec = analysis_grid_spec,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  n_cell_stim_vec = n_cell_stim_vec,
  prob_response_vec = prob_response_vec,
  transformation_vec = transformation_vec,
  mean_pos_settings_tbl = mean_pos_settings_tbl,
  sample_perturbation_sd_vec = sample_perturbation_sd_vec,
  condition_perturbation_sd_vec = condition_perturbation_sd_vec,
  cluster_perturbation_sd_vec = cluster_perturbation_sd_vec,
  background_relative_to_response_vec = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim_vec = n_cell_uns_relative_to_stim_vec,
  stimgate_bias_uns_factor = stimgate_bias_uns_factor,
  prob_exact = TRUE,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  stimgate_bw_mtd = stimgate_bw_mtd,
  stimgate_bw_scope = stimgate_bw_scope,
  stimgate_bw_ncell_max = stimgate_bw_ncell_max,
  stimgate_bw_ncell_min = stimgate_bw_ncell_min,
  stimgate_bw_fallback = stimgate_bw_fallback,
  stimgate_bw_min = stimgate_bw_min,
  stimgate_bw_max = stimgate_bw_max,
  cluster_gates = cluster_gates,
  loc_enforce_shape_threshold = loc_enforce_shape_threshold,
  calc_cyt_pos_gates = calc_cyt_pos_gates,
  fbeta_beta = fbeta_beta,
  fbeta_theta = fbeta_theta,
  fbeta_width = fbeta_width,
  tailgate_adjust = tailgate_adjust,
  tailgate_method = tailgate_method,
  tailgate_tol = tailgate_tol,
  tailgate_auto_tol = tailgate_auto_tol,
  tailgate_x = tailgate_x
)


compare_raw <- tibble::tibble()

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks
      ),
      analysis_result_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )

  if (interactive()) {
    progressr::handlers(global = TRUE)
  }

  if (nrow(sim_grid) > 0L) {
    compare_raw <- .simCompareFreqBsGrid(
      sim_grid = sim_grid,
      nSample = n_sample_sim,
      nIter = n_iter_sim,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      probExact = TRUE,
      covEvMin = 1.5,
      covEvMax = 1.5,
      clusterGates = cluster_gates,
      locEnforceShapeThreshold = loc_enforce_shape_threshold,
      calcCytPosGates = calc_cyt_pos_gates,
      pathFbeta = path_fbeta,
      fbetaBeta = fbeta_beta,
      fbetaTheta = fbeta_theta,
      fbetaWidth = fbeta_width,
      tailgateAdjust = tailgate_adjust,
      tailgateAutoTol = tailgate_auto_tol,
      tailgateMethod = tailgate_method,
      tailgateTol = tailgate_tol,
      tailgateX = tailgate_x,
      parallel = (n_workers > 1L),
      workers = n_workers,
      dirCache = dir_output,
      pathProgress = path_progress_file,
      dirJobs = run_ctx$chunk_jobs_dir,
      progressHeading = "METHOD COMPARISON SIMULATION PROGRESS",
      resume = TRUE,
      retryErrors = TRUE,
      sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
        sim_grid_chunk_index
      } else {
        NULL
      },
      sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
        sim_grid_n_chunks
      } else {
        NULL
      }
    )

    .write_rds_atomic(compare_raw, path_rds)
  } else {
    message("No simulations assigned to this chunk.")
  }

  chunk_check <- .simCompareGridOutputStatus(
    compare_raw,
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )
  chunk_collate_ok <- isTRUE(chunk_check$collate_ok)
  chunk_validation_ok <- isTRUE(chunk_check$validation_ok)
  n_failed <- length(chunk_check$failed_ids)
  n_completed <- length(chunk_check$completed_ids)

  validation_error <- if (!chunk_collate_ok) {
    paste0(
      "Chunk collation does not match the assigned simulation grid. Missing sim_id: ",
      paste(chunk_check$missing_ids, collapse = ", "),
      "; extra sim_id: ",
      paste(chunk_check$extra_ids, collapse = ", "),
      "."
    )
  } else if (length(chunk_check$failed_ids) > 0L) {
    paste0(
      "Chunk contains incomplete, non-finite or errored primary method results for sim_id: ",
      paste(chunk_check$failed_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_collate_ok,
    validation_ok = chunk_validation_ok,
    error_message = validation_error
  )

  if (!isTRUE(chunk_validation_ok)) {
    stop(validation_error)
  }

  .simComparePromoteIfReady(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )

  if (sim_grid_n_chunks > 1L) {
    message(
      "Completed this simulation chunk. ",
      "Render with run_simulations = FALSE after all chunks are promoted ",
      "to inspect summaries or write plots."
    )
    knitr::knit_exit()
  }
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
  message("Skipping simulations; using canonical promoted results.")
}
```

## Collation

```{r}
#| label: compare-collate
#| echo: false
#| results: asis
# Without `run_ctx` (simulation chunk not run in this session), read the
# canonical promoted results under `current/`.
if (!exists("run_ctx")) {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}

path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = analysis_result_params
)
compare_raw <- .analysis_read_rds(path_rds, analysis_key, analysis_qmd)

compare_tbl <- compare_raw

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(
    method %in% c(
      "stimgate",
      "fbeta",
      "tailgate"
    )
  ) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

scenario_cols <- c(
  setdiff(colnames(sim_grid), "sim_id"),
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_tbl,
  scenarioCols = scenario_cols,
  mcse = show_mcse && isTRUE(run_plots)
)

if (isTRUE(run_plots)) {
  compare_summary |>
    dplyr::arrange(rmse) |>
    dplyr::select(
      any_of(colnames(sim_grid)),
      approach,
      method,
      n_est,
      n_run_error,
      n_no_cutpoint,
      n_threshold_fallback,
      propRespTruth_mean,
      propRespEst_mean,
      bias,
      mean_abs_error,
      rmse,
      threshold_median
    ) |>
    .analysis_report_table(c(fig_key, paste0("mcse_", mcse_mode), "comparison-summary.csv"),
      "Frequency-error and threshold summaries by scenario and method.", path_root = root_dir)
}

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after the canonical comparison summary.")
  knitr::knit_exit()
}

if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(fig_key, path_root = root_dir)
}
```

## Results

Pooled means and quantiles below describe tube-level distributions across
simulated datasets. Tubes within a dataset share gating settings and are not
independent replicates; uncertainty for method comparisons uses datasets.

Every comparison figure is shown twice. "All methods" includes StimGate,
Tailgate and F-beta. "Without Tailgate" leaves Tailgate out, because it
performs very poorly without more tuning and squashes the scale for the other
two methods. The two versions are saved in the `all_methods` and `no_tailgate`
subfolders of each figure folder. Within each version there is one figure per
mean position setting.

**Monte Carlo intervals.** `show_mcse` / `SHOW_MCSE` accepts `off` or `on`
(default); false/true still selects off/on. Each render prints one version
and saves it under `mcse_off/` or `mcse_on/`, with identical points and axes. With error bars on, 95% percentile-bootstrap
intervals resample whole independent datasets (`iter`), retaining all samples
of each selected dataset with multiplicity. Each draw recalculates the actual
pooled sample statistic shown, rather than a mean of within-dataset percentiles.
Point estimates are identical with intervals on or off. At least five
contributing datasets and 95% finite bootstrap draws are required; unavailable
intervals and their coverage are reported. Final runs use 20 independent datasets
of 20 samples each; draft runs use five datasets of 20 samples and give rougher
intervals. Tiny quick-mode datasets are smoke checks only. Maxima are reported
separately through dataset occurrence and conditional severity.

```{r}
#| label: pooled-bootstrap-coverage
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .analysis_report_table(.simComparePerformanceCoverage(compare_summary, scenario_cols),
    c(fig_key, paste0("mcse_", mcse_mode), "pooled-bootstrap-coverage.csv"),
    "Pooled-statistic bootstrap coverage and unavailable intervals by scenario and method.", path_root = root_dir)
}
```

### Paired differences between datasets

The independent unit is the simulated dataset (`iter`), whose tubes are gated
with shared settings and cluster gates. Within each scenario and dataset, we
compute each method's mean absolute frequency error and mean absolute relative
frequency error, then subtract the competitor's value from
StimGate's value. Each dataset has equal weight. The tables report the mean
paired difference and a 95% interval, mean ± 1.96 SD / √D, using the spread
between the D datasets with finite paired summaries. Intervals are NA for D < 5.
Dataset-pair coverage and finite-tube coverage are reported explicitly;
undefined outcomes are omitted from the corresponding method summary and a
dataset with no finite outcomes has no paired difference.
Negative differences favour StimGate for both errors.

```{r}
#| label: dataset-paired-differences
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  dataset_differences <- .simCompareDatasetDifferences(
    compare_tbl, scenarioCols = scenario_cols,
    outcomes = c("abs_error", "abs_rel_error")
  )
  for (set in .simCompareMethodSets()) {
    .analysis_heading(set$heading, 4L)
    for (pos in unique(dataset_differences$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", pos), 5L)
      table <- dataset_differences |>
        dplyr::filter(.data$method %in% set$methods, .data$mean_pos_setting == pos) |>
        dplyr::mutate(
          interval_95 = sprintf("[%.4f, %.4f]", .data$lower, .data$upper),
          paired_datasets = paste0(.data$n_pair, " / ", .data$n_dataset)
        ) |>
        dplyr::select(dplyr::all_of(c(
          "transformation", "prob_response", "n_cell", "condition_perturbation_sd",
          "outcome", "method", "paired_datasets", "n_pair", "n_dataset", "mean_difference", "interval_95",
          "tube_coverage_stimgate", "tube_coverage_competitor"
        )))
      .analysis_report_table(table,
        c(fig_key, paste0("mcse_", mcse_mode), "paired-differences", paste0(paste(set$methods, collapse = "-"), "-", pos, ".csv")),
        "Dataset-level paired differences: StimGate minus competitor.", path_root = root_dir)
      cat("\n\n")
    }
  }
}
```

### Mean absolute error by cell count

Mean absolute frequency error is plotted against stimulated cell count on a
logarithmic axis, with colour identifying the method. Panels separate response
frequencies (rows) and expression distributions (columns). Each plot uses one
mean position setting and no random condition perturbation. Lower values
indicate closer agreement with the true frequency.

```{r}
#| label: mae-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "mean_abs_error", "Tube-level mean absolute error", percent_y = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "mae_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_mae_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### Median relative error by cell count

Median absolute relative error is plotted against stimulated cell count, with
colour identifying the method and panels separating response frequencies and
expression distributions. Each plot uses one mean position setting and no random
condition perturbation. Values nearer the dashed zero line indicate smaller
errors relative to the true frequency.

```{r}
#| label: med-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "med_abs_rel_error", "Tube-level median absolute relative error",
        zero_line = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "median_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_med_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### 90th percentile relative error by cell count

The 90th percentile of absolute relative error is plotted against stimulated
cell count, with colour identifying the method. Panels separate response
frequencies and expression distributions, with one mean position setting per
plot and no random condition perturbation. Compare the upper-tail errors with
the dashed zero line.

```{r}
#| label: q90-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "q90_abs_rel_error", "Tube-level 90th percentile absolute relative error",
        zero_line = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "q90_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_q90_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### 95th percentile relative error by cell count

The 95th percentile of absolute relative error is plotted against stimulated
cell count, with colour identifying the method. Panels separate response
frequencies and expression distributions, with one mean position setting per
plot and no random condition perturbation. This shows errors further into the
upper tail; the dashed line marks zero error.

```{r}
#| label: q95-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "q95_abs_rel_error", "Tube-level 95th percentile absolute relative error",
        zero_line = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "q95_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_q95_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

```{r}
#| label: compare-relative-error-summary
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  compare_summary |>
    dplyr::filter(
      condition_perturbation_sd == 0
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    dplyr::select(
      approach,
      transformation, prob_response, n_cell, mean_pos_setting,
      med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error
    ) |>
    dplyr::group_by(
      mean_pos_setting,
      transformation,
      prob_response,
      n_cell
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = c(med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error)
    ) |>
    dplyr::arrange(
      dplyr::desc(q95_abs_rel_error_fbeta)
    ) |>
    .analysis_report_table(c(fig_key, paste0("mcse_", mcse_mode), "relative-error-summary.csv"),
      "Relative-error statistics for each method and scenario.", path_root = root_dir)
}
```

### Unconditional percentiles of signed relative error by cell count

These are the main signed-error performance plots. Every finite error contributes, including exact agreement at zero; failed or undefined estimates remain in the separate coverage summaries.

The seven curves summarise the full signed relative error distribution, including zero errors, rather than conditioning on over- or under-estimation. They show the 2.5th, 10th, 25th, 50th, 75th, 90th and 97.5th percentiles together. If either outer percentile lacks an interval at any finite point, the whole figure uses the 5th/95th pair and labels that fallback; unavailable intervals are omitted without dropping points. The median is darkest and thickest; symmetric percentile pairs share progressively lighter, thinner tiers. Percentiles are calculated before applying the signed display scale.

Intervals resample whole independent datasets (`iter`), recomputing pooled tube percentiles. Scenario averages bootstrap the complete equal-weight average with shared draws across methods and deterministic mismatch settings; bootstrap validity is reported beside the figures.

```{r}
#| label: signed-percentiles-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  percentile_tbl <- compare_tbl |>
    dplyr::filter(condition_perturbation_sd == 0) |>
    .simCompareSignedPercentileSummary(
      scenarioCols = scenario_cols, mcse = show_mcse
    )
  .analysis_report_table(.simComparePerformanceCoverage(percentile_tbl, scenario_cols),
    c(fig_key, paste0("mcse_", mcse_mode), "signed-percentiles-cell-count.csv"),
    "Unconditional pooled-percentile coverage, including undefined bootstrap draws.", path_root = root_dir)
  .simCompareFigureLoop(
    percentile_tbl,
    make_plot = function(d) .simBandwidthSignedPercentilePlot(d, x = "n_cell",
      x_label = "Number of stimulated cells", x_log = TRUE, mcse = show_mcse) +
      ggplot2::aes(shape = method, linetype = method) +
      .analysis_scale_method(c("colour", "shape", "linetype")),
    dir = file.path(plot_dir, "signed_percentiles_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0(
        "freq_bs_signed_percentiles_meanpos_setting_", pos, "_prob_response_",
        gsub("[^0-9a-z]+", "_", .analysis_label_percent(extra)), ".png"
      )
    },
    height = 14,
    level = 4L,
    extra_col = "prob_response",
    extra_heading = function(x) {
      paste0("Response probability: ", .analysis_label_percent(x))
    }, mcse_mode = mcse_mode
  )
}
```

### Dataset-maximum occurrence and conditional severity

These figures answer two separate questions for each direction: how often does
an eligible dataset contain an over- or under-estimate, and how large is its
maximum error on average when that direction occurs? An eligible dataset has
valid relative errors for all 20 intended samples; incomplete datasets remain
in the coverage table. Methods can have different eligible cohorts, so compare
those counts before comparing curves. Tiny quick-mode smoke checks use their
single intended sample instead and do not provide reportable intervals.

Unaffected datasets count towards occurrence, but not conditional severity.
With no affected dataset occurrence is zero and severity is undefined. All
datasets are resampled together, including unaffected and incomplete ones.
Occurrence intervals require at least five eligible datasets and 95% finite
bootstrap draws; severity also needs at least five affected datasets. The table
reports undefined-bootstrap coverage and suppressed intervals. Conditional mean
dataset maxima need not exceed pooled sample percentiles.

```{r}
#| label: dataset-max-coverage
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  dataset_max <- .simCompareDatasetMaxSummary(
    compare_tbl |> dplyr::filter(condition_perturbation_sd == 0),
    scenarioCols = scenario_cols, expected_samples = n_sample_sim,
    expected_datasets = seq_len(n_iter_sim), mcse = show_mcse
  )
  .analysis_report_table(.simCompareDatasetMaxCoverage(dataset_max),
    c(fig_key, paste0("mcse_", mcse_mode), "dataset-max-coverage.csv"),
    "Method-specific eligible, incomplete and affected dataset cohorts; bootstrap coverage.", path_root = root_dir)
}
```

```{r}
#| label: dataset-max-occurrence
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .simCompareFigureLoop(dataset_max,
    make_plot = function(d) .simCompareDatasetMaxPlot(d, "occurrence", by_prob = FALSE, mcse = show_mcse),
    dir = file.path(plot_dir, "dataset_max_occurrence_by_n_cell"),
    file_fn = function(pos, extra) paste0("dataset_max_occurrence_", pos,
      "_prob_response_", extra, ".png"),
    extra_col = "prob_response",
    extra_heading = function(x) paste0("Response probability: ", .analysis_label_percent(x)),
    height = 14, level = 4L, mcse_mode = mcse_mode)
}
```

### Conditional dataset maxima: signed errors and ratio companions

Over-estimation severity appears above zero; under-estimation severity appears
below zero. The ratio companion uses the identical curves and intervals, with
1x for agreement and 0x for a zero estimate.

```{r}
#| label: dataset-max-severity
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .simCompareFigureLoop(dataset_max,
    make_plot = function(d) .simCompareDatasetMaxPlot(d, "severity", by_prob = FALSE, mcse = show_mcse),
    dir = file.path(plot_dir, "dataset_max_signed_error_severity_by_n_cell"),
    file_fn = function(pos, extra) paste0("dataset_max_severity_", pos,
      "_prob_response_", extra, ".png"),
    extra_col = "prob_response",
    extra_heading = function(x) paste0("Response probability: ", .analysis_label_percent(x)),
    ratio_twins = TRUE, height = 14, level = 4L, mcse_mode = mcse_mode)
}
```

Each signed-error figure also has an estimate/reference ratio companion in a
sibling `ratio*` folder. The curves and intervals are identical: 1x is exact,
0x is an estimate of zero and 2x is twice the reference. Negative ratios remain
visible; a capped upper tick reads “≥ 16x”.

### Conditional over- and under-estimates by cell count

These are conditional error-severity diagnostics: exact agreement (zero error) is excluded from both directions. The direction shares use all finite errors, including zeros, as their denominator; the remaining share is exact agreement. An all-exact group has no directional curve. Use the unconditional signed-error percentiles for overall performance, and the coverage summaries for failed or undefined estimates.

The relative error, (estimate - truth) / truth, is summarised separately for
over-estimates and under-estimates in each scenario, using the same simulated
rows as the figures above. Over-estimates are drawn above zero and
under-estimates below. Rows of panels show the median and 95th percentile, and columns separate expression distributions. Colour identifies the
method, and a thicker line means a larger share of the estimates lies in that
direction. The vertical scale is linear, including errors below -100% for negative background-subtracted estimates, with -100% (0x, an estimate of zero) and
doubles at each step above, so -100% and a two-fold over-estimate are equally
far from zero. Each plot uses one mean position setting, one response frequency
and no random condition perturbation. Whole-dataset bootstrap intervals
recalculate each direction's pooled percentile and retain random changes in
the number of errors in that direction. At least five contributing datasets
and 95% finite bootstrap draws are required; rare directions can leave an
interval unavailable. Line weight shows the share of valid sample estimates
in each direction; exact agreement belongs to neither direction.

```{r}
#| label: signed-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  signed_tbl <- compare_tbl |>
    dplyr::filter(condition_perturbation_sd == 0) |>
    .simCompareSignedErrorSummary(
      scenarioCols = scenario_cols, mcse = show_mcse
    )
  .analysis_report_table(.simComparePerformanceCoverage(signed_tbl, scenario_cols),
    c(fig_key, paste0("mcse_", mcse_mode), "signed-error-cell-count.csv"),
    "Directional pooled-percentile coverage, including undefined bootstrap draws.", path_root = root_dir)
  .simCompareFigureLoop(
    signed_tbl,
    make_plot = function(d) .simComparePlotSignedError(d,
      stat_cols = c(median = "Median", q95 = "95th percentile"), mcse = show_mcse),
    dir = file.path(plot_dir, "signed_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0(
        "freq_bs_signed_error_meanpos_setting_", pos, "_prob_response_",
        gsub("[^0-9a-z]+", "_", .analysis_label_percent(extra)), ".png"
      )
    },
    ratio_twins = TRUE,
    height = 14,
    level = 4L,
    extra_col = "prob_response",
    extra_heading = function(x) {
      paste0("Response probability: ", .analysis_label_percent(x))
    }, mcse_mode = mcse_mode
  )
}
```


Each dataset-maximum figure uses one response probability. Severity panels have
independent vertical ranges; occurrence panels retain the fixed 0–100% scale.

Error-by-cell-count and signed-error figures also use independent panel ranges;
compare methods within each panel. Method colour, point shape and line type agree
across comparison figures.

### Estimated versus true response frequency

Response levels have evenly spaced positions labelled by their simulated
percentages; methods have centred offsets and sina spreads within each level.
Dashed horizontal segments mark the true response frequency. The estimated
frequency uses a log10 percent axis. Zero and negative estimates are placed at
the labelled lower plotting limit, rather than discarded. Each plot pools cell
counts at one mean position, without condition perturbation.

```{r}
#| label: est-vs-truth
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl_focus |>
    dplyr::filter(
      is.finite(propRespTruth),
      is.finite(propRespEst)
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) .simComparePlotEstVsTruth(d),
    dir = file.path(plot_dir, "estimated_vs_true_frequency"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_est_vs_truth_", pos, ".png")
    },
    height = 9,
    level = 4L,
    after_plot = function(d, p) {
      .analysis_report_table(.simCompareQmd7FloorTable(d, unique(p$data$plot_floor)),
        c(fig_key, paste0("mcse_", mcse_mode), "estimated_vs_true_frequency", paste0(paste(unique(d$method), collapse = "-"), "-", unique(d$mean_pos_setting), ".csv")),
        "Estimates at or below the plotting floor; denominators count method/sample estimates.", path_root = root_dir)
      cat("\n\n")
    }
  )
}
```

### Threshold fallback rate

The table includes both condition-perturbation settings and distinguishes
threshold fallbacks, run errors and genuine no-cutpoint outcomes. Counts and
percentages use all intended method/sample outcomes as their denominator,
including failed runs. Expression distributions and mean positions remain
separate so failures in a particular biological setting stay visible; response
levels and cell counts are pooled for this coverage table. No-cutpoint outcomes
can also be threshold fallbacks, so these columns are not disjoint.

```{r}
#| label: threshold-fallback
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  .analysis_report_table(.simCompareQmd7FallbackTable(compare_summary),
    c(fig_key, paste0("mcse_", mcse_mode), "threshold-fallback.csv"),
    "Threshold provenance: count / all intended samples (percentage).", path_root = root_dir)
}
```

### Threshold distributions

Light histograms show finite threshold densities in method colours, including
finite fallbacks, on the ordinary density axis. Grey/black lines show the
shape of the unstimulated/stimulated sampled expression densities, scaled so
the highest reference peak in each panel matches the tallest histogram bar. Their
heights therefore show shape only, not density values. Because the responding
cells are often too rare to see in these densities, a dotted vertical line marks
the mean of the stimulated reference tube's positive-component cells. These seeded reference simulations use the same biological settings
(`probExact = TRUE`, covariance eigenvalues 1.5), with 100,000 cells per tube;
they are render-local references, rather than the cells from the gated datasets.
Both tubes include all sampled components. Each biological reference is computed
once and reused across method sets and cell counts. Panels have independent x
and y limits; plots pool cell counts without condition perturbation.

```{r}
#| label: threshold-density
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl |>
    dplyr::filter(
      condition_perturbation_sd == 0,
      is.finite(threshold)
    ) |>
    dplyr::select(
      approach, threshold,
      transformation, prob_response, n_cell, mean_pos_setting, mean_pos,
      sample_perturbation_sd, condition_perturbation_sd,
      cluster_perturbation_sd, background_relative_to_response
    )
  threshold_reference_densities <- .simBandwidthThresholdDensities(
    plot_tbl,
    settings = list(probExact = TRUE, covEvMin = 1.5, covEvMax = 1.5),
    seed = 271L, unstimulated_negative_only = FALSE
  )
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) .simComparePlotThresholdDensity(d, densities = threshold_reference_densities),
    dir = file.path(plot_dir, "threshold_distributions"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_threshold_density_", pos, ".png")
    },
    height = 20,
    level = 4L,
    method_col = "approach"
  )
}
```

## Rerun one scenario

A single scenario can be rerun with the production settings and RNG.
Run the setup, settings and grid chunks first.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_rerun <- 1L
row_rerun <- sim_grid_full[sim_grid_full$sim_id == sim_id_rerun, , drop = FALSE]
stopifnot(nrow(row_rerun) == 1L)
compare_one <- .simCompareRunScenario(
  row = row_rerun,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  clusterGates = cluster_gates,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  tailgateX = tailgate_x,
  resume = FALSE
)
```
