---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
  show_mcse: true
---

This analysis compares background-subtracted response frequencies from StimGate,
F-beta and Tailgate. Paired stimulated and unstimulated samples are simulated
with different response frequencies, cell counts and expression distributions.
The estimates are compared with the known response frequency and their threshold
fallbacks are recorded.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
analysis_qmd <- "analysis/7-sim-compare-freq_bs.qmd"

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "corrected-comparison-v6"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("7-sim-compare-freq_bs")
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
# Monte Carlo error bars on the summary plots (see analysis-mcse.R).
show_mcse <- .as_flag(.get_qmd_param_env("show_mcse", "SHOW_MCSE", TRUE))
sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  stop("Expected the stimgate DESCRIPTION at the repository root: ", root_dir)
}

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "flowStats", "simcyto"))
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

```

## Scientific settings

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at
   `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the `cytoUtils` package.

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but uses common histogram support spanning the paired
stimulated and unstimulated distributions. The published code derives the bin
range from the negative distribution alone, which can omit a simulated response
tail beyond the observed negative range.

For StimGate, bandwidth is estimated automatically using the `nrd0` selector
(Silverman's rule of thumb), without background normalisation, with the
bandwidth-estimation sample capped at 10,000 cells. StimGate's
unstimulated-sample bias is not fixed: it equals the bandwidth StimGate
estimates at the start (`biasUns = NULL` with `biasUnsFactor = 4`). Tailgate is deliberately left to use its usual automatic
bandwidth estimation so that it is evaluated as it would normally run.

### Simulation settings

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
condition_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1:2]
cluster_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
background_relative_to_response_vec <- c(
  2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001
)[4]
n_cell_uns_relative_to_stim_vec <- c(
  0.1, 0.2, 0.5, 1, 2, 5, 10
)[4]

# StimGate bandwidth settings.
stimgate_bw_mtd <- "nrd0"
# StimGate's unstimulated-sample bias equals its initial bandwidth estimate:
# with biasUns = NULL, gateStim() uses 0.25 * bwFallback * biasUnsFactor.
stimgate_bias_uns_factor <- 4
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# StimGate method settings for the full method comparison.
# Cluster refinement is enabled to represent the full one-marker procedure.
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
n_sample_sim <- if (analysis_quick) 1L else 10
n_iter_sim <- if (analysis_quick) 1L else 2

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- file.path(
  root_dir,
  "scripts",
  "python",
  "fbeta.py"
)
if (isTRUE(run_simulations)) {
  if (!file.exists(path_fbeta)) {
    stop("F-beta comparison script not found: ", path_fbeta)
  }
  .simCompareFbetaEnvironment(pathFbeta = path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"
```

## Simulation

### Simulation grid

```{r}
#| label: compare-grid
#| include: false
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = stimgate_bw_mtd,
  bw_ncell_max = stimgate_bw_ncell_max,
  bw_fallback = stimgate_bw_fallback,
  bw_min = stimgate_bw_min,
  bw_max = stimgate_bw_max,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::mutate(
    bias_uns = NA_real_,
    bias_uns_factor = stimgate_bias_uns_factor
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_max,
    bw_fallback,
    bw_min,
    bw_max,
    bias_uns,
    bias_uns_factor,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    n_cell,
    dplyr::desc(prob_response)
  ) |>
  dplyr::mutate(
    sim_seed = as.integer(simulation_seed + sim_id - 1L)
  )

# Keep every seeded row available for an exact interactive rerun.
sim_grid_full <- sim_grid

# Quick mode keeps cells, response and transformation; all comparison methods remain.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell %in% c(5e3, 2e4),
      prob_response %in% c(0.002, 0.2),
      transformation %in% c("gaussian", "gamma"),
      mean_pos_setting == "high",
      condition_perturbation_sd == 0
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      condition_perturbation_sd == 0
    )
}

sim_grid_all <- sim_grid

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

sim_grid
print(paste0("Total simulation scenarios: ", nrow(sim_grid)))
```

Run the simulations in parallel, with progress reporting. Each scenario has a
deterministic seed `simulation_seed + sim_id - 1`, so its simulation is stable
across worker scheduling and grid chunking.

```{r}
#| label: compare-parallel
#| include: false
n_workers <- max(1, .simGetCores())

analysis_result_params <- list(
  comparison_semantics_version = comparison_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  n_cell_stim_vec = n_cell_stim_vec,
  prob_response_vec = prob_response_vec,
  transformation_vec = transformation_vec,
  mean_pos_settings_tbl = mean_pos_settings_tbl,
  sample_perturbation_sd_vec = sample_perturbation_sd_vec,
  condition_perturbation_sd_vec = condition_perturbation_sd_vec,
  cluster_perturbation_sd_vec = cluster_perturbation_sd_vec,
  background_relative_to_response_vec = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim_vec = n_cell_uns_relative_to_stim_vec,
  stimgate_bias_uns_factor = stimgate_bias_uns_factor,
  prob_exact = TRUE,
  cov_ev_min = 1.5,
  cov_ev_max = 1.5,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  stimgate_bw_mtd = stimgate_bw_mtd,
  stimgate_bw_ncell_max = stimgate_bw_ncell_max,
  stimgate_bw_fallback = stimgate_bw_fallback,
  stimgate_bw_min = stimgate_bw_min,
  stimgate_bw_max = stimgate_bw_max,
  tol_clust = tol_clust,
  loc_enforce_shape_threshold = loc_enforce_shape_threshold,
  calc_cyt_pos_gates = calc_cyt_pos_gates,
  fbeta_beta = fbeta_beta,
  fbeta_theta = fbeta_theta,
  fbeta_width = fbeta_width,
  tailgate_adjust = tailgate_adjust,
  tailgate_method = tailgate_method,
  tailgate_tol = tailgate_tol,
  tailgate_auto_tol = tailgate_auto_tol,
  tailgate_x = tailgate_x
)


compare_raw <- tibble::tibble()

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks
      ),
      analysis_result_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )

  if (interactive()) {
    progressr::handlers(global = TRUE)
  }

  if (nrow(sim_grid) > 0L) {
    compare_raw <- .simCompareFreqBsGrid(
      sim_grid = sim_grid,
      nSample = n_sample_sim,
      nIter = n_iter_sim,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      probExact = TRUE,
      covEvMin = 1.5,
      covEvMax = 1.5,
      tolClust = tol_clust,
      locEnforceShapeThreshold = loc_enforce_shape_threshold,
      calcCytPosGates = calc_cyt_pos_gates,
      pathFbeta = path_fbeta,
      fbetaBeta = fbeta_beta,
      fbetaTheta = fbeta_theta,
      fbetaWidth = fbeta_width,
      tailgateAdjust = tailgate_adjust,
      tailgateAutoTol = tailgate_auto_tol,
      tailgateMethod = tailgate_method,
      tailgateTol = tailgate_tol,
      tailgateX = tailgate_x,
      parallel = (n_workers > 1L),
      workers = n_workers,
      dirCache = dir_output,
      pathProgress = path_progress_file,
      dirJobs = run_ctx$chunk_jobs_dir,
      progressHeading = "METHOD COMPARISON SIMULATION PROGRESS",
      resume = TRUE,
      retryErrors = TRUE,
      sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
        sim_grid_chunk_index
      } else {
        NULL
      },
      sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
        sim_grid_n_chunks
      } else {
        NULL
      }
    )

    .write_rds_atomic(compare_raw, path_rds)
  } else {
    message("No simulations assigned to this chunk.")
  }

  chunk_check <- .simCompareGridOutputStatus(
    compare_raw,
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )
  chunk_collate_ok <- isTRUE(chunk_check$collate_ok)
  chunk_validation_ok <- isTRUE(chunk_check$validation_ok)
  n_failed <- length(chunk_check$failed_ids)
  n_completed <- length(chunk_check$completed_ids)

  validation_error <- if (!chunk_collate_ok) {
    paste0(
      "Chunk collation does not match the assigned simulation grid. Missing sim_id: ",
      paste(chunk_check$missing_ids, collapse = ", "),
      "; extra sim_id: ",
      paste(chunk_check$extra_ids, collapse = ", "),
      "."
    )
  } else if (length(chunk_check$failed_ids) > 0L) {
    paste0(
      "Chunk contains incomplete, non-finite or errored primary method results for sim_id: ",
      paste(chunk_check$failed_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_collate_ok,
    validation_ok = chunk_validation_ok,
    error_message = validation_error
  )

  if (!isTRUE(chunk_validation_ok)) {
    stop(validation_error)
  }

  .simComparePromoteIfReady(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )

  if (sim_grid_n_chunks > 1L) {
    message(
      "Completed this simulation chunk. ",
      "Render with run_simulations = FALSE after all chunks are promoted ",
      "to inspect summaries or write plots."
    )
    knitr::knit_exit()
  }
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
  message("Skipping simulations; using canonical promoted results.")
}
```

## Collation

```{r}
#| label: compare-collate
#| echo: false
# Without `run_ctx` (simulation chunk not run in this session), read the
# canonical promoted results under `current/`.
if (!exists("run_ctx")) {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key, path_root = root_dir, qmd_path = analysis_qmd
  )
}

path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = analysis_result_params
)
compare_raw <- .analysis_read_rds(path_rds, analysis_key, analysis_qmd)

compare_tbl <- compare_raw

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(
    method %in% c(
      "stimgate",
      "fbeta",
      "tailgate"
    )
  ) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

scenario_cols <- c(
  setdiff(colnames(sim_grid), "sim_id"),
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_tbl,
  scenarioCols = scenario_cols,
  mcse = show_mcse
)

if (isTRUE(run_plots)) {
  compare_summary |>
    dplyr::arrange(rmse) |>
    dplyr::select(
      any_of(colnames(sim_grid)),
      approach,
      method,
      n_est,
      n_threshold_fallback,
      propRespTruth_mean,
      propRespEst_mean,
      bias,
      mean_abs_error,
      rmse,
      threshold_median
    )
}

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after the canonical comparison summary.")
  knitr::knit_exit()
}

if (isTRUE(run_plots)) {
  plot_dir <- .analysis_fig_dir(fig_key, path_root = root_dir)
}
```

## Results

Every comparison figure is shown twice. "All methods" includes StimGate,
Tailgate and F-beta. "Without Tailgate" leaves Tailgate out, because it
performs very poorly without more tuning and squashes the scale for the other
two methods. The two versions are saved in the `all_methods` and `no_tailgate`
subfolders of each figure folder. Within each version there is one figure per
mean position setting.

**Monte Carlo error bars.** With `show_mcse` on (the default; set the
`show_mcse` parameter or the `SHOW_MCSE` environment variable to false to
turn it off), the summary plots show error bars for how much each plotted
value could move if the simulation were repeated, using only the results
already simulated. Here the simulated samples are not independent of each
other: within one simulated dataset, StimGate estimates its bandwidth and
bias once across the dataset's samples and can cluster their gates. So the
unit is the simulated dataset. Each statistic (for example the median
relative error) is also calculated within each dataset, and its standard
error is the standard deviation of these per-dataset values divided by the
square root of the number of datasets; the datasets are treated as
independent draws. The bar is the plotted value plus or minus 1.96 of these
standard errors. The plotted values themselves are unchanged: they are still
calculated over all samples of all datasets. No bar is drawn with fewer than
five datasets per scenario, which is the case with the current settings (two
datasets per scenario); increase the number of datasets (`n_iter_sim`) to
see them. Maxima never have error bars: the results say little about how
much larger the maximum could be.

### Mean absolute error by cell count

Mean absolute frequency error is plotted against stimulated cell count on a
logarithmic axis, with colour identifying the method. Panels separate response
frequencies (rows) and expression distributions (columns). Each plot uses one
mean position setting and no random condition perturbation. Lower values
indicate closer agreement with the true frequency.

```{r}
#| label: mae-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "mean_abs_error", "Mean absolute error", percent_y = TRUE,
        mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "mae_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_mae_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L
  )
}
```

### Median relative error by cell count

Median absolute relative error is plotted against stimulated cell count, with
colour identifying the method and panels separating response frequencies and
expression distributions. Each plot uses one mean position setting and no random
condition perturbation. Values nearer the dashed zero line indicate smaller
errors relative to the true frequency.

```{r}
#| label: med-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "med_abs_rel_error", "Median absolute relative error",
        zero_line = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "median_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_med_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L
  )
}
```

### 90th percentile relative error by cell count

The 90th percentile of absolute relative error is plotted against stimulated
cell count, with colour identifying the method. Panels separate response
frequencies and expression distributions, with one mean position setting per
plot and no random condition perturbation. Compare the upper-tail errors with
the dashed zero line.

```{r}
#| label: q90-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "q90_abs_rel_error", "90th percentile absolute relative error",
        zero_line = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "q90_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_q90_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L
  )
}
```

### 95th percentile relative error by cell count

The 95th percentile of absolute relative error is plotted against stimulated
cell count, with colour identifying the method. Panels separate response
frequencies and expression distributions, with one mean position setting per
plot and no random condition perturbation. This shows errors further into the
upper tail; the dashed line marks zero error.

```{r}
#| label: q95-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "q95_abs_rel_error", "95th percentile absolute relative error",
        zero_line = TRUE, free_y = TRUE, mcse = show_mcse
      )
    },
    dir = file.path(plot_dir, "q95_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_q95_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L
  )
}
```

```{r}
#| label: compare-relative-error-summary
#| echo: false
if (isTRUE(run_plots)) {
  compare_summary |>
    dplyr::filter(
      condition_perturbation_sd == 0
    ) |>
    dplyr::mutate(transformation = .analysis_trans_factor(transformation)) |>
    dplyr::select(
      approach, max_abs_rel_error,
      transformation, prob_response, n_cell, mean_pos_setting,
      max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error
    ) |>
    dplyr::group_by(
      mean_pos_setting,
      transformation,
      prob_response,
      n_cell
    ) |>
    tidyr::pivot_wider(
      names_from = approach,
      values_from = c(max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error)
    ) |>
    dplyr::arrange(
      dplyr::desc(q95_abs_rel_error_fbeta)
    ) |>
    print(n = Inf, width = Inf)
}
```

### Max relative error by cell count

Maximum absolute relative error is plotted against stimulated cell count,
with colour identifying the method. Panels separate response frequencies and
expression distributions, with one mean position setting per plot and no random
condition perturbation. The dashed zero line provides a reference for the
largest observed errors. There are no error bars: the results say little
about how much larger the maximum could be.

```{r}
#| label: max-rel-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0)
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = function(d) {
      .simComparePlotByCell(
        d, "max_abs_rel_error", "Maximum absolute relative error",
        zero_line = TRUE, free_y = TRUE
      )
    },
    dir = file.path(plot_dir, "max_relative_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_max_rel_error_meanpos_setting_", pos, ".png")
    },
    height = 20,
    level = 4L
  )
}
```

### Over- and under-estimates by cell count

The relative error, (estimate - truth) / truth, is summarised separately for
over-estimates and under-estimates in each scenario, using the same simulated
rows as the figures above. Over-estimates are drawn above zero and
under-estimates below. Rows of panels show the median, 95th percentile and
maximum, and columns separate expression distributions. Colour identifies the
method, and a thicker line means a larger share of the estimates lies in that
direction. The vertical scale is linear down to -100% (nothing gated) and
doubles at each step above, so -100% and a two-fold over-estimate are equally
far from zero. Each plot uses one mean position setting, one response frequency
and no random condition perturbation. Error bars (with `show_mcse` on and at
least five datasets) come from the spread between datasets of each dataset's
statistic in that direction. The number of errors in each direction is
itself random, so the bars for a rare direction understate the uncertainty a
little (the line weight shows how rare it is).

```{r}
#| label: signed-error-cell-count
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  signed_tbl <- compare_tbl |>
    dplyr::filter(condition_perturbation_sd == 0) |>
    .simCompareSignedErrorSummary(
      scenarioCols = scenario_cols, mcse = show_mcse
    )
  .simCompareFigureLoop(
    signed_tbl,
    make_plot = function(d) .simComparePlotSignedError(d, mcse = show_mcse),
    dir = file.path(plot_dir, "signed_error_by_n_cell"),
    file_fn = function(pos, extra) {
      paste0(
        "freq_bs_signed_error_meanpos_setting_", pos, "_prob_response_",
        gsub("[^0-9a-z]+", "_", .analysis_label_percent(extra)), ".png"
      )
    },
    height = 14,
    level = 4L,
    extra_col = "prob_response",
    extra_heading = function(x) {
      paste0("Response probability: ", .analysis_label_percent(x))
    }
  )
}
```

### Estimated versus true response frequency

Estimated frequency is plotted against the true background-subtracted frequency,
with colour identifying the method and panels separating expression distributions.
Each plot uses one mean position setting and pools the simulated cell counts
without random condition perturbation. Points closer to the dashed identity
line indicate better agreement.

```{r}
#| label: est-vs-truth
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl_focus |>
    dplyr::filter(
      is.finite(propRespTruth),
      is.finite(propRespEst)
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = .simComparePlotEstVsTruth,
    dir = file.path(plot_dir, "estimated_vs_true_frequency"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_est_vs_truth_", pos, ".png")
    },
    height = 9,
    level = 4L
  )
}
```

### Threshold fallback rate

The share of estimates using a threshold fallback is plotted against
stimulated cell count, with colour identifying the method. Panels separate
response frequencies and expression distributions, with one mean position
setting per plot and no random condition perturbation. Read this alongside
frequency error to distinguish calculated gates from fallbacks. Error bars
(with `show_mcse` on and at least five datasets) come from the spread of the
fallback rate between datasets.

```{r}
#| label: threshold-fallback
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  fallback_tbl <- compare_summary |>
    dplyr::filter(condition_perturbation_sd == 0) |>
    dplyr::mutate(
      fallback_rate = n_threshold_fallback / n
    )
  .simCompareFigureLoop(
    fallback_tbl,
    make_plot = function(d) .simComparePlotFallback(d, mcse = show_mcse),
    dir = file.path(plot_dir, "threshold_fallback_rate"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_threshold_fallback_rate_", pos, ".png")
    },
    height = 20,
    level = 4L
  )
}
```

### Threshold distributions

Histograms show the density of finite thresholds, with colour and fill
identifying the method. Panels separate response frequencies and expression
distributions; each plot uses one mean position setting and pools cell counts
without random condition perturbation. Compare the spread and overlap of the
threshold distributions, including finite fallback thresholds.

```{r}
#| label: threshold-density
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl |>
    dplyr::filter(
      condition_perturbation_sd == 0,
      is.finite(threshold)
    ) |>
    dplyr::select(
      approach, threshold,
      transformation, prob_response, n_cell, mean_pos_setting
    )
  .simCompareFigureLoop(
    plot_tbl,
    make_plot = .simComparePlotThresholdDensity,
    dir = file.path(plot_dir, "threshold_distributions"),
    file_fn = function(pos, extra) {
      paste0("freq_bs_threshold_density_", pos, ".png")
    },
    height = 20,
    level = 4L,
    method_col = "approach"
  )
}
```

## Rerun one scenario

A single scenario can be rerun with the production settings and RNG.
Run the setup, settings and grid chunks first.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_rerun <- 1L
row_rerun <- sim_grid_full[sim_grid_full$sim_id == sim_id_rerun, , drop = FALSE]
stopifnot(nrow(row_rerun) == 1L)
compare_one <- .simCompareRunScenario(
  row = row_rerun,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = tol_clust,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  tailgateX = tailgate_x,
  resume = FALSE
)
```
