---
title: Compare background-subtracted frequency approaches
format:
  html:
    embed-resources: true
    toc: true
params:
  analysis_run_id: null
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(progressr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

# Tailgate comparisons use the cytoUtils package directly.
if (!requireNamespace("cytoUtils", quietly = TRUE)) {
  warning(
    "Package 'cytoUtils' is not installed. Tailgate comparisons will fail if run."
  )
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches:

1. `fbeta`, called from the Python implementation at
   `scripts/python/fbeta.py` via `reticulate`.
2. `tailgate`, called from the `cytoUtils` package.

For StimGate, bandwidth is estimated automatically using the ordinary `hpi1`
selector, without background normalisation, with the bandwidth-estimation sample
capped at 10,000 cells. Tailgate is deliberately left to use its usual automatic
bandwidth estimation so that it is evaluated as it would normally run.

### Simulation settings

```{r}
#| label: actual-settings
n_cell_stim_vec <- c(
  1e3, 2e3, 5e3, 1e4, 2e4, 5e4,
  1e5, 2e5, 5e5, 1e6, 2e6, 5e6
)[c(1, 3, 5, 7)]

prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

transformation_vec <- c("gaussian", "gamma", "skew")
mean_pos_vec <- sort(unique(mean_pos_settings_tbl$mean_pos))
sample_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
condition_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1:2]
cluster_perturbation_sd_vec <- c(
  0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2
)[1]
background_relative_to_response_vec <- c(
  2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001
)[4]
n_cell_uns_relative_to_stim_vec <- c(
  0.1, 0.2, 0.5, 1, 2, 5, 10
)[4]

# StimGate bandwidth settings.
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# StimGate method settings for the full method comparison.
# Cluster refinement is enabled to represent the full one-marker procedure.
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"
```

### Simulation grid

```{r}
#| label: compare-grid
sim_grid <- tidyr::expand_grid(
  mean_pos_settings_tbl,
  prob_response = prob_response_vec,
  n_cell = n_cell_stim_vec,
  bw_mtd = stimgate_bw_mtd,
  bw_ncell_max = stimgate_bw_ncell_max,
  bw_fallback = stimgate_bw_fallback,
  bw_min = stimgate_bw_min,
  bw_max = stimgate_bw_max,
  sample_perturbation_sd = sample_perturbation_sd_vec,
  condition_perturbation_sd = condition_perturbation_sd_vec,
  cluster_perturbation_sd = cluster_perturbation_sd_vec,
  background_relative_to_response = background_relative_to_response_vec,
  n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim_vec
) |>
  dplyr::mutate(
    bias_uns = dplyr::case_when(
      transformation == "gamma" ~ 0.01,
      transformation %in% c("gaussian", "skew") ~ 0.15,
      TRUE ~ NA_real_
    )
  ) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos_setting,
    mean_pos,
    bw_mtd,
    bw_ncell_max,
    bw_fallback,
    bw_min,
    bw_max,
    bias_uns,
    sample_perturbation_sd,
    condition_perturbation_sd,
    cluster_perturbation_sd,
    background_relative_to_response,
    n_cell_uns_relative_to_stim
  ) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  ) |>
  dplyr::arrange(
    n_cell,
    dplyr::desc(prob_response)
  )

if (.isDev()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos_setting == "high",
      transformation %in% transformation_vec,
      bias_uns == 0.05,
      condition_perturbation_sd == 0
    )
}

sim_grid_all <- sim_grid

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

sim_grid
print(paste0("Total simulation scenarios: ", nrow(sim_grid)))
```

A loop for debugging a single failure or inspecting intermediate objects.

```{r}
#| label: compare-loop
#| eval: false
compare_list_raw_loop <- vector("list", nrow(sim_grid))
source("scripts/r/sim-compare-freq_bs.R")
for (i in seq_len(nrow(sim_grid))) {
  message(paste0("Running simulation ", i, " of ", nrow(sim_grid)))

  compare_list_raw_loop[[i]] <- .simCompareFreqBs(
    nSample = 5,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 2,
    biasUns = sim_grid$bias_uns[i],
    bw = NULL,
    bwFallback = sim_grid$bw_fallback[i],
    bwMin = sim_grid$bw_min[i],
    bwMax = sim_grid$bw_max[i],
    bwMtd = sim_grid$bw_mtd[i],
    bwNcellMax = sim_grid$bw_ncell_max[i],
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = sim_grid$sample_perturbation_sd[i],
    conditionPerturbationSd = sim_grid$condition_perturbation_sd[i],
    clusterPerturbationSd = sim_grid$cluster_perturbation_sd[i],
    backgroundRelativeToResponse =
      sim_grid$background_relative_to_response[i],
    ncellUnsRelativeToStim =
      sim_grid$n_cell_uns_relative_to_stim[i],
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = tol_clust,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol,
    tailgateX = tailgate_x
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos_setting = sim_grid$mean_pos_setting[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      bw_ncell_max = sim_grid$bw_ncell_max[i],
      bw_fallback = sim_grid$bw_fallback[i],
      bw_min = sim_grid$bw_min[i],
      bw_max = sim_grid$bw_max[i],
      bias_uns = sim_grid$bias_uns[i],
      sample_perturbation_sd = sim_grid$sample_perturbation_sd[i],
      condition_perturbation_sd =
        sim_grid$condition_perturbation_sd[i],
      cluster_perturbation_sd =
        sim_grid$cluster_perturbation_sd[i],
      background_relative_to_response =
        sim_grid$background_relative_to_response[i],
      n_cell_uns_relative_to_stim =
        sim_grid$n_cell_uns_relative_to_stim[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos_setting,
      mean_pos,
      bw_mtd,
      bw_ncell_max,
      bw_fallback,
      bw_min,
      bw_max,
      bias_uns,
      sample_perturbation_sd,
      condition_perturbation_sd,
      cluster_perturbation_sd,
      background_relative_to_response,
      n_cell_uns_relative_to_stim,
      sim_id,
      dplyr::everything()
    )
}
```

Run the simulations in parallel, with progress reporting. The execution grid is
shuffled first so that the larger and smaller cell-count scenarios are mixed
across workers. `sim_id` remains unchanged.

```{r}
#| label: compare-parallel
#| results: hide
log_dir <- file.path("cache", "log", "freq_bs")
path_progress_file <- file.path(log_dir, "progress.txt")

run_ctx <- .analysis_run_context(
  analysis_key = c("sim", "compare", "freq_bs"),
  run_id = analysis_run_id,
  path_root = root_dir,
  params = list(
    run_simulations = run_simulations,
    run_plots = run_plots,
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  ),
  sim_grid_chunk_index = sim_grid_chunk_index,
  sim_grid_n_chunks = sim_grid_n_chunks
)
dir_output <- run_ctx$chunk_output_dir
dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

path_rds <- file.path(
  run_ctx$staging_collated_dir,
  if (sim_grid_n_chunks > 1L) {
    sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
  } else {
    "compare_raw.rds"
  }
)

n_sample_sim <- 10
n_iter_sim <- 2
n_workers <- max(1, .simGetCores())

if (interactive()) {
  progressr::handlers(global = TRUE)
}

compare_raw <- tibble::tibble()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  compare_raw <- .simCompareFreqBsGrid(
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    probExact = TRUE,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = tol_clust,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol,
    tailgateX = tailgate_x,
    parallel = (n_workers > 1L),
    workers = n_workers,
    dirCache = dir_output,
    pathProgress = path_progress_file,
    resume = TRUE,
    sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
      sim_grid_chunk_index
    } else {
      NULL
    },
    sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
      sim_grid_n_chunks
    } else {
      NULL
    }
  )

  .write_rds_atomic(compare_raw, path_rds)

  if (sim_grid_n_chunks > 1L) {
    path_rds_full <- file.path(run_ctx$staging_collated_dir, "compare_raw.rds")
    compare_raw_full <- .simCompareCollateScenarioOutputs(
      dirCache = run_ctx$staging_run_dir,
      sim_grid = sim_grid_all
    )
    if (nrow(compare_raw_full) > 0L) {
      .write_rds_atomic(compare_raw_full, path_rds_full)
      compare_raw <- compare_raw_full
    }
  }
} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}

if (isTRUE(run_simulations)) {
  sim_status_tbl <- if (nrow(compare_raw) > 0L && "sim_id" %in% names(compare_raw)) {
    compare_raw |>
      dplyr::filter(.data$sim_id %in% sim_grid$sim_id) |>
      dplyr::group_by(.data$sim_id) |>
      dplyr::summarise(
        has_error = if ("error" %in% names(compare_raw)) {
          any(!is.na(.data$error) & .data$error != "")
        } else {
          FALSE
        },
        .groups = "drop"
      )
  } else {
    tibble::tibble(sim_id = integer(), has_error = logical())
  }

  n_failed <- sum(sim_status_tbl$has_error)
  n_completed <- sum(!sim_status_tbl$has_error)

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = nrow(compare_raw) > 0L,
    validation_ok = nrow(compare_raw) > 0L
  )

  if (.analysis_can_promote(run_ctx)) {
    .analysis_promote_run(run_ctx)
  }
}
```

### Collate results

```{r}
#| label: compare-collate
#| error: true
if (!exists("compare_raw") || nrow(compare_raw) == 0L) {
  path_rds <- file.path(run_ctx$current_dir, "collated", "compare_raw.rds")
  if (file.exists(path_rds)) {
    compare_raw <- readRDS(path_rds)
  }
}

if (!exists("compare_raw") || nrow(compare_raw) == 0L) {
  stop("Could not find canonical compare_raw.rds in current/. Run the simulation chunk first.")
}

compare_tbl <- compare_raw

compare_tbl_focus <- compare_tbl |>
  dplyr::filter(
    method %in% c(
      "stimgate",
      "fbeta",
      "tailgate"
    )
  ) |>
  dplyr::mutate(
    freq_error = propRespEst - propRespTruth,
    abs_freq_error = abs(freq_error),
    sq_freq_error = freq_error^2
  )

scenario_cols <- c(
  setdiff(colnames(sim_grid), "sim_id"),
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_tbl,
  scenarioCols = scenario_cols
)

compare_summary |>
  dplyr::arrange(rmse) |>
  dplyr::select(
    any_of(colnames(sim_grid)),
    approach,
    method,
    n_est,
    n_threshold_fallback,
    propRespTruth_mean,
    propRespEst_mean,
    bias,
    mean_abs_error,
    rmse,
    threshold_median
  )
```

### Mean absolute error by cell count

```{r}
#| label: plot-mae-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      )

    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = mean_abs_error,
        color = method
      )
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Mean absolute error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_mae_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Median relative error by cell count

```{r}
#| label: plot-med-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    print(paste0("Plotting median relative error for mean position setting: ", mean_pos_setting_curr))
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, med_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = med_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::mutate(
        max_error = pmax(fbeta, stimgate, tailgate)
      ) |>
      # dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = med_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Median relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_med_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### 90th percentile relative error by cell count

```{r}
#| label: plot-90th-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, q90_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = q90_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = q90_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "90th percentile relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_q90_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### 95th percentile relative error by cell count

```{r}
#| label: plot-95th-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, q95_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = q95_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = q95_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "95th percentile relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_q95_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

```{r}
compare_summary |>
  dplyr::filter(
    condition_perturbation_sd == 0
  ) |>
  dplyr::select(
    approach, max_abs_rel_error,
    transformation, prob_response, n_cell, mean_pos_setting,
    max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error
  ) |>
  dplyr::group_by(
    mean_pos_setting,
    transformation,
    prob_response,
    n_cell
  ) |>
  tidyr::pivot_wider(
    names_from = approach,
    values_from = c(max_abs_rel_error, med_abs_rel_error, q90_abs_rel_error, q95_abs_rel_error)
  ) |>
  dplyr::arrange(
    dplyr::desc(q95_abs_rel_error_fbeta)
  ) |>
  print(n = Inf, width = Inf)
```

### Max relative error by cell count

```{r}
#| label: plot-max-rel-error-cell-count
if (isTRUE(run_plots)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    plot_tbl <- compare_summary |>
      dplyr::filter(
        mean_pos_setting == !!mean_pos_setting_curr
      ) |>
      dplyr::filter(
        condition_perturbation_sd == 0
      )
    plot_tbl |>
      dplyr::select(
        approach, max_abs_rel_error,
        transformation, prob_response, n_cell, mean_pos_setting,
      ) |>
      tidyr::pivot_wider(
        names_from = approach,
        values_from = max_abs_rel_error
      ) |>
      dplyr::select(-mean_pos_setting) |>
      dplyr::mutate(
        winner = dplyr::case_when(
          stimgate < fbeta &
            stimgate < tailgate ~ "stimgate",
          fbeta < stimgate &
            fbeta < tailgate ~ "fbeta",
          tailgate < stimgate &
            tailgate < fbeta ~ "tailgate",
          TRUE ~ "tie"
        )
      ) |>
      dplyr::mutate(
        big_winner = ifelse(pmax(fbeta, stimgate, tailgate) > 1, "yes", "no")
      ) |>
      dplyr::arrange(dplyr::desc(big_winner), winner) |>
      dplyr::arrange(dplyr::desc(fbeta)) |>
      print(n = Inf, width = Inf)


    p <- ggplot(
      plot_tbl,
      aes(
        x = n_cell,
        y = max_abs_rel_error,
        color = method
      )
    ) +
      geom_hline(
        yintercept = 0,
        colour = "gray25",
        linetype = "dashed"
      ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(prob_response ~ transformation, scales = "free_y") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Maximum relative error",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_max_rel_error_meanpos_setting_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Estimated versus true response frequency

```{r}
#| label: plot-est-vs-truth
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl_focus |>
    dplyr::filter(
      is.finite(propRespTruth),
      is.finite(propRespEst)
    ) |>
    dplyr::filter(
      condition_perturbation_sd == 0
    )

  for (mean_pos_setting_curr in unique(plot_tbl$mean_pos_setting)) {
    plot_tbl_curr <- plot_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p_est_truth <- ggplot(
      plot_tbl_curr,
      aes(
        x = propRespTruth,
        y = propRespEst,
        color = method
      )
    ) +
      geom_point(alpha = 0.35, size = 1) +
      expand_limits(
        y = c(
          0,
          max(plot_tbl_curr$propRespTruth, na.rm = TRUE) * 2
        ),
        x = c(
          0,
          max(plot_tbl_curr$propRespTruth, na.rm = TRUE) * 2
        )
      ) +
      geom_abline(
        slope = 1,
        intercept = 0,
        linetype = "dashed"
      ) +
      facet_wrap(~transformation) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "True background-subtracted response frequency",
        y = "Estimated background-subtracted response frequency",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p_est_truth)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_est_vs_truth_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p_est_truth,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

### Threshold fallback rate

```{r}
#| label: plot-threshold-fallback
if (isTRUE(run_plots)) {
  fallback_tbl <- compare_summary |>
    dplyr::mutate(
      fallback_rate = n_threshold_fallback / n
    )

  for (mean_pos_setting_curr in unique(fallback_tbl$mean_pos_setting)) {
    fallback_tbl_curr <- fallback_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p_fallback <- ggplot(
      fallback_tbl_curr,
      aes(
        x = n_cell,
        y = fallback_rate,
        color = method
      )
    ) +
      geom_line() +
      geom_point() +
      scale_x_log10() +
      facet_grid(
        prob_response ~ transformation
      ) +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Number of stimulated cells",
        y = "Threshold fallback rate",
        color = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p_fallback)

    ggsave(
      filename = projr::projr_path_get(
        "cache",
        "sim",
        "compare",
        "freq_bs",
        paste0(
          "freq_bs_threshold_fallback_rate_",
          mean_pos_setting_curr,
          ".png"
        )
      ),
      plot = p_fallback,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```

## Plot the thresholds

```{r}
#| label: plot-threshold-density
if (isTRUE(run_plots)) {
  plot_tbl <- compare_tbl |>
    dplyr::filter(
      condition_perturbation_sd == 0,
      is.finite(threshold)
    ) |>
    dplyr::select(
      approach, threshold,
      transformation, prob_response, n_cell, mean_pos_setting
    )

  for (mean_pos_setting_curr in unique(plot_tbl$mean_pos_setting)) {
    plot_tbl_curr <- plot_tbl |>
      dplyr::filter(mean_pos_setting == !!mean_pos_setting_curr)

    p <- ggplot(
      plot_tbl_curr,
      aes(
        x = threshold,
        fill = approach,
        colour = approach
      )
    ) +
      geom_histogram(alpha = 0.5) +
      facet_grid(prob_response ~ transformation, scales = "free") +
      labs(
        title = paste0(
          "Mean position setting: ",
          mean_pos_setting_curr
        ),
        x = "Threshold",
        y = "Density",
        fill = "Method",
        colour = "Method"
      ) +
      cowplot::theme_cowplot() +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(
          fill = "white",
          color = NA
        ),
        plot.background = element_rect(
          fill = "white",
          color = NA
        ),
        legend.position = "bottom"
      )

    print(p)

    path_p <- projr::projr_path_get(
      "cache",
      "sim",
      "compare",
      "freq_bs",
      paste0(
        "freq_bs_threshold_density_",
        mean_pos_setting_curr,
        ".png"
      )
    )
    ggplot2::ggsave(
      filename = path_p,
      plot = p,
      width = 18,
      height = 15,
      units = "cm"
    )
  }
}
```
