---
title: Compare background-subtracted frequency approaches under condition mismatch
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
---

This analysis compares StimGate, F-beta and Tailgate when stimulated and
unstimulated expression distributions are misaligned. Paired samples are
simulated for selected baseline scenarios, then the stimulated expression is
shifted or its background spread is increased. The same simulated draws are
used across mismatch settings within each baseline scenario.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)
library(dplyr)
library(tidyr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
analysis_qmd <- "analysis/8-sim-compare-freq_bs-batch.qmd"

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs_batch")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "batch-mismatch-comparison-v4"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
analysis_key <- .analysis_mode_key(analysis_key)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  stop("Expected the stimgate DESCRIPTION at the repository root: ", root_dir)
}

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "flowStats", "simcyto"))
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

```

## Scientific settings

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches (`fbeta` and `tailgate`) under controlled
**condition mismatch** (technical misalignment between the stimulated and
matched unstimulated tubes).

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but adapts its histogram support to span both paired
distributions. This prevents stimulated response-tail events outside the
observed unstimulated range from being omitted from the F-beta calculation.

The simulation isolates two deterministic post-transformation mismatch
mechanisms provided by `simcyto`:

1. **Mean shift** (`stimMeanShift`): An additive shift on the transformed
   expression scale, applied either to all stimulated cells or only to the
   stimulated negative/background component.
2. **SD inflation** (`stimSdMultiplier`): A multiplicative scaling of the
   stimulated negative/background component's within-component standard
   deviation around its realised mean.

Random perturbations are disabled
(`samplePerturbationSd = 0`, `conditionPerturbationSd = 0`,
`clusterPerturbationSd = 0`). Within each baseline scenario, all mismatch
magnitudes and variants use the same random seed, so the degradation curves
compare deterministic changes to the same simulated draws.

### Simulation settings

```{r}
#| label: actual-settings
# Method settings inherited from analysis 7:
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# Cluster refinement enabled for the full one-marker procedure
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
n_sample_sim <- if (analysis_quick) 1L else 5
n_iter_sim <- if (analysis_quick) 1L else 5

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- file.path(
  root_dir,
  "scripts",
  "python",
  "fbeta.py"
)

if (isTRUE(run_simulations)) {
  if (!file.exists(path_fbeta)) {
    stop("F-beta comparison script not found: ", path_fbeta)
  }
  .simCompareFbetaEnvironment(pathFbeta = path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"

background_relative_to_response <- 0.2
n_cell_uns_relative_to_stim <- 1

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

# Targeted scenario panel representing qualitatively distinct baselines
targeted_scenarios_tbl <- tibble::tribble(
  ~transformation, ~mean_pos_setting, ~prob_response, ~n_cell, ~scenario_desc,
  "gaussian", "high", 0.002, 20000, "Rare response, clean baseline",
  "skew", "high", 0.2, 20000, "Clean skew baseline",
  "gamma", "high", 0.2, 5000, "Very clean gamma baseline",
  "gaussian", "low", 0.2, 1000, "Overlapping Gaussian (StimGate clean advantage)",
  "gamma", "low", 0.2, 1000, "Overlapping Gamma (StimGate clean advantage)",
  "gaussian", "high", 0.0005, 100000, "Rare Gaussian (F-beta clean advantage)",
  "gamma", "high", 0.0005, 20000, "Rare Gamma (F-beta clean advantage)"
) |>
  dplyr::left_join(
    mean_pos_settings_tbl,
    by = c("transformation", "mean_pos_setting")
  ) |>
  dplyr::mutate(
    base_scenario_id = dplyr::row_number(),
    bias_uns = dplyr::case_when(
      transformation == "gamma" ~ 0.01,
      transformation %in% c("gaussian", "skew") ~ 0.15,
      TRUE ~ NA_real_
    ),
    bw_mtd = stimgate_bw_mtd,
    bw_ncell_max = stimgate_bw_ncell_max,
    bw_fallback = stimgate_bw_fallback,
    bw_min = stimgate_bw_min,
    bw_max = stimgate_bw_max,
    sample_perturbation_sd = 0,
    condition_perturbation_sd = 0,
    cluster_perturbation_sd = 0,
    background_relative_to_response = background_relative_to_response,
    n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim
  )
```

## Simulation

### Mismatch experiment grids

```{r}
#| label: mismatch-grids
#| include: false
# 1. Mean-shift grid (all components):
# Gaussian and Skew: 0, 0.025, 0.05, 0.1, 0.2, 0.5, 1
# Gamma: 0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2
mean_shift_gaussian_skew <- c(0, 0.025, 0.05, 0.1, 0.2, 0.5, 1)
mean_shift_gamma <- c(0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2)

grid_mean_shift_all <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_all",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = NA_character_,
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 2. Mean-shift grid (negative/background component only):
grid_mean_shift_negative <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_negative",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = "gn",
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 3. SD-inflation grid (negative/background component only):
# Fractional SD increase: 0, 0.01, 0.025, 0.05, 0.10, 0.20, 0.5, 1
sd_increase_vec <- c(0, 0.01, 0.025, 0.05, 0.10, 0.20, 0.5, 1)

grid_sd_inflation <- tidyr::expand_grid(
  targeted_scenarios_tbl,
  sd_increase = sd_increase_vec
) |>
  dplyr::mutate(
    mismatch_type = "sd_inflation",
    mismatch_val = sd_increase,
    stim_mean_shift = 0,
    stim_mean_shift_clusters = NA_character_,
    stim_sd_multiplier = 1 + sd_increase,
    stim_sd_multiplier_clusters = "gn"
  )

# Combined full mismatch simulation grid
sim_grid <- dplyr::bind_rows(
  grid_mean_shift_all,
  grid_mean_shift_negative,
  grid_sd_inflation
) |>
  dplyr::mutate(
    sim_id = dplyr::row_number(),
    # Use common random numbers within each baseline scenario so that
    # degradation across mismatch magnitudes/variants is paired.
    sim_seed = as.integer(simulation_seed + base_scenario_id - 1L)
  )

# Keep every seeded row available for an exact interactive rerun.
sim_grid_full <- sim_grid

# Quick mode keeps baseline scenario, mismatch magnitude (including zero) and mechanism.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      base_scenario_id %in% c(4L, 5L),
      mismatch_val %in% c(0, 0.05)
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      transformation == "gaussian",
      mean_pos_setting == "high",
      prob_response == 0.002,
      mismatch_val %in% c(0, 0.05)
    )
}

sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

sim_grid_all <- sim_grid
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-dplyr::any_of("sim_seed"))

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

print(paste0(
  "Simulation scenarios assigned to chunk ",
  sim_chunk_label,
  ": ",
  nrow(sim_grid),
  " of ",
  nrow(sim_grid_all)
))
```

### Simulation execution

```{r}
#| label: compare-batch-run
#| include: false
n_workers <- max(1, .simGetCores())

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = list(
      run_simulations = run_simulations,
      run_plots = run_plots,
      comparison_semantics_version = comparison_semantics_version,
      simulation_seed = simulation_seed,
      analysis_quick = analysis_quick,
      analysis_dev = analysis_dev,
      paired_mismatch_rng = TRUE,
      analysis_grid_spec = analysis_grid_spec,
      n_sample_sim = n_sample_sim,
      n_iter_sim = n_iter_sim,
      stimgate_bw_mtd = stimgate_bw_mtd,
      stimgate_bw_ncell_max = stimgate_bw_ncell_max,
      stimgate_bw_fallback = stimgate_bw_fallback,
      stimgate_bw_min = stimgate_bw_min,
      stimgate_bw_max = stimgate_bw_max,
      tol_clust = tol_clust,
      loc_enforce_shape_threshold = loc_enforce_shape_threshold,
      calc_cyt_pos_gates = calc_cyt_pos_gates,
      fbeta_beta = fbeta_beta,
      fbeta_theta = fbeta_theta,
      fbeta_width = fbeta_width,
      tailgate_adjust = tailgate_adjust,
      tailgate_method = tailgate_method,
      tailgate_tol = tailgate_tol,
      tailgate_auto_tol = tailgate_auto_tol,
      tailgate_x = tailgate_x,
      sim_grid_chunk_index = sim_grid_chunk_index,
      sim_grid_n_chunks = sim_grid_n_chunks
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
}

if (interactive()) {
  progressr::handlers(global = TRUE)
}

compare_raw <- tibble::tibble()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  compare_raw <- .simCompareFreqBsGrid(
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    probExact = TRUE,
    covEvMin = 1.5,
    covEvMax = 1.5,
    tolClust = tol_clust,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol,
    tailgateX = tailgate_x,
    parallel = (n_workers > 1L),
    workers = n_workers,
    dirCache = dir_output,
    pathProgress = path_progress_file,
    resume = TRUE,
    retryErrors = TRUE,
    sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
      sim_grid_chunk_index
    } else {
      NULL
    },
    sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
      sim_grid_n_chunks
    } else {
      NULL
    }
  )

  .write_rds_atomic(compare_raw, path_rds)

} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}

simulation_promoted <- FALSE

if (isTRUE(run_simulations)) {
  chunk_status <- .simCompareGridOutputStatus(
    compare_raw,
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )
  n_failed <- length(chunk_status$failed_ids)
  n_completed <- length(chunk_status$completed_ids)

  chunk_error <- if (!isTRUE(chunk_status$validation_ok)) {
    paste0(
      "Analysis 8 chunk output is incomplete or invalid. Missing sim_id: ",
      paste(chunk_status$missing_ids, collapse = ", "),
      "; failed sim_id: ",
      paste(chunk_status$failed_ids, collapse = ", "),
      "; extra sim_id: ",
      paste(chunk_status$extra_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_status$collate_ok,
    validation_ok = chunk_status$validation_ok,
    error_message = chunk_error
  )

  if (!isTRUE(chunk_status$validation_ok)) {
    stop(chunk_error)
  }

  simulation_promoted <- .simComparePromoteIfReady(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )

  if (!isTRUE(run_plots)) {
    message("run_plots is false, so stopping after simulation/collation.")
    knitr::knit_exit()
  }

  if (sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    knitr::knit_exit()
  }

  if (!isTRUE(simulation_promoted)) {
    message("The simulation run has not been promoted; skipping summary and plots.")
    knitr::knit_exit()
  }
}
```

## Collation

```{r}
#| label: compare-summary
#| echo: false
path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = list(
    comparison_semantics_version = comparison_semantics_version,
    simulation_seed = simulation_seed,
    analysis_quick = analysis_quick,
    analysis_dev = analysis_dev,
    paired_mismatch_rng = TRUE,
    analysis_grid_spec = analysis_grid_spec,
    n_sample_sim = n_sample_sim,
    n_iter_sim = n_iter_sim
  )
)
compare_raw <- .analysis_read_rds(path_rds, analysis_key, analysis_qmd)

if (nrow(compare_raw) == 0L) {
  stop("Canonical compare_raw.rds is empty. Rerun the simulation analysis.")
}

canonical_status <- .simCompareGridOutputStatus(
  compare_raw,
  sim_grid = sim_grid_all,
  nSample = n_sample_sim,
  nIter = n_iter_sim
)
if (!isTRUE(canonical_status$validation_ok)) {
  stop(
    "Canonical compare_raw.rds does not contain a complete valid result ",
    "for every analysis 8 scenario."
  )
}

results_available <- TRUE

compare_summary <- .simCompareSummariseFreqBs(
  compare_raw,
  scenarioCols = c(
    "base_scenario_id",
    "mismatch_type",
    "mismatch_val",
    "stim_mean_shift",
    "stim_mean_shift_clusters",
    "stim_sd_multiplier",
    "stim_sd_multiplier_clusters",
    "sd_increase",
    "transformation",
    "mean_pos_setting",
    "mean_pos",
    "prob_response",
    "n_cell",
    "scenario_desc",
    "approach",
    "method"
  )
)

compare_summary

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after the canonical comparison summary.")
  knitr::knit_exit()
}
```

## Results

### Mean-shift degradation curves (all components vs negative only)

```{r}
#| include: false
plot_dir <- .analysis_cache_dir(c(analysis_key, "mean_shift_degradation"), path_root = root_dir)
if (
  isTRUE(run_plots) &&
    isTRUE(results_available) &&
    dir.exists(plot_dir)
) {
  unlink(plot_dir, recursive = TRUE)
}
```

#### Median

Median absolute relative error is plotted against the stimulated mean shift
on an inverse-hyperbolic-sine vertical scale. Colour identifies the method,
line type distinguishes shifts of all cells from shifts of background cells
only, and facets separate baseline scenarios. Lower curves indicate smaller
errors at the same shift.

```{r}
#| label: fig-mean-shift-degradation-median
#| fig-cap: "Median absolute relative frequency error under mean shifts."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_mean_shift <- compare_summary |>
    dplyr::filter(
      mismatch_type %in% c("mean_shift_all", "mean_shift_negative")
    )
  for (mean_pos_setting_curr in unique(summary_mean_shift$mean_pos_setting)) {
    summary_mean_shift_curr <- summary_mean_shift |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
    p_shift <- .simComparePlotMeanShift(
      summary_mean_shift_curr,
      statistic = "med_abs_rel_error",
      label = "Median Absolute Relative Error"
    )
    path_p <- file.path(
      plot_dir, paste0("plot_med_abs_rel_error_", mean_pos_setting_curr, ".pdf")
    )
    .analysis_save_plot(p_shift, path_p, width = 20, height = 15)
  }
}
```

#### Maximum

Maximum absolute relative error is plotted against the stimulated mean shift
on an inverse-hyperbolic-sine vertical scale. Colour identifies the method,
line type identifies which cells are shifted, and facets separate baseline
scenarios. Compare how the largest errors vary with shift magnitude.

```{r}
#| label: fig-mean-shift-degradation-max
#| fig-cap: "Maximum absolute relative frequency error under mean shifts."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_mean_shift <- compare_summary |>
    dplyr::filter(
      mismatch_type %in% c("mean_shift_all", "mean_shift_negative")
    )
  for (mean_pos_setting_curr in unique(summary_mean_shift$mean_pos_setting)) {
    summary_mean_shift_curr <- summary_mean_shift |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
    p_shift <- .simComparePlotMeanShift(
      summary_mean_shift_curr,
      statistic = "max_abs_rel_error",
      label = "Maximum Absolute Relative Error"
    )
    path_p <- file.path(
      plot_dir, paste0("plot_max_", mean_pos_setting_curr, ".pdf")
    )
    .analysis_save_plot(p_shift, path_p, width = 20, height = 15)
  }
}
```

#### 90th percentile

The 90th percentile of absolute relative error is plotted against the
stimulated mean shift on an inverse-hyperbolic-sine vertical scale. Colour
identifies the method, line type identifies which cells are shifted, and
facets separate baseline scenarios. Compare upper-tail errors at each shift.

```{r}
#| label: fig-mean-shift-degradation-q90
#| fig-cap: "90th percentile absolute relative frequency error under mean shifts."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_mean_shift <- compare_summary |>
    dplyr::filter(
      mismatch_type %in% c("mean_shift_all", "mean_shift_negative")
    )
  for (mean_pos_setting_curr in unique(summary_mean_shift$mean_pos_setting)) {
    summary_mean_shift_curr <- summary_mean_shift |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
    p_shift <- .simComparePlotMeanShift(
      summary_mean_shift_curr,
      statistic = "q90_abs_rel_error",
      label = "90th Percentile Absolute Relative Error"
    )
    path_p <- file.path(
      plot_dir, paste0("plot_q90_", mean_pos_setting_curr, ".pdf")
    )
    .analysis_save_plot(p_shift, path_p, width = 20, height = 15)
  }
}
```

### SD-inflation degradation curves

```{r}
#| include: false
plot_dir <- .analysis_cache_dir(c(analysis_key, "sd_inflation_degradation"), path_root = root_dir)
if (
  isTRUE(run_plots) &&
    isTRUE(results_available) &&
    dir.exists(plot_dir)
) {
  unlink(plot_dir, recursive = TRUE)
}
```

Median absolute relative error is plotted against the fractional increase
in stimulated background standard deviation on an inverse-hyperbolic-sine
vertical scale. Colour identifies the method and facets separate baseline
scenarios, with one mean position setting per plot. Zero on the horizontal
axis is the unchanged background spread.

```{r}
#| label: fig-sd-inflation-degradation
#| fig-cap: "Median absolute relative frequency error under background SD inflation."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_sd <- compare_summary |>
    dplyr::filter(mismatch_type == "sd_inflation")

  for (mean_pos_setting_curr in unique(summary_sd$mean_pos_setting)) {
    summary_sd_curr <- summary_sd |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
    p_sd <- ggplot(
      summary_sd_curr,
      aes(
        x = mismatch_val,
        y = med_abs_rel_error,
        color = method,
        group = method
      )
    ) +
      geom_line(linewidth = 0.8) +
      geom_point(size = 2) +
      facet_wrap(~scenario_desc, scales = "free_y") +
      scale_y_continuous(
        transform = scales::asinh_trans()
      ) +
      cowplot::theme_cowplot(font_size = 10) +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        legend.position = "bottom"
      ) +
      labs(
        title = "Degradation under Stimulated SD Inflation",
        subtitle = "Median Absolute Relative Error vs Fractional SD Increase (Negative Component Only)",
        x = "Fractional SD Increase",
        y = "Median Absolute Relative Error (asinh scale)",
        color = "Method"
      )
    path_p <- file.path(
      .analysis_cache_dir(c(analysis_key, "sd_inflation_degradation"), path_root = root_dir),
      paste0(
        "plot_",
        mean_pos_setting_curr,
        ".pdf"
      )
    )
    .analysis_save_plot(
      filename = path_p,
      plot = p_sd,
      width = 20,
      height = 15,
      units = "cm"
    )
  }
}
```

### Threshold placement diagnostic (unstimulated background intrusion)

The mean proportion of unstimulated cells above the gate is plotted against
the stimulated mean shift. Colour identifies the method, line type identifies
which stimulated cells are shifted, and facets separate baseline scenarios.
Higher values indicate that more unstimulated cells lie above the chosen gate.

```{r}
#| label: fig-threshold-placement-unstim
#| fig-cap: "Unstimulated cells above the gate under stimulated mean shifts."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_shifts <- compare_summary |>
    dplyr::filter(
      mismatch_type %in% c("mean_shift_all", "mean_shift_negative")
    )

  for (mean_pos_setting_curr in unique(summary_shifts$mean_pos_setting)) {
    summary_shifts_curr <- summary_shifts |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr)

    p_placement <- ggplot(
      summary_shifts_curr,
      aes(
        x = mismatch_val,
        y = propUns_mean,
        color = method,
        linetype = mismatch_type,
        group = interaction(method, mismatch_type)
      )
    ) +
      geom_line(linewidth = 0.8) +
      geom_point(size = 2) +
      facet_wrap(~scenario_desc, scales = "free_y") +
      cowplot::theme_cowplot(font_size = 10) +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA)
      ) +
      theme(
        legend.position = "bottom",
        legend.box = "vertical",
        legend.box.just = "left"
      ) +
      guides(
        colour = guide_legend(nrow = 1, byrow = TRUE),
        linetype = guide_legend(nrow = 1, byrow = TRUE)
      ) +
      labs(
        title = "Threshold Placement: Unstimulated Background Intrusion",
        subtitle = "Proportion of unstimulated cells above threshold vs shift",
        x = "Stimulated Mean Shift",
        y = "Proportion of Unstimulated Cells Above Gate",
        color = "Method",
        linetype = "Mismatch Variant"
      )
    path_p <- file.path(
      .analysis_cache_dir(c(analysis_key, "threshold_placement_unstim"), path_root = root_dir),
      paste0(
        "plot_",
        mean_pos_setting_curr,
        ".pdf"
      )
    )
    .analysis_save_plot(
      filename = path_p,
      plot = p_placement,
      width = 20,
      height = 15,
      units = "cm"
    )
  }
}
```

### Upper-tail error analysis

The 90th percentile of absolute relative error is plotted against mismatch
magnitude on an inverse-hyperbolic-sine vertical scale, with colour identifying
the method. Facet rows separate mismatch mechanisms and columns separate
baseline scenarios, with one mean position setting per plot. Compare curves
within each panel, since mean shift and SD inflation use different units.

```{r}
#| label: fig-upper-tail-and-fallbacks
#| fig-cap: "90th percentile absolute relative frequency error across mismatch settings."
#| echo: false
#| fig-width: 10
#| fig-height: 9
if (isTRUE(run_plots) && isTRUE(results_available)) {
  for (mean_pos_setting_curr in unique(compare_summary$mean_pos_setting)) {
    compare_summary_curr <- compare_summary |>
      dplyr::filter(mean_pos_setting == mean_pos_setting_curr)

    p_tail <- ggplot(
      compare_summary_curr,
      aes(
        x = mismatch_val,
        y = q90_abs_rel_error,
        color = method,
        group = method
      )
    ) +
      geom_line(linewidth = 0.8) +
      geom_point(size = 2) +
      facet_grid(mismatch_type ~ scenario_desc, scales = "free") +
      scale_y_continuous(
        transform = scales::asinh_trans()
      ) +
      cowplot::theme_cowplot(font_size = 10) +
      cowplot::background_grid(major = "xy", minor = "none") +
      theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        legend.position = "bottom"
      ) +
      labs(
        title = "90th Percentile Relative Error under Mismatch",
        subtitle = paste0("Mean position setting: ", mean_pos_setting_curr),
        x = "Mismatch Magnitude",
        y = "90th Percentile Absolute Relative Error (asinh scale)",
        color = "Method"
      )
    path_p <- file.path(
      .analysis_cache_dir(c(analysis_key, "upper_tail_and_fallbacks"), path_root = root_dir),
      paste0(
        "plot_",
        mean_pos_setting_curr,
        ".pdf"
      )
    )
    .analysis_save_plot(
      filename = path_p,
      plot = p_tail,
      width = 20,
      height = 15,
      units = "cm"
    )
  }
}
```

## Rerun one scenario

A single scenario can be rerun with the production settings and RNG.
Run the setup, settings and grid chunks first.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_rerun <- 1L
row_rerun <- sim_grid_full[sim_grid_full$sim_id == sim_id_rerun, , drop = FALSE]
stopifnot(nrow(row_rerun) == 1L)
compare_one <- .simCompareRunScenario(
  row = row_rerun,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = tol_clust,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  tailgateX = tailgate_x,
  resume = FALSE
)
```
