---
title: Compare background-subtracted frequency approaches under condition mismatch
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
  sim_size: final
---

This analysis compares StimGate, F-beta and Tailgate when stimulated and
unstimulated expression distributions are misaligned. Paired samples are
simulated for selected baseline scenarios, then the stimulated expression is
shifted or its background spread is increased. The same simulated draws are
used across mismatch settings within each baseline scenario.

The question is whether misalignment between the stimulated and unstimulated
negative components makes a method call cells in the right-hand tail of the
stimulated negative component positive. The main outcomes are therefore the
purity and the detection rate of each method's final gate in the stimulated
tube, measured against the simulation's own labels:

- **False discovery proportion (FDP):** of the stimulated cells classified as
  positive, the proportion that are genuinely negative.
- **Sensitivity:** of the genuinely positive stimulated cells, the proportion
  classified as positive.

Both lie between 0 and 1 in every scenario. The relative error of the
background-subtracted frequency, which becomes very large when the true
response is rare, is kept as a secondary result.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)
library(dplyr)
library(tidyr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
analysis_qmd <- "analysis/8-sim-compare-freq_bs-batch.qmd"

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs_batch")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "batch-mismatch-comparison-v10"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "final" (default) or "draft" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("8-sim-compare-freq_bs-batch")
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  stop("Expected the stimgate DESCRIPTION at the repository root: ", root_dir)
}

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "flowStats", "simcyto"))
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

```

## Scientific settings

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches (`fbeta` and `tailgate`) under controlled
**condition mismatch** (technical misalignment between the stimulated and
matched unstimulated tubes).

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but adapts its histogram support to span both paired
distributions. This prevents stimulated response-tail events outside the
observed unstimulated range from being omitted from the F-beta calculation.

The simulation isolates two deterministic post-transformation mismatch
mechanisms provided by `simcyto`:

1. **Mean shift** (`stimMeanShift`): An additive shift on the transformed
   expression scale, applied either to all stimulated cells or only to the
   stimulated negative/background component. Shifts go up to 0.35 for the
   Gaussian and Skew transformations and up to 0.2 for Gamma.
2. **SD inflation** (`stimSdMultiplier`): A multiplicative scaling of the
   stimulated negative/background component's within-component standard
   deviation around its realised mean, increasing it by up to 35%.

The two mean-shift variants are different mechanisms. Shifting all
stimulated cells misaligns the background while keeping the separation
between positive and negative cells within the stimulated tube. Shifting only
the stimulated negatives misaligns the background and also moves it towards
the unchanged positive component. Neither is assumed to degrade performance
steadily as the mismatch grows.

Random perturbations are disabled
(`samplePerturbationSd = 0`, `conditionPerturbationSd = 0`,
`clusterPerturbationSd = 0`). Within each baseline scenario, all mismatch
magnitudes and variants use the same random seed, so the degradation curves
compare deterministic changes to the same simulated draws. Each replicate
draws its own seed from that scenario seed before any method runs, so later
replicates stay paired even though the methods use different amounts of
randomness in different mismatch settings.

### Simulation settings

Each scenario simulates 20 datasets of 10 samples in final runs and 5 datasets
of 10 samples in draft runs (`SIM_SIZE=draft`). The samples in a dataset are
gated together, with shared settings and cluster gates, so the dataset is the
unit for uncertainty. Draft results are kept separately and are for iterating,
not for reporting.

```{r}
#| label: actual-settings
# Method settings inherited from analysis 7:
stimgate_bw_mtd <- "nrd0"
# StimGate's unstimulated-sample bias equals its initial bandwidth estimate:
# with biasUns = NULL, gateStim() uses 0.25 * bwFallback * biasUnsFactor.
stimgate_bias_uns_factor <- 4
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# Cluster refinement enabled for the full one-marker procedure
# Share thresholds across clusters of similar samples via stimControl(clusterGates = ).
cluster_gates <- TRUE
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
# Each dataset of `n_sample_sim` samples is gated together (shared settings
# and cluster gates), so the dataset is the unit for uncertainty. Draft keeps
# the samples per dataset and uses a quarter of the datasets (at least 5).
n_sample_sim <- if (analysis_quick) 1L else 10L
n_iter_sim <- if (analysis_quick) 1L else c(final = 20L, draft = 5L)[[sim_size]]

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- file.path(
  root_dir,
  "scripts",
  "python",
  "fbeta.py"
)

if (isTRUE(run_simulations)) {
  if (!file.exists(path_fbeta)) {
    stop("F-beta comparison script not found: ", path_fbeta)
  }
  .simCompareFbetaEnvironment(pathFbeta = path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"

background_relative_to_response <- 0.2
n_cell_uns_relative_to_stim <- 1

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

# Targeted scenario panel representing qualitatively distinct baselines
targeted_scenarios_tbl <- tibble::tribble(
  ~transformation, ~mean_pos_setting, ~prob_response, ~n_cell, ~scenario_desc,
  "gaussian", "high", 0.002, 20000, "Rare response, clean baseline",
  "skew", "high", 0.2, 20000, "Clean skew baseline",
  "gamma", "high", 0.2, 5000, "Very clean gamma baseline",
  "gaussian", "low", 0.2, 1000, "Overlapping Gaussian (StimGate clean advantage)",
  "gamma", "low", 0.2, 1000, "Overlapping Gamma (StimGate clean advantage)",
  "gaussian", "high", 0.0005, 100000, "Rare Gaussian (F-beta clean advantage)",
  "gamma", "high", 0.0005, 20000, "Rare Gamma (F-beta clean advantage)"
) |>
  dplyr::left_join(
    mean_pos_settings_tbl,
    by = c("transformation", "mean_pos_setting")
  ) |>
  dplyr::mutate(
    base_scenario_id = dplyr::row_number(),
    bias_uns = NA_real_,
    bias_uns_factor = stimgate_bias_uns_factor,
    bw_mtd = stimgate_bw_mtd,
    bw_ncell_max = stimgate_bw_ncell_max,
    bw_fallback = stimgate_bw_fallback,
    bw_min = stimgate_bw_min,
    bw_max = stimgate_bw_max,
    sample_perturbation_sd = 0,
    condition_perturbation_sd = 0,
    cluster_perturbation_sd = 0,
    background_relative_to_response = background_relative_to_response,
    n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim
  )
```

## Simulation

### Mismatch experiment grids

```{r}
#| label: mismatch-grids
#| include: false
# 1. Mean-shift grid (all components):
# Gaussian and Skew: 0 to 0.35
# Gamma: 0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2
mean_shift_gaussian_skew <- c(0, 0.025, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35)
mean_shift_gamma <- c(0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2)

grid_mean_shift_all <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_all",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = NA_character_,
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 2. Mean-shift grid (negative/background component only):
grid_mean_shift_negative <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_negative",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = "gn",
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 3. SD-inflation grid (negative/background component only):
# Fractional SD increase: 0 to 0.35, matching the Gaussian and Skew mean shifts
sd_increase_vec <- c(0, 0.025, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35)

grid_sd_inflation <- tidyr::expand_grid(
  targeted_scenarios_tbl,
  sd_increase = sd_increase_vec
) |>
  dplyr::mutate(
    mismatch_type = "sd_inflation",
    mismatch_val = sd_increase,
    stim_mean_shift = 0,
    stim_mean_shift_clusters = NA_character_,
    stim_sd_multiplier = 1 + sd_increase,
    stim_sd_multiplier_clusters = "gn"
  )

# Combined full mismatch simulation grid
sim_grid <- dplyr::bind_rows(
  grid_mean_shift_all,
  grid_mean_shift_negative,
  grid_sd_inflation
) |>
  dplyr::mutate(
    sim_id = dplyr::row_number(),
    # Use common random numbers within each baseline scenario so that
    # degradation across mismatch magnitudes/variants is paired.
    sim_seed = as.integer(simulation_seed + base_scenario_id - 1L)
  )

# Keep every seeded row available for an exact interactive rerun.
sim_grid_full <- sim_grid

# Gate diagnostic: the very clean gamma baseline before and after the jump in
# StimGate's error near a shift of 0.1, for both shift variants. The rows
# come from the full grid, so they keep their production seeds.
gate_diagnostic_spec <- list(
  base_scenario_id = 3L,
  mismatch_type = c("mean_shift_all", "mean_shift_negative"),
  mismatch_val = c(0, 0.05, 0.1)
)
gate_diagnostic_grid <- sim_grid_full |>
  dplyr::filter(
    base_scenario_id == gate_diagnostic_spec$base_scenario_id,
    mismatch_type %in% gate_diagnostic_spec$mismatch_type,
    mismatch_val %in% gate_diagnostic_spec$mismatch_val
  )
stopifnot(
  nrow(gate_diagnostic_grid) ==
    length(gate_diagnostic_spec$mismatch_type) *
      length(gate_diagnostic_spec$mismatch_val)
)

# Quick mode keeps baseline scenario, mismatch magnitude (including zero) and mechanism.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      base_scenario_id %in% c(4L, 5L),
      mismatch_val %in% c(0, 0.05)
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      transformation == "gaussian",
      mean_pos_setting == "high",
      prob_response == 0.002,
      mismatch_val %in% c(0, 0.05)
    )
}

sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

sim_grid_all <- sim_grid
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-dplyr::any_of("sim_seed"))

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

print(paste0(
  "Simulation scenarios assigned to chunk ",
  sim_chunk_label,
  ": ",
  nrow(sim_grid),
  " of ",
  nrow(sim_grid_all)
))
```

### Simulation execution

```{r}
#| label: compare-batch-run
#| include: false
n_workers <- max(1, .simGetCores())

analysis_result_params <- list(
  comparison_semantics_version = comparison_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  paired_mismatch_rng = TRUE,
  analysis_grid_spec = analysis_grid_spec,
  gate_diagnostic_spec = gate_diagnostic_spec,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  stimgate_bw_mtd = stimgate_bw_mtd,
  stimgate_bias_uns_factor = stimgate_bias_uns_factor,
  stimgate_bw_ncell_max = stimgate_bw_ncell_max,
  stimgate_bw_fallback = stimgate_bw_fallback,
  stimgate_bw_min = stimgate_bw_min,
  stimgate_bw_max = stimgate_bw_max,
  cluster_gates = cluster_gates,
  loc_enforce_shape_threshold = loc_enforce_shape_threshold,
  calc_cyt_pos_gates = calc_cyt_pos_gates,
  fbeta_beta = fbeta_beta,
  fbeta_theta = fbeta_theta,
  fbeta_width = fbeta_width,
  tailgate_adjust = tailgate_adjust,
  tailgate_method = tailgate_method,
  tailgate_tol = tailgate_tol,
  tailgate_auto_tol = tailgate_auto_tol,
  tailgate_x = tailgate_x
)

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks
      ),
      analysis_result_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
}

if (interactive()) {
  progressr::handlers(global = TRUE)
}

compare_raw <- tibble::tibble()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  compare_raw <- .simCompareFreqBsGrid(
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    probExact = TRUE,
    covEvMin = 1.5,
    covEvMax = 1.5,
    clusterGates = cluster_gates,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol,
    tailgateX = tailgate_x,
    parallel = (n_workers > 1L),
    workers = n_workers,
    dirCache = dir_output,
    pathProgress = path_progress_file,
    dirJobs = run_ctx$chunk_jobs_dir,
    progressHeading = "BATCH MISMATCH COMPARISON SIMULATION PROGRESS",
    resume = TRUE,
    retryErrors = TRUE,
    sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
      sim_grid_chunk_index
    } else {
      NULL
    },
    sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
      sim_grid_n_chunks
    } else {
      NULL
    }
  )

  .write_rds_atomic(compare_raw, path_rds)

} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}

# The gate diagnostic needs the cells, which the grid run does not keep, so
# the first chunk reruns its few rows for one replicate and saves them with
# the run. It runs before the chunk is marked complete, so a promoted run
# always includes it.
path_gate_diagnostic <- NULL
if (isTRUE(run_simulations) && sim_grid_chunk_index == 1L) {
  path_gate_diagnostic <- file.path(
    run_ctx$staging_collated_dir, "gate_diagnostic.rds"
  )
  if (!file.exists(path_gate_diagnostic)) {
    gate_diagnostic <- .simCompareGateDiagnosticRun(
      gate_diagnostic_grid,
      nSample = n_sample_sim,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      probExact = TRUE,
      covEvMin = 1.5,
      covEvMax = 1.5,
      clusterGates = cluster_gates,
      locEnforceShapeThreshold = loc_enforce_shape_threshold,
      calcCytPosGates = calc_cyt_pos_gates,
      pathFbeta = path_fbeta,
      fbetaBeta = fbeta_beta,
      fbetaTheta = fbeta_theta,
      fbetaWidth = fbeta_width,
      tailgateAdjust = tailgate_adjust,
      tailgateAutoTol = tailgate_auto_tol,
      tailgateMethod = tailgate_method,
      tailgateTol = tailgate_tol,
      tailgateX = tailgate_x
    )
    .write_rds_atomic(gate_diagnostic, path_gate_diagnostic)
  }
}

simulation_promoted <- FALSE

if (isTRUE(run_simulations)) {
  chunk_status <- .simCompareGridOutputStatus(
    compare_raw,
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )
  n_failed <- length(chunk_status$failed_ids)
  n_completed <- length(chunk_status$completed_ids)

  chunk_error <- if (!isTRUE(chunk_status$validation_ok)) {
    paste0(
      "Analysis 8 chunk output is incomplete or invalid. Missing sim_id: ",
      paste(chunk_status$missing_ids, collapse = ", "),
      "; failed sim_id: ",
      paste(chunk_status$failed_ids, collapse = ", "),
      "; extra sim_id: ",
      paste(chunk_status$extra_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_status$collate_ok,
    validation_ok = chunk_status$validation_ok,
    error_message = chunk_error
  )

  if (!isTRUE(chunk_status$validation_ok)) {
    stop(chunk_error)
  }

  simulation_promoted <- .simComparePromoteIfReady(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    nSample = n_sample_sim,
    nIter = n_iter_sim,
    validate_full = .simCompareValidateMismatch
  )

  if (!isTRUE(run_plots)) {
    message("run_plots is false, so stopping after simulation/collation.")
    knitr::knit_exit()
  }

  if (sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    knitr::knit_exit()
  }

  if (!isTRUE(simulation_promoted)) {
    message("The simulation run has not been promoted; skipping summary and plots.")
    knitr::knit_exit()
  }
}
```

## Collation

```{r}
#| label: compare-summary
#| echo: false
path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = analysis_result_params
)
compare_raw <- .analysis_read_rds(path_rds, analysis_key, analysis_qmd)

if (nrow(compare_raw) == 0L) {
  stop("Canonical compare_raw.rds is empty. Rerun the simulation analysis.")
}

canonical_status <- .simCompareGridOutputStatus(
  compare_raw,
  sim_grid = sim_grid_all,
  nSample = n_sample_sim,
  nIter = n_iter_sim
)
if (!isTRUE(canonical_status$validation_ok)) {
  stop(
    "Canonical compare_raw.rds does not contain a complete valid result ",
    "for every analysis 8 scenario."
  )
}

results_available <- TRUE

scenario_cols <- c(
  "base_scenario_id",
  "mismatch_type",
  "mismatch_val",
  "stim_mean_shift",
  "stim_mean_shift_clusters",
  "stim_sd_multiplier",
  "stim_sd_multiplier_clusters",
  "sd_increase",
  "transformation",
  "mean_pos_setting",
  "mean_pos",
  "prob_response",
  "n_cell",
  "scenario_desc",
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_raw,
  scenarioCols = scenario_cols
)

compare_summary
```

## Validation checks

Three checks run on the full collated results before promotion and again before
any figure is drawn; the render stops if one fails.

- **Counts match the gates.** For every stimulated tube, replicate and method,
  the confusion-matrix counts must add up to the number of stimulated cells,
  and the cells above the gate (true plus false positives) must equal the
  number of stimulated cells the method itself gated. Results that lack the
  counts are rejected as incomplete.
- **Pairing across mismatch settings.** Within each baseline scenario,
  replicate and sample, every mismatch setting must have the same
  unstimulated tube (which no mismatch changes), the same number of genuine
  positives and the same number of stimulated cells.
- **Zero mismatch.** With no mismatch, shifting only the stimulated negatives
  must give exactly the same gates and counts as shifting all stimulated
  cells. Zero SD inflation is compared too; it rescales the background by
  one, which can change the last bit of a value, so it is reported rather
  than required to match exactly.

```{r}
#| label: validation-checks
#| echo: false
mismatch_validation <- .simCompareValidateMismatch(compare_raw)
pairing_check <- mismatch_validation$pairing_check
zero_agreement <- mismatch_validation$zero_agreement

cat(
  "Counts reproduce the gated stimulated counts in all",
  sum(compare_raw$method %in% c("stimgate", "fbeta", "tailgate")),
  "tube-level results.\n"
)
cat(
  "Paired data in all", nrow(pairing_check),
  "baseline-scenario, replicate and sample groups.\n"
)
knitr::kable(
  zero_agreement |>
    dplyr::mutate(
      mismatch_type = unname(.simCompareMismatchLabels[mismatch_type]),
      method = unname(.analysis_method_labels[method])
    ),
  col.names = c(
    "Zero-mismatch variant", "Method", "Compared", "Same gate",
    "Same counts", "Largest gate difference"
  ),
  caption = paste(
    "Agreement of each zero-mismatch variant with the zero shift of all",
    "stimulated cells, on the same replicates."
  )
)

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after the canonical comparison summary.")
  knitr::knit_exit()
}
```

## Results

Every comparison figure is shown twice. "All methods" includes StimGate,
Tailgate and F-beta. "Without Tailgate" leaves Tailgate out, because it
performs very poorly without more tuning and squashes the scale for the other
two methods. The two versions are saved in the `all_methods` and `no_tailgate`
subfolders of each figure folder. Within each version there is one figure per
mean position setting and mismatch mechanism.

Mean shifts are additive shifts on the transformed expression scale, the
scale the methods gate on. SD inflation is the fractional increase in the
standard deviation of the stimulated background (negative) cells.

```{r}
#| label: figure-helpers
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  # Start each figure folder afresh.
  .compare_fresh_fig_dir <- function(name) {
    path <- .analysis_fig_dir(c(fig_key, name), path_root = root_dir, create = FALSE)
    if (dir.exists(path)) unlink(path, recursive = TRUE)
    .analysis_fig_dir(c(fig_key, name), path_root = root_dir)
  }
  .compare_mismatch_heading <- function(extra) {
    type <- if (is.list(extra)) extra$mismatch_type else extra
    out <- paste0("Mismatch: ", unname(.simCompareMismatchLabels[type]))
    if (is.list(extra)) {
      out <- paste0(out, "; cells: ", .analysis_label_number(extra$n_cell))
    }
    out
  }
  .compare_mismatch_file <- function(pos, extra) {
    type <- if (is.list(extra)) extra$mismatch_type else extra
    cells <- if (is.list(extra)) paste0("_n_cell_", extra$n_cell) else ""
    paste0("plot_", type, "_mean_pos_setting_", pos, cells, ".pdf")
  }
  .compare_mismatch_x <- function(type) {
    unname(.simCompareMismatchAxisLabels[type])
  }
  # Plot function for one figure: the x axis label depends on the mechanism.
  .compare_mismatch_plot <- function(plot_fn, ...) {
    function(d) {
      plot_fn(d, x_label = .compare_mismatch_x(d$mismatch_type[[1]]), ...)
    }
  }
}
```

## Gate purity and detection

For every stimulated tube, replicate and method, the method's final gate is
applied to the stimulated cells and compared with the simulation labels. As
in the package, a cell is positive when its expression is strictly above the
gate; a cell exactly at the gate is negative. StimGate is scored with its
final gate, which is the cluster-refined gate where clustering applies.
Positive cells that are part of the background in the stimulated tube count
as genuine positives: these outcomes describe the purity and detection rate
of the stimulated gate, not which positive cells were induced by
stimulation. The background-subtracted frequency is a separate estimand and
is covered in the secondary results.

From the counts of true positives (TP), false positives (FP), false
negatives (FN) and true negatives (TN):

- FDP = FP / (TP + FP), undefined when the gate selects no stimulated cells;
- sensitivity = TP / (TP + FN), so an empty gate has zero sensitivity;
- false-positive rate = FP / (FP + TN);
- selected fraction = (TP + FP) / (TP + FP + FN + TN).

Each replicate gives one value of each outcome, and the figures summarise
these replicate-level values directly; cell counts are never pooled across
replicates. **FDP summaries leave out replicates whose gate selected no
cells**, because FDP is undefined there; the tables below give how many such
replicates there are. FDP and sensitivity are always shown together, so a
gate that selects almost nothing cannot look successful merely because it
contains few negatives.

```{r}
#| label: classification-summary
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  classification_cols <- c(
    "base_scenario_id", "scenario_desc", "transformation", "mean_pos_setting",
    "prob_response", "n_cell", "mismatch_type", "mismatch_val", "method"
  )
  class_scenario <- .simCompareClassificationSummary(
    compare_raw, scenarioCols = classification_cols
  )
}
```

### FDP and sensitivity against mismatch size

The upper row of each figure is FDP and the lower row sensitivity, both on
fixed 0-100% scales. Solid lines are medians over replicates; dashed lines
are the worse tail, the 90th percentile of FDP and the 10th percentile of
sensitivity. Columns are baseline scenarios, each with its own horizontal
range (on a square-root scale) because mismatch sizes differ between
transformations. Each mismatch mechanism has its own figure.

```{r}
#| label: fdp-sensitivity
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    class_scenario,
    make_plot = .compare_mismatch_plot(
      .simComparePlotClassification, outcomes = c("fdp", "sensitivity")
    ),
    dir = .compare_fresh_fig_dir("fdp_sensitivity"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 13,
    level = 4L
  )
}
```

### Observation counts

Each cell gives, for one method and setting, the number of valid replicates
(no run error), the number with a defined FDP (the gate selected at least one
cell), the number of empty gates and the number of gates that used a
threshold fallback. Fallback gates are included in the outcomes above, as
they are the gates the methods applied.

```{r}
#| label: classification-counts
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  for (type in intersect(names(.simCompareMismatchLabels), class_scenario$mismatch_type)) {
    .analysis_heading(
      paste0("Mismatch: ", unname(.simCompareMismatchLabels[type])), 4L
    )
    print(knitr::kable(
      .simCompareClassificationCountTable(
        class_scenario[class_scenario$mismatch_type == type, ]
      ),
      caption = "Valid / FDP defined / empty gates / fallbacks."
    ))
    cat("\n\n")
  }
}
```

### False-positive rate

The false-positive rate is the share of the genuinely negative stimulated
cells that lie above the gate. It is a secondary diagnostic. FDP depends on
how common genuine positives are as well as on the gate: the same
false-positive rate gives a much larger FDP when the response is rare, so
differences in FDP between rare- and common-response scenarios are not all
due to the methods. The vertical scales here are not fixed, because the
rates are small.

```{r}
#| label: false-positive-rate
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    class_scenario,
    make_plot = .compare_mismatch_plot(
      .simComparePlotClassification, outcomes = "fpr", unit_scale = FALSE
    ),
    dir = .compare_fresh_fig_dir("false_positive_rate"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 9,
    level = 4L
  )
}
```

## Very clean gamma baseline: gate diagnostic

In the very clean gamma baseline, StimGate's median absolute relative error
jumped to about 1 near a mean shift of 0.1. The direction of a relative error
of 1 is not clear from its size alone, so this section shows the cells, their
true labels and the final gates before and after the jump, for both shift
variants. The rows below were rerun during the simulation step for one
replicate with the production seeds and cells kept; they are saved with the
results, so nothing is simulated here. The replicate is the same simulated
draw in every setting, and (with the same number of samples) it is the first
replicate of the main results.

```{r}
#| label: gate-diagnostic-data
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  path_gate_diagnostic <- .analysis_current_file(
    run_ctx = run_ctx,
    relative_path = c("collated", "gate_diagnostic.rds"),
    required_params = analysis_result_params
  )
  gate_diagnostic <- .analysis_read_rds(
    path_gate_diagnostic, analysis_key, analysis_qmd
  )
  diag_results <- gate_diagnostic$results
  diag_primary <- diag_results |>
    dplyr::filter(method %in% c("stimgate", "fbeta", "tailgate")) |>
    .simCompareClassificationMetrics()
  # StimGate's sample-level gate, before cluster refinement.
  diag_sample_gate <- diag_results |>
    dplyr::filter(method == "stimgate_loc_sample") |>
    dplyr::select(
      sim_id, sample, sample_gate = threshold,
      sample_gate_source = gateReturnPoint
    )

  # Where the main results contain the same rows, the diagnostic replicate
  # must reproduce their first replicate exactly.
  diag_match <- diag_primary |>
    dplyr::select(sim_id, sample, method, threshold) |>
    dplyr::inner_join(
      compare_raw |>
        dplyr::filter(iter == 1L) |>
        dplyr::select(sim_id, sample, method, threshold_main = threshold),
      by = c("sim_id", "sample", "method")
    )
  if (nrow(diag_match) > 0L) {
    cat(
      "The diagnostic replicate reproduces the first replicate of the main",
      "results for", sum(diag_match$threshold == diag_match$threshold_main),
      "of", nrow(diag_match), "matching gates.\n"
    )
  } else {
    cat(
      "The main results of this run do not include the diagnostic scenario,",
      "so the diagnostic replicate is not compared with them.\n"
    )
  }
}
```

The table summarises each setting over samples: median FDP (over samples
with a defined FDP), median sensitivity and median selected fraction, with
the number of empty gates and fallbacks. The second table lists StimGate's
gates for every sample, with the sample-level gate chosen before cluster
refinement and how it was chosen.

```{r}
#| label: gate-diagnostic-tables
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  q <- .simCompareQuantileFinite
  diag_summary <- diag_primary |>
    dplyr::group_by(mismatch_type, mismatch_val, method) |>
    dplyr::summarise(
      samples = dplyr::n(),
      fdp_defined = sum(is.finite(fdp)),
      fdp = .analysis_label_percent(q(fdp, 0.5)),
      sensitivity = .analysis_label_percent(q(sensitivity, 0.5)),
      selected = .analysis_label_percent(q(selected_fraction, 0.5)),
      empty = sum(gate_empty %in% TRUE),
      fallback = sum(gate_status %in% c("fallback_empty", "fallback_selected")),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      mismatch_type = unname(.simCompareMismatchLabels[mismatch_type]),
      method = unname(.analysis_method_labels[method])
    )
  print(knitr::kable(
    diag_summary,
    col.names = c(
      "Mismatch", "Shift", "Method", "Samples", "FDP defined", "Median FDP",
      "Median sensitivity", "Median selected", "Empty gates", "Fallbacks"
    )
  ))
  cat("\n\n")

  diag_stimgate <- diag_primary |>
    dplyr::filter(method == "stimgate") |>
    dplyr::left_join(diag_sample_gate, by = c("sim_id", "sample")) |>
    dplyr::transmute(
      mismatch = unname(.simCompareMismatchLabels[mismatch_type]),
      shift = mismatch_val,
      sample,
      final_gate = .analysis_label_number(threshold),
      sample_gate = .analysis_label_number(sample_gate),
      sample_gate_source,
      final_gate_source = gateReturnPoint,
      fdp = .analysis_label_percent(fdp),
      sensitivity = .analysis_label_percent(sensitivity),
      selected = .analysis_label_percent(selected_fraction),
      status = gate_status
    )
  print(knitr::kable(
    diag_stimgate,
    col.names = c(
      "Mismatch", "Shift", "Sample", "Final gate", "Sample-level gate",
      "Sample-level gate source", "Final gate source", "FDP", "Sensitivity",
      "Selected", "Gate status"
    )
  ))
}
```

When this diagnostic was developed, StimGate used the `hpi1` bandwidth
selector and a fixed unstimulated-sample bias (0.01 for gamma); it now uses
`nrd0` with the bias set to its initial bandwidth estimate, so check the
tables and figures above against this earlier result before relying on it.
With those earlier settings (five samples, production seeds), it showed the
following. Without a shift, every StimGate
gate lay in the gap between the components (about 1.6-1.7), with no false
discoveries and at least 99.8% sensitivity. At a shift of 0.05 the gates
became unstable: some fell into the right-hand tail of the stimulated
negatives (about 1.03, FDP about 20%, all positives detected), others jumped
into the positive component (2.8-3.2, sensitivity 1-4%). At a shift of 0.1,
in both variants, every gate lay inside the positive component (2.3-3.7):
FDP was 0, or undefined for one empty gate, and sensitivity was mostly below
2% (at most 41%). These were StimGate's own sample-level local-FDR gates,
which cluster refinement kept; they were not fallbacks (one sample-level
fallback was replaced by a calculated cluster gate). The jump in absolute
relative error to about 1 is therefore missed positives, an under-estimate
close to -100%, caused by a change in where StimGate places its gate, not
by negative contamination or a fallback. F-beta and Tailgate gates stayed
between the components (1.1-1.4), with FDP at most 0.2% and full
sensitivity. The tables above show the same quantities for the current run.

Each figure is one sample. Grey bars are the stimulated cells, stacked by
true label; the black outline is the unstimulated tube; vertical lines are
each method's final gate. Rows are shift sizes and columns the two shift
variants.

```{r}
#| label: gate-diagnostic-figures
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  dir_diag <- .compare_fresh_fig_dir("gate_diagnostic_clean_gamma")
  for (s in sort(unique(gate_diagnostic$cells$sample))) {
    .analysis_heading(paste0("Sample ", s), 3L)
    p <- .simComparePlotGateDiagnostic(
      gate_diagnostic$cells[gate_diagnostic$cells$sample == s, ],
      diag_primary[diag_primary$sample == s, ]
    )
    .analysis_save_fig(
      p, file.path(dir_diag, paste0("plot_sample_", s, ".pdf")),
      height = 16
    )
    .analysis_print_fig(p)
  }
}
```

## Secondary results: background-subtracted frequency error

The figures below describe the accuracy of the background-subtracted
frequency estimate, which answers a different question from the gate purity
and detection above. A gate can be pure yet miss most responding cells (a
large under-estimate with zero FDP), and an impure gate can still give an
accurate background-subtracted frequency if the unstimulated tube is
contaminated to the same extent. Both are needed to judge a method.

### Relative error summaries

The relative error of an estimate is (estimate - truth) / truth. Each scenario
is one baseline (transformation, mean position setting, response probability and
number of stimulated cells), one mismatch mechanism and one mismatch size. For
each scenario and method, the median, 95th percentile and maximum are
calculated over the simulated samples, using the same scored samples as the
summary table above. Samples with a true frequency of zero are left out. The
figures below then either average these scenario statistics equally over
cell counts and response probabilities, over response probabilities only, or
not at all. Mismatch sizes differ between transformations, so scenarios are
only ever averaged within one transformation and mismatch size.

Two kinds of figure are shown, each in three views. The first kind shows the
size of the relative error, ignoring direction. The second kind splits the
relative error into over-estimates (above zero) and under-estimates (below
zero), as in analysis 2a. Every figure is shown with all methods and without
Tailgate, with colour identifying the method, rows of panels showing the
median, 95th percentile and maximum, columns separating transformations, and
one figure per mean position setting, mismatch mechanism and (where
applicable) cell count.

```{r}
#| label: mismatch-error-summaries
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  scenario_error_cols <- c(
    "base_scenario_id", "transformation", "mean_pos_setting",
    "prob_response", "n_cell", "mismatch_type", "mismatch_val", "method"
  )
  # Scenario statistics, one row per scenario and method.
  err_scenario <- .simCompareUnsignedErrorSummary(
    compare_raw, scenarioCols = scenario_error_cols
  )
  signed_scenario <- .simCompareSignedErrorSummary(
    compare_raw, scenarioCols = scenario_error_cols
  )
  # Averaging drops the baseline scenario columns left out of the groups.
  group_all <- c(
    "mean_pos_setting", "transformation", "mismatch_type", "mismatch_val",
    "method"
  )
  group_cell <- c(group_all, "n_cell")
  err_averaged <- .simCompareErrorAverage(err_scenario, group_all)
  err_by_cell <- .simCompareErrorAverage(err_scenario, group_cell)
  signed_averaged <- .simBandwidthSignedErrorAverage(signed_scenario, group_all)
  signed_by_cell <- .simBandwidthSignedErrorAverage(signed_scenario, group_cell)
}
```

### Relative error averaged over cell counts

The size of the relative error is plotted against mismatch size, with scenario
statistics averaged equally over cell counts and response probabilities within
each transformation.

```{r}
#| label: relative-error-averaged
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    err_averaged,
    make_plot = .compare_mismatch_plot(.simComparePlotMismatchError),
    dir = .compare_fresh_fig_dir("relative_error_averaged"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 14,
    level = 4L
  )
}
```

### Relative error at each cell count

As above, but keeping each cell count separate and averaging only over response
probabilities.

```{r}
#| label: relative-error-by-n-cell
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    err_by_cell,
    make_plot = .compare_mismatch_plot(.simComparePlotMismatchError),
    dir = .compare_fresh_fig_dir("relative_error_by_n_cell"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"),
    height = 14,
    level = 4L
  )
}
```

### Relative error at each cell count and response probability

As above, but without averaging: each cell count and response probability is
shown separately, with the response probability given in the column heading.

```{r}
#| label: relative-error-by-n-cell-prob
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    err_scenario,
    make_plot = .compare_mismatch_plot(
      .simComparePlotMismatchError, by_prob = TRUE
    ),
    dir = .compare_fresh_fig_dir("relative_error_by_n_cell_prob"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"),
    height = 14,
    level = 4L
  )
}
```

### Over- and under-estimates averaged over cell counts

Relative error is split by direction: over-estimates sit above zero and
under-estimates below. Within each scenario, the median, 95th percentile and
maximum are calculated separately for over- and under-estimates, then averaged
equally over cell counts and response probabilities within each
transformation. Line weight shows the share of estimates in that direction.
Under-estimates are on a linear scale down to -100% (nothing gated), and
over-estimates on a fold scale, so -100% and +100% (twice the truth) are the
same distance from zero.

```{r}
#| label: signed-error-averaged
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    signed_averaged,
    make_plot = .compare_mismatch_plot(
      .simComparePlotSignedError, x = "mismatch_val", x_log = FALSE
    ),
    dir = .compare_fresh_fig_dir("signed_error_averaged"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 16,
    level = 4L
  )
}
```

### Over- and under-estimates at each cell count

As above, but keeping each cell count separate and averaging only over response
probabilities.

```{r}
#| label: signed-error-by-n-cell
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    signed_by_cell,
    make_plot = .compare_mismatch_plot(
      .simComparePlotSignedError, x = "mismatch_val", x_log = FALSE
    ),
    dir = .compare_fresh_fig_dir("signed_error_by_n_cell"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"),
    height = 16,
    level = 4L
  )
}
```

### Over- and under-estimates at each cell count and response probability

As above, but without averaging: each cell count and response probability is
shown separately, so large over-estimates can be traced to particular response
probabilities, which are given in the column headings.

```{r}
#| label: signed-error-by-n-cell-prob
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    signed_scenario,
    make_plot = .compare_mismatch_plot(
      .simComparePlotSignedError, x = "mismatch_val", x_log = FALSE,
      by_prob = TRUE
    ),
    dir = .compare_fresh_fig_dir("signed_error_by_n_cell_prob"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"),
    height = 16,
    level = 4L
  )
}
```

### Threshold placement diagnostic (unstimulated background intrusion)

The mean share of unstimulated cells above the gate is plotted against
the stimulated mean shift. Colour identifies the method, line type identifies
which stimulated cells are shifted, and panels separate baseline scenarios.
Higher values indicate that more unstimulated cells lie above the chosen gate.

```{r}
#| label: threshold-placement-unstim
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_shifts <- compare_summary |>
    dplyr::filter(
      mismatch_type %in% c("mean_shift_all", "mean_shift_negative")
    )
  .simCompareFigureLoop(
    summary_shifts,
    make_plot = .simComparePlotPlacement,
    dir = .analysis_fig_dir(c(fig_key, "threshold_placement_unstim"), path_root = root_dir),
    file_fn = function(pos, extra) paste0("plot_", pos, ".pdf"),
    height = 16,
    level = 4L
  )
}
```

### Upper-tail error analysis

The 90th percentile of absolute relative error is plotted against mismatch
size on an inverse-hyperbolic-sine vertical scale, with colour identifying
the method. Rows of panels separate mismatch mechanisms and columns separate
baseline scenarios, with one mean position setting per plot. Compare curves
within each panel, since mean shift and spread increase use different units.

```{r}
#| label: upper-tail-and-fallbacks
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    compare_summary,
    make_plot = .simComparePlotUpperTail,
    dir = .analysis_fig_dir(c(fig_key, "upper_tail_and_fallbacks"), path_root = root_dir),
    file_fn = function(pos, extra) paste0("plot_", pos, ".pdf"),
    height = 18,
    level = 4L
  )
}
```

## Rerun one scenario

A single scenario can be rerun with the production settings and RNG.
Run the setup, settings and grid chunks first.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_rerun <- 1L
row_rerun <- sim_grid_full[sim_grid_full$sim_id == sim_id_rerun, , drop = FALSE]
stopifnot(nrow(row_rerun) == 1L)
compare_one <- .simCompareRunScenario(
  row = row_rerun,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  clusterGates = cluster_gates,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  tailgateX = tailgate_x,
  resume = FALSE
)
```
