---
title: Compare background-subtracted frequency approaches under condition mismatch
format:
  html:
    embed-resources: true
    toc: true
knitr:
  opts_chunk:
    fig.retina: 1
execute:
  warning: false
  message: false
params:
  analysis_run_id: null
  simulation_seed: 12345
  sim_grid_chunk_index: 1
  sim_grid_n_chunks: 1
  run_simulations: true
  run_plots: false
  show_mcse: "on"
---

This analysis compares StimGate, F-beta and Tailgate when stimulated and
unstimulated expression distributions are misaligned. Paired samples are
simulated for selected baseline scenarios, then the stimulated expression is
shifted or its background spread is increased. The same simulated draws are
used across mismatch settings within each baseline scenario.

The question is whether misalignment between the stimulated and unstimulated
negative components makes a method call cells in the right-hand tail of the
stimulated negative component positive. The main outcomes are therefore the
purity and the detection rate of each method's final gate in the stimulated
tube, measured against the simulation's own labels:

- **False discovery proportion (FDP):** of the stimulated cells classified as
  positive, the proportion that are genuinely negative.
- **Sensitivity:** of the genuinely positive stimulated cells, the proportion
  classified as positive.

Both lie between 0 and 1 in every scenario. The relative error of the
background-subtracted frequency, which becomes very large when the true
response is rare, is kept as a secondary result.

## Set-up

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)
library(dplyr)
library(tidyr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
analysis_qmd <- "analysis/8-sim-compare-freq_bs-batch.qmd"

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))
analysis_run_id <- .get_qmd_param_env(
  "analysis_run_id",
  "ANALYSIS_RUN_ID",
  NULL
)
analysis_key <- c("sim", "compare", "freq_bs_batch")
simulation_seed <- as.integer(.get_qmd_param_env(
  "simulation_seed",
  "SIMULATION_SEED",
  12345L
))
if (is.na(simulation_seed)) {
  stop("simulation_seed must be coercible to an integer.")
}
comparison_semantics_version <- "batch-mismatch-comparison-v16"
analysis_dev <- .analysis_is_dev()
analysis_quick <- .analysis_is_quick() && !analysis_dev
# "draft" (current default) or "final" sample/dataset counts; dev and quick win.
sim_size <- .analysis_sim_size()
analysis_key <- .analysis_mode_key(analysis_key)
fig_key <- .analysis_mode_key("8-sim-compare-freq_bs-batch")
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
# Monte Carlo error bars on the summary plots (see analysis-mcse.R).
mcse_mode <- .analysis_mcse_mode(.get_qmd_param_env("show_mcse", "SHOW_MCSE", "on"))
show_mcse <- mcse_mode != "off"

desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  stop("Expected the stimgate DESCRIPTION at the repository root: ", root_dir)
}

if (isTRUE(run_simulations)) {
  .analysis_require_packages(c("cytoUtils", "openCyto", "flowStats", "simcyto"))
}

# Source simulation helpers in dependency order:
# 1. sim-misc.R: general simulation utilities and transformation lookups
# 2. sim-bandwidth.R: location details reader and bandwidth helpers
# 3. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-bandwidth-analysis-io.R",
  "sim-bandwidth-analysis-plot.R",
  "sim-compare-freq_bs.R",
  "sim-compare-performance-plot.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

```

## Scientific settings

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches (`fbeta` and `tailgate`) under controlled
**condition mismatch** (technical misalignment between the stimulated and
matched unstimulated tubes).

The F-beta implementation retains the Richards et al. (2014) scoring procedure
and standard parameters, but adapts its histogram support to span both paired
distributions. This prevents stimulated response-tail events outside the
observed unstimulated range from being omitted from the F-beta calculation.

The simulation isolates two deterministic post-transformation mismatch
mechanisms provided by `simcyto`:

1. **Mean shift** (`stimMeanShift`): An additive shift on the transformed
   expression scale, applied either to all stimulated cells or only to the
   stimulated negative/background component. Shifts go up to 0.35 for the
   Gaussian and Skew transformations and up to 0.2 for Gamma.
2. **SD inflation** (`stimSdMultiplier`): A multiplicative scaling of the
   stimulated negative/background component's within-component standard
   deviation around its realised mean, increasing it by up to 35%.

The two mean-shift variants are different mechanisms. Shifting all
stimulated cells misaligns the background while keeping the separation
between positive and negative cells within the stimulated tube. Shifting only
the stimulated negatives misaligns the background and also moves it towards
the unchanged positive component. Neither is assumed to degrade performance
steadily as the mismatch grows.

Random perturbations are disabled
(`samplePerturbationSd = 0`, `conditionPerturbationSd = 0`,
`clusterPerturbationSd = 0`). Within each baseline scenario, all mismatch
magnitudes and variants use the same random seed, so the degradation curves
compare deterministic changes to the same simulated draws. Each replicate
draws its own seed from that scenario seed before any method runs, so later
replicates stay paired even though the methods use different amounts of
randomness in different mismatch settings.

StimGate bandwidths are shared per marker (`bwScope = "cytokine"`), using
the `nrd0` selector with the bandwidth-estimation sample capped at 10,000 cells.

### Simulation settings

Each scenario simulates 20 datasets of 20 samples in final runs and 5 datasets
of 20 samples in draft runs (`SIM_SIZE=draft`). The samples in a dataset are
gated together, with shared settings and cluster gates, so the dataset is the
unit for uncertainty. Draft results are kept separately and are for iterating,
not for reporting.

```{r}
#| label: actual-settings
# Method settings inherited from analysis 7:
stimgate_bw_mtd <- "nrd0"
stimgate_bw_scope <- "cytokine"
# StimGate's unstimulated-sample bias equals its initial bandwidth estimate:
# with biasUns = NULL, gateStim() uses 0.25 * bwFallback * biasUnsFactor.
stimgate_bias_uns_factor <- 4
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# Cluster refinement enabled for the full one-marker procedure
# Share thresholds across clusters of similar samples via stimControl(clusterGates = ).
cluster_gates <- TRUE
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE
# Each dataset of `n_sample_sim` samples is gated together (shared settings
# and cluster gates), so the dataset is the unit for uncertainty. Draft keeps
# the samples per dataset and uses a quarter of the datasets (at least 5).
n_sample_sim <- if (analysis_quick) 1L else 20L
n_iter_sim <- if (analysis_quick) 1L else c(final = 20L, draft = 5L)[[sim_size]]

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- file.path(
  root_dir,
  "scripts",
  "python",
  "fbeta.py"
)

if (isTRUE(run_simulations)) {
  if (!file.exists(path_fbeta)) {
    stop("F-beta comparison script not found: ", path_fbeta)
  }
  .simCompareFbetaEnvironment(pathFbeta = path_fbeta)
}

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
tailgate_x <- "stim"

background_relative_to_response <- 0.2
n_cell_uns_relative_to_stim <- 1

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

# Targeted scenario panel representing qualitatively distinct baselines
targeted_scenarios_tbl <- tibble::tribble(
  ~transformation, ~mean_pos_setting, ~prob_response, ~n_cell, ~scenario_desc,
  "gaussian", "high", 0.002, 20000, "Rare response, clean baseline",
  "skew", "high", 0.2, 20000, "Clean skew baseline",
  "gamma", "high", 0.2, 5000, "Very clean gamma baseline",
  "gaussian", "low", 0.2, 1000, "Overlapping Gaussian (StimGate advantage over F-beta)",
  "gamma", "low", 0.2, 1000, "Overlapping Gamma (StimGate advantage over F-beta)",
  "skew", "high", 0.0005, 100000, "Rare skew (F-beta advantage)",
  "gamma", "high", 0.0005, 20000, "Rare Gamma (F-beta gate-purity advantage)"
) |>
  dplyr::left_join(
    mean_pos_settings_tbl,
    by = c("transformation", "mean_pos_setting")
  ) |>
  dplyr::mutate(
    base_scenario_id = dplyr::row_number(),
    bias_uns = NA_real_,
    bias_uns_factor = stimgate_bias_uns_factor,
    bw_mtd = stimgate_bw_mtd,
    bw_scope = stimgate_bw_scope,
    bw_ncell_max = stimgate_bw_ncell_max,
    bw_fallback = stimgate_bw_fallback,
    bw_min = stimgate_bw_min,
    bw_max = stimgate_bw_max,
    sample_perturbation_sd = 0,
    condition_perturbation_sd = 0,
    cluster_perturbation_sd = 0,
    background_relative_to_response = background_relative_to_response,
    n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim
  )
```

## Simulation

### Mismatch experiment grids

```{r}
#| label: mismatch-grids
#| include: false
# 1. Mean-shift grid (all components):
# Gaussian and Skew: 0 to 0.35
# Gamma: 0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2
mean_shift_gaussian_skew <- c(0, 0.025, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35)
mean_shift_gamma <- c(0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2)

grid_mean_shift_all <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_all",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = NA_character_,
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 2. Mean-shift grid (negative/background component only):
grid_mean_shift_negative <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_negative",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = "gn",
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 3. SD-inflation grid (negative/background component only):
# Fractional SD increase: 0 to 0.35, matching the Gaussian and Skew mean shifts
sd_increase_vec <- c(0, 0.025, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35)

grid_sd_inflation <- tidyr::expand_grid(
  targeted_scenarios_tbl,
  sd_increase = sd_increase_vec
) |>
  dplyr::mutate(
    mismatch_type = "sd_inflation",
    mismatch_val = sd_increase,
    stim_mean_shift = 0,
    stim_mean_shift_clusters = NA_character_,
    stim_sd_multiplier = 1 + sd_increase,
    stim_sd_multiplier_clusters = "gn"
  )

# Combined full mismatch simulation grid
sim_grid <- dplyr::bind_rows(
  grid_mean_shift_all,
  grid_mean_shift_negative,
  grid_sd_inflation
) |>
  dplyr::mutate(
    sim_id = dplyr::row_number(),
    # Use common random numbers within each baseline scenario so that
    # degradation across mismatch magnitudes/variants is paired.
    sim_seed = as.integer(simulation_seed + base_scenario_id - 1L)
  )

# Keep every seeded row available for an exact interactive rerun.
sim_grid_full <- sim_grid

# Gate diagnostic: examine the very clean gamma baseline across selected
# mean shifts for both shift variants. The rows
# come from the full grid, so they keep their production seeds.
gate_diagnostic_spec <- list(
  base_scenario_id = 3L,
  mismatch_type = c("mean_shift_all", "mean_shift_negative"),
  mismatch_val = c(0, 0.05, 0.1)
)
gate_diagnostic_grid <- sim_grid_full |>
  dplyr::filter(
    base_scenario_id == gate_diagnostic_spec$base_scenario_id,
    mismatch_type %in% gate_diagnostic_spec$mismatch_type,
    mismatch_val %in% gate_diagnostic_spec$mismatch_val
  )
stopifnot(
  nrow(gate_diagnostic_grid) ==
    length(gate_diagnostic_spec$mismatch_type) *
      length(gate_diagnostic_spec$mismatch_val)
)

# Quick mode keeps baseline scenario, mismatch magnitude (including zero) and mechanism.
# Smallest grid covering every figure; dev takes precedence.
if (analysis_quick) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      base_scenario_id %in% c(4L, 5L),
      mismatch_val %in% c(0, 0.05)
    )
}

if (analysis_dev) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      transformation == "gaussian",
      mean_pos_setting == "high",
      prob_response == 0.002,
      mismatch_val %in% c(0, 0.05)
    )
}

sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

sim_grid_all <- sim_grid
analysis_grid_spec <- sim_grid_all |>
  dplyr::select(-dplyr::any_of("sim_seed"))

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

print(paste0(
  "Simulation scenarios assigned to chunk ",
  sim_chunk_label,
  ": ",
  nrow(sim_grid),
  " of ",
  nrow(sim_grid_all)
))
```

### Simulation execution

```{r}
#| label: compare-batch-run
#| include: false
n_workers <- max(1, .simGetCores())

analysis_result_params <- list(
  comparison_semantics_version = comparison_semantics_version,
  simulation_seed = simulation_seed,
  analysis_quick = analysis_quick,
  analysis_dev = analysis_dev,
  sim_size = sim_size,
  paired_mismatch_rng = TRUE,
  analysis_grid_spec = analysis_grid_spec,
  gate_diagnostic_spec = gate_diagnostic_spec,
  n_sample_sim = n_sample_sim,
  n_iter_sim = n_iter_sim,
  stimgate_bw_mtd = stimgate_bw_mtd,
  stimgate_bw_scope = stimgate_bw_scope,
  stimgate_bias_uns_factor = stimgate_bias_uns_factor,
  stimgate_bw_ncell_max = stimgate_bw_ncell_max,
  stimgate_bw_fallback = stimgate_bw_fallback,
  stimgate_bw_min = stimgate_bw_min,
  stimgate_bw_max = stimgate_bw_max,
  cluster_gates = cluster_gates,
  loc_enforce_shape_threshold = loc_enforce_shape_threshold,
  calc_cyt_pos_gates = calc_cyt_pos_gates,
  fbeta_beta = fbeta_beta,
  fbeta_theta = fbeta_theta,
  fbeta_width = fbeta_width,
  tailgate_adjust = tailgate_adjust,
  tailgate_method = tailgate_method,
  tailgate_tol = tailgate_tol,
  tailgate_auto_tol = tailgate_auto_tol,
  tailgate_x = tailgate_x
)

if (isTRUE(run_simulations)) {
  run_ctx <- .analysis_run_context(
    analysis_key = analysis_key,
    run_id = analysis_run_id,
    qmd_path = analysis_qmd,
    path_root = root_dir,
    params = c(
      list(
        run_simulations = run_simulations,
        run_plots = run_plots,
        sim_grid_chunk_index = sim_grid_chunk_index,
        sim_grid_n_chunks = sim_grid_n_chunks
      ),
      analysis_result_params
    ),
    sim_grid_chunk_index = sim_grid_chunk_index,
    sim_grid_n_chunks = sim_grid_n_chunks
  )

  path_progress_file <- run_ctx$progress_file
  dir_output <- run_ctx$chunk_output_dir
  dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)
  dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

  path_rds <- file.path(
    run_ctx$staging_collated_dir,
    if (sim_grid_n_chunks > 1L) {
      sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
    } else {
      "compare_raw.rds"
    }
  )
} else {
  run_ctx <- .analysis_results_context(
    analysis_key = analysis_key,
    path_root = root_dir,
    qmd_path = analysis_qmd
  )
}

if (interactive()) {
  progressr::handlers(global = TRUE)
}

compare_raw <- tibble::tibble()

if (isTRUE(run_simulations) && nrow(sim_grid) > 0L) {
  compare_raw <- .simCompareFreqBsGrid(
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    probExact = TRUE,
    covEvMin = 1.5,
    covEvMax = 1.5,
    clusterGates = cluster_gates,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates,
    pathFbeta = path_fbeta,
    fbetaBeta = fbeta_beta,
    fbetaTheta = fbeta_theta,
    fbetaWidth = fbeta_width,
    tailgateAdjust = tailgate_adjust,
    tailgateAutoTol = tailgate_auto_tol,
    tailgateMethod = tailgate_method,
    tailgateTol = tailgate_tol,
    tailgateX = tailgate_x,
    parallel = (n_workers > 1L),
    workers = n_workers,
    dirCache = dir_output,
    pathProgress = path_progress_file,
    dirJobs = run_ctx$chunk_jobs_dir,
    progressHeading = "BATCH MISMATCH COMPARISON SIMULATION PROGRESS",
    resume = TRUE,
    retryErrors = TRUE,
    sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
      sim_grid_chunk_index
    } else {
      NULL
    },
    sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
      sim_grid_n_chunks
    } else {
      NULL
    }
  )

  .write_rds_atomic(compare_raw, path_rds)

} else if (isTRUE(run_simulations)) {
  message("No simulations assigned to this chunk.")
} else {
  message("Skipping simulations because run_simulations is FALSE.")
}

# The gate diagnostic needs the cells, which the grid run does not keep, so
# the first chunk reruns its few rows for one replicate and saves them with
# the run. It runs before the chunk is marked complete, so a promoted run
# always includes it.
path_gate_diagnostic <- NULL
if (isTRUE(run_simulations) && sim_grid_chunk_index == 1L) {
  path_gate_diagnostic <- file.path(
    run_ctx$staging_collated_dir, "gate_diagnostic.rds"
  )
  if (!file.exists(path_gate_diagnostic)) {
    gate_diagnostic <- .simCompareGateDiagnosticRun(
      gate_diagnostic_grid,
      nSample = n_sample_sim,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      probExact = TRUE,
      covEvMin = 1.5,
      covEvMax = 1.5,
      clusterGates = cluster_gates,
      locEnforceShapeThreshold = loc_enforce_shape_threshold,
      calcCytPosGates = calc_cyt_pos_gates,
      pathFbeta = path_fbeta,
      fbetaBeta = fbeta_beta,
      fbetaTheta = fbeta_theta,
      fbetaWidth = fbeta_width,
      tailgateAdjust = tailgate_adjust,
      tailgateAutoTol = tailgate_auto_tol,
      tailgateMethod = tailgate_method,
      tailgateTol = tailgate_tol,
      tailgateX = tailgate_x
    )
    .write_rds_atomic(gate_diagnostic, path_gate_diagnostic)
  }
}

simulation_promoted <- FALSE

if (isTRUE(run_simulations)) {
  chunk_status <- .simCompareGridOutputStatus(
    compare_raw,
    sim_grid = sim_grid,
    nSample = n_sample_sim,
    nIter = n_iter_sim
  )
  n_failed <- length(chunk_status$failed_ids)
  n_completed <- length(chunk_status$completed_ids)

  chunk_error <- if (!isTRUE(chunk_status$validation_ok)) {
    paste0(
      "Analysis 8 chunk output is incomplete or invalid. Missing sim_id: ",
      paste(chunk_status$missing_ids, collapse = ", "),
      "; failed sim_id: ",
      paste(chunk_status$failed_ids, collapse = ", "),
      "; reasons: ", paste(names(chunk_status$failure_reasons), chunk_status$failure_reasons, collapse = "; "),
      "; extra sim_id: ",
      paste(chunk_status$extra_ids, collapse = ", "),
      "."
    )
  } else {
    NULL
  }

  .analysis_mark_chunk(
    run_ctx = run_ctx,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    collate_ok = chunk_status$collate_ok,
    validation_ok = chunk_status$validation_ok,
    error_message = chunk_error
  )

  if (!isTRUE(chunk_status$validation_ok)) {
    stop(chunk_error)
  }

  simulation_promoted <- .simComparePromoteIfReady(
    run_ctx = run_ctx,
    sim_grid_all = sim_grid_all,
    total_sims = nrow(sim_grid),
    completed_sims = n_completed,
    failed_sims = n_failed,
    nSample = n_sample_sim,
    nIter = n_iter_sim,
    validate_full = .simCompareValidateMismatch
  )

  if (!isTRUE(run_plots)) {
    message("run_plots is false, so stopping after simulation/collation.")
    knitr::knit_exit()
  }

  if (sim_grid_n_chunks > 1L) {
    message(
      "Skipping plots during a multi-chunk simulation render. ",
      "Render again with run_simulations = FALSE after promotion."
    )
    knitr::knit_exit()
  }

  if (!isTRUE(simulation_promoted)) {
    message("The simulation run has not been promoted; skipping summary and plots.")
    knitr::knit_exit()
  }
}
```

## Collation

```{r}
#| label: compare-summary
#| echo: false
path_rds <- .analysis_current_file(
  run_ctx = run_ctx,
  relative_path = c("collated", "compare_raw.rds"),
  required_params = analysis_result_params
)
compare_raw <- .analysis_read_rds(path_rds, analysis_key, analysis_qmd)

if (nrow(compare_raw) == 0L) {
  stop("Canonical compare_raw.rds is empty. Rerun the simulation analysis.")
}

canonical_status <- .simCompareGridOutputStatus(
  compare_raw,
  sim_grid = sim_grid_all,
  nSample = n_sample_sim,
  nIter = n_iter_sim
)
if (!isTRUE(canonical_status$validation_ok)) {
  stop(
    "Canonical compare_raw.rds does not contain a complete valid result ",
    "for every analysis 8 scenario."
  )
}

results_available <- TRUE

scenario_cols <- c(
  "base_scenario_id",
  "mismatch_type",
  "mismatch_val",
  "stim_mean_shift",
  "stim_mean_shift_clusters",
  "stim_sd_multiplier",
  "stim_sd_multiplier_clusters",
  "sd_increase",
  "transformation",
  "mean_pos_setting",
  "mean_pos",
  "prob_response",
  "n_cell",
  "scenario_desc",
  "approach",
  "method"
)

compare_summary <- .simCompareSummariseFreqBs(
  compare_raw,
  scenarioCols = scenario_cols,
  mcse = show_mcse
)

compare_summary
```

## Validation checks

Three checks run on the full collated results before promotion and again before
any figure is drawn; the render stops if one fails.

- **Counts match the gates.** For every stimulated tube, replicate and method,
  the confusion-matrix counts must add up to the number of stimulated cells,
  and the cells above the gate (true plus false positives) must equal the
  number of stimulated cells the method itself gated. Results that lack the
  counts are rejected as incomplete.
- **Pairing across mismatch settings.** Within each baseline scenario,
  replicate and sample, every mismatch setting must have the same
  unstimulated tube (which no mismatch changes), the same number of genuine
  positives and the same number of stimulated cells.
- **Zero mismatch.** With no mismatch, shifting only the stimulated negatives
  must give exactly the same gates and counts as shifting all stimulated
  cells. Zero SD inflation is compared too; it rescales the background by
  one, which can change the last bit of a value, so it is reported rather
  than required to match exactly.

```{r}
#| label: validation-checks
#| echo: false
mismatch_validation <- .simCompareValidateMismatch(compare_raw)
pairing_check <- mismatch_validation$pairing_check
zero_agreement <- mismatch_validation$zero_agreement

knitr::kable(
  .simCompareMethodOutcomeCounts(compare_raw) |>
    dplyr::select(method, n, n_valid, n_run_error, n_no_cutpoint, n_fallback),
  caption = "Method outcomes: runtime errors are unscored; no-cutpoint empty gates are recorded fallbacks."
)

cat(
  "Counts reproduce the gated stimulated counts in all",
  sum(compare_raw$method %in% c("stimgate", "fbeta", "tailgate") &
    !.simCompareRecordedComparatorErrors(compare_raw)),
  "successful tube-level results; recorded comparator failures remain missing.\n"
)
cat(
  "Paired data in all", nrow(pairing_check),
  "baseline-scenario, replicate and sample groups.\n"
)
knitr::kable(
  zero_agreement |>
    dplyr::mutate(
      mismatch_type = unname(.simCompareMismatchLabels[mismatch_type]),
      method = unname(.analysis_method_labels[method])
    ),
  col.names = c(
    "Zero-mismatch variant", "Method", "Pairs", "Failed pairs", "Compared", "Same gate",
    "Same counts", "Largest gate difference"
  ),
  caption = paste(
    "Agreement of each zero-mismatch variant with the zero shift of all",
    "stimulated cells, on the same replicates. Failed pairs retain missing outcomes; only finite pairs are compared."
  )
)

if (!isTRUE(run_plots)) {
  message("run_plots is false, so stopping after the canonical comparison summary.")
  knitr::knit_exit()
}
```

## Results

Every comparison figure is shown twice. "All methods" includes StimGate,
Tailgate and F-beta. "Without Tailgate" leaves Tailgate out, because it
performs very poorly without more tuning and squashes the scale for the other
two methods. The two versions are saved in the `all_methods` and `no_tailgate`
subfolders of each figure folder. Within each version there is one figure per
mean position setting and mismatch mechanism.

Mean shifts are additive shifts on the transformed expression scale, the
scale the methods gate on. SD inflation is the fractional increase in the
standard deviation of the stimulated background (negative) cells.

**Monte Carlo intervals.** `show_mcse` / `SHOW_MCSE` accepts `off` or `on`
(default); false/true still selects off/on. Each render prints one version
and saves it under `mcse_off/` or `mcse_on/`, with identical points and axes. With error bars on, 95% percentile-bootstrap
intervals resample independent whole datasets (`iter`) with all selected samples
and their multiplicity. Each draw recalculates the actual pooled tube statistic
shown, including classification distributions. Point estimates are unchanged
when intervals are disabled. At least five contributing datasets and 95% finite
bootstrap draws are required; coverage and suppressed intervals are reported.
Final runs use 20 datasets of 20 samples and draft runs use five datasets of 20
samples, so draft intervals are rough. Tiny quick-mode datasets are smoke checks.
Averaged figures recompute the full equal-weight scenario average in each draw,
with the same dataset indices for methods and deterministic mismatch settings
sharing biological random draws. They do not combine scenario SEs as though
paired settings were independent. Maxima have their own occurrence and
conditional-severity figures below.

```{r}
#| label: figure-helpers
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  # Preserve the sibling MC mode from the other render.
  .compare_fresh_fig_dir <- function(name) {
    .analysis_fig_dir(c(fig_key, name), path_root = root_dir)
  }
  .compare_mismatch_heading <- function(extra) {
    type <- if (is.list(extra)) extra$mismatch_type else extra
    out <- paste0("Mismatch: ", unname(.simCompareMismatchLabels[type]))
    if (is.list(extra)) {
      out <- paste0(out, "; cells: ", .analysis_label_number(extra$n_cell))
    }
    out
  }
  .compare_mismatch_file <- function(pos, extra) {
    type <- if (is.list(extra)) extra$mismatch_type else extra
    cells <- if (is.list(extra)) paste0("_n_cell_", extra$n_cell) else ""
    paste0("plot_", type, "_mean_pos_setting_", pos, cells, ".pdf")
  }
  .compare_mismatch_x <- function(type) {
    unname(.simCompareMismatchAxisLabels[type])
  }
  .compare_baseline_note <- function(d) {
    pos <- unique(d$mean_pos_setting)
    baselines <- targeted_scenarios_tbl |>
      dplyr::filter(
        base_scenario_id %in% compare_raw$base_scenario_id,
        mean_pos_setting == pos, transformation %in% d$transformation
      )
    for (col in intersect(c("base_scenario_id", "n_cell", "prob_response"), names(d))) {
      baselines <- baselines[baselines[[col]] %in% d[[col]], , drop = FALSE]
    }
    cat("\n\n", paste0(
      "Baselines: ", paste0(
        baselines$base_scenario_id, " (", baselines$transformation,
        ", cells=", .analysis_label_number(baselines$n_cell),
        ", response=", .analysis_label_percent(baselines$prob_response), ")",
        collapse = "; "
      )
    ), ".\n\n", sep = "")
  }
  # Plot function for one figure: the x axis label depends on the mechanism.
  .compare_mismatch_plot <- function(plot_fn, ...) {
    function(d) {
      .compare_baseline_note(d)
      if ("n_scenario_median" %in% names(d)) {
        count_cols <- c(
          "transformation", "mismatch_val", "method", "n_cell", "direction",
          "n_scenario_median", "n_scenario_q95"
        )
        cat("\n\nContributing scenarios per statistic:\n\n",
            knitr::kable(dplyr::select(d, dplyr::any_of(count_cols))), sep = "\n")
        plot_fn(
          d, x_label = .compare_mismatch_x(d$mismatch_type[[1]]),
          stat_cols = c(
            median = "Mean of scenario medians",
            q95 = "Mean of scenario 95th percentiles"
          ), ...
        ) + ggplot2::labs(caption = .simBandwidthScenarioCaption(d))
      } else {
        plot_fn(d, x_label = .compare_mismatch_x(d$mismatch_type[[1]]), ...)
      }
    }
  }
}
```

## Gate purity and detection

For every stimulated tube, replicate and method, the method's final gate is
applied to the stimulated cells and compared with the simulation labels. As
in the package, a cell is positive when its expression is strictly above the
gate; a cell exactly at the gate is negative. StimGate is scored with its
final gate, which is the cluster-refined gate where clustering applies.
Positive cells that are part of the background in the stimulated tube count
as genuine positives: these outcomes describe the purity and detection rate
of the stimulated gate, not which positive cells were induced by
stimulation. The background-subtracted frequency is a separate estimand and
is covered in the secondary results.

From the counts of true positives (TP), false positives (FP), false
negatives (FN) and true negatives (TN):

- FDP = FP / (TP + FP), undefined when the gate selects no stimulated cells;
- sensitivity = TP / (TP + FN), so an empty gate has zero sensitivity;
- false-positive rate = FP / (FP + TN);
- selected fraction = (TP + FP) / (TP + FP + FN + TN).

Each stimulated tube in each dataset gives one value of each outcome. The
pooled quantiles describe tube-level distributions; cell counts are never
pooled across tubes. Tubes within a dataset share gating settings and are not
independent replicates. **FDP summaries leave out tubes whose gate selected no
cells**, because FDP is undefined there; the tables below count these tubes.
FDP and sensitivity are always shown together, so a
gate that selects almost nothing cannot look successful merely because it
contains few negatives.

```{r}
#| label: classification-summary
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  classification_cols <- c(
    "base_scenario_id", "scenario_desc", "transformation", "mean_pos_setting",
    "prob_response", "n_cell", "mismatch_type", "mismatch_val", "method"
  )
  class_scenario <- .simCompareClassificationSummary(
    compare_raw,
    scenarioCols = classification_cols, mcse = show_mcse
  )
  print(knitr::kable(.simComparePerformanceCoverage(class_scenario, classification_cols),
    caption = "Tube-level classification bootstrap coverage and unavailable intervals."))
}
```

### Paired differences between datasets

The independent unit is the simulated dataset (`iter`), whose tubes are gated
with shared settings and cluster gates. Within each scenario and dataset, we
compute each method's median FDP and median sensitivity, then subtract the
competitor's value from
StimGate's value. Each dataset has equal weight. The tables report the mean
paired difference and a 95% interval, mean ± 1.96 SD / √D, using the spread
between the D datasets with finite paired summaries. Intervals are NA for D < 5.
Dataset-pair coverage and finite-tube coverage are reported explicitly;
undefined outcomes are omitted from the corresponding method summary and a
dataset with no finite outcomes has no paired difference.
Negative FDP differences and positive sensitivity differences favour StimGate.

```{r}
#| label: dataset-paired-differences
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  dataset_differences <- .simCompareDatasetDifferences(
    compare_raw, scenarioCols = classification_cols,
    outcomes = c("fdp", "sensitivity")
  )
  for (set in .simCompareMethodSets()) {
    .analysis_heading(set$heading, 4L)
    for (pos in unique(dataset_differences$mean_pos_setting)) {
      .analysis_heading(paste0("Mean position: ", pos), 5L)
      table <- dataset_differences |>
        dplyr::filter(.data$method %in% set$methods, .data$mean_pos_setting == pos) |>
        dplyr::mutate(
          interval_95 = sprintf("[%.4f, %.4f]", .data$lower, .data$upper),
          paired_datasets = paste0(.data$n_pair, " / ", .data$n_dataset)
        ) |>
        dplyr::select(dplyr::all_of(c(
          "scenario_desc", "mismatch_type", "mismatch_val",
          "outcome", "method", "paired_datasets", "mean_difference", "interval_95",
          "tube_coverage_stimgate", "tube_coverage_competitor"
        )))
      print(knitr::kable(table, digits = 4,
        caption = "Dataset-level paired differences: StimGate minus competitor."))
      cat("\n\n")
    }
  }
}
```

### FDP and sensitivity against mismatch size

Each scenario occupies a row with FDP and sensitivity side by side, both on
fixed 0-100% scales. Solid lines are tube-level medians; dashed lines
are the worse tail, the 90th percentile of FDP and the 10th percentile of
sensitivity. Wrapped strips name each baseline scenario, each with its own horizontal
range (on a square-root scale) because mismatch sizes differ between
transformations. Each mismatch mechanism has its own figure. Error bars (with
`show_mcse` on) resample whole datasets and recalculate the pooled tube-level
median or tail percentile, as described above. Undefined bootstrap draws and
suppressed intervals are reported in the coverage table.

```{r}
#| label: fdp-sensitivity
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    class_scenario,
    make_plot = .compare_mismatch_plot(
      .simComparePlotClassification,
      outcomes = c("fdp", "sensitivity"), mcse = show_mcse
    ),
    dir = .compare_fresh_fig_dir("fdp_sensitivity"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 13,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### Observation counts

Each cell gives, for one method and setting, the number of valid tubes
(no run error), the number with a defined FDP (the gate selected at least one
cell), the number of empty gates and the number of gates that used a
threshold fallback. Fallback gates are included in the outcomes above, as
they are the gates the methods applied.

```{r}
#| label: classification-counts
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  for (type in intersect(names(.simCompareMismatchLabels), class_scenario$mismatch_type)) {
    .analysis_heading(
      paste0("Mismatch: ", unname(.simCompareMismatchLabels[type])), 4L
    )
    print(knitr::kable(
      .simCompareClassificationCountTable(
        class_scenario[class_scenario$mismatch_type == type, ]
      ),
      caption = "Tube counts: valid / FDP defined / empty gates / fallbacks."
    ))
    cat("\n\n")
  }
}
```

### False-positive rate

The false-positive rate is the share of the genuinely negative stimulated
cells that lie above the gate. It is a secondary diagnostic. FDP depends on
how common genuine positives are as well as on the gate: the same
false-positive rate gives a much larger FDP when the response is rare, so
differences in FDP between rare- and common-response scenarios are not all
due to the methods. The vertical scales here are not fixed, because the
rates are small. Each panel has independent ranges, arranged in two columns.

```{r}
#| label: false-positive-rate
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    class_scenario,
    make_plot = .compare_mismatch_plot(
      .simComparePlotClassification,
      outcomes = "fpr", unit_scale = FALSE, mcse = show_mcse
    ),
    dir = .compare_fresh_fig_dir("false_positive_rate"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 9,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

## Very clean gamma baseline: gate diagnostic

This diagnostic shows all 20 samples of replicate dataset 1 (`iter == 1`)
of base scenario 3, “Very clean gamma baseline”: gamma transformation, high
separation, response probability 0.2 and 5000 stimulated cells per sample.
It is a deliberate rerun of the main run's first replicate with production
seeds, not a random selection of samples. The same simulated dataset is used
for shifts 0, 0.05 and 0.1, applied either to all stimulated cells or only to
stimulated negatives. The diagnostic keeps the cells and each method's final
gates during simulation and saves them with the results; nothing is simulated
here. Quick mode uses one sample instead of 20, and quick/dev main grids can
omit this baseline, in which case there are no main-run rows to compare.

```{r}
#| label: gate-diagnostic-data
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  path_gate_diagnostic <- .analysis_current_file(
    run_ctx = run_ctx,
    relative_path = c("collated", "gate_diagnostic.rds"),
    required_params = analysis_result_params
  )
  gate_diagnostic <- .analysis_read_rds(
    path_gate_diagnostic, analysis_key, analysis_qmd
  )
  diag_results <- gate_diagnostic$results
  diag_primary <- diag_results |>
    dplyr::filter(method %in% c("stimgate", "fbeta", "tailgate")) |>
    .simCompareClassificationMetrics()
  # StimGate's sample-level gate, before cluster refinement.
  diag_sample_gate <- diag_results |>
    dplyr::filter(method == "stimgate_loc_sample") |>
    dplyr::select(
      sim_id, sample, sample_gate = threshold,
      sample_gate_source = gateReturnPoint
    )

  # Where the main results contain the same rows, the diagnostic replicate
  # must reproduce their first replicate exactly.
  diag_match <- diag_primary |>
    dplyr::select(sim_id, sample, method, threshold) |>
    dplyr::inner_join(
      compare_raw |>
        dplyr::filter(iter == 1L) |>
        dplyr::select(sim_id, sample, method, threshold_main = threshold),
      by = c("sim_id", "sample", "method")
    )
  if (nrow(diag_match) > 0L) {
    cat(
      "The diagnostic replicate reproduces the first replicate of the main",
      "results for", sum(diag_match$threshold == diag_match$threshold_main),
      "of", nrow(diag_match), "matching gates.\n"
    )
  } else {
    cat(
      "The main results of this run do not include the diagnostic scenario,",
      "so the diagnostic replicate is not compared with them.\n"
    )
  }
}
```

The table summarises each setting over samples: median FDP (over samples
with a defined FDP), median sensitivity and median selected fraction, with
the number of empty gates and fallbacks. The second table lists StimGate's
gates for every sample, with the sample-level gate chosen before cluster
refinement and how it was chosen.

```{r}
#| label: gate-diagnostic-tables
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  q <- .simCompareQuantileFinite
  diag_summary <- diag_primary |>
    dplyr::group_by(mismatch_type, mismatch_val, method) |>
    dplyr::summarise(
      samples = dplyr::n(),
      fdp_defined = sum(is.finite(fdp)),
      fdp = .analysis_label_percent(q(fdp, 0.5)),
      sensitivity = .analysis_label_percent(q(sensitivity, 0.5)),
      selected = .analysis_label_percent(q(selected_fraction, 0.5)),
      empty = sum(gate_empty %in% TRUE),
      fallback = sum(gate_status %in% c("fallback_empty", "fallback_selected")),
      .groups = "drop"
    ) |>
    dplyr::mutate(
      mismatch_type = unname(.simCompareMismatchLabels[mismatch_type]),
      method = unname(.analysis_method_labels[method])
    )
  print(knitr::kable(
    diag_summary,
    col.names = c(
      "Mismatch", "Shift", "Method", "Samples", "FDP defined", "Median FDP",
      "Median sensitivity", "Median selected", "Empty gates", "Fallbacks"
    )
  ))
  cat("\n\n")

  diag_stimgate <- diag_primary |>
    dplyr::filter(method == "stimgate") |>
    dplyr::left_join(diag_sample_gate, by = c("sim_id", "sample")) |>
    dplyr::transmute(
      mismatch = unname(.simCompareMismatchLabels[mismatch_type]),
      shift = mismatch_val,
      sample,
      final_gate = .analysis_label_number(threshold),
      sample_gate = .analysis_label_number(sample_gate),
      sample_gate_source,
      final_gate_source = gateReturnPoint,
      fdp = .analysis_label_percent(fdp),
      sensitivity = .analysis_label_percent(sensitivity),
      selected = .analysis_label_percent(selected_fraction),
      status = gate_status
    ) |>
    dplyr::arrange(mismatch, shift, as.integer(sample))
  print(knitr::kable(
    diag_stimgate,
    col.names = c(
      "Mismatch", "Shift", "Sample", "Final gate", "Sample-level gate",
      "Sample-level gate source", "Final gate source", "FDP", "Sensitivity",
      "Selected", "Gate status"
    )
  ))
}
```

Historical context: an earlier diagnostic used `hpi1`, a fixed gamma bias
of 0.01 and five samples. Its gate-placement and error observations do not
represent this run. Assess the current settings using the saved tables and
figures here.

Each figure is one sample. Grey bars are the stimulated cells, stacked by
true label; the black outline is the unstimulated tube; vertical lines are
each method's final gate. Rows are shift sizes and columns the two shift
variants.

```{r}
#| label: gate-diagnostic-figures
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  dir_diag <- .compare_fresh_fig_dir("gate_diagnostic_clean_gamma")
  diagnostic_samples <- unique(gate_diagnostic$cells$sample)
  diagnostic_samples <- diagnostic_samples[order(as.integer(diagnostic_samples))]
  for (s in diagnostic_samples) {
    .analysis_heading(paste0("Sample ", s), 3L)
    p <- .simComparePlotGateDiagnostic(
      gate_diagnostic$cells[gate_diagnostic$cells$sample == s, ],
      diag_primary[diag_primary$sample == s, ]
    )
    .analysis_print_save_fig(
      p, file.path(dir_diag, paste0("plot_sample_", s, ".pdf")),
      height = 16
    )
  }
}
```

## Secondary results: background-subtracted frequency error

The figures below describe the accuracy of the background-subtracted
frequency estimate, which answers a different question from the gate purity
and detection above. A gate can be pure yet miss most responding cells (a
large under-estimate with zero FDP), and an impure gate can still give an
accurate background-subtracted frequency if the unstimulated tube is
contaminated to the same extent. Both are needed to judge a method.

### Relative error summaries

The relative error is (estimate - truth) / truth. Within each biological
scenario and method, the median and existing tail percentiles pool all valid
sample estimates across independent datasets. Samples with zero true frequency
have no relative error. Main percentile figures omit maxima.

The baseline-average views weight baseline scenario statistics equally.
Each baseline has one cell count and response probability, but the high
Gaussian and high gamma settings each contain two baselines. The per-cell
views keep these baselines separate; each cell count identifies one response
probability within its transformation and mean-position setting.
Averages stay within one transformation, mismatch mechanism and mismatch size;
labels say **Mean of scenario medians** and **Mean of scenario 95th percentiles**.
Counts beside figures report finite contributing scenarios. Whole-dataset
bootstrap draws recalculate the entire average, preserving shared biological
random draws across methods and deterministic mismatch settings. A draw in
which an originally contributing scenario has no finite statistic is undefined;
its coverage is reported and unreliable intervals are suppressed. Point
estimates and their contributing scenarios stay the same with intervals on/off.

Absolute-error figures ignore direction; signed figures split over- and
under-estimates and retain the share of valid samples in each direction.
Exact agreement belongs to neither direction. Every figure has all-method and
no-Tailgate versions. Main rows show median and 95th percentile, while the
separate dataset-maximum figures below show occurrence and conditional severity.
Whole-dataset intervals need five contributing datasets and 95% finite draws.
Bars and points share the +1500% cap; negative signed estimates remain visible.

```{r}
#| label: mismatch-error-summaries
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  scenario_error_cols <- c(
    "base_scenario_id", "transformation", "mean_pos_setting",
    "prob_response", "n_cell", "mismatch_type", "mismatch_val", "method"
  )
  # Scenario statistics, one row per scenario and method.
  err_scenario <- .simCompareUnsignedErrorSummary(
    compare_raw,
    scenarioCols = scenario_error_cols, mcse = show_mcse
  )
  signed_scenario <- .simCompareSignedErrorSummary(
    compare_raw,
    scenarioCols = scenario_error_cols, mcse = show_mcse
  )
  # Averaging drops the baseline scenario columns left out of the groups.
  group_all <- c(
    "mean_pos_setting", "transformation", "mismatch_type", "mismatch_val",
    "method"
  )
  percentile_scenario <- .simCompareSignedPercentileSummary(
    compare_raw, scenario_error_cols, mcse = show_mcse)
  percentile_averaged <- .analysis_mcse_bootstrap_average(
    percentile_scenario, group_all, names(.simBandwidthSignedErrorProbs), mcse = TRUE)
  print(knitr::kable(.simComparePerformanceCoverage(percentile_scenario, scenario_error_cols),
    caption = "Unconditional pooled-percentile bootstrap validity."))
  print(knitr::kable(.simComparePerformanceCoverage(percentile_averaged, group_all),
    caption = "Unconditional scenario-percentile average bootstrap validity."))
  group_cell <- c(group_all, "n_cell")
  err_averaged <- .simComparePerformanceAverage(compare_raw, scenario_error_cols, group_all, mcse = show_mcse)
  err_by_cell <- .simComparePerformanceAverage(compare_raw, scenario_error_cols, group_cell, mcse = show_mcse)
  signed_averaged <- .simComparePerformanceAverage(compare_raw, scenario_error_cols, group_all, signed = TRUE, mcse = show_mcse)
  signed_by_cell <- .simComparePerformanceAverage(compare_raw, scenario_error_cols, group_cell, signed = TRUE, mcse = show_mcse)
  coverage_summaries <- list(err_scenario, signed_scenario, err_averaged,
    err_by_cell, signed_averaged, signed_by_cell)
  coverage_keys <- list(scenario_error_cols, scenario_error_cols, group_all,
    group_cell, group_all, group_cell)
  for (i in seq_along(coverage_summaries)) {
    print(knitr::kable(.simComparePerformanceCoverage(coverage_summaries[[i]], coverage_keys[[i]]),
      caption = "Pooled or scenario-average bootstrap coverage; undefined draws and unavailable intervals."))
  }
}
```

### Relative error across baseline scenarios

The size of the relative error is plotted against mismatch size, with scenario
statistics averaged equally over baseline scenarios within each transformation
and mean-position setting.

```{r}
#| label: relative-error-averaged
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    err_averaged,
    make_plot = .compare_mismatch_plot(
      .simComparePlotMismatchError,
      mcse = show_mcse
    ),
    dir = .compare_fresh_fig_dir("relative_error_averaged"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 14,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### Relative error at each cell count

As above, but keeping each baseline separate by its cell count. Each plotted
cell count has a single response probability within its transformation and
mean-position setting.

```{r}
#| label: relative-error-by-n-cell
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    err_by_cell,
    make_plot = .compare_mismatch_plot(
      .simComparePlotMismatchError,
      mcse = show_mcse
    ),
    dir = .compare_fresh_fig_dir("relative_error_by_n_cell"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"),
    height = 14,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

Each signed-error figure also has an estimate/reference ratio companion in a
sibling `ratio*` folder. The curves and intervals are identical: 1x is exact,
0x is an estimate of zero and 2x is twice the reference. Negative ratios remain
visible; a capped upper tick reads “≥ 16x”.

### Over- and under-estimates across baseline scenarios

Relative error is split by direction: over-estimates sit above zero and
under-estimates below. Within each scenario, the median and 95th percentile are calculated separately for over- and under-estimates, then averaged
equally over baseline scenarios within each transformation and mean-position
setting. Line weight shows the share of estimates in that direction.
Under-estimates are on a linear scale, including errors below -100% for negative background-subtracted estimates, with -100% (0x, an estimate of zero), and
over-estimates on a fold scale, so -100% and +100% (twice the truth) are the
same distance from zero.

```{r}
#| label: signed-error-averaged
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    signed_averaged,
    make_plot = .compare_mismatch_plot(
      .simComparePlotSignedError,
      x = "mismatch_val", x_log = FALSE, mcse = show_mcse
    ),
    dir = .compare_fresh_fig_dir("signed_error_averaged"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    ratio_twins = TRUE,
    height = 16,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### Percentiles of signed relative error averaged over cell counts

The seven curves summarise the full signed relative error distribution, including zero errors, rather than conditioning on over- or under-estimation. They show the 2.5th, 10th, 25th, 50th, 75th, 90th and 97.5th percentiles together. If either outer percentile lacks an interval at any finite point, the whole figure uses the 5th/95th pair and labels that fallback; unavailable intervals are omitted without dropping points. The median is darkest and thickest; symmetric percentile pairs share progressively lighter, thinner tiers. Percentiles are calculated before applying the signed display scale.

Intervals resample whole independent datasets (`iter`), recomputing pooled tube percentiles. Scenario averages bootstrap the complete equal-weight average with shared draws across methods and deterministic mismatch settings; bootstrap validity is reported beside the figures.

```{r}
#| label: signed-percentiles-averaged
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    percentile_averaged,
    make_plot = function(d) .simBandwidthSignedPercentilePlot(
      d, x = "mismatch_val", x_label = .compare_mismatch_x(d$mismatch_type[[1]]),
      x_log = FALSE, mcse = show_mcse) +
      ggplot2::aes(shape = method, linetype = method) +
      .analysis_scale_method(c("colour", "shape", "linetype")),
    dir = .compare_fresh_fig_dir("signed_percentiles_averaged"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = "mismatch_type",
    height = 16,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

Signed- and absolute-error panels have independent ranges across transformations.
Method colours, shapes and line types follow the shared comparison encoding.

### Over- and under-estimates at each cell count

As above, but keeping each baseline separate by its cell count. Each plotted
cell count has a single response probability within its transformation and
mean-position setting.

```{r}
#| label: signed-error-by-n-cell
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    signed_by_cell,
    make_plot = .compare_mismatch_plot(
      .simComparePlotSignedError,
      x = "mismatch_val", x_log = FALSE, mcse = show_mcse
    ),
    dir = .compare_fresh_fig_dir("signed_error_by_n_cell"),
    file_fn = .compare_mismatch_file,
    extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"),
    ratio_twins = TRUE,
    height = 16,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### Dataset-maximum occurrence and conditional severity

The fixed-size maximum summaries use only datasets with valid relative errors
for all 20 intended samples. Incomplete datasets stay in the coverage table;
methods can have different eligible cohorts. Tiny quick-mode smoke checks use
their one intended sample and cannot provide reportable intervals. For each
direction, occurrence is the share of eligible datasets containing at least
one such error. Severity is the mean within-dataset maximum among affected
datasets. No affected datasets means zero occurrence and undefined severity.

Bootstrap draws resample all datasets, including unaffected and incomplete ones,
and jointly recalculate occurrence and severity. Occurrence intervals need five
eligible datasets and 95% finite draws; severity additionally needs five affected
datasets. Suppressed intervals and undefined-bootstrap coverage appear beside
method-specific eligible, incomplete and affected counts. Conditional maxima
need not exceed the pooled sample percentiles above.

```{r}
#| label: dataset-max-coverage
#| echo: false
if (isTRUE(run_plots) && isTRUE(results_available)) {
  dataset_max <- .simCompareDatasetMaxSummary(compare_raw,
    scenarioCols = scenario_error_cols, expected_samples = n_sample_sim,
    expected_datasets = seq_len(n_iter_sim), mcse = show_mcse)
  print(knitr::kable(.simCompareDatasetMaxCoverage(dataset_max), digits = 3,
    caption = "Method-specific eligible, incomplete and affected dataset cohorts; bootstrap coverage."))
}
```

```{r}
#| label: dataset-max-occurrence
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(dataset_max,
    make_plot = function(d) {
      .compare_baseline_note(d)
      .simCompareDatasetMaxPlot(d, "occurrence",
      x = "mismatch_val", x_label = .compare_mismatch_x(d$mismatch_type[[1]]),
      x_log = FALSE, mcse = show_mcse)
    },
    dir = .compare_fresh_fig_dir("dataset_max_occurrence_by_n_cell_prob"),
    file_fn = .compare_mismatch_file, extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"), height = 14, level = 4L,
    mcse_mode = mcse_mode)
}
```

### Conditional dataset maxima: signed errors and ratio companions

Severity is shown with its direction: over-estimates above zero, under-estimates
below. Each signed-error figure has an identical estimate/reference ratio twin.

```{r}
#| label: dataset-max-severity
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(dataset_max,
    make_plot = function(d) {
      .compare_baseline_note(d)
      .simCompareDatasetMaxPlot(d, "severity",
      x = "mismatch_val", x_label = .compare_mismatch_x(d$mismatch_type[[1]]),
      x_log = FALSE, mcse = show_mcse)
    },
    dir = .compare_fresh_fig_dir("dataset_max_signed_error_severity_by_n_cell_prob"),
    file_fn = .compare_mismatch_file, extra_heading = .compare_mismatch_heading,
    extra_col = c("mismatch_type", "n_cell"), ratio_twins = TRUE,
    height = 14, level = 4L, mcse_mode = mcse_mode)
}
```

### Threshold placement diagnostic (unstimulated background intrusion)

The mean share of unstimulated cells above the gate is plotted against
the stimulated mean shift. Colour, point shape and line type identify the
method, and panels separate baseline scenarios.
Each mismatch mechanism has its own figure; scenarios wrap into two columns
with independent vertical ranges. Higher values indicate that more unstimulated
cells lie above the chosen gate.
Error bars (with `show_mcse` on) are whole-dataset bootstrap intervals for
the plotted pooled mean.

```{r}
#| label: threshold-placement-unstim
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  summary_shifts <- compare_summary |>
    dplyr::filter(
      mismatch_type %in% c("mean_shift_all", "mean_shift_negative")
    )
  .simCompareFigureLoop(
    summary_shifts,
    make_plot = function(d) {
      .compare_baseline_note(d)
      .simComparePlotPlacement(d, mcse = show_mcse)
    },
    dir = .analysis_fig_dir(c(fig_key, "threshold_placement_unstim"), path_root = root_dir),
    file_fn = .compare_mismatch_file,
    extra_col = "mismatch_type",
    extra_heading = .compare_mismatch_heading,
    height = 16,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

### Upper-tail error analysis

The 90th percentile of absolute relative error is plotted against mismatch
size on an inverse-hyperbolic-sine vertical scale, with colour identifying
the method. Each mismatch mechanism has its own figure and heading, with baseline scenarios
wrapped into two columns and one mean position setting per plot. Panel ranges
are independent. Compare curves
within each panel, since mean shift and spread increase use different units.
Error bars (with `show_mcse` on) resample whole datasets and recalculate the
pooled tube-level 90th percentile in each draw.

```{r}
#| label: upper-tail-and-fallbacks
#| echo: false
#| results: asis
if (isTRUE(run_plots) && isTRUE(results_available)) {
  .simCompareFigureLoop(
    compare_summary,
    make_plot = function(d) {
      .compare_baseline_note(d)
      .simComparePlotUpperTail(d, mcse = show_mcse)
    },
    dir = .analysis_fig_dir(c(fig_key, "upper_tail_and_fallbacks"), path_root = root_dir),
    file_fn = .compare_mismatch_file,
    extra_col = "mismatch_type",
    extra_heading = .compare_mismatch_heading,
    height = 18,
    level = 4L, mcse_mode = mcse_mode
  )
}
```

## Rerun one scenario

A single scenario can be rerun with the production settings and RNG.
Run the setup, settings and grid chunks first.

```{r}
#| label: rerun-one-simulation
#| include: false
#| eval: false
sim_id_rerun <- 1L
row_rerun <- sim_grid_full[sim_grid_full$sim_id == sim_id_rerun, , drop = FALSE]
stopifnot(nrow(row_rerun) == 1L)
compare_one <- .simCompareRunScenario(
  row = row_rerun,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  clusterGates = cluster_gates,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  tailgateX = tailgate_x,
  resume = FALSE
)
```
