---
title: Compare background-subtracted frequency approaches under condition mismatch
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr)
library(dplyr)
library(tidyr)

root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

# Tailgate comparisons use the cytoUtils package directly.
if (!requireNamespace("cytoUtils", quietly = TRUE)) {
  warning(
    "Package 'cytoUtils' is not installed. Tailgate comparisons will fail if run."
  )
}

scripts_r_dir <- file.path(root_dir, "scripts", "r")

# Source simulation helpers in dependency order:
# 1. analysis-runtime.R: shared QMD execution/runtime plumbing
# 2. sim-misc.R: general simulation utilities and transformation lookups
# 3. sim-bandwidth.R: location details reader and bandwidth helpers
# 4. sim-compare-freq_bs.R: comparison functions requiring the helpers above
for (fn in c(
  "analysis-runtime.R",
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}
```

### Aim

Compare the StimGate background-subtracted frequency estimate against two
alternative thresholding approaches (`fbeta` and `tailgate`) under controlled
**condition mismatch** (technical misalignment between the stimulated and
matched unstimulated tubes).

The simulation isolates two deterministic post-transformation mismatch
mechanisms provided by `simcyto`:

1. **Mean shift** (`stimMeanShift`): An additive shift applied to all stimulated
   cells on the transformed expression scale.
2. **SD inflation** (`stimSdMultiplier`): A multiplicative scaling of the
   within-component standard deviation for stimulated cells around each
   component's realised mean.

In all primary mismatch comparisons, random perturbations are disabled
(`samplePerturbationSd = 0`, `conditionPerturbationSd = 0`,
`clusterPerturbationSd = 0`) to isolate the effects of the controlled mismatch.

### Simulation settings

```{r}
#| label: actual-settings
# Method settings inherited from analysis 7:
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"

# Cluster refinement enabled for the full one-marker procedure
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE

background_relative_to_response <- 0.2
n_cell_uns_relative_to_stim <- 1

mean_pos_settings_tbl <- .simMiscGetMeanPosTbl()

# Targeted scenario panel representing qualitatively distinct baselines
targeted_scenarios_tbl <- tibble::tribble(
  ~transformation, ~mean_pos_setting, ~prob_response, ~n_cell, ~scenario_desc,
  "gaussian", "high", 0.002, 20000, "Rare response, clean baseline",
  "skew", "high", 0.2, 20000, "Clean skew baseline",
  "gamma", "high", 0.2, 5000, "Very clean gamma baseline",
  "gaussian", "low", 0.2, 1000, "Overlapping Gaussian (StimGate clean advantage)",
  "gamma", "low", 0.2, 1000, "Overlapping Gamma (StimGate clean advantage)",
  "gaussian", "high", 0.0005, 100000, "Rare Gaussian (F-beta clean advantage)",
  "gamma", "high", 0.0005, 20000, "Rare Gamma (F-beta clean advantage)"
) |>
  dplyr::left_join(
    mean_pos_settings_tbl,
    by = c("transformation", "mean_pos_setting")
  ) |>
  dplyr::mutate(
    bias_uns = dplyr::case_when(
      transformation == "gamma" ~ 0.01,
      transformation %in% c("gaussian", "skew") ~ 0.15,
      TRUE ~ NA_real_
    ),
    bw_mtd = stimgate_bw_mtd,
    bw_ncell_max = stimgate_bw_ncell_max,
    bw_fallback = stimgate_bw_fallback,
    bw_min = stimgate_bw_min,
    bw_max = stimgate_bw_max,
    sample_perturbation_sd = 0,
    condition_perturbation_sd = 0,
    cluster_perturbation_sd = 0,
    background_relative_to_response = background_relative_to_response,
    n_cell_uns_relative_to_stim = n_cell_uns_relative_to_stim
  )
```

### Mismatch experiment grids

```{r}
#| label: mismatch-grids
# 1. Mean-shift grid (all components):
# Gaussian and Skew: 0, 0.025, 0.05, 0.1, 0.2
# Gamma: 0, 0.005, 0.01, 0.025, 0.05
mean_shift_gaussian_skew <- c(0, 0.025, 0.05, 0.1, 0.2, 0.5, 1)
mean_shift_gamma <- c(0, 0.005, 0.01, 0.025, 0.05, 0.1, 0.2)

grid_mean_shift_all <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_all",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = NA_character_,
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 2. Mean-shift grid (negative/background component only):
grid_mean_shift_negative <- purrr::map_df(
  seq_len(nrow(targeted_scenarios_tbl)),
  function(i) {
    row <- targeted_scenarios_tbl[i, , drop = FALSE]
    shifts <- if (row$transformation[[1]] == "gamma") {
      mean_shift_gamma
    } else {
      mean_shift_gaussian_skew
    }
    tidyr::expand_grid(
      row,
      stim_mean_shift = shifts
    ) |>
      dplyr::mutate(
        mismatch_type = "mean_shift_negative",
        mismatch_val = stim_mean_shift,
        stim_mean_shift_clusters = "gn",
        stim_sd_multiplier = 1,
        sd_increase = 0
      )
  }
)

# 3. SD-inflation grid (negative/background component only):
# Fractional SD increase: 0, 0.01, 0.025, 0.05, 0.10, 0.20
sd_increase_vec <- c(0, 0.01, 0.025, 0.05, 0.10, 0.20, 0.5, 1)

grid_sd_inflation <- tidyr::expand_grid(
  targeted_scenarios_tbl,
  sd_increase = sd_increase_vec
) |>
  dplyr::mutate(
    mismatch_type = "sd_inflation",
    mismatch_val = sd_increase,
    stim_mean_shift = 0,
    stim_mean_shift_clusters = NA_character_,
    stim_sd_multiplier = 1 + sd_increase,
    stim_sd_multiplier_clusters = "gn"
  )

# Combined full mismatch simulation grid
sim_grid <- dplyr::bind_rows(
  grid_mean_shift_all,
  grid_mean_shift_negative,
  grid_sd_inflation
) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

if (.isDev()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      transformation == "gaussian",
      mean_pos_setting == "high",
      prob_response == 0.002,
      mismatch_val %in% c(0, 0.05)
    )
}

sim_grid_chunk_index <- as.integer(
  .get_qmd_param_env("sim_grid_chunk_index", "SIM_GRID_CHUNK_INDEX", 1L)
)
sim_grid_n_chunks <- as.integer(
  .get_qmd_param_env("sim_grid_n_chunks", "SIM_GRID_N_CHUNKS", 1L)
)
.validate_sim_grid_chunk_settings(sim_grid_chunk_index, sim_grid_n_chunks)
sim_chunk_label <- .sim_chunk_label(sim_grid_chunk_index, sim_grid_n_chunks)

sim_grid_all <- sim_grid

if (sim_grid_n_chunks > 1L && nrow(sim_grid_all) > 0L) {
  sim_grid <- sim_grid_all |>
    dplyr::filter(
      ((dplyr::row_number() - 1L) %% sim_grid_n_chunks) + 1L ==
        sim_grid_chunk_index
    )
}

sim_grid
print(paste0("Total simulation scenarios: ", nrow(sim_grid)))
```

### Simulation execution

```{r}
#| label: compare-batch-run
#| results: hide
path_cache_dir <- projr::projr_path_get(
  "cache",
  "sim",
  "compare",
  "freq_bs_batch"
)
dir_output <- file.path(path_cache_dir, "output")
dir.create(dir_output, recursive = TRUE, showWarnings = FALSE)

path_progress_file <- projr::projr_path_get(
  "cache",
  "log",
  "analysis",
  "sim",
  "compare",
  "freq_bs_batch",
  if (sim_grid_n_chunks > 1L) {
    paste0("progress-chunk_", sim_chunk_label, ".txt")
  } else {
    "progress.txt"
  }
)
dir.create(dirname(path_progress_file), recursive = TRUE, showWarnings = FALSE)

path_rds <- file.path(
  path_cache_dir,
  if (sim_grid_n_chunks > 1L) {
    sprintf("compare_raw_chunk_%s.rds", sim_chunk_label)
  } else {
    "compare_raw.rds"
  }
)

n_sample_sim <- 5
n_iter_sim <- 5
n_workers <- max(1, .simGetCores())

if (interactive()) {
  progressr::handlers(global = TRUE)
}

compare_raw <- .simCompareFreqBsGrid(
  sim_grid = sim_grid,
  nSample = n_sample_sim,
  nIter = n_iter_sim,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  probExact = TRUE,
  covEvMin = 1.5,
  covEvMax = 1.5,
  tolClust = tol_clust,
  locEnforceShapeThreshold = loc_enforce_shape_threshold,
  calcCytPosGates = calc_cyt_pos_gates,
  includeLocCondition = TRUE,
  pathFbeta = path_fbeta,
  fbetaBeta = fbeta_beta,
  fbetaTheta = fbeta_theta,
  fbetaWidth = fbeta_width,
  tailgateAdjust = tailgate_adjust,
  tailgateAutoTol = tailgate_auto_tol,
  tailgateMethod = tailgate_method,
  tailgateTol = tailgate_tol,
  parallel = (n_workers > 1L),
  workers = n_workers,
  dirCache = dir_output,
  pathProgress = path_progress_file,
  resume = TRUE,
  sim_grid_chunk_index = if (sim_grid_n_chunks > 1L) {
    sim_grid_chunk_index
  } else {
    NULL
  },
  sim_grid_n_chunks = if (sim_grid_n_chunks > 1L) {
    sim_grid_n_chunks
  } else {
    NULL
  }
)

.write_rds_atomic(compare_raw, path_rds)

if (sim_grid_n_chunks > 1L) {
  path_rds_full <- file.path(path_cache_dir, "compare_raw.rds")
  compare_raw_full <- .simCompareCollateScenarioOutputs(
    dirCache = dir_output,
    sim_grid = sim_grid_all
  )
  if (nrow(compare_raw_full) > 0L) {
    .write_rds_atomic(compare_raw_full, path_rds_full)
    compare_raw <- compare_raw_full
  }
}
```

### Results summary

```{r}
#| label: compare-summary
if (!exists("compare_raw")) {
  path_rds <- projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs_batch",
    "compare_raw.rds"
  )
  compare_raw <- readRDS(path_rds)
}
compare_summary <- .simCompareSummariseFreqBs(
  compare_raw,
  scenarioCols = c(
    "mismatch_type",
    "mismatch_val",
    "stim_mean_shift",
    "stim_mean_shift_clusters",
    "stim_sd_multiplier",
    "stim_sd_multiplier_clusters",
    "sd_increase",
    "transformation",
    "mean_pos_setting",
    "mean_pos",
    "prob_response",
    "n_cell",
    "scenario_desc",
    "approach",
    "method"
  )
)

compare_summary
```

### Mean-shift degradation curves (all components vs negative only)

#### Median

```{r}
#| label: plot-mean-shift-degradation-median
#| fig-width: 10
#| fig-height: 9
summary_mean_shift <- compare_summary |>
  dplyr::filter(
    mismatch_type %in% c("mean_shift", "mean_shift_all", "mean_shift_negative")
  )

for (mean_pos_setting_curr in unique(summary_mean_shift$mean_pos_setting)) {
  summary_mean_shift_curr <- summary_mean_shift |>
    dplyr::filter(mean_pos_setting == mean_pos_setting_curr)

  p_shift <- ggplot(
    summary_mean_shift_curr,
    aes(
      x = mismatch_val,
      y = med_abs_rel_error,
      color = method,
      linetype = mismatch_type,
      group = interaction(method, mismatch_type)
    )
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    facet_wrap(~scenario_desc, scales = "free_y") +
    scale_y_continuous(
      transform = scales::asinh_trans()
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = "Degradation under Stimulated Mean Shift",
      subtitle = "Median Absolute Relative Error vs Mean Shift",
      x = "Stimulated Mean Shift",
      y = "Median Absolute Relative Error (asinh scale)",
      color = "Method",
      linetype = "Mismatch Variant"
    )
  path_p <- projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs_batch",
    paste0(
      "mean_shift_degradation_plot_",
      mean_pos_setting_curr,
      ".pdf"
    )

  )
  ggsave(
    path_p,
    p_shift,
    width = 20,
    height = 15,
    units = "cm"
  )
  print(p_shift)
}
```

#### Max

```{r}
#| label: plot-mean-shift-degradation-max
#| fig-width: 10
#| fig-height: 9
summary_mean_shift <- compare_summary |>
  dplyr::filter(
    mismatch_type %in% c("mean_shift", "mean_shift_all", "mean_shift_negative")
  )

summary_mean_shift <- compare_summary |>
  dplyr::filter(mismatch_type == "mean_shift")

for (mean_pos_setting_curr in unique(summary_mean_shift$mean_pos_setting)) {
  summary_mean_shift_curr <- summary_mean_shift |>
    dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
  p_shift_max <- ggplot(
    summary_mean_shift_curr,
    aes(
      x = mismatch_val,
      y = max_abs_rel_error,
      color = method,
      linetype = mismatch_type,
      group = interaction(method, mismatch_type)
    )
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    facet_wrap(~scenario_desc, scales = "free_y") +
    scale_y_continuous(
      transform = scales::asinh_trans()
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = "Degradation under Stimulated Mean Shift",
      subtitle = "Maximum Absolute Relative Error vs Mean Shift",
      x = "Stimulated Mean Shift",
      y = "Maximum Absolute Relative Error (asinh scale)",
      color = "Method",
      linetype = "Mismatch Variant"
    )
  path_p <- projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs_batch",
    paste0(
      "mean_shift_degradation_plot_max_",
      mean_pos_setting_curr,
      ".pdf"
    )
  )
  ggsave(
    path_p,
    p_shift_max,
    width = 20,
    height = 15,
    units = "cm"
  )
  print(p_shift_max)
}
```

#### 90-th percentile

```{r}
#| label: plot-mean-shift-degradation-q90
#| fig-width: 10
#| fig-height: 9
summary_mean_shift <- compare_summary |>
  dplyr::filter(
    mismatch_type %in% c("mean_shift", "mean_shift_all", "mean_shift_negative")
  )

for (mean_pos_setting_curr in unique(summary_mean_shift$mean_pos_setting)) {
  summary_mean_shift_curr <- summary_mean_shift |>
    dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
  p_shift_90 <- ggplot(
    summary_mean_shift_curr,
    aes(
      x = mismatch_val,
      y = q90_abs_rel_error,
      color = method,
      linetype = mismatch_type,
      group = interaction(method, mismatch_type)
    )
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    facet_wrap(~scenario_desc, scales = "free_y") +
    scale_y_continuous(
      transform = scales::asinh_trans()
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = "Degradation under Stimulated Mean Shift",
      subtitle = "90th Percentile Absolute Relative Error vs Mean Shift",
      x = "Stimulated Mean Shift",
      y = "90th Percentile Absolute Relative Error (asinh scale)",
      color = "Method",
      linetype = "Mismatch Variant"
    )
  path_p <- projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs_batch",
    paste0(
      "mean_shift_degradation_plot_q90_",
      mean_pos_setting_curr,
      ".pdf"
    )
  )
  ggsave(
    path_p,
    p_shift_90,
    width = 20,
    height = 15,
    units = "cm"
  )
  print(p_shift_90)
}
```

### SD-inflation degradation curves

```{r}
#| label: plot-sd-inflation-degradation
#| fig-width: 10
#| fig-height: 8
summary_sd <- compare_summary |>
  dplyr::filter(mismatch_type == "sd_inflation")

for (mean_pos_setting_curr in unique(summary_sd$mean_pos_setting)) {
  summary_sd_curr <- summary_sd |>
    dplyr::filter(mean_pos_setting == mean_pos_setting_curr)
  p_sd <- ggplot(
    summary_sd_curr,
    aes(
      x = mismatch_val,
      y = med_abs_rel_error,
      color = method,
      group = method
    )
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    facet_wrap(~scenario_desc, scales = "free_y") +
    scale_y_continuous(
      transform = scales::asinh_trans()
    ) +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = "Degradation under Stimulated SD Inflation",
      subtitle = "Median Absolute Relative Error vs Fractional SD Increase (Negative Component Only)",
      x = "Fractional SD Increase",
      y = "Median Absolute Relative Error (asinh scale)",
      color = "Method"
    )
  path_p <- projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs_batch",
     paste0(
        "sd_inflation_degradation_plot_",
        mean_pos_setting_curr,
        ".pdf"
     )
  )
  ggsave(
    path_p,
    p_sd,
    width = 20,
    height = 15,
    units = "cm"
  )
  print(p_sd)
}
```

### Threshold placement diagnostic (unstimulated background intrusion)

```{r}
#| label: plot-threshold-placement-unstim
#| fig-width: 10
#| fig-height: 9
summary_shifts <- compare_summary |>
  dplyr::filter(
    mismatch_type %in% c("mean_shift", "mean_shift_all", "mean_shift_negative")
  )

if (nrow(summary_shifts) > 0) {
  p_placement <- ggplot(
    summary_shifts,
    aes(
      x = mismatch_val,
      y = propUns_mean,
      color = method,
      linetype = mismatch_type,
      group = interaction(method, mismatch_type)
    )
  ) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    facet_wrap(~scenario_desc, scales = "free_y") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "bottom"
    ) +
    labs(
      title = "Threshold Placement: Unstimulated Background Intrusion",
      subtitle = "Proportion of unstimulated cells above threshold vs shift",
      x = "Stimulated Mean Shift",
      y = "Proportion of Unstimulated Cells Above Gate",
      color = "Method",
      linetype = "Mismatch Variant"
    )
  path_p <- projr::projr_path_get(
    "cache",
    "sim",
    "compare",
    "freq_bs_batch",
    "threshold_placement_unstim_plot.pdf"
  )
  ggsave(
    path_p,
    p_placement,
    width = 20,
    height = 15,
    units = "cm"
  )
  print(p_placement)
}
```

### Upper-tail error analysis

```{r}
#| label: plot-upper-tail-and-fallbacks
#| fig-width: 10
#| fig-height: 8
p_tail <- ggplot(
  compare_summary,
  aes(
    x = mismatch_val,
    y = q90_abs_rel_error,
    color = method,
    group = method
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(mismatch_type ~ scenario_desc, scales = "free") +
  scale_y_continuous(
    transform = scales::asinh_trans()
  ) +
  cowplot::theme_cowplot() +
  cowplot::background_grid(major = "xy", minor = "none") +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "bottom"
  ) +
  labs(
    title = "90th Percentile Relative Error under Mismatch",
    x = "Mismatch Magnitude",
    y = "90th Percentile Absolute Relative Error (asinh scale)",
    color = "Method"
  )
print(p_tail)
```
