---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_preprocessing: null
  run_stimgate: null
  run_comparators: null
  run_plots: false
  tester_n_sample: 20
  n_workers: 2
---

Validate StimGate on the ACS CyTOF dataset and compare its estimated
background-subtracted response frequencies with Tailgate and F-beta.
Each batch contains one unstimulated and four stimulated samples, analysed separately for six cell populations.

## Set-up

`RUN_SIMULATIONS=true` runs preprocessing and all three methods; `RUN_SIMULATIONS=false RUN_PLOTS=true` reads the saved manual-comparison table. The individual `RUN_PREPROCESSING`, `RUN_STIMGATE` and `RUN_COMPARATORS` controls override those stages when supplied. Computation requires the population FCS directories under the `raw-data-large/comparison_data/acscytof/fcs` projr path, the manual `Cytokines_10Jun2021_2.csv` under `raw-data-small/comparison_data/acscytof`, and the filename map from `DataTidyACSCyTOFFAUST`.

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- normalizePath(if (file.exists("DESCRIPTION")) "." else "..", winslash = "/")
stopifnot(file.exists(file.path(root_dir, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = root_dir)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "analysis-plot-style.R"))
source(file.path(scripts_r_dir, "analysis-mcse.R"))
source(file.path(scripts_r_dir, "sim-bandwidth-analysis-plot.R"))
fig_key <- .analysis_mode_key("9-real-compare-acs-cytof", sized = FALSE)
analysis_qmd <- "analysis/9-real-compare-acs-cytof.qmd"
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-preprocess.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-misc.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "acs_cytof-manual.R"))
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  run_simulations
))
run_stimgate <- .as_flag(.get_qmd_param_env(
  "run_stimgate",
  "RUN_STIMGATE",
  run_simulations
))
run_comparators <- .as_flag(.get_qmd_param_env(
  "run_comparators",
  "RUN_COMPARATORS",
  run_simulations
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
tester_n_sample <- as.integer(.get_qmd_param_env(
  "tester_n_sample",
  "TESTER_N_SAMPLE",
  20L
))
n_workers <- as.integer(.get_qmd_param_env(
  "n_workers",
  "ACS_N_WORKERS",
  2L
))

if (is.na(n_workers) || n_workers < 1L) {
  stop("n_workers must be coercible to an integer of at least one.")
}
.acsCytofValidateSampleCount(tester_n_sample)

.acsCytofEnsureCurrentCheckout(root_dir)

# Setup starts in the QMD directory; later chunks and workers use root_dir.
# Keep configured paths absolute across that working-directory change.
path_fcs_base <- projr::projr_path_get(
  "raw-data-large",
  "comparison_data", "acscytof", "fcs",
  format = "absolute"
)
if (isTRUE(run_preprocessing) && !dir.exists(path_fcs_base)) {
  stop("Path to FCS files does not exist: ", path_fcs_base)
}
path_gs_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "gs",
  format = "absolute"
)
path_scratch_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "scratch",
  format = "absolute"
)
path_fbeta <- file.path(root_dir, "scripts", "python", "fbeta.py")
if (isTRUE(run_comparators)) {
  if (!requireNamespace("cytoUtils", quietly = TRUE) ||
      !requireNamespace("reticulate", quietly = TRUE) ||
      !file.exists(path_fbeta)) {
    stop("ACS comparisons require cytoUtils, reticulate and scripts/python/fbeta.py.")
  }
  if (!reticulate::py_module_available("numpy")) {
    stop("ACS F-beta requires numpy in the Python environment selected by reticulate.")
  }
}
```

## Scientific settings

```{r}
#| label: scientific-settings
pop_vec <- c("tcrgd", "cd4", "cd8", "nk", "nk_pre", "b")
.by_pop <- function(x) stats::setNames(rep(x, length(pop_vec)), pop_vec)
run_preprocessing_vec_by_pop <- .by_pop(run_preprocessing)
run_stimgate_vec_by_pop <- .by_pop(run_stimgate)
run_comparators_vec_by_pop <- .by_pop(run_comparators)
# Display diagnostics sequentially below, including when using cached runs.
run_stimgate_plots_vec_by_pop <- .by_pop(FALSE)
run_preprocessing_plots_vec_by_pop <- .by_pop(FALSE)
stimgate_bw_mtd <- "nrd0"
stimgate_bw_scope <- "cytokine"
# NA: StimGate's unstimulated-sample bias equals its initial bandwidth estimate.
bias_uns <- NA_real_
bias_uns_vec_by_pop <- .by_pop(bias_uns)
analysis_seed <- 20260823L
```

## Gating

The TCRgd population is first run on a limited set of complete five-sample
batches. If that tester completes successfully, the same preprocessing and
StimGate flow is applied to every population configured in `pop_vec`.

The reusable helpers fix the six cytokine channels and the method settings.
StimGate uses `bwMtd = "nrd0"` (with `bwNcellMax = 1e4`, `bwFallback = "auto"`
and no bandwidth bounds); bandwidths are shared per marker
(`bwScope = "cytokine"`). StimGate uses `gateCombn = "min"`, `minCell = 100`,
threshold clustering on (`clusterGates = TRUE`) and cytokine-positive refinement
on (`calcCytPosGates = TRUE`). The comparison scores the final clustered
`loc_minClust` gates after cytokine-positive refinement. Both settings are
recorded with the saved run.
Its unstimulated-sample bias
is not fixed: it equals the bandwidth StimGate estimates at the start
(`biasUns = NULL` with `biasUnsFactor = 4`). F-beta uses its standard
parameters (`beta = 0.8`, `theta = 2`, `width = 10` and automatic bin count).
The F-beta implementation is adapted from Richards et al. (2014) only in the
histogram support: common bin edges span the matched stimulated and unstimulated
distributions, instead of being derived from the unstimulated range alone. This
prevents stimulated response-tail events outside the observed unstimulated range
from being omitted when no explicit strong positive control is available.
Tailgate uses the stimulated distribution, automatic bandwidth estimation and
automatic tolerance tuning.

### TCRgd tester

The tester uses only the first `tester_n_sample` files, which must comprise
complete batches of one unstimulated and four stimulated samples. Its cached
GatingSet and StimGate project are kept under a separate `tester/tcrgd`
subdirectory, so they cannot overwrite the full TCRgd run.

```{r}
#| label: run-tcrgd-tester
#| include: false
set.seed(analysis_seed)

tcrgd_tester <- .acsCytofRunPopulation(
  pop = "tcrgd",
  pathFcsBase = path_fcs_base,
  pathGsBase = path_gs_base,
  pathScratchBase = path_scratch_base,
  runPreprocessing = run_preprocessing,
  runMethods = run_stimgate,
  runPlots = FALSE,
  runPreprocessingPlots = FALSE,
  nSample = tester_n_sample,
  biasUns = bias_uns_vec_by_pop[["tcrgd"]],
  bwMtd = stimgate_bw_mtd,
  bwScope = stimgate_bw_scope,
  outputGroup = "tester"
)

if (isTRUE(run_comparators)) {
  tcrgd_comparison_tester <- .acsCytofRunComparisonMethods(
    pop = "tcrgd",
    pathFcsBase = path_fcs_base,
    pathGsBase = path_gs_base,
    pathScratchBase = path_scratch_base,
    runMethods = TRUE,
    nSample = tester_n_sample,
    outputGroup = "tester",
    pathFbeta = path_fbeta
  )
}
```

### Across populations, in parallel

Each population writes to its own cached GatingSet and scratch directory. The
number of concurrent population runs defaults to two and can be changed with
the `n_workers` parameter or `ACS_N_WORKERS` environment variable.

```{r}
#| label: run-all-populations
#| include: false
if (isTRUE(run_preprocessing) || isTRUE(run_stimgate)) {
  population_runs <- .acsCytofMapPopulations(
    pop_vec,
    function(pop) {
      .acsCytofRunPopulationSafe(
        pop = pop,
        pathFcsBase = path_fcs_base,
        pathGsBase = path_gs_base,
        pathScratchBase = path_scratch_base,
        runPreprocessing = run_preprocessing_vec_by_pop[[pop]],
        runMethods = run_stimgate_vec_by_pop[[pop]],
        runPlots = run_stimgate_plots_vec_by_pop[[pop]],
        runPreprocessingPlots = run_preprocessing_plots_vec_by_pop[[pop]],
        bwMtd = stimgate_bw_mtd,
        bwScope = stimgate_bw_scope,
        biasUns = bias_uns_vec_by_pop[[pop]]
      )
    },
    nWorkers = n_workers,
    seed = analysis_seed,
    label = "ACS population preprocessing/StimGate runs"
  )
}
```

## Tailgate and F-beta

Tailgate and F-beta run concurrently across populations, using the same
`n_workers` control as StimGate. Within each population, the two methods remain
sequential so that they do not compete for the same GatingSet. Each method is
saved to its own `result.rds`. Whenever `run_comparators` is true, both methods
are recomputed for every population and replace the previous results only once
both have finished, so a failed rerun keeps the last good results. When
`run_comparators` is false, existing results remain available to
the downstream comparison and plotting sections.

```{r}
#| label: run-comparison-methods-in-parallel
#| include: false
if (isTRUE(run_comparators)) {
  comparison_method_runs <- .acsCytofMapPopulations(
    pop_vec,
    function(pop) {
      .acsCytofRunComparisonMethodsSafe(
        pop = pop,
        pathFcsBase = path_fcs_base,
        pathGsBase = path_gs_base,
        pathScratchBase = path_scratch_base,
        runMethods = run_comparators_vec_by_pop[[pop]],
        pathFbeta = path_fbeta
      )
    },
    nWorkers = n_workers,
    seed = analysis_seed,
    label = "ACS Tailgate/F-beta runs"
  )
}
```

## Collation

### Compare methods with the ACS reference counts

The cached combination tables for all three methods are reduced to one
frequency per cytokine with `UtilsCytoRSV::sum_over_markers()`. Combination
labels are then normalised to the standard `+`/`-` form with
`UtilsCompassSV::convert_cyt_combn_format()` before the automated and manual
tables are joined. `nk_pre` is still run above, but is omitted here because the
manual file has no matching population.

```{r}
#| label: format-manual-comparison
#| include: false
path_manual_output <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "manual-comparison",
  format = "absolute"
)
path_comparison <- file.path(path_manual_output, "manual-comparison.rds")
if (isTRUE(run_preprocessing) || isTRUE(run_stimgate) || isTRUE(run_comparators)) {
  data("fn_to_sampleid_map", package = "DataTidyACSCyTOFFAUST")

  # Build the complete comparison before touching the last good saved output.
  manual_comparison_tbl <- comp_against_manual_cyt(
    fn = "Cytokines_10Jun2021_2.csv",
    path_scratch_base = path_scratch_base,
    path_fcs_base = path_fcs_base,
    pop = pop_vec,
    methods = c("stimgate", "tailgate", "fbeta"),
    path_dir_save = NULL,
    save_plots = FALSE,
    sample_map = fn_to_sampleid_map
  )
  # Save the complete table without generating figures in this plumbing chunk.
  .acsCytofManualSave(
    comparisonTbl = manual_comparison_tbl,
    pathDirSave = path_manual_output,
    savePlots = FALSE
  )
  .write_rds_atomic(
    list(
      analysis_semantics_version = "acs-cytof-v2",
      run_id = .sanitize_run_id(Sys.getenv("ANALYSIS_RUN_ID", unset = ""))
    ),
    file.path(path_manual_output, "manifest.rds")
  )
} else {
  .analysis_check_expected_run(
    .analysis_read_rds(
      file.path(path_manual_output, "manifest.rds"), "acs_cytof", analysis_qmd
    ),
    "acs_cytof", analysis_qmd, semantics_version = "acs-cytof-v2"
  )
  manual_comparison_tbl <- .analysis_read_rds(
    path_comparison, "acs_cytof", analysis_qmd
  )
}
.acsCytofValidateComparisonManifest(manual_comparison_tbl)
manual_summary_tbl <- .acsCytofManualSummaryTable(manual_comparison_tbl)
```

## Results

### Correlation with the manual gating

The comparison table, its RDS copy and a method/population/cytokine summary
are written under `path_manual_output`. When `run_plots` is true, the
figures are saved under `output/fig/9-real-compare-acs-cytof/`. Failed thresholds (including finite fallback gates) remain in the comparison
with missing scored frequencies. Summary tables state total, failed and scored
counts; agreement statistics use successful estimates only. The exclusion CSV
records unmatched manual/automated keys. Methods must have identical sample and
stratum keys and matching data/preprocessing manifests before joining.
Pearson correlations of the StimGate, Tailgate and F-beta estimates with the manual frequencies
follow.

```{r}
#| label: show-manual-summary
#| echo: false
manual_summary_tbl |>
  dplyr::select(
    method:stim, pcc
  ) |>
  tidyr::pivot_wider(
    names_from = method,
    values_from = pcc
  ) |>
  dplyr::mutate(
    diff_abs = abs(stimgate - fbeta)
  ) |>
  dplyr::arrange(desc(diff_abs)) |>
  print(n = Inf)
```

### Estimated versus reference response frequency

Each figure is shown for all methods and again without Tailgate, which performs poorly without more tuning. Manual background-subtracted frequency (%) is on the horizontal axis and automated frequency (%) on the vertical axis. Colour identifies the method, shapes identify stimuli, and panels separate populations and cytokines. Points closer to the dashed identity line indicate closer agreement.

```{r}
#| label: manual-comparison
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  method_sets <- .acsCytofMethodSets()
  for (set_name in names(method_sets)) {
    method_set <- method_sets[[set_name]]
    .analysis_heading(method_set$label, 4)
    p <- .acsCytofManualPlotScatter(
      manual_comparison_tbl |> dplyr::filter(method %in% method_set$methods)
    )
    .analysis_save_fig(
      p,
      file.path(
        .analysis_fig_dir(
          c(fig_key, "manual_comparison_scatter", set_name),
          root_dir
        ),
        "manual-comparison-scatter.png"
      ),
      height = 22
    )
    .analysis_print_fig(p)
  }
}
```

### Relative error

Each figure is shown for all methods and again without Tailgate, which performs poorly without more tuning. Boxplots show absolute relative error by method, with separate panels for populations and cytokines. Smaller values indicate closer agreement; rows with non-positive manual background-subtracted frequencies have no relative-error estimate.

```{r}
#| label: manual-relative-error
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  method_sets <- .acsCytofMethodSets()
  for (set_name in names(method_sets)) {
    method_set <- method_sets[[set_name]]
    .analysis_heading(method_set$label, 4)
    p <- .acsCytofManualPlotRelativeError(
      manual_comparison_tbl |> dplyr::filter(method %in% method_set$methods)
    )
    .analysis_save_fig(
      p,
      file.path(
        .analysis_fig_dir(
          c(fig_key, "manual_comparison_relative_error", set_name),
          root_dir
        ),
        "manual-comparison-relative-error.png"
      ),
      height = 22
    )
    .analysis_print_fig(p)
  }
}
```

### Over- and under-estimates

Each figure is shown for all methods and again without Tailgate, which performs poorly without more tuning. Relative error is (automated - manual) / manual for each sample, using the same samples as the figure above. Over-estimates are plotted above zero and under-estimates below, summarised separately by median, 95th percentile and maximum for each method, population (columns) and cytokine. Larger points mean more samples fell in that direction. The scale is linear down to -100% (0x, an estimate of zero) and doubles above zero, so -100% and a two-fold over-estimate are equally far from zero.

The companion ratio figures show automated/manual frequency with identical
curves: 1x agrees with the manual reference, 0x is zero automated frequency
and 2x is twice the manual frequency. Manual gating remains an imperfect reference.

```{r}
#| label: manual-signed-error
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  method_sets <- .acsCytofMethodSets()
  for (set_name in names(method_sets)) {
    method_set <- method_sets[[set_name]]
    .analysis_heading(method_set$label, 4)
    p <- .acsCytofManualPlotSignedError(
      manual_comparison_tbl |> dplyr::filter(method %in% method_set$methods)
    )
    .analysis_save_fig(
      p,
      file.path(
        .analysis_fig_dir(
          c(fig_key, "manual_comparison_signed_error", set_name),
          root_dir
        ),
        "manual-comparison-signed-error.png"
      ),
      height = 22
    )
    .analysis_print_fig(p)
    .simBandwidthPrintRatioTwin(p,
      file.path(.analysis_fig_dir(c(fig_key, "manual_comparison_signed_error", set_name), root_dir),
        "manual-comparison-ratio.png"), height = 22, level = 5L)
  }
}
```

### Percentiles of signed relative error against manual gating

The seven curves summarise the full signed relative error distribution, including zero errors, rather than conditioning on over- or under-estimation. They show the 2.5th, 10th, 25th, 50th, 75th, 90th and 97.5th percentiles together. If either outer percentile lacks an interval at any finite point, the whole figure uses the 5th/95th pair and labels that fallback; unavailable intervals are omitted without dropping points. The median is darkest and thickest; symmetric percentile pairs share progressively lighter, thinner tiers. Percentiles are calculated before applying the signed display scale.

Intervals bootstrap donors, carrying all their tubes together and using common donor draws across methods and strata. They describe agreement with the imperfect manual reference. All seven percentiles share each population panel; cytokine remains on the x axis.

```{r}
#| label: manual-signed-percentiles
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  # Summarise all methods together before subsetting to keep common donor draws.
  manual_percentiles <- .acsCytofManualSignedPercentiles(manual_comparison_tbl)
  method_sets <- .acsCytofMethodSets()
  for (set_name in names(method_sets)) {
    method_set <- method_sets[[set_name]]
    .analysis_heading(method_set$label, 4)
    p <- .simBandwidthSignedPercentilePlot(
      manual_percentiles |> dplyr::filter(method %in% method_set$methods),
      x = "cyt", x_label = "Cytokine", mcse = TRUE,
      facet = ggplot2::facet_wrap(~pop, scales = "free_y"),
      subtitle = paste0(method_set$label, "; all stimuli; donor-bootstrap intervals")
    )
    .analysis_save_fig(p,
      file.path(.analysis_fig_dir(c(fig_key, "signed_percentiles", set_name), root_dir),
        "manual-comparison-signed-percentiles.png"), height = 22)
    .analysis_print_fig(p)
    print(knitr::kable(manual_percentiles |> dplyr::filter(method %in% method_set$methods) |>
      dplyr::select(method, pop, cyt, dplyr::starts_with("n_"),
        dplyr::starts_with("bootstrap_coverage_"), dplyr::starts_with("interval_available_")), caption = "Donor-bootstrap validity for unconditional percentiles."))
  }
}
```

### Absolute frequency error and results by stimulus

Manual gating is an imperfect reference: these errors describe agreement with it,
not error against known biological truth. Absolute frequency differences, in
percentage points, retain zero manual net frequencies. The current manual and
automated net frequencies are truncated at zero after background subtraction. Relative
errors require a positive manual net frequency and can be large when it is tiny.
The coverage table reports every matched row, method failure, missing manual
frequency and relative-error exclusion by method, population, cytokine and stimulus;
unmatched cohort keys remain in the separate exclusion CSV.

```{r}
#| label: manual-error-coverage
#| echo: false
print(manual_summary_tbl, n = Inf)
```

The following figures separate stimuli so a strong response in one stimulus does
not conceal poorer agreement in another. Both error measures use successful
method estimates only; the table above supplies their denominators.

```{r}
#| label: manual-errors-by-stimulus
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (stimulus in unique(as.character(manual_comparison_tbl$stim))) {
    .analysis_heading(paste("Stimulus:", stimulus), 4)
    rows <- manual_comparison_tbl |> dplyr::filter(stim == stimulus)
    for (metric in c("absolute", "relative")) {
      .analysis_heading(paste(tools::toTitleCase(metric), "error"), 5)
      p <- if (metric == "absolute") .acsCytofManualPlotAbsoluteError(rows) else .acsCytofManualPlotRelativeError(rows)
      .analysis_save_fig(p, file.path(
        .analysis_fig_dir(c(fig_key, "manual_comparison_errors_by_stimulus", stimulus, metric), root_dir),
        paste0("manual-comparison-", metric, "-error.png")
      ), height = 22)
      .analysis_print_fig(p)
    }
  }
}
```

### Donor uncertainty for mean errors

The intervals below are percentile donor-bootstrap intervals for **mean absolute
frequency error** and **mean absolute relative error**, rather than the medians
or tail quantiles in the figures above. Each of 999 replicates resamples donors,
keeping all their stimulated tubes together with their shared, already-subtracted
unstimulated control. The same donor draws are used for all methods, stimuli,
populations and cytokines. Means use finite successful estimates; relative means
also require positive manual net frequency. Fewer than two contributing donors
leaves an interval unavailable. These intervals reflect variation across donors,
not uncertainty about whether manual gating is biologically correct.

```{r}
#| label: manual-donor-uncertainty
#| echo: false
if (isTRUE(run_plots)) {
  manual_uncertainty_tbl <- .acsCytofManualUncertainty(manual_comparison_tbl)
  print(manual_uncertainty_tbl, n = Inf)
}
```

### Expression and gate diagnostics

Histograms compare untransformed and asinh-transformed expression in the first sample of each cached GatingSet, with one panel per marker. The gate plots show the second sample for two cytokine channels, with the saved StimGate thresholds overlaid. These checks require cached GatingSets and gate projects; the frequency figures above use only the canonical comparison table.

```{r}
#| label: acs-expression-diagnostics
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  diagnostic_paths <- c(
    list(.acsCytofPopulationPaths(
      "tcrgd", path_fcs_base, path_gs_base, path_scratch_base,
      outputGroup = "tester"
    )),
    lapply(pop_vec, .acsCytofPopulationPaths,
      pathFcsBase = path_fcs_base, pathGsBase = path_gs_base,
      pathScratchBase = path_scratch_base
    )
  )
  names(diagnostic_paths) <- c("tester/tcrgd", pop_vec)
  for (pop in names(diagnostic_paths)) {
    paths <- diagnostic_paths[[pop]]
    if (!dir.exists(paths$gs)) {
      cat("\nCached GatingSet unavailable for ", pop, "; expression diagnostics skipped.\n\n", sep = "")
      next
    }
    .analysis_heading(paste0("Population: ", pop), 4)
    plots <- .acsCytofPlotGatingSetCheck(paths$gs)
    plot_gatingset_check(paths$gs, paths$gsCheck, plots = plots)
    trans_labels <- c(none = "Untransformed", asinh = "Asinh-transformed")
    for (trans in names(plots)) {
      .analysis_heading(trans_labels[[trans]], 5)
      .analysis_print_fig(plots[[trans]])
    }
  }
}
```

Scatter panels show expression in the two cytokine channels, and marginal panels show their densities in the second sample. Lines mark the saved gate thresholds; compare their position with the expression distributions.

```{r}
#| label: acs-gate-diagnostics
#| echo: false
#| results: asis
if (isTRUE(run_plots)) {
  for (pop in names(diagnostic_paths)) {
    paths <- diagnostic_paths[[pop]]
    if (!dir.exists(paths$gs) || !dir.exists(paths$stimgate)) {
      cat("\nCached GatingSet or gate project unavailable for ", pop, "; gate diagnostics skipped.\n\n", sep = "")
      next
    }
    .analysis_heading(paste0("Population: ", pop), 4)
    gs <- flowWorkspace::load_gs(paths$gs)
    p <- stimgate::plotStim(
      ind = 2L, .data = gs, pathProject = paths$stimgate,
      pop = "root", chnl = c("Ho165Di", "Nd146Di")
    )
    path_check <- file.path(
      .analysis_fig_dir(
        c(fig_key, "gate_diagnostics", gsub("[^A-Za-z0-9_.-]+", "_", pop)),
        root_dir
      ),
      "stimgate_check.pdf"
    )
    .analysis_save_fig(p, path_check, height = 14)
    .analysis_print_fig(p)
  }
}
```
