---
title: ACS CyTOF validation outputs
format:
  html:
    embed-resources: true
    toc: true
params:
  run_plots: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)
scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "acs_cytof-plot_cyt.R"))

run_plots <- isTRUE(params$run_plots)
path_manual_output <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "manual-comparison"
)
path_comparison <- file.path(path_manual_output, "manual-comparison.rds")
if (!file.exists(path_comparison)) {
  stop(
    "ACS manual comparison table not found at: ",
    path_comparison,
    ". Run analysis/9-real-compare-acs-cytof.qmd first."
  )
}
manual_comparison_tbl <- readRDS(path_comparison)
```

## Aim

Reproduce the ACS manuscript-style validation plots for StimGate, F-beta and
Tailgate using the comparison table produced by the main ACS analysis.

The raw-frequency plots show automated against manual background-subtracted
frequencies for every matched population and cytokine. The correlation outputs
show both Pearson correlation (PCC) and concordance correlation (CCC), with one
set restricted to CD4, CD8 and TCRgd T cells and another using every matched
population. This keeps the original manuscript comparison available while also
showing the additional populations in the current benchmark.

## Correlation table

```{r}
#| label: correlation-table
correlation_tbl <- .acsCytofValidationCorrelationTable(manual_comparison_tbl)
correlation_tbl
```

## Validation figures

```{r}
#| label: save-validation-figures
#| results: hide
if (run_plots) {
  .acsCytofValidationSavePlots(
    comparisonTbl = manual_comparison_tbl,
    pathDirSave = path_manual_output
  )
}
```

### Raw background-subtracted frequencies

```{r}
#| label: show-scatter
if (run_plots) {
  .acsCytofValidationPlotScatter(manual_comparison_tbl, "stimgate")
  .acsCytofValidationPlotScatter(manual_comparison_tbl, "fbeta")
  .acsCytofValidationPlotScatter(manual_comparison_tbl, "tailgate")
}
```

### Correlation heatmaps for CD4, CD8 and TCRgd T cells

```{r}
#| label: show-real-pop-heatmaps
if (run_plots) {
  for (method in c("stimgate", "fbeta", "tailgate")) {
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "pcc",
      realPopulationsOnly = TRUE
    ))
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "ccc",
      realPopulationsOnly = TRUE
    ))
  }
}
```

### Correlation heatmaps for all matched populations

```{r}
#| label: show-all-pop-heatmaps
if (run_plots) {
  for (method in c("stimgate", "fbeta", "tailgate")) {
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "pcc",
      realPopulationsOnly = FALSE
    ))
    print(.acsCytofValidationPlotCorrelation(
      correlationTbl = correlation_tbl,
      method = method,
      metric = "ccc",
      realPopulationsOnly = FALSE
    ))
  }
}
```
