---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
params:
  run_simulations: true
  run_preprocessing: null
  run_stimgate: null
  run_comparators: null
  run_plots: false
  tester_n_sample: 20
  n_workers: 2
---

This analysis compares StimGate, Tailgate and F-beta estimates of background-subtracted cytokine frequencies with manual gating in the ACS CyTOF dataset. Each batch contains one unstimulated and four stimulated samples, analysed separately for six cell populations. Automated marker-combination counts are reduced to single-cytokine frequencies and matched to the manual reference counts.

## Set-up

`RUN_SIMULATIONS=true` runs preprocessing and all three methods; `RUN_SIMULATIONS=false RUN_PLOTS=true` reads the saved manual-comparison table. The individual `RUN_PREPROCESSING`, `RUN_STIMGATE` and `RUN_COMPARATORS` controls override those stages when supplied. Computation requires the population FCS directories under the `raw-data-large/comparison_data/acscytof/fcs` projr path, the manual `Cytokines_10Jun2021_2.csv` under `raw-data-small/comparison_data/acscytof`, and the filename map from `DataTidyACSCyTOFFAUST`.

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

# Quarto may start in analysis/ when no project execution directory is set.
if (!file.exists(file.path(root_dir, "DESCRIPTION"))) {
  root_dir <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
}
if (!file.exists(file.path(root_dir, "DESCRIPTION"))) {
  stop("Cannot find the stimgate repository root; render from the checkout.")
}

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-preprocess.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-misc.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "acs_cytof-manual.R"))

run_simulations <- .as_flag(.get_qmd_param_env(
  "run_simulations",
  "RUN_SIMULATIONS",
  FALSE
))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  run_simulations
))
run_stimgate <- .as_flag(.get_qmd_param_env(
  "run_stimgate",
  "RUN_STIMGATE",
  run_simulations
))
run_comparators <- .as_flag(.get_qmd_param_env(
  "run_comparators",
  "RUN_COMPARATORS",
  run_simulations
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
tester_n_sample <- as.integer(.get_qmd_param_env(
  "tester_n_sample",
  "TESTER_N_SAMPLE",
  20L
))
n_workers <- as.integer(.get_qmd_param_env(
  "n_workers",
  "ACS_N_WORKERS",
  2L
))

if (is.na(n_workers) || n_workers < 1L) {
  stop("n_workers must be coercible to an integer of at least one.")
}
.acsCytofBatchList(tester_n_sample)

.acsCytofEnsureCurrentCheckout(root_dir)

# common paths
path_fcs_base <- projr::projr_path_get(
  "raw-data-large",
  "comparison_data", "acscytof", "fcs"
)
if (isTRUE(run_preprocessing) && !dir.exists(path_fcs_base)) {
  stop("Path to FCS files does not exist: ", path_fcs_base)
}
path_gs_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "gs"
)
path_scratch_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "scratch"
)
path_fbeta <- file.path(root_dir, "scripts", "python", "fbeta.py")
if (isTRUE(run_comparators)) {
  if (!requireNamespace("cytoUtils", quietly = TRUE) ||
      !requireNamespace("reticulate", quietly = TRUE) ||
      !file.exists(path_fbeta)) {
    stop("ACS comparisons require cytoUtils, reticulate and scripts/python/fbeta.py.")
  }
  if (!reticulate::py_module_available("numpy")) {
    stop("ACS F-beta requires numpy in the Python environment selected by reticulate.")
  }
}
```

## Scientific settings

```{r}
#| label: scientific-settings
pop_vec <- c("tcrgd", "cd4", "cd8", "nk", "nk_pre", "b")
.by_pop <- function(x) stats::setNames(rep(x, length(pop_vec)), pop_vec)
run_preprocessing_vec_by_pop <- .by_pop(run_preprocessing)
run_stimgate_vec_by_pop <- .by_pop(run_stimgate)
run_comparators_vec_by_pop <- .by_pop(run_comparators)
# Display diagnostics sequentially below, including when using cached runs.
run_stimgate_plots_vec_by_pop <- .by_pop(FALSE)
run_preprocessing_plots_vec_by_pop <- .by_pop(FALSE)
bias_uns <- 0.15
bias_uns_vec_by_pop <- .by_pop(bias_uns)
analysis_seed <- 20260823L
```

## Gating

The TCRgd population is first run on a limited set of complete five-sample
batches. If that tester completes successfully, the same preprocessing and
StimGate flow is applied to every population configured in `pop_vec`.

The reusable helpers fix the six cytokine channels and the method settings.
StimGate uses `bwMtd = "hpi1"` (with `bwNcellMax = 1e4`, `bwFallback = "auto"`
and no bandwidth bounds), `gateCombn = "min"`, `minCell = 100`,
`tolClust = NULL`, `calcCytPosGates = TRUE` and an unstimulated-sample bias
adjustment `biasUns = 0.15`. F-beta uses its standard
parameters (`beta = 0.8`, `theta = 2`, `width = 10` and automatic bin count).
The F-beta implementation is adapted from Richards et al. (2014) only in the
histogram support: common bin edges span the matched stimulated and unstimulated
distributions, instead of being derived from the unstimulated range alone. This
prevents stimulated response-tail events outside the observed unstimulated range
from being omitted when no explicit strong positive control is available.
Tailgate uses the stimulated distribution, automatic bandwidth estimation and
automatic tolerance tuning.

### TCRgd tester

The tester uses only the first `tester_n_sample` files, which must comprise
complete batches of one unstimulated and four stimulated samples. Its cached
GatingSet and StimGate project are kept under a separate `tester/tcrgd`
subdirectory, so they cannot overwrite the full TCRgd run.

```{r}
#| label: run-tcrgd-tester
#| include: false
set.seed(analysis_seed)

tcrgd_tester <- .acsCytofRunPopulation(
  pop = "tcrgd",
  pathFcsBase = path_fcs_base,
  pathGsBase = path_gs_base,
  pathScratchBase = path_scratch_base,
  runPreprocessing = run_preprocessing,
  runMethods = run_stimgate,
  runPlots = FALSE,
  runPreprocessingPlots = FALSE,
  nSample = tester_n_sample,
  biasUns = bias_uns_vec_by_pop[["tcrgd"]],
  outputGroup = "tester"
)

if (isTRUE(run_comparators)) {
  tcrgd_comparison_tester <- .acsCytofRunComparisonMethods(
    pop = "tcrgd",
    pathFcsBase = path_fcs_base,
    pathGsBase = path_gs_base,
    pathScratchBase = path_scratch_base,
    runMethods = TRUE,
    nSample = tester_n_sample,
    outputGroup = "tester",
    pathFbeta = path_fbeta
  )
}
```

### Across populations, in parallel

Each population writes to its own cached GatingSet and scratch directory. The
number of concurrent population runs defaults to two and can be changed with
the `n_workers` parameter or `ACS_N_WORKERS` environment variable.

```{r}
#| label: run-all-populations
#| include: false
if (isTRUE(run_preprocessing) || isTRUE(run_stimgate)) {
  population_runs <- .acsCytofMapPopulations(
    pop_vec,
    function(pop) {
      .acsCytofRunPopulationSafe(
        pop = pop,
        pathFcsBase = path_fcs_base,
        pathGsBase = path_gs_base,
        pathScratchBase = path_scratch_base,
        runPreprocessing = run_preprocessing_vec_by_pop[[pop]],
        runMethods = run_stimgate_vec_by_pop[[pop]],
        runPlots = run_stimgate_plots_vec_by_pop[[pop]],
        runPreprocessingPlots = run_preprocessing_plots_vec_by_pop[[pop]],
        biasUns = bias_uns_vec_by_pop[[pop]]
      )
    },
    nWorkers = n_workers,
    seed = analysis_seed,
    label = "ACS population preprocessing/StimGate runs"
  )
}
```

## Tailgate and F-beta

Tailgate and F-beta run concurrently across populations, using the same
`n_workers` control as StimGate. Within each population, the two methods remain
sequential so that they do not compete for the same GatingSet. Each method is
saved to its own `result.rds`. Whenever `run_comparators` is true, both methods
are recomputed for every population and replace the previous results only once
both have finished, so a failed rerun keeps the last good results. When
`run_comparators` is false, existing results remain available to
the downstream comparison and plotting sections.

```{r}
#| label: run-comparison-methods-in-parallel
#| include: false
if (isTRUE(run_comparators)) {
  comparison_method_runs <- .acsCytofMapPopulations(
    pop_vec,
    function(pop) {
      .acsCytofRunComparisonMethodsSafe(
        pop = pop,
        pathFcsBase = path_fcs_base,
        pathGsBase = path_gs_base,
        pathScratchBase = path_scratch_base,
        runMethods = run_comparators_vec_by_pop[[pop]],
        pathFbeta = path_fbeta
      )
    },
    nWorkers = n_workers,
    seed = analysis_seed,
    label = "ACS Tailgate/F-beta runs"
  )
}
```

## Compare methods with the ACS reference counts

The cached combination tables for all three methods are reduced to one
frequency per cytokine with `UtilsCytoRSV::sum_over_markers()`. Combination
labels are then normalised to the standard `+`/`-` form with
`UtilsCompassSV::convert_cyt_combn_format()` before the automated and manual
tables are joined. `nk_pre` is still run above, but is omitted here because the
manual file has no matching population.

```{r}
#| label: format-manual-comparison
#| include: false
path_manual_output <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "manual-comparison"
)
path_comparison <- file.path(path_manual_output, "manual-comparison.rds")
if (isTRUE(run_preprocessing) || isTRUE(run_stimgate) || isTRUE(run_comparators)) {
  data("fn_to_sampleid_map", package = "DataTidyACSCyTOFFAUST")

  # Build the complete comparison before touching the last good saved output.
  manual_comparison_tbl <- comp_against_manual_cyt(
    fn = "Cytokines_10Jun2021_2.csv",
    path_scratch_base = path_scratch_base,
    path_fcs_base = path_fcs_base,
    pop = pop_vec,
    methods = c("stimgate", "tailgate", "fbeta"),
    path_dir_save = NULL,
    save_plots = FALSE,
    sample_map = fn_to_sampleid_map
  )
  # Save the complete table without generating figures in this plumbing chunk.
  .acsCytofManualSave(
    comparisonTbl = manual_comparison_tbl,
    pathDirSave = path_manual_output,
    savePlots = FALSE
  )
} else {
  if (!file.exists(path_comparison)) {
    stop(
      "No canonical ACS comparison found at: ", path_comparison,
      ". Run RUN_SIMULATIONS=true RUN_PLOTS=false quarto render ",
      "analysis/9-real-compare-acs-cytof.qmd first with the ACS data available."
    )
  }
  manual_comparison_tbl <- readRDS(path_comparison)
}
manual_summary_tbl <- .acsCytofManualSummaryTable(manual_comparison_tbl)
```

## Output tables and figures

### Correlation with the manual gating

The comparison table, its RDS copy, a method/population/cytokine summary and
(when `run_plots` is true) the comparison PNGs are written under
`path_manual_output`. Pearson correlations of the
StimGate and F-beta estimates with the manual frequencies follow.

```{r}
#| label: show-manual-summary
#| echo: false
manual_summary_tbl |>
  dplyr::filter(!method == "tailgate") |>
  dplyr::select(
    method:stim, pcc
  ) |>
  tidyr::pivot_wider(
    names_from = method,
    values_from = pcc
  ) |>
  dplyr::mutate(
    diff_abs = abs(stimgate - fbeta)
  ) |>
  dplyr::arrange(desc(diff_abs)) |>
  print(n = Inf)
```

### Estimated versus reference response frequency

Manual background-subtracted frequency is on the horizontal axis and automated frequency on the vertical axis. Colour identifies the method, shapes identify stimuli, and panels separate populations and cytokines. Points closer to the dashed identity line indicate closer agreement.

```{r}
#| label: fig-manual-comparison-all-methods
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Automated and manual background-subtracted frequencies for all three methods."
if (isTRUE(run_plots)) {
  manual_plots <- .acsCytofManualSave(
    comparisonTbl = manual_comparison_tbl,
    pathDirSave = path_manual_output,
    savePlots = TRUE
  )
  print(manual_plots$scatter)
}
```

### Relative error across all methods

Boxplots show absolute relative error by method, with separate panels for populations and cytokines. Smaller values indicate closer agreement; rows with non-positive manual background-subtracted frequencies have no relative-error estimate.

```{r}
#| label: fig-manual-relative-error-all-methods
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Absolute relative error for StimGate, Tailgate and F-beta."
if (isTRUE(run_plots)) {
  print(manual_plots$relativeError)
}
```

### StimGate and F-beta frequencies

The same frequency comparison is shown for StimGate and F-beta alone. Colours identify methods and shapes identify stimuli; the dashed line marks equal automated and manual frequencies.

```{r}
#| label: fig-manual-comparison
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "StimGate and F-beta frequencies against manual gating."
if (isTRUE(run_plots)) {
  p <- .acsCytofManualPlotScatter(
    manual_comparison_tbl |> dplyr::filter(method != "tailgate")
  )
  ggsave(
    file.path(path_manual_output, "scatter-no_tg.pdf"),
    plot = p, units = "cm", width = 20, height = 20
  )
  print(p)
}
```

### StimGate and F-beta relative error

Absolute relative error is shown by method, population and cytokine after omitting Tailgate. Compare the boxes and their spread, allowing for the separate vertical scales across panels.

```{r}
#| label: fig-manual-relative-error
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Absolute relative error for StimGate and F-beta."
if (isTRUE(run_plots)) {
  p <- .acsCytofManualPlotRelativeError(
    manual_comparison_tbl |> dplyr::filter(method != "tailgate")
  )
  ggsave(
    file.path(path_manual_output, "rel_error-no_tg.pdf"),
    plot = p, units = "cm", width = 20, height = 20
  )
  print(p)
}
```

### Expression and gate diagnostics

Histograms compare untransformed and asinh-transformed expression in the first sample of each cached GatingSet, with one panel per marker. The gate plots show the second sample for two cytokine channels, with the saved StimGate thresholds overlaid. These checks require cached GatingSets and gate projects; the frequency figures above use only the canonical comparison table.

```{r}
#| label: fig-acs-expression-diagnostics
#| echo: false
#| fig-width: 12
#| fig-height: 9
#| fig-cap: "Marker expression before and after transformation, by population."
if (isTRUE(run_plots)) {
  diagnostic_paths <- c(
    list(.acsCytofPopulationPaths(
      "tcrgd", path_fcs_base, path_gs_base, path_scratch_base,
      outputGroup = "tester"
    )),
    lapply(pop_vec, .acsCytofPopulationPaths,
      pathFcsBase = path_fcs_base, pathGsBase = path_gs_base,
      pathScratchBase = path_scratch_base
    )
  )
  names(diagnostic_paths) <- c("tester/tcrgd", pop_vec)
  for (pop in names(diagnostic_paths)) {
    paths <- diagnostic_paths[[pop]]
    if (!dir.exists(paths$gs)) {
      cat("Cached GatingSet unavailable for ", pop, "; expression diagnostics skipped.\n", sep = "")
      next
    }
    plots <- .acsCytofPlotGatingSetCheck(paths$gs)
    plot_gatingset_check(paths$gs, paths$gsCheck, plots = plots)
    for (p in plots) print(p + ggplot2::labs(subtitle = pop))
  }
}
```

Scatter panels show expression in the two cytokine channels, and marginal panels show their densities in the second sample. Lines mark the saved gate thresholds; compare their position with the expression distributions.

```{r}
#| label: fig-acs-gate-diagnostics
#| echo: false
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Saved StimGate thresholds for the second sample of each population run."
if (isTRUE(run_plots)) {
  for (pop in names(diagnostic_paths)) {
    paths <- diagnostic_paths[[pop]]
    if (!dir.exists(paths$gs) || !dir.exists(paths$stimgate)) {
      cat("Cached GatingSet or gate project unavailable for ", pop, "; gate diagnostics skipped.\n", sep = "")
      next
    }
    gs <- flowWorkspace::load_gs(paths$gs)
    p <- stimgate::plotStim(
      ind = 2L, .data = gs, pathProject = paths$stimgate,
      pop = "root", chnl = c("Ho165Di", "Nd146Di")
    )
    ggplot2::ggsave(paths$stimgateCheck, plot = p,
      width = 18, height = 16, units = "cm"
    )
    print(p + ggplot2::labs(subtitle = pop))
  }
}
```
