---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
params:
  run_preprocessing: true
  run_methods: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "acs_cytof-runtime.R"))
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-preprocess.R"))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  TRUE
))
run_methods <- .as_flag(.get_qmd_param_env(
  "run_methods",
  "RUN_METHODS",
  TRUE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Aim

Validate StimGate on the ACS CyTOF dataset and compare its estimated
background-subtracted response frequencies with Tailgate and F-beta.

## Gating

### Common settings

Here we specify settings common across all the populations gated.

```{r}
path_fcs_base <- projr::projr_path_get(
  "raw-data",
  "comparison_data", "acscytof", "fcs"
)
if (!dir.exists(path_fcs_base)) {
  stop("Path to fcs files does not exist: ", path_fcs_base)
}
path_gs_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "gs"
)
path_scratch_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "scratch"
)
```

```{r}
#| label: settings-common-gatings
chnl_vec <- c("Ho165Di", "Gd158Di", "Nd146Di", "Dy164Di", "Gd156Di", "Nd150Di")

# StimGate settings. gateCombn is intentionally explicit.
stimgate_gate_combn <- "min"
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"
stimgate_bias_uns <- NULL
stimgate_min_cell <- 100
tol_clust <- NULL
calc_cyt_pos_gates <- TRUE

# Comparator settings, matching the corrected analyses 7 and 8 semantics.
fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
```

### TCRgd

```{r}
# directories
path_fcs_tcrgd <- file.path(path_fcs_base, "tcrgd")
path_gs_tcrgd <- file.path(path_gs_base, "tcrgd")
path_scratch_tcrgd <- file.path(path_scratch_base, "tcrgd")
if (!dir.exists(path_scratch_tcrgd)) {
  dir.create(path_scratch_tcrgd, recursive = TRUE, showWarnings = FALSE)
}
path_stimgate_tcrgd <- file.path(path_scratch_tcrgd, "stimgate")
# gating settings
bias_uns_tcrgd <- 0.15
```

#### Pre-processing

##### Create GatingSet

```{r}
#| label: create-gatingset-tcrgd
#| results: hide
if (isTRUE(run_preprocessing)) {
  # ensure fcs directory exists
  if (!dir.exists(path_fcs_tcrgd)) {
    stop("Path to TCRgd FCS files does not exist: ", path_fcs_tcrgd)
  }
  # ensure that there are many fcs files in there
  fcs_files_tcrgd <- list.files(
    path_fcs_tcrgd,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  if (length(fcs_files_tcrgd) < 10L) {
    stop(
      "Expected at least 10 TCRgd FCS files in the directory, but found
  only ", length(fcs_files_tcrgd), "."
    )
  }
  # ensure that the gating set directory exists and is empty
  if (dir.exists(path_gs_tcrgd)) {
    unlink(path_gs_tcrgd, recursive = TRUE)
  }
  dir.create(path_gs_tcrgd, recursive = TRUE, showWarnings = FALSE)
  # create the gatingset
  create_gatingset(
    path_fcs = path_fcs_tcrgd,
    path_gs = path_gs_tcrgd# ,
    # n_sample = 20
  )

  # plot the gatingset to check transformation
  # is appropriate
  # create the scratch directory
  path_check_tcrgd <- file.path(path_scratch_tcrgd, "gatingset")
  if (!dir.exists(path_check_tcrgd)) {
    dir.create(path_check_tcrgd, recursive = TRUE, showWarnings = FALSE)
  }
  plot_gatingset_check(
    path_gs = path_gs_tcrgd,
    path_plot_dir = path_check_tcrgd
  )
  # okay, looking here, we clearly need to use the transformation,
  # as e.g. the DNA markers are just vertical bars otherwise
  # and I know from preprocessing that they should not be so.
}
```

#### Gate

##### StimGate

```{r}
#| label: gate-stimgate-tcrgd
if (run_methods) {
  # ensure a clean stimgate run
  if (dir.exists(path_stimgate_tcrgd)) {
    unlink(path_stimgate_tcrgd, recursive = TRUE)
  }
  dir.create(path_stimgate_tcrgd, recursive = TRUE, showWarnings = FALSE)

  # load gatingset
  gs_tcrgd <- flowWorkspace::load_gs(path_gs_tcrgd)

  # create a batch_list object,
  # which for each element has five indices,
  # the first of which specifies unstim.
  # the elements covered go from 
  # 1 to the length of gs_tcrgd.
  # and every fifth element in gs_tcrgd is an unstim sample.
  # but its index must go FIRST in batch_list
  # (and all the stims come afterwards)
  batch_list_tcrgd <- lapply(seq(1, length(gs_tcrgd), by = 5), function(i) {
    c(i, seq(i + 1, min(i + 4, length(gs_tcrgd))))
  })

  invisible(stimgate::gateStim(
    pathProject = path_stimgate_tcrgd,
    .data = gs_tcrgd,
    popGate = "root",
    batchList = batch_list_tcrgd,
    chnl = chnl_vec,
    biasUns = bias_uns_tcrgd,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  p <- plotStim(
    ind = 2L,
    .data = gs_tcrgd,
    pathProject = path_stimgate_tcrgd,
    pop = "root",
    chnl = c("Ho165Di", "Nd146Di"),
  )

  path_p_stimgate_check_tcrgd <- projr::projr_path_get(
    "cache",
    "acs_cytof",
    "scratch",
    "tcrgd",
    "stimgate_check.pdf"
  )
  ggsave(
    path_p_stimgate_check_tcrgd,
    p,
    width = 18,
    height = 16,
    units = "cm"
  )
}
```

#### Tailgate

```{r}
#| label: run-tailgate-tcrgd
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

#### F-beta

```{r}
#| label: run-fbeta-tcrgd
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

### cd4

```{r}
# directories
path_fcs_cd4 <- file.path(path_fcs_base, "cd4")
path_gs_cd4 <- file.path(path_gs_base, "cd4")
path_scratch_cd4 <- file.path(path_scratch_base, "cd4")
if (!dir.exists(path_scratch_cd4)) {
  dir.create(path_scratch_cd4, recursive = TRUE, showWarnings = FALSE)
}
path_stimgate_cd4 <- file.path(path_scratch_cd4, "stimgate")
# gating settings
bias_uns_cd4 <- 0.15
```

#### Pre-processing

##### Create GatingSet

```{r}
#| label: create-gatingset-cd4
#| results: hide
if (isTRUE(run_preprocessing)) {
  # ensure fcs directory exists
  if (!dir.exists(path_fcs_cd4)) {
    stop("Path to cd4 FCS files does not exist: ", path_fcs_cd4)
  }
  # ensure that there are many fcs files in there
  fcs_files_cd4 <- list.files(
    path_fcs_cd4,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  if (length(fcs_files_cd4) < 10L) {
    stop(
      "Expected at least 10 cd4 FCS files in the directory, but found
  only ", length(fcs_files_cd4), "."
    )
  }
  # ensure that the gating set directory exists and is empty
  if (dir.exists(path_gs_cd4)) {
    unlink(path_gs_cd4, recursive = TRUE)
  }
  dir.create(path_gs_cd4, recursive = TRUE, showWarnings = FALSE)
  # create the gatingset
  create_gatingset(
    path_fcs = path_fcs_cd4,
    path_gs = path_gs_cd4# ,
    # n_sample = 20
  )

  # plot the gatingset to check transformation
  # is appropriate
  # create the scratch directory
  path_check_cd4 <- file.path(path_scratch_cd4, "gatingset")
  if (!dir.exists(path_check_cd4)) {
    dir.create(path_check_cd4, recursive = TRUE, showWarnings = FALSE)
  }
  plot_gatingset_check(
    path_gs = path_gs_cd4,
    path_plot_dir = path_check_cd4
  )
  # okay, looking here, we clearly need to use the transformation,
  # as e.g. the DNA markers are just vertical bars otherwise
  # and I know from preprocessing that they should not be so.
}
```

#### Gate

##### StimGate

```{r}
#| label: gate-stimgate-cd4
if (run_methods) {
  # ensure a clean stimgate run
  if (dir.exists(path_stimgate_cd4)) {
    unlink(path_stimgate_cd4, recursive = TRUE)
  }
  dir.create(path_stimgate_cd4, recursive = TRUE, showWarnings = FALSE)

  # load gatingset
  gs_cd4 <- flowWorkspace::load_gs(path_gs_cd4)

  # create a batch_list object,
  # which for each element has five indices,
  # the first of which specifies unstim.
  # the elements covered go from 
  # 1 to the length of gs_cd4.
  # and every fifth element in gs_cd4 is an unstim sample.
  # but its index must go FIRST in batch_list
  # (and all the stims come afterwards)
  batch_list_cd4 <- lapply(seq(1, length(gs_cd4), by = 5), function(i) {
    c(i, seq(i + 1, min(i + 4, length(gs_cd4))))
  })

  invisible(stimgate::gateStim(
    pathProject = path_stimgate_cd4,
    .data = gs_cd4,
    popGate = "root",
    batchList = batch_list_cd4,
    chnl = chnl_vec,
    biasUns = bias_uns_cd4,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  p <- plotStim(
    ind = 2L,
    .data = gs_cd4,
    pathProject = path_stimgate_cd4,
    pop = "root",
    chnl = c("Ho165Di", "Nd146Di"),
  )

  path_p_stimgate_check_cd4 <- projr::projr_path_get(
    "cache",
    "acs_cytof",
    "scratch",
    "cd4",
    "stimgate_check.pdf"
  )
  ggsave(
    path_p_stimgate_check_cd4,
    p,
    width = 18,
    height = 16,
    units = "cm"
  )
}
```

#### Tailgate

```{r}
#| label: run-tailgate-cd4
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

#### F-beta

```{r}
#| label: run-fbeta-cd4
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

### cd8

```{r}
# directories
path_fcs_cd8 <- file.path(path_fcs_base, "cd8")
path_gs_cd8 <- file.path(path_gs_base, "cd8")
path_scratch_cd8 <- file.path(path_scratch_base, "cd8")
if (!dir.exists(path_scratch_cd8)) {
  dir.create(path_scratch_cd8, recursive = TRUE, showWarnings = FALSE)
}
path_stimgate_cd8 <- file.path(path_scratch_cd8, "stimgate")
# gating settings
bias_uns_cd8 <- 0.15
```

#### Pre-processing

##### Create GatingSet

```{r}
#| label: create-gatingset-cd8
#| results: hide
if (isTRUE(run_preprocessing)) {
  # ensure fcs directory exists
  if (!dir.exists(path_fcs_cd8)) {
    stop("Path to cd8 FCS files does not exist: ", path_fcs_cd8)
  }
  # ensure that there are many fcs files in there
  fcs_files_cd8 <- list.files(
    path_fcs_cd8,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  if (length(fcs_files_cd8) < 10L) {
    stop(
      "Expected at least 10 cd8 FCS files in the directory, but found
  only ", length(fcs_files_cd8), "."
    )
  }
  # ensure that the gating set directory exists and is empty
  if (dir.exists(path_gs_cd8)) {
    unlink(path_gs_cd8, recursive = TRUE)
  }
  dir.create(path_gs_cd8, recursive = TRUE, showWarnings = FALSE)
  # create the gatingset
  create_gatingset(
    path_fcs = path_fcs_cd8,
    path_gs = path_gs_cd8# ,
    # n_sample = 20
  )

  # plot the gatingset to check transformation
  # is appropriate
  # create the scratch directory
  path_check_cd8 <- file.path(path_scratch_cd8, "gatingset")
  if (!dir.exists(path_check_cd8)) {
    dir.create(path_check_cd8, recursive = TRUE, showWarnings = FALSE)
  }
  plot_gatingset_check(
    path_gs = path_gs_cd8,
    path_plot_dir = path_check_cd8
  )
  # okay, looking here, we clearly need to use the transformation,
  # as e.g. the DNA markers are just vertical bars otherwise
  # and I know from preprocessing that they should not be so.
}
```

#### Gate

##### StimGate

```{r}
#| label: gate-stimgate-cd8
if (run_methods) {
  # ensure a clean stimgate run
  if (dir.exists(path_stimgate_cd8)) {
    unlink(path_stimgate_cd8, recursive = TRUE)
  }
  dir.create(path_stimgate_cd8, recursive = TRUE, showWarnings = FALSE)

  # load gatingset
  gs_cd8 <- flowWorkspace::load_gs(path_gs_cd8)

  # create a batch_list object,
  # which for each element has five indices,
  # the first of which specifies unstim.
  # the elements covered go from 
  # 1 to the length of gs_cd8.
  # and every fifth element in gs_cd8 is an unstim sample.
  # but its index must go FIRST in batch_list
  # (and all the stims come afterwards)
  batch_list_cd8 <- lapply(seq(1, length(gs_cd8), by = 5), function(i) {
    c(i, seq(i + 1, min(i + 4, length(gs_cd8))))
  })

  invisible(stimgate::gateStim(
    pathProject = path_stimgate_cd8,
    .data = gs_cd8,
    popGate = "root",
    batchList = batch_list_cd8,
    chnl = chnl_vec,
    biasUns = bias_uns_cd8,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  p <- plotStim(
    ind = 2L,
    .data = gs_cd8,
    pathProject = path_stimgate_cd8,
    pop = "root",
    chnl = c("Ho165Di", "Nd146Di"),
  )

  path_p_stimgate_check_cd8 <- projr::projr_path_get(
    "cache",
    "acs_cytof",
    "scratch",
    "cd8",
    "stimgate_check.pdf"
  )
  ggsave(
    path_p_stimgate_check_cd8,
    p,
    width = 18,
    height = 16,
    units = "cm"
  )
}
```

#### Tailgate

```{r}
#| label: run-tailgate-cd8
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

#### F-beta

```{r}
#| label: run-fbeta-cd8
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

### nk

```{r}
# directories
path_fcs_nk <- file.path(path_fcs_base, "nk")
path_gs_nk <- file.path(path_gs_base, "nk")
path_scratch_nk <- file.path(path_scratch_base, "nk")
if (!dir.exists(path_scratch_nk)) {
  dir.create(path_scratch_nk, recursive = TRUE, showWarnings = FALSE)
}
path_stimgate_nk <- file.path(path_scratch_nk, "stimgate")
# gating settings
bias_uns_nk <- 0.15
```

#### Pre-processing

##### Create GatingSet

```{r}
#| label: create-gatingset-nk
#| results: hide
if (isTRUE(run_preprocessing)) {
  # ensure fcs directory exists
  if (!dir.exists(path_fcs_nk)) {
    stop("Path to nk FCS files does not exist: ", path_fcs_nk)
  }
  # ensure that there are many fcs files in there
  fcs_files_nk <- list.files(
    path_fcs_nk,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  if (length(fcs_files_nk) < 10L) {
    stop(
      "Expected at least 10 nk FCS files in the directory, but found
  only ", length(fcs_files_nk), "."
    )
  }
  # ensure that the gating set directory exists and is empty
  if (dir.exists(path_gs_nk)) {
    unlink(path_gs_nk, recursive = TRUE)
  }
  dir.create(path_gs_nk, recursive = TRUE, showWarnings = FALSE)
  # create the gatingset
  create_gatingset(
    path_fcs = path_fcs_nk,
    path_gs = path_gs_nk# ,
    # n_sample = 20
  )

  # plot the gatingset to check transformation
  # is appropriate
  # create the scratch directory
  path_check_nk <- file.path(path_scratch_nk, "gatingset")
  if (!dir.exists(path_check_nk)) {
    dir.create(path_check_nk, recursive = TRUE, showWarnings = FALSE)
  }
  plot_gatingset_check(
    path_gs = path_gs_nk,
    path_plot_dir = path_check_nk
  )
  # okay, looking here, we clearly need to use the transformation,
  # as e.g. the DNA markers are just vertical bars otherwise
  # and I know from preprocessing that they should not be so.
}
```

#### Gate

##### StimGate

```{r}
#| label: gate-stimgate-nk
if (run_methods) {
  # ensure a clean stimgate run
  if (dir.exists(path_stimgate_nk)) {
    unlink(path_stimgate_nk, recursive = TRUE)
  }
  dir.create(path_stimgate_nk, recursive = TRUE, showWarnings = FALSE)

  # load gatingset
  gs_nk <- flowWorkspace::load_gs(path_gs_nk)

  # create a batch_list object,
  # which for each element has five indices,
  # the first of which specifies unstim.
  # the elements covered go from 
  # 1 to the length of gs_nk.
  # and every fifth element in gs_nk is an unstim sample.
  # but its index must go FIRST in batch_list
  # (and all the stims come afterwards)
  batch_list_nk <- lapply(seq(1, length(gs_nk), by = 5), function(i) {
    c(i, seq(i + 1, min(i + 4, length(gs_nk))))
  })

  invisible(stimgate::gateStim(
    pathProject = path_stimgate_nk,
    .data = gs_nk,
    popGate = "root",
    batchList = batch_list_nk,
    chnl = chnl_vec,
    biasUns = bias_uns_nk,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  p <- plotStim(
    ind = 2L,
    .data = gs_nk,
    pathProject = path_stimgate_nk,
    pop = "root",
    chnl = c("Ho165Di", "Nd146Di"),
  )

  path_p_stimgate_check_nk <- projr::projr_path_get(
    "cache",
    "acs_cytof",
    "scratch",
    "nk",
    "stimgate_check.pdf"
  )
  ggsave(
    path_p_stimgate_check_nk,
    p,
    width = 18,
    height = 16,
    units = "cm"
  )
}
```

#### Tailgate

```{r}
#| label: run-tailgate-nk
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

#### F-beta

```{r}
#| label: run-fbeta-nk
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

### nk_pre

```{r}
# directories
path_fcs_nk_pre <- file.path(path_fcs_base, "nk_pre")
path_gs_nk_pre <- file.path(path_gs_base, "nk_pre")
path_scratch_nk_pre <- file.path(path_scratch_base, "nk_pre")
if (!dir.exists(path_scratch_nk_pre)) {
  dir.create(path_scratch_nk_pre, recursive = TRUE, showWarnings = FALSE)
}
path_stimgate_nk_pre <- file.path(path_scratch_nk_pre, "stimgate")
# gating settings
bias_uns_nk_pre <- 0.15
```

#### Pre-processing

##### Create GatingSet

```{r}
#| label: create-gatingset-nk_pre
#| results: hide
if (isTRUE(run_preprocessing)) {
  # ensure fcs directory exists
  if (!dir.exists(path_fcs_nk_pre)) {
    stop("Path to nk_pre FCS files does not exist: ", path_fcs_nk_pre)
  }
  # ensure that there are many fcs files in there
  fcs_files_nk_pre <- list.files(
    path_fcs_nk_pre,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  if (length(fcs_files_nk_pre) < 10L) {
    stop(
      "Expected at least 10 nk_pre FCS files in the directory, but found
  only ", length(fcs_files_nk_pre), "."
    )
  }
  # ensure that the gating set directory exists and is empty
  if (dir.exists(path_gs_nk_pre)) {
    unlink(path_gs_nk_pre, recursive = TRUE)
  }
  dir.create(path_gs_nk_pre, recursive = TRUE, showWarnings = FALSE)
  # create the gatingset
  create_gatingset(
    path_fcs = path_fcs_nk_pre,
    path_gs = path_gs_nk_pre# ,
    # n_sample = 20
  )

  # plot the gatingset to check transformation
  # is appropriate
  # create the scratch directory
  path_check_nk_pre <- file.path(path_scratch_nk_pre, "gatingset")
  if (!dir.exists(path_check_nk_pre)) {
    dir.create(path_check_nk_pre, recursive = TRUE, showWarnings = FALSE)
  }
  plot_gatingset_check(
    path_gs = path_gs_nk_pre,
    path_plot_dir = path_check_nk_pre
  )
  # okay, looking here, we clearly need to use the transformation,
  # as e.g. the DNA markers are just vertical bars otherwise
  # and I know from preprocessing that they should not be so.
}
```

#### Gate

##### StimGate

```{r}
#| label: gate-stimgate-nk_pre
if (run_methods) {
  # ensure a clean stimgate run
  if (dir.exists(path_stimgate_nk_pre)) {
    unlink(path_stimgate_nk_pre, recursive = TRUE)
  }
  dir.create(path_stimgate_nk_pre, recursive = TRUE, showWarnings = FALSE)

  # load gatingset
  gs_nk_pre <- flowWorkspace::load_gs(path_gs_nk_pre)

  # create a batch_list object,
  # which for each element has five indices,
  # the first of which specifies unstim.
  # the elements covered go from 
  # 1 to the length of gs_nk_pre.
  # and every fifth element in gs_nk_pre is an unstim sample.
  # but its index must go FIRST in batch_list
  # (and all the stims come afterwards)
  batch_list_nk_pre <- lapply(seq(1, length(gs_nk_pre), by = 5), function(i) {
    c(i, seq(i + 1, min(i + 4, length(gs_nk_pre))))
  })

  invisible(stimgate::gateStim(
    pathProject = path_stimgate_nk_pre,
    .data = gs_nk_pre,
    popGate = "root",
    batchList = batch_list_nk_pre,
    chnl = chnl_vec,
    biasUns = bias_uns_nk_pre,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  p <- plotStim(
    ind = 2L,
    .data = gs_nk_pre,
    pathProject = path_stimgate_nk_pre,
    pop = "root",
    chnl = c("Ho165Di", "Nd146Di"),
  )

  path_p_stimgate_check_nk_pre <- projr::projr_path_get(
    "cache",
    "acs_cytof",
    "scratch",
    "nk_pre",
    "stimgate_check.pdf"
  )
  ggsave(
    path_p_stimgate_check_nk_pre,
    p,
    width = 18,
    height = 16,
    units = "cm"
  )
}
```

#### Tailgate

```{r}
#| label: run-tailgate-nk_pre
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

#### F-beta

```{r}
#| label: run-fbeta-nk_pre
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

### b

```{r}
# directories
path_fcs_b <- file.path(path_fcs_base, "b")
path_gs_b <- file.path(path_gs_base, "b")
path_scratch_b <- file.path(path_scratch_base, "b")
if (!dir.exists(path_scratch_b)) {
  dir.create(path_scratch_b, recursive = TRUE, showWarnings = FALSE)
}
path_stimgate_b <- file.path(path_scratch_b, "stimgate")
# gating settings
bias_uns_b <- 0.15
```

#### Pre-processing

##### Create GatingSet

```{r}
#| label: create-gatingset-b
#| results: hide
if (isTRUE(run_preprocessing)) {
  # ensure fcs directory exists
  if (!dir.exists(path_fcs_b)) {
    stop("Path to b FCS files does not exist: ", path_fcs_b)
  }
  # ensure that there are many fcs files in there
  fcs_files_b <- list.files(
    path_fcs_b,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  if (length(fcs_files_b) < 10L) {
    stop(
      "Expected at least 10 b FCS files in the directory, but found
  only ", length(fcs_files_b), "."
    )
  }
  # ensure that the gating set directory exists and is empty
  if (dir.exists(path_gs_b)) {
    unlink(path_gs_b, recursive = TRUE)
  }
  dir.create(path_gs_b, recursive = TRUE, showWarnings = FALSE)
  # create the gatingset
  create_gatingset(
    path_fcs = path_fcs_b,
    path_gs = path_gs_b# ,
    # n_sample = 20
  )

  # plot the gatingset to check transformation
  # is appropriate
  # create the scratch directory
  path_check_b <- file.path(path_scratch_b, "gatingset")
  if (!dir.exists(path_check_b)) {
    dir.create(path_check_b, recursive = TRUE, showWarnings = FALSE)
  }
  plot_gatingset_check(
    path_gs = path_gs_b,
    path_plot_dir = path_check_b
  )
  # okay, looking here, we clearly need to use the transformation,
  # as e.g. the DNA markers are just vertical bars otherwise
  # and I know from preprocessing that they should not be so.
}
```

#### Gate

##### StimGate

```{r}
#| label: gate-stimgate-b
if (run_methods) {
  # ensure a clean stimgate run
  if (dir.exists(path_stimgate_b)) {
    unlink(path_stimgate_b, recursive = TRUE)
  }
  dir.create(path_stimgate_b, recursive = TRUE, showWarnings = FALSE)

  # load gatingset
  gs_b <- flowWorkspace::load_gs(path_gs_b)

  # create a batch_list object,
  # which for each element has five indices,
  # the first of which specifies unstim.
  # the elements covered go from 
  # 1 to the length of gs_b.
  # and every fifth element in gs_b is an unstim sample.
  # but its index must go FIRST in batch_list
  # (and all the stims come afterwards)
  batch_list_b <- lapply(seq(1, length(gs_b), by = 5), function(i) {
    c(i, seq(i + 1, min(i + 4, length(gs_b))))
  })

  invisible(stimgate::gateStim(
    pathProject = path_stimgate_b,
    .data = gs_b,
    popGate = "root",
    batchList = batch_list_b,
    chnl = chnl_vec,
    biasUns = bias_uns_b,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  p <- plotStim(
    ind = 2L,
    .data = gs_b,
    pathProject = path_stimgate_b,
    pop = "root",
    chnl = c("Ho165Di", "Nd146Di"),
  )

  path_p_stimgate_check_b <- projr::projr_path_get(
    "cache",
    "acs_cytof",
    "scratch",
    "b",
    "stimgate_check.pdf"
  )
  ggsave(
    path_p_stimgate_check_b,
    p,
    width = 18,
    height = 16,
    units = "cm"
  )
}
```

#### Tailgate

```{r}
#| label: run-tailgate-b
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

#### F-beta

```{r}
#| label: run-fbeta-b
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

## Compare methods with the ACS reference counts

## Output tables and figures

### Save analysis tables

### Estimated versus reference response frequency

### Relative error by method
