---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
params:
  run_preprocessing: true
  run_methods: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  FALSE
))
run_methods <- .as_flag(.get_qmd_param_env(
  "run_methods",
  "RUN_METHODS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

if (!requireNamespace("cytoUtils", quietly = TRUE)) {
  warning(
    "Package 'cytoUtils' is not installed. Tailgate comparisons will fail if run."
  )
}

# Reuse the current comparator implementations from analyses 7 and 8.
# These helpers should eventually be split into generic comparison helpers if
# the real-data analysis needs to use them more broadly.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

.acs_assert_file <- function(path, description) {
  if (!file.exists(path)) {
    stop(description, " not found at: ", path)
  }
  invisible(path)
}

.acs_assert_dir <- function(path, description) {
  if (!dir.exists(path)) {
    stop(description, " not found at: ", path)
  }
  invisible(path)
}

.acs_get_expr <- function(gs, ind, chnl, pop = "root") {
  gh <- gs[[as.integer(ind)]]
  ff <- flowWorkspace::gh_pop_get_data(gh, pop)
  ex <- flowCore::exprs(ff)
  if (!chnl %in% colnames(ex)) {
    stop("Channel '", chnl, "' is not present in sample index ", ind, ".")
  }
  as.numeric(ex[, chnl])
}

.acs_estimate_from_threshold <- function(x_stim, x_uns, threshold) {
  threshold <- suppressWarnings(as.numeric(threshold))[1]
  n_stim <- length(x_stim)
  n_uns <- length(x_uns)

  if (!is.finite(threshold)) {
    return(tibble::tibble(
      threshold = NA_real_,
      nCellStim = n_stim,
      nCellUns = n_uns,
      nPosStim = NA_integer_,
      nPosUns = NA_integer_,
      propStim = NA_real_,
      propUns = NA_real_,
      propRespEst = NA_real_
    ))
  }

  n_pos_stim <- sum(x_stim > threshold, na.rm = TRUE)
  n_pos_uns <- sum(x_uns > threshold, na.rm = TRUE)
  prop_stim <- n_pos_stim / n_stim
  prop_uns <- n_pos_uns / n_uns

  tibble::tibble(
    threshold = threshold,
    nCellStim = n_stim,
    nCellUns = n_uns,
    nPosStim = n_pos_stim,
    nPosUns = n_pos_uns,
    propStim = prop_stim,
    propUns = prop_uns,
    propRespEst = prop_stim - prop_uns
  )
}

.acs_stim_pair_tbl <- function(batch_list, sample_metadata) {
  purrr::imap_dfr(batch_list, function(ind_batch, batch_nm) {
    ind_uns <- ind_batch[[1]]
    ind_stim <- ind_batch[-1]

    tibble::tibble(
      batch = batch_nm,
      indUns = ind_uns,
      indStim = ind_stim,
      sampleUns = sample_metadata$sample_name[ind_uns],
      sampleStim = sample_metadata$sample_name[ind_stim],
      conditionStim = sample_metadata$condition[ind_stim]
    )
  })
}
```

### Aim

Validate StimGate on the ACS CyTOF dataset and compare its estimated
background-subtracted response frequencies with Tailgate and F-beta.

This analysis is deliberately organised differently from the simulation analyses.
The raw FCS files are first read and converted into a cached `GatingSet`. Once the
input scale, transformations, sample structure and marker mapping have been
verified, the three methods are run sequentially on the same data. No part of the
analysis is expected to require parallel execution.

The primary StimGate comparison must use `gateCombn = "min"`, so that within each
batch the lowest genuinely generated local-FDR threshold across stimulated
conditions is used for the batch-level gate combination. This setting is explicit
below and should not be left to a default.

## Pre-processing

### Data and cache paths

```{r}
#| label: preprocessing-settings
# TODO #346: confirm the exact raw-data subdirectory and file layout.
acs_raw_subdir <- "acs-cytof"
acs_raw_dir <- projr::projr_path_get("raw-data", acs_raw_subdir)

# The analysis currently expects these small metadata files beside the FCS data.
# Adjust the names here once the existing ACS files/reference material have been
# located. The FCS files themselves may live anywhere below acs_raw_dir.
acs_sample_metadata_path <- file.path(acs_raw_dir, "sample_metadata.csv")
acs_marker_metadata_path <- file.path(acs_raw_dir, "marker_metadata.csv")
acs_reference_counts_path <- file.path(acs_raw_dir, "reference_counts.csv")

acs_cache_dir <- projr::projr_path_get("cache", "real", "acs-cytof")
acs_gs_path <- file.path(acs_cache_dir, "gatingset")
acs_sample_metadata_cache <- file.path(acs_cache_dir, "sample_metadata.csv")
acs_marker_metadata_cache <- file.path(acs_cache_dir, "marker_metadata.csv")
acs_batch_list_path <- file.path(acs_cache_dir, "batch_list.rds")

acs_output_dir <- projr::projr_path_get("cache-analysis", "9-acs-cytof")
acs_table_dir <- file.path(acs_output_dir, "tables")
acs_figure_dir <- file.path(acs_output_dir, "figures")

dir.create(acs_cache_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(acs_table_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(acs_figure_dir, recursive = TRUE, showWarnings = FALSE)

# TODO #346: set these only after the ACS input scale has been checked.
# Use "none" when the stored FCS values are already on the analysis scale.
# Use "asinh" only if the raw values need an asinh transform here.
acs_transform_name <- NA_character_
acs_transform_cofactor <- NA_real_

# Metadata contract for this QMD.
acs_unstim_label <- "unstim"
acs_pop_gate <- "root"

# StimGate settings. Keep the batch combination explicit.
stimgate_gate_combn <- "min"
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"
stimgate_bias_uns <- NULL
stimgate_min_cell <- 100
tol_clust <- 1e-7
loc_enforce_shape_threshold <- FALSE
calc_cyt_pos_gates <- FALSE

# Comparator settings, matching the corrected analyses 7 and 8 semantics.
fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
```

### Read FCS files and construct the analysis GatingSet

```{r}
#| label: preprocessing-fcs
#| results: hide
if (isTRUE(run_preprocessing)) {
  .acs_assert_dir(acs_raw_dir, "ACS raw-data directory")
  .acs_assert_file(acs_sample_metadata_path, "ACS sample metadata")
  .acs_assert_file(acs_marker_metadata_path, "ACS marker metadata")

  sample_metadata <- utils::read.csv(
    acs_sample_metadata_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )
  marker_metadata <- utils::read.csv(
    acs_marker_metadata_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  required_sample_cols <- c("fcs_file", "sample_name", "batch", "condition")
  missing_sample_cols <- setdiff(required_sample_cols, names(sample_metadata))
  if (length(missing_sample_cols) > 0L) {
    stop(
      "ACS sample metadata is missing columns: ",
      paste(missing_sample_cols, collapse = ", ")
    )
  }

  required_marker_cols <- c("channel", "marker")
  missing_marker_cols <- setdiff(required_marker_cols, names(marker_metadata))
  if (length(missing_marker_cols) > 0L) {
    stop(
      "ACS marker metadata is missing columns: ",
      paste(missing_marker_cols, collapse = ", ")
    )
  }

  fcs_paths_all <- list.files(
    acs_raw_dir,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  fcs_lookup <- stats::setNames(fcs_paths_all, basename(fcs_paths_all))

  missing_fcs <- setdiff(sample_metadata$fcs_file, names(fcs_lookup))
  if (length(missing_fcs) > 0L) {
    stop(
      "Could not find ACS FCS files listed in sample_metadata.csv: ",
      paste(missing_fcs, collapse = ", ")
    )
  }

  fcs_paths <- unname(fcs_lookup[sample_metadata$fcs_file])

  ff_list <- lapply(
    fcs_paths,
    flowCore::read.FCS,
    transformation = FALSE,
    truncate_max_range = FALSE
  )

  sample_metadata$n_cell <- vapply(ff_list, nrow, integer(1))

  if (is.na(acs_transform_name)) {
    stop(
      "ACS transformation has not yet been verified. Set acs_transform_name ",
      "to 'none' or 'asinh' after checking whether the FCS values are already ",
      "transformed. No analysis-scale GatingSet has been cached."
    )
  }

  if (identical(acs_transform_name, "asinh")) {
    if (!is.finite(acs_transform_cofactor) || acs_transform_cofactor <= 0) {
      stop("A positive acs_transform_cofactor is required for the asinh transform.")
    }

    transform_channels <- marker_metadata$channel
    ff_list <- lapply(ff_list, function(ff) {
      ex <- flowCore::exprs(ff)
      missing_channels <- setdiff(transform_channels, colnames(ex))
      if (length(missing_channels) > 0L) {
        stop(
          "Marker channels missing from an ACS FCS file: ",
          paste(missing_channels, collapse = ", ")
        )
      }
      ex[, transform_channels] <- asinh(
        ex[, transform_channels, drop = FALSE] / acs_transform_cofactor
      )
      flowCore::exprs(ff) <- ex
      ff
    })
  } else if (!identical(acs_transform_name, "none")) {
    stop(
      "Unsupported acs_transform_name: ", acs_transform_name,
      ". Add the verified transformation explicitly rather than guessing."
    )
  }

  fs <- flowCore::flowSet(ff_list)
  flowCore::sampleNames(fs) <- sample_metadata$sample_name
  gs <- flowWorkspace::GatingSet(fs)

  if (dir.exists(acs_gs_path)) {
    unlink(acs_gs_path, recursive = TRUE)
  }
  flowWorkspace::save_gs(gs, path = acs_gs_path)

  utils::write.csv(
    sample_metadata,
    acs_sample_metadata_cache,
    row.names = FALSE
  )
  utils::write.csv(
    marker_metadata,
    acs_marker_metadata_cache,
    row.names = FALSE
  )

  batch_list <- stimgate::getBatchList(
    fnTblInfo = sample_metadata,
    colGrp = "batch",
    colStim = "condition",
    unsChr = acs_unstim_label,
    colNCell = "n_cell",
    minCell = stimgate_min_cell
  )

  if (length(batch_list) == 0L) {
    stop("No valid ACS batches remain after constructing batch_list.")
  }
  saveRDS(batch_list, acs_batch_list_path)
}
```

### Load cached pre-processing outputs

```{r}
#| label: preprocessing-load
if (!dir.exists(acs_gs_path)) {
  stop(
    "Cached ACS GatingSet not found. Run the pre-processing section after ",
    "the transformation has been verified."
  )
}

.acs_assert_file(acs_sample_metadata_cache, "Cached ACS sample metadata")
.acs_assert_file(acs_marker_metadata_cache, "Cached ACS marker metadata")
.acs_assert_file(acs_batch_list_path, "Cached ACS batch list")

gs_acs <- flowWorkspace::load_gs(acs_gs_path)
sample_metadata <- utils::read.csv(
  acs_sample_metadata_cache,
  stringsAsFactors = FALSE,
  check.names = FALSE
)
marker_metadata <- utils::read.csv(
  acs_marker_metadata_cache,
  stringsAsFactors = FALSE,
  check.names = FALSE
)
batch_list <- readRDS(acs_batch_list_path)
stim_pair_tbl <- .acs_stim_pair_tbl(batch_list, sample_metadata)

stim_pair_tbl
marker_metadata
```

## StimGate

```{r}
#| label: run-stimgate
#| results: hide
path_stimgate <- file.path(acs_cache_dir, "stimgate")
path_stimgate_results <- file.path(acs_cache_dir, "stimgate_results.rds")
path_stimgate_gates <- file.path(acs_cache_dir, "stimgate_gates.rds")
path_stimgate_gates_detailed <- file.path(
  acs_cache_dir,
  "stimgate_gates_detailed.rds"
)

if (isTRUE(run_methods)) {
  chnl_settings <- stats::setNames(
    lapply(marker_metadata$channel, function(chnl_curr) {
      list(cut = chnl_curr)
    }),
    marker_metadata$marker
  )

  if (dir.exists(path_stimgate)) {
    unlink(path_stimgate, recursive = TRUE)
  }

  invisible(stimgate::gateStim(
    pathProject = path_stimgate,
    .data = gs_acs,
    popGate = acs_pop_gate,
    batchList = batch_list,
    chnl = chnl_settings,
    biasUns = stimgate_bias_uns,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    locEnforceShapeThreshold = loc_enforce_shape_threshold,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  stimgate_gates <- stimgate::getStimGates(path_stimgate)
  stimgate_gates_detailed <- stimgate::getStimGatesDetailed(path_stimgate)

  saveRDS(stimgate_gates, path_stimgate_gates)
  saveRDS(stimgate_gates_detailed, path_stimgate_gates_detailed)

  stimgate_results <- purrr::pmap_dfr(
    tidyr::crossing(
      stim_pair_tbl,
      marker_metadata[, c("channel", "marker"), drop = FALSE]
    ),
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      gate_rows <- stimgate_gates |>
        dplyr::filter(
          as.character(.data$ind) == as.character(indStim),
          .data$chnl == channel
        )

      if (nrow(gate_rows) > 0L && "gateName" %in% names(gate_rows)) {
        if (any(grepl("Clust$", gate_rows$gateName))) {
          gate_rows <- gate_rows[grepl("Clust$", gate_rows$gateName), , drop = FALSE]
        }
        gate_rows <- gate_rows[nrow(gate_rows), , drop = FALSE]
      }

      threshold <- if (nrow(gate_rows) > 0L) {
        suppressWarnings(as.numeric(gate_rows$gate[[1]]))
      } else {
        NA_real_
      }
      gate_name <- if (nrow(gate_rows) > 0L && "gateName" %in% names(gate_rows)) {
        as.character(gate_rows$gateName[[1]])
      } else {
        NA_character_
      }

      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      .acs_estimate_from_threshold(x_stim, x_uns, threshold) |>
        dplyr::mutate(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "stimgate",
          thresholdOrigin = gate_name,
          thresholdFallbackUsed = !is.finite(threshold),
          .before = 1
        )
    }
  )

  saveRDS(stimgate_results, path_stimgate_results)
} else {
  stimgate_results <- readRDS(path_stimgate_results)
  stimgate_gates <- readRDS(path_stimgate_gates)
  stimgate_gates_detailed <- readRDS(path_stimgate_gates_detailed)
}
```

### Check the `min` batch gate combination

```{r}
#| label: check-stimgate-min
if (!identical(stimgate_gate_combn, "min")) {
  stop("ACS StimGate comparison must use gateCombn = 'min'.")
}

# Keep the detailed threshold diagnostics available for explicit inspection of
# the condition-level generated thresholds and the subsequent batch-combined
# threshold. This is intentionally a visible analysis check rather than relying
# only on the gateStim() call argument.
min_gate_check_tbl <- stimgate_gates_detailed |>
  dplyr::filter(.data$detailLevel %in% c("condition", "sample", "cluster_final")) |>
  dplyr::arrange(.data$chnl, .data$ind, .data$detailLevel)

saveRDS(
  min_gate_check_tbl,
  file.path(acs_table_dir, "stimgate_min_gate_diagnostics.rds")
)
min_gate_check_tbl
```

## Tailgate

```{r}
#| label: run-tailgate
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    tidyr::crossing(
      stim_pair_tbl,
      marker_metadata[, c("channel", "marker"), drop = FALSE]
    ),
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = thresholdFallbackUsed,
          nCellStim = nCellStim,
          nCellUns = nCellUns,
          nPosStim = nPosStim,
          nPosUns = nPosUns,
          propStim = propStim,
          propUns = propUns,
          propRespEst = propRespEst
        )
    }
  )

  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

## F-beta

```{r}
#| label: run-fbeta
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    tidyr::crossing(
      stim_pair_tbl,
      marker_metadata[, c("channel", "marker"), drop = FALSE]
    ),
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = thresholdFallbackUsed,
          nCellStim = nCellStim,
          nCellUns = nCellUns,
          nPosStim = nPosStim,
          nPosUns = nPosUns,
          propStim = propStim,
          propUns = propUns,
          propRespEst = propRespEst
        )
    }
  )

  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

## Compare methods with the ACS reference counts

```{r}
#| label: compare-methods
method_results <- dplyr::bind_rows(
  stimgate_results,
  tailgate_results,
  fbeta_results
)

if (file.exists(acs_reference_counts_path)) {
  reference_counts <- utils::read.csv(
    acs_reference_counts_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  required_reference_cols <- c("sample_name", "channel")
  missing_reference_cols <- setdiff(
    required_reference_cols,
    names(reference_counts)
  )
  if (length(missing_reference_cols) > 0L) {
    stop(
      "ACS reference table is missing columns: ",
      paste(missing_reference_cols, collapse = ", ")
    )
  }

  if (!"reference_prop" %in% names(reference_counts)) {
    if (all(c("reference_count", "reference_n") %in% names(reference_counts))) {
      reference_counts <- reference_counts |>
        dplyr::mutate(
          reference_prop = .data$reference_count / .data$reference_n
        )
    } else {
      stop(
        "ACS reference table must contain reference_prop, or both ",
        "reference_count and reference_n."
      )
    }
  }

  comparison_tbl <- method_results |>
    dplyr::left_join(
      reference_counts,
      by = c("sampleStim" = "sample_name", "channel")
    ) |>
    dplyr::mutate(
      error = .data$propRespEst - .data$reference_prop,
      abs_error = abs(.data$error),
      rel_error = dplyr::if_else(
        .data$reference_prop == 0,
        NA_real_,
        .data$error / .data$reference_prop
      )
    )
} else {
  warning(
    "ACS reference counts have not yet been located. Method outputs are ",
    "available, but reference-based comparison is deferred."
  )
  comparison_tbl <- method_results
}

comparison_tbl
```

## Output tables and figures

### Save analysis tables

```{r}
#| label: save-output-tables
saveRDS(method_results, file.path(acs_table_dir, "method_results.rds"))
saveRDS(comparison_tbl, file.path(acs_table_dir, "comparison_with_reference.rds"))

utils::write.csv(
  method_results,
  file.path(acs_table_dir, "method_results.csv"),
  row.names = FALSE
)
utils::write.csv(
  comparison_tbl,
  file.path(acs_table_dir, "comparison_with_reference.csv"),
  row.names = FALSE
)

if (exists("reference_counts")) {
  saveRDS(reference_counts, file.path(acs_table_dir, "reference_counts.rds"))
}
```

### Method estimates against reference values

```{r}
#| label: plot-estimate-v-reference
#| fig-width: 9
#| fig-height: 7
if (
  isTRUE(run_plots) &&
    all(c("reference_prop", "propRespEst") %in% names(comparison_tbl))
) {
  p_est_ref <- ggplot(
    comparison_tbl,
    aes(
      x = reference_prop,
      y = propRespEst,
      colour = method
    )
  ) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    geom_point(alpha = 0.75) +
    facet_wrap(~marker, scales = "free") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    labs(
      title = "ACS CyTOF: Estimated vs Reference Response Frequency",
      x = "Reference response frequency",
      y = "Estimated background-subtracted response frequency",
      colour = "Method"
    )

  ggsave(
    file.path(acs_figure_dir, "estimate_vs_reference.pdf"),
    p_est_ref,
    width = 18,
    height = 14,
    units = "cm"
  )
  print(p_est_ref)
}
```

### Relative error by method

```{r}
#| label: plot-relative-error
#| fig-width: 9
#| fig-height: 7
if (
  isTRUE(run_plots) &&
    "rel_error" %in% names(comparison_tbl)
) {
  p_rel_error <- ggplot(
    comparison_tbl,
    aes(
      x = method,
      y = rel_error
    )
  ) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.15, alpha = 0.5) +
    facet_wrap(~marker, scales = "free_y") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "y", minor = "none") +
    labs(
      title = "ACS CyTOF: Relative Error by Method",
      x = NULL,
      y = "Relative error"
    )

  ggsave(
    file.path(acs_figure_dir, "relative_error_by_method.pdf"),
    p_rel_error,
    width = 18,
    height = 14,
    units = "cm"
  )
  print(p_rel_error)
}
```

### Notes for final outputs

The plots above are deliberately initial comparison outputs. Once the ACS
reference counts and their exact denominator are recovered, this section should
be reduced to the final thesis figures and tables rather than retaining every
exploratory diagnostic.
