---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
params:
  run_preprocessing: false
  run_stimgate: false
  run_comparators: true
  overwrite_comparator_cache: true
  run_plots: true
  tester_n_sample: 20
  n_workers: 2
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-preprocess.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-misc.R"))
source(file.path(scripts_r_dir, "sim-compare-freq_bs.R"))
source(file.path(scripts_r_dir, "acs_cytof-methods.R"))
source(file.path(scripts_r_dir, "acs_cytof-manual.R"))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  FALSE
))
pop_vec <- c("tcrgd", "cd4", "cd8", "nk", "nk_pre", "b")
run_preprocessing_vec_by_pop <- c(
  tcrgd = NULL %||% run_preprocessing,
  cd4 = NULL %||% run_preprocessing,
  cd8 = NULL %||% run_preprocessing,
  nk = NULL %||% run_preprocessing,
  nk_pre = NULL %||% run_preprocessing,
  b = NULL %||% run_preprocessing
)
run_stimgate <- .as_flag(.get_qmd_param_env(
  "run_stimgate",
  "RUN_STIMGATE",
  FALSE
))
run_stimgate_vec_by_pop <- c(
  tcrgd = NULL %||% run_stimgate,
  cd4 = NULL %||% run_stimgate,
  cd8 = NULL %||% run_stimgate,
  nk = NULL %||% run_stimgate,
  nk_pre = NULL %||% run_stimgate,
  b = NULL %||% run_stimgate
)
run_comparators <- .as_flag(.get_qmd_param_env(
  "run_comparators",
  "RUN_COMPARATORS",
  TRUE
))
run_comparators_vec_by_pop <- c(
  tcrgd = NULL %||% run_comparators,
  cd4 = NULL %||% run_comparators,
  cd8 = NULL %||% run_comparators,
  nk = NULL %||% run_comparators,
  nk_pre = NULL %||% run_comparators,
  b = NULL %||% run_comparators
)
overwrite_comparator_cache <- .as_flag(.get_qmd_param_env(
  "overwrite_comparator_cache",
  "OVERWRITE_COMPARATOR_CACHE",
  TRUE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
run_plots_vec_by_pop <- c(
  tcrgd = NULL %||% run_plots,
  cd4 = NULL %||% run_plots,
  cd8 = NULL %||% run_plots,
  nk = NULL %||% run_plots,
  nk_pre = NULL %||% run_plots,
  b = NULL %||% run_plots
)
tester_n_sample <- .get_qmd_param_env(
  "tester_n_sample",
  "TESTER_N_SAMPLE",
  20L
)
n_workers <- max(
  1L,
  as.integer(.get_qmd_param_env(
    "n_workers",
    "ACS_N_WORKERS",
    2L
  ))
)
bias_uns <- 0.15
bias_uns_vec_by_pop <- c(
  tcrgd = NULL %||% bias_uns,
  cd4 = NULL %||% bias_uns,
  cd8 = NULL %||% bias_uns,
  nk = NULL %||% bias_uns,
  nk_pre = NULL %||% bias_uns,
  b = NULL %||% bias_uns
)

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Aim

Validate StimGate on the ACS CyTOF dataset and compare its estimated
background-subtracted response frequencies with Tailgate and F-beta.

## Gating

The TCRgd population is first run on a limited set of complete five-sample
batches. If that tester completes successfully, the same preprocessing and
StimGate flow is applied to every population configured in `pop_vec`.

The reusable helpers fix the six cytokine channels and the method settings.
StimGate uses `bwMtd = "hpi1"` and `gateCombn = "min"`. F-beta uses its standard
parameters (`beta = 0.8`, `theta = 2`, `width = 10` and automatic bin count).
Tailgate uses the stimulated distribution, automatic bandwidth estimation and
automatic tolerance tuning.

### Common paths

```{r}
#| label: settings-common-gating-paths
path_fcs_base <- projr::projr_path_get(
  "raw-data-large",
  "comparison_data", "acscytof", "fcs"
)
if (!dir.exists(path_fcs_base)) {
  stop("Path to FCS files does not exist: ", path_fcs_base)
}
path_gs_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "gs"
)
path_scratch_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "scratch"
)
```

### TCRgd tester

The tester uses only the first `tester_n_sample` files, which must comprise
complete batches of one unstimulated and four stimulated samples. Its cached
GatingSet and StimGate project are kept under a separate `tester/tcrgd`
subdirectory, so they cannot overwrite the full TCRgd run.

```{r}
#| label: run-tcrgd-tester
#| results: hide
tcrgd_tester <- .acsCytofRunPopulation(
  pop = "tcrgd",
  pathFcsBase = path_fcs_base,
  pathGsBase = path_gs_base,
  pathScratchBase = path_scratch_base,
  runPreprocessing = run_preprocessing,
  runMethods = run_stimgate,
  runPlots = run_plots,
  nSample = tester_n_sample,
  biasUns = bias_uns_vec_by_pop[["tcrgd"]],
  outputGroup = "tester"
)

tcrgd_comparison_tester <- .acsCytofRunComparisonMethods(
  pop = "tcrgd",
  pathFcsBase = path_fcs_base,
  pathGsBase = path_gs_base,
  pathScratchBase = path_scratch_base,
  runMethods = run_comparators,
  nSample = tester_n_sample,
  outputGroup = "tester",
  overwrite = overwrite_comparator_cache
)
```

### Across populations, in parallel

Each population writes to its own cached GatingSet and scratch directory. The
number of concurrent population runs defaults to two and can be changed with
the `n_workers` parameter or `ACS_N_WORKERS` environment variable.

```{r}
#| label: run-all-populations
#| results: hide
n_workers_use <- min(n_workers, length(pop_vec))
old_plan <- future::plan()

Sys.setenv("STIMGATE_DEBUG" = "TRUE")

population_runs <- tryCatch(
  {
    if (n_workers_use > 1L) {
      future::plan(future::multisession, workers = n_workers_use)
    } else {
      future::plan(future::sequential)
    }

    furrr::future_map(
      pop_vec,
      function(pop) {
        .acsCytofRunPopulationSafe(
          pop = pop,
          pathFcsBase = path_fcs_base,
          pathGsBase = path_gs_base,
          pathScratchBase = path_scratch_base,
          runPreprocessing = run_preprocessing_vec_by_pop[[pop]],
          runMethods = run_stimgate_vec_by_pop[[pop]],
          runPlots = run_plots_vec_by_pop[[pop]],
          biasUns = bias_uns_vec_by_pop[[pop]]
        )
      },
      .options = furrr::furrr_options(seed = TRUE, scheduling = Inf)
    )
  },
  finally = future::plan(old_plan)
)
names(population_runs) <- pop_vec
failed_pop_vec <- names(population_runs)[
  !vapply(
    population_runs,
    function(x) isTRUE(x$success),
    logical(1)
  )
]

if (length(failed_pop_vec) > 0L) {
  warning(
    "ACS CyTOF populations failed: ",
    paste(failed_pop_vec, collapse = ", ")
  )

  for (pop in failed_pop_vec) {
    message(
      pop,
      ": ",
      population_runs[[pop]]$error
    )
  }
}
```

## Tailgate and F-beta

Tailgate and F-beta run one population at a time. Each method is saved
immediately to its own `result.rds`; a later render reuses a complete cache when
the sample count and fixed settings still match. Set
`overwrite_comparator_cache` to `true` only when those method outputs need to
be recomputed.

```{r}
#| label: run-comparison-methods-sequentially
#| results: hide
comparison_method_runs <- purrr::map(
  pop_vec,
  function(pop) {
    .acsCytofRunComparisonMethodsSafe(
      pop = pop,
      pathFcsBase = path_fcs_base,
      pathGsBase = path_gs_base,
      pathScratchBase = path_scratch_base,
      runMethods = run_comparators_vec_by_pop[[pop]],
      overwrite = overwrite_comparator_cache
    )
  }
)
names(comparison_method_runs) <- pop_vec

failed_comparator_pop_vec <- names(comparison_method_runs)[
  !vapply(
    comparison_method_runs,
    function(x) isTRUE(x$success),
    logical(1)
  )
]
if (length(failed_comparator_pop_vec) > 0L) {
  stop(
    "ACS Tailgate/F-beta runs failed for: ",
    paste(failed_comparator_pop_vec, collapse = ", "),
    ". See the returned error entries in comparison_method_runs."
  )
}
```

## Compare methods with the ACS reference counts

The cached combination tables for all three methods are reduced to one
frequency per cytokine with `UtilsCytoRSV::sum_over_markers()`. Combination
labels are then normalised to the standard `+`/`-` form with
`UtilsCompassSV::convert_cyt_combn_format()` before the automated and manual
tables are joined. `nk_pre` is still run above, but is omitted here because the
manual file has no matching population.

```{r}
#| label: format-manual-comparison
if (isTRUE(run_plots)) {
  path_manual_output <- file.path(
    path_scratch_base,
    "manual-comparison"
  )
  if (dir.exists(path_manual_output)) {
    unlink(path_manual_output, recursive = TRUE)
  }
  dir.create(path_manual_output, showWarnings = FALSE, recursive = TRUE)
  manual_comparison_tbl <- comp_against_manual_cyt(
    fn = "Cytokines_10Jun2021_2.csv",
    path_scratch_base = path_scratch_base,
    path_fcs_base = path_fcs_base,
    pop = pop_vec,
    methods = c("stimgate", "tailgate", "fbeta"),
    path_dir_save = path_manual_output,
    save_plots = run_plots
  )
  manual_summary_tbl <- attr(manual_comparison_tbl, "summary")
}
```

## Output tables and figures

### Save analysis tables

The comparison table, its RDS copy and a method/population/cytokine summary are
written under `path_manual_output` by the formatting step above.

```{r}
#| label: show-manual-summary
manual_summary_tbl
```

### Estimated versus reference response frequency

```{r}
#| label: plot-manual-comparison
if (isTRUE(run_plots)) {
  .acsCytofManualPlotScatter(manual_comparison_tbl)
}
```

### Relative error by method

```{r}
#| label: plot-manual-relative-error
if (isTRUE(run_plots)) {
  .acsCytofManualPlotRelativeError(manual_comparison_tbl)
}
```
