---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
params:
  run_preprocessing: true
  run_methods: true
  run_plots: false
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  FALSE
))
run_methods <- .as_flag(.get_qmd_param_env(
  "run_methods",
  "RUN_METHODS",
  FALSE
))
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}

if (!requireNamespace("cytoUtils", quietly = TRUE)) {
  warning(
    "Package 'cytoUtils' is not installed. Tailgate comparisons will fail if run."
  )
}

# Reuse the corrected comparator implementations from analyses 7 and 8.
for (fn in c(
  "sim-misc.R",
  "sim-bandwidth.R",
  "sim-compare-freq_bs.R"
)) {
  source(file.path(scripts_r_dir, fn))
}

.acs_assert_file <- function(path, description) {
  if (!file.exists(path)) {
    stop(description, " not found at: ", path)
  }
  invisible(path)
}

.acs_assert_dir <- function(path, description) {
  if (!dir.exists(path)) {
    stop(description, " not found at: ", path)
  }
  invisible(path)
}

.acs_get_expr <- function(gs, ind, chnl, pop = "root") {
  gh <- gs[[as.integer(ind)]]
  ff <- flowWorkspace::gh_pop_get_data(gh, pop)
  ex <- flowCore::exprs(ff)
  if (!chnl %in% colnames(ex)) {
    stop("Channel '", chnl, "' is not present in sample index ", ind, ".")
  }
  as.numeric(ex[, chnl])
}

.acs_estimate_from_threshold <- function(x_stim, x_uns, threshold) {
  threshold <- suppressWarnings(as.numeric(threshold))[1]
  n_stim <- length(x_stim)
  n_uns <- length(x_uns)

  if (!is.finite(threshold)) {
    return(tibble::tibble(
      threshold = NA_real_,
      nCellStim = n_stim,
      nCellUns = n_uns,
      nPosStim = NA_integer_,
      nPosUns = NA_integer_,
      propStim = NA_real_,
      propUns = NA_real_,
      propRespEst = NA_real_
    ))
  }

  n_pos_stim <- sum(x_stim > threshold, na.rm = TRUE)
  n_pos_uns <- sum(x_uns > threshold, na.rm = TRUE)
  prop_stim <- n_pos_stim / n_stim
  prop_uns <- n_pos_uns / n_uns

  tibble::tibble(
    threshold = threshold,
    nCellStim = n_stim,
    nCellUns = n_uns,
    nPosStim = n_pos_stim,
    nPosUns = n_pos_uns,
    propStim = prop_stim,
    propUns = prop_uns,
    propRespEst = prop_stim - prop_uns
  )
}

.acs_stim_pair_tbl <- function(batch_list, sample_metadata) {
  purrr::imap_dfr(batch_list, function(ind_batch, batch_nm) {
    ind_uns <- ind_batch[[1]]
    ind_stim <- ind_batch[-1]

    tibble::tibble(
      batch = batch_nm,
      indUns = ind_uns,
      indStim = ind_stim,
      sampleUns = sample_metadata$sample_name[ind_uns],
      sampleStim = sample_metadata$sample_name[ind_stim],
      conditionStim = sample_metadata$condition[ind_stim]
    )
  })
}
```

### Aim

Validate StimGate on the ACS CyTOF dataset and compare its estimated
background-subtracted response frequencies with Tailgate and F-beta.

Unlike the simulation analyses, this analysis starts with a non-parallel
pre-processing stage. The raw FCS files are read from `raw-data`, the input scale
and transformation are made explicit, and the resulting `GatingSet` is cached.
The three methods are then run sequentially on exactly the same processed data.

The primary StimGate analysis explicitly uses `gateCombn = "min"`. Within each
batch this should choose the lowest genuinely generated local-FDR threshold from
the stimulated conditions, rather than allowing a fallback cutpoint to determine
the minimum.

## Pre-processing

### Data, transformation and method settings

```{r}
#| label: preprocessing-settings
# TODO #346: confirm the exact raw-data location and metadata filenames.
acs_raw_subdir <- "acs-cytof"
acs_raw_dir <- projr::projr_path_get("raw-data", acs_raw_subdir)
acs_sample_metadata_path <- file.path(acs_raw_dir, "sample_metadata.csv")
acs_marker_metadata_path <- file.path(acs_raw_dir, "marker_metadata.csv")
acs_reference_counts_path <- file.path(acs_raw_dir, "reference_counts.csv")

acs_cache_dir <- projr::projr_path_get("cache", "real", "acs-cytof")
acs_gs_path <- file.path(acs_cache_dir, "gatingset")
acs_sample_metadata_cache <- file.path(acs_cache_dir, "sample_metadata.csv")
acs_marker_metadata_cache <- file.path(acs_cache_dir, "marker_metadata.csv")
acs_batch_list_path <- file.path(acs_cache_dir, "batch_list.rds")

acs_output_dir <- projr::projr_path_get("cache-analysis", "9-acs-cytof")
acs_table_dir <- file.path(acs_output_dir, "tables")
acs_figure_dir <- file.path(acs_output_dir, "figures")

dir.create(acs_cache_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(acs_table_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(acs_figure_dir, recursive = TRUE, showWarnings = FALSE)

# TODO #346: set these only after inspecting the stored ACS values.
# "none" means that the FCS values are already on the analysis scale.
# "asinh" applies asinh(x / cofactor) to the listed marker channels.
acs_transform_name <- NA_character_
acs_transform_cofactor <- NA_real_

acs_unstim_label <- "unstim"
acs_pop_gate <- "root"

# StimGate settings. gateCombn is intentionally explicit.
stimgate_gate_combn <- "min"
stimgate_bw_mtd <- "hpi1"
stimgate_bw_ncell_max <- 1e4
stimgate_bw_fallback <- "auto"
stimgate_bw_min <- "none"
stimgate_bw_max <- "none"
stimgate_bias_uns <- NULL
stimgate_min_cell <- 100
tol_clust <- 1e-7
calc_cyt_pos_gates <- FALSE

# Comparator settings, matching the corrected analyses 7 and 8 semantics.
fbeta_beta <- 0.8
fbeta_theta <- 2
fbeta_width <- 10
path_fbeta <- projr::projr_path_get(
  "project",
  "scripts",
  "python",
  "fbeta.py"
)

tailgate_adjust <- 1
tailgate_method <- "firstDeriv"
tailgate_tol <- 1e-2
tailgate_auto_tol <- TRUE
```

### Read the FCS files and inspect their stored scale

```{r}
#| label: preprocessing-fcs
#| results: hide
if (isTRUE(run_preprocessing)) {
  .acs_assert_dir(acs_raw_dir, "ACS raw-data directory")
  .acs_assert_file(acs_sample_metadata_path, "ACS sample metadata")
  .acs_assert_file(acs_marker_metadata_path, "ACS marker metadata")

  sample_metadata <- utils::read.csv(
    acs_sample_metadata_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )
  marker_metadata <- utils::read.csv(
    acs_marker_metadata_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  required_sample_cols <- c("fcs_file", "sample_name", "batch", "condition")
  missing_sample_cols <- setdiff(required_sample_cols, names(sample_metadata))
  if (length(missing_sample_cols) > 0L) {
    stop(
      "ACS sample metadata is missing columns: ",
      paste(missing_sample_cols, collapse = ", ")
    )
  }

  required_marker_cols <- c("channel", "marker")
  missing_marker_cols <- setdiff(required_marker_cols, names(marker_metadata))
  if (length(missing_marker_cols) > 0L) {
    stop(
      "ACS marker metadata is missing columns: ",
      paste(missing_marker_cols, collapse = ", ")
    )
  }

  fcs_paths_all <- list.files(
    acs_raw_dir,
    pattern = "\\.fcs$",
    recursive = TRUE,
    full.names = TRUE,
    ignore.case = TRUE
  )
  fcs_lookup <- stats::setNames(fcs_paths_all, basename(fcs_paths_all))

  missing_fcs <- setdiff(sample_metadata$fcs_file, names(fcs_lookup))
  if (length(missing_fcs) > 0L) {
    stop(
      "Could not find ACS FCS files listed in sample_metadata.csv: ",
      paste(missing_fcs, collapse = ", ")
    )
  }

  fcs_paths <- unname(fcs_lookup[sample_metadata$fcs_file])
  ff_list <- lapply(
    fcs_paths,
    flowCore::read.FCS,
    transformation = FALSE,
    truncate_max_range = FALSE
  )

  sample_metadata$n_cell <- vapply(ff_list, nrow, integer(1))

  scale_summary <- purrr::map2_dfr(
    ff_list,
    sample_metadata$sample_name,
    function(ff, sample_name) {
      ex <- flowCore::exprs(ff)
      purrr::map_dfr(marker_metadata$channel, function(channel) {
        if (!channel %in% colnames(ex)) {
          return(tibble::tibble(
            sample_name = sample_name,
            channel = channel,
            min = NA_real_,
            q50 = NA_real_,
            q95 = NA_real_,
            max = NA_real_
          ))
        }
        x <- as.numeric(ex[, channel])
        tibble::tibble(
          sample_name = sample_name,
          channel = channel,
          min = min(x, na.rm = TRUE),
          q50 = stats::median(x, na.rm = TRUE),
          q95 = stats::quantile(x, 0.95, na.rm = TRUE),
          max = max(x, na.rm = TRUE)
        )
      })
    }
  )
  utils::write.csv(
    scale_summary,
    file.path(acs_table_dir, "stored_fcs_scale_summary.csv"),
    row.names = FALSE
  )

  if (is.na(acs_transform_name)) {
    stop(
      "ACS transformation has not yet been verified. Inspect ",
      "stored_fcs_scale_summary.csv, then set acs_transform_name to 'none' ",
      "or 'asinh'. No analysis-scale GatingSet has been cached."
    )
  }

  if (identical(acs_transform_name, "asinh")) {
    if (!is.finite(acs_transform_cofactor) || acs_transform_cofactor <= 0) {
      stop("A positive acs_transform_cofactor is required for the asinh transform.")
    }

    transform_channels <- marker_metadata$channel
    ff_list <- lapply(ff_list, function(ff) {
      ex <- flowCore::exprs(ff)
      missing_channels <- setdiff(transform_channels, colnames(ex))
      if (length(missing_channels) > 0L) {
        stop(
          "Marker channels missing from an ACS FCS file: ",
          paste(missing_channels, collapse = ", ")
        )
      }
      ex[, transform_channels] <- asinh(
        ex[, transform_channels, drop = FALSE] / acs_transform_cofactor
      )
      flowCore::exprs(ff) <- ex
      ff
    })
  } else if (!identical(acs_transform_name, "none")) {
    stop(
      "Unsupported acs_transform_name: ", acs_transform_name,
      ". Add the verified transformation explicitly rather than guessing."
    )
  }

  fs <- flowCore::flowSet(ff_list)
  flowCore::sampleNames(fs) <- sample_metadata$sample_name
  gs <- flowWorkspace::GatingSet(fs)

  if (dir.exists(acs_gs_path)) {
    unlink(acs_gs_path, recursive = TRUE)
  }
  flowWorkspace::save_gs(gs, path = acs_gs_path)

  utils::write.csv(
    sample_metadata,
    acs_sample_metadata_cache,
    row.names = FALSE
  )
  utils::write.csv(
    marker_metadata,
    acs_marker_metadata_cache,
    row.names = FALSE
  )

  batch_list <- stimgate::getBatchList(
    fnTblInfo = sample_metadata,
    colGrp = "batch",
    colStim = "condition",
    unsChr = acs_unstim_label,
    colNCell = "n_cell",
    minCell = stimgate_min_cell
  )
  if (length(batch_list) == 0L) {
    stop("No valid ACS batches remain after constructing batch_list.")
  }
  saveRDS(batch_list, acs_batch_list_path)
}
```

### Load the cached pre-processing outputs

```{r}
#| label: preprocessing-load
if (!dir.exists(acs_gs_path)) {
  stop(
    "Cached ACS GatingSet not found. Run pre-processing after the ",
    "transformation has been verified."
  )
}

.acs_assert_file(acs_sample_metadata_cache, "Cached ACS sample metadata")
.acs_assert_file(acs_marker_metadata_cache, "Cached ACS marker metadata")
.acs_assert_file(acs_batch_list_path, "Cached ACS batch list")

gs_acs <- flowWorkspace::load_gs(acs_gs_path)
sample_metadata <- utils::read.csv(
  acs_sample_metadata_cache,
  stringsAsFactors = FALSE,
  check.names = FALSE
)
marker_metadata <- utils::read.csv(
  acs_marker_metadata_cache,
  stringsAsFactors = FALSE,
  check.names = FALSE
)
batch_list <- readRDS(acs_batch_list_path)
stim_pair_tbl <- .acs_stim_pair_tbl(batch_list, sample_metadata)
method_grid <- merge(
  stim_pair_tbl,
  marker_metadata[, c("channel", "marker"), drop = FALSE],
  by = NULL
)

stim_pair_tbl
marker_metadata
```

## StimGate

```{r}
#| label: run-stimgate
#| results: hide
path_stimgate <- file.path(acs_cache_dir, "stimgate")
path_stimgate_results <- file.path(acs_cache_dir, "stimgate_results.rds")
path_stimgate_gates <- file.path(acs_cache_dir, "stimgate_gates.rds")
path_stimgate_gates_detailed <- file.path(
  acs_cache_dir,
  "stimgate_gates_detailed.rds"
)

if (isTRUE(run_methods)) {
  chnl_settings <- stats::setNames(
    lapply(marker_metadata$channel, function(channel) list(cut = channel)),
    marker_metadata$marker
  )

  if (dir.exists(path_stimgate)) {
    unlink(path_stimgate, recursive = TRUE)
  }

  invisible(stimgate::gateStim(
    pathProject = path_stimgate,
    .data = gs_acs,
    popGate = acs_pop_gate,
    batchList = batch_list,
    chnl = chnl_settings,
    biasUns = stimgate_bias_uns,
    bwMtd = stimgate_bw_mtd,
    bwNcellMax = stimgate_bw_ncell_max,
    bwFallback = stimgate_bw_fallback,
    bwMin = stimgate_bw_min,
    bwMax = stimgate_bw_max,
    minCell = stimgate_min_cell,
    gateCombn = stimgate_gate_combn,
    tolClust = tol_clust,
    calcCytPosGates = calc_cyt_pos_gates
  ))

  stimgate_gates <- stimgate::getStimGates(path_stimgate)
  stimgate_gates_detailed <- stimgate::getStimGatesDetailed(path_stimgate)
  saveRDS(stimgate_gates, path_stimgate_gates)
  saveRDS(stimgate_gates_detailed, path_stimgate_gates_detailed)

  stimgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      gate_rows <- stimgate_gates |>
        dplyr::filter(
          as.character(.data$ind) == as.character(indStim),
          .data$chnl == channel,
          .data$gateCombn %in% c("min", "minClust")
        )

      if (nrow(gate_rows) > 0L && any(gate_rows$gateCombn == "minClust")) {
        gate_rows <- gate_rows[gate_rows$gateCombn == "minClust", , drop = FALSE]
      }
      if (nrow(gate_rows) > 0L) {
        gate_rows <- gate_rows[nrow(gate_rows), , drop = FALSE]
      }

      threshold <- if (nrow(gate_rows) > 0L) {
        suppressWarnings(as.numeric(gate_rows$gate[[1]]))
      } else {
        NA_real_
      }
      gate_name <- if (nrow(gate_rows) > 0L) {
        as.character(gate_rows$gateName[[1]])
      } else {
        NA_character_
      }

      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      .acs_estimate_from_threshold(x_stim, x_uns, threshold) |>
        dplyr::mutate(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "stimgate",
          thresholdOrigin = gate_name,
          thresholdFallbackUsed = !is.finite(threshold),
          .before = 1
        )
    }
  )
  saveRDS(stimgate_results, path_stimgate_results)
} else {
  stimgate_results <- readRDS(path_stimgate_results)
  stimgate_gates <- readRDS(path_stimgate_gates)
  stimgate_gates_detailed <- readRDS(path_stimgate_gates_detailed)
}
```

### Verify the `min` gate combination

```{r}
#| label: check-stimgate-min
if (!identical(stimgate_gate_combn, "min")) {
  stop("ACS StimGate comparison must use gateCombn = 'min'.")
}

condition_generated <- stimgate_gates_detailed |>
  dplyr::filter(
    .data$detailLevel == "condition",
    .data$locGenerated %in% TRUE,
    is.finite(.data$threshold)
  ) |>
  dplyr::mutate(indStim = as.integer(.data$ind)) |>
  dplyr::inner_join(
    stim_pair_tbl |>
      dplyr::select(.data$batch, .data$indStim),
    by = "indStim"
  ) |>
  dplyr::group_by(.data$batch, .data$chnl) |>
  dplyr::summarise(
    expected_min = min(.data$threshold),
    n_generated = dplyr::n(),
    .groups = "drop"
  )

actual_min <- stimgate_gates |>
  dplyr::filter(.data$gateType == "loc", .data$gateCombn == "min") |>
  dplyr::mutate(indStim = as.integer(.data$ind)) |>
  dplyr::inner_join(
    stim_pair_tbl |>
      dplyr::select(.data$batch, .data$indStim),
    by = c("batch", "indStim")
  ) |>
  dplyr::group_by(.data$batch, .data$chnl) |>
  dplyr::summarise(
    actual_min = dplyr::first(.data$gate),
    n_distinct_gates = dplyr::n_distinct(.data$gate),
    .groups = "drop"
  )

min_gate_check_tbl <- condition_generated |>
  dplyr::left_join(actual_min, by = c("batch", "chnl")) |>
  dplyr::mutate(
    min_matches = is.finite(.data$actual_min) &
      abs(.data$expected_min - .data$actual_min) < 1e-10,
    one_gate_within_batch = .data$n_distinct_gates == 1L
  )

if (
  nrow(min_gate_check_tbl) > 0L &&
    any(
      !min_gate_check_tbl$min_matches |
        !min_gate_check_tbl$one_gate_within_batch,
      na.rm = TRUE
    )
) {
  stop("StimGate gateCombn = 'min' did not match the generated-condition minimum.")
}

if (!is.null(tol_clust)) {
  n_min_clust <- stimgate_gates |>
    dplyr::filter(.data$gateCombn == "minClust") |>
    nrow()
  if (n_min_clust == 0L) {
    stop("Cluster refinement was requested but no minClust gates were saved.")
  }
}

saveRDS(
  min_gate_check_tbl,
  file.path(acs_table_dir, "stimgate_min_gate_check.rds")
)
min_gate_check_tbl
```

## Tailgate

```{r}
#| label: run-tailgate
#| results: hide
path_tailgate_results <- file.path(acs_cache_dir, "tailgate_results.rds")

if (isTRUE(run_methods)) {
  tailgate_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareTailgateThreshold(
          x = x_stim,
          adjust = tailgate_adjust,
          method = tailgate_method,
          tol = tailgate_tol,
          autoTol = tailgate_auto_tol,
          strict = FALSE
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "tailgate",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(tailgate_results, path_tailgate_results)
} else {
  tailgate_results <- readRDS(path_tailgate_results)
}
```

## F-beta

```{r}
#| label: run-fbeta
#| results: hide
path_fbeta_results <- file.path(acs_cache_dir, "fbeta_results.rds")

if (isTRUE(run_methods)) {
  fbeta_results <- purrr::pmap_dfr(
    method_grid,
    function(
        batch,
        indUns,
        indStim,
        sampleUns,
        sampleStim,
        conditionStim,
        channel,
        marker) {
      x_stim <- .acs_get_expr(gs_acs, indStim, channel, acs_pop_gate)
      x_uns <- .acs_get_expr(gs_acs, indUns, channel, acs_pop_gate)

      threshold_obj <- tryCatch(
        .simCompareFbetaThreshold(
          xUns = x_uns,
          xStim = x_stim,
          pathFbeta = path_fbeta,
          beta = fbeta_beta,
          theta = fbeta_theta,
          width = fbeta_width
        ),
        error = function(e) {
          list(
            threshold = NA_real_,
            thresholdMetric = NA_real_,
            thresholdOrigin = paste0("error: ", e$message)
          )
        }
      )

      est <- .simCompareEstimateFromThreshold(
        xStim = x_stim,
        xUns = x_uns,
        threshold = threshold_obj$threshold,
        fallbackHighValue = TRUE
      )

      tibble::as_tibble(est) |>
        dplyr::transmute(
          batch = batch,
          indUns = indUns,
          indStim = indStim,
          sampleUns = sampleUns,
          sampleStim = sampleStim,
          conditionStim = conditionStim,
          channel = channel,
          marker = marker,
          method = "fbeta",
          threshold = .data$threshold,
          thresholdOrigin = threshold_obj$thresholdOrigin,
          thresholdMetric = threshold_obj$thresholdMetric,
          thresholdFallbackUsed = .data$thresholdFallbackUsed,
          nCellStim = .data$nCellStim,
          nCellUns = .data$nCellUns,
          nPosStim = .data$nPosStim,
          nPosUns = .data$nPosUns,
          propStim = .data$propStim,
          propUns = .data$propUns,
          propRespEst = .data$propRespEst
        )
    }
  )
  saveRDS(fbeta_results, path_fbeta_results)
} else {
  fbeta_results <- readRDS(path_fbeta_results)
}
```

## Compare methods with the ACS reference counts

```{r}
#| label: compare-methods
method_results <- dplyr::bind_rows(
  stimgate_results,
  tailgate_results,
  fbeta_results
)

if (file.exists(acs_reference_counts_path)) {
  reference_counts <- utils::read.csv(
    acs_reference_counts_path,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  required_reference_cols <- c("sample_name", "channel")
  missing_reference_cols <- setdiff(
    required_reference_cols,
    names(reference_counts)
  )
  if (length(missing_reference_cols) > 0L) {
    stop(
      "ACS reference table is missing columns: ",
      paste(missing_reference_cols, collapse = ", ")
    )
  }

  if (!"reference_prop" %in% names(reference_counts)) {
    if (all(c("reference_count", "reference_n") %in% names(reference_counts))) {
      reference_counts <- reference_counts |>
        dplyr::mutate(
          reference_prop = .data$reference_count / .data$reference_n
        )
    } else {
      stop(
        "ACS reference table must contain reference_prop, or both ",
        "reference_count and reference_n."
      )
    }
  }

  comparison_tbl <- method_results |>
    dplyr::left_join(
      reference_counts,
      by = c("sampleStim" = "sample_name", "channel")
    ) |>
    dplyr::mutate(
      error = .data$propRespEst - .data$reference_prop,
      abs_error = abs(.data$error),
      rel_error = dplyr::if_else(
        .data$reference_prop == 0,
        NA_real_,
        .data$error / .data$reference_prop
      )
    )
} else {
  warning(
    "ACS reference counts have not yet been located. Method outputs are ",
    "available, but reference-based comparison is deferred."
  )
  comparison_tbl <- method_results
}

comparison_tbl
```

## Output tables and figures

### Save analysis tables

```{r}
#| label: save-output-tables
saveRDS(method_results, file.path(acs_table_dir, "method_results.rds"))
saveRDS(comparison_tbl, file.path(acs_table_dir, "comparison_with_reference.rds"))

utils::write.csv(
  method_results,
  file.path(acs_table_dir, "method_results.csv"),
  row.names = FALSE
)
utils::write.csv(
  comparison_tbl,
  file.path(acs_table_dir, "comparison_with_reference.csv"),
  row.names = FALSE
)

if (exists("reference_counts")) {
  saveRDS(reference_counts, file.path(acs_table_dir, "reference_counts.rds"))
}
```

### Estimated versus reference response frequency

```{r}
#| label: plot-estimate-v-reference
#| fig-width: 9
#| fig-height: 7
if (
  isTRUE(run_plots) &&
    all(c("reference_prop", "propRespEst") %in% names(comparison_tbl))
) {
  p_est_ref <- ggplot(
    comparison_tbl,
    aes(
      x = reference_prop,
      y = propRespEst,
      colour = method
    )
  ) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    geom_point(alpha = 0.75) +
    facet_wrap(~marker, scales = "free") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "xy", minor = "none") +
    labs(
      title = "ACS CyTOF: Estimated vs Reference Response Frequency",
      x = "Reference response frequency",
      y = "Estimated background-subtracted response frequency",
      colour = "Method"
    )

  ggsave(
    file.path(acs_figure_dir, "estimate_vs_reference.pdf"),
    p_est_ref,
    width = 18,
    height = 14,
    units = "cm"
  )
  print(p_est_ref)
}
```

### Relative error by method

```{r}
#| label: plot-relative-error
#| fig-width: 9
#| fig-height: 7
if (isTRUE(run_plots) && "rel_error" %in% names(comparison_tbl)) {
  p_rel_error <- ggplot(
    comparison_tbl,
    aes(x = method, y = rel_error)
  ) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.15, alpha = 0.5) +
    facet_wrap(~marker, scales = "free_y") +
    cowplot::theme_cowplot() +
    cowplot::background_grid(major = "y", minor = "none") +
    labs(
      title = "ACS CyTOF: Relative Error by Method",
      x = NULL,
      y = "Relative error"
    )

  ggsave(
    file.path(acs_figure_dir, "relative_error_by_method.pdf"),
    p_rel_error,
    width = 18,
    height = 14,
    units = "cm"
  )
  print(p_rel_error)
}
```

The figures above are initial comparison outputs. Once the ACS reference counts
and their exact denominator are recovered, this final section should be reduced
to the thesis figures and tables that are actually needed.
