---
title: Compare StimGate, Tailgate and F-beta on the ACS CyTOF dataset
format:
  html:
    embed-resources: true
    toc: true
params:
  run_preprocessing: true
  run_methods: true
  run_plots: false
  tester_n_sample: 20
  n_workers: 2
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)

root_dir <- tryCatch(
  projr::projr_path_get("project"),
  error = function(e) getwd()
)

scripts_r_dir <- file.path(root_dir, "scripts", "r")
source(file.path(scripts_r_dir, "analysis-runtime.R"))
source(file.path(scripts_r_dir, "acs_cytof-helper.R"))
source(file.path(scripts_r_dir, "acs_cytof-preprocess.R"))
source(file.path(scripts_r_dir, "acs_cytof-gate.R"))
source(file.path(scripts_r_dir, "sim-misc.R"))

run_preprocessing <- .as_flag(.get_qmd_param_env(
  "run_preprocessing",
  "RUN_PREPROCESSING",
  TRUE
))
pop_vec <- c("tcrgd", "cd4", "cd8", "nk", "pre_nk", "b")
run_preprocessing_vec_by_pop <- c(
  tcrgd = NULL %||% run_preprocessing,
  cd4 = NULL %||% run_preprocessing,
  cd8 = NULL %||% run_preprocessing,
  nk = NULL %||% run_preprocessing,
  pre_nk = NULL %||% run_preprocessing,
  b = NULL %||% run_preprocessing
)
run_methods <- .as_flag(.get_qmd_param_env(
  "run_methods",
  "RUN_METHODS",
  TRUE
))
run_methods_vec_by_pop <- c(
  tcrgd = NULL %||% run_methods,
  cd4 = NULL %||% run_methods,
  cd8 = NULL %||% run_methods,
  nk = NULL %||% run_methods,
  pre_nk = NULL %||% run_methods,
  b = NULL %||% run_methods
)
run_plots <- .as_flag(.get_qmd_param_env(
  "run_plots",
  "RUN_PLOTS",
  TRUE
))
run_plots_vec_by_pop <- c(
  tcrgd = NULL %||% run_plots,
  cd4 = NULL %||% run_plots,
  cd8 = NULL %||% run_plots,
  nk = NULL %||% run_plots,
  pre_nk = NULL %||% run_plots,
  b = NULL %||% run_plots
)
tester_n_sample <- .get_qmd_param_env(
  "tester_n_sample",
  "TESTER_N_SAMPLE",
  20L
)
n_workers <- max(1, .simGetCores())
bias_uns <- 0.15
bias_uns_vec_by_pop <- c(
  tcrgd = NULL %||% bias_uns,
  cd4 = NULL %||% bias_uns,
  cd8 = NULL %||% bias_uns,
  nk = NULL %||% bias_uns,
  pre_nk = NULL %||% bias_uns,
  b = NULL %||% bias_uns
)

desc_path <- file.path(root_dir, "DESCRIPTION")
if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

### Aim

Validate StimGate on the ACS CyTOF dataset and compare its estimated
background-subtracted response frequencies with Tailgate and F-beta.

## Gating

The TCRgd population is first run on a limited set of complete five-sample
batches. If that tester completes successfully, the same preprocessing and
StimGate flow is applied to every population configured in `pop_vec`.

The reusable helper fixes the settings that are shared by these runs, including
the six cytokine channels, `bwMtd = "hpi1"`, debug profiling and the required
`gateCombn = "min"`. The population, stage controls, tester sample limit and
`biasUns` remain visible here because they are the settings most likely to vary.

### Common paths

```{r}
#| label: settings-common-gating-paths
path_fcs_base <- projr::projr_path_get(
  "raw-data",
  "comparison_data", "acscytof", "fcs"
)
if (!dir.exists(path_fcs_base)) {
  stop("Path to FCS files does not exist: ", path_fcs_base)
}
path_gs_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "gs"
)
path_scratch_base <- projr::projr_path_get(
  "cache",
  "acs_cytof",
  "scratch"
)
```

### TCRgd tester

The tester uses only the first `tester_n_sample` files, which must comprise
complete batches of one unstimulated and four stimulated samples. Its cached
GatingSet and StimGate project are kept under a separate `tester/tcrgd`
subdirectory, so they cannot overwrite the full TCRgd run.

```{r}
#| label: run-tcrgd-tester
#| results: hide
tcrgd_tester <- .acsCytofRunPopulation(
  pop = "tcrgd",
  pathFcsBase = path_fcs_base,
  pathGsBase = path_gs_base,
  pathScratchBase = path_scratch_base,
  runPreprocessing = run_preprocessing,
  runMethods = run_methods,
  runPlots = run_plots,
  nSample = tester_n_sample,
  biasUns = bias_uns_vec_by_pop[["tcrgd"]],
  outputGroup = "tester"
)
```

### Across populations, in parallel

Each population writes to its own cached GatingSet and scratch directory. The
number of concurrent population runs defaults to two and can be changed with
the `n_workers` parameter or `ACS_N_WORKERS` environment variable.

```{r}
#| label: run-all-populations
#| results: hide
n_workers_use <- min(n_workers, length(pop_vec))
old_plan <- future::plan()

Sys.setenv("STIMGATE_DEBUG" = "TRUE")

population_runs <- tryCatch(
  {
    if (n_workers_use > 1L) {
      future::plan(future::multisession, workers = n_workers_use)
    } else {
      future::plan(future::sequential)
    }

    furrr::future_map(
      pop_vec,
      function(pop) {
        .acsCytofRunPopulation(
          pop = pop,
          pathFcsBase = path_fcs_base,
          pathGsBase = path_gs_base,
          pathScratchBase = path_scratch_base,
          runPreprocessing = run_preprocessing_vec_by_pop[[pop]],
          runMethods = run_methods_vec_by_pop[[pop]],
          runPlots = run_plots_vec_by_pop[[pop]],
          biasUns = bias_uns_vec_by_pop[[pop]]
        )
      },
      .options = furrr::furrr_options(seed = TRUE)
    )
  },
  finally = future::plan(old_plan)
)
names(population_runs) <- pop_vec
```

## Tailgate and F-beta

The reusable run above covers the currently working ACS preprocessing and
StimGate path. Tailgate and F-beta will be added to the same population helper
once their ACS method grid and result-cache contract have been completed.

## Compare methods with the ACS reference counts

## Output tables and figures

### Save analysis tables

### Estimated versus reference response frequency

### Relative error by method
