---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| include: false
library(ggplot2)
library(tibble)

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

## Checks

### Transformations

```{r}
#| eval: false
calc_gamma <- function(x) {
  gamma(1 + abs(x / 4))
}
calc_gaussian <- function(x) {
  x
}
calc_skew <- function(x, epsilon = 2, delta = 1) {
  cluster_mean <- mean(x)
  cluster_sd <- sd(x)
  weight <- 1 / (1 + exp(1.5 * (cluster_mean - 3.5)))
  epsilon <- epsilon * weight
  out <- sinh(epsilon + delta * asinh(x))
  gamma_divisor <- rgamma(length(x), shape = 5, rate = 5)
  out_fat <- out / sqrt(gamma_divisor)
  (out - mean(out)) / sd(out) * sd(x) + cluster_mean
}
mean_vec <- c(0, 8)
plot_tbl <- purrr::map_df(mean_vec, function(mean_curr) {
  sim_vec <- rnorm(1e3, mean = mean_curr, sd = sqrt(2))
  tibble::tibble(
    trans = rep(c("gamma", "gaussian", "skew"), each = length(sim_vec)),
    sim = c(calc_gamma(sim_vec), calc_gaussian(sim_vec), calc_skew(sim_vec))
  )
}) |>
  dplyr::filter(!trans == "gamma")
ggplot(plot_tbl, aes(x = sim, col = trans)) +
  geom_density(alpha = 0.7) +
  cowplot::theme_cowplot()
```

### Simulations

```{r}
#| eval: false
# debugonce(simCytCondition)
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    "skew" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  out_list_condition <- simCytCondition(
    nMarker = 2,
    nCell = 1e4,
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probVec = c(0.05, 0.02, 0.075, 0.855)
  )
  plot_tbl <- tibble::tibble(
    pop = out_list_condition$conditionLabels
  ) |>
    dplyr::bind_cols(
      out_list_condition$conditionMatrix
    )
  pop_vec <- unique(plot_tbl$pop)
  p_list <- lapply(pop_vec, function(x) {
    UtilsCytoRSV::plot_cyto(
      plot_tbl |>
        dplyr::filter(pop == x),
      marker = c("F1", "F2"),
      limits_expand = limits_expand
    ) +
      labs(title = x) +
      geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
      geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
  })
  plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA)
    )
  path_save <- projr::projr_path_get(
    "cache",
    "sim-bandwidth",
    "check",
    "condition",
    paste0("p-condition-", trans, ".png")
  )
  ggplot2::ggsave(
    path_save,
    units = "cm",
    width = 16,
    height = 16
  )
}
```

```{r}
#| eval: false
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  # debugonce(simCytSample)
  out_list_sample <- simCytSample(
    nMarker = 2L,
    nCondition = 3L,
    nCluster = 4L,
    nCellByCondition = c(1e4, 1e4, 1e4),
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probResponseVecByStimCondition = list(
      c(0.1, 0.05, 0.0025, 0.8425),
      c(0.00001, 0.0001, 0.0001, 0.99978)
    ),
    probVecUns = c(0.05, 0.02, 0.075, 0.855),
    conditionPerturbationSd = 0,
    clusterPerturbationSd = 0
  )
  for (i in seq_along(out_list_sample$flowFrameList)) {
    stim <- names(out_list_sample$flowFrameList)[[i]]
    plot_tbl <- tibble::tibble(
      pop = out_list_sample$conditionLabelsList[[i]]
    ) |>
      dplyr::bind_cols(
        as.data.frame(flowCore::exprs(out_list_sample$flowFrameList[[i]]))
      )
    pop_vec <- unique(plot_tbl$pop)
    p_list <- lapply(pop_vec, function(x) {
      UtilsCytoRSV::plot_cyto(
        plot_tbl |>
          dplyr::filter(pop == x),
        marker = c("F1", "F2"),
        limits_expand = limits_expand
      ) +
        labs(title = x) +
        geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
        geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
    })
    plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
      )
    path_save <- projr::projr_path_get(
      "cache",
      "sim-bandwidth",
      "check",
      "sample",
      paste0("p-sample-", trans, "-", stim, ".png")
    )
    ggplot2::ggsave(
      path_save,
      units = "cm",
      width = 16,
      height = 16
    )
  }
}
```

```{r}
#| eval: false
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  out_list_experiment <- simCytExperiment(
    nSample = 2L,
    nMarker = 2L,
    nCondition = 3L,
    nCluster = 4L,
    nCellByCondition = c(1e4, 1e4, 1e4),
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probResponseVecByStimCondition = list(
      c(0.1, 0.05, 0.0025, 0.8425),
      c(0.00001, 0.0001, 0.0001, 0.99978)
    ),
    probVecUns = c(0.05, 0.02, 0.075, 0.855),
    samplePerturbationSd = 0.1,
    conditionPerturbationSd = 0.1,
    clusterPerturbationSd = 0.1,
    covEvMin = 2,
    covEvMax = 2
  )
  for (i in seq_along(out_list_experiment$flowFrameList)) {
    smpl_cond <- names(out_list_experiment$flowFrameList)[[i]]
    plot_tbl <- tibble::tibble(
      pop = out_list_experiment$conditionLabelsList[[i]]
    ) |>
      dplyr::bind_cols(
        as.data.frame(flowCore::exprs(out_list_experiment$flowFrameList[[i]]))
      )
    pop_vec <- unique(plot_tbl$pop)
    p_list <- lapply(pop_vec, function(x) {
      UtilsCytoRSV::plot_cyto(
        plot_tbl |>
          dplyr::filter(pop == x),
        marker = c("F1", "F2"),
        limits_expand = limits_expand
      ) +
        labs(title = x) +
        geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
        geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
    })
    plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
      )
    path_save <- projr::projr_path_get(
      "cache",
      "sim-bandwidth",
      "check",
      "experiment",
      paste0("p-experiment-", trans, "-", smpl_cond, ".png")
    )
    ggplot2::ggsave(
      path_save,
      units = "cm",
      width = 16,
      height = 16
    )
  }
}
```

### Analysis

#### Run gateStim

```{r}
#| eval: false
# loop parameters
n_cell_stim <- 1e5
prob_response <- 0.02
mean_pos <- 8
transformation <- "gaussian"
sample_perturbation_sd <- 0
condition_perturbation_sd <- 0
cluster_perturbation_sd <- 0
background_relative_to_response <- 0.2
ncell_uns_relative_to_stim <- 1

# stimgate parameters
bias_uns <- 0.2

# calculated parameters
n_cell_uns <- round(n_cell_stim * ncell_uns_relative_to_stim)
n_cell_by_condition <- c(n_cell_uns, n_cell_stim)
transformation_func <- switch(
  transformation,
  "gamma" = calc_gamma,
  "gaussian" = calc_gaussian,
  "skew" = calc_skew,
  stop("transformation not recognised")
)
mean_expr_mat <- matrix(
  c(0, 8),
  byrow = TRUE,
  ncol = 1
)
cluster_label_vec <- c("gn", "gp")
prob_response_uns <- prob_response * background_relative_to_response
prob_vec_uns <- c(1 - prob_response_uns, prob_response_uns)
prob_response_vec_by_stim_condition <- list(c(-prob_response, prob_response))
iter_num <- 1

n_sample <- 5L
n_marker <- 1L
n_condition <- 2L
n_cluster <- 2L

print(paste0("iter ", iter_num))
# debugonce(simCytCondition)
out_list_experiment <- simCytExperiment(
  nSample = n_sample,
  nMarker = n_marker,
  nCondition = n_condition,
  nCluster = n_cluster,
  nCellByCondition = n_cell_by_condition,
  transformationFunc = transformation_func,
  mixtureType = "gaussianOnly",
  meanExprMat = mean_expr_mat,
  clusterLabelVec = cluster_label_vec,
  probVecUns = prob_vec_uns,
  probResponseVecByStimCondition = prob_response_vec_by_stim_condition,
  conditionPerturbationSd = condition_perturbation_sd,
  clusterPerturbationSd = cluster_perturbation_sd,
  samplePerturbationSd = sample_perturbation_sd,
  covEvMin = 2,
  covEvMax = 2
)

# apply stimgate
fs <- as(out_list_experiment[["flowFrameList"]], "flowSet")
gs <- flowWorkspace::GatingSet(fs)

path_project <- projr::projr_path_get(
  "cache",
  "sim-bandwidth",
  "check",
  "gate",
  paste0("iter_", iter_num)
)
unlink(path_project, recursive = TRUE)
dir.create(path_project, recursive = TRUE)
batch_list <- lapply(seq(n_sample), function(i) {
  seq((i - 1) * n_condition + 1, i * n_condition)
})

# debugonce(stimgate:::.complete_chnl_list)
gateStim(
  pathProject = path_project,
  chnl = "F1",
  .data = gs,
  batchList = batch_list,
  chnlSettings = list(),
  biasUns = bias_uns
)

stats_tbl <- readRDS(
  file.path(path_project, "gateStats.rds")
)
stats_tbl
```

#### Comparing actual to estimated frequencies

This is for when we're varying the bandwidth manually.

```{r}
#| eval: false
# loop parameters
n_cell_stim <- 1e3
prob_response <- 0.2
mean_pos <- 8
transformation <- "gaussian"
sample_perturbation_sd <- 0
condition_perturbation_sd <- 0
cluster_perturbation_sd <- 0
background_relative_to_response <- 0.2
ncell_uns_relative_to_stim <- 1

# stimgate parameters
bias_uns <- 0.2

# calculated parameters
n_cell_uns <- round(n_cell_stim * ncell_uns_relative_to_stim)
n_cell_by_condition <- c(n_cell_uns, n_cell_stim)
transformation_func <- .simMiscGetTrans(transformation)
mean_expr_mat <- matrix(
  c(0, 8),
  byrow = TRUE,
  ncol = 1
)
cluster_label_vec <- c("gn", "gp")
prob_response_uns <- prob_response * background_relative_to_response
prob_vec_uns <- c(1 - prob_response_uns, prob_response_uns)
prob_response_vec_by_stim_condition <- list(c(-prob_response, prob_response))
iter_num <- 1

n_sample <- 1L
n_marker <- 1L
n_condition <- 2L
n_cluster <- 2L

print(paste0("iter ", iter_num))
# debugonce(simCytCondition)
out_list_experiment <- simCytExperiment(
  nSample = n_sample,
  nMarker = n_marker,
  nCondition = n_condition,
  nCluster = n_cluster,
  nCellByCondition = n_cell_by_condition,
  transformationFunc = transformation_func,
  mixtureType = "gaussianOnly",
  meanExprMat = mean_expr_mat,
  clusterLabelVec = cluster_label_vec,
  probVecUns = prob_vec_uns,
  probResponseVecByStimCondition = prob_response_vec_by_stim_condition,
  conditionPerturbationSd = condition_perturbation_sd,
  clusterPerturbationSd = cluster_perturbation_sd,
  samplePerturbationSd = sample_perturbation_sd,
  covEvMin = 2,
  covEvMax = 2
)
# 4. Construct flowSet and GatingSet
fs <- as(out_list_experiment[["flowFrameList"]], "flowSet")
gs <- flowWorkspace::GatingSet(fs)
labels_list <- out_list_experiment[["labelsList"]]

path_project <- projr::projr_path_get(
  "cache",
  "sim-bandwidth",
  "check",
  "gate",
  paste0("iter_", iter_num)
)
unlink(path_project, recursive = TRUE)
dir.create(path_project, recursive = TRUE)
batch_list <- lapply(seq(n_sample), function(i) {
  seq((i - 1) * n_condition + 1, i * n_condition)
})

Sys.setenv("STIMGATE_INTERMEDIATE" = "TRUE")
invisible(gateStim(
  .data = gs,
  pathProject = path_project,
  popGate = "root",
  batchList = batch_list,
  marker = paste0("MarkerF", seq_len(n_marker)),
  bw = 11
))

stats_tbl <- readRDS(
  file.path(path_project, "gateStats.rds")
)
stats_tbl
```

```{r}
# truth table
prob_bs_tbl_truth <- purrr::map_df(seq_len(n_sample), function(sample_curr) {
  labels_vec <- (sample_curr - 1) * n_condition + seq(1, n_condition)
  purrr::map_df(labels_vec, function(i) {
    label_vec <- labels_list[[i]]
    stim_condition <- ifelse(i %% n_condition == 1, "unstim", "stim")
    f1p <- sum(grepl("^gp$", label_vec)) / length(label_vec)
    tibble::tibble(
      ind = sample_curr |> as.character(),
      stim = stim_condition,
      chnl = "F1",
      F1 = f1p
    )
  })
}) |>
  tidyr::pivot_wider(
    names_from = stim,
    values_from = F1
  ) |>
  dplyr::mutate(
    prop_resp_truth = stim - unstim
  ) |>
  dplyr::rename(
    sample = ind
  ) |>
  dplyr::select(
    all_of(c("chnl", "sample", "prop_resp_truth"))
  )

path_dir_int_init <- file.path(path_project, "intermediateData", "init")
chnl_vec <- list.dirs(path_dir_int_init, full.names = FALSE, recursive = FALSE)
chnl <- chnl_vec[[1]]
ind <- 2
prob_bs_tbl_est <- purrr::map_df(chnl_vec, function(chnl) {
  ind_vec_stim <- seq(2, n_sample * n_condition, by = 2) |>
    as.character()
  purrr::map_df(ind_vec_stim, function(ind) {
    path_ind <- file.path(path_dir_int_init, chnl, "ind", ind)
    prob_smooth <- file.path(path_ind, "dataMod.rds") |>
      readRDS()
    ncell_resp_smooth <- if (nrow(prob_smooth) > 0L) {
      sum(prob_smooth$probSmooth)
    } else {
      0
    }
    ncell_resp_pred <- if (nrow(prob_smooth) > 0L) {
      sum(prob_smooth$pred)
    } else {
      0
    }
    tibble::tibble(
      sample = as.character(as.numeric(ind) / n_condition),
      ind = ind,
      chnl = chnl,
      ncell_condition = n_cell_stim,
      ncell_resp_smooth = ncell_resp_smooth,
      ncell_resp_pred = ncell_resp_pred,
      prop_resp_smooth = ncell_resp_smooth / ncell_condition,
      prop_resp_pred = ncell_resp_pred / ncell_condition
    )
  })
}) |>
  dplyr::select(
    sample,
    chnl,
    prop_resp_smooth,
    prop_resp_pred
  ) |>
  tidyr::pivot_longer(
    cols = c(prop_resp_smooth, prop_resp_pred),
    names_to = "method",
    values_to = "prop_resp"
  )
comparison_tbl <- prob_bs_tbl_truth |>
  dplyr::left_join(
    prob_bs_tbl_est,
    by = c("sample", "chnl")
  )
comparison_tbl |>
  dplyr::group_by(method) |>
  dplyr::summarise(
    ccc = DescTools::CCC(prop_resp_truth, prop_resp)$rho.c$est,
    pearson = cor(prop_resp_truth, prop_resp, method = "pearson"),
    spearman = cor(prop_resp_truth, prop_resp, method = "spearman")
  )
chnlSettings <- stimgateMetaReadSettingsChnls(path_project)
chnlSettings |>
  lapply(
    function(x) x[c("bwMin", "bwMax", "bwCluster", "biasUns")] |>
      lapply(function(y) signif(y, 2))
  )

# let's plot it
plotlist <- lapply(1:8, function(i) {
  ind_vec <- c((i - 1) * n_condition + 1, i * n_condition)
  plotStim(
    ind = ind_vec,
    .data = gs,
    pathProject = path_project,
    chnl = "BC1(La139)Dd"
  )
})
p_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 2,
  labels = paste0("sample ", seq_len(8))
)
ggsave(
  filename = file.path(dir_cache_check_bw, "p_grid.png"),
  plot = p_grid,
  width = 30,
  height = 30,
  units = "cm"
)
```

## Varying bandwidth manually

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
So, we can vary the bandwidth from a small number (say 0.1) to a large number (say 10), and see how estimated background-subtracted frequencies compare to actual.

It should be doable with just the `bw` parameter, but ah, we haven't set that yet.

Okay, we've got the code working above. Now we'll vary the fixed bandwidth from 0.01 to 4, along with making the bias_unstim some reasonably large value (we'll have to fix it).

And we'll do this for varying numbers of cells. 

```{r}
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)
mean_pos_vec <- seq_len(8)
transformation_vec <- c("gaussian", "gamma", "skew")
sample_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
condition_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
cluster_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
background_relative_to_response <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)
n_cell_uns_relative_to_stim <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)

# bandwidth settings
bandwidth_method_vec <- c("hpi_0", "hpi_1", "hpi_2", "hpi_3", "sj", "nrd0")
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- seq(0, 2, by = 0.2)

# fix the bandwidth
bandwidth_value_vec <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2, 5)
```

```{r}
library(tidyr)
library(dplyr)
library(purrr)
library(furrr)

# 1. Set up the parallel backend (uses all available cores minus one)
plan(multisession, workers = availableCores() - 1)

path_progress_file <- projr::projr_path_get(
  "cache", "sim-bandwidth", "check", "progress.txt"
)

# Clear or initialize the progress file
writeLines("Starting parallel simulations...", path_progress_file)

# 2. Flatten your nested loops into a single combination grid
sim_grid <- expand_grid(
  transformation = transformation_vec,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  mean_pos       = mean_pos_vec,
  bw             = bandwidth_value_vec
)

# 3. Use future_pmap to run the grid rows in parallel
bw_list_raw <- future_pmap(sim_grid, function(transformation, prob_response, n_cell, mean_pos, bw) {
  
  # Safe logging: since multiple workers write at once, construct the line in one go
  log_msg <- paste0(
    "Running: trans=", transformation, 
    " | prob=", prob_response, 
    " | n_cell=", n_cell, 
    " | mean=", mean_pos, 
    " | bw=", bw, "\n"
  )
  cat(log_msg, file = path_progress_file, append = TRUE)
  
  # Run the simulation
  out <- tryCatch({
    .simBandwidth(
      nSample = 10,
      nMarker = 1,
      nCondition = 2,
      nCluster = 2,
      nIter = 10,
      biasUns = 0.05,
      bw = bw,
      nCellStim = n_cell,
      probResponse = prob_response,
      meanPos = mean_pos,
      transformation = transformation,
      samplePerturbationSd = 0,
      conditionPerturbationSd = 0,
      clusterPerturbationSd = 0,
      backgroundRelativeToResponse = 0.2, 
      ncellUnsRelativeToStim = 1,
      covEvMin = 2, 
      covEvMax = 2
    )
  }, error = function(e) {
    # Keep error logging self-contained per worker
    err_msg <- paste0("Error [bw=", bw, ", n_cell=", n_cell, "]: ", conditionMessage(e), "\n")
    cat(err_msg, file = path_progress_file, append = TRUE)
    return(paste0("Error: ", conditionMessage(e)))
  })
  
  return(out)
}, .options = furrr_options(seed = TRUE)) # Crucial if .simBandwidth uses random numbers

# Close the parallel workers when done
plan(sequential)
```
