---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tibble)
library(furrr)
library(progressr) 

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

## Checks

### Transformations

```{r}
#| eval: false
#| label: check-transformations
mean_vec <- c(0, 8)
plot_tbl <- purrr::map_df(mean_vec, function(mean_curr) {
  sim_vec <- rnorm(1e3, mean = mean_curr, sd = sqrt(2))
  tibble::tibble(
    trans = rep(c("gamma", "gaussian", "skew"), each = length(sim_vec)),
    sim = c(calc_gamma(sim_vec), calc_gaussian(sim_vec), calc_skew(sim_vec))
  )
}) |>
  dplyr::filter(!trans == "gamma")
ggplot(plot_tbl, aes(x = sim, col = trans)) +
  geom_density(alpha = 0.7) +
  cowplot::theme_cowplot()
```

### Simulations

```{r}
#| eval: false
#| label: check-simulations
# debugonce(simCytCondition)
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    "skew" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  out_list_condition <- simCytCondition(
    nMarker = 2,
    nCell = 1e4,
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probVec = c(0.05, 0.02, 0.075, 0.855)
  )
  plot_tbl <- tibble::tibble(
    pop = out_list_condition$conditionLabels
  ) |>
    dplyr::bind_cols(
      out_list_condition$conditionMatrix
    )
  pop_vec <- unique(plot_tbl$pop)
  p_list <- lapply(pop_vec, function(x) {
    UtilsCytoRSV::plot_cyto(
      plot_tbl |>
        dplyr::filter(pop == x),
      marker = c("F1", "F2"),
      limits_expand = limits_expand
    ) +
      labs(title = x) +
      geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
      geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
  })
  plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA)
    )
  path_save <- projr::projr_path_get(
    "cache",
    "sim-bw",
    "check",
    "condition",
    paste0("p-condition-", trans, ".png")
  )
  ggplot2::ggsave(
    path_save,
    units = "cm",
    width = 16,
    height = 16
  )
}
```

```{r}
#| eval: false
#| label: check-sample
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  # debugonce(simCytSample)
  out_list_sample <- simCytSample(
    nMarker = 2L,
    nCondition = 3L,
    nCluster = 4L,
    nCellByCondition = c(1e4, 1e4, 1e4),
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probResponseVecByStimCondition = list(
      c(0.1, 0.05, 0.0025, 0.8425),
      c(0.00001, 0.0001, 0.0001, 0.99978)
    ),
    probVecUns = c(0.05, 0.02, 0.075, 0.855),
    conditionPerturbationSd = 0,
    clusterPerturbationSd = 0
  )
  for (i in seq_along(out_list_sample$flowFrameList)) {
    stim <- names(out_list_sample$flowFrameList)[[i]]
    plot_tbl <- tibble::tibble(
      pop = out_list_sample$conditionLabelsList[[i]]
    ) |>
      dplyr::bind_cols(
        as.data.frame(flowCore::exprs(out_list_sample$flowFrameList[[i]]))
      )
    pop_vec <- unique(plot_tbl$pop)
    p_list <- lapply(pop_vec, function(x) {
      UtilsCytoRSV::plot_cyto(
        plot_tbl |>
          dplyr::filter(pop == x),
        marker = c("F1", "F2"),
        limits_expand = limits_expand
      ) +
        labs(title = x) +
        geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
        geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
    })
    plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
      )
    path_save <- projr::projr_path_get(
      "cache",
      "sim-bw",
      "check",
      "sample",
      paste0("p-sample-", trans, "-", stim, ".png")
    )
    ggplot2::ggsave(
      path_save,
      units = "cm",
      width = 16,
      height = 16
    )
  }
}
```

```{r}
#| eval: false
#| label: check-experiment
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  out_list_experiment <- simCytExperiment(
    nSample = 2L,
    nMarker = 2L,
    nCondition = 3L,
    nCluster = 4L,
    nCellByCondition = c(1e4, 1e4, 1e4),
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probResponseVecByStimCondition = list(
      c(0.1, 0.05, 0.0025, 0.8425),
      c(0.00001, 0.0001, 0.0001, 0.99978)
    ),
    probVecUns = c(0.05, 0.02, 0.075, 0.855),
    samplePerturbationSd = 0.1,
    conditionPerturbationSd = 0.1,
    clusterPerturbationSd = 0.1,
    covEvMin = 2,
    covEvMax = 2
  )
  for (i in seq_along(out_list_experiment$flowFrameList)) {
    smpl_cond <- names(out_list_experiment$flowFrameList)[[i]]
    plot_tbl <- tibble::tibble(
      pop = out_list_experiment$conditionLabelsList[[i]]
    ) |>
      dplyr::bind_cols(
        as.data.frame(flowCore::exprs(out_list_experiment$flowFrameList[[i]]))
      )
    pop_vec <- unique(plot_tbl$pop)
    p_list <- lapply(pop_vec, function(x) {
      UtilsCytoRSV::plot_cyto(
        plot_tbl |>
          dplyr::filter(pop == x),
        marker = c("F1", "F2"),
        limits_expand = limits_expand
      ) +
        labs(title = x) +
        geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
        geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
    })
    plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
      )
    path_save <- projr::projr_path_get(
      "cache",
      "sim-bw",
      "check",
      "experiment",
      paste0("p-experiment-", trans, "-", smpl_cond, ".png")
    )
    ggplot2::ggsave(
      path_save,
      units = "cm",
      width = 16,
      height = 16
    )
  }
}
```

### Analysis

#### Bs-freqs

```{r}
#| eval: false
#| label: check-bs-freqs
# test indices
std_check <- FALSE
if (std_check) {
  transformation <- "gaussian"
  prob_response <- 0.05
  n_cell <- 1e4
  mean_pos <- 8
  bw <- 0.3
} else {
  transformation <- "gaussian"
  prob_response <- 0.002
  n_cell <- 5e3
  mean_pos <- 2
  bw <- 0.1
}

sim_res <- .simBandwidthBsFreq(
  nSample = 10,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = 10,
  biasUns = 0.05,
  bw = bw,
  nCellStim = n_cell,
  probResponse = prob_response,
  probExact = TRUE,
  meanPos = mean_pos,
  transformation = transformation,
  samplePerturbationSd = 0,
  conditionPerturbationSd = 0,
  clusterPerturbationSd = 0,
  backgroundRelativeToResponse = 0.2, 
  ncellUnsRelativeToStim = 1,
  covEvMin = 2, 
  covEvMax = 2,
  tolClust = NULL
)
```

#### Bandwidths

```{r}
#| eval: false
#| label: check-bandwidths
# test indices
std_check <- TRUE
if (std_check) {
  transformation <- "gaussian"
  prob_response <- 0.05
  n_cell <- 1e4
  mean_pos <- 8
  bw_mtd <- "hpi1"
} else {
  transformation <- "gaussian"
  prob_response <- 0.002
  n_cell <- 5e3
  mean_pos <- 2
  bw_mtd <- "hpi1"
}

sim_res <- .simBandwidthEstBw(
  nSample = 10,
  nMarker = 1,
  nCondition = 2,
  nCluster = 2,
  nIter = 10,
  biasUns = 0.05,
  bwMtd = bw_mtd,
  nCellStim = n_cell,
  probResponse = prob_response,
  probExact = TRUE,
  meanPos = mean_pos,
  transformation = transformation,
  samplePerturbationSd = 0,
  conditionPerturbationSd = 0,
  clusterPerturbationSd = 0,
  backgroundRelativeToResponse = 0.2, 
  ncellUnsRelativeToStim = 1,
  covEvMin = 2, 
  covEvMax = 2,
  tolClust = NULL
)
```

## Actual

```{r}
#| label: actual-settings
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)[c(1, 3, 5, 7)]
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)[c(2, 4, 6, 8, 10, 16)]
mean_pos_vec <- seq_len(8)[c(2, 4, 6, 8)]
transformation_vec <- c("gaussian", "gamma", "skew")
sample_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
condition_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
cluster_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
background_relative_to_response <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)
n_cell_uns_relative_to_stim <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)

# bandwidth settings
bandwidth_method_vec <- c("hpi0", "hpi1", "hpi2", "hpi3", "sj", "nrd0")
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- seq(0, 2, by = 0.2)

# fix the bandwidth
bandwidth_value_vec <- c(0.01, 0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 2, 5)[c(3:8)]

# number of iterations
length(n_cell_stim_vec) * length(prob_response_vec) * length(mean_pos_vec) * length(transformation_vec) * length(bandwidth_value_vec)
```

### Varying bandwidth manually

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
So, we can vary the bandwidth from a small number (say 0.1) to a large number (say 10), and see how estimated background-subtracted frequencies compare to actual.

It should be doable with just the `bw` parameter, but ah, we haven't set that yet.

Okay, we've got the code working above. Now we'll vary the fixed bandwidth from 0.01 to 4, along with making the bias_unstim some reasonably large value (we'll have to fix it).

And we'll do this for varying numbers of cells. 

Generate the grid of parameters to simulate.

```{r}
#| label: bw-manual-grid
# 2. Flatten your nested loops into a single combination grid
sim_grid <- tidyr::expand_grid(
  transformation = transformation_vec,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  mean_pos       = mean_pos_vec,
  bw             = bandwidth_value_vec
) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))


if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos == 8,
      transformation %in% transformation_vec,
      bw == 0.5
    )
}
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-manual-loop
#| eval: false
for (i in seq_len(nrow(sim_grid))) {
  sim_res <- stimgate:::.simBandwidthBsFreq(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 10,
    biasUns = 0.05,
    bw = sim_grid$bw[i],
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    samplePerturbationSd = 0,
    conditionPerturbationSd = 0,
    clusterPerturbationSd = 0,
    backgroundRelativeToResponse = 0.2, 
    ncellUnsRelativeToStim = 1,
    covEvMin = 2, 
    covEvMax = 2,
    tolClust = NULL
  ) |>
    dplyr::mutate(
      transformation = transformation,
      prob_response = prob_response,
      n_cell = n_cell,
      mean_pos = mean_pos,
      bw = bw
    ) |>
    dplyr::select(
      transformation, prob_response, n_cell,
      mean_pos, bw, dplyr::everything()
    )
}
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-manual-parallel
# Activate global progress bars in your console
if (interactive()) {
  handlers(global = TRUE)
}


path_progress_file <- projr::projr_path_get(
  "cache", "sim-bw", "bs-freq", "progress.txt"
)
if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

# Set up the parallel backend
plan(multisession, workers = availableCores() - 1)

# 3. Wrap your parallel execution in with_progress()
bw_list_raw <- with_progress({
  
  # Create a progressor function with the total number of steps
  p <- progressor(steps = nrow(sim_grid))
  
  future_pmap(sim_grid, function(transformation, prob_response, n_cell, mean_pos, bw, sim_id) {

    settings_row <- paste0("trans = ", transformation, 
                           " | prob = ", prob_response, 
                           " | n_cell = ", n_cell, 
                           " | mean = ", mean_pos, 
                           " | bw = ", bw,
                           " | sim_id = ", sim_id)

    log_msg <- paste0("Running: ", settings_row, "\n")
    cat(log_msg, file = path_progress_file, append = TRUE)
    
    # Run the simulation
    out <- tryCatch({
      sim_res <- stimgate:::.simBandwidthBsFreq(
        nSample = 10,
        nMarker = 1,
        nCondition = 2,
        nCluster = 2,
        nIter = 10,
        biasUns = 0.05,
        bw = bw,
        nCellStim = n_cell,
        probResponse = prob_response,
        probExact = TRUE,
        meanPos = mean_pos,
        transformation = transformation,
        samplePerturbationSd = 0,
        conditionPerturbationSd = 0,
        clusterPerturbationSd = 0,
        backgroundRelativeToResponse = 0.2, 
        ncellUnsRelativeToStim = 1,
        covEvMin = 2, 
        covEvMax = 2,
        tolClust = NULL
      ) |>
        dplyr::mutate(
          transformation = transformation,
          prob_response = prob_response,
          n_cell = n_cell,
          mean_pos = mean_pos,
          bw = bw
        ) |>
        dplyr::select(
          transformation, prob_response, n_cell,
          mean_pos, bw, dplyr::everything()
        )

      cat("Completed: ", settings_row, "\n", file = path_progress_file, append = TRUE)
      
      # Signal successful step to the progress bar with a custom message
      p(sprintf("Completed sim_id: %s", sim_id))
      
      sim_res
    }, error = function(e) {
      
      # Signal failed step to the progress bar so it keeps moving
      err_msg <- paste0("Error [", settings_row, "]: ", e$message, "\n")
      cat(err_msg, file = path_progress_file, append = TRUE)
      p(sprintf("ERROR on sim_id: %s", sim_id))
      
      tibble::tibble(
        transformation = transformation,
        prob_response = prob_response,
        n_cell = n_cell,
        mean_pos = mean_pos,
        bw = bw,
        iter = NA,
        chnl = NA,
        sample = NA
      )
    })
    
    return(out)
  }, .options = furrr_options(seed = TRUE))
})

# Close the parallel workers when done
plan(sequential)

saveRDS(
  bw_list_raw,
  file = projr::projr_path_get(
    "output", "sim-bw", "bs-freq", "bw_list_raw.rds"
  )
)
```

Collate the results and calculate the metrics of interest.

```{r}
#| label: bw-manual-collate
#| error: true
bw_list_raw <- readRDS(
  file = projr::projr_path_get(
    "output", "sim-bw", "bs-freq", "bw_list_raw.rds"
  )
)

bw_tbl <- bw_list_raw |>
  dplyr::bind_rows() |>
  dplyr::filter(!is.na(iter))

# Extract gates
bw_gate_detail_tbl <- bw_tbl |>
  dplyr::filter(grepl("^loc_", .data$method)) |>
  dplyr::select(
    transformation,
    prob_response,
    n_cell,
    mean_pos,
    bw,
    iter,
    chnl,
    sample,
    method,
    propRespTruth,
    propRespEst,
    threshold,
    thresholdOrigin,
    locGenerated,
    locGeneratedDirect,
    locSource,
    locReason,
    detailLevel
  )

bw_gate_detail_tbl

bw_tbl |>
  dplyr::group_by(
    transformation, prob_response, n_cell, mean_pos,
    bw, method, chnl
    ) |>
  dplyr::summarise(
    mean_diff = mean(abs(propRespEst - propRespTruth)),
    rel_diff = mean(abs(propRespEst - propRespTruth)) / mean(propRespTruth),
    ccc = DescTools::CCC(propRespTruth, propRespEst)$rho.c$est,
    pearson = cor(propRespTruth, propRespEst, method = "pearson"),
    spearman = cor(propRespTruth, propRespEst, method = "spearman"),
    .groups = "drop"
  )  |>
  dplyr::arrange(transformation, prob_response, n_cell, mean_pos, bw)
```

```{r}
#| label: bw-manual-collate-1
#| error: true
bw_1 <- bw_list_raw[[1]] |>
  dplyr::filter(
    transformation == transformation[[1]],
    prob_response == max(prob_response),
    n_cell == n_cell[[1]],
    mean_pos == max(mean_pos),
    chnl == chnl[[1]],
    method == method[[1]]
  )
bw_1
bw_1 |>
  dplyr::group_by(
    transformation, prob_response, n_cell, mean_pos,
    bw, method, chnl
    ) |>
  dplyr::summarise(
    ccc = DescTools::CCC(propRespTruth, propRespEst)$rho.c$est,
    pearson = cor(propRespTruth, propRespEst, method = "pearson"),
    spearman = cor(propRespTruth, propRespEst, method = "spearman"),
    .groups = "drop"
  ) |>
  dplyr::arrange(transformation, prob_response, n_cell, mean_pos, bw)
```

### Checking what the estimators actually estimate

Get the parameters over which we simulate.

```{r}
#| label: bw-estimate-grid
sim_grid <- tidyr::expand_grid(
  transformation = transformation_vec,
  prob_response  = prob_response_vec,
  n_cell         = n_cell_stim_vec,
  mean_pos       = mean_pos_vec,
  bw_mtd         = bandwidth_method_vec
) |>
  dplyr::filter(
    !(n_cell * prob_response < 5 & prob_response < 0.04)
  ) |>
  dplyr::mutate(
    sim_id = dplyr::row_number()
  )

sim_grid <- sim_grid |>
  dplyr::arrange(n_cell, dplyr::desc(prob_response))

if ("dev" %in% projr::projr_profile_get()) {
  sim_grid <- sim_grid |>
    dplyr::filter(
      n_cell == 1e5,
      prob_response == 0.002,
      mean_pos == 8,
      transformation %in% transformation_vec,
      bw_mtd == "hpi1"
    )
}
```

A loop we only run when we want to see why particular runs failed.

```{r}
#| label: bw-estimate-loop
#| eval: false
bw_list_raw_mtd <- lapply(seq_len(nrow(sim_grid)), function(i) {
  NULL
})

for (i in seq_len(nrow(sim_grid))) {
  settings_row <- paste0(
    "trans = ",
    sim_grid$transformation[i],
    " | prob = ",
    sim_grid$prob_response[i],
    " | n_cell = ",
    sim_grid$n_cell[i],
    " | mean = ",
    sim_grid$mean_pos[i],
    " | bw_mtd = ",
    sim_grid$bw_mtd[i],
    " | sim_id = ",
    sim_grid$sim_id[i]
  )

  message("Running: ", settings_row)

  bw_list_raw_mtd[[i]] <- stimgate:::.simBandwidthEstBwDirect(
    nSample = 10,
    nMarker = 1,
    nCondition = 2,
    nCluster = 2,
    nIter = 10,
    biasUns = 0.05,
    bwMtd = sim_grid$bw_mtd[i],
    bwMin = 1e-10,
    bwMax = 1e10,
    bwCluster = 0.5,
    nCellStim = sim_grid$n_cell[i],
    probResponse = sim_grid$prob_response[i],
    probExact = TRUE,
    meanPos = sim_grid$mean_pos[i],
    transformation = sim_grid$transformation[i],
    backgroundRelativeToResponse = 0.2,
    ncellUnsRelativeToStim = 1,
    covEvMin = 2,
    covEvMax = 2,
    tolClust = NULL,
    summarise = FALSE
  ) |>
    dplyr::mutate(
      transformation = sim_grid$transformation[i],
      prob_response = sim_grid$prob_response[i],
      n_cell = sim_grid$n_cell[i],
      mean_pos = sim_grid$mean_pos[i],
      bw_mtd = sim_grid$bw_mtd[i],
      sim_id = sim_grid$sim_id[i]
    ) |>
    dplyr::select(
      transformation,
      prob_response,
      n_cell,
      mean_pos,
      bw_mtd,
      sim_id,
      dplyr::everything()
    )

  message("Completed: ", settings_row)
}

saveRDS(
  bw_list_raw_mtd,
  file = projr::projr_path_get(
    "output",
    "sim-bw",
    "mtd",
    "bw_list_raw_mtd.rds"
  )
)

bw_list_raw_mtd |>
  dplyr::bind_rows() |>
  dplyr::group_by(
    transformation, prob_response, n_cell,
    mean_pos, bw_mtd, chnl,
    n_cell_stim, n_cell_uns
  ) |>
  dplyr::summarise(
    mean_bw_stim = mean(bw_stim),
    mean_bw_uns = mean(bw_uns),
    mean_bw = mean(bw),
    .groups = "drop"
  ) |>
  dplyr::select(
    transformation, mean_bw
  )
```

Run the simulations in parallel, with progress reporting.

```{r}
#| label: bw-estimate-parallel

if (interactive()) {
  progressr::handlers(global = TRUE)
}

path_progress_file <- projr::projr_path_get(
  "cache",
  "sim-bw",
  "mtd",
  "progress.txt"
)

if (file.exists(path_progress_file)) {
  file.remove(path_progress_file)
}

future::plan(
  future::multisession,
  workers = max(1, future::availableCores() - 1)
)

bw_list_raw_mtd <- progressr::with_progress({
  p <- progressr::progressor(steps = nrow(sim_grid))

  furrr::future_pmap(
    sim_grid,
    function(transformation, prob_response, n_cell, mean_pos, bw_mtd, sim_id) {
      settings_row <- paste0(
        "trans = ",
        transformation,
        " | prob = ",
        prob_response,
        " | n_cell = ",
        n_cell,
        " | mean = ",
        mean_pos,
        " | bw_mtd = ",
        bw_mtd,
        " | sim_id = ",
        sim_id
      )

      log_msg <- paste0("Running: ", settings_row, "\n")
      cat(log_msg, file = path_progress_file, append = TRUE)

      out <- tryCatch(
        {
          sim_res <- stimgate:::.simBandwidthEstBwDirect(
            nSample = 10,
            nMarker = 1,
            nCondition = 2,
            nCluster = 2,
            nIter = 10,
            biasUns = 0.05,
            bwMtd = bw_mtd,
            bwMin = 1e-10,
            bwMax = 1e10,
            bwCluster = 0.5,
            nCellStim = n_cell,
            probResponse = prob_response,
            probExact = TRUE,
            meanPos = mean_pos,
            transformation = transformation,
            backgroundRelativeToResponse = 0.2,
            ncellUnsRelativeToStim = 1,
            covEvMin = 2,
            covEvMax = 2,
            tolClust = NULL,
            summarise = FALSE
          ) |>
            dplyr::mutate(
              transformation = transformation,
              prob_response = prob_response,
              n_cell = n_cell,
              mean_pos = mean_pos,
              bw_mtd = bw_mtd,
              sim_id = sim_id
            ) |>
            dplyr::select(
              transformation,
              prob_response,
              n_cell,
              mean_pos,
              bw_mtd,
              sim_id,
              dplyr::everything()
            )

          cat(
            "Completed: ",
            settings_row,
            "\n",
            file = path_progress_file,
            append = TRUE
          )

          p(sprintf("Completed sim_id: %s", sim_id))

          sim_res
        },
        error = function(e) {
          err_msg <- paste0("Error [", settings_row, "]: ", e$message, "\n")
          cat(err_msg, file = path_progress_file, append = TRUE)

          p(sprintf("ERROR on sim_id: %s", sim_id))

          tibble::tibble(
            transformation = transformation,
            prob_response = prob_response,
            n_cell = n_cell,
            mean_pos = mean_pos,
            bw_mtd = bw_mtd,
            sim_id = sim_id,
            iter = NA_integer_,
            chnl = NA_character_,
            sample = NA_character_,
            ind = NA_character_,
            bw_uns = NA_real_,
            bw_stim = NA_real_,
            bw = NA_real_,
            bw_source = NA_character_,
            error = e$message
          )
        }
      )

      out
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
})

future::plan(future::sequential)

saveRDS(
  bw_list_raw_mtd,
  file = projr::projr_path_get(
    "output",
    "sim-bw",
    "mtd",
    "bw_list_raw_mtd.rds"
  )
)
```

Collate the results.

```{r}
#| label: bw-estimate-collate
#| error: true
bw_list_raw_mtd <- readRDS(
  file = projr::projr_path_get(
    "output",
    "sim-bw",
    "mtd",
    "bw_list_raw_mtd.rds"
  )
)
bw_tbl_mtd <- bw_list_raw_mtd |>
  dplyr::bind_rows() |>
  dplyr::filter(!is.na(iter))
```


