---
title: Simulate best bandwidth
format:
  html:
    embed-resources: true
    toc: true
---

```{r}
#| include: false
library(ggplot2)
library(tibble)

# Find the build directory root
root_dir <- projr::projr_path_get("project")
desc_path <- file.path(root_dir, "DESCRIPTION")

if (
  file.exists(desc_path) &&
    grepl("Package: stimgate", readLines(desc_path, n = 1))
) {
  devtools::load_all(root_dir)
} else {
  fn_vec <- list.files(file.path(root_dir, "R"), full.names = TRUE)
  for (fn in fn_vec) {
    source(fn)
  }
}
```

## Checks

### Transformations

```{r}
calc_gamma <- function(x) {
  gamma(1 + abs(x / 4))
}
calc_gaussian <- function(x) {
  x
}
calc_skew <- function(x, epsilon = 2, delta = 1) {
  cluster_mean <- mean(x)
  cluster_sd <- sd(x)
  weight <- 1 / (1 + exp(1.5 * (cluster_mean - 3.5)))
  epsilon <- epsilon * weight
  out <- sinh(epsilon + delta * asinh(x))
  gamma_divisor <- rgamma(length(x), shape = 5, rate = 5)
  out_fat <- out / sqrt(gamma_divisor)
  (out - mean(out)) / sd(out) * sd(x) + cluster_mean
}
mean_vec <- c(0, 8)
plot_tbl <- purrr::map_df(mean_vec, function(mean_curr) {
  sim_vec <- rnorm(1e3, mean = mean_curr, sd = sqrt(2))
  tibble::tibble(
    trans = rep(c("gamma", "gaussian", "skew"), each = length(sim_vec)),
    sim = c(calc_gamma(sim_vec), calc_gaussian(sim_vec), calc_skew(sim_vec))
  )
}) |>
  dplyr::filter(!trans == "gamma")
ggplot(plot_tbl, aes(x = sim, col = trans)) +
  geom_density(alpha = 0.7) +
  cowplot::theme_cowplot()
```

### Simulations

```{r}
#| eval: false
# debugonce(simCytCondition)
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    "skew" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  out_list_doncition <- simCytCondition(
    nMarker = 2,
    nCell = 1e4,
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probVec = c(0.05, 0.02, 0.075, 0.855)
  )
  plot_tbl <- tibble::tibble(
    pop = out_list_condition$conditionLabels
  ) |>
    dplyr::bind_cols(
      out_list_condition$conditionMatrix
    )
  pop_vec <- unique(plot_tbl$pop)
  p_list <- lapply(pop_vec, function(x) {
    UtilsCytoRSV::plot_cyto(
      plot_tbl |>
        dplyr::filter(pop == x),
      marker = c("F1", "F2"),
      limits_expand = limits_expand
    ) +
      labs(title = x) +
      geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
      geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
  })
  plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = NA)
    )
  path_save <- projr::projr_path_get(
    "cache",
    "sim-bandwidth",
    "check",
    "condition",
    paste0("p-condition-", trans, ".png")
  )
  ggplot2::ggsave(
    path_save,
    units = "cm",
    width = 16,
    height = 16
  )
}
```

```{r}
#| eval: false
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  # debugonce(simCytSample)
  out_list_sample <- simCytSample(
    nMarker = 2L,
    nCondition = 3L,
    nCluster = 4L,
    nCellByCondition = c(1e4, 1e4, 1e4),
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probResponseVecByStimCondition = list(
      c(0.1, 0.05, 0.0025, 0.8425),
      c(0.00001, 0.0001, 0.0001, 0.99978)
    ),
    probVecUns = c(0.05, 0.02, 0.075, 0.855),
    conditionPerturbationSd = 0,
    clusterPerturbationSd = 0
  )
  for (i in seq_along(out_list_sample$flowFrameList)) {
    stim <- names(out_list_sample$flowFrameList)[[i]]
    plot_tbl <- tibble::tibble(
      pop = out_list_sample$conditionLabelsList[[i]]
    ) |>
      dplyr::bind_cols(
        as.data.frame(flowCore::exprs(out_list_sample$flowFrameList[[i]]))
      )
    pop_vec <- unique(plot_tbl$pop)
    p_list <- lapply(pop_vec, function(x) {
      UtilsCytoRSV::plot_cyto(
        plot_tbl |>
          dplyr::filter(pop == x),
        marker = c("F1", "F2"),
        limits_expand = limits_expand
      ) +
        labs(title = x) +
        geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
        geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
    })
    plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
      )
    path_save <- projr::projr_path_get(
      "cache",
      "sim-bandwidth",
      "check",
      "sample",
      paste0("p-sample-", trans, "-", stim, ".png")
    )
    ggplot2::ggsave(
      path_save,
      units = "cm",
      width = 16,
      height = 16
    )
  }
}
```

```{r}
#| eval: false
for (trans in c("gaussian", "gamma", "skew")) {
  trans_func <- switch(
    trans,
    "gamma" = calc_gamma,
    "gaussian" = calc_gaussian,
    "skew" = calc_skew,
    stop("transformation not recognised")
  )
  limits_expand <- switch(
    trans,
    "gamma" = list(c(0, 4)),
    "gaussian" = list(c(-4, 12)),
    stop("transformation not recognised")
  )
  out_list_experiment <- simCytExperiment(
    nSample = 2L,
    nMarker = 2L,
    nCondition = 3L,
    nCluster = 4L,
    nCellByCondition = c(1e4, 1e4, 1e4),
    transformationFunc = trans_func,
    mixtureType = "gaussianOnly",
    meanExprMat = matrix(
      c(c(8, 8), c(8, 0), c(0, 8), c(0, 0)),
      byrow = TRUE,
      ncol = 2
    ),
    clusterLabelVec = c("gp2p", "gp2n", "gn2p", "gn2n"),
    probResponseVecByStimCondition = list(
      c(0.1, 0.05, 0.0025, 0.8425),
      c(0.00001, 0.0001, 0.0001, 0.99978)
    ),
    probVecUns = c(0.05, 0.02, 0.075, 0.855),
    samplePerturbationSd = 0.1,
    conditionPerturbationSd = 0.1,
    clusterPerturbationSd = 0.1
  )
  for (i in seq_along(out_list_experiment$flowFrameList)) {
    smpl_cond <- names(out_list_experiment$flowFrameList)[[i]]
    plot_tbl <- tibble::tibble(
      pop = out_list_experiment$conditionLabelsList[[i]]
    ) |>
      dplyr::bind_cols(
        as.data.frame(flowCore::exprs(out_list_experiment$flowFrameList[[i]]))
      )
    pop_vec <- unique(plot_tbl$pop)
    p_list <- lapply(pop_vec, function(x) {
      UtilsCytoRSV::plot_cyto(
        plot_tbl |>
          dplyr::filter(pop == x),
        marker = c("F1", "F2"),
        limits_expand = limits_expand
      ) +
        labs(title = x) +
        geom_hline(yintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4) +
        geom_vline(xintercept = trans_func(c(0, 8)), col = "red", alpha = 0.4)
    })
    plot_grid <- cowplot::plot_grid(plotlist = p_list, ncol = 2) +
      theme(
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
      )
    path_save <- projr::projr_path_get(
      "cache",
      "sim-bandwidth",
      "check",
      "experiment",
      paste0("p-experiment-", trans, "-", smpl_cond, ".png")
    )
    ggplot2::ggsave(
      path_save,
      units = "cm",
      width = 16,
      height = 16
    )
  }
}
```

### Analysis

```{r}
# simulation settings
n_cell_stim_vec <- c(1e3, 2e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, 1e6, 2e6, 5e6)
prob_response_vec <- c(
  1 / 5e5,
  1 / 2e5,
  1 / 1e5,
  1 / 5e4,
  1 / 2e4,
  1 / 1e4,
  1 / 5e3,
  1 / 2e3,
  1 / 1e3,
  1 / 5e2,
  1 / 2e2,
  1 / 1e2,
  1 / 5e1,
  1 / 2e1,
  1 / 1e1,
  1 / 5,
  1 / 2
)
mean_pos_vec <- seq_len(8)
transformation_vec <- c("gaussian", "gamma", "skew")
sample_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
condition_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
cluster_perturbation_sd <- c(0, sqrt(1 / 4), sqrt(1 / 2), 1, sqrt(2), 2)
background_relative_to_response <- c(2, 1, 1 / 2, 0.2, 0.1, 0.05, 0.01, 0.001)
n_cell_uns_relative_to_stim <- c(0.1, 0.2, 0.5, 1, 2, 5, 10)

# bandwidth settings
bandwidth_method_vec <- c("hpi_0", "hpi_1", "hpi_2", "hpi_3", "sj", "nrd0")
bandwidth_ncell_upper_vec <- c(2.5e3, 5e3, 1e4, 2e4, 5e4, 1e5, 2e5, 5e5, Inf)
bandwidth_ncell_lower_vec <- c(0, 10, 20, 30, 50, 100, 200, 500, 1e3)
bias_uns_vec <- seq(0, 2, by = 0.2)

# fix the bandwidth
bandwidth_value_vec <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2, 5)
```

```{r}
# loop parameters
n_cell_stim <- 1e5
prob_response <- 0.02
mean_pos <- 8
transformation <- "gaussian"
sample_perturbation_sd <- 0
condition_perturbation_sd <- 0
cluster_perturbation_sd <- 0
background_relative_to_response <- 0.2
ncell_uns_relative_to_stim <- 1

# stimgate parameters
bias_uns <- 0.2

# calculated parameters
n_cell_uns <- round(n_cell_stim * ncell_uns_relative_to_stim)
n_cell_by_condition <- c(n_cell_uns, n_cell_stim)
transformation_func <- switch(
  transformation,
  "gamma" = calc_gamma,
  "gaussian" = calc_gaussian,
  "skew" = calc_skew,
  stop("transformation not recognised")
)
mean_expr_mat <- matrix(
  c(0, 8),
  byrow = TRUE,
  ncol = 1
)
cluster_label_vec <- c("gn", "gp")
prob_responseUns <- prob_response * background_relative_to_response
prob_vec_uns <- c(1 - prob_responseUns, prob_responseUns)
prob_response_vec_by_stim_condition <- list(c(-prob_response, prob_response))
iter_num <- 1

n_sample <- 1L
n_marker <- 1L
n_condition <- 2L
nCluster <- 2L

print(paste0("iter ", iter_num))
# debugonce(simCytCondition)
out_list_experiment <- simCytExperiment(
  nSample = n_sample,
  nMarker = n_marker,
  nCondition = n_condition,
  nCluster = nCluster,
  nCellByCondition = n_cell_by_condition,
  transformationFunc = transformation_func,
  mixtureType = "gaussianOnly",
  meanExprMat = mean_expr_mat,
  clusterLabelVec = cluster_label_vec,
  probVecUns = prob_vec_uns,
  probResponseVecByStimCondition = prob_response_vec_by_stim_condition,
  conditionPerturbationSd = condition_perturbation_sd,
  clusterPerturbationSd = cluster_perturbation_sd,
  samplePerturbationSd = sample_perturbation_sd
)

# apply stimgate
fs <- as(out_list_experiment[["flowFrameList"]], "flowSet")
gs <- flowWorkspace::GatingSet(fs)

pathProject <- projr::projr_path_get(
  "cache",
  "sim-bandwidth",
  "check",
  "gate",
  paste0("iter_", iter_num)
)
unlink(pathProject, recursive = TRUE)
dir.create(pathProject, recursive = TRUE)
batchList <- lapply(seq(n_sample), function(i) {
  seq((i - 1) * n_condition + 1, i * n_condition)
})
chnl_list_stimgate <- list(
  list("cut" = "F1")
)

# debugonce(stimgate:::.complete_chnl_list)
gateStim(
  pathProject = pathProject,
  chnl = "F1",
  .data = gs,
  batchList = batchList,
  chnlSettings = chnl_list_stimgate,
  biasUns = bias_uns
)

stats_tbl <- readRDS(
  file.path(pathProject, "gateStats.rds")
)
stats_tbl
```

#### Comparing actual to estimated frequencies

This is for when we're varying the bandwidth manually.

```{r}
n_cell_sim_check_bw <- 1e4
test_data <- simCytExperiment(
  nSample = 10L,
  nMarker = 1L,
  nCondition = 2L,
  nCluster = 2L,
  nCellByCondition = c(n_cell_sim_check_bw, n_cell_sim_check_bw),
  transformationFunc = calc_gaussian,
  mixtureType = "gaussianOnly",
  meanExprMat = matrix(
    c(0, 8),
    byrow = TRUE,
    ncol = 1
  ),
  clusterLabelVec = c("gn", "gp"),
  probVecUns = c(0.98, 0.02),
  probResponseVecByStimCondition = list(c(-0.02, 0.02)),
  covEvMin = 2,
  covEvMax = 2
)

labelsList <- file.path(test_data$path_cache, "labelsList.rds") |>
  readRDS()
gs <- flowWorkspace::load_gs(test_data$path_gs)
pathProject <- file.path(test_data$path_cache, "stimgate")
# debugonce(.get_cp_uns_loc_get_prob)
Sys.setenv("STIMGATE_INTERMEDIATE" = "TRUE")
invisible(gateStim(
  .data = gs,
  pathProject = pathProject,
  popGate = "root",
  batchList = test_data$batchList,
  marker = test_data$marker,
  bwAdj = 2
))

n_sample <- length(labelsList) / 2
n_condition <- 2
# truth table
prob_bs_tbl_truth <- purrr::map_df(seq_len(n_sample), function(sampleCurr) {
  labelsIdx <- (sampleCurr - 1) * n_condition + seq(1, n_condition)
  purrr::map_df(labelsIdx, function(i) {
    labelVec <- labelsList[[i]]
    stimCondition <- ifelse(i %% 2 == 0, "stim", "unstim")
    f1p <- sum(grepl("^pos", labelVec)) / length(labelVec)
    f2p <- sum(grepl("pos$", labelVec)) / length(labelVec)
    tibble::tibble(
      ind = sampleCurr |> as.character(),
      stim = stimCondition,
      f1p = f1p,
      f2p = f2p
    )
  })
}) |>
  tidyr::pivot_wider(
    names_from = stim,
    values_from = c(f1p, f2p)
  ) |>
  dplyr::mutate(
    f1p_bs = f1p_stim - f1p_unstim,
    f2p_bs = f2p_stim - f2p_unstim
  ) |>
  dplyr::rename(
    sample = ind
  ) |>
  dplyr::select(
    all_of(c("sample", "f1p_bs", "f2p_bs"))
  ) |>
  tidyr::pivot_longer(
    cols = c(f1p_bs, f2p_bs),
    names_to = "chnl",
    values_to = "prop_resp_truth"
  ) |>
  dplyr::mutate(
    chnl = gsub("p_bs", "", chnl)
  )

path_dir_int_init <- file.path(pathProject, "intermediateData", "init")
chnl_vec <- list.dirs(path_dir_int_init, full.names = FALSE, recursive = FALSE)
chnl <- chnl_vec[[1]]
ind <- n_sample * n_condition
prob_bs_tbl_est <- purrr::map_df(chnl_vec, function(chnl) {
  ind_vec_stim <- seq(2, n_sample * n_condition, by = 2) |>
    as.character()
  purrr::map_df(ind_vec_stim, function(ind) {
    path_ind <- file.path(path_dir_int_init, chnl, "ind", ind)
    prob_tbl_list <- file.path(path_ind, "prob_tbl_list.rds") |>
      readRDS()
    prob_norm <- prob_tbl_list[["pos"]]
    ncell_resp_norm <- if (nrow(prob_norm) > 0L) {
      sum(prob_norm$probStimNorm)
    } else {
      0
    }
    prob_smooth <- file.path(path_ind, "prob_smooth_out.rds") |>
      readRDS()
    ncell_resp_smooth <- if (nrow(prob_smooth) > 0L) {
      sum(prob_smooth$prob_smooth)
    } else {
      0
    }
    ncell_resp_pred <- if (nrow(prob_smooth) > 0L) {
      sum(prob_smooth$pred)
    } else {
      0
    }
    ncell_condition <- n_cell_sim_check_bw
    tibble::tibble(
      sample = as.character(as.numeric(ind) / n_condition),
      ind = ind,
      chnl = chnl,
      ncell_condition = ncell_condition,
      ncell_resp_norm = ncell_resp_norm,
      ncell_resp_smooth = ncell_resp_smooth,
      ncell_resp_pred = ncell_resp_pred,
      prop_resp_norm = ncell_resp_norm / ncell_condition,
      prop_resp_smooth = ncell_resp_smooth / ncell_condition,
      prop_resp_pred = ncell_resp_pred / ncell_condition
    )
  })
}) |>
  dplyr::mutate(
    chnl = dplyr::case_when(
      chnl == "BC1(La139)Dd" ~ "f1",
      chnl == "BC2(Pr141)Dd" ~ "f2",
      .unmatched = "error"
    )
  ) |>
  dplyr::select(
    sample,
    chnl,
    prop_resp_norm,
    prop_resp_smooth,
    prop_resp_pred
  ) |>
  tidyr::pivot_longer(
    cols = c(prop_resp_norm, prop_resp_smooth, prop_resp_pred),
    names_to = "method",
    values_to = "prop_resp"
  )
comparison_tbl <- prob_bs_tbl_truth |>
  dplyr::left_join(
    prob_bs_tbl_est,
    by = c("sample", "chnl")
  )
comparison_tbl |>
  dplyr::group_by(method) |>
  dplyr::summarise(
    ccc = DescTools::CCC(prop_resp_truth, prop_resp)$rho.c$est,
    pearson = cor(prop_resp_truth, prop_resp, method = "pearson"),
    spearman = cor(prop_resp_truth, prop_resp, method = "spearman")
  )
chnlSettings <- stimgate_meta_read_settings_chnls(pathProject)
chnlSettings |>
  lapply(
    calc_gaussian[c("bwMin", "bwMax", "bwCluster", "bias_uns")] |>
      lapply(function(y) signif(y, 2))
  )

# let's plot it
plotlist <- lapply(1:8, function(i) {
  ind_vec <- c((i - 1) * n_condition + 1, i * n_condition)
  plotStim(
    ind = ind_vec,
    .data = gs,
    pathProject = pathProject,
    chnl = "BC1(La139)Dd"
  )
})
p_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 2,
  labels = paste0("sample ", seq_len(8))
)
ggsave(
  filename = file.path(dir_cache_check_bw, "p_grid.png"),
  plot = p_grid,
  width = 30,
  height = 30,
  units = "cm"
)
```

## Varying bandwidth manually

So, now we'll vary the bandwidth from a small value up to to a large value, where "small" is relative to the known spread of the data.

The spread of the data (without any batch effects) around the mean is on average 2, and the spread between the means is up to 8.
So, we can vary the bandwidth from a small number (say 0.1) to a large number (say 10), and see how estimated background-subtracted frequencies compare to actual.

It should be doable with just the `bw` parameter, but ah, we haven't set that yet.

Okay, we've got the code working above. Now we'll vary the fixed bandwidth from 0.01 to 4, along with making the bias_unstim some reasonably large value (we'll have to fix it).

And we'll do this for varying numbers of cells. 

```{r}

```